### 2022-08-31 rebuttal.tex
Update on Overleaf.

%!TeX TS-program = Lualatex
%!TeX encoding = UTF-8 Unicode
%!TeX spellcheck = en-US
% https://tex.stackexchange.com/questions/354194/how-to-edit-newlfm 
\documentclass[9pt]{article}

\usepackage{charter,graphicx}
\usepackage[margin=1in]{geometry}

\usepackage{fancyhdr}
\pagestyle{fancy}

\newcommand{\AuthorAG}{Antoine Grimaldi}
\newcommand{\AuthorCB}{Camille Besnainou}
\newcommand{\AuthorHL}{Hugo Ladret}
\newcommand{\AuthorLP}{Laurent U Perrinet}
\newcommand{\EmailLP}{laurent.perrinet@univ-amu.fr}
\newcommand{\EmailAG}{antoine.grimaldi@univ-amu.fr}
\newcommand{\EmailHL}{hugo.ladret@univ-amu.fr}
\newcommand{\orcidLP}{0000-0002-9536-010X}
\newcommand{\orcidHL}{0000-0001-7999-3751}
\newcommand{\orcidAG}{0000-0002-3107-4788}
\newcommand{\Address}{Institut de Neurosciences de la Timone (UMR 7289); Aix Marseille Univ, CNRS; Marseille, France}%
\newcommand{\Affiliation}{Institut de Neurosciences de la Timone (UMR 7289); Aix Marseille Univ, CNRS;}%
\newcommand{\CityLP}{Marseille, France}%
\newcommand{\WebsiteLP}{https://laurentperrinet.github.io}%

\renewcommand{\headrulewidth}{.4pt}% Default header rule
\renewcommand{\footrulewidth}{0pt}% No footer rule
\fancyhf{}% Clear header/footer
\fancypagestyle{plain}{
  \renewcommand{\headrulewidth}{0pt}% No header rule
  \renewcommand{\footrulewidth}{.4pt}% Default footer rule
  \fancyhf{}% Clear header/footer
}
\AtBeginDocument{\thispagestyle{plain}}

\setlength{\parindent}{0pt}
\setlength{\parskip}{.5\baselineskip plus 1pt minus 1pt}
%: %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\Journal}{IEEE International Conference on Image Processing}%

\bibliography{biblio}

\begin{document}

\hrulefill

\vspace{.1\baselineskip}

\begin{tabular}[b]{@{} l @{}}
  \today\\[12pt] % Date

\AuthorAG\ \\

\Address\\%[6pt]
%\AuthorA\\[6pt]
%%    \AddressA
%    \LongAddressA
%        \phonefrom{\PhoneA}
  Email: \EmailAG %\\[6pt]

\end{tabular}
%\hfill
%\begin{tabular}[t]{@{} l @{}}
%  Mrs. Jane Smith \\ % Addressee of the letter above the to address
%  Recruitment Officer \\ % To address
%  The Corporation \\
%  123 Pleasant Lane \\
%  City, State 12345
%\end{tabular}
%\hspace*{7em}

\hrulefill

\vspace{1\baselineskip}

To the Area Chair of \emph{\Journal},%

\vspace{2\baselineskip}

We would like to thank the reviewers for their comments that help us improve the clarity of our short paper. We are thankful to the reviewers for their global interest and the positive feedbacks raised by our study. %Even if some reviewers explicitly mention their lack of knowledge in the field preventing them to provide a confident evaluation, reviews are encouraging. 
We report in this rebuttal three main remarks brought by reviewers: 1/ about the pertinence of the contribution of this work to the image processing field (0357), 2/ about the clarity of the description of our model and about the figure illustrating it (1B91) and 3/ about the generalization of this model to more realistic data (1B91).

First, we would like to highlight the fact that our work is to be presented at the Special Session about \textit{Neuromorphic and perception-based image acquisition and analysis}. Thus, using as a visual signal the output of an event-based camera fits the scope of this session. This is why we convert motion clouds movies into spikes. We are mostly ready to make some modifications to the manuscript to improve this description. We strongly believe such event-based algorithms will benefit the image processing community.

Second, we will improve the formalism and the figure to clarify some confusing points. From the question, we understand there is some confusion about the scale of Figure 2-(a) and (c). Figure 2-(c) is the spiking output of the layer of hetero-synaptic delays neurons with the raster plot in Figure 2-(a) as input. Indeed, the output spike could not happen before the integration of the whole spatio-temporal motif.

Third, the reviewer points to a really interesting question about generalization, and we would clearly like to go into this direction. Our future goal is to apply this model for more realistic data, and preliminary results are encouraging. As it stands, our model may not be able to infer a velocity that is different from the 8 motion directions learned during training. Indeed, we use a softmax activation function, which implies some lateral inhibition of the neurons of a layer. By changing the spiking mechanism, we believe that this model will be able to infer a linear combination of the basis motion directions learned to generalize to a continuum of motions.

We hope these points clarify our contribution. Thank you for your consideration,
\vspace{\baselineskip}

Sincerely,

\vspace{2\baselineskip}

On behalf of the authors, \\
\AuthorAG

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{.5\baselineskip}
\begingroup
\renewcommand{\section}[2]{}%
%\printbibliography
\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-08-31 2022-07-11_Grimaldi-etal-ICIP.tex
Update on Overleaf.

% !TeX TS-program = Lualatex
% !TeX encoding = UTF-8 Unicode
% !TeX spellcheck = en
% !BIB TS-program = bibtex
% -*- coding: UTF-8; -*-
% vim: set fenc=utf-8
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% CONF: https://2022.ieeeicip.org/
% dates: https://2022.ieeeicip.org/important-dates/
%
% paper kit: https://cmsworkshops.com/ICIP2022/papers/paper_kit.php#Templates
% actual submission : https://cmsworkshops.com/icip2022/Papers/Submission.asp
%
% 4 pages + one of references
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%Review #0357
%
%*General Comments to Authors*
%This paper proposes a model for detection of temporal spiking motifs, which can potentially improve the performance of spiking neural networks (SNN). The topic is out of my research scope. However, I am not sure how this paper is related to ICIP. The authors applied the proposed model in motion detection. Such motion is captured by imaging or not? The overall presentation of the paper is acceptable, but I would suggest the authors to more explicitly represent the contribution of the proposed model and how it is expected to benefit image processing studies.
%
%-----------
%Review #1B91
%
%*General Comments to Authors*
%The authors propose a Spiking Neural Network-like method that consists of a single layer of spiking neurons aiming to learn the heterosexual-synaptic delays in order to detect motion. The learning happens in a supervised way as the authors transform the Motion Cloud dataset into spike trains in order to control the direction of motion which is the ground-truth information. The authors considered 8 different directions of motion that are assigned to the 8 learnt kernels they introduce in their model. According to the experimental results, the model is able to capture the direction of motion which is evident to the selectivity along the specific axes of the different kernels. Last but not least, The authors assess their model in terms of the classification accuracy as a function of the computational cost. They show that on the one hand, their method is able to highly reduce the computational load by thresholding some of the weights but at the same time achieve the same performance to computationally more greedy algorithms. The paper is very interesting and promising as this method can be easily extended to hardware.
%
%Some suggestions to improve the presentation of the results are given below:
%1. Figure 2 is a bit confusing. First of all, in (b) it's not that evident the different size of the radius but in few cases - maybe another example could be selected. In addition, if (a)-(b) and (c) is a single experiment how the delay is related to the time we observe the spike trains? Maybe that could be better explained in the text too.
%
%Last but not least, are these 8 directions enough to capture every motion? Could we use this model in real data or the computational cost would increase as one should consider much higher number of motion directions and as a results much larger set of kernels to be trained?
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\Title}{
Learning hetero-synaptic delays for motion detection in a single layer of spiking neurons
%FAST MOTION DETECTION USING HETERO-SYNAPTIC DELAYS IN A LAYER OF SPIKING NEURONS
}
\newcommand{\AuthorAG}{Antoine Grimaldi}
\newcommand{\AuthorCB}{Camille Besnainou}
\newcommand{\AuthorHL}{Hugo Ladret}
\newcommand{\AuthorLP}{Laurent U Perrinet}
\newcommand{\EmailLP}{laurent.perrinet@univ-amu.fr}
\newcommand{\EmailAG}{antoine.grimaldi@univ-amu.fr}
\newcommand{\EmailHL}{hugo.ladret@univ-amu.fr}
\newcommand{\orcidLP}{0000-0002-9536-010X}
\newcommand{\orcidHL}{0000-0001-7999-3751}
\newcommand{\orcidAG}{0000-0002-3107-4788}
\newcommand{\Address}{Institut de Neurosciences de la Timone (UMR 7289); Aix Marseille Univ, CNRS; Marseille, France}%
\newcommand{\Affiliation}{Institut de Neurosciences de la Timone (UMR 7289); Aix Marseille Univ, CNRS;}%
\newcommand{\CityLP}{Marseille, France}%
\newcommand{\WebsiteLP}{https://laurentperrinet.github.io}%
\newcommand{\Abstract}{
The response of a biological neuron depends on the precise timing of afferent spikes. This temporal aspect of the neuronal code is essential in understanding information processing in neurobiology and applies particularly well to the output of neuromorphic hardware such as event-based cameras. However, most artificial neuronal models do not take advantage of this minute temporal dimension and here, we develop a model for the efficient detection of temporal spiking motifs based on a layer of neurons with hetero-synaptic delays. Indeed, the variety of synaptic delays on the dendritic tree allows to synchronize synaptic inputs as they reach the basal dendritic tree. We show this can be formalized as a time-invariant logistic regression which can be trained using labelled data. We apply this model to solve the specific computer vision problem of motion detection, and demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of event-based cameras. In particular, we quantify how its accuracy can vary with the total computational load. This end-to-end event-driven computational brick could help improve the performance of future Spiking Neural Network (SNN) algorithms and their prospective use in neuromorphic chips.
}
\newcommand{\Keywords}{time code, event-based computations, spiking neural networks, motion detection, efficient coding, logistic regression
}
\newcommand{\Acknowledgments}{This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}''). %HL and LP received funding from the ANR project  N° ANR-20-CE23-0021 (''\href{https://laurentperrinet.github.io/grant/anr-anr/}{AgileNeuroBot}''). HL received funding from a University of Montreal Artificial Intelligence scholarship. For the purpose of open access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript version arising from this submission. This research was funded, in whole or in part, by [Organisation name, Grant #]. 
A CC-BY public copyright license has been applied by the authors to the present document and will be applied to all subsequent versions up to the Author Accepted Manuscript arising from this submission, in accordance with the grant’s open access conditions. 
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}

\usepackage[utf8]{inputenc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% NOTATIONS
% Pixel world
\newcommand{\presynaddr}{a} % pre address
\newcommand{\postsynaddr}{n} % post address
\newcommand{\numevent}{N_{ev}} % total number of events
\newcommand{\presynaddrspace}{\mathcal{A}} %presynaptic address space
\newcommand{\postsynaddrspace}{\mathcal{B}} %postsynaptic address space
\newcommand{\Npol}{N_\text{p}} % number of polarity
\newcommand{\Nneuron}{N_\text{n}} % number of output neurons in the layer
\newcommand{\arank}{r} % address index
\newcommand{\bias}{b} % bias for the MLR model
\newcommand{\synapse}{\sigma} % synapse
\newcommand{\synapticweight}{w} % synaptic weight
\newcommand{\synapticdelay}{\delta} % synaptic delay
\newcommand{\ranksyn}{s} % synapse index
\newcommand{\Nsyn}{N_{s}} % total number of synapses
\newcommand{\activeweights}{\mathcal{W}} 
\newcommand{\timev}{t} % time
\newcommand{\polev}{p} % polarity
\newcommand{\event}{\epsilon} % event
\newcommand{\eventstream}{\xi} % stream of events
\newcommand{\TS}{S} % time surface
\newcommand{\neuron}{\mathbf{n}} % neuron in the SNN (defined by the spatial position and the channel)
\newcommand{\postneuron}{\mathbf{m}} % post synaptic neuron in the SNN (defined by the spatial position and the kernel)
\newcommand{\channel}{\mathbf{p}} % channel
\newcommand{\layer}{\mathbf{L}} % layer
\newcommand{\ms}{\si{\milli\second}}%
\newcommand{\us}{\si{\micro\second}}%
\newcommand{\timecontext}{T} % time context (cf HOTS) matrice gathering last event times
\newcommand{\current}{I} % post synaptic current
\newcommand{\volt}{u} % membrane potential
\newcommand{\volts}{V} % matrix of membrane potentials
\newcommand{\gain}{\gamma} % homeostatic gain
\newcommand{\simil}{\beta} % similarity value
\newcommand{\Nclass}{N_\text{class}} % number of classes for MLR:
\newcommand{\Nx}{N_\text{X}}
\newcommand{\Ny}{N_\text{Y}}
\newcommand{\Ntime}{N_\text{t}}
\newcommand{\kernel}{K} % convolution kernel
%\newcommand{\kernelind}{\mathbf{k}} % indice of the kernel
\newcommand{\kernelind}{k} % indice of the kernel
\newcommand{\Kx}{K_\text{x}}
\newcommand{\Ky}{K_\text{y}}
\newcommand{\Ktime}{K_\text{t}}
\newcommand{\classiflayer}{\mathbf{C}}
\newcommand{\class}{c} % class k of the MLR
\newcommand{\lrweights}{\theta} % matrix of MLR weights
\newcommand{\lrtrue}{y} % true value of the prediction for MLR
\newcommand{\loss}{J} % cost function for MLR
\newcommand{\softmax}{\sigma}
\newcommand{\actfreq}{f}
\newcommand{\decision}{\hat{y}}
\newcommand{\colorsec}{black}
\newcommand{\colorsubsec}{black}
\newcommand{\speed}{v}
\newcommand{\Nspeed}{N_v}
% Example definitions.
% --------------------
\def\x{{\mathbf x}}
\def\L{{\cal L}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\citep}[1]{\cite{#1}}
\newcommand{\citet}[1]{\cite{#1}}
\newcommand{\seeFig}[1]{see Fig.~\ref{fig:#1}}%{see Figure~\ref{fig:#1}}
%% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
%\usepackage{cite}
\usepackage[usenames,dvipsnames]{xcolor} %The package starts from the basic facilities of the color package, and provides easy driver-independent access to several kinds of color tints, shades, tones, and mixes of arbitrary colors.
\usepackage{tcolorbox} % define boxes that can be customed, has to be loaded after xcolor package
\usepackage{amsmath,amssymb,amsfonts} % amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{float} %allows me to use [H] with figures to insert figures in multicol env (poster)
%\usepackage{ulem} %The package provides an \ul (underline) command which will break over line ends; this technique may be used to replace \em (both in that form and as the \emph command), so as to make output look as if it comes from a typewrite. Caution: with this package, bibliography will make unbreakable and underlined journal titles
\usepackage[switch]{lineno} % add line numbers
\modulolinenumbers[5]
\usepackage{physics} %The  goal  of  this  package  is  to  make  typesetting  equations  for  physics  simpler,  faster,  and  more  human-readable.
%\usepackage{lineno} %Adds line numbers to selected paragraphs with reference possible through the LaTeX \ref and \pageref cross reference mechanism.
\usepackage{hyperref} %The hyperref package is used to handle cross-referencing commands in LaTeX to produce hypertext links in the document
\hypersetup{
    colorlinks = false,
    hidelinks
    %linkbordercolor = {white},
}
%\usepackage{array} %An extended implementation of the array and tabular environments which extends the options for column formats, and provides “programmable” format specifications.
\usepackage{multicol} %Multicol defines a multicols environment which typesets text in multiple columns
%\usepackage{wrapfig} %Allows figures or tables to have text wrapped around them. wrapfig is used when calling multicols
% \usepackage[position=top]{subfig} %The package provides support for the manipulation and reference of small or ‘sub’ figures and tables within a single figure or table environment. Linked to subfloat.
% for units
\usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage%[disable]	% uncomment to hide all the notes
	{todonotes}
\newcommand{\noteAG}[1]{\todo[color=green!40, author=\textbf{Antoine}, inline]{#1}}
\newcommand{\noteLP}[1]{\todo[color=red!20, author=\textbf{Laurent}, inline]{#1}}
\newcommand{\modif}[1]{\todo[color=orange!20, author=\textbf{Proposition}, inline]{#1}}

%\newcommand{\mycaption}[1]{\captionof{figure}{#1}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \usepackage{orcidlink}

%\bibliography{references}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \newcommand{\mycaption}[1]{\caption*{#1}}

\usepackage{titlesec}
% \titlespacing*{<command>}{<left>}{<before-sep>}{<after-sep>}
\titlespacing*{\section}
{0pt}{1.5ex}{0.8ex}
\titlespacing*{\subsection}
{0pt}{0.9ex}{0.4ex}
\titlespacing*{\subsubsection}
{0pt}{0.5ex}{0.3ex}
\titlespacing*{\paragraph}{%
  0pt}{%              left margin
  0.0\baselineskip}{% space before (vertical)
  1em}%               space after (horizontal)

\usepackage{setspace}
%%%-----------------------------------------------------------------
% Title.
% ------
\title{\Title}
%
% Single address.
% ---------------
\name{
\AuthorAG, \AuthorLP
\thanks{\Acknowledgments}
}
\address{\Affiliation}
%
\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
\Abstract
\end{abstract}
%
\begin{keywords}
\Keywords
\end{keywords}
%%%%%%%%% BODY TEXT
\section{Introduction}
\label{sec:intro}
%
In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\cite{abeles1982role}. While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code. Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\cite{riehle1997spike, davis2021spontaneous}, efficient encoding thanks to the use of spike latencies~\cite{perrinet2004coding, gollisch2008rapid} or precise timing in the auditory system~\cite{deweese2002binary, carr1990circuit}. All these findings, and more~\cite{bohte2004evidence}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in biological spike trains.

In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\cite{gutig2006tempotron, guise2014bayesian, zhang2020supervised}. Notably, Izhikevich~\citet{izhikevich2006polychronization} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays. This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
However, most current neuroscience-inspired computer vision algorithms (for instance convolution neural networks) do not make use of this dynamic aspect. A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance. The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations. Yet, flexible and robust event-driven algorithms for classical computer vision tasks, such as motion detection, are still not able to compete with the state-of-the-art of dense computer vision solutions.

In this work, we study the emergence of such spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motionclouds}). We develop a SNN-like method able to learn hetero-synaptic delays to perform motion detection on a synthetic event-based dataset. Because neuromorphic devices are, by design, good candidates to integrate computations with time, we highlight the fact that this event-driven algorithm is transferable to such hardware.
%
\section{Methods}
\label{sec:methods}
%
In a recent study, we have introduced a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{Grimaldi22pami}. To represent temporal relationships between events, this model first builds ``time surfaces'', an % spatio-temporal 2D
image computed using the time difference to the last recorded events~\citet{Lagorce17}. By transforming each event in the stream as a vectorial input, this MLR classifier is able to make a decision for every single event. We have demonstrated on several datasets that it provides with online computations resulting in ultra-fast classification. Additionally, we made a formal bridge between the event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.

Here, we propose to extend such a model to a layer of spiking neurons which include, in addition to synaptic weights, hetero-synaptic delays. In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay as a computational process. The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier. In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these hetero-synaptic delays.
%
\subsection{Task definition: fast motion detection}
%
To test our model, we will quantify its ability to categorize different motions. In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\cite{leon2012motion} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}). In particular, we will set the spatial size to $(\Nx, \Ny)$ and consider a discretization of time with a time step of $1$ such that $\timev \in \mathbb{N}$. Movies' duration are here set to $\Ntime=400$.
This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions:
% of a motion cloud is determined by a constant $k_v$: with a constant speed $\speed$ (see \fig{motionclouds}-(a)): 
%\begin{equation*}
%%\begin{gathered}
%\va{v}_k = %\binom{\cos(\frac{2\pi\cdot k_v}{N_v})}{\sin(\frac{2\pi\cdot k_v}{N_v})}
%  \begin{pmatrix}
%    \cos(2\pi\cdot \frac{k}{\Nclass})\\
%    \sin(2\pi\cdot \frac{k}{\Nclass})
%  \end{pmatrix}
%%\end{gathered}
%\end{equation*}
$
\speed_\kernelind = 
  ( 
    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
  )
$ (\seeFig{motionclouds}-(a) for an illustration).
For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content. This method provides a rich dataset of textured movies for which we know the ground truth for motion.
%
\begin{figure}[h!]%[ht!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/motionclouds.pdf}
    \caption{{\bf Motion detection task.} {\bf (a)} The motion direction represented as the plain red vector, other possible motion directions are represented in light red. {\bf (b)} A screenshot of one generated naturalistic textured stimulus at a specific time. % $\timev_\arank$. 
    {\bf (c)} The corresponding ON (in red) and OFF (in blue) event stream generated from the stimuli on (b) and constituting the input to the spiking neural network.}
    \label{fig:motionclouds}
\end{figure}

To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames. If, on a specific pixel at that specific timestamp, the absolute value of this gradient exceeds a threshold, an event is generated. The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive. This signed threshold value is then subtracted from the residual gradient image. When applied to the whole movie, the event stream is then similar to the output of a neuromorphic camera~\cite{Rasetto2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motionclouds}-(c)). The goal here is to infer the correct motion solely by observing these events.

%
%\begin{figure}[t!]
%    \centering
%    \includegraphics[width=0.90\linewidth]{figures/figure_1a.pdf}
%    \includegraphics[width=0.90\linewidth]{figures/figure_1b.pdf}
%    \includegraphics[width=0.90\linewidth]{figures/figure_1c.pdf}
%	    \caption{The hetero-synaptic delays model applied on synthetic data. {\bf (a)}~The afferent information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs). We highlight them by different colors, an information hidden to a detection model. {\bf (b)}~The model is defined as an assembly of neurons (here for $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay. Note that each afferent may be connected with multiple weights at different delays. The propagation of the afferent information through these delay may generate at each time step a synchronous pattern on a subset of synapses. %These PG patterns were efficiently learned by a logistic regression.
%	    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG pattern at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences. 
%	    \noteAG{clarify the figure - size of the dots for the synaptic weights - how the delay is related to the time we observe the spike trains?}
%	    }
%	    \noteLP{cette figure me parait superflue maintenant, et elle n'eest pas citée dan sle corps du texte - on pourrait simplement l'enlever. c'est la seule aussi (avec le paragraphe en entrée) qui parle de "polychronous groups" - l'enlever permet de simplifier l'ensemble du texte. }
%    \label{fig:model}
%\end{figure}
%
\subsection{Hetero-synaptic delays model}
%
\begin{figure*}[ht!]
    {\centering
    \vspace{-3cm}
    \includegraphics[width=\linewidth]{figures/3D_kernels_oneline.png}
    }
    %\vspace{-.5cm}
    \caption{
    	Representation of the weights for the $8$ learned kernels of the model corresponding to the OFF polarities and selective to the different motion directions
	(because of the symmetry observed between the ON and OFF event streams, kernels are similar for the ON polarities). These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown. One sees positive (excitatory) coefficients for the specific direction of motion and negative (inhibitory) coefficients for all other directions.
	}
    \label{fig:kernels}
\end{figure*} 
%
The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$. 
This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).

We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$. In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron. A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then not only described by synaptic weights connecting to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays. For each neuron $\postsynaddr$, we define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$. Note that a neuron may contact an afferent neuron with multiple different delays.

The corresponding input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible patterns.
For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
$
\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
$.
%
The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\cite{Grimaldi22pami}, in analogy to a spiking Winner-Take-All network~\cite{nessler2013bayesian}. It transforms this list of weights into a probability with the following formula:
$
Pr(k=\postsynaddr \; | \; \timev) =
\frac 1 Z
{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
$ 
where $\mathcal{C}^\postsynaddr(\timev) = \sum
\activeweights^\postsynaddr(t)
$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
%
\subsection{
Temporal
Logistic Regression}
%
In our MLR model with $\Nclass$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function described in the above hetero-synpatic delays model. Such a probability can be computed for each neuron $\postsynaddr$ in association to a specific class. 
From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation. As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$. In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the nonzero weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$. 
Using this dense representation, the counting defined above becomes:
$$
\mathcal{C}^\postsynaddr(a,t)
= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
$$
%
This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different patterns in the event stream.
%
In particular, we may extend the convolution to a 3D convolution 
such that the representation would also benefit from spatial invariance. The use of spatio-temporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\cite{ghosh2022conv3D}.
For that, we design 3D kernels of shape  $(\Kx, \Ky, \Ktime) = (15, 15, 8)$, respectively representing the two spatial dimensions and the range of delays. An example of such kernels is given in Figure~\ref{fig:kernels}.
Computations are performed on spatio-temporal windows, defined by the kernels, sliding around the events, that is, the center of the spatio-temporal window around the current event $\epsilon_\arank$.
%
Finally, the output of the MLR model results in an event with the highest probability class, keeping the same timing as the event as input.
%
The loss function of the MLR model is the binary cross entropy on the output of the classification layer. Simulations are performed thanks to the PyTorch library using gradient descent with Adam (with $2^{12}$ epochs and a learning rate of $10^{-5}$). 
%
\section{Results}
\label{sec:results}
%
\subsection{Kernels learned for motion detection}
%
After training our model, we first observe the weights learned for the different neurons (\seeFig{kernels}). Focusing on the positive weights, a strong selectivity is observed along specific axes for the different kernels. These directions can be easily associated to the direction of motion controlled in the motion clouds. For instance, the third and the seventh kernels show a horizontal selectivity to motion directions.
%
With the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to. This qualitative look at the 3D kernels allows the reader to infer for the $8$ different motion directions used to generate our synthetic event streams.

If one focuses on the interpretation of these kernels in terms of spatio-temporal patterns embedded in the event stream, it can lead to interesting outcomes. In~\cite{Grimaldi22pami}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions. Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spike pattern. Each voxel in the 3D kernels defines a specific timestamp and a specific address. Then, our model is able to detect precise spatio-temporal patterns embedded in the spike train and associated to the different motion directions. The cone shape for the positive weights distribution highlights a loss of precision for longer delays, i.e. events away in the past. For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Figure~\ref{fig:kernels}. 
%
\subsection{Accuracy for the motion detection task}
%
Once our MLR is trained, we obtain 3D kernels corresponding to the weights associated to the hetero-synaptic delays of our layer of spiking neurons and which may be used for detection. We observed that the distribution of the kernels' weights is sparse, with most values near zero. As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the number of spikes times the number of synapses. This scales with the number of nonzero synaptic weights. To  assess the robustness of the classification as a function of the computational load, we will prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold.

In Figure~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations, or active weights, per decision for each neuron of the layer. As a comparison and to account for the gain in performance by using hetero-synaptic delays, we provide the accuracy obtained with a MLR model using 2D time surface (in red) as in~\cite{Grimaldi22pami}. This latter method is based on delays from the last recorded events and uses fewer computations (in our case $15\times15$) than the dense 3D kernels without any pruning ($15\times15\times8$). While less computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
By pruning weights, we observe that the evolution of accuracy as a function of the log percentage of active weights follows a sigmoid curve. Half-saturation level is reached at about $3.5\times 10^{-3}\%$ of active weights, corresponding to a total amount of $6$ computations per decision. Compared to the full kernels, the accuracy of our method is maintained to its top performances when dividing the number of computations by a factor up to about $200$. In this case, the number of computations is greatly reduced compared to~\cite{Grimaldi22pami}, thus demonstrating the efficiency of the presented method. 
%
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/accuracy.pdf}
    \caption{Accuracy as a function of the number of computation load for the hetero-synaptic delays model (in blue) and for a method using 2D time surfaces (in red). The relative computational load (on a log axis) is controlled by changing the percentage of active weights relative to the dense convolution kernel. We observe a similar accuracy than HOTS, yet that our model could achieve a similar accuracy with significantly fewer coefficients.}
    \label{fig:accuracy}
\end{figure}
%
\section{Discussion}

Here, we have introduced a generic SNN using hetero-synaptic delays and shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification. The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification. This shows that we may use the precise timing of spikes to enhance neural computations.

%One advantage of our model is the generality of the possible kernels. However, this can be a limit as this increases the number of free parameters in the learning algorithm. As we observe that the timing of events in natural scene may be slightly variable (in the order of the synaptic time constant of $5~\ms$) an extension of the model would be to include a fixed filtering stage similar to that in the tempotron model~\citep{gutig2006tempotron}. This could be efficiently added as a temporal convolution on $\mathcal{C}^\postsynaddr(t)$.
%\noteAG{C'est pas clair pour moi ce que veut dire ou ce qu'ajoute le paragraphe précédent. La généralité des kernels est données dans le suivant et celui là fait appel à des concepts de filtrage pas très explicités. Aussi la référence au timing des events des scènes naturelles peut être justifié par une référence. }

Note that this supervised learning scheme can be extended to a variety of tasks. It would follow from the emergence of new kernels adapted to this new task after supervised learning.  This constitutes a major advantage over other algorithms which derive event-based algorithms from specific physical rules (see for instance~\citep{Benosman12} for computing the optic flow using the luminance conservation rule). We aim at extending the application of this model on more generic datasets acquired in natural conditions for progressively more complex tasks from motion extraction, optic flow or time-to-contact maps.
%

\bibliographystyle{IEEEbib}
\bibliography{references}
\end{document}

### 2022-08-31 Grimaldi-etal-BiolCybernetics.tex
premier draft

--- 2022-08-31 2022-07-11_Grimaldi-etal-ICIP.tex
@@
-% !TeX TS-program = Lualatex
-% !TeX encoding = UTF-8 Unicode
-% !TeX spellcheck = en
-% !BIB TS-program = bibtex
+%!TeX TS-program = pdfLaTeX
+%!TeX encoding = UTF-8 Unicode
+%!TeX spellcheck = en-US
+%!BIB TS-program = bibtex
 % -*- coding: UTF-8; -*-
 % vim: set fenc=utf-8
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% CONF: https://2022.ieeeicip.org/
-% dates: https://2022.ieeeicip.org/important-dates/
-%
-% paper kit: https://cmsworkshops.com/ICIP2022/papers/paper_kit.php#Templates
-% actual submission : https://cmsworkshops.com/icip2022/Papers/Submission.asp
-%
-% 4 pages + one of references
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%Review #0357
-%
-%*General Comments to Authors*
-%This paper proposes a model for detection of temporal spiking motifs, which can potentially improve the performance of spiking neural networks (SNN).
-The topic is out of my research scope.
-However, I am not sure how this paper is related to ICIP.
-The authors applied the proposed model in motion detection.
-Such motion is captured by imaging or not?
-The overall presentation of the paper is acceptable, but I would suggest the authors to more explicitly represent the contribution of the proposed model and how it is expected to benefit image processing studies.
-%
-%-----------
-%Review #1B91
-%
-%*General Comments to Authors*
-%The authors propose a Spiking Neural Network-like method that consists of a single layer of spiking neurons aiming to learn the heterosexual-synaptic delays in order to detect motion.
-The learning happens in a supervised way as the authors transform the Motion Cloud dataset into spike trains in order to control the direction of motion which is the ground-truth information.
-The authors considered 8 different directions of motion that are assigned to the 8 learnt kernels they introduce in their model.
-According to the experimental results, the model is able to capture the direction of motion which is evident to the selectivity along the specific axes of the different kernels.
-Last but not least, The authors assess their model in terms of the classification accuracy as a function of the computational cost.
-They show that on the one hand, their method is able to highly reduce the computational load by thresholding some of the weights but at the same time achieve the same performance to computationally more greedy algorithms.
-The paper is very interesting and promising as this method can be easily extended to hardware.
-%
-%Some suggestions to improve the presentation of the results are given below:
-%1.
-Figure 2 is a bit confusing.
-First of all, in (b) it's not that evident the different size of the radius but in few cases - maybe another example could be selected.
-In addition, if (a)-(b) and (c) is a single experiment how the delay is related to the time we observe the spike trains?
-Maybe that could be better explained in the text too.
-%
-%Last but not least, are these 8 directions enough to capture every motion?
-Could we use this model in real data or the computational cost would increase as one should consider much higher number of motion directions and as a results much larger set of kernels to be trained?
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                                 %%
+%% Please do not use \input{...} to include other tex files.       %%
+%% Submit your LaTeX manuscript as one .tex document.              %%
+%%                                                                 %%
+%% All additional figures and files should be attached             %%
+%% separately and not embedded in the \TeX\ document itself.       %%
+%%                                                                 %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing
+
+%%=======================================================%%
+%% to print line numbers in the margin use lineno option %%
+%%=======================================================%%
+
+%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+
+%%======================================================%%
+%% to compile with pdflatex/xelatex use pdflatex option %%
+%%======================================================%%
+
+%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+
+%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+%%\documentclass[sn-mathphys]{sn-jnl}% Math and Physical Sciences Reference Style
+%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
+%%\documentclass[sn-vancouver]{sn-jnl}% Vancouver Reference Style
+%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style
+%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
+%%\documentclass[sn-standardnature]{sn-jnl}% Standard Nature Portfolio Reference Style
+%%\documentclass[default]{sn-jnl}% Default
+%%\documentclass[default,iicol]{sn-jnl}% Default with double column layout
+\documentclass[iicol]{sn-jnl}
+
+%%%% Standard Packages
+%%<additional latex packages if required can be included here>
+\usepackage[lofdepth,lotdepth,position=top]{subfig}
+\usepackage{tabularx}
+%%%%
+
+%%%%%=============================================================================%%%%
+%%%%  Remarks: This template is provided to aid authors with the preparation
+%%%%  of original research articles intended for submission to journals published 
+%%%%  by Springer Nature. The guidance has been prepared in partnership with 
+%%%%  production teams to conform to Springer Nature technical requirements. 
+%%%%  Editorial and presentation requirements differ among journal portfolios and 
+%%%%  research disciplines. You may find sections in this template are irrelevant 
+%%%%  to your work and are empowered to omit any such section if allowed by the 
+%%%%  journal you intend to submit to. The submission guidelines and policies 
+%%%%  of the journal take precedence. A detailed User Manual is available in the 
+%%%%  template package for technical guidance.
+%%%%%=============================================================================%%%%
+
+\jyear{2022}%
+
+%% as per the requirement new theorem styles can be included as shown below
+\theoremstyle{thmstyleone}%
+\newtheorem{theorem}{Theorem}%  meant for continuous numbers
+%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
+%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
+\newtheorem{proposition}[theorem]{Proposition}% 
+%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
+
+\theoremstyle{thmstyletwo}%
+\newtheorem{example}{Example}%
+\newtheorem{remark}{Remark}%
+
+\theoremstyle{thmstylethree}%
+\newtheorem{definition}{Definition}%
+
+\raggedbottom
+%%\unnumbered% uncomment this for unnumbered level heads
+
+%%\unnumbered% uncomment this for unnumbered level heads%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \newcommand{\Title}{
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass{article}
-\usepackage{spconf,amsmath,graphicx}
-
 \usepackage[utf8]{inputenc}
@@
 %\usepackage{cite}
-\usepackage[usenames,dvipsnames]{xcolor} %The package starts from the basic facilities of the color package, and provides easy driver-independent access to several kinds of color tints, shades, tones, and mixes of arbitrary colors.
-\usepackage{tcolorbox} % define boxes that can be customed, has to be loaded after xcolor package
+%\usepackage[usenames,dvipsnames]{xcolor} %The package starts from the basic facilities of the color package, and provides easy driver-independent access to several kinds of color tints, shades, tones, and mixes of arbitrary colors.
+%\usepackage{tcolorbox} % define boxes that can be customed, has to be loaded after xcolor package
 \usepackage{amsmath,amssymb,amsfonts} % amsmath and amssymb packages, useful for mathematical formulas and symbols
@@
 Caution: with this package, bibliography will make unbreakable and underlined journal titles
-\usepackage[switch]{lineno} % add line numbers
-\modulolinenumbers[5]
+%\usepackage[switch]{lineno} % add line numbers
+%\modulolinenumbers[5]
 \usepackage{physics} %The  goal  of  this  package  is  to  make  typesetting  equations  for  physics  simpler,  faster,  and  more  human-readable.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%\usepackage%[disable]	% uncomment to hide all the notes
+%	{todonotes}
+%\newcommand{\noteAG}[1]{\todo[color=green!40, author=\textbf{Antoine}, inline]{#1}}
+%\newcommand{\noteLP}[1]{\todo[color=red!20, author=\textbf{Laurent}, inline]{#1}}
+%\newcommand{\modif}[1]{\todo[color=orange!20, author=\textbf{Proposition}, inline]{#1}}
 
-\usepackage%[disable]	% uncomment to hide all the notes
-	{todonotes}
-\newcommand{\noteAG}[1]{\todo[color=green!40, author=\textbf{Antoine}, inline]{#1}}
-\newcommand{\noteLP}[1]{\todo[color=red!20, author=\textbf{Laurent}, inline]{#1}}
-\newcommand{\modif}[1]{\todo[color=orange!20, author=\textbf{Proposition}, inline]{#1}}
-
-%\newcommand{\mycaption}[1]{\captionof{figure}{#1}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \usepackage{orcidlink}
-
-%\bibliography{references}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \newcommand{\mycaption}[1]{\caption*{#1}}
-
-\usepackage{titlesec}
-% \titlespacing*{<command>}{<left>}{<before-sep>}{<after-sep>}
-\titlespacing*{\section}
-{0pt}{1.5ex}{0.8ex}
-\titlespacing*{\subsection}
-{0pt}{0.9ex}{0.4ex}
-\titlespacing*{\subsubsection}
-{0pt}{0.5ex}{0.3ex}
-\titlespacing*{\paragraph}{%
-  0pt}{%              left margin
-  0.0\baselineskip}{% space before (vertical)
-  1em}%               space after (horizontal)
-
-\usepackage{setspace}
+\begin{document}
 %%%-----------------------------------------------------------------
@@
 % ---------------
-\name{
-\AuthorAG, \AuthorLP
-\thanks{\Acknowledgments}
-}
-\address{\Affiliation}
+
+\author*[1]{\fnm{Antoine} \sur{Grimaldi}}\email{antoine.grimaldi@univ-amu.fr}
+
+\author[2]{\fnm{Amélie} \sur{Gruel}}\email{amelie.gruel@univ-cotedazur.fr}
+
+\author[2]{\fnm{Jean} \sur{Martinet}}\email{jean.martinet@univ-cotedazur.fr}
+
+\author[1]{\fnm{Laurent} \sur{Perrinet}}\email{laurent.perrinet@univ-amu.fr}
+
+\affil[1]{\orgdiv{Department}, \orgname{\Affiliation}, \orgaddress{\street{\Street}, \city{Marseille}, \postcode{13005}, \country{France}}}
+\affil[2]{\orgdiv{SPARKS}, \orgname{Université Côte d'Azur, CNRS, I3S}, \orgaddress{\street{2000 Rte des Lucioles}, \city{Sophia-Antipolis}, \postcode{06900}, \country{France}}}
+
 %
-\begin{document}
 %\ninept
 %
+%
+\abstract{\Abstract}
+%
+\keywords{\Keywords}
+%
 \maketitle
-%
-\begin{abstract}
-\Abstract
-\end{abstract}
-%
-\begin{keywords}
-\Keywords
-\end{keywords}
 %%%%%%%%% BODY TEXT
@@
 $
-\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
+toto
+%\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
 $.
@@
 $
-Pr(k=\postsynaddr \; | \; \timev) =
+toto
+%Pr(k=\postsynaddr \; | \; \timev) =
 \frac 1 Z
@@
 
-\bibliographystyle{IEEEbib}
-\bibliography{references}
+\bibliographystyle{apalike}
+\bibliography{Grimaldi-etal-BiolCybernetics}
 \end{document}

### 2022-08-31 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
 % vim: set fenc=utf-8
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\Title}{
+Learning hetero-synaptic delays for motion detection in a single layer of spiking neurons
+%FAST MOTION DETECTION USING HETERO-SYNAPTIC DELAYS IN A LAYER OF SPIKING NEURONS
+}
+\newcommand{\AuthorAG}{Antoine Grimaldi}
+\newcommand{\AuthorCB}{Camille Besnainou}
+\newcommand{\AuthorHL}{Hugo Ladret}
+\newcommand{\AuthorLP}{Laurent U Perrinet}
+\newcommand{\EmailLP}{laurent.perrinet@univ-amu.fr}
+\newcommand{\EmailAG}{antoine.grimaldi@univ-amu.fr}
+\newcommand{\EmailHL}{hugo.ladret@univ-amu.fr}
+\newcommand{\orcidLP}{0000-0002-9536-010X}
+\newcommand{\orcidHL}{0000-0001-7999-3751}
+\newcommand{\orcidAG}{0000-0002-3107-4788}
+\newcommand{\Address}{Institut de Neurosciences de la Timone (UMR 7289); Aix Marseille Univ, CNRS; Marseille, France}%
+\newcommand{\Affiliation}{Institut de Neurosciences de la Timone (UMR 7289); Aix Marseille Univ, CNRS;}%
+\newcommand{\CityLP}{Marseille, France}%
+\newcommand{\WebsiteLP}{https://laurentperrinet.github.io}%
+\newcommand{\Abstract}{
+The response of a biological neuron depends on the precise timing of afferent spikes.
+This temporal aspect of the neuronal code is essential in understanding information processing in neurobiology and applies particularly well to the output of neuromorphic hardware such as event-based cameras.
+However, most artificial neuronal models do not take advantage of this minute temporal dimension and here, we develop a model for the efficient detection of temporal spiking motifs based on a layer of neurons with hetero-synaptic delays.
+Indeed, the variety of synaptic delays on the dendritic tree allows to synchronize synaptic inputs as they reach the basal dendritic tree.
+We show this can be formalized as a time-invariant logistic regression which can be trained using labelled data.
+We apply this model to solve the specific computer vision problem of motion detection, and demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of event-based cameras.
+In particular, we quantify how its accuracy can vary with the total computational load.
+This end-to-end event-driven computational brick could help improve the performance of future Spiking Neural Network (SNN) algorithms and their prospective use in neuromorphic chips.
+}
+\newcommand{\Keywords}{time code, event-based computations, spiking neural networks, motion detection, efficient coding, logistic regression
+}
+\newcommand{\Acknowledgments}{This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}'').
+%HL and LP received funding from the ANR project  N° ANR-20-CE23-0021 (''\href{https://laurentperrinet.github.io/grant/anr-anr/}{AgileNeuroBot}'').
+HL received funding from a University of Montreal Artificial Intelligence scholarship.
+For the purpose of open access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript version arising from this submission.
+This research was funded, in whole or in part, by [Organisation name, Grant #].
+A CC-BY public copyright license has been applied by the authors to the present document and will be applied to all subsequent versions up to the Author Accepted Manuscript arising from this submission, in accordance with the grant’s open access conditions.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% NOTATIONS
+% Pixel world
+\newcommand{\presynaddr}{a} % pre address
+\newcommand{\postsynaddr}{n} % post address
+\newcommand{\numevent}{N_{ev}} % total number of events
+\newcommand{\presynaddrspace}{\mathcal{A}} %presynaptic address space
+\newcommand{\postsynaddrspace}{\mathcal{B}} %postsynaptic address space
+\newcommand{\Npol}{N_\text{p}} % number of polarity
+\newcommand{\Nneuron}{N_\text{n}} % number of output neurons in the layer
+\newcommand{\arank}{r} % address index
+\newcommand{\bias}{b} % bias for the MLR model
+\newcommand{\synapse}{\sigma} % synapse
+\newcommand{\synapticweight}{w} % synaptic weight
+\newcommand{\synapticdelay}{\delta} % synaptic delay
+\newcommand{\ranksyn}{s} % synapse index
+\newcommand{\Nsyn}{N_{s}} % total number of synapses
+\newcommand{\activeweights}{\mathcal{W}} 
+\newcommand{\timev}{t} % time
+\newcommand{\polev}{p} % polarity
+\newcommand{\event}{\epsilon} % event
+\newcommand{\eventstream}{\xi} % stream of events
+\newcommand{\TS}{S} % time surface
+\newcommand{\neuron}{\mathbf{n}} % neuron in the SNN (defined by the spatial position and the channel)
+\newcommand{\postneuron}{\mathbf{m}} % post synaptic neuron in the SNN (defined by the spatial position and the kernel)
+\newcommand{\channel}{\mathbf{p}} % channel
+\newcommand{\layer}{\mathbf{L}} % layer
+\newcommand{\ms}{\si{\milli\second}}%
+\newcommand{\us}{\si{\micro\second}}%
+\newcommand{\timecontext}{T} % time context (cf HOTS) matrice gathering last event times
+\newcommand{\current}{I} % post synaptic current
+\newcommand{\volt}{u} % membrane potential
+\newcommand{\volts}{V} % matrix of membrane potentials
+\newcommand{\gain}{\gamma} % homeostatic gain
+\newcommand{\simil}{\beta} % similarity value
+\newcommand{\Nclass}{N_\text{class}} % number of classes for MLR:
+\newcommand{\Nx}{N_\text{X}}
+\newcommand{\Ny}{N_\text{Y}}
+\newcommand{\Ntime}{N_\text{t}}
+\newcommand{\kernel}{K} % convolution kernel
+%\newcommand{\kernelind}{\mathbf{k}} % indice of the kernel
+\newcommand{\kernelind}{k} % indice of the kernel
+\newcommand{\Kx}{K_\text{x}}
+\newcommand{\Ky}{K_\text{y}}
+\newcommand{\Ktime}{K_\text{t}}
+\newcommand{\classiflayer}{\mathbf{C}}
+\newcommand{\class}{c} % class k of the MLR
+\newcommand{\lrweights}{\theta} % matrix of MLR weights
+\newcommand{\lrtrue}{y} % true value of the prediction for MLR
+\newcommand{\loss}{J} % cost function for MLR
+\newcommand{\softmax}{\sigma}
+\newcommand{\actfreq}{f}
+\newcommand{\decision}{\hat{y}}
+\newcommand{\colorsec}{black}
+\newcommand{\colorsubsec}{black}
+\newcommand{\speed}{v}
+\newcommand{\Nspeed}{N_v}
+% Example definitions.
+% --------------------
+\def\x{{\mathbf x}}
+\def\L{{\cal L}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%%\unnumbered% uncomment this for unnumbered level heads%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\Title}{
-Learning hetero-synaptic delays for motion detection in a single layer of spiking neurons
-%FAST MOTION DETECTION USING HETERO-SYNAPTIC DELAYS IN A LAYER OF SPIKING NEURONS
-}
-\newcommand{\AuthorAG}{Antoine Grimaldi}
-\newcommand{\AuthorCB}{Camille Besnainou}
-\newcommand{\AuthorHL}{Hugo Ladret}
-\newcommand{\AuthorLP}{Laurent U Perrinet}
-\newcommand{\EmailLP}{laurent.perrinet@univ-amu.fr}
-\newcommand{\EmailAG}{antoine.grimaldi@univ-amu.fr}
-\newcommand{\EmailHL}{hugo.ladret@univ-amu.fr}
-\newcommand{\orcidLP}{0000-0002-9536-010X}
-\newcommand{\orcidHL}{0000-0001-7999-3751}
-\newcommand{\orcidAG}{0000-0002-3107-4788}
-\newcommand{\Address}{Institut de Neurosciences de la Timone (UMR 7289); Aix Marseille Univ, CNRS; Marseille, France}%
-\newcommand{\Affiliation}{Institut de Neurosciences de la Timone (UMR 7289); Aix Marseille Univ, CNRS;}%
-\newcommand{\CityLP}{Marseille, France}%
-\newcommand{\WebsiteLP}{https://laurentperrinet.github.io}%
-\newcommand{\Abstract}{
-The response of a biological neuron depends on the precise timing of afferent spikes.
-This temporal aspect of the neuronal code is essential in understanding information processing in neurobiology and applies particularly well to the output of neuromorphic hardware such as event-based cameras.
-However, most artificial neuronal models do not take advantage of this minute temporal dimension and here, we develop a model for the efficient detection of temporal spiking motifs based on a layer of neurons with hetero-synaptic delays.
-Indeed, the variety of synaptic delays on the dendritic tree allows to synchronize synaptic inputs as they reach the basal dendritic tree.
-We show this can be formalized as a time-invariant logistic regression which can be trained using labelled data.
-We apply this model to solve the specific computer vision problem of motion detection, and demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of event-based cameras.
-In particular, we quantify how its accuracy can vary with the total computational load.
-This end-to-end event-driven computational brick could help improve the performance of future Spiking Neural Network (SNN) algorithms and their prospective use in neuromorphic chips.
-}
-\newcommand{\Keywords}{time code, event-based computations, spiking neural networks, motion detection, efficient coding, logistic regression
-}
-\newcommand{\Acknowledgments}{This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}'').
-%HL and LP received funding from the ANR project  N° ANR-20-CE23-0021 (''\href{https://laurentperrinet.github.io/grant/anr-anr/}{AgileNeuroBot}'').
-HL received funding from a University of Montreal Artificial Intelligence scholarship.
-For the purpose of open access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript version arising from this submission.
-This research was funded, in whole or in part, by [Organisation name, Grant #].
-A CC-BY public copyright license has been applied by the authors to the present document and will be applied to all subsequent versions up to the Author Accepted Manuscript arising from this submission, in accordance with the grant’s open access conditions.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage[utf8]{inputenc}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% NOTATIONS
-% Pixel world
-\newcommand{\presynaddr}{a} % pre address
-\newcommand{\postsynaddr}{n} % post address
-\newcommand{\numevent}{N_{ev}} % total number of events
-\newcommand{\presynaddrspace}{\mathcal{A}} %presynaptic address space
-\newcommand{\postsynaddrspace}{\mathcal{B}} %postsynaptic address space
-\newcommand{\Npol}{N_\text{p}} % number of polarity
-\newcommand{\Nneuron}{N_\text{n}} % number of output neurons in the layer
-\newcommand{\arank}{r} % address index
-\newcommand{\bias}{b} % bias for the MLR model
-\newcommand{\synapse}{\sigma} % synapse
-\newcommand{\synapticweight}{w} % synaptic weight
-\newcommand{\synapticdelay}{\delta} % synaptic delay
-\newcommand{\ranksyn}{s} % synapse index
-\newcommand{\Nsyn}{N_{s}} % total number of synapses
-\newcommand{\activeweights}{\mathcal{W}} 
-\newcommand{\timev}{t} % time
-\newcommand{\polev}{p} % polarity
-\newcommand{\event}{\epsilon} % event
-\newcommand{\eventstream}{\xi} % stream of events
-\newcommand{\TS}{S} % time surface
-\newcommand{\neuron}{\mathbf{n}} % neuron in the SNN (defined by the spatial position and the channel)
-\newcommand{\postneuron}{\mathbf{m}} % post synaptic neuron in the SNN (defined by the spatial position and the kernel)
-\newcommand{\channel}{\mathbf{p}} % channel
-\newcommand{\layer}{\mathbf{L}} % layer
-\newcommand{\ms}{\si{\milli\second}}%
-\newcommand{\us}{\si{\micro\second}}%
-\newcommand{\timecontext}{T} % time context (cf HOTS) matrice gathering last event times
-\newcommand{\current}{I} % post synaptic current
-\newcommand{\volt}{u} % membrane potential
-\newcommand{\volts}{V} % matrix of membrane potentials
-\newcommand{\gain}{\gamma} % homeostatic gain
-\newcommand{\simil}{\beta} % similarity value
-\newcommand{\Nclass}{N_\text{class}} % number of classes for MLR:
-\newcommand{\Nx}{N_\text{X}}
-\newcommand{\Ny}{N_\text{Y}}
-\newcommand{\Ntime}{N_\text{t}}
-\newcommand{\kernel}{K} % convolution kernel
-%\newcommand{\kernelind}{\mathbf{k}} % indice of the kernel
-\newcommand{\kernelind}{k} % indice of the kernel
-\newcommand{\Kx}{K_\text{x}}
-\newcommand{\Ky}{K_\text{y}}
-\newcommand{\Ktime}{K_\text{t}}
-\newcommand{\classiflayer}{\mathbf{C}}
-\newcommand{\class}{c} % class k of the MLR
-\newcommand{\lrweights}{\theta} % matrix of MLR weights
-\newcommand{\lrtrue}{y} % true value of the prediction for MLR
-\newcommand{\loss}{J} % cost function for MLR
-\newcommand{\softmax}{\sigma}
-\newcommand{\actfreq}{f}
-\newcommand{\decision}{\hat{y}}
-\newcommand{\colorsec}{black}
-\newcommand{\colorsubsec}{black}
-\newcommand{\speed}{v}
-\newcommand{\Nspeed}{N_v}
-% Example definitions.
-% --------------------
-\def\x{{\mathbf x}}
-\def\L{{\cal L}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand{\citep}[1]{\cite{#1}}
-\newcommand{\citet}[1]{\cite{#1}}
+%%\unnumbered% uncomment this for unnumbered level heads
+%\newcommand{\citep}[1]{\cite{#1}}
+%\newcommand{\citet}[1]{\cite{#1}}
 \newcommand{\seeFig}[1]{see Fig.~\ref{fig:#1}}%{see Figure~\ref{fig:#1}}
@@
 \usepackage{amsmath,amssymb,amsfonts} % amsmath and amssymb packages, useful for mathematical formulas and symbols
-\usepackage{float} %allows me to use [H] with figures to insert figures in multicol env (poster)
+%%\usepackage{float} %allows me to use [H] with figures to insert figures in multicol env (poster)
 %\usepackage{ulem} %The package provides an \ul (underline) command which will break over line ends; this technique may be used to replace \em (both in that form and as the \emph command), so as to make output look as if it comes from a typewrite.
@@
 %\usepackage{lineno} %Adds line numbers to selected paragraphs with reference possible through the LaTeX \ref and \pageref cross reference mechanism.
-\usepackage{hyperref} %The hyperref package is used to handle cross-referencing commands in LaTeX to produce hypertext links in the document
-\hypersetup{
-    colorlinks = false,
-    hidelinks
-    %linkbordercolor = {white},
-}
+%\usepackage{hyperref} %The hyperref package is used to handle cross-referencing commands in LaTeX to produce hypertext links in the document
+%\hypersetup{
+%    colorlinks = false,
+%    hidelinks
+%    %linkbordercolor = {white},
+%}
 %\usepackage{array} %An extended implementation of the array and tabular environments which extends the options for column formats, and provides “programmable” format specifications.
-\usepackage{multicol} %Multicol defines a multicols environment which typesets text in multiple columns
-%\usepackage{wrapfig} %Allows figures or tables to have text wrapped around them. wrapfig is used when calling multicols
+%\usepackage{multicol} %Multicol defines a multicols environment which typesets text in multiple columns
+%%\usepackage{wrapfig} %Allows figures or tables to have text wrapped around them. wrapfig is used when calling multicols
 % \usepackage[position=top]{subfig} %The package provides support for the manipulation and reference of small or ‘sub’ figures and tables within a single figure or table environment.
@@
 
-\affil[1]{\orgdiv{Department}, \orgname{\Affiliation}, \orgaddress{\street{\Street}, \city{Marseille}, \postcode{13005}, \country{France}}}
+\affil[1]{\orgdiv{Department}, \orgname{\Affiliation}, \orgaddress{\street{Street}, \city{Marseille}, \postcode{13005}, \country{France}}}
 \affil[2]{\orgdiv{SPARKS}, \orgname{Université Côte d'Azur, CNRS, I3S}, \orgaddress{\street{2000 Rte des Lucioles}, \city{Sophia-Antipolis}, \postcode{06900}, \country{France}}}
@@
 %
-\abstract{\Abstract}
+%\abstract{
+%Abstract
+%}
 %
@@
 
+\backmatter
+
+\bmhead{Acknowledgments}
+
+% Acknowledgement hidden for review.
+This work was supported by the European Union’s ERA-NET CHIST-ERA 2018 research and innovation programme under grant agreement ANR-19-CHR3-0008.
+
+% Nef acknowledgements 
+The authors are grateful to the OPAL infrastructure from Université Côte d'Azur for providing resources and support.
+
+\section*{Declarations}
+
+\textcolor{purple}{TBD - What the template says about it:}
+
+\textcolor{purple}{Some journals require declarations to be submitted in a standardised format.
+Please check the Instructions for Authors of the journal to which you are submitting to see if you need to complete this section.
+If yes, your manuscript must contain the following sections under the heading `Declarations': Funding; Conflict of interest/Competing interests (check journal-specific guidelines for which heading to use); Ethics approval; Consent to participate; Consent for publication; Availability of data and materials; Code availability; Authors' contributions.
+}
+
+\textcolor{purple}{If any of the sections are not relevant to your manuscript, please include the heading and write `Not applicable' for that section. }
+
 \bibliographystyle{apalike}

### 2022-09-01 Grimaldi-etal-BiolCybernetics.tex
abstract

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%Call for papers: Biological Cybernetics Special Issue: What can Computer Vision learn from Visual Neuroscience?
+%
+%Computer vision still struggles to solve many problems as quickly and accurately as our brain does, despite its tremendous progress using deep learning over the last decade.
+Current challenges include geometry understanding, dynamic scene understanding, few shot learning of novel objects, anomaly and counterfeit detection, stable reproduction of imaginary visuals, and more.
+The brain achieves all these on a low computational cost and power budget (i.e., tens of watts), whereas the state-of-the-art computer vision models require multiple graphics processors consuming kilowatts of power.
+Hence, new inspirations from the brain could make computer vision more efficient, robust, and capable of continuous learning and adaptation.
+Although visual neuroscience literature provides knowledge about neuronal selectivities and early processing of visual inputs, computational modelling of neuronal and network behaviours with some degree of biological constraints and characteristics is an important tool to address many of the challenges facing computer vision.
+With new neuroscientific insights about deep cortical vision processing using the latest stimulation and recording technologies, it is important that we strive to understand the underlying computations and learning mechanisms, so that artificial vision systems with similar efficiency can be developed.
+This special issue invites original research and review articles related to topics in biological vision that can potentially benefit computer vision systems.
+The following is a non-exhaustive list of topics in biological vision that can also benefit computer vision systems.
+%
+%● Active vision’s role in visual search, scene understanding, social interactions, etc.
+%
+%● Learning in the visual system. Learning in biology is continual, few-shot, and adversarially robust.
+%
+%● The roles of recurrent and top-down connections in the visual cortex.
+%
+%● Spike based spatiotemporal processing and its implications for neuromorphic vision.
+%
+%● Motion perception in dynamic environments.
+%
+%● Neural coding schemes in the visual system (e.g., sparse coding, predictive coding, and temporal coding.)
+%
+%● The roles of attention mechanisms in biological vision.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \newcommand{\Title}{
@@
 }
-\newcommand{\AuthorAG}{Antoine Grimaldi}
-\newcommand{\AuthorCB}{Camille Besnainou}
-\newcommand{\AuthorHL}{Hugo Ladret}
-\newcommand{\AuthorLP}{Laurent U Perrinet}
+\newcommand{\FirstAG}{Antoine}
+\newcommand{\LastAG}{Grimaldi}
+\newcommand{\AuthorAG}{\FirstAG \LastAG}
+\newcommand{\FirstLP}{Laurent U}
+\newcommand{\LastLP}{Perrinet}
+\newcommand{\AuthorLP}{\FirstLP \LastLP}
 \newcommand{\EmailLP}{laurent.perrinet@univ-amu.fr}
 \newcommand{\EmailAG}{antoine.grimaldi@univ-amu.fr}
-\newcommand{\EmailHL}{hugo.ladret@univ-amu.fr}
 \newcommand{\orcidLP}{0000-0002-9536-010X}
-\newcommand{\orcidHL}{0000-0001-7999-3751}
 \newcommand{\orcidAG}{0000-0002-3107-4788}
-\newcommand{\Address}{Institut de Neurosciences de la Timone (UMR 7289); Aix Marseille Univ, CNRS; Marseille, France}%
-\newcommand{\Affiliation}{Institut de Neurosciences de la Timone (UMR 7289); Aix Marseille Univ, CNRS;}%
-\newcommand{\CityLP}{Marseille, France}%
+\newcommand{\Department}{Institut de Neurosciences de la Timone (UMR 7289)}%
+\newcommand{\Affiliation}{Aix Marseille Univ, CNRS}%
+\newcommand{\Street}{27 boulevard Jean Moulin}%
+\newcommand{\City}{Marseille}%
+\newcommand{\Country}{France}%
 \newcommand{\WebsiteLP}{https://laurentperrinet.github.io}%
@@
 %%\unnumbered% uncomment this for unnumbered level heads
-%\newcommand{\citep}[1]{\cite{#1}}
-%\newcommand{\citet}[1]{\cite{#1}}
+\newcommand{\citep}[1]{\cite{#1}}
+\newcommand{\citet}[1]{\cite{#1}}
 \newcommand{\seeFig}[1]{see Fig.~\ref{fig:#1}}%{see Figure~\ref{fig:#1}}
@@
 
-\author*[1]{\fnm{Antoine} \sur{Grimaldi}}\email{antoine.grimaldi@univ-amu.fr}
+\author*{\fnm{Antoine} \sur{Grimaldi}}\email{antoine.grimaldi@univ-amu.fr}
 
-\author[2]{\fnm{Amélie} \sur{Gruel}}\email{amelie.gruel@univ-cotedazur.fr}
+\author{\fnm{Laurent} \sur{Perrinet}}\email{laurent.perrinet@univ-amu.fr}
 
-\author[2]{\fnm{Jean} \sur{Martinet}}\email{jean.martinet@univ-cotedazur.fr}
+\affil{\orgdiv{\Department}, \orgname{\Affiliation}\\ \orgaddress{\street{\Street}, \city{\City}, \postcode{13005}, \country{\Country}}}
 
-\author[1]{\fnm{Laurent} \sur{Perrinet}}\email{laurent.perrinet@univ-amu.fr}
+%\author*[1]{\fnm{Antoine} \sur{Grimaldi}}\email{antoine.grimaldi@univ-amu.fr}
+%
+%\author[2]{\fnm{Amélie} \sur{Gruel}}\email{amelie.gruel@univ-cotedazur.fr}
+%
+%\author[2]{\fnm{Jean} \sur{Martinet}}\email{jean.martinet@univ-cotedazur.fr}
+%
+%\author[1]{\fnm{Laurent} \sur{Perrinet}}\email{laurent.perrinet@univ-amu.fr}
+%
+%\affil[1]{\orgdiv{Department}, \orgname{\Affiliation}, \orgaddress{\street{Street}, \city{Marseille}, \postcode{13005}, \country{France}}}
+%\affil[2]{\orgdiv{SPARKS}, \orgname{Université Côte d'Azur, CNRS, I3S}, \orgaddress{\street{2000 Rte des Lucioles}, \city{Sophia-Antipolis}, \postcode{06900}, \country{France}}}
 
-\affil[1]{\orgdiv{Department}, \orgname{\Affiliation}, \orgaddress{\street{Street}, \city{Marseille}, \postcode{13005}, \country{France}}}
-\affil[2]{\orgdiv{SPARKS}, \orgname{Université Côte d'Azur, CNRS, I3S}, \orgaddress{\street{2000 Rte des Lucioles}, \city{Sophia-Antipolis}, \postcode{06900}, \country{France}}}
+%%==================================%%
+%% sample for unstructured abstract %%
+%%==================================%%
+
+\abstract{
+\Abstract
+}
 
@@
 %
-%
-%\abstract{
-%Abstract
-%}
+%\abstract{Abstract}
 %

### 2022-09-13 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
bib
hetero-synaptic

@@
 \newcommand{\Title}{
-Learning hetero-synaptic delays for motion detection in a single layer of spiking neurons
-%FAST MOTION DETECTION USING HETERO-SYNAPTIC DELAYS IN A LAYER OF SPIKING NEURONS
+Learning hetero-synaptic delays for fast motion detection in a single layer of spiking neurons
+}
+\newcommand{\ShortTitle}{
+Spiking neurons with hetero-synaptic delays for motion detection
 }
@@
 %%\documentclass[sn-standardnature]{sn-jnl}% Standard Nature Portfolio Reference Style
-%%\documentclass[default]{sn-jnl}% Default
+\documentclass[default]{sn-jnl}% Default
 %%\documentclass[default,iicol]{sn-jnl}% Default with double column layout
-\documentclass[iicol]{sn-jnl}
+%\documentclass[iicol]{sn-jnl}
 
@@
 %%\unnumbered% uncomment this for unnumbered level heads
-\newcommand{\citep}[1]{\cite{#1}}
-\newcommand{\citet}[1]{\cite{#1}}
+%\newcommand{\citep}[1]{\cite{#1}}
+%\newcommand{\citet}[1]{\cite{#1}}
 \newcommand{\seeFig}[1]{see Fig.~\ref{fig:#1}}%{see Figure~\ref{fig:#1}}
@@
 % ------
-\title{\Title}
+\title[\ShortTitle]{\Title}
 %
@@
 %
-In a recent study, we have introduced a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{Grimaldi22pami}.
-To represent temporal relationships between events, this model first builds ``time surfaces'', an % spatio-temporal 2D
-image computed using the time difference to the last recorded events~\citet{Lagorce17}.
-By transforming each event in the stream as a vectorial input, this MLR classifier is able to make a decision for every single event.
-We have demonstrated on several datasets that it provides with online computations resulting in ultra-fast classification.
-Additionally, we made a formal bridge between the event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.
+In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{Grimaldi22pami} which extended the HOTS model~\citet{Lagorce17}.
+This HOTS model represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events.
+By transforming each event in the stream as a vectorial input, our MLR classifier is able to make a decision for every single event, that is in an online fashion.
+We have demonstrated on several datasets that it provides with efficient computations resulting in ultra-fast categorization.
+Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.
 
-Here, we propose to extend such a model to a layer of spiking neurons which include, in addition to synaptic weights, hetero-synaptic delays.
+Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
+Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, hetero-synaptic delays.
 In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay as a computational process.
+As a consequence, this extends our previous model by encoding spiking patterns as 3D images.
 The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier.

### 2022-09-13 Grimaldi-etal-BiolCybernetics.tex
task

@@
 %
+%
+We expect our model to be able to efficiently detect temporal patterns and here we will challenge it by defining a realistic task of motion detection in a stream generated by a natural scene with known ground truth motion.
+%
 \subsection{Task definition: fast motion detection}
 %
-To test our model, we will quantify its ability to categorize different motions.
-In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\cite{leon2012motion} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}).
+Taking analogy with biological conditions, a common use case for motion detection is to estimate the motion generated on a sensor by saccade-like movements.
+Saccades are rapid movements of the eyeball which allow to reposition the axis of view in a short period.
+In humans, these are particularly rapid
+To test our model, we will quantify its ability to categorize different motions. In that order, we will generate a natural
+
+use a set of synthetic visual stimuli, \textit{Motion Clouds}~\cite{leon2012motion} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}).
 In particular, we will set the spatial size to $(\Nx, \Ny)$ and consider a discretization of time with a time step of $1$ such that $\timev \in \mathbb{N}$.
@@
 
-%
-%\begin{figure}[t!]
-%    \centering
-%    \includegraphics[width=0.90\linewidth]{figures/figure_1a.pdf}
-%    \includegraphics[width=0.90\linewidth]{figures/figure_1b.pdf}
-%    \includegraphics[width=0.90\linewidth]{figures/figure_1c.pdf}
-%	    \caption{The hetero-synaptic delays model applied on synthetic data.
-{\bf (a)}~The afferent information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs).
-We highlight them by different colors, an information hidden to a detection model.
-{\bf (b)}~The model is defined as an assembly of neurons (here for $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
-Note that each afferent may be connected with multiple weights at different delays.
-The propagation of the afferent information through these delay may generate at each time step a synchronous pattern on a subset of synapses.
-%These PG patterns were efficiently learned by a logistic regression.
-%	    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG pattern at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
-%	    \noteAG{clarify the figure - size of the dots for the synaptic weights - how the delay is related to the time we observe the spike trains?}
-%	    }
-%	    \noteLP{cette figure me parait superflue maintenant, et elle n'eest pas citée dan sle corps du texte - on pourrait simplement l'enlever.
-c'est la seule aussi (avec le paragraphe en entrée) qui parle de "polychronous groups" - l'enlever permet de simplifier l'ensemble du texte.
-}
-%    \label{fig:model}
-%\end{figure}
-%
 \subsection{Hetero-synaptic delays model}
@@
 %
+
+\subsection{Testing with natural-like textures}
+%
+To test our model, we will quantify its ability to categorize different motions.
+In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\cite{leon2012motion} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}).
+In particular, we will set the spatial size to $(\Nx, \Ny)$ and consider a discretization of time with a time step of $1$ such that $\timev \in \mathbb{N}$.
+Movies' duration are here set to $\Ntime=400$.
+This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions:
+% of a motion cloud is determined by a constant $k_v$: with a constant speed $\speed$ (see \fig{motionclouds}-(a)): 
+%\begin{equation*}
+%%\begin{gathered}
+%\va{v}_k = %\binom{\cos(\frac{2\pi\cdot k_v}{N_v})}{\sin(\frac{2\pi\cdot k_v}{N_v})}
+%  \begin{pmatrix}
+%    \cos(2\pi\cdot \frac{k}{\Nclass})\\
+%    \sin(2\pi\cdot \frac{k}{\Nclass})
+%  \end{pmatrix}
+%%\end{gathered}
+%\end{equation*}
+$
+\speed_\kernelind = 
+  ( 
+    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
+    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
+  )
+$ (\seeFig{motionclouds}-(a) for an illustration).
+For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content.
+This method provides a rich dataset of textured movies for which we know the ground truth for motion.
+%
+\begin{figure}[h!]%[ht!]
+    \centering
+    \includegraphics[width=0.95\linewidth]{figures/motionclouds.pdf}
+    \caption{{\bf Motion detection task.} {\bf (a)} The motion direction represented as the plain red vector, other possible motion directions are represented in light red.
+{\bf (b)} A screenshot of one generated naturalistic textured stimulus at a specific time.
+% $\timev_\arank$.
+    {\bf (c)} The corresponding ON (in red) and OFF (in blue) event stream generated from the stimuli on (b) and constituting the input to the spiking neural network.}
+    \label{fig:motionclouds}
+\end{figure}
+
+To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames.
+If, on a specific pixel at that specific timestamp, the absolute value of this gradient exceeds a threshold, an event is generated.
+The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive.
+This signed threshold value is then subtracted from the residual gradient image.
+When applied to the whole movie, the event stream is then similar to the output of a neuromorphic camera~\cite{Rasetto2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motionclouds}-(c)).
+The goal here is to infer the correct motion solely by observing these events.
+
 \section{Discussion}

### 2022-09-13 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %
-In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\cite{abeles1982role}.
+In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\cite{Abe}.
 While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code.
@@
 \bibliographystyle{apalike}
-\bibliography{Grimaldi-etal-BiolCybernetics}
+\bibliography{FastMotionDetection}
 \end{document}

### 2022-09-13 Grimaldi-etal-BiolCybernetics.tex
task

@@
 %
-In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\cite{Abe}.
+In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\cite{Abeles82}.
 While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code.

### 2022-09-13 Grimaldi-etal-BiolCybernetics.tex
task

--- 2022-09-13 Grimaldi-etal-BiolCybernetics.tex
@@
 %
-In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\cite{Abe}.
+In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\cite{abeles_role_1982}.
 While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code.

### 2022-09-13 Grimaldi-etal-BiolCybernetics.tex
task

--- 2022-09-13 Grimaldi-etal-BiolCybernetics.tex
@@
 \raggedbottom
-%%\unnumbered% uncomment this for unnumbered level heads
 
-%%\unnumbered% uncomment this for unnumbered level heads
-%\newcommand{\citep}[1]{\cite{#1}}
-%\newcommand{\citet}[1]{\cite{#1}}
 \newcommand{\seeFig}[1]{see Fig.~\ref{fig:#1}}%{see Figure~\ref{fig:#1}}
@@
 %	{todonotes}
-%\newcommand{\noteAG}[1]{\todo[color=green!40, author=\textbf{Antoine}, inline]{#1}}
-%\newcommand{\noteLP}[1]{\todo[color=red!20, author=\textbf{Laurent}, inline]{#1}}
-%\newcommand{\modif}[1]{\todo[color=orange!20, author=\textbf{Proposition}, inline]{#1}}
+%\newcommand{\noteAG}[1]{\todo[author=\textbf{Antoine}, inline]{#1}}
+\usepackage{soul}
+\usepackage{color}
+
+\DeclareRobustCommand{\note}[1]{{\sethlcolor{yellow}\hl{#1}}}
+%\newcommand{\noteLP}[1]{\todo[author=\textbf{Laurent}, inline]{#1}}
+%\newcommand{\modif}[1]{\todo[author=\textbf{Proposition}, inline]{#1}}
 
@@
 %
-In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\cite{Abe}.
+In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\citep{abeles_role_1982}.
 While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code.
-Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\cite{riehle1997spike, davis2021spontaneous}, efficient encoding thanks to the use of spike latencies~\cite{perrinet2004coding, gollisch2008rapid} or precise timing in the auditory system~\cite{deweese2002binary, carr1990circuit}.
-All these findings, and more~\cite{bohte2004evidence}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in biological spike trains.
+Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
+All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in biological spike trains.
 
-In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\cite{gutig2006tempotron, guise2014bayesian, zhang2020supervised}.
-Notably, Izhikevich~\citet{izhikevich2006polychronization} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays.
+In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.
+Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays.
 This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
@@
 
-use a set of synthetic visual stimuli, \textit{Motion Clouds}~\cite{leon2012motion} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}).
+use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}).
+In particular, we will set the spatial size to $(\Nx, \Ny)$ and consider a discretization of time with a time step of $1$ such that $\timev \in \mathbb{N}$.
+Movies' duration are here set to $\Ntime=400$.
+This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions:
+% of a motion cloud is determined by a constant $k_v$: with a constant speed $\speed$ (see \fig{motionclouds}-(a)): 
+%\begin{equation*}
+%%\begin{gathered}
+%\va{v}_k = %\binom{\cos(\frac{2\pi\cdot k_v}{N_v})}{\sin(\frac{2\pi\cdot k_v}{N_v})}
+%  \begin{pmatrix}
+%    \cos(2\pi\cdot \frac{k}{\Nclass})\\
+%    \sin(2\pi\cdot \frac{k}{\Nclass})
+%  \end{pmatrix}
+%%\end{gathered}
+%\end{equation*}
+$
+\speed_\kernelind = 
+  ( 
+    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
+    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
+  )
+$ (\seeFig{motionclouds}-(a) for an illustration).
+For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content.
+This method provides a rich dataset of textured movies for which we know the ground truth for motion.
+%
+\begin{figure}[h!]%[ht!]
+    \centering
+    \includegraphics[width=0.95\linewidth]{figures/motionclouds.pdf}
+    \note{TODO: redo figure according to caption}
+    \caption{
+    {\bf Motion detection task.} {\bf (a)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (here the cross at the center of the image).
+{\bf (b)} To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like Brownian motion, and we show here on example path of the trajectory.
+{\bf (c)} Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation.
+Mimicking the retina, this representation codes for proportional increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events.
+This will constitute the input to the spiking neural network.}
+    \label{fig:motionclouds}
+\end{figure}
+
+To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames.
+If, on a specific pixel at that specific timestamp, the absolute value of this gradient exceeds a threshold, an event is generated.
+The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive.
+This signed threshold value is then subtracted from the residual gradient image.
+When applied to the whole movie, the event stream is then similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motionclouds}-(c)).
+The goal here is to infer the correct motion solely by observing these events.
+
+\subsection{Hetero-synaptic delays model}
+%
+\begin{figure*}[ht!]
+    {\centering
+    \vspace{-3cm}
+    \includegraphics[width=\linewidth]{figures/3D_kernels_oneline.png}
+    }
+    %\vspace{-.5cm}
+    \caption{
+    	Representation of the weights for the $8$ learned kernels of the model corresponding to the OFF polarities and selective to the different motion directions
+	(because of the symmetry observed between the ON and OFF event streams, kernels are similar for the ON polarities).
+These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
+	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.
+One sees positive (excitatory) coefficients for the specific direction of motion and negative (inhibitory) coefficients for all other directions.
+	}
+    \label{fig:kernels}
+\end{figure*} 
+%
+The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
+This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
+This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
+
+We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then not only described by synaptic weights connecting to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+For each neuron $\postsynaddr$, we define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
+Note that a neuron may contact an afferent neuron with multiple different delays.
+
+The corresponding input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible patterns.
+For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
+$
+toto
+%\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
+$.
+%
+The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
+It transforms this list of weights into a probability with the following formula:
+$
+toto
+%Pr(k=\postsynaddr \; | \; \timev) =
+\frac 1 Z
+{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
+$ 
+where $\mathcal{C}^\postsynaddr(\timev) = \sum
+\activeweights^\postsynaddr(t)
+$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
+In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
+%
+\subsection{
+Temporal
+Logistic Regression}
+%
+In our MLR model with $\Nclass$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function described in the above hetero-synpatic delays model.
+Such a probability can be computed for each neuron $\postsynaddr$ in association to a specific class.
+From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.
+As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
+In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the nonzero weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
+Using this dense representation, the counting defined above becomes:
+$$
+\mathcal{C}^\postsynaddr(a,t)
+= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
+$$
+%
+This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
+This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different patterns in the event stream.
+%
+In particular, we may extend the convolution to a 3D convolution 
+such that the representation would also benefit from spatial invariance.
+The use of spatio-temporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\citep{ghosh_spatiotemporal_2019}.
+For that, we design 3D kernels of shape  $(\Kx, \Ky, \Ktime) = (15, 15, 8)$, respectively representing the two spatial dimensions and the range of delays.
+An example of such kernels is given in Figure~\ref{fig:kernels}.
+Computations are performed on spatio-temporal windows, defined by the kernels, sliding around the events, that is, the center of the spatio-temporal window around the current event $\epsilon_\arank$.
+%
+Finally, the output of the MLR model results in an event with the highest probability class, keeping the same timing as the event as input.
+%
+The loss function of the MLR model is the binary cross entropy on the output of the classification layer.
+Simulations are performed thanks to the PyTorch library using gradient descent with Adam (with $2^{12}$ epochs and a learning rate of $10^{-5}$).
+%
+\section{Results}
+\label{sec:results}
+%
+\subsection{Kernels learned for motion detection}
+%
+After training our model, we first observe the weights learned for the different neurons (\seeFig{kernels}).
+Focusing on the positive weights, a strong selectivity is observed along specific axes for the different kernels.
+These directions can be easily associated to the direction of motion controlled in the motion clouds.
+For instance, the third and the seventh kernels show a horizontal selectivity to motion directions.
+%
+With the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
+This qualitative look at the 3D kernels allows the reader to infer for the $8$ different motion directions used to generate our synthetic event streams.
+
+If one focuses on the interpretation of these kernels in terms of spatio-temporal patterns embedded in the event stream, it can lead to interesting outcomes.
+In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions.
+Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spike pattern.
+Each voxel in the 3D kernels defines a specific timestamp and a specific address.
+Then, our model is able to detect precise spatio-temporal patterns embedded in the spike train and associated to the different motion directions.
+The cone shape for the positive weights distribution highlights a loss of precision for longer delays, i.e.
+events away in the past.
+For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Figure~\ref{fig:kernels}.
+%
+\subsection{Accuracy for the motion detection task}
+%
+Once our MLR is trained, we obtain 3D kernels corresponding to the weights associated to the hetero-synaptic delays of our layer of spiking neurons and which may be used for detection.
+We observed that the distribution of the kernels' weights is sparse, with most values near zero.
+As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the number of spikes times the number of synapses.
+This scales with the number of nonzero synaptic weights.
+To  assess the robustness of the classification as a function of the computational load, we will prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold.
+
+In Figure~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations, or active weights, per decision for each neuron of the layer.
+As a comparison and to account for the gain in performance by using hetero-synaptic delays, we provide the accuracy obtained with a MLR model using 2D time surface (in red) as in~\citep{grimaldi_robust_2022}.
+This latter method is based on delays from the last recorded events and uses fewer computations (in our case $15\times15$) than the dense 3D kernels without any pruning ($15\times15\times8$).
+While less computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
+By pruning weights, we observe that the evolution of accuracy as a function of the log percentage of active weights follows a sigmoid curve.
+Half-saturation level is reached at about $3.5\times 10^{-3}\%$ of active weights, corresponding to a total amount of $6$ computations per decision.
+Compared to the full kernels, the accuracy of our method is maintained to its top performances when dividing the number of computations by a factor up to about $200$.
+In this case, the number of computations is greatly reduced compared to~\citep{grimaldi_robust_2022}, thus demonstrating the efficiency of the presented method.
+%
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.95\linewidth]{figures/accuracy.pdf}
+    \caption{Accuracy as a function of the number of computation load for the hetero-synaptic delays model (in blue) and for a method using 2D time surfaces (in red).
+The relative computational load (on a log axis) is controlled by changing the percentage of active weights relative to the dense convolution kernel.
+We observe a similar accuracy than HOTS, yet that our model could achieve a similar accuracy with significantly fewer coefficients.}
+    \label{fig:accuracy}
+\end{figure}
+%
+
+\subsection{Testing with natural-like textures}
+%
+To test our model, we will quantify its ability to categorize different motions.
+In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}).
 In particular, we will set the spatial size to $(\Nx, \Ny)$ and consider a discretization of time with a time step of $1$ such that $\timev \in \mathbb{N}$.
@@
 This signed threshold value is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is then similar to the output of a neuromorphic camera~\cite{Rasetto2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motionclouds}-(c)).
-The goal here is to infer the correct motion solely by observing these events.
-
-\subsection{Hetero-synaptic delays model}
-%
-\begin{figure*}[ht!]
-    {\centering
-    \vspace{-3cm}
-    \includegraphics[width=\linewidth]{figures/3D_kernels_oneline.png}
-    }
-    %\vspace{-.5cm}
-    \caption{
-    	Representation of the weights for the $8$ learned kernels of the model corresponding to the OFF polarities and selective to the different motion directions
-	(because of the symmetry observed between the ON and OFF event streams, kernels are similar for the ON polarities).
-These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
-	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.
-One sees positive (excitatory) coefficients for the specific direction of motion and negative (inhibitory) coefficients for all other directions.
-	}
-    \label{fig:kernels}
-\end{figure*} 
-%
-The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
-This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
-This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
-
-We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then not only described by synaptic weights connecting to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-For each neuron $\postsynaddr$, we define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
-Note that a neuron may contact an afferent neuron with multiple different delays.
-
-The corresponding input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible patterns.
-For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
-$
-toto
-%\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
-$.
-%
-The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\cite{Grimaldi22pami}, in analogy to a spiking Winner-Take-All network~\cite{nessler2013bayesian}.
-It transforms this list of weights into a probability with the following formula:
-$
-toto
-%Pr(k=\postsynaddr \; | \; \timev) =
-\frac 1 Z
-{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
-$ 
-where $\mathcal{C}^\postsynaddr(\timev) = \sum
-\activeweights^\postsynaddr(t)
-$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
-In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
-%
-\subsection{
-Temporal
-Logistic Regression}
-%
-In our MLR model with $\Nclass$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function described in the above hetero-synpatic delays model.
-Such a probability can be computed for each neuron $\postsynaddr$ in association to a specific class.
-From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.
-As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
-In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the nonzero weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
-Using this dense representation, the counting defined above becomes:
-$$
-\mathcal{C}^\postsynaddr(a,t)
-= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
-$$
-%
-This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
-This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different patterns in the event stream.
-%
-In particular, we may extend the convolution to a 3D convolution 
-such that the representation would also benefit from spatial invariance.
-The use of spatio-temporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\cite{ghosh2022conv3D}.
-For that, we design 3D kernels of shape  $(\Kx, \Ky, \Ktime) = (15, 15, 8)$, respectively representing the two spatial dimensions and the range of delays.
-An example of such kernels is given in Figure~\ref{fig:kernels}.
-Computations are performed on spatio-temporal windows, defined by the kernels, sliding around the events, that is, the center of the spatio-temporal window around the current event $\epsilon_\arank$.
-%
-Finally, the output of the MLR model results in an event with the highest probability class, keeping the same timing as the event as input.
-%
-The loss function of the MLR model is the binary cross entropy on the output of the classification layer.
-Simulations are performed thanks to the PyTorch library using gradient descent with Adam (with $2^{12}$ epochs and a learning rate of $10^{-5}$).
-%
-\section{Results}
-\label{sec:results}
-%
-\subsection{Kernels learned for motion detection}
-%
-After training our model, we first observe the weights learned for the different neurons (\seeFig{kernels}).
-Focusing on the positive weights, a strong selectivity is observed along specific axes for the different kernels.
-These directions can be easily associated to the direction of motion controlled in the motion clouds.
-For instance, the third and the seventh kernels show a horizontal selectivity to motion directions.
-%
-With the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
-This qualitative look at the 3D kernels allows the reader to infer for the $8$ different motion directions used to generate our synthetic event streams.
-
-If one focuses on the interpretation of these kernels in terms of spatio-temporal patterns embedded in the event stream, it can lead to interesting outcomes.
-In~\cite{Grimaldi22pami}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions.
-Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spike pattern.
-Each voxel in the 3D kernels defines a specific timestamp and a specific address.
-Then, our model is able to detect precise spatio-temporal patterns embedded in the spike train and associated to the different motion directions.
-The cone shape for the positive weights distribution highlights a loss of precision for longer delays, i.e.
-events away in the past.
-For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Figure~\ref{fig:kernels}.
-%
-\subsection{Accuracy for the motion detection task}
-%
-Once our MLR is trained, we obtain 3D kernels corresponding to the weights associated to the hetero-synaptic delays of our layer of spiking neurons and which may be used for detection.
-We observed that the distribution of the kernels' weights is sparse, with most values near zero.
-As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the number of spikes times the number of synapses.
-This scales with the number of nonzero synaptic weights.
-To  assess the robustness of the classification as a function of the computational load, we will prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold.
-
-In Figure~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations, or active weights, per decision for each neuron of the layer.
-As a comparison and to account for the gain in performance by using hetero-synaptic delays, we provide the accuracy obtained with a MLR model using 2D time surface (in red) as in~\cite{Grimaldi22pami}.
-This latter method is based on delays from the last recorded events and uses fewer computations (in our case $15\times15$) than the dense 3D kernels without any pruning ($15\times15\times8$).
-While less computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
-By pruning weights, we observe that the evolution of accuracy as a function of the log percentage of active weights follows a sigmoid curve.
-Half-saturation level is reached at about $3.5\times 10^{-3}\%$ of active weights, corresponding to a total amount of $6$ computations per decision.
-Compared to the full kernels, the accuracy of our method is maintained to its top performances when dividing the number of computations by a factor up to about $200$.
-In this case, the number of computations is greatly reduced compared to~\cite{Grimaldi22pami}, thus demonstrating the efficiency of the presented method.
-%
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.95\linewidth]{figures/accuracy.pdf}
-    \caption{Accuracy as a function of the number of computation load for the hetero-synaptic delays model (in blue) and for a method using 2D time surfaces (in red).
-The relative computational load (on a log axis) is controlled by changing the percentage of active weights relative to the dense convolution kernel.
-We observe a similar accuracy than HOTS, yet that our model could achieve a similar accuracy with significantly fewer coefficients.}
-    \label{fig:accuracy}
-\end{figure}
-%
-
-\subsection{Testing with natural-like textures}
-%
-To test our model, we will quantify its ability to categorize different motions.
-In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\cite{leon2012motion} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}).
-In particular, we will set the spatial size to $(\Nx, \Ny)$ and consider a discretization of time with a time step of $1$ such that $\timev \in \mathbb{N}$.
-Movies' duration are here set to $\Ntime=400$.
-This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions:
-% of a motion cloud is determined by a constant $k_v$: with a constant speed $\speed$ (see \fig{motionclouds}-(a)): 
-%\begin{equation*}
-%%\begin{gathered}
-%\va{v}_k = %\binom{\cos(\frac{2\pi\cdot k_v}{N_v})}{\sin(\frac{2\pi\cdot k_v}{N_v})}
-%  \begin{pmatrix}
-%    \cos(2\pi\cdot \frac{k}{\Nclass})\\
-%    \sin(2\pi\cdot \frac{k}{\Nclass})
-%  \end{pmatrix}
-%%\end{gathered}
-%\end{equation*}
-$
-\speed_\kernelind = 
-  ( 
-    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
-    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
-  )
-$ (\seeFig{motionclouds}-(a) for an illustration).
-For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content.
-This method provides a rich dataset of textured movies for which we know the ground truth for motion.
-%
-\begin{figure}[h!]%[ht!]
-    \centering
-    \includegraphics[width=0.95\linewidth]{figures/motionclouds.pdf}
-    \caption{{\bf Motion detection task.} {\bf (a)} The motion direction represented as the plain red vector, other possible motion directions are represented in light red.
-{\bf (b)} A screenshot of one generated naturalistic textured stimulus at a specific time.
-% $\timev_\arank$.
-    {\bf (c)} The corresponding ON (in red) and OFF (in blue) event stream generated from the stimuli on (b) and constituting the input to the spiking neural network.}
-    \label{fig:motionclouds}
-\end{figure}
-
-To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames.
-If, on a specific pixel at that specific timestamp, the absolute value of this gradient exceeds a threshold, an event is generated.
-The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive.
-This signed threshold value is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is then similar to the output of a neuromorphic camera~\cite{Rasetto2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motionclouds}-(c)).
+When applied to the whole movie, the event stream is then similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motionclouds}-(c)).
 The goal here is to infer the correct motion solely by observing these events.
@@
 It would follow from the emergence of new kernels adapted to this new task after supervised learning.
-This constitutes a major advantage over other algorithms which derive event-based algorithms from specific physical rules (see for instance~\citep{Benosman12} for computing the optic flow using the luminance conservation rule).
+This constitutes a major advantage over other algorithms which derive event-based algorithms from specific physical rules (see for instance~\citep{benosman_asynchronous_2012} for computing the optic flow using the luminance conservation rule).
 We aim at extending the application of this model on more generic datasets acquired in natural conditions for progressively more complex tasks from motion extraction, optic flow or time-to-contact maps.

### 2022-09-13 Grimaldi-etal-BiolCybernetics.tex

@@
 %
-In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{Grimaldi22pami} which extended the HOTS model~\citet{Lagorce17}.
+In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}.
 This HOTS model represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events.

### 2022-09-13 Grimaldi-etal-BiolCybernetics.tex
task

@@
 %
+\subsection{Dynamics of sensory input and spiking neurons}
 In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\citep{abeles_role_1982}.
@@
 
+\subsection{Hetero-synaptic delays}
 In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.
@@
 
+\subsection{Outline}
+
+In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}.
+This HOTS model represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events.
+By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is in an online fashion.
+We have demonstrated on several datasets that it provides with efficient computations resulting in ultra-fast categorization.
+Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.
+
+Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
+Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, hetero-synaptic delays.
+In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay in the computational process.
+As a consequence, this extends our previous model by encoding spiking motifs as 3D images.
+The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier.
+In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these hetero-synaptic delays.
+%
+
 In this work, we study the emergence of such spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motionclouds}).
@@
 \label{sec:methods}
-%
-In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}.
-This HOTS model represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events.
-By transforming each event in the stream as a vectorial input, our MLR classifier is able to make a decision for every single event, that is in an online fashion.
-We have demonstrated on several datasets that it provides with efficient computations resulting in ultra-fast categorization.
-Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.
-
-Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
-Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, hetero-synaptic delays.
-In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay as a computational process.
-As a consequence, this extends our previous model by encoding spiking patterns as 3D images.
-The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier.
-In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these hetero-synaptic delays.
 %
@@
 
-\subsection{Hetero-synaptic delays model}
+\subsection{Hetero-synaptic delays model as a temporal Logistic Regression}
 %
@@
 %
-\subsection{
-Temporal
-Logistic Regression}
+\subsection{A Spiking Neural Network detecting motion as a CNN}
 %

### 2022-09-13 Grimaldi-etal-BiolCybernetics.tex
task

@@
 %
-\subsection{Dynamics of sensory input and spiking neurons}
+\subsection{Processing of dynamical sensory inputs}
 In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\citep{abeles_role_1982}.
@@
 
-\subsection{Hetero-synaptic delays}
+\subsection{Hetero-synaptic delays in spiking neurons}
 In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.
@@
 %
-We expect our model to be able to efficiently detect temporal patterns and here we will challenge it by defining a realistic task of motion detection in a stream generated by a natural scene with known ground truth motion.
+We expect our model to be able to efficiently detect temporal patterns.
+Here, we will challenge it by defining a realistic task of motion detection in a stream generated by a natural scene with known ground truth motion.
+This ground truth will be used to learn the weights and delays of the network in a supervised fashion.
 %
@@
     \caption{
-    {\bf Motion detection task.} {\bf (a)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (here the cross at the center of the image).
-{\bf (b)} To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like Brownian motion, and we show here on example path of the trajectory.
+    {\bf Motion detection task.} {\bf (a)} To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like Brownian motion, and we show here on example path of the trajectory.
+{\bf (b)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (here the cross at the center of the image) such as to reproduce the effect of displacing the eye.
 {\bf (c)} Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation.

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
from 2022-05-08_Grimaldi-etal-NeuroVision_626a9d254400eb0219f5879a/2022-05-08_Grimaldi-etal-NeuroVision.tex

@@
 This temporal aspect of the neuronal code is essential in understanding information processing in neurobiology and applies particularly well to the output of neuromorphic hardware such as event-based cameras.
-However, most artificial neuronal models do not take advantage of this minute temporal dimension and here, we develop a model for the efficient detection of temporal spiking motifs based on a layer of neurons with hetero-synaptic delays.
+However, most artificial neuronal models do not take advantage of this minute temporal dimension.
+Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of neurons with hetero-synaptic delays.
 Indeed, the variety of synaptic delays on the dendritic tree allows to synchronize synaptic inputs as they reach the basal dendritic tree.
@@
 %
-In our MLR model with $\Nclass$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function described in the above hetero-synpatic delays model.
-Such a probability can be computed for each neuron $\postsynaddr$ in association to a specific class.
+In our MLR model with $\Nclass=\Nspeed$ classes, 
+a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model. 
 From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %
+Our SNN model will be tested on a behaviorally relevant task consisting in detecting the motion within an event - based sensory input.
+Taking analogy with biological constraints We Will design a paradigm simulating eye movements where an input image is displaced by a parameterized jittered trajectory.
+The task is then to guess at best the motion (see fig : motion - task).
+This is, for instance useful in real-life situations to compensate for this motion and provide with motion-invariant image categorization by stabilizing the image on the retina.
+
 \subsection{Task definition: fast motion detection}

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 
+Let's first define a biologically inspired trajectory
+inspired by Gye movements. Indeed,
+these allow to dynamically actuate
+the axis of vision in the visual field of view.
+Ton instance, Saccades au rapid movements of the eyeball
+which reposition the center of vision. In humans,
+these are Very frequent (on average every 3 seconds
+[ ]) They reproduced very rapidly.
+C- 80ms) and the peak speed is proportional to the
+amplitude of the displacement (up to 6000/1 for
+a large saccade of 30" of amplitude). At a more
+microscopic scale, the gaze moves with au
+incessant drift similar to a Brownian-like
+trajectory (Poletti pois).
+To keep with the generality of the task,
+we defined eye movements using a form of
+Random walk (Engbent 20"]. This
+consists first dro define a finite set of
+possible motion. Bayed on the distribution of
+biological motions, we simplified it
+by selectingasetmotionsetofaye motions as the
+cartesian product of N-V- phi-8 linearly spaced motion directions
+and NY-speed - 7 different speeds.
+(see inset of fg : motion_task, a)
+Note that speeds are sampled on a geometric
+scale between Yq and v-speed_base : 4.
+second, we define the trajectory as segments
+Whose duration are drawn from a Poisson
+distribution with an average length of
+block-length : 20 msflevyfl.int). Finally, the trajectory
+is integrated by assuming ① that speeds are
+uniformly and independently sampled from the
+et of different motions motionset and ② motion
+is uniform during a time segment. resulting
+instances are giving qualitatively similar
+trajectories as those observed for eye movements
+as observed for instance in humans (see
+figure fig : motion_task - a) .
+
 use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}).

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
task

--- 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
@@
 %
-We expect our model to be able to efficiently detect temporal patterns.
+% We expect our model to be able to efficiently detect temporal patterns.
 Here, we will challenge it by defining a realistic task of motion detection in a stream generated by a natural scene with known ground truth motion.
 This ground truth will be used to learn the weights and delays of the network in a supervised fashion.
+%
+Our SNN model will be tested on an appropriate computational task of detecting motion in an event-based sensory input.
+By analogy with biological requirements, we will design a paradigm simulating eye movements where an input image is moved by a parameterized saccadic trajectory.
+The task is then to guess the movement as accurately as possible (see fig: movement - task).
+This is, for example, useful in real life situations to compensate for this motion and provide a motion invariant image categorization by stabilizing the image on the retina.
 %

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
@@
 %
-Taking analogy with biological conditions, a common use case for motion detection is to estimate the motion generated on a sensor by saccade-like movements.
+%Taking analogy with biological conditions, a common use case for motion detection is to estimate the motion generated on a sensor by saccade-like movements.
 Saccades are rapid movements of the eyeball which allow to reposition the axis of view in a short period.
 In humans, these are particularly rapid
-To test our model, we will quantify its ability to categorize different motions. In that order, we will generate a natural
+%To test our model, we will quantify its ability to categorize different motions. In that order, we will generate a natural
 
-use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}).
-In particular, we will set the spatial size to $(\Nx, \Ny)$ and consider a discretization of time with a time step of $1$ such that $\timev \in \mathbb{N}$.
-Movies' duration are here set to $\Ntime=400$.
-This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions:
-% of a motion cloud is determined by a constant $k_v$: with a constant speed $\speed$ (see \fig{motionclouds}-(a)): 
-%\begin{equation*}
-%%\begin{gathered}
-%\va{v}_k = %\binom{\cos(\frac{2\pi\cdot k_v}{N_v})}{\sin(\frac{2\pi\cdot k_v}{N_v})}
-%  \begin{pmatrix}
-%    \cos(2\pi\cdot \frac{k}{\Nclass})\\
-%    \sin(2\pi\cdot \frac{k}{\Nclass})
-%  \end{pmatrix}
-%%\end{gathered}
-%\end{equation*}
-$
-\speed_\kernelind = 
-  ( 
-    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
-    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
-  )
-$ (\seeFig{motionclouds}-(a) for an illustration).
-For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content.
-This method provides a rich dataset of textured movies for which we know the ground truth for motion.
-%
+Let's first define a biologically inspired trajectory inspired by eye movements.
+Indeed, these allow to dynamically actuate the axis of vision in the visual field of view.
+For instance, Saccades are rapid movements of the eyeball which reposition the center of vision.
+In humans, these are Very frequent (on average every 3 seconds [ ]).
+They are produced very rapidly C- 80ms) and the peak speed is proportional to the amplitude of the displacement (up to 600 deg/sec for a large saccade of 30deg of amplitude).
+At a more microscopic scale, the gaze moves with are incessant drift similar to a Brownian-like trajectory (Poletti 2015).
+To keep with the generality of the task, we defined eye movements using a form of Random walk (Engbert 211).
+This consists first to define a finite set of
+possible 2D motions in polar coordinates.
+Based on the distribution of biological motions, we simplified it by selecting a set motion set of eye motions as the cartesian product of N-V-phi=8 linearly spaced motion directions and N-Y-speed=7 different speeds (see inset of fig : motion - task, a).
+Note that speeds are sampled on a geometric scale between $1/4$ and $v-speed_base = 4$.
+Second, we define the trajectory as segments whose duration are drawn from a Poisson distribution with an average length of block-length = 20 ms - similar to a Levy flight.
+Finally, the trajectory is integrated by assuming 1/that speeds are uniformly and independently sampled from the set of different motions motion-set and 2/ that motion is uniform during a time segment.
+resulting instances are giving qualitatively similar trajectories as those observed for eye movements as observed for instance in humans (see figure fig : motion-task - a) .
+
 \begin{figure}[h!]%[ht!]
@@
 \end{figure}
+
+Once these eye-movement trajectories are simulated, we can use them on a visual scene.
+For this, we selected a database of 10 grayscale natural images which are typically used for studying tu statistics of natural images [olshausen 96].
+Note that these are reprocessed in order to equalize the energy in each frequency band (that is, whitened).
+This process is known to occur as early as the retina-thalamic pathway [dan 1996].
+These images are of size 512 × 512 and we will extract sub-images of
+1281 128 which will be positionned around
+the center of gaze at every time step (see
+figure fig : motion -task - b).
+
+We will discretize time in 1ms true bins and produce movies of duration N-T = 250 ms (that is, approximately corresponding to the mean duration between two saccades).
+To avoid border effects, we will position one draw of the full trajectory at random in image space such that the sub-image is translated using the position given by the trajectory at each time step.
+The translation is computed using a roll of coordinates in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space [Perrinet, 2015].
+Note that the amplitude of the displacement is relative to the time bin and we defined speed such that a speed of V-1 corresponds to a motion of one pixel per frame (that is, per
+time bin).
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex

@@
 %
-We expect our model to be able to efficiently detect temporal patterns.
+% We expect our model to be able to efficiently detect temporal patterns.
 Here, we will challenge it by defining a realistic task of motion detection in a stream generated by a natural scene with known ground truth motion.
@@
 %
-Our SNN model will be tested on a behaviorally relevant task consisting in detecting the motion within an event - based sensory input.
-Taking analogy with biological constraints We Will design a paradigm simulating eye movements where an input image is displaced by a parameterized jittered trajectory.
-The task is then to guess at best the motion (see fig : motion - task).
-This is, for instance useful in real-life situations to compensate for this motion and provide with motion-invariant image categorization by stabilizing the image on the retina.
-
+Our SNN model will be tested on an appropriate computational task of detecting motion in an event-based sensory input.
+By analogy with biological requirements, we will design a paradigm simulating eye movements where an input image is moved by a parameterized saccadic trajectory.
+The task is then to guess the movement as accurately as possible (see fig: movement - task).
+This is, for example, useful in real life situations to compensate for this motion and provide a motion invariant image categorization by stabilizing the image on the retina.
+%
 \subsection{Task definition: fast motion detection}

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 
-Let's first define a biologically inspired trajectory inspired by eye movements.
-Indeed, these allow to dynamically actuate the axis of vision in the visual field of view.
-For instance, Saccades are rapid movements of the eyeball which reposition the center of vision.
-In humans, these are Very frequent (on average every 3 seconds [ ]).
-They are produced very rapidly C- 80ms) and the peak speed is proportional to the amplitude of the displacement (up to 600 deg/sec for a large saccade of 30deg of amplitude).
-At a more microscopic scale, the gaze moves with are incessant drift similar to a Brownian-like trajectory (Poletti 2015).
-To keep with the generality of the task, we defined eye movements using a form of Random walk (Engbert 211).
-This consists first to define a finite set of
-possible 2D motions in polar coordinates.
-Based on the distribution of biological motions, we simplified it by selecting a set motion set of eye motions as the cartesian product of N-V-phi=8 linearly spaced motion directions and N-Y-speed=7 different speeds (see inset of fig : motion - task, a).
-Note that speeds are sampled on a geometric scale between $1/4$ and $v-speed_base = 4$.
-Second, we define the trajectory as segments whose duration are drawn from a Poisson distribution with an average length of block-length = 20 ms - similar to a Levy flight.
-Finally, the trajectory is integrated by assuming 1/that speeds are uniformly and independently sampled from the set of different motions motion-set and 2/ that motion is uniform during a time segment.
-resulting instances are giving qualitatively similar trajectories as those observed for eye movements as observed for instance in humans (see figure fig : motion-task - a) .
+Définissons au préalable une trajectoire biologiquement inspirée par les mouvements oculaires.
+En effet, ceux-ci permettent d'actionner dynamiquement le centre de vision dans le champ de vision.
+Par exemple, les saccades sont des mouvements rapides du globe oculaire qui repositionnent le centre de la vision vers des points d'interet.
+Chez l'homme, celles-ci sont très fréquentes (en moyenne toutes les 3 secondes [ ]).
+Elles sont produites très rapidement (environ 80ms) et la vitesse maximale est proportionnelle à l'amplitude du déplacement (jusqu'à 600 deg/sec pour une grande saccade de 30deg d'amplitude).
+À une échelle plus microscopique, le regard se déplace avec une dérive incessante similaire à une trajectoire de type brownien (Poletti 2015).
+Pour conserver toute la généralité de la tâche, nous avons défini les mouvements oculaires en utilisant une forme de marche aléatoire (Engbert 211).
+Cette démarche consiste d'abord à définir un ensemble défini de mouvements 2D possibles en coordonnées polaires.
+En nous basant sur la distribution des mouvements biologiques, nous l'avons simplifié en sélectionnant un ensemble de mouvements oculaires comme le produit cartésien de N-V-phi=8 directions de mouvement linéairement espacées et de N-Y-vitesse=7 vitesses différentes (voir encart de la fig : mouvement - tâche, a).
+Notez que les vitesses sont échantillonnées sur une échelle géométrique comprise entre $1/4$ et $v-speed_base = 4$.
+Ensuite, nous définissons la trajectoire comme des segments dont la durée est tirée d'une distribution de Poisson avec une longueur moyenne de bloc = 20 ms - similaire à un vol de Lévy.
+Enfin, la trajectoire est intégrée en supposant 1/que les vitesses sont échantillonnées de manière uniforme et indépendante à partir de l'ensemble de mouvements différents motion-set et 2/que le mouvement est uniforme pendant un segment de temps.
+Les instances résultantes donnent des trajectoires qualitativement similaires à celles observées pour les mouvements oculaires tels qu'observés par exemple chez les humains (voir figure fig : motion-task - a).
 
@@
 This process is known to occur as early as the retina-thalamic pathway [dan 1996].
-These images are of size 512 × 512 and we will extract sub-images of
-1281 128 which will be positionned around
-the center of gaze at every time step (see
-figure fig : motion -task - b).
-
+These images are of size 512 × 512 and we will extract sub-images of 1281 128 which will be positionned around the center of gaze at every time step (see figure fig : motion -task - b).
 We will discretize time in 1ms true bins and produce movies of duration N-T = 250 ms (that is, approximately corresponding to the mean duration between two saccades).

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
@@
 %To test our model, we will quantify its ability to categorize different motions. In that order, we will generate a natural
-
-Let's first define a biologically inspired trajectory inspired by eye movements.
-Indeed, these allow to dynamically actuate the axis of vision in the visual field of view.
-For instance, Saccades are rapid movements of the eyeball which reposition the center of vision.
-In humans, these are Very frequent (on average every 3 seconds [ ]).
-They are produced very rapidly C- 80ms) and the peak speed is proportional to the amplitude of the displacement (up to 600 deg/sec for a large saccade of 30deg of amplitude).
-At a more microscopic scale, the gaze moves with are incessant drift similar to a Brownian-like trajectory (Poletti 2015).
-To keep with the generality of the task, we defined eye movements using a form of Random walk (Engbert 211).
-This consists first to define a finite set of
-possible 2D motions in polar coordinates.
-Based on the distribution of biological motions, we simplified it by selecting a set motion set of eye motions as the cartesian product of N-V-phi=8 linearly spaced motion directions and N-Y-speed=7 different speeds (see inset of fig : motion - task, a).
-Note that speeds are sampled on a geometric scale between $1/4$ and $v-speed_base = 4$.
-Second, we define the trajectory as segments whose duration are drawn from a Poisson distribution with an average length of block-length = 20 ms - similar to a Levy flight.
-Finally, the trajectory is integrated by assuming 1/that speeds are uniformly and independently sampled from the set of different motions motion-set and 2/ that motion is uniform during a time segment.
-resulting instances are giving qualitatively similar trajectories as those observed for eye movements as observed for instance in humans (see figure fig : motion-task - a) .
+Let us first define a trajectory inspired by the biological movements of the eyes.
+Indeed, these movements allow us to dynamically actuate the center of vision in the field of vision.
+For example, saccades are rapid movements of the eyeball that reposition the center of vision towards points of interest.
+In humans, these are very frequent (on average every 3 seconds [ ]).
+They are produced very rapidly (about 80ms) and the maximum speed is proportional to the amplitude of the displacement (up to 600 deg/sec for a large saccade of 30deg amplitude).
+At a more microscopic scale, the gaze moves with an incessant drift similar to a Brownian-like trajectory (Poletti 2015).
+To maintain the full generality of the task, we defined eye movements using a form of random walk (Engbert 211).
+This approach first defines a defined set of possible 2D motions in polar coordinates.
+Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of N-V-phi=8 linearly spaced motion directions and N-Y-velocity=7 different velocities (see inset of Fig: motion-task, a).
+Note that the velocities are sampled on a geometric scale between $1/4$ and $v-speed_base = 4$.
+Next, we define the trajectory as segments whose duration is drawn from a Poisson distribution with an average block length = 20 ms - similar to a Lévy flight.
+Finally, the trajectory is integrated by assuming 1/ that velocities are uniformly and independently sampled from the set of different motion-sets and 2/ that motion is uniform during a time segment.
+The resulting instances yield trajectories qualitatively similar to those observed for eye movements as observed for example in humans (see figure fig: motion-task - a).
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 
-Once these eye-movement trajectories are simulated, we can use them on a visual scene.
-For this, we selected a database of 10 grayscale natural images which are typically used for studying tu statistics of natural images [olshausen 96].
-Note that these are reprocessed in order to equalize the energy in each frequency band (that is, whitened).
-This process is known to occur as early as the retina-thalamic pathway [dan 1996].
-These images are of size 512 × 512 and we will extract sub-images of 1281 128 which will be positionned around the center of gaze at every time step (see figure fig : motion -task - b).
-We will discretize time in 1ms true bins and produce movies of duration N-T = 250 ms (that is, approximately corresponding to the mean duration between two saccades).
-To avoid border effects, we will position one draw of the full trajectory at random in image space such that the sub-image is translated using the position given by the trajectory at each time step.
-The translation is computed using a roll of coordinates in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space [Perrinet, 2015].
-Note that the amplitude of the displacement is relative to the time bin and we defined speed such that a speed of V-1 corresponds to a motion of one pixel per frame (that is, per
-time bin).
+Une fois que ces trajectoires de mouvements oculaires sont générées, nous pouvons les appliquer à une scène visuelle.
+Pour cela, nous avons sélectionné une base de données de 10 images naturelles en niveaux de gris qui sont typiquement utilisées pour étudier les statistiques des images naturelles [olshausen 96].
+Notons que celles-ci sont prétraitées afin d'égaliser l'énergie dans chaque bande de fréquence (c'est-à-dire blanchies).
+Ce processus est connu pour se manifester dès la voie rétino-thalamique [dan 1996].
+Ces images sont de taille 512 × 512 et nous allons extraire des sous-images de taille 128 x 128 qui seront positionnées autour du centre du regard à chaque pas de temps (voir figure fig : motion -task - b).
+Nous discrétiserons le temps en bins de 1ms réels et produirons des films de durée N-T = 250 ms (c'est-à-dire correspondant approximativement à la durée moyenne entre deux saccades).
+Pour éviter les effets de bord, nous positionnerons un tirage de la trajectoire complète au hasard dans l'espace image de telle sorte que la sous-image soit déplacée en utilisant la position donnée par la trajectoire à chaque pas de temps.
+La translation est calculée en utilisant un roulement de coordonnées dans les dimensions horizontale et verticale, suivi d'une translation sous-pixel définie dans l'espace de Fourier [Perrinet, 2015].
+Notez que l'amplitude du déplacement est relative au bin de temps et nous avons défini la vitesse de telle sorte qu'une vitesse de V=1 correspond à un mouvement d'un pixel par image (c'est-à-dire par bin de temps).
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
@@
 
-In this work, we study the emergence of such spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motionclouds}).
+In this work, we study the emergence of such spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motion-task}).
 We develop a SNN-like method able to learn hetero-synaptic delays to perform motion detection on a synthetic event-based dataset.
@@
 This approach first defines a defined set of possible 2D motions in polar coordinates.
-Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of N-V-phi=8 linearly spaced motion directions and N-Y-velocity=7 different velocities (see inset of Fig: motion-task, a).
+Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of N-V-phi=8 linearly spaced motion directions and N-Y-velocity=7 different velocities (see inset of \seeFig{motion-task}-(a)).
 Note that the velocities are sampled on a geometric scale between $1/4$ and $v-speed_base = 4$.
@@
 Finally, the trajectory is integrated by assuming 1/ that velocities are uniformly and independently sampled from the set of different motion-sets and 2/ that motion is uniform during a time segment.
-The resulting instances yield trajectories qualitatively similar to those observed for eye movements as observed for example in humans (see figure fig: motion-task - a).
+The resulting instances yield trajectories qualitatively similar to those observed for eye movements as observed for example in humans (\seeFig{motion-task}-(a)).
 
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{figures/motionclouds.pdf}
+    \includegraphics[width=0.95\linewidth]{figures/motion-task.pdf}
     \note{TODO: redo figure according to caption}
@@
 This will constitute the input to the spiking neural network.}
-    \label{fig:motionclouds}
+    \label{fig:motion-task}
 \end{figure}
 
-Once these eye-movement trajectories are simulated, we can use them on a visual scene.
-For this, we selected a database of 10 grayscale natural images which are typically used for studying tu statistics of natural images [olshausen 96].
-Note that these are reprocessed in order to equalize the energy in each frequency band (that is, whitened).
-This process is known to occur as early as the retina-thalamic pathway [dan 1996].
-These images are of size 512 × 512 and we will extract sub-images of 1281 128 which will be positionned around the center of gaze at every time step (see figure fig : motion -task - b).
-We will discretize time in 1ms true bins and produce movies of duration N-T = 250 ms (that is, approximately corresponding to the mean duration between two saccades).
-To avoid border effects, we will position one draw of the full trajectory at random in image space such that the sub-image is translated using the position given by the trajectory at each time step.
-The translation is computed using a roll of coordinates in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space [Perrinet, 2015].
-Note that the amplitude of the displacement is relative to the time bin and we defined speed such that a speed of V-1 corresponds to a motion of one pixel per frame (that is, per
-time bin).
+Once these eye movement trajectories are generated, we can apply them to a visual scene.
+For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images [olshausen 96].
+Note that these are pre-processed to equalize the energy in each frequency band (i.e., whitened).
+This process is known to occur as early as the retino-thalamic pathway [dan 1996].
+These images are 512 × 512 in size and we will extract sub-images of size 128 x 128 which will be positioned around the center of the gaze at each time step (see \seeFig{motion-task}-(b)).
+We will discretize the time in 1ms bins and produce movies of duration N-T = 250 ms (i.e.
+approximately corresponding to the average duration between two saccades).
+To avoid border effects, we will position a draw of the complete trajectory at random in the image space so that the sub-image is translated using the position given by the trajectory at each time step.
+The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space [Perrinet, 2015].
+Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that a velocity of V=1 corresponds to a movement of one pixel per frame (i.e., per time bin).
 
@@
 This signed threshold value is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is then similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motionclouds}-(c)).
+When applied to the whole movie, the event stream is then similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion-task}-(c)).
 The goal here is to infer the correct motion solely by observing these events.

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
task

@@
 %
-% We expect our model to be able to efficiently detect temporal patterns.
-Here, we will challenge it by defining a realistic task of motion detection in a stream generated by a natural scene with known ground truth motion.
-This ground truth will be used to learn the weights and delays of the network in a supervised fashion.
-%
+Let's define formally the model, the task and how we will solve it.
+First, we will define the hetero-synaptic model which allows to learn in a  supervised manner the weights and delays of a feed-forward layer of spiking neural network (SNN).
 Our SNN model will be tested on an appropriate computational task of detecting motion in an event-based sensory input.
 By analogy with biological requirements, we will design a paradigm simulating eye movements where an input image is moved by a parameterized saccadic trajectory.
-The task is then to guess the movement as accurately as possible (see fig: movement - task).
-This is, for example, useful in real life situations to compensate for this motion and provide a motion invariant image categorization by stabilizing the image on the retina.
+The task is then to guess the movement as accurately as possible (see \seeFig{motion-task}).
+This is, for example, useful in real life situations to compensate for eye, head or body motions and provide a robust image categorization by stabilizing the image on the retina.
+For this, we will extend the SNN model to be efficiently adapted to solve the task by including convolutional kernels and describe its implementation on conventional computers.
 %

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
task

--- 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
@@
 %
-% We expect our model to be able to efficiently detect temporal patterns.
-Here, we will challenge it by defining a realistic task of motion detection in a stream generated by a natural scene with known ground truth motion.
-This ground truth will be used to learn the weights and delays of the network in a supervised fashion.
-%
-Our SNN model will be tested on an appropriate computational task of detecting motion in an event-based sensory input.
-By analogy with biological requirements, we will design a paradigm simulating eye movements where an input image is moved by a parameterized saccadic trajectory.
-The task is then to guess the movement as accurately as possible (see fig: movement - task).
-This is, for example, useful in real life situations to compensate for this motion and provide a motion invariant image categorization by stabilizing the image on the retina.
+Let us now formally define the model, as well as the task and how we will solve it.
+First, we will define the hetero-synaptic model for supervised learning of weights and delays in a feed-forward layer of a spiking neural network (SNN).
+Our SNN model will be tested on a computationally relevant task of motion detection in an event-based sensory input.
+By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory.
+The task is then to guess the motion as accurately as possible (see \seeFig{motion-task}).
+This is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina.
+To this end, we will extend the SNN model to be efficient for the resolution of the task by including convolutional kernels and will describe its implementation on conventional computers.
 %

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
task

@@
 %
-Let us now formally define the model, as well as the task and how we will solve it.
-First, we will define the hetero-synaptic model for supervised learning of weights and delays in a feed-forward layer of a spiking neural network (SNN).
+Let us now formally define the hetero-synaptic model, as well as the task and how we will solve it.
+First, we will define the model for supervised learning of weights and delays in a feed-forward layer of a spiking neural network (SNN).
 Our SNN model will be tested on a computationally relevant task of motion detection in an event-based sensory input.
@@
 This is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina.
-To this end, we will extend the SNN model to be efficient for the resolution of the task by including convolutional kernels and will describe its implementation on conventional computers.
+To this end, we will extend the SNN model to be efficient for the resolution of the task by including convolutional kernels.
+We will finally describe its implementation on conventional computers.
 %
@@
 %
-%Taking analogy with biological conditions, a common use case for motion detection is to estimate the motion generated on a sensor by saccade-like movements.
-Saccades are rapid movements of the eyeball which allow to reposition the axis of view in a short period.
-In humans, these are particularly rapid
-%To test our model, we will quantify its ability to categorize different motions. In that order, we will generate a natural
-Let us first define a trajectory inspired by the biological movements of the eyes.
-Indeed, these movements allow us to dynamically actuate the center of vision in the field of vision.
-For example, saccades are rapid movements of the eyeball that reposition the center of vision towards points of interest.
-In humans, these are very frequent (on average every 3 seconds [ ]).
+Let us now define a trajectory inspired by the biological movements of the eyes.
+Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the field of vision.
+In animals with a fovea, this is particularly useful as it allows to move the area with the highest density of photoreceptors in the environnement, for example at a point of interest.
+For example, saccades are rapid movements of the eyeball that reposition the center of vision.
+In humans, these are very frequent (on average one every 3 seconds~\citep{}).
 They are produced very rapidly (about 80ms) and the maximum speed is proportional to the amplitude of the displacement (up to 600 deg/sec for a large saccade of 30deg amplitude).
-At a more microscopic scale, the gaze moves with an incessant drift similar to a Brownian-like trajectory (Poletti 2015).
-To maintain the full generality of the task, we defined eye movements using a form of random walk (Engbert 211).
-This approach first defines a defined set of possible 2D motions in polar coordinates.
+At a more microscopic scale, the human gaze moves with an incessant drift similar to a Brownian-like trajectory~\citep{Poletti 2015}.
+To maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{Engbert 2011}.
+This approach first defines a finite set of possible 2D motions in polar coordinates.
 Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of N-V-phi=8 linearly spaced motion directions and N-Y-velocity=7 different velocities (see inset of \seeFig{motion-task}-(a)).
 Note that the velocities are sampled on a geometric scale between $1/4$ and $v-speed_base = 4$.
-Next, we define the trajectory as segments whose duration is drawn from a Poisson distribution with an average block length = 20 ms - similar to a Lévy flight.
+Next, we define one trajectory of gaze as segments whose duration is drawn from a Poisson distribution with an average block length = 20 ms - similar to a Lévy flight~\citep{}.
 Finally, the trajectory is integrated by assuming 1/ that velocities are uniformly and independently sampled from the set of different motion-sets and 2/ that motion is uniform during a time segment.
-The resulting instances yield trajectories qualitatively similar to those observed for eye movements as observed for example in humans (\seeFig{motion-task}-(a)).
+The resulting instances yield trajectories qualitatively similar to those observed for human eye movements (\seeFig{motion-task}-(a)).
 
@@
     \caption{
-    {\bf Motion detection task.} {\bf (a)} To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like Brownian motion, and we show here on example path of the trajectory.
+    {\bf Motion detection task.} {\bf (a)} To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like random walk, and we show here on example path of the trajectory.
 {\bf (b)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (here the cross at the center of the image) such as to reproduce the effect of displacing the eye.
@@
 Once these eye movement trajectories are generated, we can apply them to a visual scene.
-For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images [olshausen 96].
+For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images~\citep{olshausen 96}.
 Note that these are pre-processed to equalize the energy in each frequency band (i.e., whitened).
 This process is known to occur as early as the retino-thalamic pathway [dan 1996].
-These images are 512 × 512 in size and we will extract sub-images of size 128 x 128 which will be positioned around the center of the gaze at each time step (see \seeFig{motion-task}-(b)).
-We will discretize the time in 1ms bins and produce movies of duration N-T = 250 ms (i.e.
-approximately corresponding to the average duration between two saccades).
+These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step (see \seeFig{motion-task}-(b)).
+We will discretize the time in $1~\ms$ bins and produce movies of duration N-T = $250~\ms$.
 To avoid border effects, we will position a draw of the complete trajectory at random in the image space so that the sub-image is translated using the position given by the trajectory at each time step.
-The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space [Perrinet, 2015].
-Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that a velocity of V=1 corresponds to a movement of one pixel per frame (i.e., per time bin).
+The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{Perrinet, 2015}.
+Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that a velocity of $V=1$ corresponds to a movement of one pixel per frame (i.e., per time bin).
 
+% TODO: formula / say it is driven by image gradients / give scaling rules
 To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames.

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
model

@@
 \subsection{Hetero-synaptic delays model as a temporal Logistic Regression}
-%
-\begin{figure*}[ht!]
-    {\centering
-    \vspace{-3cm}
-    \includegraphics[width=\linewidth]{figures/3D_kernels_oneline.png}
-    }
-    %\vspace{-.5cm}
-    \caption{
-    	Representation of the weights for the $8$ learned kernels of the model corresponding to the OFF polarities and selective to the different motion directions
-	(because of the symmetry observed between the ON and OFF event streams, kernels are similar for the ON polarities).
-These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
-	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.
-One sees positive (excitatory) coefficients for the specific direction of motion and negative (inhibitory) coefficients for all other directions.
-	}
-    \label{fig:kernels}
-\end{figure*} 
-%
 The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
@@
 %
-\subsection{A Spiking Neural Network detecting motion as a CNN}
+
 %
@@
 This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different patterns in the event stream.
+\subsection{Extension as 3D convolutions for motion detection}
 %
@@
 \subsection{Kernels learned for motion detection}
+%
+\begin{figure*}[ht!]
+    {\centering
+    \vspace{-3cm}
+    \includegraphics[width=\linewidth]{figures/3D_kernels_oneline.png}
+    }
+    %\vspace{-.5cm}
+    \caption{
+    	Representation of the weights for the $8$ learned kernels of the model corresponding to the OFF polarities and selective to the different motion directions
+	(because of the symmetry observed between the ON and OFF event streams, kernels are similar for the ON polarities).
+These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
+	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.
+One sees positive (excitatory) coefficients for the specific direction of motion and negative (inhibitory) coefficients for all other directions.
+	}
+    \label{fig:kernels}
+\end{figure*} 
+%
 %

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
model

@@
 We will finally describe its implementation on conventional computers.
+
+\subsection{Hetero-synaptic delays model as a temporal Logistic Regression}
+The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
+This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
+This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
+
+We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then not only described by synaptic weights connecting to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+For each neuron $\postsynaddr$, we define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
+Note that a neuron may contact an afferent neuron with multiple different delays.
+
+The corresponding input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible patterns.
+For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
+$
+toto
+%\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
+$.
+%
+The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
+It transforms this list of weights into a probability with the following formula:
+$
+toto
+%Pr(k=\postsynaddr \; | \; \timev) =
+\frac 1 Z
+{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
+$ 
+where $\mathcal{C}^\postsynaddr(\timev) = \sum
+\activeweights^\postsynaddr(t)
+$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
+In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
+%
+
+%
+In our MLR model with $\Nclass=\Nspeed$ classes, 
+a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model. 
+From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.
+As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
+In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the nonzero weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
+Using this dense representation, the counting defined above becomes:
+$$
+\mathcal{C}^\postsynaddr(a,t)
+= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
+$$
+%
+This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
+This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different patterns in the event stream.
 %
@@
 
-\subsection{Hetero-synaptic delays model as a temporal Logistic Regression}
-The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
-This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
-This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
-
-We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then not only described by synaptic weights connecting to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-For each neuron $\postsynaddr$, we define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
-Note that a neuron may contact an afferent neuron with multiple different delays.
-
-The corresponding input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible patterns.
-For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
-$
-toto
-%\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
-$.
-%
-The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
-It transforms this list of weights into a probability with the following formula:
-$
-toto
-%Pr(k=\postsynaddr \; | \; \timev) =
-\frac 1 Z
-{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
-$ 
-where $\mathcal{C}^\postsynaddr(\timev) = \sum
-\activeweights^\postsynaddr(t)
-$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
-In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
-%
-
-%
-In our MLR model with $\Nclass=\Nspeed$ classes, 
-a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
-The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model. 
-From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.
-As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
-In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the nonzero weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
-Using this dense representation, the counting defined above becomes:
-$$
-\mathcal{C}^\postsynaddr(a,t)
-= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
-$$
-%
-This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
-This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different patterns in the event stream.
 \subsection{Extension as 3D convolutions for motion detection}

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
task

@@
 Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the field of vision.
-In animals with a fovea, this is particularly useful as it allows to move the area with the highest density of photoreceptors in the environnement, for example at a point of interest.
+In animals with a fovea, this is particularly useful as it allows to move the area with the highest density of photoreceptors in the environment, for example at a point of interest.
 For example, saccades are rapid movements of the eyeball that reposition the center of vision.
@@
 Note that these are pre-processed to equalize the energy in each frequency band (i.e., whitened).
-This process is known to occur as early as the retino-thalamic pathway [dan 1996].
+This process is known to occur as early as the retino-thalamic pathway~\citep{dan 1996}.
 These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step (see \seeFig{motion-task}-(b)).
@@
 
-% TODO: formula / say it is driven by image gradients / give scaling rules
+\note{ TODO: formula / say it is driven by image gradients / give scaling rules }
 To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames.
@@
 The goal here is to infer the correct motion solely by observing these events.
+
+\note{ the task is full-field rigid translations, the model resolves it locally (analogy to the input from V1 to MT) }
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
task

@@
 \subsection{Hetero-synaptic delays model as a temporal Logistic Regression}
-The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
-This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
-This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
 
-We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+Let's first define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
 In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
@@
 $
-toto
-%\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
+\activeweights^\postsynaddr (t) = % \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
 $.
@@
 The goal here is to infer the correct motion solely by observing these events.
-
+The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
+This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
+This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
 \note{ the task is full-field rigid translations, the model resolves it locally (analogy to the input from V1 to MT) }
@@
 %
-In particular, we may extend the convolution to a 3D convolution 
-such that the representation would also benefit from spatial invariance.
+In particular, we may extend the convolution to a 3D convolution such that the representation would also benefit from spatial invariance which is inherent to the task's construction.
 The use of spatio-temporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\citep{ghosh_spatiotemporal_2019}.

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
model

@@
 
-\subsection{Hetero-synaptic delays model as a temporal Logistic Regression}
+\subsection{Hetero-synaptic delays model as a Logistic Regression with a temporal convolution}
 
@@
 In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then not only described by synaptic weights connecting to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-For each neuron $\postsynaddr$, we define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
 Note that a neuron may contact an afferent neuron with multiple different delays.
+For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
+This defines the full connectivity of the hetero-synaptic model.
 
-The corresponding input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible patterns.
+We can now define an event-based input to the model as a set on incoming spikes reaching the presynaptic neurons at specified times, like the sensory signal representing the output of an event-based camera.
+It thus forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
+Each event has a time of occurrence $\timev_\arank$.
+The corresponding set of input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible patterns.
 For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
 $
-\activeweights^\postsynaddr (t) = % \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
+\activeweights^\postsynaddr (t) = 
+\{ \synapticweight_\ranksyn^\postsynaddr  \vert 
+\presynaddr_\arank = 
+ \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + 
+\synapticdelay_\ranksyn^\postsynaddr \}_{ \arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]} 
+\}
 $.
-%
-The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
-It transforms this list of weights into a probability with the following formula:
-$
-toto
-%Pr(k=\postsynaddr \; | \; \timev) =
-\frac 1 Z
-{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
-$ 
-where $\mathcal{C}^\postsynaddr(\timev) = \sum
-\activeweights^\postsynaddr(t)
-$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
-In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
-%
+\note{finish by saying that it is effortless in biology but difficult on conventional computers}
 
-%
-In our MLR model with $\Nclass=\Nspeed$ classes, 
-a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
-The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model. 
+\note{ TODO: show why this formula implements a polychrony detector }
 From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.
@@
 %
+
+The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
+It transforms this list of weights into a probability with the following formula:
+$
+Pr(k=\postsynaddr \; \vert \; \timev) =
+\frac 1 Z
+{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
+$ 
+where $\mathcal{C}^\postsynaddr(\timev) = \sum
+\activeweights^\postsynaddr(t)
+$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
+In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
+%
+In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model.
+
 \subsection{Task definition: fast motion detection}
@@
 The goal here is to infer the correct motion solely by observing these events.
-The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
-This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
+This sensory signal representing the output of this event-based camera forms a discrete stream of events, $\event$ as defined above.   
+Each event has an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
 This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{figures/motionclouds.pdf}
+    \includegraphics[width=0.95\linewidth]{figures/motion-clouds.pdf}
     \caption{{\bf Motion detection task.} {\bf (a)} The motion direction represented as the plain red vector, other possible motion directions are represented in light red.

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 
+\subsection{Synthesis and main contributions}
+
+To synthesize, we have introduced in this paper a spiking neural network model with hetero-synaptic delays which we have trained and evaluated on a challenging motion detection task.
+The model was defined as optimally detecting spatio-temporal event-based motifs.
+and we have shown that by training on a dataset of natural images with realistic eye movements, it learns kernels similar to those found in the early visual cortex, of humans for instance.
+We have then shown that the response of the model has similar responses as found in biology.
+We have further evaluated the computational cost for this model if it would be implemented on neuromorphic hardware, showing that including hetero-synaptic delays may prove
+to be a frugal computational solution for future neuroomorphic hardware but also a key to understand why spikes is one universal currency in neural information processing.
+
+Let us highlight some novelties of the contributions reported in this paper. First, the generic hetero-synaptic model is formalized from first
+principles of optimally detecting spatio-temporal event-based signal, while []  use a heuristic based on correlation, which we have proved to be less efficient.
+\note{compared to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly give the logit (inverse sigmoid of the probability) for contributing to detect each spatio-temporal spiking motif.} The model learns simultaneously the weights and delays, while the polychronization model (Izhikevich 2006) only learns weights using STDP while delays are drawn randomly and theit values frozen.
+Moreover, the model is evaluated on a realistic task, while models
+such as the tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}.
+Foremost, the main contribution is to bring forward a model which is adapted to spatio-temporal spiking motifs and able to be trained in a supervised manner.
+This allows a more flexible definition of the model using a properly labelled dataset instead of relying on a minute description of the physical rules governing a task, for instance the
+luminance conservation principle for motion detection (Bensman 2012, Dardeletz).
+
 %One advantage of our model is the generality of the possible kernels.

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 \note{compared to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly give the logit (inverse sigmoid of the probability) for contributing to detect each spatio-temporal spiking motif.} The model learns simultaneously the weights and delays, while the polychronization model (Izhikevich 2006) only learns weights using STDP while delays are drawn randomly and theit values frozen.
-Moreover, the model is evaluated on a realistic task, while models
-such as the tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}.
-Foremost, the main contribution is to bring forward a model which is adapted to spatio-temporal spiking motifs and able to be trained in a supervised manner.
-This allows a more flexible definition of the model using a properly labelled dataset instead of relying on a minute description of the physical rules governing a task, for instance the
-luminance conservation principle for motion detection (Bensman 2012, Dardeletz).
+Moreover, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}.
+Foremost, the main contribution is to bring forward a model which is adapted to any spatio-temporal spiking motifs and able to be trained in a supervised manner by giving a dataset of supervison pairs.
+This allows a more flexible definition of the model using this properly labelled dataset instead of relying on a minute description of the physical rules governing a task, for instance the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet}  (Benosman 2012, Dardelet 2021).
+
+\subsection{Main limits}
 
@@
 
-Note that this supervised learning scheme can be extended to a variety of tasks.
+\subsection{Perspectives}
+
+%Note that this supervised learning scheme can be extended to a variety of tasks.
 It would follow from the emergence of new kernels adapted to this new task after supervised learning.

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Discussion: Synthesis and main contributions
Update on Overleaf.

@@
 
+We ientified several limits to our model,
+the major of which we will now detail. First,
+the full framework is based on discrete time which
+is incompatible with the analogous nature of biological
+time. We have used the binning of time to ease
+the implementation of the computational framework
+on conventional hardware, especially GPUs,
+and in particular by using the three-dimensional
+convolution. We have tested the effect of this time
+bin size and have shown it does not impact
+the results shown here. This is consistent with the
+relative robustness of other ff event-based
+frameworks such as thots, where accuracy was
+not changed when jittering input spikes by
+up to 1ms noise (grimaldi zz]. This hints
+at including analytically a precision term in
+the temporal value of incoming spikes, a
+mechanism potentially implemented by the
+n 5 ms smoothing at the level of the PSC
+of conventional synapses. Furthermore, it is
+possible to overcome the problem using a purely
+even d- based scheme. Indeed, it is not necessary to
+compute voltage males between two spikes (Morrison)
+and to define a purely event-driven framework.
+A further limit is that the model that
+we defined is purely feed-forward. It
+results that the spikes generated by post-synaptic
+neurons are produced solely based on the
+information in the classical receptive field.
+However, it is known that neurons in a
+layer may interact using lateral interaction
+[charane Lorry and that this may be the
+basis of computational principles, for instance
+of binding neighboring orientations (Bednar).
+Moreover, neural information is modulated by
+feedback information, for instance do distinguish
+figure from background (Roefsema J and
+has proven to be essential in building realistic
+models of the primary visual areas (Boutin 2020
+Boutin rozo feedback , Boutin '2oz]. It is nd
+currently possible to implant these recurrent
+connections (lateral or feedback), mainly because
+of our use of convolutions. However the generic
+model is capable of including this by
+inserting spikes in the list of spikes reaching
+the pipe-synaptic addresses. While possible
+in theory, it should be properly adapted
+in practice such that these recurrent
+connections do not reach outside an
+homeostatic state ( a death or a
+blow-up of activity).
+Suchreccarrentatvityweential
+to implement predictive or anticipative
+processes . This is essential in a neural
+system as it contains multiple different
+delays which necessitates a temporal
+alignment (Hoogendoorn 2019]. This was
+previously modeled to explain for instance
+the flash - lag visual illusion (Khoei 2017)
+As was previously started " ... " (reminet
+20174], and our hetero-synaptic model
+would provide with a simple and elegant
+solution to that problem.
+
 \subsection{Perspectives}

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Discussion: Synthesis and main contributions

@@
 We have then shown that the response of the model has similar responses as found in biology.
-We have further evaluated the computational cost for this model if it would be implemented on neuromorphic hardware, showing that including hetero-synaptic delays may prove
-to be a frugal computational solution for future neuroomorphic hardware but also a key to understand why spikes is one universal currency in neural information processing.
+We have further evaluated the computational cost for this model if it would be implemented on neuromorphic hardware, showing that including hetero-synaptic delays may prove to be a frugal computational solution for future neuromorphic hardware but also a key to understand why spikes is one universal currency in neural information processing.
 
-Let us highlight some novelties of the contributions reported in this paper. First, the generic hetero-synaptic model is formalized from first
-principles of optimally detecting spatio-temporal event-based signal, while []  use a heuristic based on correlation, which we have proved to be less efficient.
-\note{compared to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly give the logit (inverse sigmoid of the probability) for contributing to detect each spatio-temporal spiking motif.} The model learns simultaneously the weights and delays, while the polychronization model (Izhikevich 2006) only learns weights using STDP while delays are drawn randomly and theit values frozen.
+Let us highlight some novelties of the contributions reported in this paper.
+First, the generic hetero-synaptic model is formalized from first principles of optimally detecting spatio-temporal event-based signal, while []  use a heuristic based on correlation, which we have proved to be less efficient.
+In addition, compared to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly give the logit (inverse sigmoid of the probability) for contributing to detect each spatio-temporal spiking motif.
+Another novelty is that the model learns simultaneously the weights and delays, while for instance the polychronization model~\citep{Izhikevich 2006} only learns weights using STDP while delays are drawn randomly and their values frozen.
 Moreover, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}.
-Foremost, the main contribution is to bring forward a model which is adapted to any spatio-temporal spiking motifs and able to be trained in a supervised manner by giving a dataset of supervison pairs.
-This allows a more flexible definition of the model using this properly labelled dataset instead of relying on a minute description of the physical rules governing a task, for instance the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet}  (Benosman 2012, Dardelet 2021).
+Foremost, the main contribution is to bring forward a model which is adapted to any spatio-temporal spiking motifs and able to be trained in a supervised manner by giving a dataset of supervision pairs.
+This allows a more flexible definition of the model using this properly labelled dataset instead of relying on a minute description of the physical rules governing a task, for instance the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet}.
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Discussion: Synthesis and main contributions

--- 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
@@
 \subsection{Synthesis and main contributions}
+Pour synthétiser, nous avons introduit dans cet article un modèle de réseau de neurones à spikes avec des retards hétérosynaptiques que nous avons entraîné et évalué sur une tâche complexe de détection de mouvement.
+Le modèle a été défini comme devant détecter de manière optimale des motifs spatio-temporels événementiels.
+Nous avons montré qu'en l'entraînant sur un dataset d'images naturelles avec des mouvements oculaires réalistes, le modèle apprend des kernels similaires à ceux trouvés dans le cortex visuel précoce, des humains par exemple.
+Nous avons ensuite montré que la réponse du modèle a des réponses similaires à celles trouvées en biologie.
+Nous avons ensuite évalué le coût de calcul de ce modèle s'il était implémenté sur du matériel neuromorphique, montrant que le recours à des délais hétérosynaptiques peut s'avérer une solution de calcul frugale pour le futur hardware neuromorphique, mais aussi une clé pour comprendre pourquoi les spikes sont une composante universelle du traitement de l'information neuronale.
 
-To synthesize, we have introduced in this paper a spiking neural network model with hetero-synaptic delays which we have trained and evaluated on a challenging motion detection task.
-The model was defined as optimally detecting spatio-temporal event-based motifs.
-and we have shown that by training on a dataset of natural images with realistic eye movements, it learns kernels similar to those found in the early visual cortex, of humans for instance.
-We have then shown that the response of the model has similar responses as found in biology.
-We have further evaluated the computational cost for this model if it would be implemented on neuromorphic hardware, showing that including hetero-synaptic delays may prove
-to be a frugal computational solution for future neuroomorphic hardware but also a key to understand why spikes is one universal currency in neural information processing.
-
-Let us highlight some novelties of the contributions reported in this paper. First, the generic hetero-synaptic model is formalized from first
-principles of optimally detecting spatio-temporal event-based signal, while []  use a heuristic based on correlation, which we have proved to be less efficient.
-\note{compared to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly give the logit (inverse sigmoid of the probability) for contributing to detect each spatio-temporal spiking motif.} The model learns simultaneously the weights and delays, while the polychronization model (Izhikevich 2006) only learns weights using STDP while delays are drawn randomly and theit values frozen.
-Moreover, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}.
-Foremost, the main contribution is to bring forward a model which is adapted to any spatio-temporal spiking motifs and able to be trained in a supervised manner by giving a dataset of supervison pairs.
-This allows a more flexible definition of the model using this properly labelled dataset instead of relying on a minute description of the physical rules governing a task, for instance the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet}  (Benosman 2012, Dardelet 2021).
+Soulignons quelques innovations dans les contributions présentées dans cet article.
+Tout d'abord, le modèle hétérosynaptique générique est formalisé à partir des premiers principes de détection optimale du signal spatio-temporel basé sur les événements, alors que [] utilise une heuristique basée sur la corrélation, dont nous avons prouvé qu'elle était moins efficace.
+De plus, par rapport à HOTS~\citep{lagorce_hots_2017} les poids sont explicables car ils donnent directement le logit (sigmoïde inverse de la probabilité) pour contribuer à détecter chaque motif de spiking spatio-temporel.
+Une autre nouveauté est que le modèle apprend simultanément les poids et les délais, alors que par exemple le modèle de polychronisation~\citep{Izhikevich 2006} n'apprend que les poids en utilisant la STDP alors que les délais sont tirés aléatoirement et leurs valeurs gelées.
+De plus, le modèle est évalué sur une tâche réaliste, alors que des modèles tels que le tempotron sont testés sur des problèmes simplifiés~\citep{gutig_tempotron_2006}.
+La contribution principale est d'apporter un modèle qui est adapté à n'importe quel type de motifs de spiking spatio-temporels et qui peut être entraîné de manière supervisée en fournissant un dataset de paires de supervision.
+Cela permet une définition plus flexible du modèle en utilisant ce dataset correctement étiqueté au lieu de s'appuyer sur une description minutieuse des règles physiques régissant une tâche, par exemple le principe de conservation de la luminance pour la détection de mouvement~\citep{benosman_asynchronous_2012, dardelet}.
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Discussion: Synthesis and main contributions

--- 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
@@
 \subsection{Synthesis and main contributions}
+To synthesize, we have introduced in this paper a spiking neural network model with hetero-synaptic delays that we have trained and evaluated on a complex motion detection task.
+The model was defined to optimally detect event-driven spatiotemporal patterns.
+We showed that when trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example.
+We then showed that the model has similar responses to those found in biology.
+We then evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of hetero-synaptic delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understanding why spikes represent a universal component of neuronal information processing.
 
-To synthesize, we have introduced in this paper a spiking neural network model with hetero-synaptic delays which we have trained and evaluated on a challenging motion detection task.
-The model was defined as optimally detecting spatio-temporal event-based motifs.
-and we have shown that by training on a dataset of natural images with realistic eye movements, it learns kernels similar to those found in the early visual cortex, of humans for instance.
-We have then shown that the response of the model has similar responses as found in biology.
-We have further evaluated the computational cost for this model if it would be implemented on neuromorphic hardware, showing that including hetero-synaptic delays may prove
-to be a frugal computational solution for future neuroomorphic hardware but also a key to understand why spikes is one universal currency in neural information processing.
-
-Let us highlight some novelties of the contributions reported in this paper. First, the generic hetero-synaptic model is formalized from first
-principles of optimally detecting spatio-temporal event-based signal, while []  use a heuristic based on correlation, which we have proved to be less efficient.
-\note{compared to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly give the logit (inverse sigmoid of the probability) for contributing to detect each spatio-temporal spiking motif.} The model learns simultaneously the weights and delays, while the polychronization model (Izhikevich 2006) only learns weights using STDP while delays are drawn randomly and theit values frozen.
-Moreover, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}.
-Foremost, the main contribution is to bring forward a model which is adapted to any spatio-temporal spiking motifs and able to be trained in a supervised manner by giving a dataset of supervison pairs.
-This allows a more flexible definition of the model using this properly labelled dataset instead of relying on a minute description of the physical rules governing a task, for instance the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet}  (Benosman 2012, Dardelet 2021).
+Let us highlight some innovations in the contributions presented in this paper.
+First, the generic heterosynaptic model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs, whereas [] uses a correlation-based heuristic, which we have proven to be less efficient.
+Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking pattern.
+Another novelty is that the model simultaneously learns the weights and the delays, while for example the polychronization model~\citep{Izhikevich 2006} only learns the weights using STDP while the delays are randomly drawn and their values frozen.
+Moreover, the model is evaluated on a realistic task, while models such as tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}.
+The main contribution is to provide a model that is suitable for any type of spatio-temporal spiking patterns and that can be trained in a supervised way by providing a dataset of supervision pairs.
+This allows for a more flexible definition of the model using this properly labeled dataset instead of relying on a careful description of the physical rules governing a task, e.g., the luminance conservation principle for motion detection~~citep{benosman_asynchronous_2012, dardelet}.
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Discussion: Limits

@@
 The main contribution is to provide a model that is suitable for any type of spatio-temporal spiking patterns and that can be trained in a supervised way by providing a dataset of supervision pairs.
-This allows for a more flexible definition of the model using this properly labeled dataset instead of relying on a careful description of the physical rules governing a task, e.g., the luminance conservation principle for motion detection~~citep{benosman_asynchronous_2012, dardelet}.
+This allows for a more flexible definition of the model using this properly labeled dataset instead of relying on a careful description of the physical rules governing a task, e.g., the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet}.
 
@@
 
-We ientified several limits to our model,
-the major of which we will now detail. First,
-the full framework is based on discrete time which
-is incompatible with the analogous nature of biological
-time. We have used the binning of time to ease
-the implementation of the computational framework
-on conventional hardware, especially GPUs,
-and in particular by using the three-dimensional
-convolution. We have tested the effect of this time
-bin size and have shown it does not impact
-the results shown here. This is consistent with the
-relative robustness of other ff event-based
-frameworks such as thots, where accuracy was
-not changed when jittering input spikes by
-up to 1ms noise (grimaldi zz]. This hints
-at including analytically a precision term in
-the temporal value of incoming spikes, a
-mechanism potentially implemented by the
-n 5 ms smoothing at the level of the PSC
-of conventional synapses. Furthermore, it is
-possible to overcome the problem using a purely
-even d- based scheme. Indeed, it is not necessary to
-compute voltage males between two spikes (Morrison)
-and to define a purely event-driven framework.
-A further limit is that the model that
-we defined is purely feed-forward. It
-results that the spikes generated by post-synaptic
-neurons are produced solely based on the
-information in the classical receptive field.
-However, it is known that neurons in a
-layer may interact using lateral interaction
-[charane Lorry and that this may be the
-basis of computational principles, for instance
-of binding neighboring orientations (Bednar).
-Moreover, neural information is modulated by
-feedback information, for instance do distinguish
-figure from background (Roefsema J and
-has proven to be essential in building realistic
-models of the primary visual areas (Boutin 2020
-Boutin rozo feedback , Boutin '2oz]. It is nd
-currently possible to implant these recurrent
-connections (lateral or feedback), mainly because
-of our use of convolutions. However the generic
-model is capable of including this by
-inserting spikes in the list of spikes reaching
-the pipe-synaptic addresses. While possible
-in theory, it should be properly adapted
-in practice such that these recurrent
-connections do not reach outside an
-homeostatic state ( a death or a
-blow-up of activity).
-Suchreccarrentatvityweential
-to implement predictive or anticipative
-processes . This is essential in a neural
-system as it contains multiple different
-delays which necessitates a temporal
-alignment (Hoogendoorn 2019]. This was
-previously modeled to explain for instance
-the flash - lag visual illusion (Khoei 2017)
-As was previously started " ... " (reminet
-20174], and our hetero-synaptic model
-would provide with a simple and elegant
-solution to that problem.
+We identified several limits to our model, the major of which we will now detail.
+First, the full framework is based on discrete binning of time which is incompatible with the analogous nature of biological time.
+We have used this binning to ease the implementation of the computational framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions.
+We have tested the effect of the time bin size and have shown it does not impact the results presented in this paper.
+This is consistent with the relative robustness of other  event-based frameworks such as HOTS (Lagorce 2017), where accuracy was not changed when jittering input spikes by up to 1ms of noise (grimaldi 2022).
+This hints at including analytically a precision term in the temporal value of incoming spikes, a mechanism potentially implemented by the approx 5 ms smoothing which is implemented at the level of the PSC of synapses.
+Furthermore, it is possible to overcome the problem using a purely even-driven based scheme.
+Indeed, it is not necessary to compute voltage males between two spikes (Morrison) and to define a purely event-driven framework.
+Such an architecture should provide with promising speed-ups for the computations.
+
+A further limit is that the model that we defined is purely feed-forward.
+As a consequence, the spikes generated by post-synaptic neurons are produced solely based on the information in the classical receptive field.
+However, it is known that neurons in a layer may interact using lateral interaction (chavane 2022) and that this may be the basis of computational principles, for instance of binding of neighboring orientations which may be a basis for image categorization (Bednar).
+Moreover, neural information is modulated by feedback information, for instance do distinguish figure from background (Roelfsema 2016) and has proven to be essential in building realistic models of the primary visual areas (Boutin 2020 Boutin 2020 feedback , Boutin 2022).
+It is not currently possible to implement these recurrent connections in our implementation (lateral or feedback), mainly because of our use of convolutions.
+However the generic theoretical model is capable of including these by inserting spikes in the list of spikes reaching the pre-synaptic addresses.
+While possible in theory, it should be properly adapted in practice such that these recurrent connections do not amplify neural activity outside an homeostatic state (by a death or a blow-up of activity).
+
+Such recurrent activity would be essential to implement predictive or anticipative processes .
+This is essential in a neural system as it contains multiple different delays which necessitates a temporal alignment (Hoogendoorn 2019].
+This was previously modeled to explain for instance the flash - lag visual illusion (Khoei 2017) As was previously started " ...
+" (Perrinet 2014], and our hetero-synaptic model would provide with a simple and elegant solution to that problem.
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Discussion: Limits

@@
 }
+Nous avons identifié plusieurs limites à notre modèle, dont nous détaillons à présent les principales.
+Tout d'abord, le cadre complet est basé sur un binning discret du temps qui est incompatible avec la nature analogue du temps biologique.
+Nous avons utilisé ce binning pour faciliter l'implémentation du cadre de calcul sur du matériel conventionnel, notamment des GPU, et notamment pour être en mesure d'utiliser des convolutions tridimensionnelles.
+Nous avons testé l'effet de la taille du bin de temps et nous avons montré qu'il n'a pas d'impact sur les résultats présentés dans cet article.
+Cela est cohérent avec la robustesse relative d'autres cadres basés sur des événements tels que HOTS (Lagorce 2017), où la précision n'a pas été modifiée lorsque les spikes d'entrée ont été soumis à un bruit allant jusqu'à 1ms (grimaldi 2022).
+Cela suggère la possibilité d'inclure analytiquement un terme de précision dans la valeur temporelle des spikes en entrée, un mécanisme potentiellement mis en œuvre par le filtrage d'environ 5 ms qui est mis en œuvre au niveau du PSC des synapses.
+De plus, il est possible de surmonter le problème en utilisant un schéma purement basé sur les événements.
+En effet, il n'est pas nécessaire de calculer les traces de voltage entre deux spikes (Morrison) et donc de définir un cadre purement événementiel.
+Une telle architecture pourrait fournir des gains de vitesse prometteurs pour les calculs.
 
-We identified several limits to our model, the major of which we will now detail.
-First, the full framework is based on discrete binning of time which is incompatible with the analogous nature of biological time.
-We have used this binning to ease the implementation of the computational framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions.
-We have tested the effect of the time bin size and have shown it does not impact the results presented in this paper.
-This is consistent with the relative robustness of other  event-based frameworks such as HOTS (Lagorce 2017), where accuracy was not changed when jittering input spikes by up to 1ms of noise (grimaldi 2022).
-This hints at including analytically a precision term in the temporal value of incoming spikes, a mechanism potentially implemented by the approx 5 ms smoothing which is implemented at the level of the PSC of synapses.
-Furthermore, it is possible to overcome the problem using a purely even-driven based scheme.
-Indeed, it is not necessary to compute voltage males between two spikes (Morrison) and to define a purely event-driven framework.
-Such an architecture should provide with promising speed-ups for the computations.
+Une autre limite est que le modèle que nous avons défini est purement feed-forward.
+Par conséquent, les spikes générés par les neurones post-synaptiques sont produits uniquement sur la base des informations contenues dans le champ réceptif classique.
+Cependant, il est reconnu que les neurones d'une couche peuvent interagir en utilisant des interactions latérales (chavane 2022) et que cela peut être la base de principes computationnels, par exemple de la combinaison d'orientations voisines qui peut être une composante du mécanisme de catégorisation des images (Bednar).
+De plus, l'information neuronale est modulée par l'information en feed-back, par exemple pour distinguer une figure de l'arrière-plan (Roelfsema 2016) et s'est avérée essentielle pour construire des modèles réalistes des aires visuelles primaires (Boutin 2020 Boutin 2020 feedback , Boutin 2022).
+Il n'est actuellement pas possible d'implémenter ces connexions récurrentes dans notre implémentation (latérale ou feedback), principalement en raison de notre utilisation des convolutions.
+Cependant le modèle théorique générique est capable de les inclure en insérant des spikes dans la liste des spikes qui atteignent les adresses pré-synaptiques.
+Bien que cela soit possible en théorie, il convient de l'adapter correctement en pratique afin que ces connexions récurrentes n'amplifient pas l'activité neuronale en dehors d'un état homéostatique (par une mort ou une explosion de l'activité).
 
-A further limit is that the model that we defined is purely feed-forward.
-As a consequence, the spikes generated by post-synaptic neurons are produced solely based on the information in the classical receptive field.
-However, it is known that neurons in a layer may interact using lateral interaction (chavane 2022) and that this may be the basis of computational principles, for instance of binding of neighboring orientations which may be a basis for image categorization (Bednar).
-Moreover, neural information is modulated by feedback information, for instance do distinguish figure from background (Roelfsema 2016) and has proven to be essential in building realistic models of the primary visual areas (Boutin 2020 Boutin 2020 feedback , Boutin 2022).
-It is not currently possible to implement these recurrent connections in our implementation (lateral or feedback), mainly because of our use of convolutions.
-However the generic theoretical model is capable of including these by inserting spikes in the list of spikes reaching the pre-synaptic addresses.
-While possible in theory, it should be properly adapted in practice such that these recurrent connections do not amplify neural activity outside an homeostatic state (by a death or a blow-up of activity).
-
-Such recurrent activity would be essential to implement predictive or anticipative processes .
-This is essential in a neural system as it contains multiple different delays which necessitates a temporal alignment (Hoogendoorn 2019].
-This was previously modeled to explain for instance the flash - lag visual illusion (Khoei 2017) As was previously started " ...
-" (Perrinet 2014], and our hetero-synaptic model would provide with a simple and elegant solution to that problem.
+Une telle activité récurrente serait essentielle pour mettre en œuvre des processus prédictifs ou anticipatifs.
+Ceci est essentiel dans un système neuronal car il contient de multiples retards différents qui nécessitent un alignement temporel (Hoogendoorn 2019].
+Ceci a été précédemment modélisé pour expliquer par exemple l'illusion de flash - lag (Khoei 2017) Comme cela a été précédemment commencé " ....
+" (Perrinet 2014], et notre modèle hétérosynaptique fournirait une solution simple et élégante à ce problème.
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Discussion: Limits

--- 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
@@
 }
+We have identified several limits to our model, the most important being now detailed.
+First, the complete framework is based on a discrete binning of time that is incompatible with the continuous nature of biological time.
+We used this binning to be able to efficiently implement the framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions.
+We tested the effect of the size of the time bin and showed that it has essentially no impact on the results presented in this paper.
+This is consistent with the relative robustness of other event-based frameworks such as HOTS (Lagorce 2017), where the accuracy was not affected when the input spikes were subjected to noise up to 1ms (grimaldi 2022).
+This suggests the possibility of analytically including a precision term in the temporal value of the input spikes, a mechanism potentially implemented by the approximately 5 ms filtering that is implemented at the PSC of synapses.
+Furthermore, it is possible to overcome the problem of time discretization by using a purely event-based scheme.
+Indeed, it is not necessary to compute voltage traces between two spikes ( Hanuschkin 2010) and thus define a purely event-based framework.
+Such an architecture could provide promising speed gains for the calculations.
 
-We identified several limits to our model, the major of which we will now detail.
-First, the full framework is based on discrete binning of time which is incompatible with the analogous nature of biological time.
-We have used this binning to ease the implementation of the computational framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions.
-We have tested the effect of the time bin size and have shown it does not impact the results presented in this paper.
-This is consistent with the relative robustness of other  event-based frameworks such as HOTS (Lagorce 2017), where accuracy was not changed when jittering input spikes by up to 1ms of noise (grimaldi 2022).
-This hints at including analytically a precision term in the temporal value of incoming spikes, a mechanism potentially implemented by the approx 5 ms smoothing which is implemented at the level of the PSC of synapses.
-Furthermore, it is possible to overcome the problem using a purely even-driven based scheme.
-Indeed, it is not necessary to compute voltage males between two spikes (Morrison) and to define a purely event-driven framework.
-Such an architecture should provide with promising speed-ups for the computations.
+Another restriction is that this model is purely feed-forward.
+Therefore, the spikes generated by the postsynaptic neurons are produced solely on the basis of information contained in the classical receptive field.
+However, it is well known that neurons in a layer can interact using lateral interactions, for instance in V1 (Chavane 2022) and that this can be the basis for computational principles, for example, the combination of neighbouring orientations which is a candidate for the image categorization mechanism (Bednar).
+Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background (Roelfsema 2016) and has been shown to be essential for building realistic models of primary visual areas (Boutin 2020 Boutin 2020 feedback , Boutin 2022).
+It is currently not possible to implement these recurrent connections in our implementation (lateral or feedback), mainly due to our use of convolutions.
+However, the generic theoretical model is able to include them by inserting new spikes into the list of spikes that are reaching pre-synaptic addresses.
+While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside of a homeostatic state (through a death or explosion of activity).
 
-A further limit is that the model that we defined is purely feed-forward.
-As a consequence, the spikes generated by post-synaptic neurons are produced solely based on the information in the classical receptive field.
-However, it is known that neurons in a layer may interact using lateral interaction (chavane 2022) and that this may be the basis of computational principles, for instance of binding of neighboring orientations which may be a basis for image categorization (Bednar).
-Moreover, neural information is modulated by feedback information, for instance do distinguish figure from background (Roelfsema 2016) and has proven to be essential in building realistic models of the primary visual areas (Boutin 2020 Boutin 2020 feedback , Boutin 2022).
-It is not currently possible to implement these recurrent connections in our implementation (lateral or feedback), mainly because of our use of convolutions.
-However the generic theoretical model is capable of including these by inserting spikes in the list of spikes reaching the pre-synaptic addresses.
-While possible in theory, it should be properly adapted in practice such that these recurrent connections do not amplify neural activity outside an homeostatic state (by a death or a blow-up of activity).
-
-Such recurrent activity would be essential to implement predictive or anticipative processes .
-This is essential in a neural system as it contains multiple different delays which necessitates a temporal alignment (Hoogendoorn 2019].
-This was previously modeled to explain for instance the flash - lag visual illusion (Khoei 2017) As was previously started " ...
-" (Perrinet 2014], and our hetero-synaptic model would provide with a simple and elegant solution to that problem.
+Such recurrent activity would be essential for implementing predictive or anticipatory processes.
+This is essential in a neural system as it contains multiple different delays that require temporal alignment (Hoogendoorn 2019).
+This has been previously modeled to explain for example the flash-lag illusion (Khoei 2017).
+As previously stated "....
+" (Perrinet 2014), and our hetero-synaptic model would provide a simple and elegant solution to this problem.
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Discussion: Limits

@@
 We tested the effect of the size of the time bin and showed that it has essentially no impact on the results presented in this paper.
-This is consistent with the relative robustness of other event-based frameworks such as HOTS (Lagorce 2017), where the accuracy was not affected when the input spikes were subjected to noise up to 1ms (grimaldi 2022).
+This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{} (Lagorce 2017), where the accuracy was not affected when the input spikes were subjected to noise up to 1ms~\citep{} (grimaldi 2022).
 This suggests the possibility of analytically including a precision term in the temporal value of the input spikes, a mechanism potentially implemented by the approximately 5 ms filtering that is implemented at the PSC of synapses.
 Furthermore, it is possible to overcome the problem of time discretization by using a purely event-based scheme.
-Indeed, it is not necessary to compute voltage traces between two spikes ( Hanuschkin 2010) and thus define a purely event-based framework.
+Indeed, it is not necessary to compute voltage traces between two spikes~\citep{} (Hanuschkin 2010) and thus define a purely event-based framework.
 Such an architecture could provide promising speed gains for the calculations.
@@
 However, it is well known that neurons in a layer can interact using lateral interactions, for instance in V1 (Chavane 2022) and that this can be the basis for computational principles, for example, the combination of neighbouring orientations which is a candidate for the image categorization mechanism (Bednar).
-Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background (Roelfsema 2016) and has been shown to be essential for building realistic models of primary visual areas (Boutin 2020 Boutin 2020 feedback , Boutin 2022).
+Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background (Roelfsema 2016) and has been shown to be essential for building realistic models of primary visual areas~\citep{} (Boutin 2020 Boutin 2020 feedback , Boutin 2022).
 It is currently not possible to implement these recurrent connections in our implementation (lateral or feedback), mainly due to our use of convolutions.
@@
 Such recurrent activity would be essential for implementing predictive or anticipatory processes.
-This is essential in a neural system as it contains multiple different delays that require temporal alignment (Hoogendoorn 2019).
+This is essential in a neural system as it contains multiple different delays that require temporal alignment~\citep{} (Hoogendoorn 2019).
 This has been previously modeled to explain for example the flash-lag illusion (Khoei 2017).
 As previously stated "....
-" (Perrinet 2014), and our hetero-synaptic model would provide a simple and elegant solution to this problem.
+"~\citep{} (Perrinet 2014), and our hetero-synaptic model would provide a simple and elegant solution to this problem.
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
@@
 We tested the effect of the size of the time bin and showed that it has essentially no impact on the results presented in this paper.
-This is consistent with the relative robustness of other event-based frameworks such as HOTS (Lagorce 2017), where the accuracy was not affected when the input spikes were subjected to noise up to 1ms (grimaldi 2022).
+This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where the accuracy was not affected when the input spikes were subjected to noise up to 1ms~\citep{grimaldi_robust_2022}.
 This suggests the possibility of analytically including a precision term in the temporal value of the input spikes, a mechanism potentially implemented by the approximately 5 ms filtering that is implemented at the PSC of synapses.
 Furthermore, it is possible to overcome the problem of time discretization by using a purely event-based scheme.
-Indeed, it is not necessary to compute voltage traces between two spikes ( Hanuschkin 2010) and thus define a purely event-based framework.
+Indeed, it is not necessary to compute voltage traces between two spikes~\citep{hanuschkin_general_2010} and thus define a purely event-based framework.
 Such an architecture could provide promising speed gains for the calculations.
@@
 Therefore, the spikes generated by the postsynaptic neurons are produced solely on the basis of information contained in the classical receptive field.
-However, it is well known that neurons in a layer can interact using lateral interactions, for instance in V1 (Chavane 2022) and that this can be the basis for computational principles, for example, the combination of neighbouring orientations which is a candidate for the image categorization mechanism (Bednar).
-Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background (Roelfsema 2016) and has been shown to be essential for building realistic models of primary visual areas (Boutin 2020 Boutin 2020 feedback , Boutin 2022).
+However, it is well known that neurons in a layer can interact using lateral interactions, for instance in V1~\citep{chavane_revisiting_2022} and that this can be the basis for computational principles, for example, the combination of neighbouring orientations which is a candidate for the image categorization mechanism~\citep{perrinet_edge_2015}.
+Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background~\citep{roelfsema_early_2016} and has been shown to be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020, boutin_pooling_2022}.
 It is currently not possible to implement these recurrent connections in our implementation (lateral or feedback), mainly due to our use of convolutions.
@@
 Such recurrent activity would be essential for implementing predictive or anticipatory processes.
-This is essential in a neural system as it contains multiple different delays that require temporal alignment (Hoogendoorn 2019).
-This has been previously modeled to explain for example the flash-lag illusion (Khoei 2017).
+This is essential in a neural system as it contains multiple different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}.
+This has been previously modeled to explain for example the flash-lag illusion~\citep{khoei_flash-lag_2017}.
 As previously stated "....
-" (Perrinet 2014), and our hetero-synaptic model would provide a simple and elegant solution to this problem.
+"~\citep{perrinet_active_2014}, and our hetero-synaptic model would provide a simple and elegant solution to this problem.
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 The main contribution is to provide a model that is suitable for any type of spatio-temporal spiking patterns and that can be trained in a supervised way by providing a dataset of supervision pairs.
-This allows for a more flexible definition of the model using this properly labeled dataset instead of relying on a careful description of the physical rules governing a task, e.g., the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet}.
+This allows for a more flexible definition of the model using this properly labeled dataset instead of relying on a careful description of the physical rules governing a task, e.g., the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}.
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking pattern.
-Another novelty is that the model simultaneously learns the weights and the delays, while for example the polychronization model~\citep{Izhikevich 2006} only learns the weights using STDP while the delays are randomly drawn and their values frozen.
+Another novelty is that the model simultaneously learns the weights and the delays, while for example the polychronization model~\citep{izhikevich_polychronization_2006} only learns the weights using STDP while the delays are randomly drawn and their values frozen.
 Moreover, the model is evaluated on a realistic task, while models such as tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}.

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
Discussion: Limits

@@
 This has been previously modeled to explain for example the flash-lag illusion~\citep{khoei_flash-lag_2017}.
-As previously stated "....
-"~\citep{perrinet_active_2014}, and our hetero-synaptic model would provide a simple and elegant solution to this problem.
+As previously stated, this could be implemented using generalized coordinates (position, speed, acceleration, ...) and ``neurobiologically, the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
+Our proposed hetero-synaptic model provides an alternate elegant solution to this problem.
 

### 2022-09-14 Grimaldi-etal-BiolCybernetics.tex
bib
Update on Overleaf.

@@
 \newcommand{\ShortTitle}{
-Spiking neurons with hetero-synaptic delays for motion detection
+Learning synaptic delays of spiking neurons for motion detection
 }
@@
 For example, saccades are rapid movements of the eyeball that reposition the center of vision.
-In humans, these are very frequent (on average one every 3 seconds~\citep{}).
-They are produced very rapidly (about 80ms) and the maximum speed is proportional to the amplitude of the displacement (up to 600 deg/sec for a large saccade of 30deg amplitude).
+In humans, these are very frequent (on average one every 3 per second~\citep{dandekar_neural_2012}).
+They are produced very rapidly (about $80~\ms$) and display a whole range of speeds.
 At a more microscopic scale, the human gaze moves with an incessant drift similar to a Brownian-like trajectory~\citep{Poletti 2015}.
@@
 We aim at extending the application of this model on more generic datasets acquired in natural conditions for progressively more complex tasks from motion extraction, optic flow or time-to-contact maps.
-%
+% >>>>>>>>>>> LuP is here <<<<<<<<<<<
 

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
discussion

@@
 Let us highlight some innovations in the contributions presented in this paper.
-First, the generic heterosynaptic model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs, whereas [] uses a correlation-based heuristic, which we have proven to be less efficient.
+First, the generic heterosynaptic model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs, whereas~\citet{ghosh_spatiotemporal_2019} use a correlation-based heuristic, which we have proven to be less efficient.
 Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking pattern.
@@
 This has been previously modeled to explain for example the flash-lag illusion~\citep{khoei_flash-lag_2017}.
-As previously stated, this could be implemented using generalized coordinates (position, speed, acceleration, ...) and ``neurobiologically, the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
-Our proposed hetero-synaptic model provides an alternate elegant solution to this problem.
+As previously stated, this could be implemented using generalized coordinates (that is variables such as position which are complemented by speed, acceleration, jerk, ...) and ``neurobiologically, the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
+Our proposed hetero-synaptic model provides an alternate and elegant implementation solution to this problem.
 

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
bib

@@
 \newcommand{\Acknowledgments}{This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}'').
-%HL and LP received funding from the ANR project  N° ANR-20-CE23-0021 (''\href{https://laurentperrinet.github.io/grant/anr-anr/}{AgileNeuroBot}'').
-HL received funding from a University of Montreal Artificial Intelligence scholarship.
-For the purpose of open access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript version arising from this submission.
+%HL and LP received funding from the ANR project  N° ANR-20-CE23-0021 (''\href{https://laurentperrinet.github.io/grant/anr-anr/}
+LP received funding from the ANR project  N° ANR-22-XXX-XXX (''\href{https://laurentperrinet.github.io/grant/polychronies/}{polychronies}''). 
+%HL received funding from a University of Montreal Artificial Intelligence scholarship. 
+%
+%For the purpose of open access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript version arising from this submission.
 This research was funded, in whole or in part, by [Organisation name, Grant #].
@@
 % >>>>>>>>>>> LuP is here <<<<<<<<<<<
+Lors de la définition de notre tâche, nous avons montré comment la génération d'événements dépendait du gradient dans les images.
+Ce gradient est défini dans les dimensions horizontale et verticale et les maxima de ce gradient sont généralement orientés.
+Considérés ensemble, ces bords orientés forment les contours des objets visuels présents dans la scène (Koenderink).
+Il existe donc une dépendance au sein du flux d'événements entre les informations relatives au mouvement et à l'orientation.
+Il serait crucial d'étudier plus en profondeur cette interdépendance, ce qui pourrait être amorcé en entraînant le modèle sur un ensemble de données comportant des labels d'orientation.
+L'étude de cette dépendance permettra de dissocier ces deux formes d'information visuelle et de pouvoir les intégrer, et notamment de considérer que la définition du mouvement est plus précise perpendiculairement à un contour orienté (aka, le problème de l'ouverture), mais aussi d'implémenter des règles prédictives récurrentes, telles que celles qui sont identifiées pour dissocier ce problème (Perrinet 2012).
 
+Le modèle est entraîné sur une tâche de bas niveau de détection de mouvement local, et on peut se demander s'il pourrait être entraîné sur des tâches de plus haut niveau.
+Un exemple de tâche serait par exemple celui de l'estimation de la profondeur dans la scène visuelle.
+Il existe de multiples sources d'information pour déduire la profondeur, comme la disparité binoculaire ou les changements de texture ou d'ombrage, mais dans notre cas, la parallaxe du mouvement serait l'indice le plus significatif (Rogers).
+En effet, les objets proches d'un observateur se déplacent relativement plus rapidement sur la rétine qu'un objet situé à une grande distance.
+En utilisant cette information, on peut segmenter les objets et estimer leur profondeur (Yoonessi).
+Cependant, cela nécessiterait d'abord de calculer le flux optique.
+Une implémentation possible est donc d'ajouter une nouvelle couche à notre modèle, en analogie avec l'organisation hiérarchique mise en évidence dans le cortex visuel.
+Cela est en théorie possible en utilisant la sortie de notre modèle (qui estime la vitesse dans l'espace rétinotopique) comme entrée d'une nouvelle couche de neurones qui estimerait la vitesse dans le champ visuel, en incluant la dimension de profondeur comme étiquette de sortie.
+Cela pourrait avoir des applications directes et importantes, par exemple dans la conduite autonome, afin de détecter les obstacles de manière rapide et robuste.
+Une autre extension serait de produire activement des mouvements du capteur (physiquement ou virtuellement) pour produire cette estimation (Nawrot).
+
+Pour conclure, le modèle que nous avons présenté offre un moyen de traiter efficacement les signaux basés sur des événements.
+Nous avons montré que nous pouvons entraîner le modèle de manière semi-supervisée, en connaissant les étiquettes de sortie mais sans savoir quand elles se produisent.
+Une autre perspective consisterait à étendre le modèle à un paradigme d'apprentissage entièrement auto-supervisé, c'est-à-dire sans aucune donnée étiquetée (Barlow).
+On pense que ce type d'apprentissage est prédominant dans le système nerveux central et, en supposant que le signal est parcimonieux (Olshausen), on pourrait étendre les schémas d'apprentissage hebbien parcimonieux aux spikes (Perrinet Masquelier).
+Un tel schéma serait particulièrement bien adapté à l'exploration de données neurobiologiques.
+En effet, il existe une littérature importante indiquant que la dynamique cérébrale s'organise souvent en séquences stéréotypées comme des chaînes synfire (Ikegaya), des paquets (Luczak) ou des séquences hippocampiques (Pastalkora 2008 Villette 2015).
+Ces motifs sont stéréotypés et robustes, car ils peuvent être activés selon le même motif d'un jour à l'autre (Haimerl 2019).
+Contrairement aux méthodes classiques qui sont utilisées pour traiter les données neurobiologiques, un tel modèle serait capable de répondre à des questions clés concernant la représentation de l'information dans les données neuronales.
+En outre, il ouvrirait des possibilités dans le domaine de l'apprentissage automatique, en particulier dans celui de la vision par ordinateur, pour résoudre les principales incertitudes actuelles telles que la robustesse aux attaques, la scalabilité, l'interprétabilité ou la consommation d'énergie.
+
+%%%%%%%%%%%%%%%%%%%%%%%%
 \backmatter
@@
 % Acknowledgement hidden for review.
-This work was supported by the European Union’s ERA-NET CHIST-ERA 2018 research and innovation programme under grant agreement ANR-19-CHR3-0008.
-
-% Nef acknowledgements 
-The authors are grateful to the OPAL infrastructure from Université Côte d'Azur for providing resources and support.
+\Acknowledgments
 

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
perspectives

--- 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
@@
 % >>>>>>>>>>> LuP is here <<<<<<<<<<<
+In the definition of our task, we highlighted how the generation of events depends on the spatial gradient in each image.
+This gradient is defined in both horizontal and vertical dimensions and its maxima are generally oriented.
+Taken together, these oriented edges form the contours of visual objects in the scene (Koenderink).
+Thus, there is a dependence within the event stream between motion and orientation information.
+It would be crucial to study this interdependence further, which could be initiated by training the model on a dataset with orientation labels.
+Studying this dependence will allow us to dissociate these two forms of visual information and to enable us to integrate them.
+In particular it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka, the aperture problem), thus allowing us to implement recurrent predictive rules, such as those identified to dissociate this problem (Perrinet 2012).
 
+The model is trained on a low-level local motion detection task, and one might wonder if it could be trained on higher-level tasks.
+An example task would be depth estimation in the visual scene.
+There are multiple sources of information for inferring depth, such as binocular disparity or changes in texture or shading, but in our case, motion parallax would be the most significant cue (Rogers).
+This is because objects close to an observer move relatively faster on the retina than an object at a great distance.
+Using this information, one can segment objects and estimate their depth (Yoonessi).
+However, this would first require computing the optical flow.
+A possible implementation is therefore to add a new layer to our model, in analogy with the hierarchical organization highlighted in the visual cortex.
+This is theoretically possible by using the output of our model (which estimates velocity in retinotopic space) as input to a new layer of neurons that would estimate velocity in the visual field, including the depth dimension as an output label.
+This could have direct and important applications, for example in autonomous driving, to detect obstacles in a fast and robust way.
+Another extension would be to actively generate sensor movements (physically or virtually) to yield better depths estimates notably to resolve uncertain estimates (Nawrot).
+
+In conclusion, the model we have presented provides a way to efficiently process event-based signals.
+We have shown that we can train the model in a semi-supervised manner, knowing the output labels but not knowing when they occur.
+Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data (Barlow).
+This type of learning is thought to be predominant in the central nervous system and, assuming that the signal is sparse (Olshausen), one could extend those hebbian sparse learning schemes to spikes (Perrinet Masquelier).
+This would be particularly well suited to the exploration of neurobiological data.
+In fact, there is a substantial literature indicating that brain dynamics often organize into stereotyped sequences such as synfire chains (Ikegaya), packets (Luczak), or hippocampal sequences (Pastalkora 2008 Villette 2015).
+These patterns are stereotypical and robust, as they can be activated in the same motif from day to day (Haimerl 2019).
+In contrast to conventional methodologies that are used to process neurobiological data, such a event-based model would be able to answer key questions regarding the representation of information in neurobiological data.
+Furthermore, it would open up possibilities in the field of machine learning, especially in computer vision, to address current key concerns such as robustness to attacks, scalability, interpretability, or energy consumption.
+
+%%%%%%%%%%%%%%%%%%%%%%%%
 \backmatter

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
perspectives

@@
 This gradient is defined in both horizontal and vertical dimensions and its maxima are generally oriented.
-Taken together, these oriented edges form the contours of visual objects in the scene (Koenderink).
+Taken together, these oriented edges form the contours of visual objects in the scene~\citep{} (Koenderink).
 Thus, there is a dependence within the event stream between motion and orientation information.
@@
 Studying this dependence will allow us to dissociate these two forms of visual information and to enable us to integrate them.
-In particular it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka, the aperture problem), thus allowing us to implement recurrent predictive rules, such as those identified to dissociate this problem (Perrinet 2012).
+In particular it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka, the aperture problem), thus allowing us to implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{} (Perrinet 2012).
 
@@
 This is because objects close to an observer move relatively faster on the retina than an object at a great distance.
-Using this information, one can segment objects and estimate their depth (Yoonessi).
+Using this information, one can segment objects and estimate their depth~\citep{} (Yoonessi).
 However, this would first require computing the optical flow.
@@
 This could have direct and important applications, for example in autonomous driving, to detect obstacles in a fast and robust way.
-Another extension would be to actively generate sensor movements (physically or virtually) to yield better depths estimates notably to resolve uncertain estimates (Nawrot).
+Another extension would be to actively generate sensor movements (physically or virtually) to yield better depths estimates notably to resolve uncertain estimates~\citep{} (Nawrot).
 
@@
 This would be particularly well suited to the exploration of neurobiological data.
-In fact, there is a substantial literature indicating that brain dynamics often organize into stereotyped sequences such as synfire chains (Ikegaya), packets (Luczak), or hippocampal sequences (Pastalkora 2008 Villette 2015).
-These patterns are stereotypical and robust, as they can be activated in the same motif from day to day (Haimerl 2019).
+In fact, there is a substantial literature indicating that brain dynamics often organize into stereotyped sequences such as synfire chains (Ikegaya), packets~\citep{} (Luczak), or hippocampal sequences~\citep{} (Pastalkora 2008 Villette 2015).
+These patterns are stereotypical and robust, as they can be activated in the same motif from day to day~\citep{} (Haimerl 2019).
 In contrast to conventional methodologies that are used to process neurobiological data, such a event-based model would be able to answer key questions regarding the representation of information in neurobiological data.

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
@@
 This gradient is defined in both horizontal and vertical dimensions and its maxima are generally oriented.
-Taken together, these oriented edges form the contours of visual objects in the scene (Koenderink).
+Taken together, these oriented edges form the contours of visual objects in the scene~\citep{koenderink_representation_1987}.
 Thus, there is a dependence within the event stream between motion and orientation information.
@@
 Studying this dependence will allow us to dissociate these two forms of visual information and to enable us to integrate them.
-In particular it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka, the aperture problem), thus allowing us to implement recurrent predictive rules, such as those identified to dissociate this problem (Perrinet 2012).
+In particular it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka, the aperture problem), thus allowing us to implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
 
@@
 This is because objects close to an observer move relatively faster on the retina than an object at a great distance.
-Using this information, one can segment objects and estimate their depth (Yoonessi).
+Using this information, one can segment objects and estimate their depth~\citep{yoonessi_contribution_2011}.
 However, this would first require computing the optical flow.
@@
 This could have direct and important applications, for example in autonomous driving, to detect obstacles in a fast and robust way.
-Another extension would be to actively generate sensor movements (physically or virtually) to yield better depths estimates notably to resolve uncertain estimates (Nawrot).
+Another extension would be to actively generate sensor movements (physically or virtually) to yield better depths estimates notably to resolve uncertain estimates~\citep{nawrot_eye_2003}.
 
@@
 We have shown that we can train the model in a semi-supervised manner, knowing the output labels but not knowing when they occur.
-Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data (Barlow).
-This type of learning is thought to be predominant in the central nervous system and, assuming that the signal is sparse (Olshausen), one could extend those hebbian sparse learning schemes to spikes (Perrinet Masquelier).
+Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data~\citep{barlow_unsupervised_1989}.
+This type of learning is thought to be predominant in the central nervous system and, assuming that the signal is sparse~\citep{olshausen_emergence_1996}, one could extend those Hebbian sparse learning schemes to spikes~\citep{perrinet_emergence_2004, masquelier_competitive_2009}.
 This would be particularly well suited to the exploration of neurobiological data.
-In fact, there is a substantial literature indicating that brain dynamics often organize into stereotyped sequences such as synfire chains (Ikegaya), packets (Luczak), or hippocampal sequences (Pastalkora 2008 Villette 2015).
-These patterns are stereotypical and robust, as they can be activated in the same motif from day to day (Haimerl 2019).
+In fact, there is a substantial literature indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}.
+These patterns are stereotypical and robust, as they can be activated in the same motif from day to day~\citep{haimerl_internal_2019}.
 In contrast to conventional methodologies that are used to process neurobiological data, such a event-based model would be able to answer key questions regarding the representation of information in neurobiological data.

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 They are produced very rapidly (about $80~\ms$) and display a whole range of speeds.
-At a more microscopic scale, the human gaze moves with an incessant drift similar to a Brownian-like trajectory~\citep{Poletti 2015}.
-To maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{Engbert 2011}.
+At a more microscopic scale, the human gaze moves with an incessant drift similar to a Brownian-like trajectory~\citep{poletti_head-eye_2015}.
+To maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{eng}.
 This approach first defines a finite set of possible 2D motions in polar coordinates.

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
scan

--- 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
@@
 \section{Introduction}
+%%%-----------------------------------------------------------------
+%%%-----------------------------------------------------------------
 \label{sec:intro}
+% >>>>>>>>>>> LuP is here <<<<<<<<<<<
 %
 \subsection{Processing of dynamical sensory inputs}
-In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\citep{abeles_role_1982}.
+%%%-----------------------------------------------------------------
+In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking motifs~\citep{abeles_role_1982}.
 While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code.
 Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
-All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in biological spike trains.
+All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal motifs in biological spike trains.
 
 \subsection{Hetero-synaptic delays in spiking neurons}
-In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.
+%%%-----------------------------------------------------------------
+In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.
 Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays.
@@
 \subsection{Outline}
-
+%%%-----------------------------------------------------------------
 In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}.
@@
 \label{sec:methods}
-%
-%
+%%%-----------------------------------------------------------------
+%%%-----------------------------------------------------------------
 Let us now formally define the hetero-synaptic model, as well as the task and how we will solve it.
@@
 We will finally describe its implementation on conventional computers.
-
+%
 \subsection{Hetero-synaptic delays model as a Logistic Regression with a temporal convolution}
-
+%%%-----------------------------------------------------------------
 Let's first define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
@@
 Each event has a time of occurrence $\timev_\arank$.
-The corresponding set of input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible patterns.
+The corresponding set of input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible motifs.
 For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
@@
 This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
-This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different patterns in the event stream.
+This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
 %
@@
 The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model.
-
+%
 \subsection{Task definition: fast motion detection}
-%
+%%%-----------------------------------------------------------------
 Let us now define a trajectory inspired by the biological movements of the eyes.
@@
 Note that the velocities are sampled on a geometric scale between $1/4$ and $v-speed_base = 4$.
-Next, we define one trajectory of gaze as segments whose duration is drawn from a Poisson distribution with an average block length = 20 ms - similar to a Lévy flight~\citep{}.
+Next, we define one trajectory of gaze as segments whose duration is drawn from a Poisson distribution with an average block length = $20~\ms$ - similar to a Lévy flight~\citep{}.
 Finally, the trajectory is integrated by assuming 1/ that velocities are uniformly and independently sampled from the set of different motion-sets and 2/ that motion is uniform during a time segment.
@@
 
+%%%-----------------------------------------------------------------
 \begin{figure}[h!]%[ht!]
@@
 \end{figure}
-
+%%%-----------------------------------------------------------------
 Once these eye movement trajectories are generated, we can apply them to a visual scene.
@@
 This signed threshold value is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is then similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion-task}-(c)).
+When applied to the whole movie, the event stream is as a consequence similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion-task}-(c)).
 The goal here is to infer the correct motion solely by observing these events.
@@
 \note{ the task is full-field rigid translations, the model resolves it locally (analogy to the input from V1 to MT) }
-
+%
 \subsection{Extension as 3D convolutions for motion detection}
-%
+%%%-----------------------------------------------------------------
 In particular, we may extend the convolution to a 3D convolution such that the representation would also benefit from spatial invariance which is inherent to the task's construction.
@@
 %
+\note{kernel size has to be adapted to the characteristics of motion see Samonds %~\citep{samonds}
+}
+%
 \section{Results}
 \label{sec:results}
-%
+%%%-----------------------------------------------------------------
+%%%-----------------------------------------------------------------
 \subsection{Kernels learned for motion detection}
-%
+%%%-----------------------------------------------------------------
 \begin{figure*}[ht!]
@@
 
-If one focuses on the interpretation of these kernels in terms of spatio-temporal patterns embedded in the event stream, it can lead to interesting outcomes.
+If one focuses on the interpretation of these kernels in terms of spatio-temporal motifs embedded in the event stream, it can lead to interesting outcomes.
 In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions.
-Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spike pattern.
+Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spike motif.
 Each voxel in the 3D kernels defines a specific timestamp and a specific address.
-Then, our model is able to detect precise spatio-temporal patterns embedded in the spike train and associated to the different motion directions.
+Consequently, our model is able to detect precise spatio-temporal motifs embedded in the spike train and associated to the different motion directions.
 The cone shape for the positive weights distribution highlights a loss of precision for longer delays, i.e.
@@
 \subsection{Accuracy for the motion detection task}
-%
+%%%-----------------------------------------------------------------
+
 Once our MLR is trained, we obtain 3D kernels corresponding to the weights associated to the hetero-synaptic delays of our layer of spiking neurons and which may be used for detection.
@@
 In this case, the number of computations is greatly reduced compared to~\citep{grimaldi_robust_2022}, thus demonstrating the efficiency of the presented method.
+
 %
+%%%-----------------------------------------------------------------
 \begin{figure}[h!]
@@
 \end{figure}
+%%%-----------------------------------------------------------------
 %
-
 \subsection{Testing with natural-like textures}
-%
+%%%-----------------------------------------------------------------
 To test our model, we will quantify its ability to categorize different motions.
@@
 This signed threshold value is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is then similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motionclouds}-(c)).
+When applied to the whole movie, the event stream is threfore similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motionclouds}-(c)).
 The goal here is to infer the correct motion solely by observing these events.
-
+%
 \section{Discussion}
-
-Here, we have introduced a generic SNN using hetero-synaptic delays and shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification.
+%%%-----------------------------------------------------------------
+%%%-----------------------------------------------------------------
+Here, we have introduced a generic SNN using hetero-synaptic delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification \note{to change if we compare to ghosh instead of HOTs}.
 The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification.
 This shows that we may use the precise timing of spikes to enhance neural computations.
-
+%
 \subsection{Synthesis and main contributions}
+%%%-----------------------------------------------------------------
 To synthesize, we have introduced in this paper a spiking neural network model with hetero-synaptic delays that we have trained and evaluated on a complex motion detection task.
-The model was defined to optimally detect event-driven spatiotemporal patterns.
-We showed that when trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example.
-We then showed that the model has similar responses to those found in biology.
-We then evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of hetero-synaptic delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understanding why spikes represent a universal component of neuronal information processing.
+The model was defined to optimally detect event-driven spatiotemporal motifs.
+We have shown that when the model is trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example.
+We have then shown that the model displays similarities with the responses observed in biology.
+We have evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of hetero-synaptic delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understand why spikes represent a universal component of neuronal information processing.
 
@@
 First, the generic heterosynaptic model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs, whereas~\citet{ghosh_spatiotemporal_2019} use a correlation-based heuristic, which we have proven to be less efficient.
-Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking pattern.
+Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking motif.
 Another novelty is that the model simultaneously learns the weights and the delays, while for example the polychronization model~\citep{izhikevich_polychronization_2006} only learns the weights using STDP while the delays are randomly drawn and their values frozen.
 Moreover, the model is evaluated on a realistic task, while models such as tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}.
-The main contribution is to provide a model that is suitable for any type of spatio-temporal spiking patterns and that can be trained in a supervised way by providing a dataset of supervision pairs.
+The main contribution is to provide a model that is suitable for any type of spatio-temporal spiking motifs and that can be trained in a supervised way by providing a dataset of supervision pairs.
 This allows for a more flexible definition of the model using this properly labeled dataset instead of relying on a careful description of the physical rules governing a task, e.g., the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}.
-
+%
 \subsection{Main limits}
-
-%One advantage of our model is the generality of the possible kernels.
-However, this can be a limit as this increases the number of free parameters in the learning algorithm.
-As we observe that the timing of events in natural scene may be slightly variable (in the order of the synaptic time constant of $5~\ms$) an extension of the model would be to include a fixed filtering stage similar to that in the tempotron model~\citep{gutig2006tempotron}.
-This could be efficiently added as a temporal convolution on $\mathcal{C}^\postsynaddr(t)$.
-%\noteAG{C'est pas clair pour moi ce que veut dire ou ce qu'ajoute le paragraphe précédent.
-La généralité des kernels est données dans le suivant et celui là fait appel à des concepts de filtrage pas très explicités.
-Aussi la référence au timing des events des scènes naturelles peut être justifié par une référence.
-}
+%%%-----------------------------------------------------------------
 We have identified several limits to our model, the most important being now detailed.
 First, the complete framework is based on a discrete binning of time that is incompatible with the continuous nature of biological time.
-We used this binning to be able to efficiently implement the framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions.
-We tested the effect of the size of the time bin and showed that it has essentially no impact on the results presented in this paper.
-This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where the accuracy was not affected when the input spikes were subjected to noise up to 1ms~\citep{grimaldi_robust_2022}.
-This suggests the possibility of analytically including a precision term in the temporal value of the input spikes, a mechanism potentially implemented by the approximately 5 ms filtering that is implemented at the PSC of synapses.
-Furthermore, it is possible to overcome the problem of time discretization by using a purely event-based scheme.
-Indeed, it is not necessary to compute voltage traces between two spikes~\citep{hanuschkin_general_2010} and thus define a purely event-based framework.
+We have used this binning to be able to efficiently implement the framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions.
+We have tested the effect of the size of the time bin and shown that it has essentially no impact on the results presented in this paper.
+This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where the accuracy was not affected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}.
+This suggests the possibility of analytically including a precision term in the temporal value of the input spikes, a mechanism potentially implemented by the filtering that is implemented by the synaptic time constant of about $5~\ms$.
+Furthermore, it is possible to circonvent the necessity of using a time discretization by using a purely event-based scheme.
+Indeed, it is not necessary to compute voltage traces between two spikes~\citep{hanuschkin_general_2010} and it is thus possible to define a purely event-based framework.
 Such an architecture could provide promising speed gains for the calculations.
@@
 Therefore, the spikes generated by the postsynaptic neurons are produced solely on the basis of information contained in the classical receptive field.
-However, it is well known that neurons in a layer can interact using lateral interactions, for instance in V1~\citep{chavane_revisiting_2022} and that this can be the basis for computational principles, for example, the combination of neighbouring orientations which is a candidate for the image categorization mechanism~\citep{perrinet_edge_2015}.
-Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background~\citep{roelfsema_early_2016} and has been shown to be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020, boutin_pooling_2022}.
+However, it is well known that neurons in a layer can interact using lateral interactions, for instance in V1 and that this can be the basis for computational principles~\citep{chavane_revisiting_2022}.
+For example, the combination of neighboring orientations may contribute to image categorization~\citep{perrinet_edge_2015}.
+Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background~\citep{roelfsema_early_2016} and it has been shown that feedback may be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020, boutin_pooling_2022}.
 It is currently not possible to implement these recurrent connections in our implementation (lateral or feedback), mainly due to our use of convolutions.
@@
 This has been previously modeled to explain for example the flash-lag illusion~\citep{khoei_flash-lag_2017}.
-As previously stated, this could be implemented using generalized coordinates (that is variables such as position which are complemented by speed, acceleration, jerk, ...) and ``neurobiologically, the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
+\note{Say that this could be implemented by properly setting the timing of the label} As previously stated, this could be implemented using generalized coordinates (that is variables such as position which are complemented by speed, acceleration, jerk, ...) and ``neurobiologically, the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
 Our proposed hetero-synaptic model provides an alternate and elegant implementation solution to this problem.
-
+%
 \subsection{Perspectives}
-
-%Note that this supervised learning scheme can be extended to a variety of tasks.
-It would follow from the emergence of new kernels adapted to this new task after supervised learning.
-This constitutes a major advantage over other algorithms which derive event-based algorithms from specific physical rules (see for instance~\citep{benosman_asynchronous_2012} for computing the optic flow using the luminance conservation rule).
-We aim at extending the application of this model on more generic datasets acquired in natural conditions for progressively more complex tasks from motion extraction, optic flow or time-to-contact maps.
-% >>>>>>>>>>> LuP is here <<<<<<<<<<<
+%%%-----------------------------------------------------------------
 In the definition of our task, we highlighted how the generation of events depends on the spatial gradient in each image.
@@
 Thus, there is a dependence within the event stream between motion and orientation information.
-It would be crucial to study this interdependence further, which could be initiated by training the model on a dataset with orientation labels.
+It would be crucial to study this interdependence further.
+This could be initiated by training the model on a dataset with orientation labels.
 Studying this dependence will allow us to dissociate these two forms of visual information and to enable us to integrate them.
@@
 In fact, there is a substantial literature indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}.
-These patterns are stereotypical and robust, as they can be activated in the same motif from day to day~\citep{haimerl_internal_2019}.
+These motifs are stereotypical and robust, as they can be activated in the same motif from day to day~\citep{haimerl_internal_2019}.
 In contrast to conventional methodologies that are used to process neurobiological data, such a event-based model would be able to answer key questions regarding the representation of information in neurobiological data.
 Furthermore, it would open up possibilities in the field of machine learning, especially in computer vision, to address current key concerns such as robustness to attacks, scalability, interpretability, or energy consumption.
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%
 \backmatter
-
+%
 \bmhead{Acknowledgments}
-
+%
 % Acknowledgement hidden for review.
 \Acknowledgments
-
+%
 \section*{Declarations}

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.
-As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
+As such, we may first write any event-based input as the Boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
 In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the nonzero weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
@@
 At a more microscopic scale, the human gaze moves with an incessant drift similar to a Brownian-like trajectory~\citep{poletti_head-eye_2015}.
-To maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{eng}.
+To maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.
 This approach first defines a finite set of possible 2D motions in polar coordinates.

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
scan pass

@@
 It would be crucial to study this interdependence further.
-This could be initiated by training the model on a dataset with orientation labels.
+This could be initiated by training the model on a dataset with labels providing local orientation.
+\note{we can say we tried and that it works well}.
 Studying this dependence will allow us to dissociate these two forms of visual information and to enable us to integrate them.
@@
 An example task would be depth estimation in the visual scene.
-There are multiple sources of information for inferring depth, such as binocular disparity or changes in texture or shading, but in our case, motion parallax would be the most significant cue (Rogers).
+There are multiple sources of information for inferring depth, such as binocular disparity or changes in texture or shading, but in our case, motion parallax would be the most significant cue~\citep{} (Rogers).
 This is because objects close to an observer move relatively faster on the retina than an object at a great distance.
 Using this information, one can segment objects and estimate their depth~\citep{yoonessi_contribution_2011}.
-However, this would first require computing the optical flow.
+However, this would require computing first the optical flow.
 A possible implementation is therefore to add a new layer to our model, in analogy with the hierarchical organization highlighted in the visual cortex.
-This is theoretically possible by using the output of our model (which estimates velocity in retinotopic space) as input to a new layer of neurons that would estimate velocity in the visual field, including the depth dimension as an output label.
+This is theoretically possible by using the output of our model (which estimates velocity in retinotopic space) as input to a new layer of neurons that would estimate velocity in the visual field, including the depth dimension in the output supervision labels.
 This could have direct and important applications, for example in autonomous driving, to detect obstacles in a fast and robust way.

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 An example task would be depth estimation in the visual scene.
-There are multiple sources of information for inferring depth, such as binocular disparity or changes in texture or shading, but in our case, motion parallax would be the most significant cue~\citep{} (Rogers).
+There are multiple sources of information for inferring depth, such as binocular disparity or changes in texture or shading, but in our case, motion parallax would be the most significant cue~\citep{rogers_motion_1979}.
 This is because objects close to an observer move relatively faster on the retina than an object at a great distance.

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
scan pass

@@
 There are multiple sources of information for inferring depth, such as binocular disparity or changes in texture or shading, but in our case, motion parallax would be the most significant cue~\citep{rogers_motion_1979}.
-This is because objects close to an observer move relatively faster on the retina than an object at a great distance.
+This is because objects close to an observer move relatively faster on the retina than an object at a great distance, and also that visual occlusions depend on depth ordering.
 Using this information, one can segment objects and estimate their depth~\citep{yoonessi_contribution_2011}.
@@
 This could have direct and important applications, for example in autonomous driving, to detect obstacles in a fast and robust way.
-Another extension would be to actively generate sensor movements (physically or virtually) to yield better depths estimates notably to resolve uncertain estimates~\citep{nawrot_eye_2003}.
+Another extension would be to actively generate sensor movements (physically or virtually) to yield better depths estimates, notably to disambiguate uncertain estimates~\citep{nawrot_eye_2003}.
 
 In conclusion, the model we have presented provides a way to efficiently process event-based signals.
-We have shown that we can train the model in a semi-supervised manner, knowing the output labels but not knowing when they occur.
+We have shown that we can train the model in a semi-supervised manner, knowing \emph{what} output label but not knowing \emph{when} it occurs.
 Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data~\citep{barlow_unsupervised_1989}.
 This type of learning is thought to be predominant in the central nervous system and, assuming that the signal is sparse~\citep{olshausen_emergence_1996}, one could extend those Hebbian sparse learning schemes to spikes~\citep{perrinet_emergence_2004, masquelier_competitive_2009}.
-This would be particularly well suited to the exploration of neurobiological data.
+We anticipate that this would be particularly well suited to the exploration of neurobiological data.
 In fact, there is a substantial literature indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}.

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 \newcommand{\Street}{27 boulevard Jean Moulin}%
+\newcommand{\PostCode}{13005}%
 \newcommand{\City}{Marseille}%
@@
 % ---------------
-
-\author*{\fnm{Antoine} \sur{Grimaldi}}\email{antoine.grimaldi@univ-amu.fr}
-
-\author{\fnm{Laurent} \sur{Perrinet}}\email{laurent.perrinet@univ-amu.fr}
-
-\affil{\orgdiv{\Department}, \orgname{\Affiliation}\\ \orgaddress{\street{\Street}, \city{\City}, \postcode{13005}, \country{\Country}}}
-
-%\author*[1]{\fnm{Antoine} \sur{Grimaldi}}\email{antoine.grimaldi@univ-amu.fr}
-%
-%\author[2]{\fnm{Amélie} \sur{Gruel}}\email{amelie.gruel@univ-cotedazur.fr}
-%
-%\author[2]{\fnm{Jean} \sur{Martinet}}\email{jean.martinet@univ-cotedazur.fr}
-%
-%\author[1]{\fnm{Laurent} \sur{Perrinet}}\email{laurent.perrinet@univ-amu.fr}
-%
-%\affil[1]{\orgdiv{Department}, \orgname{\Affiliation}, \orgaddress{\street{Street}, \city{Marseille}, \postcode{13005}, \country{France}}}
-%\affil[2]{\orgdiv{SPARKS}, \orgname{Université Côte d'Azur, CNRS, I3S}, \orgaddress{\street{2000 Rte des Lucioles}, \city{Sophia-Antipolis}, \postcode{06900}, \country{France}}}
+\author*{\fnm{\FirstAG} \sur{\LastAG}}\email{\EmailAG}
+\author{\fnm{\FirstLP} \sur{\LastLP}}\email{\EmailLP}
+\affil{\orgdiv{\Department}, \orgname{\Affiliation}\\ \orgaddress{\street{\Street}, \city{\City}, \postcode{\PostCode}, \country{\Country}}}
 
@@
 
-%
 %\ninept
@@
 To avoid border effects, we will position a draw of the complete trajectory at random in the image space so that the sub-image is translated using the position given by the trajectory at each time step.
-The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{Perrinet, 2015}.
+The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
 Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that a velocity of $V=1$ corresponds to a movement of one pixel per frame (i.e., per time bin).
@@
 %
+%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Declarations}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliographystyle{apalike}
@@
 \end{document}
+%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
scan pass

@@
 \newcommand{\orcidAG}{0000-0002-3107-4788}
-\newcommand{\Department}{Institut de Neurosciences de la Timone (UMR 7289)}%
+\newcommand{\Department}{Institut de Neurosciences de la Timone}% (UMR 7289)}%
 \newcommand{\Affiliation}{Aix Marseille Univ, CNRS}%
@@
 To avoid border effects, we will position a draw of the complete trajectory at random in the image space so that the sub-image is translated using the position given by the trajectory at each time step.
-The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
+The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{Perrinet, 2015}.
 Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that a velocity of $V=1$ corresponds to a movement of one pixel per frame (i.e., per time bin).

### 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2022-09-15 Grimaldi-etal-BiolCybernetics.tex
@@
 Once these eye movement trajectories are generated, we can apply them to a visual scene.
-For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images~\citep{olshausen 96}.
+For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images~\citep{olshausen_emergence_1996}.
 Note that these are pre-processed to equalize the energy in each frequency band (i.e., whitened).
-This process is known to occur as early as the retino-thalamic pathway~\citep{dan 1996}.
+This process is known to occur as early as the retino-thalamic pathway~\citep{dan_efficient_1996}.
 These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step (see \seeFig{motion-task}-(b)).

### 2022-09-16 Grimaldi-etal-BiolCybernetics.tex
new figures

@@
 }
-\newcommand{\Acknowledgments}{This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}'').
+\newcommand{\Acknowledgments}{
+Thanks to CB, TS, ...
+This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}'').
 %HL and LP received funding from the ANR project  N° ANR-20-CE23-0021 (''\href{https://laurentperrinet.github.io/grant/anr-anr/}
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{figures/motion-task.pdf}
+    \includegraphics[width=0.95\linewidth]{figures/motion_task.pdf}
     \note{TODO: redo figure according to caption}
@@
     \vspace{-3cm}
-    \includegraphics[width=\linewidth]{figures/3D_kernels_oneline.png}
+    \includegraphics[width=\linewidth]{figures/motion_kernels.pdf}
     }
@@
 Such an architecture could provide promising speed gains for the calculations.
+\note{pourrait parler du poster bernstein?
+de nadafian?}
 
@@
 Using this information, one can segment objects and estimate their depth~\citep{yoonessi_contribution_2011}.
-However, this would require computing first the optical flow.
+However, this would require computing first the optical flow, that is extending the framework described here for a rigid full-field motion to a more generic one where motion may vary in the visual field.
 A possible implementation is therefore to add a new layer to our model, in analogy with the hierarchical organization highlighted in the visual cortex.

### 2022-09-16 Grimaldi-etal-BiolCybernetics.tex
figures

@@
 
-In this work, we study the emergence of such spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motion-task}).
+In this work, we study the emergence of such spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motion_task}).
 We develop a SNN-like method able to learn hetero-synaptic delays to perform motion detection on a synthetic event-based dataset.
@@
 By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory.
-The task is then to guess the motion as accurately as possible (see \seeFig{motion-task}).
+The task is then to guess the motion as accurately as possible (see \seeFig{motion_task}).
 This is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina.
@@
 This approach first defines a finite set of possible 2D motions in polar coordinates.
-Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of N-V-phi=8 linearly spaced motion directions and N-Y-velocity=7 different velocities (see inset of \seeFig{motion-task}-(a)).
+Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of N-V-phi=8 linearly spaced motion directions and N-Y-velocity=7 different velocities (see inset of \seeFig{motion_task}-(a)).
 Note that the velocities are sampled on a geometric scale between $1/4$ and $v-speed_base = 4$.
@@
 Finally, the trajectory is integrated by assuming 1/ that velocities are uniformly and independently sampled from the set of different motion-sets and 2/ that motion is uniform during a time segment.
-The resulting instances yield trajectories qualitatively similar to those observed for human eye movements (\seeFig{motion-task}-(a)).
+The resulting instances yield trajectories qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(a)).
 
+%%%-----------------------------------------------------------------
+%: fig:motion_task
 %%%-----------------------------------------------------------------
@@
 This will constitute the input to the spiking neural network.}
-    \label{fig:motion-task}
+    \label{fig:motion_task}
 \end{figure}
@@
 This process is known to occur as early as the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step (see \seeFig{motion-task}-(b)).
+These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step (see \seeFig{motion_task}-(b)).
 We will discretize the time in $1~\ms$ bins and produce movies of duration N-T = $250~\ms$.
@@
 This signed threshold value is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is as a consequence similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion-task}-(c)).
+When applied to the whole movie, the event stream is as a consequence similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion_task}-(c)).
 The goal here is to infer the correct motion solely by observing these events.
@@
 %%%-----------------------------------------------------------------
+%: fig:kernels
 %%%-----------------------------------------------------------------
@@
     {\centering
-    \vspace{-3cm}
+    %\vspace{-3cm}
     \includegraphics[width=\linewidth]{figures/motion_kernels.pdf}
@@
 %%%-----------------------------------------------------------------
+%: fig:accuracy
+%%%-----------------------------------------------------------------
 \begin{figure}[h!]
@@
 %
-\begin{figure}[h!]%[ht!]
+%%%-----------------------------------------------------------------
+%: fig:motion_clouds
+%%%-----------------------------------------------------------------
+\begin{figure}[h!]
     \centering
-    \includegraphics[width=0.95\linewidth]{figures/motion-clouds.pdf}
+    \includegraphics[width=0.95\linewidth]{figures/motion_clouds.pdf}
     \caption{{\bf Motion detection task.} {\bf (a)} The motion direction represented as the plain red vector, other possible motion directions are represented in light red.
@@
     {\bf (c)} The corresponding ON (in red) and OFF (in blue) event stream generated from the stimuli on (b) and constituting the input to the spiking neural network.}
-    \label{fig:motionclouds}
+    \label{fig:motion_clouds}
 \end{figure}
+%%%-----------------------------------------------------------------
 

### 2022-09-16 Grimaldi-etal-BiolCybernetics.tex
minor

@@
     \includegraphics[width=0.95\linewidth]{figures/motion_clouds.pdf}
-    \caption{{\bf Motion detection task.} {\bf (a)} The motion direction represented as the plain red vector, other possible motion directions are represented in light red.
+    \caption{{\bf Role of image parameters in the motion detection accuracy.} {\bf (a)} The motion direction represented as the plain red vector, other possible motion directions are represented in light red.
 {\bf (b)} A screenshot of one generated naturalistic textured stimulus at a specific time.

### 2022-09-19 Grimaldi-etal-BiolCybernetics.tex
notes

@@
 %%%-----------------------------------------------------------------
+\note{c'est la meme expe faite par baudot et modelisee par kremkow}
 Let us now define a trajectory inspired by the biological movements of the eyes.
@@
 %%%-----------------------------------------------------------------
+\note{la tache ne met pas en avant les capacites des motifs hetero synaptiques =trop simple et peut se faire en HOTS}
 In the definition of our task, we highlighted how the generation of events depends on the spatial gradient in each image.

### 2022-09-22 Grimaldi-etal-BiolCybernetics.tex
up
import from 2022-07-11_Grimaldi-etal-ICIP.tex

@@
+% !TEX encoding = UTF-8 Unicode
 %!TeX TS-program = pdfLaTeX
-%!TeX encoding = UTF-8 Unicode
 %!TeX spellcheck = en-US
@@
 \newcommand{\Title}{
-Learning hetero-synaptic delays for fast motion detection in a single layer of spiking neurons
+Learning hetero-synaptic delays in a layer of spiking neurons for fast motion detection
 }
@@
 
+\usepackage[T1]{fontenc}
+%\usepackage{textcomp}
+\usepackage[utf8]{inputenc}
+
 %%%% Standard Packages
 %%<additional latex packages if required can be included here>
-\usepackage[lofdepth,lotdepth,position=top]{subfig}
-\usepackage{tabularx}
+%\usepackage[lofdepth,lotdepth,position=top]{subfig}
+%\usepackage{tabularx}
 %%%%
@@
 %\newcommand{\noteAG}[1]{\todo[author=\textbf{Antoine}, inline]{#1}}
-\usepackage{soul}
+\usepackage{soulutf8}
 \usepackage{color}
 
-\DeclareRobustCommand{\note}[1]{{\sethlcolor{yellow}\hl{#1}}}
+%\DeclareRobustCommand
+\newcommand{\note}[1]{{\sethlcolor{yellow}\hl{#1}}}
 %\newcommand{\noteLP}[1]{\todo[author=\textbf{Laurent}, inline]{#1}}
@@
 \note{finish by saying that it is effortless in biology but difficult on conventional computers}
+%
+\begin{figure}[t!]
+    \centering
+    \includegraphics[width=0.90\linewidth]{figures/figure_1a.pdf}
+    \includegraphics[width=0.90\linewidth]{figures/figure_1b.pdf}
+    \includegraphics[width=0.90\linewidth]{figures/figure_1c.pdf}
+	    \caption{The hetero-synaptic delays model applied on synthetic data.
+{\bf (a)}~The afferent information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs).
+We highlight them by different colors, an information hidden to a detection model.
+{\bf (b)}~The model is defined as an assembly of neurons (here for $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
+Note that each afferent may be connected with multiple weights at different delays.
+The propagation of the afferent information through these delay may generate at each time step a synchronous pattern on a subset of synapses.
+%These PG patterns were efficiently learned by a logistic regression.
+	    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG pattern at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
+}
+    \label{fig:model}
+\end{figure}
+
+% 
+The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
+This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
+This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
+
+We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then not only described by synaptic weights connecting to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+For each neuron $\postsynaddr$, we define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
+Note that a neuron may contact an afferent neuron with multiple, different delays.
+
+The corresponding input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible patterns.
+For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
+$
+\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
+$.
+%
+The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\cite{Grimaldi22pami}, in analogy to a spiking Winner-Take-All network~\cite{nessler2013bayesian}.
+It transforms this list of weights into a probability with the following formula:
+$
+Pr(k=\postsynaddr \; | \; \timev) =
+\frac 1 Z
+{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
+$ 
+where $\mathcal{C}^\postsynaddr(\timev) = \sum
+\activeweights^\postsynaddr(t)
+$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
+In particular, some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
+%
 
@@
 %
-In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+%In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
 The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model.
 %
+%
+In our MLR model with $\Nclass=\Nspeed$ classes, 
+a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model. 
+From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representations.
+As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
+In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
+Using this dense representation, the counting defined above becomes:
+\begin{equation*}
+\mathcal{C}^\postsynaddr(a,t)
+= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
+\end{equation*}
+%
+This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
+This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different patterns in the event stream.
+%
+In particular, we may extend the convolution to a 3D convolution 
+such that the representation would also benefit from spatial invariance.
+The use of spatiotemporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\cite{ghosh2022conv3D}.
+For that, we design 3D kernels of shape  $(\Kx, \Ky, \Ktime) = (15, 15, 8)$, respectively representing the two spatial dimensions and the range of delays.
+Computations are performed on spatio-temporal windows, defined by the kernels, sliding around the events, that is, the center of the spatio-temporal window around the current event $\epsilon_\arank$.
+%
+Finally, the output of the MLR model results in an event with the highest probability class, keeping the same timing as the event as input.
+%
+The loss function of the MLR model is the binary cross entropy on the output of the classification layer.
+Simulations are performed thanks to the PyTorch library using gradient descent with Adam ($2^{12}$ epochs and a learning rate of $10^{-5}$).
+%
+
 \subsection{Task definition: fast motion detection}
 %%%-----------------------------------------------------------------
-\note{c'est la meme expe faite par baudot et modelisee par kremkow}
+\note{c'est la meme expe faite par baudot et modelisée par kremkow}
 Let us now define a trajectory inspired by the biological movements of the eyes.

### 2022-09-23 Grimaldi-etal-BiolCybernetics.tex
import from 2021-11-20_COSYNE_abstract.tex
Methods

@@
 %%%-----------------------------------------------------------------
-Let us now formally define the hetero-synaptic model, as well as the task and how we will solve it.
-First, we will define the model for supervised learning of weights and delays in a feed-forward layer of a spiking neural network (SNN).
+Let us now formally define the hetero-synaptic model, as well as the task and how we will learn to solve it.
+First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neural network (SNN) with hetero-synaptic delays.
 Our SNN model will be tested on a computationally relevant task of motion detection in an event-based sensory input.
 By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory.
-The task is then to guess the motion as accurately as possible (see \seeFig{motion_task}).
-This is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina.
+Solving is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina.
+The task is thus to guess the motion as accurately as possible (see \seeFig{motion_task}) and we will describe a supervised learning mechanism for weights and delays.
 To this end, we will extend the SNN model to be efficient for the resolution of the task by including convolutional kernels.
@@
 This defines the full connectivity of the hetero-synaptic model.
+%In general, a spiking neuron is defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree.
+A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree.
+A more detailed examination of these synaptic contacts reveals the added importance of hetero-synaptic delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{Izhikevich06}.
+Thus, we can parameterize each neuron by the set of tuples defining the weight and delay of each synaptic contact.
+By discretization of time (with here an arbitrary unitary time unit), we can also define the dendrite as a matrix giving the weight corresponding to the different delays $d \in [0, D]$ (where $D$ is the maximum delay) on different pre-synaptic addresses $a \in [1, N]$ defining the list of the $N$ dendrites.
+We will denote as $W(a, d)$ these weights.
+When a neuron is activated by a discharge pattern that corresponds to this specific set of delays and addresses, the discharge probability will increase.
+Following the observations of~\citet{Izhikevich06}, let us assume that such precise discharge pattern defines a polychronous group (PG).
+Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
+This allows then to derive a generative model for raster plots (see Figure~\ref{fig:1}).
+
+%Indeed, the probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
+Assuming that the presence of PGs is conditionally independent, this bias can be written as the combination of these factors, whose values are given by the corresponding weights.
+PGs may be activated independently at random times and  we write that $B(b, t)=1$ if $b$ is activated at $t$ (and else $B(b, t)=0$).
+We can thus write the probability bias as %the joint probability given these factors as
+%$%$
+%p(t, a) = \sigma\big(W_0 + \sum_{b, t} B(b, t) \cdot W_b(a, t-d) \big)  
+%$, %$
+%where $\sigma$ is the sigmoid function.
+We will further assume that the weights are balanced (their mean is zero) and that $W_0$ is a bias such that $p_0=\sigma(W_0)$ is the average background firing rate.
+Conveniently, one can write this summation as a one-dimensional temporal convolution operator such that we may simply write $p = \sigma(W_0 + B \ast W )$ where  $p\in [ 0, 1]^{N\times T}$ and $B\in \{0, 1\}^{M\times T}$ is the raster plot corresponding to the temporal activation of the PGs.
+Finally, we obtain the raster plot $A\in \{0, 1\}^{N\times T}$ by drawing spikes using independent Bernoulli trials $A \sim \mathcal{B}(p)$.
+Note that, depending on the shape of the kernels, the generative model can model a Poisson process, generate rhythmic activity or more generally propagating waves.
+This formulation thus defines a simple generative model for raster plots as a combination of independent PGs.
 
@@
 %
+
+% %---------------------------
+% \begin{wrapfigure}{r}{.275\textwidth}
+%     \vspace{-15pt}
+%     % ❯ git add figure_N_PG_time.pdf figure_N_PGs.pdf figure_N_pre.pdf
+%     %\subfloat[]{
+%     \includegraphics[width=\linewidth]{figure_N_PGs.pdf}
+%     %}%
+%     %\subfloat[]{\includegraphics[width=.49\linewidth]{Figures/fig_methods_nat.pdf}}%
+%     \vspace{-25pt}
+%     {
+%     \caption{Accuracy of PG detection as a function of the number $M$ of kernels.
+%     }
+%     \label{fig:2}
+%     }
+%     \vspace{-10pt}
+%     \end{wrapfigure}
+%     %---------------------------
+%     This generative model allows to define an inference model for guessing sources $B$ when observing a raster plot $A$.
+This assumes that we know the PGs as defined by the $W_b$ matrices.
+The underlying metric is the binary cross-entropy, as used in the logistic regression model.
+In particular, if we consider PG kernels with similar decreasing exponential time profile, we can prove that this is similar to %finding the tuning function of L-NL neurons, as used in
+%     the method of~\citet{berens_fast_2012}.
+In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator.
+Using this forward model, it possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator.
+It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.
+To quantify the efficiency of this operation, we generated $M=55$ synthetic PGs as random independent kernels over $128$ presynaptic inputs and $D=127$ possible delays.
+We drew random independent instances of $B$ with a length of $T=1000$ time steps and with on average $2.0$ occurrences each.
+This allowed us to generate raster plots which we use to infer $\hat{B}$.
+We compute the accuracy as the rate of true positive detections and observe on average $\approx 98\%$ correct PGs both for the address and %(exact)
+%     timing.
+
+%     %---------------------------
+%     \begin{wrapfigure}{r}{.275\textwidth}
+%     \vspace{-10pt}
+%     % ❯ git add figure_N_PG_time.pdf figure_N_PGs.pdf figure_N_pre.pdf
+%     %\subfloat[]{
+%     \includegraphics[width=\linewidth]{figure_N_PG_time.pdf}
+%     %}%
+%     %\subfloat[]{\includegraphics[width=.49\linewidth]{Figures/fig_methods_nat.pdf}}%
+%     \vspace{-25pt}
+%     {
+%     \caption{Accuracy of PG detection as a function of the temporal depth $D$ of kernels.
+%     }
+%     \label{fig:3}
+%     }
+%     \vspace{-15pt}
+%     \end{wrapfigure}
+%     %---------------------------
+%     We further extended this result by showing how the accuracy would evolve as a function of the number of simultaneous PGs, while keeping the same frequency of occurrence.
+We show in Figure~\ref{fig:2} that the accuracy of finding the right PG is still above $80\%$ accuracy with more than $1364$ PGs.
+Moreover, we show in Figure~\ref{fig:3} that (with $M=55$ PGs fixed) the accuracy increases notably as the temporal depth $D$ of the PG kernel increased, demonstrating quantitatively the potential of hetero-synaptic delays.
+These results were obtained while assuming that we know $W$.
+However, this is in general not the case, for instance when observing the raster plot of a population of neurons.
+Inspired by the k-means algorithm, it is possible to devise a self-supervised learning algorithm.
+Our preliminary results show that it is possible to retrieve PGs embedded in the data, yet that further analysis is necessary to improve the convergence of the algorithm.
+In particular, it seems promising to use a sparseness constraint in the inference mechanism such as to remove spurious correlations in the inference.
 

### 2022-09-23 Grimaldi-etal-BiolCybernetics.tex
Methods: caption model

@@
 %
-\subsection{Hetero-synaptic delays model as a Logistic Regression with a temporal convolution}
+\subsection{Detecting event-based motifs using spiking neurons with hetero-synaptic delays}% model as a Logistic Regression with a temporal convolution}
 %%%-----------------------------------------------------------------
-Let's first define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-Note that a neuron may contact an afferent neuron with multiple different delays.
-For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
-This defines the full connectivity of the hetero-synaptic model.
-%In general, a spiking neuron is defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree.
+%: Generative model
+%
+\begin{figure}[t!]
+    \centering
+    \includegraphics[width=0.490\linewidth]{figures/figure_1a_k.pdf}
+    \includegraphics[width=0.490\linewidth]{figures/figure_1c.pdf}
+    \includegraphics[width=0.490\linewidth]{figures/figure_1b.pdf}
+    \includegraphics[width=0.490\linewidth]{figures/figure_1a.pdf}
+	    \caption{Detecting event-based motifs using spiking neurons with hetero-synaptic delays. 
+	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs).
+	    {\bf (b)}~The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
+	    {\bf (c)}~We will define a layer of neurons that will detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).
+	    Note that each afferent may be connected with multiple weights at different delays.
+The propagation of the afferent information through these delay may generate at each time step a synchronous pattern on a subset of synapses.
+%These PG patterns were efficiently learned by a logistic regression.
+	    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG pattern at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
+	    {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG.
+	     }
+    \label{fig:model}
+\end{figure}
+
+% 
+The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
+This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
+This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
+
+Indeed, the probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
+Assuming that the presence of PGs is conditionally independent, this bias can be written as the combination of these factors, whose values are given by the corresponding weights.
+PGs may be activated independently at random times and  we write that $B(b, t)=1$ if $b$ is activated at $t$ (and else $B(b, t)=0$).
+We can thus write the probability bias as %the joint probability given these factors as
+$%$
+p(t, a) = \sigma\big(W_0 + \sum_{b, t} B(b, t) \cdot W_b(a, t-d) \big)  
+$, %$
+where $\sigma$ is the sigmoid function.
+We will further assume that the weights are balanced (their mean is zero) and that $W_0$ is a bias such that $p_0=\sigma(W_0)$ is the average background firing rate.
+Conveniently, one can write this summation as a one-dimensional temporal convolution operator such that we may simply write $p = \sigma(W_0 + B \ast W )$ where  $p\in [ 0, 1]^{N\times T}$ and $B\in \{0, 1\}^{M\times T}$ is the raster plot corresponding to the temporal activation of the PGs.
+Finally, we obtain the raster plot $A\in \{0, 1\}^{N\times T}$ by drawing spikes using independent Bernoulli trials $A \sim \mathcal{B}(p)$.
+Note that, depending on the shape of the kernels, the generative model can model a Poisson process, generate rhythmic activity or more generally propagating waves.
+This formulation thus defines a simple generative model for raster plots as a combination of independent PGs.
+
+%: Detection model
+
+In general, a spiking neuron is defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree.
 A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree.
@@
 
-%Indeed, the probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
-Assuming that the presence of PGs is conditionally independent, this bias can be written as the combination of these factors, whose values are given by the corresponding weights.
-PGs may be activated independently at random times and  we write that $B(b, t)=1$ if $b$ is activated at $t$ (and else $B(b, t)=0$).
-We can thus write the probability bias as %the joint probability given these factors as
-%$%$
-%p(t, a) = \sigma\big(W_0 + \sum_{b, t} B(b, t) \cdot W_b(a, t-d) \big)  
-%$, %$
-%where $\sigma$ is the sigmoid function.
-We will further assume that the weights are balanced (their mean is zero) and that $W_0$ is a bias such that $p_0=\sigma(W_0)$ is the average background firing rate.
-Conveniently, one can write this summation as a one-dimensional temporal convolution operator such that we may simply write $p = \sigma(W_0 + B \ast W )$ where  $p\in [ 0, 1]^{N\times T}$ and $B\in \{0, 1\}^{M\times T}$ is the raster plot corresponding to the temporal activation of the PGs.
-Finally, we obtain the raster plot $A\in \{0, 1\}^{N\times T}$ by drawing spikes using independent Bernoulli trials $A \sim \mathcal{B}(p)$.
-Note that, depending on the shape of the kernels, the generative model can model a Poisson process, generate rhythmic activity or more generally propagating waves.
-This formulation thus defines a simple generative model for raster plots as a combination of independent PGs.
+Let's first define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+Note that a neuron may contact an afferent neuron with multiple different delays.
+For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
+This defines the full connectivity of the hetero-synaptic model.
+% We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then not only described by synaptic weights connecting to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+For each neuron $\postsynaddr$, we define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
+Note that a neuron may contact an afferent neuron with multiple, different delays.
+
+\note{say that it is effortless in biology but difficult on conventional computers}
 
@@
 $.
-\note{finish by saying that it is effortless in biology but difficult on conventional computers}
-%
-\begin{figure}[t!]
-    \centering
-    \includegraphics[width=0.90\linewidth]{figures/figure_1a.pdf}
-    \includegraphics[width=0.90\linewidth]{figures/figure_1b.pdf}
-    \includegraphics[width=0.90\linewidth]{figures/figure_1c.pdf}
-	    \caption{The hetero-synaptic delays model applied on synthetic data.
-{\bf (a)}~The afferent information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs).
-We highlight them by different colors, an information hidden to a detection model.
-{\bf (b)}~The model is defined as an assembly of neurons (here for $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
-Note that each afferent may be connected with multiple weights at different delays.
-The propagation of the afferent information through these delay may generate at each time step a synchronous pattern on a subset of synapses.
-%These PG patterns were efficiently learned by a logistic regression.
-	    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG pattern at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
-}
-    \label{fig:model}
-\end{figure}
 
 %
-The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
-This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
-This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
-
-We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then not only described by synaptic weights connecting to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-For each neuron $\postsynaddr$, we define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
-Note that a neuron may contact an afferent neuron with multiple, different delays.
-
-The corresponding input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible patterns.
-For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
-$
-\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
-$.
-%
-The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\cite{Grimaldi22pami}, in analogy to a spiking Winner-Take-All network~\cite{nessler2013bayesian}.
+%The corresponding input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible patterns.
+%For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
+%$
+%\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
+%$.
+%%
+%The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\cite{Grimaldi22pami}, in analogy to a spiking Winner-Take-All network~\cite{nessler2013bayesian}.
 It transforms this list of weights into a probability with the following formula:
-$
-Pr(k=\postsynaddr \; | \; \timev) =
-\frac 1 Z
-{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
-$ 
-where $\mathcal{C}^\postsynaddr(\timev) = \sum
-\activeweights^\postsynaddr(t)
-$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
-In particular, some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
+%$
+%Pr(k=\postsynaddr \; | \; \timev) =
+%\frac 1 Z
+%{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
+%$ 
+%where $\mathcal{C}^\postsynaddr(\timev) = \sum
+%\activeweights^\postsynaddr(t)
+%$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
+%In particular, some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
 %
@@
 %
-This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
+This shows that $\mathcal{C}^\postsynaddr$ is a discrete temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
 This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
@@
 This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different patterns in the event stream.
-%
-In particular, we may extend the convolution to a 3D convolution 
-such that the representation would also benefit from spatial invariance.
-The use of spatiotemporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\cite{ghosh2022conv3D}.
-For that, we design 3D kernels of shape  $(\Kx, \Ky, \Ktime) = (15, 15, 8)$, respectively representing the two spatial dimensions and the range of delays.
-Computations are performed on spatio-temporal windows, defined by the kernels, sliding around the events, that is, the center of the spatio-temporal window around the current event $\epsilon_\arank$.
-%
-Finally, the output of the MLR model results in an event with the highest probability class, keeping the same timing as the event as input.
-%
-The loss function of the MLR model is the binary cross entropy on the output of the classification layer.
-Simulations are performed thanks to the PyTorch library using gradient descent with Adam ($2^{12}$ epochs and a learning rate of $10^{-5}$).
-%
 
@@
 %%%-----------------------------------------------------------------
+
+%
+In particular, we may extend the convolution to a 3D convolution 
+such that the representation would also benefit from spatial invariance.
+The use of spatiotemporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\cite{ghosh2022conv3D}.
+For that, we design 3D kernels of shape  $(\Kx, \Ky, \Ktime) = (15, 15, 8)$, respectively representing the two spatial dimensions and the range of delays.
+Computations are performed on spatio-temporal windows, defined by the kernels, sliding around the events, that is, the center of the spatio-temporal window around the current event $\epsilon_\arank$.
+%
+Finally, the output of the MLR model results in an event with the highest probability class, keeping the same timing as the event as input.
+%
+The loss function of the MLR model is the binary cross entropy on the output of the classification layer.
+Simulations are performed thanks to the PyTorch library using gradient descent with Adam ($2^{12}$ epochs and a learning rate of $10^{-5}$).
+%
+
 In particular, we may extend the convolution to a 3D convolution such that the representation would also benefit from spatial invariance which is inherent to the task's construction.

### 2022-09-23 Grimaldi-etal-BiolCybernetics.tex
Methods: 1b

@@
 	    Note that each afferent may be connected with multiple weights at different delays.
-The propagation of the afferent information through these delay may generate at each time step a synchronous pattern on a subset of synapses.
-%These PG patterns were efficiently learned by a logistic regression.
-	    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG pattern at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
+The propagation of the afferent information through these delay may generate at each time step a synchronous motif on a subset of synapses.
+%These PG motifs were efficiently learned by a logistic regression.
+	    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
 	    {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG.
@@
 We will denote as $W(a, d)$ these weights.
-When a neuron is activated by a discharge pattern that corresponds to this specific set of delays and addresses, the discharge probability will increase.
-Following the observations of~\citet{Izhikevich06}, let us assume that such precise discharge pattern defines a polychronous group (PG).
+When a neuron is activated by a discharge motif that corresponds to this specific set of delays and addresses, the discharge probability will increase.
+Following the observations of~\citet{Izhikevich06}, let us assume that such precise discharge motif defines a polychronous group (PG).
 Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
@@
 %
-%The corresponding input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible patterns.
+%The corresponding input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible motifs.
 %For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
@@
 From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.
-As such, we may first write any event-based input as the Boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
-In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the nonzero weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
+As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
+In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
 Using this dense representation, the counting defined above becomes:
-$$
+\begin{equation*}
 \mathcal{C}^\postsynaddr(a,t)
 = \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
-$$
+\end{equation*}
 %
-This shows that $\mathcal{C}^\postsynaddr$ is a discrete temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
+This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
 This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
@@
 The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model.
-From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representations.
-As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
-In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
-Using this dense representation, the counting defined above becomes:
-\begin{equation*}
-\mathcal{C}^\postsynaddr(a,t)
-= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
-\end{equation*}
-%
-This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
-This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different patterns in the event stream.
 

### 2022-09-23 Grimaldi-etal-BiolCybernetics.tex
Methods: 1c

@@
 %%%-----------------------------------------------------------------
-%: Generative model
 %
@@
 \end{figure}
-
 % 
-The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
+%: Generative model
+A generic raster plot, such as the sensory signal representing the output of an event-based camera, consists a discrete stream of events (see Figure~\ref{fig:model}-(a)).
+This can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
 This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
-This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
-
-Indeed, the probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
-Assuming that the presence of PGs is conditionally independent, this bias can be written as the combination of these factors, whose values are given by the corresponding weights.
-PGs may be activated independently at random times and  we write that $B(b, t)=1$ if $b$ is activated at $t$ (and else $B(b, t)=0$).
-We can thus write the probability bias as %the joint probability given these factors as
-$%$
-p(t, a) = \sigma\big(W_0 + \sum_{b, t} B(b, t) \cdot W_b(a, t-d) \big)  
-$, %$
-where $\sigma$ is the sigmoid function.
-We will further assume that the weights are balanced (their mean is zero) and that $W_0$ is a bias such that $p_0=\sigma(W_0)$ is the average background firing rate.
-Conveniently, one can write this summation as a one-dimensional temporal convolution operator such that we may simply write $p = \sigma(W_0 + B \ast W )$ where  $p\in [ 0, 1]^{N\times T}$ and $B\in \{0, 1\}^{M\times T}$ is the raster plot corresponding to the temporal activation of the PGs.
-Finally, we obtain the raster plot $A\in \{0, 1\}^{N\times T}$ by drawing spikes using independent Bernoulli trials $A \sim \mathcal{B}(p)$.
-Note that, depending on the shape of the kernels, the generative model can model a Poisson process, generate rhythmic activity or more generally propagating waves.
-This formulation thus defines a simple generative model for raster plots as a combination of independent PGs.
-
-%: Detection model
-
-In general, a spiking neuron is defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree.
-A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree.
-A more detailed examination of these synaptic contacts reveals the added importance of hetero-synaptic delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{Izhikevich06}.
-Thus, we can parameterize each neuron by the set of tuples defining the weight and delay of each synaptic contact.
-By discretization of time (with here an arbitrary unitary time unit), we can also define the dendrite as a matrix giving the weight corresponding to the different delays $d \in [0, D]$ (where $D$ is the maximum delay) on different pre-synaptic addresses $a \in [1, N]$ defining the list of the $N$ dendrites.
-We will denote as $W(a, d)$ these weights.
-When a neuron is activated by a discharge motif that corresponds to this specific set of delays and addresses, the discharge probability will increase.
-Following the observations of~\citet{Izhikevich06}, let us assume that such precise discharge motif defines a polychronous group (PG).
-Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
-This allows then to derive a generative model for raster plots (see Figure~\ref{fig:1}).
-
-Let's first define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-Note that a neuron may contact an afferent neuron with multiple different delays.
-For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
-This defines the full connectivity of the hetero-synaptic model.
-% We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then not only described by synaptic weights connecting to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-For each neuron $\postsynaddr$, we define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
-Note that a neuron may contact an afferent neuron with multiple, different delays.
-
-\note{say that it is effortless in biology but difficult on conventional computers}
+This defines a presynaptic address space $\presynaddrspace$ which in the particular case of event-based cameras is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for ON and OFF polarities).
 
@@
 
+%: Detection model
+
+In general, a spiking neuron is defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree.
+A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree.
+A more detailed examination of these synaptic contacts reveals the added importance of hetero-synaptic delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{Izhikevich06}.
+Thus, we can parameterize each neuron by the set of tuples defining the weight and delay of each synaptic contact.
+By discretization of time (with here an arbitrary unitary time unit), we can also define the dendrite as a matrix giving the weight corresponding to the different delays $d \in [0, D]$ (where $D$ is the maximum delay) on different pre-synaptic addresses $a \in [1, N]$ defining the list of the $N$ dendrites.
+We will denote as $W(a, d)$ these weights.
+When a neuron is activated by a discharge motif that corresponds to this specific set of delays and addresses, the discharge probability will increase.
+Following the observations of~\citet{Izhikevich06}, let us assume that such precise discharge motif defines a polychronous group (PG).
+Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
+This allows then to derive a generative model for raster plots (see Figure~\ref{fig:1}).
+
+Let's first define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+Note that a neuron may contact an afferent neuron with multiple different delays.
+For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
+This defines the full connectivity of the hetero-synaptic model.
+% We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then not only described by synaptic weights connecting to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+For each neuron $\postsynaddr$, we define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
+Note that a neuron may contact an afferent neuron with multiple, different delays.
+
+Indeed, the probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
+Assuming that the presence of PGs is conditionally independent, this bias can be written as the combination of these factors, whose values are given by the corresponding weights.
+PGs may be activated independently at random times and  we write that $B(b, t)=1$ if $b$ is activated at $t$ (and else $B(b, t)=0$).
+We can thus write the probability bias as %the joint probability given these factors as
+$%$
+p(t, a) = \sigma\big(W_0 + \sum_{b, t} B(b, t) \cdot W_b(a, t-d) \big)  
+$, %$
+where $\sigma$ is the sigmoid function.
+We will further assume that the weights are balanced (their mean is zero) and that $W_0$ is a bias such that $p_0=\sigma(W_0)$ is the average background firing rate.
+Conveniently, one can write this summation as a one-dimensional temporal convolution operator such that we may simply write $p = \sigma(W_0 + B \ast W )$ where  $p\in [ 0, 1]^{N\times T}$ and $B\in \{0, 1\}^{M\times T}$ is the raster plot corresponding to the temporal activation of the PGs.
+Finally, we obtain the raster plot $A\in \{0, 1\}^{N\times T}$ by drawing spikes using independent Bernoulli trials $A \sim \mathcal{B}(p)$.
+Note that, depending on the shape of the kernels, the generative model can model a Poisson process, generate rhythmic activity or more generally propagating waves.
+This formulation thus defines a simple generative model for raster plots as a combination of independent PGs.
+
+\note{say that it is effortless in biology but difficult on conventional computers}
+
 \note{ TODO: show why this formula implements a polychrony detector }
@@
 The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model. 
+%
+\begin{figure}[t!]
+    \centering
+    \includegraphics[width=0.490\linewidth]{figures/figure_N_PGs.pdf}
+    \includegraphics[width=0.490\linewidth]{figures/figure_N_PG_time.pdf}
+	    \caption{Detecting event-based motifs using spiking neurons with hetero-synaptic delays. 
+	    {\bf (a)}~Accuracy of PG detection as a function of the number $M$ of kernels.
+	    {\bf (b)}~Accuracy of PG detection as a function of the temporal depth $D$ of kernels.
+	    	     }
+    \label{fig:model_results}
+\end{figure}
+%
 
-% %---------------------------
-% \begin{wrapfigure}{r}{.275\textwidth}
-%     \vspace{-15pt}
-%     % ❯ git add figure_N_PG_time.pdf figure_N_PGs.pdf figure_N_pre.pdf
-%     %\subfloat[]{
-%     \includegraphics[width=\linewidth]{figure_N_PGs.pdf}
-%     %}%
-%     %\subfloat[]{\includegraphics[width=.49\linewidth]{Figures/fig_methods_nat.pdf}}%
-%     \vspace{-25pt}
-%     {
-%     \caption{Accuracy of PG detection as a function of the number $M$ of kernels.
-%     }
-%     \label{fig:2}
-%     }
-%     \vspace{-10pt}
-%     \end{wrapfigure}
-%     %---------------------------
-%     This generative model allows to define an inference model for guessing sources $B$ when observing a raster plot $A$.
+This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$.
 This assumes that we know the PGs as defined by the $W_b$ matrices.
 The underlying metric is the binary cross-entropy, as used in the logistic regression model.
-In particular, if we consider PG kernels with similar decreasing exponential time profile, we can prove that this is similar to %finding the tuning function of L-NL neurons, as used in
-%     the method of~\citet{berens_fast_2012}.
+In particular, if we consider PG kernels with similar decreasing exponential time profile, we can prove that this is similar to the method of~\citet{berens_fast_2012}.
 In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator.
@@
 This allowed us to generate raster plots which we use to infer $\hat{B}$.
-We compute the accuracy as the rate of true positive detections and observe on average $\approx 98\%$ correct PGs both for the address and %(exact)
-%     timing.
-
-%     %---------------------------
-%     \begin{wrapfigure}{r}{.275\textwidth}
-%     \vspace{-10pt}
-%     % ❯ git add figure_N_PG_time.pdf figure_N_PGs.pdf figure_N_pre.pdf
-%     %\subfloat[]{
-%     \includegraphics[width=\linewidth]{figure_N_PG_time.pdf}
-%     %}%
-%     %\subfloat[]{\includegraphics[width=.49\linewidth]{Figures/fig_methods_nat.pdf}}%
-%     \vspace{-25pt}
-%     {
-%     \caption{Accuracy of PG detection as a function of the temporal depth $D$ of kernels.
-%     }
-%     \label{fig:3}
-%     }
-%     \vspace{-15pt}
-%     \end{wrapfigure}
-%     %---------------------------
-%     We further extended this result by showing how the accuracy would evolve as a function of the number of simultaneous PGs, while keeping the same frequency of occurrence.
-We show in Figure~\ref{fig:2} that the accuracy of finding the right PG is still above $80\%$ accuracy with more than $1364$ PGs.
-Moreover, we show in Figure~\ref{fig:3} that (with $M=55$ PGs fixed) the accuracy increases notably as the temporal depth $D$ of the PG kernel increased, demonstrating quantitatively the potential of hetero-synaptic delays.
+We compute the accuracy as the rate of true positive detections and observe on average $\approx 98\%$ correct PGs both for the address and (exact) timing.
+We further extended this result by showing how the accuracy would evolve as a function of the number of simultaneous PGs, while keeping the same frequency of occurrence.
+We show in Figure~\ref{fig:model_results}-(a) that the accuracy of finding the right PG is still above $80\%$ accuracy with more than $1364$ PGs.
+Moreover, we show in Figure~\ref{fig:model_results}-(b) that (with $M=55$ PGs fixed) the accuracy increases notably as the temporal depth $D$ of the PG kernel increased, demonstrating quantitatively the potential of hetero-synaptic delays.
 These results were obtained while assuming that we know $W$.

### 2022-09-23 Grimaldi-etal-BiolCybernetics.tex
Methods: 1c

@@
 This allowed us to generate raster plots which we use to infer $\hat{B}$.
-We compute the accuracy as the rate of true positive detections and observe on average $\approx 98\%$ correct PGs both for the address and (exact) timing.
+We compute the accuracy as the rate of true positive detections (both for inferring the address and its exact timing) and observe on average $\approx 98\%$ correct detections.
 We further extended this result by showing how the accuracy would evolve as a function of the number of simultaneous PGs, while keeping the same frequency of occurrence.

### 2022-09-23 Grimaldi-etal-BiolCybernetics.tex
Methods: 1c

@@
 We further extended this result by showing how the accuracy would evolve as a function of the number of simultaneous PGs, while keeping the same frequency of occurrence.
-We show in Figure~\ref{fig:model_results}-(a) that the accuracy of finding the right PG is still above $80\%$ accuracy with more than $1364$ PGs.
-Moreover, we show in Figure~\ref{fig:model_results}-(b) that (with $M=55$ PGs fixed) the accuracy increases notably as the temporal depth $D$ of the PG kernel increased, demonstrating quantitatively the potential of hetero-synaptic delays.
+We show in Figure~\ref{fig:model_results}~(a) that the accuracy of finding the right PG is still above $80\%$ accuracy with more than $1364$ overlapping PGs.
+Moreover, we show in Figure~\ref{fig:model_results}~(b) that (with $M=55$ PGs fixed) the accuracy increases notably as the temporal depth $D$ of the PG kernel increased, demonstrating quantitatively the potential of hetero-synaptic delays.
 These results were obtained while assuming that we know $W$.
 However, this is in general not the case, for instance when observing the raster plot of a population of neurons.
-Inspired by the k-means algorithm, it is possible to devise a self-supervised learning algorithm.
+In the following, we will define a generic visual task and determine a learning algorithm to solve it.
+%Inspired by the k-means algorithm, it is possible to devise a self-supervised learning algorithm.
 Our preliminary results show that it is possible to retrieve PGs embedded in the data, yet that further analysis is necessary to improve the convergence of the algorithm.

### 2022-09-23 Grimaldi-etal-BiolCybernetics.tex
Methods: 1a

@@
 % 
-%: Generative model
-A generic raster plot, such as the sensory signal representing the output of an event-based camera, consists a discrete stream of events (see Figure~\ref{fig:model}-(a)).
+\subsection{A generative model for raster plots}
+%
+For any generic raster plot, such as the sensory signal representing the output of an event-based camera, consists a stream of events (see Figure~\ref{fig:model}-(a)).
 This can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
-This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
+That event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
 This defines a presynaptic address space $\presynaddrspace$ which in the particular case of event-based cameras is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for ON and OFF polarities).
-
-We can now define an event-based input to the model as a set on incoming spikes reaching the presynaptic neurons at specified times, like the sensory signal representing the output of an event-based camera.
-It thus forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
-Each event has a time of occurrence $\timev_\arank$.
+In general, a spiking neuron is defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree.
+A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree.
+A more detailed examination of these synaptic contacts reveals the added importance of hetero-synaptic delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{Izhikevich06}.
+Thus, we can parameterize each neuron by the set of tuples defining the weight and delay of each synaptic contact.
 The corresponding set of input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible motifs.
 For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
-$
+\begin{equation*}
 \activeweights^\postsynaddr (t) = 
@@
 \}
-$.
+\end{equation*}
 
-%
-%The corresponding input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible motifs.
-%For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
-%$
-%\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}
-%$.
-%%
+When a neuron is activated by a discharge motif that corresponds to this specific set of delays and addresses, the discharge probability will increase.
+Following the observations of~\citet{Izhikevich06}, let us assume that such precise discharge motif defines a polychronous group (PG).
+Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
+This allows then to derive a generative model for raster plots (see Figure~\ref{fig:1}).
+
+%: Detection model
+Let's first define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+Note that a neuron may contact an afferent neuron with multiple different delays.
+For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
+This defines the full connectivity of the hetero-synaptic model.
+
+ By discretization of time (with here an arbitrary unitary time unit), we can also define the dendrite as a matrix giving the weight corresponding to the different delays $d \in [0, D]$ (where $D$ is the maximum delay) on different pre-synaptic addresses $a \in [1, N]$ defining the list of the $N$ dendrites.
+We will denote as $W(a, d)$ these weights.
+
+Indeed, the probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
+Assuming that the presence of PGs is conditionally independent, this bias can be written as the combination of these factors, whose values are given by the corresponding weights.
+PGs may be activated independently at random times and  we write that $B(b, t)=1$ if $b$ is activated at $t$ (and else $B(b, t)=0$).
+We can thus write the probability bias as the accumulated evidence given these factors as
+\begin{equation*}
+p(t, a) = \sigma\big(W_0 + \sum_{b, t} B(b, t) \cdot W_b(a, t-d) \big)  
+\end{equation*}
+where $\sigma$ is the sigmoid function.
+We will further assume that the weights are balanced (their mean is zero) and that $W_0$ is a bias such that $p_0=\sigma(W_0)$ is the average background firing rate.
+Conveniently, one can write this summation as a one-dimensional temporal convolution operator such that we may simply write $p = \sigma(W_0 + B \ast W )$ where  $p\in [ 0, 1]^{N\times T}$ and $B\in \{0, 1\}^{M\times T}$ is the raster plot corresponding to the temporal activation of the PGs.
+Finally, we obtain the raster plot $A\in \{0, 1\}^{N\times T}$ by drawing spikes using independent Bernoulli trials $A \sim \mathcal{B}(p)$.
+Note that, depending on the shape of the kernels, the generative model can model a Poisson process, generate rhythmic activity or more generally propagating waves.
+This formulation thus defines a simple generative model for raster plots as a combination of independent PGs.
+
 %The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\cite{Grimaldi22pami}, in analogy to a spiking Winner-Take-All network~\cite{nessler2013bayesian}.
@@
 %
-
-%: Detection model
-
-In general, a spiking neuron is defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree.
-A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree.
-A more detailed examination of these synaptic contacts reveals the added importance of hetero-synaptic delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{Izhikevich06}.
-Thus, we can parameterize each neuron by the set of tuples defining the weight and delay of each synaptic contact.
-By discretization of time (with here an arbitrary unitary time unit), we can also define the dendrite as a matrix giving the weight corresponding to the different delays $d \in [0, D]$ (where $D$ is the maximum delay) on different pre-synaptic addresses $a \in [1, N]$ defining the list of the $N$ dendrites.
-We will denote as $W(a, d)$ these weights.
-When a neuron is activated by a discharge motif that corresponds to this specific set of delays and addresses, the discharge probability will increase.
-Following the observations of~\citet{Izhikevich06}, let us assume that such precise discharge motif defines a polychronous group (PG).
-Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
-This allows then to derive a generative model for raster plots (see Figure~\ref{fig:1}).
-
-Let's first define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-Note that a neuron may contact an afferent neuron with multiple different delays.
-For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
-This defines the full connectivity of the hetero-synaptic model.
-% We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
 In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
@@
 Note that a neuron may contact an afferent neuron with multiple, different delays.
-
-Indeed, the probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
-Assuming that the presence of PGs is conditionally independent, this bias can be written as the combination of these factors, whose values are given by the corresponding weights.
-PGs may be activated independently at random times and  we write that $B(b, t)=1$ if $b$ is activated at $t$ (and else $B(b, t)=0$).
-We can thus write the probability bias as %the joint probability given these factors as
-$%$
-p(t, a) = \sigma\big(W_0 + \sum_{b, t} B(b, t) \cdot W_b(a, t-d) \big)  
-$, %$
-where $\sigma$ is the sigmoid function.
-We will further assume that the weights are balanced (their mean is zero) and that $W_0$ is a bias such that $p_0=\sigma(W_0)$ is the average background firing rate.
-Conveniently, one can write this summation as a one-dimensional temporal convolution operator such that we may simply write $p = \sigma(W_0 + B \ast W )$ where  $p\in [ 0, 1]^{N\times T}$ and $B\in \{0, 1\}^{M\times T}$ is the raster plot corresponding to the temporal activation of the PGs.
-Finally, we obtain the raster plot $A\in \{0, 1\}^{N\times T}$ by drawing spikes using independent Bernoulli trials $A \sim \mathcal{B}(p)$.
-Note that, depending on the shape of the kernels, the generative model can model a Poisson process, generate rhythmic activity or more generally propagating waves.
-This formulation thus defines a simple generative model for raster plots as a combination of independent PGs.
 

### 2022-09-23 Grimaldi-etal-BiolCybernetics.tex
Methods: 1a

@@
 % 
-\subsection{A generative model for raster plots}
+\subsubsection{A generative model for raster plots}
 %
-For any generic raster plot, such as the sensory signal representing the output of an event-based camera, consists a stream of events (see Figure~\ref{fig:model}-(a)).
+As in a neurobiological recording, or in the sensory signal representing the output of an event-based camera,  any generic raster plot consists a stream of spikes (see Figure~\ref{fig:model}-(a)).
 This can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
-That event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
-This defines a presynaptic address space $\presynaddrspace$ which in the particular case of event-based cameras is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for ON and OFF polarities).
-In general, a spiking neuron is defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree.
-A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree.
-A more detailed examination of these synaptic contacts reveals the added importance of hetero-synaptic delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{Izhikevich06}.
-Thus, we can parameterize each neuron by the set of tuples defining the weight and delay of each synaptic contact.
-The corresponding set of input presynaptic spikes $\event$ will be integrated by this synaptic set and notably by the respective delays, which will multiplex in time all possible motifs.
-For each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
+That event has a time of occurrence $\timev_\arank$  and an associated address (which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
+This defines an address space $\presynaddrspace$ which consists the identified set of neurons in the neurobiological recording, or is denoted for event-based cameras by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for ON and OFF polarities).
+
+A spiking neuron is defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree.
+A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree, the so-called weight matrix.
+A more detailed description of these synaptic contacts reveals the added importance of hetero-synaptic delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{Izhikevich06}.
+In that hetero-synaptic description, we can parameterize each neuron by the set of tuples defining the weight and delay of each synaptic contact.
+As a consequence, a set of input presynaptic spikes $\event$ will be processed by the dendrite defined by this synaptic set and notably by the respective delays, which will multiplex in time all events.
+
+Let's first define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+Note that a neuron may contact an afferent neuron with multiple different delays.
+For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
+This defines the full connectivity of the hetero-synaptic model.
+
+Formally, for each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
 \begin{equation*}
@@
 
+This raster plot is generated either by a event generation process or by other spiking neurons.
+
 When a neuron is activated by a discharge motif that corresponds to this specific set of delays and addresses, the discharge probability will increase.
@@
 Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
-This allows then to derive a generative model for raster plots (see Figure~\ref{fig:1}).
+This allows then to derive a generative model for raster plots (see Figure~\ref{fig:model}-(b)).
 
+\subsubsection{Detecting spiking motifs}
 %: Detection model
-Let's first define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-Note that a neuron may contact an afferent neuron with multiple different delays.
-For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
-This defines the full connectivity of the hetero-synaptic model.
 
@@
 We will further assume that the weights are balanced (their mean is zero) and that $W_0$ is a bias such that $p_0=\sigma(W_0)$ is the average background firing rate.
-Conveniently, one can write this summation as a one-dimensional temporal convolution operator such that we may simply write $p = \sigma(W_0 + B \ast W )$ where  $p\in [ 0, 1]^{N\times T}$ and $B\in \{0, 1\}^{M\times T}$ is the raster plot corresponding to the temporal activation of the PGs.
+Conveniently, one can write this summation as a one-dimensional temporal convolution operator such that we may simply write
+\begin{equation*}
+p = \sigma(W_0 + B \ast W )
+\end{equation*}
+where  $p\in [ 0, 1]^{N\times T}$ and $B\in \{0, 1\}^{M\times T}$ is the raster plot corresponding to the temporal activation of the PGs.
 Finally, we obtain the raster plot $A\in \{0, 1\}^{N\times T}$ by drawing spikes using independent Bernoulli trials $A \sim \mathcal{B}(p)$.
@@
 
+\subsubsection{Implementation using spiking neurons with hetero-synaptic delays}
+%
 %The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\cite{Grimaldi22pami}, in analogy to a spiking Winner-Take-All network~\cite{nessler2013bayesian}.
@@
 The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model.
+
+\subsubsection{Performance on synthetic data}
 %
@@
 It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.
+
 To quantify the efficiency of this operation, we generated $M=55$ synthetic PGs as random independent kernels over $128$ presynaptic inputs and $D=127$ possible delays.

### 2022-09-23 Grimaldi-etal-BiolCybernetics.tex
Methods: 1a

@@
 %: Detection model
+\note{ TODO: show why this formula implements a polychrony detector }
+
+From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.
+As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
+In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
+Using this dense representation, the counting defined above becomes:
+\begin{equation*}
+\mathcal{C}^\postsynaddr(a,t)
+= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
+\end{equation*}
+%
+This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
+This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
+%
 
@@
 %
-%The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\cite{Grimaldi22pami}, in analogy to a spiking Winner-Take-All network~\cite{nessler2013bayesian}.
+The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\cite{Grimaldi22pami}, in analogy to a spiking Winner-Take-All network~\cite{nessler2013bayesian}.
 It transforms this list of weights into a probability with the following formula:
-%$
-%Pr(k=\postsynaddr \; | \; \timev) =
-%\frac 1 Z
-%{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
-%$ 
-%where $\mathcal{C}^\postsynaddr(\timev) = \sum
-%\activeweights^\postsynaddr(t)
-%$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
-%In particular, some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
-%
-We may now define a layer of neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then not only described by synaptic weights connecting to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-For each neuron $\postsynaddr$, we define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
-Note that a neuron may contact an afferent neuron with multiple, different delays.
+$
+Pr(k=\postsynaddr \; | \; \timev) =
+\frac 1 Z
+{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
+$ 
+where $\mathcal{C}^\postsynaddr(\timev) = \sum
+\activeweights^\postsynaddr(t)
+$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
+In particular, some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
 
 \note{say that it is effortless in biology but difficult on conventional computers}
-
-\note{ TODO: show why this formula implements a polychrony detector }
-From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.
-As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
-In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
-Using this dense representation, the counting defined above becomes:
-\begin{equation*}
-\mathcal{C}^\postsynaddr(a,t)
-= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
-\end{equation*}
-%
-This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
-This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
-%
 

### 2022-09-23 Grimaldi-etal-BiolCybernetics.tex
Methods: 2.1.3

@@
 %
-The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\cite{Grimaldi22pami}, in analogy to a spiking Winner-Take-All network~\cite{nessler2013bayesian}.
-It transforms this list of weights into a probability with the following formula:
-$
-Pr(k=\postsynaddr \; | \; \timev) =
-\frac 1 Z
-{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
-$ 
-where $\mathcal{C}^\postsynaddr(\timev) = \sum
-\activeweights^\postsynaddr(t)
-$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
-In particular, some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
-
-\note{say that it is effortless in biology but difficult on conventional computers}
-
 The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
@@
 %
+
+\note{say that it is effortless in biology but difficult on conventional computers}
+
 %In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.

### 2022-09-23 Grimaldi-etal-BiolCybernetics.tex
Methods: 2.1.3

@@
 
-A spiking neuron is defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree.
+This raster plot is generated by a spiking neuron which is defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree, or on the sensory processes (such as the pixel on the sensor of a camera, or the photoreceptors in the retina).
 A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree, the so-called weight matrix.
-A more detailed description of these synaptic contacts reveals the added importance of hetero-synaptic delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{Izhikevich06}.
+A more detailed description of these synaptic contacts reveals the added importance of hetero-synaptic delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}.
 In that hetero-synaptic description, we can parameterize each neuron by the set of tuples defining the weight and delay of each synaptic contact.
@@
 When a neuron is activated by a discharge motif that corresponds to this specific set of delays and addresses, the discharge probability will increase.
-Following the observations of~\citet{Izhikevich06}, let us assume that such precise discharge motif defines a polychronous group (PG).
+Following the observations of~\citet{izhikevich_polychronization_2006}, let us assume that such precise discharge motif defines a polychronous group (PG).
 Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
@@
 \note{say that it is effortless in biology but difficult on conventional computers}
-
-%In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
 The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model.
 %
-%
-In our MLR model with $\Nclass=\Nspeed$ classes, 
-a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
-The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model.
-
 \subsubsection{Performance on synthetic data}
@@
 % 
-
 This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$.

### 2022-09-26 Grimaldi-etal-BiolCybernetics.tex
import from 2021-11-20_COSYNE_abstract.tex

@@
 
+\note{PGs coccur becaus they share a common cause like chirp ob bird interaural delays Goodman}
 From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.
@@
 While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside of a homeostatic state (through a death or explosion of activity).
+\note{ dire que la prédiction permet de faire de la prediction perrinet_prediction_2012 et que ça explique les résultats de leBec}
 

### 2022-09-26 Grimaldi-etal-BiolCybernetics.tex
Methods: 2.1.3

--- 2022-09-23 Grimaldi-etal-BiolCybernetics.tex
@@
 %
-As in a neurobiological recording, or in the sensory signal representing the output of an event-based camera,  any generic raster plot consists a stream of spikes (see Figure~\ref{fig:model}-(a)).
-This can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$.
-That event has a time of occurrence $\timev_\arank$  and an associated address (which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
-This defines an address space $\presynaddrspace$ which consists the identified set of neurons in the neurobiological recording, or is denoted for event-based cameras by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for ON and OFF polarities).
+As in a neurobiological recording or in the sensory signal representing the output of an event-based camera,  any generic raster plot consists a stream of \emph{spikes} (see Figure~\ref{fig:model}-(a)).
+This can be formalized as an ordered set $\event$ of addresses and timestamps $\{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event.
+That event has a time of occurrence $\timev_\arank$  and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
+This defines an address space $\presynaddrspace$ which consists of the set of possible addresses.
+This can be the identified set of neurons in a neurobiological recording.
+For event-based cameras, it is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for ON and OFF polarities).
 
-This raster plot is generated by a spiking neuron which is defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree, or on the sensory processes (such as the pixel on the sensor of a camera, or the photoreceptors in the retina).
+This raster plot is generated by spiking neurons which are defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree, or on the sensory processes (such as the pixel on the sensor of a camera, or the photoreceptors in the retina).
 A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree, the so-called weight matrix.
 A more detailed description of these synaptic contacts reveals the added importance of hetero-synaptic delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}.
-In that hetero-synaptic description, we can parameterize each neuron by the set of tuples defining the weight and delay of each synaptic contact.
+In that hetero-synaptic description, we can parameterize each neuron by the set of tuples defining both the weight and the delay of each synaptic contact.
 As a consequence, a set of input presynaptic spikes $\event$ will be processed by the dendrite defined by this synaptic set and notably by the respective delays, which will multiplex in time all events.
 
-Let's first define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+Formally, let's define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
 In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
@@
 This defines the full connectivity of the hetero-synaptic model.
+As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic adress space is transformed into a new event stream $\event_\postsynaddrspace$ which is defined by $\{ (\presynaddr_\arank =
+ \presynaddr_\ranksyn^\postsynaddr, \timev=\timev_\arank + 
+\synapticdelay_\ranksyn^\postsynaddr \}_{ \arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]} )
+\}$. \note{formaliser le multiplexage}
+When post-synaptic neurons are activated on their soma by this discharge motif the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.
 
-Formally, for each time $\timev$ the integration of $\event$ is defined by a list of weights $\activeweights^\postsynaddr$ linked to the synapses that match a precise spatio-temporal motif as input:
-\begin{equation*}
-\activeweights^\postsynaddr (t) = 
-\{ \synapticweight_\ranksyn^\postsynaddr  \vert 
-\presynaddr_\arank = 
- \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + 
-\synapticdelay_\ranksyn^\postsynaddr \}_{ \arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]} 
-\}
-\end{equation*}
+Taking the argument the other way around, one may deduce a generative model for generic raster plots.
+Indeed, any spike  in the presynaptic address space is generated by sensory neurons (for instance in the retina) or by afferent neurons.
+These are connected to the spiking cell by a set of weights and delays, and by assuming that these connections are sparsely distributes, one may deduce that this firing was caused by a specific temporal motif.
+, this will generate the propagation in
 
+\note{PGs coccur becaus they share a common cause like chirp ob bird interaural delays Goodman}
 This raster plot is generated either by a event generation process or by other spiking neurons.
-
-When a neuron is activated by a discharge motif that corresponds to this specific set of delays and addresses, the discharge probability will increase.
-Following the observations of~\citet{izhikevich_polychronization_2006}, let us assume that such precise discharge motif defines a polychronous group (PG).
-Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
-This allows then to derive a generative model for raster plots (see Figure~\ref{fig:model}-(b)).
 
@@
 %: Detection model
+
 \note{ TODO: show why this formula implements a polychrony detector }
-
 From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.
@@
 We will denote as $W(a, d)$ these weights.
+
+Following the observations of~\citet{izhikevich_polychronization_2006}, let us assume that such precise discharge motif defines a polychronous group (PG).
+Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
+This allows then to derive a generative model for raster plots (see Figure~\ref{fig:model}-(b)).
 
@@
 While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside of a homeostatic state (through a death or explosion of activity).
+\note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}~\citep{perrinet_motion-based_2012}
 

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Methods: 2.1.3

@@
 This defines the full connectivity of the hetero-synaptic model.
-As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic adress space is transformed into a new event stream $\event_\postsynaddrspace$ which is defined by $\{ (\presynaddr_\arank =
+As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic address space is transformed into a new event stream $\event_\postsynaddrspace$ which is defined by
+$\{ (\presynaddr_\arank = 
  \presynaddr_\ranksyn^\postsynaddr, \timev=\timev_\arank + 
@@
 
-Taking the argument the other way around, one may deduce a generative model for generic raster plots.
-Indeed, any spike  in the presynaptic address space is generated by sensory neurons (for instance in the retina) or by afferent neurons.
-These are connected to the spiking cell by a set of weights and delays, and by assuming that these connections are sparsely distributes, one may deduce that this firing was caused by a specific temporal motif.
-, this will generate the propagation in
-
-\note{PGs coccur becaus they share a common cause like chirp ob bird interaural delays Goodman}
-This raster plot is generated either by a event generation process or by other spiking neurons.
-
+Taking the argument the other way around, one may form a generative model for generic raster plots.
+Indeed, any spike in the presynaptic address space is generated by sensory neurons (for instance photoreceptors in the retina, sensors in a CMOS chip) or by afferent spiking neurons.
+In the latter case, these are connected to the spiking cell by a set of weights and delays, whose structure is relatively stable relative to the coding time-scale.
+When these connections are high and sparsely distributed, this firing will cause a specific temporal motif.
+Another example is given for the barn owl auditory system: When it hears the sound of a mouse, this will generate a specific spiking response in both ears, and specifically, the precise timing between these signal generated by the the left relative to the right ear can be for instance used to determine the position of the prey~\citep{goodman_spike-timing-based_2010}.
+Overall, these examples show that raster plots may be considered as a mixture of the effects of different elementary causes, and that each event triggers a specific spatio-temporal motif.
+%
 \subsubsection{Detecting spiking motifs}
@@
 %
+
+\note{formalize the following:}
+More generally, every cause may be considered as occurring independently and one might write the generative model for the generation of the presynaptic events as:
+
+Using this formalization, one might now deduce an optimal algorithm for the detection of such temporal motifs.
 

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Methods: 2.3

@@
 %%%-----------------------------------------------------------------
-
 %
-In particular, we may extend the convolution to a 3D convolution 
+Having defined our task, we may extend the convolution-based detection model to a 3D convolution 
 such that the representation would also benefit from spatial invariance.
-The use of spatiotemporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\cite{ghosh2022conv3D}.
-For that, we design 3D kernels of shape  $(\Kx, \Ky, \Ktime) = (15, 15, 8)$, respectively representing the two spatial dimensions and the range of delays.
+The use of spatiotemporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\cite{ghosh2022conv3D} and we will here use a similar strategy, yet with a focus on using an event-based representation.
+For that, we design 3D kernels of shape  $(\Kx, \Ky, \Ktime) = (9, 9, 11)$, respectively representing the two spatial dimensions and the range of delays.
 Computations are performed on spatio-temporal windows, defined by the kernels, sliding around the events, that is, the center of the spatio-temporal window around the current event $\epsilon_\arank$.
@@
 Finally, the output of the MLR model results in an event with the highest probability class, keeping the same timing as the event as input.
+Importantly, we include the prior knowledge that the event-based input is spatio-temporally clustered, and that these clusters are sparse.
+As a consequence, we only keep a given proportion of the most-active cells which we set here to $.01\%$ of all output voxels.
+Given a spiking input, the model thus gives a spiking output  in the post-synaptic address space.
+
 %
-The loss function of the MLR model is the binary cross entropy on the output of the classification layer.
+As the model is fully differentiable, we may now implement a supervised learning rule.
+The loss function of the MLR model is the binary cross entropy on the output of the classification layer .
 Simulations are performed thanks to the PyTorch library using gradient descent with Adam ($2^{12}$ epochs and a learning rate of $10^{-5}$).

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Methods: 2.3

@@
 %
-Having defined our task, we may extend the convolution-based detection model to a 3D convolution 
+Having defined our task in the visual space for motion detection, we may extend the convolution-based detection model to a 3D convolution 
 such that the representation would also benefit from spatial invariance.
-The use of spatiotemporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\cite{ghosh2022conv3D} and we will here use a similar strategy, yet with a focus on using an event-based representation.
+The use of spatiotemporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\cite{ghosh_spatiotemporal_2019} and we will here use a similar strategy, yet with a focus on using an event-based representation.
 For that, we design 3D kernels of shape  $(\Kx, \Ky, \Ktime) = (9, 9, 11)$, respectively representing the two spatial dimensions and the range of delays.
-Computations are performed on spatio-temporal windows, defined by the kernels, sliding around the events, that is, the center of the spatio-temporal window around the current event $\epsilon_\arank$.
+Computations are performed on the spatio-temporal windows which are defined by the kernels and sliding around the events, that is, the center of the spatio-temporal window around the current event $\epsilon_\arank$.
+Note that to keep with a causal computation, the kernel are shifted in time, such that only past information is giving a response at the present time.
 %
@@
 As the model is fully differentiable, we may now implement a supervised learning rule.
-The loss function of the MLR model is the binary cross entropy on the output of the classification layer .
-Simulations are performed thanks to the PyTorch library using gradient descent with Adam ($2^{12}$ epochs and a learning rate of $10^{-5}$).
+The loss function of the MLR model is the binary cross entropy on the output of the classification layer and can be extended to the spatio-temporal convolutions.
+The supervision was implemented using the input binary events as defined above and the labels as the desired output.
+The labels were defined at each point in time as a one-hot encoding of the present motion in the channel corresponding to the current motion, and for all positions.
+Importantly, we weighted the loss functions only on the support corresponding to the active cells, such that the error was only back-propagated at the spatial locations of these active cells.
+Simulations are performed thanks to the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$).
+We tested the effect of different parameters on a validation set to check the role of each parameter.
+
 %
-
-In particular, we may extend the convolution to a 3D convolution such that the representation would also benefit from spatial invariance which is inherent to the task's construction.
-The use of spatio-temporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\citep{ghosh_spatiotemporal_2019}.
-For that, we design 3D kernels of shape  $(\Kx, \Ky, \Ktime) = (15, 15, 8)$, respectively representing the two spatial dimensions and the range of delays.
-An example of such kernels is given in Figure~\ref{fig:kernels}.
-Computations are performed on spatio-temporal windows, defined by the kernels, sliding around the events, that is, the center of the spatio-temporal window around the current event $\epsilon_\arank$.
-%
-Finally, the output of the MLR model results in an event with the highest probability class, keeping the same timing as the event as input.
-%
-The loss function of the MLR model is the binary cross entropy on the output of the classification layer.
-Simulations are performed thanks to the PyTorch library using gradient descent with Adam (with $2^{12}$ epochs and a learning rate of $10^{-5}$).
+Finally, the output of the MLR model results in an event-based representation predicting at each position and time the probability for each motion.
+Such an output gives a form of optic flow which can be exploited for non-rigid motions, but here we defined for the sake of simplicity a readout method which applies to the full field.
+We have shown above that when different independent observations (here, motion estimated at different spatial locations) are known to have a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of that probability is the sum of the logit of the independent probabilities.
+Taking the logit of the ouput probability given by the model, one can thus compute the probability for the output.
+This allows to compute the accuracy (as the percentage of times when the motion is exactly predicted) or the squared error $|| \speed - ||^2$ of the velocity estimate given by:
+$
+(u, v) = \sum_{} (u_k \cdot Pr(k \vert image) , v_k \cdot Pr(k \vert image) )
+$ 
+These metrics will be useful to compare our methods with other methods on our defined task.
+The full code to reproduce the results of this paper are openly available at \url{https://github.com/SpikeAI/pyTERtorch}.
 %

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Methods: 2.3

@@
 Note that the velocities are sampled on a geometric scale between $1/4$ and $v-speed_base = 4$.
-Next, we define one trajectory of gaze as segments whose duration is drawn from a Poisson distribution with an average block length = $20~\ms$ - similar to a Lévy flight~\citep{}.
+Next, we define one trajectory of gaze as segments whose duration is drawn from a Poisson distribution with an average block length = $20~\ms$ - similar to a Lévy flight~\citep[p.
+289]{mandelbrot_fractal_1982}.
 Finally, the trajectory is integrated by assuming 1/ that velocities are uniformly and independently sampled from the set of different motion-sets and 2/ that motion is uniform during a time segment.

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Methods: 2.3

@@
 $ 
+These computations are performed on a set of input which is different from that used at the training or validation stages.
 These metrics will be useful to compare our methods with other methods on our defined task.
 The full code to reproduce the results of this paper are openly available at \url{https://github.com/SpikeAI/pyTERtorch}.
-%
-\note{kernel size has to be adapted to the characteristics of motion see Samonds %~\citep{samonds}
-}
 %
@@
 For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Figure~\ref{fig:kernels}.
+
+%
+\note{kernel size has to be adapted to the characteristics of motion see Samonds %~\citep{samonds}
+say that we have tested }
+
 %

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Methods: 2.3

@@
 $
-(u, v) = \sum_{} (u_k \cdot Pr(k \vert image) , v_k \cdot Pr(k \vert image) )
+\speed = \sum_{\class} \speed_\class \cdot Pr(\class \vert image)
 $ 

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Methods: 2.3

@@
 The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
-Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that a velocity of $V=1$ corresponds to a movement of one pixel per frame (i.e., per time bin).
+Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that a velocity of $V=1$ corresponds to a motion of one pixel per frame (i.e., per time bin).
 
@@
 %
-Having defined our task in the visual space for motion detection, we may extend the convolution-based detection model to a 3D convolution 
-such that the representation would also benefit from spatial invariance.
-The use of spatiotemporal filters on a stream of events was shown to improve CNN performances for an action recognition task in~\cite{ghosh_spatiotemporal_2019} and we will here use a similar strategy, yet with a focus on using an event-based representation.
-For that, we design 3D kernels of shape  $(\Kx, \Ky, \Ktime) = (9, 9, 11)$, respectively representing the two spatial dimensions and the range of delays.
-Computations are performed on the spatio-temporal windows which are defined by the kernels and sliding around the events, that is, the center of the spatio-temporal window around the current event $\epsilon_\arank$.
-Note that to keep with a causal computation, the kernel are shifted in time, such that only past information is giving a response at the present time.
+Now that our motion detection task is defined in visual space, we can extend the convolution-based detection model to a 3D convolution so that the resulting model would also benefit from spatial invariance.
+The use of spatiotemporal filters on an event stream has improved the performance of CNNs for an action recognition task in~\cite{ghosh_spatiotemporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.
+For this purpose, we design 3D kernels of shape $(\Kx, \Ky, \Ktime) = (9, 9, 11)$, representing the two spatial dimensions and the range of time delays, respectively.
+The computations are performed on the spatiotemporal windows that are defined by the kernels and are "moving" around the events, i.e.
+the spatial center of the spatiotemporal window around the current event $\epsilon_\arank$.
+Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
 %
-Finally, the output of the MLR model results in an event with the highest probability class, keeping the same timing as the event as input.
-Importantly, we include the prior knowledge that the event-based input is spatio-temporally clustered, and that these clusters are sparse.
-As a consequence, we only keep a given proportion of the most-active cells which we set here to $.01\%$ of all output voxels.
-Given a spiking input, the model thus gives a spiking output  in the post-synaptic address space.
+The output of the MLR model corresponds to an event with the highest probability class, keeping the same timing as the input event.
+It is important to note that we include the prior knowledge that the event-based input is spatio-temporally clustered and that these clusters are sparse.
+Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels.
+Given a stimulus input, the model thus yields a spike output in the postsynaptic address space.
 
 %
-As the model is fully differentiable, we may now implement a supervised learning rule.
-The loss function of the MLR model is the binary cross entropy on the output of the classification layer and can be extended to the spatio-temporal convolutions.
-The supervision was implemented using the input binary events as defined above and the labels as the desired output.
-The labels were defined at each point in time as a one-hot encoding of the present motion in the channel corresponding to the current motion, and for all positions.
-Importantly, we weighted the loss functions only on the support corresponding to the active cells, such that the error was only back-propagated at the spatial locations of these active cells.
-Simulations are performed thanks to the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$).
-We tested the effect of different parameters on a validation set to check the role of each parameter.
+Since the model is fully differentiable, we can now implement a supervised learning rule.
+The loss function of the MLR model is the binary cross-entropy on the output of the classification layer and can be extended to space-time convolutions.
+Supervision was implemented using the input binary events as defined above and the labels as the desired output.
+The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions.
+It is important to note that we weighted the cost function by considering only active cells, such that the error is only backpropagated to the spatial locations of these most active cells.
+Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$).
+We tested the effect of different parameters on a validation set to quantify the role of each parameter.
 
 %
-Finally, the output of the MLR model results in an event-based representation predicting at each position and time the probability for each motion.
-Such an output gives a form of optic flow which can be exploited for non-rigid motions, but here we defined for the sake of simplicity a readout method which applies to the full field.
-We have shown above that when different independent observations (here, motion estimated at different spatial locations) are known to have a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of that probability is the sum of the logit of the independent probabilities.
-Taking the logit of the ouput probability given by the model, one can thus compute the probability for the output.
-This allows to compute the accuracy (as the percentage of times when the motion is exactly predicted) or the squared error $|| \speed - ||^2$ of the velocity estimate given by:
+Finally, the output of the MLR model is an event-based representation predicting at each position and at each time the probability of each motion.
+Such an output gives a form of optical flow that can be exploited for non-rigid motions, but we have defined here, for simplicity, an evaluation method that applies to our task with a full field motion.
+We have shown above that when different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities.
+By taking the logit of the probability of the output given by the model, we can therefore calculate the probability of the output.
+This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error $|| \speed - ||^2$ of the velocity estimate given by :
 $
@@
 $ 
-These computations are performed on a set of input which is different from that used at the training or validation stages.
+These calculations are performed on a different input dataset than the one used in the training or validation steps.
 These metrics will be useful to compare our methods with other methods on our defined task.
-The full code to reproduce the results of this paper are openly available at \url{https://github.com/SpikeAI/pyTERtorch}.
+The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/pyTERtorch}.
 %

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Methods: 2.3
Introduction: 1.1.1

@@
 %
-\subsection{Processing of dynamical sensory inputs}
+\subsection{Recent challenges in machine learning and computer vision}
 %%%-----------------------------------------------------------------
+Machine learning, and computer vision in particular, have seen a sharp transition in modern times from moving from science fiction movies to a mainstream status, present in the smartphones that sit in each and everybody's pocket.
+More hidden but even more important to our economy and security, these automated tools are now prominent on the servers that empowers essentially all sectors of our daily life, especially medicine for which it empowers practionners with now diagnostic tools to heal people sooner and better.
+This particular example illustrates also some important challenges increasingly paced by these tools.
+One of these is their sensibility to adversarial attacks, that is the sensibility of fire models output to minute changes of the input.
+This Could rain important issues on the reliability of the results y an automated diagnostic model.
+More generally, these models back also to offer why it forms a given output for a class of inputs.
+This back.
+of generality hinders the general advances and acceptance of these methods for secure-sensitive tasks.
+
+\subsection{What can neuroscience bring to computer vision?}% and the Hetero-synaptic delays in spiking neurons}
+%%%-----------------------------------------------------------------
+
 In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking motifs~\citep{abeles_role_1982}.
@@
 
-\subsection{Hetero-synaptic delays in spiking neurons}
-%%%-----------------------------------------------------------------
-In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.
-Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays.
-This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
-However, most current neuroscience-inspired computer vision algorithms (for instance convolution neural networks) do not make use of this dynamic aspect.
-A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance.
-The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations.
-Yet, flexible and robust event-driven algorithms for classical computer vision tasks, such as motion detection, are still not able to compete with the state-of-the-art of dense computer vision solutions.
-
-\subsection{Outline}
-%%%-----------------------------------------------------------------
 In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}.
@@
 %
+
+In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.
+Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays.
+This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
+However, most current neuroscience-inspired computer vision algorithms (for instance convolution neural networks) do not make use of this dynamic aspect.
+A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance.
+The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations.
+Yet, flexible and robust event-driven algorithms for classical computer vision tasks, such as motion detection, are still not able to compete with the state-of-the-art of dense computer vision solutions.
+
+\subsection{Outline}
+%%%-----------------------------------------------------------------
 
@@
 say that we have tested }
+\note{tester la meme expe faite par baudot et modelisée par kremkow}
 

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.1.1 & 1.1.2 a

@@
 %%%-----------------------------------------------------------------
+% 1.1.1
 Machine learning, and computer vision in particular, have seen a sharp transition in modern times from moving from science fiction movies to a mainstream status, present in the smartphones that sit in each and everybody's pocket.
@@
 of generality hinders the general advances and acceptance of these methods for secure-sensitive tasks.
+
+% 1.1.2 a
+Taking some perspective, one remembers that these tools are inherited by theoretical models made to understand the brain.
+while these first neural network were mainly applied on toy-like problems, they were also offering a growing number of applications in computer vision but also more generally for automating tasks.
+In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates and reaches now super-human performances for categorizing images.
+One challenge with these models is to be able to scale them up for prospective applications, for instance in autonomous driving.
+Indeed, in such applications, the input does not consist of a single image (let's say of about a M byte) but of a continuous stream of images from multiple cameras.
+
+% 1.1.2 b
 

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.1

@@
 % 1.1.2 b
+Challenges are numerous in
+that case and we should
+put forward two major ones:
+rapidity and energy budget.
+First, most existing solutions are
+being built incrementally from
+low-definition images such as
+Le Net for MN is T (Le Cun, 98),
+to high definition images, such as
+V94 on Imagen et (V9 416).
+This comes at a cost and notably
+that the computation time increases
+with image size. This has been tampered
+by the increase of computational power
+available, notably with the parallel
+processing implemented in GPU cards.
+However, transitioning from images
+to videos, and then to multiple cameras
+involves a change of scale in the
+amount of data processed. It
+raises also a question which is often
+of little importance in image
+categorization. Indeed, in the setting
+of autonomous driving, an important
+task is to respond as fast as possible
+be a eventual collision, for instance as
+a pedestrian crosses the street. As such,
+future systems should always be able
+to give a response at any time, instead
+of having to wait until the pipeline
+delivers the correct response. This
+online processing is also a hallmark
+of processing in biological systems.
+This could be a first important
+contribution if we could translate
+these principles into future automated
+processes.
+
+second, a major constraint on
+futur systems is energy budget.
+As we saw, a fait computation
+time was kept for high-dimension
+images, but this come at the cost
+of a higher energy consumption.
+Also, the need he solve increasingly
+more complex tasks necessitated the
+construction of increasingly more
+complex and deepen architectures.
+As en exaupte the top performing
+net on the Imagent challenge is
+using billions of parameters which
+tipycally cost 1Wh to train and
+approximately 1Wh per inference
+of a single image. Less greedy
+solutions have emerged such
+as Mobile Net, yet these solutions
+are yet not pit to be used
+in practice in autonomous
+Lars, as the added consumption
+would be approximately Bow
+that is, 200 to of the energy
+of an electric can in standaro
+conditions.
+
+Neuroscience could bring
+breakthrough answers to these
+two intertwined problems-Indeed
+the human brain has the remarkable
+property of being mostly autonomous
+in it a learning, to be able to act
+and respond at any time, and get to
+consume en reasonable amount of energy,
+estimated to 5W at rest. This system
+is the result of million of years of natural
+selection, get a striking property of
+biological neural networks compared
+to an artificial one is their respective
+representation. CNN s for instance
+represent the information passing from
+one lage to the other as tensons representing
+visual information in a dense manner on
+the visual topography, while representing
+different properties in different channels.
+On the other side, information in avait
+majority of biological neural networks
+is represented as spikes, that is, as
+prototypical all-or-none events whose
+only parameters are their timing and
+the address of the neuron that emitted
+that spite. Why spikes ? It is not
+yet known the reason why most systems
+evolved to use that representation, but
+one major advantage sans to be a
+form of efficiency. In the human brain,
+each of the about to trillion nerve cells
+emit on average 1 spike per second,
+but with a high variability and up to
+300 spikes per second. There has been
+numerous attempts to build artificial
+spiking neural networks, and
+notabley to have the same performance
+en that observed in classical cans, yet
+with a lower energy budget. Yet, un
+such SNN has yet surpassed state-of
+the art performances. Here, we will attempt
+to focus on a neurobiologically inspired
+property of spiking neurons to advance
+in that endeavour.
 
@@
 %%%-----------------------------------------------------------------
+% 1.2.a
 

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.2.a

@@
 % 1.2.a
+Let's focus first on some existing
+spiking neural network models. one
+first motivation for these models was
+to understand the biological observations
+at different scales. For instance the Hodgkin
+Huxley mode describes the 'dynamic
+of the membrane potential and the role
+of the different voltage-gated ion channels. It
+also explains how the all may emit
+a spike. Larger scale models introduce
+an explicit spiking mechanism, for
+instance in the Integrate-and-Fine
+model which allows to predict the
+emerging states of a network of recurrently
+connected neurons (Brunel). Several
+models have sealed up the system to
+include more neurons such as the
+Virtual Brain (Sanz) which allows
+to mix spiking neural nets with
+analog neural mass models. However
+these methods are merely descriptive
+and have application in neuroscience
+and few have had an impact in
+computer vision.
+
+Another class of attempts have designed
+spiking neural networks that aim to
+have applications in machine learning.
+One of these is the Spike Nap algorithm
+which uses a radical approach by
+anding information using one spite
+per nen non :(Delorme). Another branch
+of spiking neural networks attempt
+to determine the structure of the
+network such as to minimize a cot
+function. This has been implemented
+in the Spike Prop algorithm (Bonte)
+and is now widely used in methods
+Trying to transfer the performance
+of CNN s to Sn Ns, for instance by
+using a surrogate gradient approach.
+However, the performance of SNN always
+fall short of that of analog networks
+and the question of the advantage of
+using spin is still open.
+A novel
+generation of algorithms has
+emerged with so-called silicon retina
+which mimi de the output of the retina,
+by representing the information using
+espibes. one of this approach is that
+of Lagorce, which uses the time
+to the last spike to encode information in
+in so-called temporal surfaces. We
+have recently extended this Mots
+algorithm to provide with an online
+response and shown near-son -
+performances on categwitha notable more
+energy-efficient budget. A naige
+of similar algorithms are applied
+to real-life applications like optic
+flow (Benotman), depth estimation
+(Dandelot) on Gesture-Recently such
+algorithm was used in the space
+industry which is famous for its stringeant
+selection process.
 

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.2.a
Introduction: 1.1

@@
 %%%-----------------------------------------------------------------
-% 1.1.1
 Machine learning, and computer vision in particular, have seen a sharp transition in modern times from moving from science fiction movies to a mainstream status, present in the smartphones that sit in each and everybody's pocket.
-More hidden but even more important to our economy and security, these automated tools are now prominent on the servers that empowers essentially all sectors of our daily life, especially medicine for which it empowers practionners with now diagnostic tools to heal people sooner and better.
-This particular example illustrates also some important challenges increasingly paced by these tools.
-One of these is their sensibility to adversarial attacks, that is the sensibility of fire models output to minute changes of the input.
-This Could rain important issues on the reliability of the results y an automated diagnostic model.
-More generally, these models back also to offer why it forms a given output for a class of inputs.
-This back.
-of generality hinders the general advances and acceptance of these methods for secure-sensitive tasks.
+Less visible but even more important to our economy and security, numerous automated tools are now prominent in the services that empowers essentially all sectors of our daily life.
+This is most acute in medicine for which it empowers practionners with novel diagnostic tools to heal people sooner and better.
+This particular example illustrates also some important challenges increasingly faced by these tools.
+One of these is their sensibility to adversarial attacks, that is the sensibility of the models' output to minute changes in the input.
+This Could raise important issues on the reliability about the results any automated diagnostic model.
+More generally, these models lack also to offer why it forms a given output for a class of inputs.
+This lack of explainability hinders the general advances and acceptance of these methods for most secure-sensitive tasks.
 
-% 1.1.2 a
 Taking some perspective, one remembers that these tools are inherited by theoretical models made to understand the brain.
-while these first neural network were mainly applied on toy-like problems, they were also offering a growing number of applications in computer vision but also more generally for automating tasks.
+While the first artificial neural network were mainly applied on toy-like problems, they were also offering a growing number of applications in computer vision but also more generally for automating tasks.
 In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates and reaches now super-human performances for categorizing images.
 One challenge with these models is to be able to scale them up for prospective applications, for instance in autonomous driving.
-Indeed, in such applications, the input does not consist of a single image (let's say of about a M byte) but of a continuous stream of images from multiple cameras.
+Indeed, in such applications, the input does not consist of a single image (let's say of about a few million pixels) but of a continuous stream of images from multiple cameras.
 
-% 1.1.2 b
-Challenges are numerous in
-that case and we should
-put forward two major ones:
-rapidity and energy budget.
-First, most existing solutions are
-being built incrementally from
-low-definition images such as
-Le Net for MN is T (Le Cun, 98),
-to high definition images, such as
-V94 on Imagen et (V9 416).
-This comes at a cost and notably
-that the computation time increases
-with image size. This has been tampered
-by the increase of computational power
-available, notably with the parallel
-processing implemented in GPU cards.
-However, transitioning from images
-to videos, and then to multiple cameras
-involves a change of scale in the
-amount of data processed. It
-raises also a question which is often
-of little importance in image
-categorization. Indeed, in the setting
-of autonomous driving, an important
-task is to respond as fast as possible
-be a eventual collision, for instance as
-a pedestrian crosses the street. As such,
-future systems should always be able
-to give a response at any time, instead
-of having to wait until the pipeline
-delivers the correct response. This
-online processing is also a hallmark
-of processing in biological systems.
-This could be a first important
-contribution if we could translate
-these principles into future automated
-processes.
+Challenges are numerous in that case and we should put forward two major hurdls: rapidity and energy budget.
+First, most existing solutions are being built incrementally from low-definition images such as Le Net for MNIST (Le Cun, 98), to high definition images, such as VGG on Imagenet (VGG16).
+This comes at a cost and notably that the computation time increases with image size.
+This has been tampered by the increase of computational power available, notably with the parallel processing implemented in GPU cards.
+However, transitioning from images to videos, and then to multiple cameras involves a change of scale in the amount of data processed.
+It raises also a question which is often of little importance when dealing with image categorization.
+Indeed, in the setting of autonomous driving, an important task is to respond as fast as possible to an eventual collision, for instance as a pedestrian crosses the street.
+As such, future systems should always be able to give a response at any time, instead of having to wait until the pipeline delivers the correct response.
+This online processing is also a hallmark of processing in biological systems.
+This could be a first important contribution if we could translate these principles into future automated processes.
 
-second, a major constraint on
-futur systems is energy budget.
-As we saw, a fait computation
-time was kept for high-dimension
-images, but this come at the cost
-of a higher energy consumption.
-Also, the need he solve increasingly
-more complex tasks necessitated the
-construction of increasingly more
-complex and deepen architectures.
-As en exaupte the top performing
-net on the Imagent challenge is
-using billions of parameters which
-tipycally cost 1Wh to train and
-approximately 1Wh per inference
-of a single image. Less greedy
-solutions have emerged such
-as Mobile Net, yet these solutions
-are yet not pit to be used
-in practice in autonomous
-Lars, as the added consumption
-would be approximately Bow
-that is, 200 to of the energy
-of an electric can in standaro
-conditions.
+Second, a major constraint on future systems is energy budget.
+As we saw, a fast computation time was maintained for high-dimension images, but this come at the cost of a higher energy consumption.
+Also, the need to solve increasingly more complex tasks necessitated the construction of increasingly more complex and deeper architectures.
+As en example, the top performing model on the Imagenet challenge is using billions of parameters which typically cost 1Wh to train and approximately 1Wh per inference of a single image.
+Less greedy solutions have emerged such as Mobile Net, yet these solutions are yet not ready to be used in practice in autonomous cars, as the added consumption would be approximately 300W that is, 20\% of the energy of an electric can in standard conditions.
 
-Neuroscience could bring
-breakthrough answers to these
-two intertwined problems-Indeed
-the human brain has the remarkable
-property of being mostly autonomous
-in it a learning, to be able to act
-and respond at any time, and get to
-consume en reasonable amount of energy,
-estimated to 5W at rest. This system
-is the result of million of years of natural
-selection, get a striking property of
-biological neural networks compared
-to an artificial one is their respective
-representation. CNN s for instance
-represent the information passing from
-one lage to the other as tensons representing
-visual information in a dense manner on
-the visual topography, while representing
-different properties in different channels.
-On the other side, information in avait
-majority of biological neural networks
-is represented as spikes, that is, as
-prototypical all-or-none events whose
-only parameters are their timing and
-the address of the neuron that emitted
-that spite. Why spikes ? It is not
-yet known the reason why most systems
-evolved to use that representation, but
-one major advantage sans to be a
-form of efficiency. In the human brain,
-each of the about to trillion nerve cells
-emit on average 1 spike per second,
-but with a high variability and up to
-300 spikes per second. There has been
-numerous attempts to build artificial
-spiking neural networks, and
-notabley to have the same performance
-en that observed in classical cans, yet
-with a lower energy budget. Yet, un
-such SNN has yet surpassed state-of
-the art performances. Here, we will attempt
-to focus on a neurobiologically inspired
-property of spiking neurons to advance
-in that endeavour.
+Neuroscience could bring breakthrough answers to these two intertwined problems.
+Indeed, the human brain has the remarkable property of being able to act and respond at any time, while consuming a reasonable amount of energy, estimated to 5W at rest.
+This system is the result of million of years of natural selection, and a striking difference of biological neural networks compared to an artificial one is the representation that they use.
+CNNs for instance represent the information passing from one layer to the other as tensors representing visual information in a dense manner on the visual topography, while representing different properties in different channels.
+On the other side, information in a vast majority of biological neural networks is represented as spikes, that is, as prototypical all-or-none events whose only parameters are their timing and the address of the neuron that emitted that spite.
+Why spikes ?
+It is not yet known the reason why most systems evolved to use that representation, but one major advantage seems to be a form of efficiency.
+In the human brain, each of the about ten trillion nerve cells emit on average 1 spike per second, but each neurons fires with a high variability, and up to 300 spikes per second.
+There has been numerous attempts to build artificial spiking neural networks, and notably to have the same performance as that observed in classical CNNs, yet with a lower energy budget.
+Yet, no such SNN has yet surpassed state-of the art performances.
+Here, we will attempt to focus on a neurobiologically inspired property of spiking neurons to advance in that endeavor.
 
@@
 % 1.2.a
-Let's focus first on some existing
-spiking neural network models. one
-first motivation for these models was
-to understand the biological observations
-at different scales. For instance the Hodgkin
-Huxley mode describes the 'dynamic
-of the membrane potential and the role
-of the different voltage-gated ion channels. It
-also explains how the all may emit
-a spike. Larger scale models introduce
-an explicit spiking mechanism, for
-instance in the Integrate-and-Fine
-model which allows to predict the
-emerging states of a network of recurrently
-connected neurons (Brunel). Several
-models have sealed up the system to
-include more neurons such as the
-Virtual Brain (Sanz) which allows
-to mix spiking neural nets with
-analog neural mass models. However
-these methods are merely descriptive
-and have application in neuroscience
-and few have had an impact in
-computer vision.
+Let's focus first on some existing spiking neural network models.
+one first motivation for these models was to understand the biological observations at different scales.
+For instance the Hodgkin Huxley mode describes the 'dynamic of the membrane potential and the role of the different voltage-gated ion channels.
+It also explains how the all may emit a spike.
+Larger scale models introduce an explicit spiking mechanism, for instance in the Integrate-and-Fine model which allows to predict the emerging states of a network of recurrently connected neurons (Brunel).
+Several models have sealed up the system to include more neurons such as the Virtual Brain (Sanz) which allows to mix spiking neural nets with analog neural mass models.
+However these methods are merely descriptive and have application in neuroscience and few have had an impact in computer vision.
 
-Another class of attempts have designed
-spiking neural networks that aim to
-have applications in machine learning.
-One of these is the Spike Nap algorithm
-which uses a radical approach by
-anding information using one spite
-per nen non :(Delorme). Another branch
-of spiking neural networks attempt
-to determine the structure of the
-network such as to minimize a cot
-function. This has been implemented
-in the Spike Prop algorithm (Bonte)
-and is now widely used in methods
-Trying to transfer the performance
-of CNN s to Sn Ns, for instance by
-using a surrogate gradient approach.
-However, the performance of SNN always
-fall short of that of analog networks
-and the question of the advantage of
-using spin is still open.
-A novel
-generation of algorithms has
-emerged with so-called silicon retina
-which mimi de the output of the retina,
-by representing the information using
-espibes. one of this approach is that
-of Lagorce, which uses the time
-to the last spike to encode information in
-in so-called temporal surfaces. We
-have recently extended this Mots
-algorithm to provide with an online
-response and shown near-son -
-performances on categwitha notable more
-energy-efficient budget. A naige
-of similar algorithms are applied
-to real-life applications like optic
-flow (Benotman), depth estimation
-(Dandelot) on Gesture-Recently such
-algorithm was used in the space
-industry which is famous for its stringeant
-selection process.
+Another class of attempts have designed spiking neural networks that aim to have applications in machine learning.
+One of these is the Spike Nap algorithm which uses a radical approach by anding information using one spite per nen non :(Delorme).
+Another branch of spiking neural networks attempt to determine the structure of the network such as to minimize a cot function.
+This has been implemented in the Spike Prop algorithm (Bonte) and is now widely used in methods Trying to transfer the performance of CNN s to Sn Ns, for instance by using a surrogate gradient approach.
+However, the performance of SNN always fall short of that of analog networks and the question of the advantage of using spin is still open.
+A novel generation of algorithms has emerged with so-called silicon retina which mimi de the output of the retina, by representing the information using espibes.
+one of this approach is that of Lagorce, which uses the time to the last spike to encode information in in so-called temporal surfaces.
+We have recently extended this Mots algorithm to provide with an online response and shown near-son - performances on categwitha notable more energy-efficient budget.
+A naige of similar algorithms are applied to real-life applications like optic flow (Benotman), depth estimation (Dandelot) on Gesture-Recently such algorithm was used in the space industry which is famous for its stringeant selection process.
 

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.1

@@
 %%%-----------------------------------------------------------------
-Machine learning, and computer vision in particular, have seen a sharp transition in modern times from moving from science fiction movies to a mainstream status, present in the smartphones that sit in each and everybody's pocket.
-Less visible but even more important to our economy and security, numerous automated tools are now prominent in the services that empowers essentially all sectors of our daily life.
-This is most acute in medicine for which it empowers practionners with novel diagnostic tools to heal people sooner and better.
-This particular example illustrates also some important challenges increasingly faced by these tools.
-One of these is their sensibility to adversarial attacks, that is the sensibility of the models' output to minute changes in the input.
-This Could raise important issues on the reliability about the results any automated diagnostic model.
-More generally, these models lack also to offer why it forms a given output for a class of inputs.
-This lack of explainability hinders the general advances and acceptance of these methods for most secure-sensitive tasks.
+Machine learning, and computer vision in particular, have undergone a major transformation in modern times, leaving science fiction movies to become mainstream tools found in the smartphones in everyone's pocket.
+Less visible, but even more crucial to our economy and security, are the many automated tools that now dominate the range of services that equip virtually all sectors of our daily lives.
+This trend is most pronounced in the field of medicine, as it provides practitioners with new diagnostic tools to treat people faster and more effectively.
+This particular example also illustrates some important challenges that these tools are increasingly facing.
+One of these is their sensitivity to adversarial attacks, that is, the sensitivity of the models' output to minute changes in the input.
+This could raise important questions about the reliability of the results of in any automated diagnostic model.
+More generally, these models do not explain why they produce a given output for a class of inputs.
+This lack of explainability is a hindrance to the advancement of these methodologies as well as to their acceptance for most safety-relevant tasks.
 
-Taking some perspective, one remembers that these tools are inherited by theoretical models made to understand the brain.
-While the first artificial neural network were mainly applied on toy-like problems, they were also offering a growing number of applications in computer vision but also more generally for automating tasks.
-In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates and reaches now super-human performances for categorizing images.
-One challenge with these models is to be able to scale them up for prospective applications, for instance in autonomous driving.
-Indeed, in such applications, the input does not consist of a single image (let's say of about a few million pixels) but of a continuous stream of images from multiple cameras.
+If we take a step back, we will remember that these tools are inherited from theoretical models designed to understand the brain.
+While the first artificial neural networks were mainly applied to " toy " type problems, they also offered a growing number of applications in computer vision, but also more generally for task automation.
+In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates, and today they reach higher performances than humans for image categorization.
+One of the challenges of these models is to be able to scale them up for future applications, for example in autonomous driving.
+Indeed, in such applications, the input does not consist of a single image (say a few million pixels) but of a continuous stream of images from several cameras.
 
-Challenges are numerous in that case and we should put forward two major hurdls: rapidity and energy budget.
-First, most existing solutions are being built incrementally from low-definition images such as Le Net for MNIST (Le Cun, 98), to high definition images, such as VGG on Imagenet (VGG16).
-This comes at a cost and notably that the computation time increases with image size.
-This has been tampered by the increase of computational power available, notably with the parallel processing implemented in GPU cards.
-However, transitioning from images to videos, and then to multiple cameras involves a change of scale in the amount of data processed.
-It raises also a question which is often of little importance when dealing with image categorization.
-Indeed, in the setting of autonomous driving, an important task is to respond as fast as possible to an eventual collision, for instance as a pedestrian crosses the street.
-As such, future systems should always be able to give a response at any time, instead of having to wait until the pipeline delivers the correct response.
-This online processing is also a hallmark of processing in biological systems.
-This could be a first important contribution if we could translate these principles into future automated processes.
+The challenges are numerous in such situations and two major obstacles must be highlighted: the computational delay and the energy budget.
+First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST (Le Cun, 98), to high definition images, such as VGG on Imagenet (VGG16).
+This has a cost and notably the fact that the computation time increases with the size of the image.
+This phenomenon has been mitigated by the increase in available computing power, notably thanks to the parallel processing implemented in GPU cards.
+However, the transition from images to videos, then to multiple cameras implies a further change of scale in the amount of data processed.
+It also raises an issue that is often not addressed when it comes to image categorization.
+Indeed, in autonomous driving, an important task is to react as quickly as possible to a possible collision, for example when a pedestrian crosses the street.
+As such, future systems should always be able to give a response at any time, instead of having to wait for the pipeline to provide the correct response.
+This online processing is also a feature of processing in biological systems and could be an important first contribution if we could translate these principles into future automated processes.
 

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.1

@@
 This lack of explainability is a hindrance to the advancement of these methodologies as well as to their acceptance for most safety-relevant tasks.
-
 If we take a step back, we will remember that these tools are inherited from theoretical models designed to understand the brain.
@@
 This online processing is also a feature of processing in biological systems and could be an important first contribution if we could translate these principles into future automated processes.
-
-Second, a major constraint on future systems is energy budget.
-As we saw, a fast computation time was maintained for high-dimension images, but this come at the cost of a higher energy consumption.
-Also, the need to solve increasingly more complex tasks necessitated the construction of increasingly more complex and deeper architectures.
-As en example, the top performing model on the Imagenet challenge is using billions of parameters which typically cost 1Wh to train and approximately 1Wh per inference of a single image.
-Less greedy solutions have emerged such as Mobile Net, yet these solutions are yet not ready to be used in practice in autonomous cars, as the added consumption would be approximately 300W that is, 20\% of the energy of an electric can in standard conditions.
-
-Neuroscience could bring breakthrough answers to these two intertwined problems.
-Indeed, the human brain has the remarkable property of being able to act and respond at any time, while consuming a reasonable amount of energy, estimated to 5W at rest.
-This system is the result of million of years of natural selection, and a striking difference of biological neural networks compared to an artificial one is the representation that they use.
-CNNs for instance represent the information passing from one layer to the other as tensors representing visual information in a dense manner on the visual topography, while representing different properties in different channels.
-On the other side, information in a vast majority of biological neural networks is represented as spikes, that is, as prototypical all-or-none events whose only parameters are their timing and the address of the neuron that emitted that spite.
-Why spikes ?
-It is not yet known the reason why most systems evolved to use that representation, but one major advantage seems to be a form of efficiency.
-In the human brain, each of the about ten trillion nerve cells emit on average 1 spike per second, but each neurons fires with a high variability, and up to 300 spikes per second.
-There has been numerous attempts to build artificial spiking neural networks, and notably to have the same performance as that observed in classical CNNs, yet with a lower energy budget.
-Yet, no such SNN has yet surpassed state-of the art performances.
-Here, we will attempt to focus on a neurobiologically inspired property of spiking neurons to advance in that endeavor.
-
+Second, a major constraint for future systems is energy cost.
+As we have seen, fast computation time has been maintained for high-dimensional images, but at the cost of higher energy consumption.
+In addition, the need to solve increasingly complex tasks has required the construction of increasingly complex and deep architectures.
+As an example, the best performing model in the Imagenet challenge uses billions of parameters that typically cost 1Wh to train and about 1Wh to infer a single image.
+Less power-hungry solutions have emerged, such as Mobile Net, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be about 300W, or 20\% of the energy of an electric car under standard conditions.
+Neuroscience could provide revolutionary answers to these two closely related problems.
+Indeed, the human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, estimated at 5W at rest.
+This system is the result of millions of years of natural selection, and a striking difference between biological neural networks and artificial ones is the representation they use.
+CNNs, for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels.
+On the other hand, information in a large majority of biological neural networks is represented as spikes, i.e.
+prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted this spike.
+Why spikes?
+It is not yet known why most biological systems have evolved to use this representation, but one major advantage seems to be some form of efficiency.
+In the human brain, each of the approximately ten trillion nerve cells emits on average one spike per second, but each neuron fires with great variability, and up to 300 spikes per second.
+Many attempts have been made to build artificial spiking neural networks, and in particular to achieve the same performance as observed in classical CNNs, but with a lower energy budget.
+However, none of these networks has yet surpassed the state of the art performances.
+Here, we will try to focus on a property of spike neurons inspired by neurobiology to progress in this direction.
 \subsection{What can neuroscience bring to computer vision?}% and the Hetero-synaptic delays in spiking neurons}

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.2

@@
 Less power-hungry solutions have emerged, such as Mobile Net, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be about 300W, or 20\% of the energy of an electric car under standard conditions.
+
 Neuroscience could provide revolutionary answers to these two closely related problems.
@@
 Here, we will try to focus on a property of spike neurons inspired by neurobiology to progress in this direction.
+%
 \subsection{What can neuroscience bring to computer vision?}% and the Hetero-synaptic delays in spiking neurons}
 %%%-----------------------------------------------------------------
-% 1.2.a
-Let's focus first on some existing spiking neural network models.
-one first motivation for these models was to understand the biological observations at different scales.
-For instance the Hodgkin Huxley mode describes the 'dynamic of the membrane potential and the role of the different voltage-gated ion channels.
-It also explains how the all may emit a spike.
-Larger scale models introduce an explicit spiking mechanism, for instance in the Integrate-and-Fine model which allows to predict the emerging states of a network of recurrently connected neurons (Brunel).
-Several models have sealed up the system to include more neurons such as the Virtual Brain (Sanz) which allows to mix spiking neural nets with analog neural mass models.
-However these methods are merely descriptive and have application in neuroscience and few have had an impact in computer vision.
+Let us first focus on some existing spiking neural network models.
+A first motivation for these models was to understand biological observations at different scales.
+For example, the Hodgkin-Huxley model describes the dynamics of the membrane potential and the role of different voltage-dependent ion channels.
+It also explains how the model can produce a spike.
+Larger scale models introduce an explicit spiking mechanism, for example in the Integrate-and-Fire model.
+It predicts the emergent states of a recurrently connected neural network (Brunel).
+Several models have increased the scale of the system to include more neurons, such as the Virtual Brain (Sanz), which mixes spike neural networks with analog neural mass models.
+However, these methods are mainly descriptive with applications in neuroscience but few have had an application in computer vision.
 
-Another class of attempts have designed spiking neural networks that aim to have applications in machine learning.
-One of these is the Spike Nap algorithm which uses a radical approach by anding information using one spite per nen non :(Delorme).
-Another branch of spiking neural networks attempt to determine the structure of the network such as to minimize a cot function.
-This has been implemented in the Spike Prop algorithm (Bonte) and is now widely used in methods Trying to transfer the performance of CNN s to Sn Ns, for instance by using a surrogate gradient approach.
-However, the performance of SNN always fall short of that of analog networks and the question of the advantage of using spin is still open.
-A novel generation of algorithms has emerged with so-called silicon retina which mimi de the output of the retina, by representing the information using espibes.
-one of this approach is that of Lagorce, which uses the time to the last spike to encode information in in so-called temporal surfaces.
-We have recently extended this Mots algorithm to provide with an online response and shown near-son - performances on categwitha notable more energy-efficient budget.
-A naige of similar algorithms are applied to real-life applications like optic flow (Benotman), depth estimation (Dandelot) on Gesture-Recently such algorithm was used in the space industry which is famous for its stringeant selection process.
+Another class of approaches have designed spiking neural networks that aim to have applications in machine learning.
+One of these is the SpikeNet algorithm which uses a radical approach by encoding information using one spike per neuron (Delorme).
+Another type of spiking neural network attempts to determine the structure of the network in order to minimize a cost function.
+This was implemented in the SpikeProp algorithm (Bohte) and is now widely used in methods that attempt to transfer performance from CNNs to SNNs, for example using a surrogate gradient approach.
+However, the performance of SNNs is still lower than that of analog networks and the question of the advantage of using spikes remains open.
+A new generation of algorithms has emerged with the advent of the so-called silicon retina, which mimics the output of the retina by representing information using spikes.
+One such approach is that of Lagorce, which uses the time to the last spike to encode information into so-called temporal surfaces.
+We have recently extended this HOTS algorithm to provide an online response and have shown performance close to that of SOTA for categorization, with a much more efficient energy budget.
+A series of similar algorithms are applied to real-life applications such as optical flow (Benosman), depth estimation (Dardelet), or gesture recognition (Marro).
+Recently, such an algorithm has been used in the space industry which is famous for its stringent selection process.
 

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.2

@@
 A new generation of algorithms has emerged with the advent of the so-called silicon retina, which mimics the output of the retina by representing information using spikes.
-One such approach is that of Lagorce, which uses the time to the last spike to encode information into so-called temporal surfaces.
-We have recently extended this HOTS algorithm to provide an online response and have shown performance close to that of SOTA for categorization, with a much more efficient energy budget.
+In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}.
+This HOTS model represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events.
+By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is in an online fashion.
+We have demonstrated on several datasets that it provides with efficient computations resulting in ultra-fast categorization.
+Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.
 A series of similar algorithms are applied to real-life applications such as optical flow (Benosman), depth estimation (Dardelet), or gesture recognition (Marro).
@@
 All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal motifs in biological spike trains.
-
-In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}.
-This HOTS model represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events.
-By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is in an online fashion.
-We have demonstrated on several datasets that it provides with efficient computations resulting in ultra-fast categorization.
-Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.
+In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.
+Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays.
+This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
+However, most current neuroscience-inspired computer vision algorithms (for instance convolution neural networks) do not make use of this dynamic aspect.
+A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance.
+The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations.
+Yet, flexible and robust event-driven algorithms for classical computer vision tasks, such as motion detection, are still not able to compete with the state-of-the-art of dense computer vision solutions.
 
@@
 %
-
-In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.
-Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays.
-This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
-However, most current neuroscience-inspired computer vision algorithms (for instance convolution neural networks) do not make use of this dynamic aspect.
-A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance.
-The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations.
-Yet, flexible and robust event-driven algorithms for classical computer vision tasks, such as motion detection, are still not able to compete with the state-of-the-art of dense computer vision solutions.
 

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.3

@@
 However, these methods are mainly descriptive with applications in neuroscience but few have had an application in computer vision.
-
 Another class of approaches have designed spiking neural networks that aim to have applications in machine learning.
@@
 However, the performance of SNNs is still lower than that of analog networks and the question of the advantage of using spikes remains open.
+
 A new generation of algorithms has emerged with the advent of the so-called silicon retina, which mimics the output of the retina by representing information using spikes.
+%A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance.
+The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations.
 In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}.
@@
 
+These computational results suggest that the temporal representation underlying spiking representation may be a key ingredient in the efficiency of the algorithms, but one may wonder how this can be supported by neurobiological data.
 In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking motifs~\citep{abeles_role_1982}.
@@
 This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
-However, most current neuroscience-inspired computer vision algorithms (for instance convolution neural networks) do not make use of this dynamic aspect.
-A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance.
-The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations.
-Yet, flexible and robust event-driven algorithms for classical computer vision tasks, such as motion detection, are still not able to compete with the state-of-the-art of dense computer vision solutions.
 
+\note{show the gap:}
 Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
@@
 %
-
+%
 \subsection{Outline}
 %%%-----------------------------------------------------------------
-
 In this work, we study the emergence of such spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motion_task}).
+The paper is organized as fellows.
 We develop a SNN-like method able to learn hetero-synaptic delays to perform motion detection on a synthetic event-based dataset.
+We will first details the  méthodologie by first defining the  core spiking mechanism using  hetero-synaptic delays.
+We will  then define formally the computational  task of notion detection.
+This  will allow us to formalize the  final spiking neuron used to learn  the model and test its efficiency.
+In the results section, we will  detail the emerging structure  of the Sun solving the task and  in particular show the similarities  with neurobiological observations.
+We  will also test the network with  classical stimuli used in neuroscience,  and notably bitrate the response  to frequencies or contrast.
+We will further study the efficiency  of the motion detection mechanism  and in particular how it is  resilient to the pruning of synaptic  Weights.
+This will allow to show  how such model may provide an efficient solution to the energy/  accuracy trade-off.
+Finally, we will conclude lay  highlighting major contributions  of this paper, while delineating  some limits.
+This will open up  to perspectives for future SNN.
 Because neuromorphic devices are, by design, good candidates to integrate computations with time, we highlight the fact that this event-driven algorithm is transferable to such hardware.

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.3

--- 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
@@
 %
-\subsection{What can neuroscience bring to computer vision?}% and the Hetero-synaptic delays in spiking neurons}
+\subsection{What can neuroscience bring to computer vision?}% and the hetero-synaptic delays in spiking neurons}
 %%%-----------------------------------------------------------------
@@
 %
-
+%
 \subsection{Outline}
 %%%-----------------------------------------------------------------
-
-In this work, we study the emergence of such spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motion_task}).
-We develop a SNN-like method able to learn hetero-synaptic delays to perform motion detection on a synthetic event-based dataset.
-Because neuromorphic devices are, by design, good candidates to integrate computations with time, we highlight the fact that this event-driven algorithm is transferable to such hardware.
+In this work, we study the emergence of such spatiotemporal spike patterns when training a single layer of spike neurons on a supervised classification task (\seeFig{motion_task}).
+The paper is organized as follows.
+We develop an SNN-like method capable of learning hetero-synaptic delays to perform motion detection on an event-based synthetic dataset.
+We will first detail the methodology by defining the core spike mechanism using hetero-synaptic delays.
+We will then formally define the computational task of motion detection.
+This will allow us to formalize the final spiking neuron used to learn the model and test its effectiveness.
+In the results section, we will detail the emergent structure of the model solving the task and show in particular the similarities with neurobiological observations.
+We will also test the network with classical stimuli used in neuroscience, and in particular the frequency or contrast responses.
+We will also study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.
+This will allow us to show how such a model can provide an effective solution to the energy/accuracy trade-off.
+Finally, we will conclude by highlighting the main contributions of this paper, while defining some limitations.
+This will open up perspectives for future SNNs.
+As neuromorphic devices are, by design, good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is transferable to this type of hardware.
 %
@@
 Let us highlight some innovations in the contributions presented in this paper.
-First, the generic heterosynaptic model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs, whereas~\citet{ghosh_spatiotemporal_2019} use a correlation-based heuristic, which we have proven to be less efficient.
+First, the generic hetero-synaptic model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs, whereas~\citet{ghosh_spatiotemporal_2019} use a correlation-based heuristic, which we have proven to be less efficient.
 Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking motif.

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction

@@
 If we take a step back, we will remember that these tools are inherited from theoretical models designed to understand the brain.
-While the first artificial neural networks were mainly applied to " toy " type problems, they also offered a growing number of applications in computer vision, but also more generally for task automation.
+While the first artificial neural networks were mainly applied to toy-like problems, they also offered a growing number of applications in computer vision, but also more generally for task automation.
 In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates, and today they reach higher performances than humans for image categorization.
@@
 \note{show the gap:}
+This paper explores the possibility that spiking motifs may be a useful representation used in neurobiology and that we may transsfer this to neuromorphic algorithms.
 Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
@@
 %%%-----------------------------------------------------------------
-In this work, we study the emergence of such spatiotemporal spike patterns when training a single layer of spike neurons on a supervised classification task (\seeFig{motion_task}).
+In this work, we study the emergence of such spatiotemporal spike motifs when training a single layer of spike neurons on a supervised classification task (\seeFig{motion_task}).
 The paper is organized as follows.

### 2022-09-28 Grimaldi-etal-BiolCybernetics.tex
Introduction
Update on Overleaf.

@@
 It is not yet known why most biological systems have evolved to use this representation, but one major advantage seems to be some form of efficiency.
-In the human brain, each of the approximately ten trillion nerve cells emits on average one spike per second, but each neuron fires with great variability, and up to 300 spikes per second.
+In the human brain, each of the approximately 90 trillion nerve cells emits on average one spike per second to an average of 10000 synapses, but each neuron fires with great variability, and up to 300 spikes per second.
 Many attempts have been made to build artificial spiking neural networks, and in particular to achieve the same performance as observed in classical CNNs, but with a lower energy budget.
@@
 
+normarive models
+
 These computational results suggest that the temporal representation underlying spiking representation may be a key ingredient in the efficiency of the algorithms, but one may wonder how this can be supported by neurobiological data.
@@
 
-\note{show the gap:}
 This paper explores the possibility that spiking motifs may be a useful representation used in neurobiology and that we may transsfer this to neuromorphic algorithms.

### 2022-09-28 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 \newcommand{\Title}{
-Learning hetero-synaptic delays in a layer of spiking neurons for fast motion detection
+Learning heterogeneous delays in a layer of spiking neurons for fast motion detection
 }
@@
 However, most artificial neuronal models do not take advantage of this minute temporal dimension.
-Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of neurons with hetero-synaptic delays.
+Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of neurons with heterogeneous delays.
 Indeed, the variety of synaptic delays on the dendritic tree allows to synchronize synaptic inputs as they reach the basal dendritic tree.
@@
 %
-\subsection{What can neuroscience bring to computer vision?}% and the hetero-synaptic delays in spiking neurons}
+\subsection{What can neuroscience bring to computer vision?}% and the heterogeneous delays in spiking neurons}
 %%%-----------------------------------------------------------------
@@
 Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
-Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, hetero-synaptic delays.
+Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, heterogeneous delays.
 In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay in the computational process.
@@
 The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier.
-In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these hetero-synaptic delays.
+In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these heterogeneous delays.
 %
@@
 The paper is organized as follows.
-We develop an SNN-like method capable of learning hetero-synaptic delays to perform motion detection on an event-based synthetic dataset.
-We will first detail the methodology by defining the core spike mechanism using hetero-synaptic delays.
+We develop an SNN-like method capable of learning heterogeneous delays to perform motion detection on an event-based synthetic dataset.
+We will first detail the methodology by defining the core spike mechanism using heterogeneous delays.
 We will then formally define the computational task of motion detection.
@@
 %%%-----------------------------------------------------------------
-Let us now formally define the hetero-synaptic model, as well as the task and how we will learn to solve it.
-First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neural network (SNN) with hetero-synaptic delays.
+Let us now formally define the heterogeneous model, as well as the task and how we will learn to solve it.
+First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neural network (SNN) with heterogeneous delays.
 Our SNN model will be tested on a computationally relevant task of motion detection in an event-based sensory input.
@@
 %
-\subsection{Detecting event-based motifs using spiking neurons with hetero-synaptic delays}% model as a Logistic Regression with a temporal convolution}
+\subsection{Detecting event-based motifs using spiking neurons with heterogeneous delays}% model as a Logistic Regression with a temporal convolution}
 %%%-----------------------------------------------------------------
@@
     \includegraphics[width=0.490\linewidth]{figures/figure_1a.pdf}
-	    \caption{Detecting event-based motifs using spiking neurons with hetero-synaptic delays. 
+	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
 	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs).
@@
 A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree, the so-called weight matrix.
-A more detailed description of these synaptic contacts reveals the added importance of hetero-synaptic delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}.
-In that hetero-synaptic description, we can parameterize each neuron by the set of tuples defining both the weight and the delay of each synaptic contact.
+A more detailed description of these synaptic contacts reveals the added importance of heterogeneous delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}.
+In that heterogeneous description, we can parameterize each neuron by the set of tuples defining both the weight and the delay of each synaptic contact.
 As a consequence, a set of input presynaptic spikes $\event$ will be processed by the dendrite defined by this synaptic set and notably by the respective delays, which will multiplex in time all events.
 
-Formally, let's define a layer of hetero-synaptic spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+Formally, let's define a layer of heterogeneous spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
 In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
@@
 For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
-This defines the full connectivity of the hetero-synaptic model.
+This defines the full connectivity of the heterogeneous model.
 As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic address space is transformed into a new event stream $\event_\postsynaddrspace$ which is defined by
@@
 As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
-In this simplified model, we will consider that hetero-synaptic delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
+In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
 Using this dense representation, the counting defined above becomes:
@@
 
-\subsubsection{Implementation using spiking neurons with hetero-synaptic delays}
+\subsubsection{Implementation using spiking neurons with heterogeneous delays}
 %
@@
 In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
-The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the hetero-synaptic delays model.
+The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the heterogeneous delays model.
 %
@@
     \includegraphics[width=0.490\linewidth]{figures/figure_N_PG_time.pdf}
-	    \caption{Detecting event-based motifs using spiking neurons with hetero-synaptic delays. 
+	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
 	    {\bf (a)}~Accuracy of PG detection as a function of the number $M$ of kernels.
@@
 We show in Figure~\ref{fig:model_results}~(a) that the accuracy of finding the right PG is still above $80\%$ accuracy with more than $1364$ overlapping PGs.
-Moreover, we show in Figure~\ref{fig:model_results}~(b) that (with $M=55$ PGs fixed) the accuracy increases notably as the temporal depth $D$ of the PG kernel increased, demonstrating quantitatively the potential of hetero-synaptic delays.
+Moreover, we show in Figure~\ref{fig:model_results}~(b) that (with $M=55$ PGs fixed) the accuracy increases notably as the temporal depth $D$ of the PG kernel increased, demonstrating quantitatively the potential of heterogeneous delays.
 These results were obtained while assuming that we know $W$.
@@
 
-Once our MLR is trained, we obtain 3D kernels corresponding to the weights associated to the hetero-synaptic delays of our layer of spiking neurons and which may be used for detection.
+Once our MLR is trained, we obtain 3D kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection.
 We observed that the distribution of the kernels' weights is sparse, with most values near zero.
@@
 In Figure~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations, or active weights, per decision for each neuron of the layer.
-As a comparison and to account for the gain in performance by using hetero-synaptic delays, we provide the accuracy obtained with a MLR model using 2D time surface (in red) as in~\citep{grimaldi_robust_2022}.
+As a comparison and to account for the gain in performance by using heterogeneous delays, we provide the accuracy obtained with a MLR model using 2D time surface (in red) as in~\citep{grimaldi_robust_2022}.
 This latter method is based on delays from the last recorded events and uses fewer computations (in our case $15\times15$) than the dense 3D kernels without any pruning ($15\times15\times8$).
@@
     \includegraphics[width=0.95\linewidth]{figures/accuracy.pdf}
-    \caption{Accuracy as a function of the number of computation load for the hetero-synaptic delays model (in blue) and for a method using 2D time surfaces (in red).
+    \caption{Accuracy as a function of the number of computation load for the heterogeneous delays model (in blue) and for a method using 2D time surfaces (in red).
 The relative computational load (on a log axis) is controlled by changing the percentage of active weights relative to the dense convolution kernel.
@@
 %%%-----------------------------------------------------------------
-Here, we have introduced a generic SNN using hetero-synaptic delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification \note{to change if we compare to ghosh instead of HOTs}.
+Here, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification \note{to change if we compare to ghosh instead of HOTs}.
 The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification.
@@
 %%%-----------------------------------------------------------------
-To synthesize, we have introduced in this paper a spiking neural network model with hetero-synaptic delays that we have trained and evaluated on a complex motion detection task.
+To synthesize, we have introduced in this paper a spiking neural network model with heterogeneous delays that we have trained and evaluated on a complex motion detection task.
 The model was defined to optimally detect event-driven spatiotemporal motifs.
@@
 We have then shown that the model displays similarities with the responses observed in biology.
-We have evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of hetero-synaptic delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understand why spikes represent a universal component of neuronal information processing.
+We have evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of heterogeneous delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understand why spikes represent a universal component of neuronal information processing.
 
 Let us highlight some innovations in the contributions presented in this paper.
-First, the generic hetero-synaptic model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs, whereas~\citet{ghosh_spatiotemporal_2019} use a correlation-based heuristic, which we have proven to be less efficient.
+First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs, whereas~\citet{ghosh_spatiotemporal_2019} use a correlation-based heuristic, which we have proven to be less efficient.
 Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking motif.
@@
 \note{Say that this could be implemented by properly setting the timing of the label} As previously stated, this could be implemented using generalized coordinates (that is variables such as position which are complemented by speed, acceleration, jerk, ...) and ``neurobiologically, the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
-Our proposed hetero-synaptic model provides an alternate and elegant implementation solution to this problem.
+Our proposed heterogeneous model provides an alternate and elegant implementation solution to this problem.
 %

### 2022-09-28 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 The challenges are numerous in such situations and two major obstacles must be highlighted: the computational delay and the energy budget.
-First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST (Le Cun, 98), to high definition images, such as VGG on Imagenet (VGG16).
+First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun1998gradient}, to high definition images, such as VGG on Imagenet~\citep{Simonyan:2014cmh}.
 This has a cost and notably the fact that the computation time increases with the size of the image.
@@
 
+% Schrimpf2020 brainScores does not count energy
 Neuroscience could provide revolutionary answers to these two closely related problems.
@@
 %
-\subsection{What can neuroscience bring to computer vision?}% and the heterogeneous delays in spiking neurons}
+\subsection{What can neuroscience bring to computer vision?}% and the heterogeneous delays in spiking neurons} ~\citep{}
 %%%-----------------------------------------------------------------
@@
 Larger scale models introduce an explicit spiking mechanism, for example in the Integrate-and-Fire model.
-It predicts the emergent states of a recurrently connected neural network (Brunel).
-Several models have increased the scale of the system to include more neurons, such as the Virtual Brain (Sanz), which mixes spike neural networks with analog neural mass models.
+It predicts the emergent states of a recurrently connected neural network~\citep{Brunel2000}.
+Several models have increased the scale of the system to include more neurons, such as the Virtual Brain~\citep{SanzLeon2013}, which mixes spike neural networks with analog neural mass models.
 However, these methods are mainly descriptive with applications in neuroscience but few have had an application in computer vision.
 Another class of approaches have designed spiking neural networks that aim to have applications in machine learning.
-One of these is the SpikeNet algorithm which uses a radical approach by encoding information using one spike per neuron (Delorme).
+One of these is the SpikeNet algorithm which uses a radical approach by encoding information using one spike per neuron~\citep{Delorme1999}.
 Another type of spiking neural network attempts to determine the structure of the network in order to minimize a cost function.
-This was implemented in the SpikeProp algorithm (Bohte) and is now widely used in methods that attempt to transfer performance from CNNs to SNNs, for example using a surrogate gradient approach.
+This was implemented in the SpikeProp algorithm~\citep{Bohte2002} and is now widely used in methods that attempt to transfer performance from CNNs to SNNs, for example using a surrogate gradient approach.
 However, the performance of SNNs is still lower than that of analog networks and the question of the advantage of using spikes remains open.
@@
 Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.
-A series of similar algorithms are applied to real-life applications such as optical flow (Benosman), depth estimation (Dardelet), or gesture recognition (Marro).
+A series of similar algorithms are applied to real-life applications such as optical flow~\citep{Benosman2012}, depth estimation~\citep{Dardelet2021}, or gesture recognition~\citep{Maro2020}.
 Recently, such an algorithm has been used in the space industry which is famous for its stringent selection process.
 
-normarive models
+% normative models
 

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Larger scale models introduce an explicit spiking mechanism, for example in the Integrate-and-Fire model.
-It predicts the emergent states of a recurrently connected neural network~\citep{Brunel2000}.
-Several models have increased the scale of the system to include more neurons, such as the Virtual Brain~\citep{SanzLeon2013}, which mixes spike neural networks with analog neural mass models.
+It predicts the emergent states of a recurrently connected neural network~\citep{brunel_phase_2000}.
+Several models have increased the scale of the system to include more neurons, such as the Virtual Brain~\citep{sanz_leon_virtual_2013}, which mixes spike neural networks with analog neural mass models.
 However, these methods are mainly descriptive with applications in neuroscience but few have had an application in computer vision.
 Another class of approaches have designed spiking neural networks that aim to have applications in machine learning.
-One of these is the SpikeNet algorithm which uses a radical approach by encoding information using one spike per neuron~\citep{Delorme1999}.
+One of these is the SpikeNet algorithm which uses a radical approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
 Another type of spiking neural network attempts to determine the structure of the network in order to minimize a cost function.
-This was implemented in the SpikeProp algorithm~\citep{Bohte2002} and is now widely used in methods that attempt to transfer performance from CNNs to SNNs, for example using a surrogate gradient approach.
+This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and is now widely used in methods that attempt to transfer performance from CNNs to SNNs, for example using a surrogate gradient approach.
 However, the performance of SNNs is still lower than that of analog networks and the question of the advantage of using spikes remains open.
@@
 Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.
-A series of similar algorithms are applied to real-life applications such as optical flow~\citep{Benosman2012}, depth estimation~\citep{Dardelet2021}, or gesture recognition~\citep{Maro2020}.
+A series of similar algorithms are applied to real-life applications such as optical flow~\citep{benosman_asynchronous_2012}, depth estimation~\citep{dardelet_event-by-event_2021}, or gesture recognition~\citep{maro_event-based_2020}.
 Recently, such an algorithm has been used in the space industry which is famous for its stringent selection process.
@@
 	    {\bf (a)}~Accuracy of PG detection as a function of the number $M$ of kernels.
-	    {\bf (b)}~Accuracy of PG detection as a function of the temporal depth $D$ of kernels.
+	    {\bf (b)}~Accuracy of PG detection as a function of the temporal depth $D$ of kernels among $M=144$ kernels.
 	    	     }
@@
 Now that our motion detection task is defined in visual space, we can extend the convolution-based detection model to a 3D convolution so that the resulting model would also benefit from spatial invariance.
-The use of spatiotemporal filters on an event stream has improved the performance of CNNs for an action recognition task in~\cite{ghosh_spatiotemporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.
+The use of spatiotemporal filters on an event stream has improved the performance of CNNs for an action recognition task by~\citet{ghosh_spatiotemporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.
 For this purpose, we design 3D kernels of shape $(\Kx, \Ky, \Ktime) = (9, 9, 11)$, representing the two spatial dimensions and the range of time delays, respectively.

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
introduction

@@
 %%%-----------------------------------------------------------------
-Machine learning, and computer vision in particular, have undergone a major transformation in modern times, leaving science fiction movies to become mainstream tools found in the smartphones in everyone's pocket.
-Less visible, but even more crucial to our economy and security, are the many automated tools that now dominate the range of services that equip virtually all sectors of our daily lives.
-This trend is most pronounced in the field of medicine, as it provides practitioners with new diagnostic tools to treat people faster and more effectively.
+Machine learning, and computer vision in particular, have undergone a major transformation in modern times, transitioning from an active actor in science fiction movies to become mainstream tools found in the smartphones in everyone's pocket.
+Less visible, but even more crucial to our economy and security, these automated tools now dominate the range of services that equip virtually most sectors of our daily lives.
+This trend is profoundly pronounced in the field of medicine, as it provides practitioners with new diagnostic tools to treat people faster and more effectively.
 This particular example also illustrates some important challenges that these tools are increasingly facing.
 One of these is their sensitivity to adversarial attacks, that is, the sensitivity of the models' output to minute changes in the input.
-This could raise important questions about the reliability of the results of in any automated diagnostic model.
+This could raise important questions about the reliability of the results of any automated diagnostic model.
 More generally, these models do not explain why they produce a given output for a class of inputs.
 This lack of explainability is a hindrance to the advancement of these methodologies as well as to their acceptance for most safety-relevant tasks.
-If we take a step back, we will remember that these tools are inherited from theoretical models designed to understand the brain.
-While the first artificial neural networks were mainly applied to toy-like problems, they also offered a growing number of applications in computer vision, but also more generally for task automation.
-In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates, and today they reach higher performances than humans for image categorization.
+Having said that, we may remember that these tools are inherited from theoretical models originally designed to understand the brain.
+While the first artificial neural networks were mainly applied to toy-like problems, they also offered a growing number of applications in machine learning and computer vision, but also more generally for task automation.
+In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates, and today they reach higher performances than humans for key image categorization challenges.
 One of the challenges of these models is to be able to scale them up for future applications, for example in autonomous driving.
@@
 The challenges are numerous in such situations and two major obstacles must be highlighted: the computational delay and the energy budget.
-First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun1998gradient}, to high definition images, such as VGG on Imagenet~\citep{Simonyan:2014cmh}.
+First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun1998gradient}, then high definition images, such as VGG on Imagenet~\citep{Simonyan:2014cmh}.
 This has a cost and notably the fact that the computation time increases with the size of the image.

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Introduction

@@
 This phenomenon has been mitigated by the increase in available computing power, notably thanks to the parallel processing implemented in GPU cards.
-However, the transition from images to videos, then to multiple cameras implies a further change of scale in the amount of data processed.
+However, the transition from images to videos, then to multiple cameras implies a further level in the change of scale in the amount of data processed.
 It also raises an issue that is often not addressed when it comes to image categorization.
-Indeed, in autonomous driving, an important task is to react as quickly as possible to a possible collision, for example when a pedestrian crosses the street.
+Indeed, in autonomous driving, an important task is to react as quickly as possible to a possible collision, for example when a pedestrian unexpectedly crosses the street.
 As such, future systems should always be able to give a response at any time, instead of having to wait for the pipeline to provide the correct response.
 This online processing is also a feature of processing in biological systems and could be an important first contribution if we could translate these principles into future automated processes.
+%
 Second, a major constraint for future systems is energy cost.
@@
 In addition, the need to solve increasingly complex tasks has required the construction of increasingly complex and deep architectures.
-As an example, the best performing model in the Imagenet challenge uses billions of parameters that typically cost 1Wh to train and about 1Wh to infer a single image.
-Less power-hungry solutions have emerged, such as Mobile Net, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be about 300W, or 20\% of the energy of an electric car under standard conditions.
+As an example, the best performing model in the Imagenet challenge uses 2100 million parameters.
+Training the model typically cost over 2500 GFLOPS and about 1 joule to infer a single image.
+Less power-hungry solutions have emerged, such as Mobile Net, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significative percentage of the energy consumed by an electric car under standard conditions.
 

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.1

@@
 
-% Schrimpf2020 brainScores does not count energy
 Neuroscience could provide revolutionary answers to these two closely related problems.
@@
 CNNs, for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels.
+These networks have been known to mimick many propoerties of biological systems~\citep{Schrimpf2020}, yet, this score does not count rapidity or energy.
 On the other hand, information in a large majority of biological neural networks is represented as spikes, i.e.
@@
 It is not yet known why most biological systems have evolved to use this representation, but one major advantage seems to be some form of efficiency.
-In the human brain, each of the approximately 90 trillion nerve cells emits on average one spike per second to an average of 10000 synapses, but each neuron fires with great variability, and up to 300 spikes per second.
-Many attempts have been made to build artificial spiking neural networks, and in particular to achieve the same performance as observed in classical CNNs, but with a lower energy budget.
+In the human brain, each of the approximately 90 trillion nerve cells emits on average one spike per second which fan out to an average of 10000 synapses.
+Importantly, each neuron fires with great variability, and up to 300 spikes per second.
+Many attempts have been made to build artificial spiking neural networks, and in particular to achieve the same performance as observed in classical CNNs with a lower energy budget.
 However, none of these networks has yet surpassed the state of the art performances.
-Here, we will try to focus on a property of spike neurons inspired by neurobiology to progress in this direction.
+Here, we will try to focus on one core property of spiking neurons inspired by neurobiology to progress in this direction.
 %
@@
 %%%-----------------------------------------------------------------
-In this work, we study the emergence of such spatiotemporal spike motifs when training a single layer of spike neurons on a supervised classification task (\seeFig{motion_task}).
+In this work, we study the emergence of such spatiotemporal spike motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motion_task}).
 The paper is organized as follows.

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.2

@@
 %%%-----------------------------------------------------------------
-Let us first focus on some existing spiking neural network models.
-A first motivation for these models was to understand biological observations at different scales.
-For example, the Hodgkin-Huxley model describes the dynamics of the membrane potential and the role of different voltage-dependent ion channels.
+Let us first review some existing spiking neural network models.
+A first motivation for these models was to understand biological observations at the different scales of the central nervous system.
+For example, the Hodgkin-Huxley model describes the dynamics of the membrane potential of a single nerve cell and specifically explains the role of the different voltage-dependent ion channels.
 It also explains how the model can produce a spike.
-Larger scale models introduce an explicit spiking mechanism, for example in the Integrate-and-Fire model.
+Larger scale models introduce an simpler dynamics and an explicit spiking mechanism, for example in the Integrate-and-Fire model.
 It predicts the emergent states of a recurrently connected neural network~\citep{brunel_phase_2000}.
-Several models have increased the scale of the system to include more neurons, such as the Virtual Brain~\citep{sanz_leon_virtual_2013}, which mixes spike neural networks with analog neural mass models.
-However, these methods are mainly descriptive with applications in neuroscience but few have had an application in computer vision.
-Another class of approaches have designed spiking neural networks that aim to have applications in machine learning.
+Several models have increased the scale of the system to include more neurons, such as The Virtual Brain (TVB)~\citep{sanz_leon_virtual_2013}, which mixes different scales and includes spiking neuron networks with (analog) neural mass models.
+However, these methods are mainly descriptive with applications in neuroscience but few spiking neurall models have had an application in computer vision.
+Another class of approaches have designed normative models of spiking neuron networks that aim to have applications in machine learning.
 One of these is the SpikeNet algorithm which uses a radical approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
-Another type of spiking neural network attempts to determine the structure of the network in order to minimize a cost function.
-This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and is now widely used in methods that attempt to transfer performance from CNNs to SNNs, for example using a surrogate gradient approach.
+Another type of spiking neural network using precise spike timing attempts to determine the structure of the network in order to minimize a cost function.
+This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extanded in novel gradient-based methods.
+Surrogate gradient for instance is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke}.
 However, the performance of SNNs is still lower than that of analog networks and the question of the advantage of using spikes remains open.
@@
 Recently, such an algorithm has been used in the space industry which is famous for its stringent selection process.
-
-% normative models
 

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
Introduction: 1.2

@@
 
-A new generation of algorithms has emerged with the advent of the so-called silicon retina, which mimics the output of the retina by representing information using spikes.
+A new generation of algorithms has emerged with the advent of so-called silicon retinas, which mimics the output of the retina by representing visual information using spikes.
 %A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance.
 The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations.
-In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}.
-This HOTS model represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events.
-By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is in an online fashion.
-We have demonstrated on several datasets that it provides with efficient computations resulting in ultra-fast categorization.
+The HOTS model~\citet{lagorce_hots_2017} represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events.
+In a recent study, we have introduced a spiking neuron network as a classification model applied to event-based streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model.
+By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is, in an online fashion.
+We have demonstrated on several benchmark datasets that it provides with efficient computations resulting in ultra-fast categorization.
 Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.
 A series of similar algorithms are applied to real-life applications such as optical flow~\citep{benosman_asynchronous_2012}, depth estimation~\citep{dardelet_event-by-event_2021}, or gesture recognition~\citep{maro_event-based_2020}.
-Recently, such an algorithm has been used in the space industry which is famous for its stringent selection process.
+Recently, such an algorithm was used in the CASPR mission on the ISS, illustrating, given the stringent selection processes used by the space sector, that such algorithms have reached maturity for industry.
 
-These computational results suggest that the temporal representation underlying spiking representation may be a key ingredient in the efficiency of the algorithms, but one may wonder how this can be supported by neurobiological data.
+These computational results suggest that the temporal representation underlying processing in spiking neurons may be a key ingredient in the efficiency of the algorithms, but one may wonder how this can be supported by neurobiological data.
 In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking motifs~\citep{abeles_role_1982}.
@@
 Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
-All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal motifs in biological spike trains.
+All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
 In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.3

@@
 This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extanded in novel gradient-based methods.
-Surrogate gradient for instance is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke}.
+Surrogate gradient for instance is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}.
 However, the performance of SNNs is still lower than that of analog networks and the question of the advantage of using spikes remains open.
@@
 
-This paper explores the possibility that spiking motifs may be a useful representation used in neurobiology and that we may transsfer this to neuromorphic algorithms.
+This paper explores the possibility that prototypical spiking motifs may be a useful representation used in neurobiology and that we may transfer this hypothesis to neuromorphic algorithms.
 Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
@@
 The paper is organized as follows.
-We develop an SNN-like method capable of learning heterogeneous delays to perform motion detection on an event-based synthetic dataset.
+We develop a theroretically-defined SNN capable of learning heterogeneous delays which we first evaluate on an event-based synthetic dataset.
 We will first detail the methodology by defining the core spike mechanism using heterogeneous delays.
 We will then formally define the computational task of motion detection.
-This will allow us to formalize the final spiking neuron used to learn the model and test its effectiveness.
+This will allow us to formalize the  spiking neuron used to learn the model and test its effectiveness.
+We aim at applying this model to a real-world task of motion detection which we will first define.
+Then we will show how the model can be adapted to solve the motion detection task.
 In the results section, we will detail the emergent structure of the model solving the task and show in particular the similarities with neurobiological observations.
@@
 This will open up perspectives for future SNNs.
-As neuromorphic devices are, by design, good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is transferable to this type of hardware.
+In particular, as neuromorphic devices are by design good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is transferable to this type of hardware.
 %

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.3

@@
 %%%-----------------------------------------------------------------
-In this work, we study the emergence of such spatiotemporal spike motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motion_task}).
+In this work, we study the emergence of spatio-temporal spiking motifs when training a single layer of spiking neurons on a semi-supervised classification task (\seeFig{motion_task}).
 The paper is organized as follows.
-We develop a theroretically-defined SNN capable of learning heterogeneous delays which we first evaluate on an event-based synthetic dataset.
-We will first detail the methodology by defining the core spike mechanism using heterogeneous delays.
-We will then formally define the computational task of motion detection.
-This will allow us to formalize the  spiking neuron used to learn the model and test its effectiveness.
-We aim at applying this model to a real-world task of motion detection which we will first define.
-Then we will show how the model can be adapted to solve the motion detection task.
+We develop a theoretically defined SNN capable of learning from heterogeneous delays that we first evaluate on a synthetic event-based dataset.
+We first detail the methodology by defining the basic mechanism of spiking neurons that utilize heterogeneous delays.
+This will allow us to formalize the spike neuron used to learn the model's parameters and test its effectiveness.
+We aim to apply this model to a realistic motion detection task which we will first define.
+Then, we will show how the model can be adapted to solve the motion detection task.
 In the results section, we will detail the emergent structure of the model solving the task and show in particular the similarities with neurobiological observations.
-We will also test the network with classical stimuli used in neuroscience, and in particular the frequency or contrast responses.
+We will also test the network with classical stimuli used in neuroscience, and in particular with frequency or contrast responses.
 We will also study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.
-This will allow us to show how such a model can provide an effective solution to the energy/accuracy trade-off.
+This will allow us to show how such a model can provide an efficient solution to the energy/accuracy trade-off.
 Finally, we will conclude by highlighting the main contributions of this paper, while defining some limitations.
-This will open up perspectives for future SNNs.
+This will open perspectives for future SNNs.
 In particular, as neuromorphic devices are by design good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is transferable to this type of hardware.

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
supervised >> semi-supervised

@@
 Solving is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina.
-The task is thus to guess the motion as accurately as possible (see \seeFig{motion_task}) and we will describe a supervised learning mechanism for weights and delays.
+The task is thus to guess the motion as accurately as possible (see \seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays.
 To this end, we will extend the SNN model to be efficient for the resolution of the task by including convolutional kernels.
@@
 %
-Since the model is fully differentiable, we can now implement a supervised learning rule.
+Since the model is fully differentiable, we can now implement a semi-supervised learning rule.
 The loss function of the MLR model is the binary cross-entropy on the output of the classification layer and can be extended to space-time convolutions.
@@
 The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions.
+As in this semi-supervised context, the label is lknown but the timing is not, we used a selection of the temporal suppport in an unsuperviesd manner.
 It is important to note that we weighted the cost function by considering only active cells, such that the error is only backpropagated to the spatial locations of these most active cells.
+This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier2007}.
 Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$).

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Results: 3.1

@@
     \caption{
-    	Representation of the weights for the $8$ learned kernels of the model corresponding to the OFF polarities and selective to the different motion directions
-	(because of the symmetry observed between the ON and OFF event streams, kernels are similar for the ON polarities).
+    	Representation of the weights for $3$ of the $32$ different learned kernels of the model.
+Each corresponding to the OFF polarities and selective to the different motion directions (because of the symmetry observed between the ON and OFF event streams, we observed that kernels are very similar for the ON polarities).
 These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
-	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.
-One sees positive (excitatory) coefficients for the specific direction of motion and negative (inhibitory) coefficients for all other directions.
+%	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.
+        One sees positive (excitatory) coefficients for the specific direction of motion but also negative (inhibitory) coefficients for other directions.
 	}
@@
 Focusing on the positive weights, a strong selectivity is observed along specific axes for the different kernels.
-These directions can be easily associated to the direction of motion controlled in the motion clouds.
-For instance, the third and the seventh kernels show a horizontal selectivity to motion directions.
+These directions can be easily associated to the direction of motion controlled in the natural images.
+For instance, the first kernel shows a selectivity to horizontal motion directions.
 %
-With the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
+We observed a high dependendance between the weights reaching the ON polarities and that reaching the OFF polarities.
+ith the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
 This qualitative look at the 3D kernels allows the reader to infer for the $8$ different motion directions used to generate our synthetic event streams.

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Results: 3.1

@@
 The resulting instances yield trajectories qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(a)).
+A similar procedure of animating of natural scene by virtual eye-movements was previously in neurobiliological~\citep{Baudot2013} and computational neuroscience~\citet{Kremkow2016} studies.
 
@@
 %
-After training our model, we first observe the weights learned for the different neurons (\seeFig{kernels}).
-Focusing on the positive weights, a strong selectivity is observed along specific axes for the different kernels.
+After training our model, we first analyze the weights learned for the different neurons (\seeFig{kernels}). 
+We first observed a high dependendance between the weights reaching the ON polarities and that reaching the OFF polarities.
+In particular, whenever a weight for a given position and delay is positive for one poalrity, it will be negative in the other.
+This property comes from the way the events are generated and that the luminance can not at the same time increase and decrease.
+%ith the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
+We also observe that these cells show an orientation selectivity, similar to that observe in MT neurons~\citep{DeAngelis1999}.
+Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{Kremkow2016} to explain neurophysiological results~\citep{Baudot2013}.
+Focusing on the positive weights, a strong selectivity is observed along specific axes of motion for each of the different kernels.
 These directions can be easily associated to the direction of motion controlled in the natural images.
-For instance, the first kernel shows a selectivity to horizontal motion directions.
+For instance, the first kernel shows a strong selectivity to horizontal motion directions.
 %
-We observed a high dependendance between the weights reaching the ON polarities and that reaching the OFF polarities.
-ith the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
-This qualitative look at the 3D kernels allows the reader to infer for the $8$ different motion directions used to generate our synthetic event streams.
+%This qualitative look at the 3D kernels allows the reader to infer for the $8$ different motion directions used to generate our synthetic event streams.
 
@@
 say that we have tested }
-\note{tester la meme expe faite par baudot et modelisée par kremkow}
+\note{tester la meme expe faite par baudot et modelisée par kremkow en montrant des raster plots du modele avec NI versus Gratings }
 

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Results: 3.2

@@
 
+%%%-----------------------------------------------------------------
+%
+\subsection{Testing with natural-like textures}
+%%%-----------------------------------------------------------------
+To test our model, we will quantify its ability to categorize different motions.
+Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli.
+In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds})~\citep{Vacher2018}.
+In particular, we will set the spatial size and duration similarly to the motion task defined above.
+This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions $
+\speed_\kernelind = 
+  ( 
+    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
+    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
+  )
+$.
+For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content to provide with some naturalistic variability.
+This method provides a rich dataset of textured movies for which we know the ground truth for motion.
+
+%%%-----------------------------------------------------------------
+%: fig:motion_clouds
+%%%-----------------------------------------------------------------
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.95\linewidth]{figures/motion_clouds.pdf}
+    \caption{{\bf Role of image parameters in the motion detection accuracy.} Accuracy as a function of {\bf (a)} the mean spatial frequency, {\bf (b)} the bandwidth in orientation: from a grating-like (left) to isotropic textures (right)).
+{\bf (c)} the bandwidth in speed, from a rigid motion (left) to independent frames (right).
+Note that these accuracy is computed both in the case is necessarily perpendiular to the motion (aperture) or aribtrary (no aperture), showing that accuracy decreases in the latter case.}
+    \label{fig:motion_clouds}
+\end{figure}
+%%%-----------------------------------------------------------------
+
+We plot here main axis of interests.
+First, as we change the mean spatial frequency of the texture, we observe a montonous decrease in accuracy.
+This comes as a similar trend as that observed in the primary visual areas~\citep{Priebe2006}.
+Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively ressemble a more textured stimuls) than for a grating-like stimulation, reminiscent to the behavioural response of humans' eye movemnts to such stimuli~\citep{Simoncini2012}.
+Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.
+When the stimulus is grating-like and that the orientation is arbitrary with respect to the direction of motion, the system is faced with the aperture problem and see a decrease of accuracy.
+This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion.
+Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter.
+This shows a progressive decrease in accuracy similar to that observed in the amplitude of humans' eye movements~\citep{MansourPour2018}.
 %
@@
 
-Once our MLR is trained, we obtain 3D kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection.
+Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection.
 We observed that the distribution of the kernels' weights is sparse, with most values near zero.
@@
 \end{figure}
-%%%-----------------------------------------------------------------
-%
-\subsection{Testing with natural-like textures}
-%%%-----------------------------------------------------------------
-To test our model, we will quantify its ability to categorize different motions.
-In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}).
-In particular, we will set the spatial size to $(\Nx, \Ny)$ and consider a discretization of time with a time step of $1$ such that $\timev \in \mathbb{N}$.
-Movies' duration are here set to $\Ntime=400$.
-This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions:
-% of a motion cloud is determined by a constant $k_v$: with a constant speed $\speed$ (see \fig{motionclouds}-(a)): 
-%\begin{equation*}
-%%\begin{gathered}
-%\va{v}_k = %\binom{\cos(\frac{2\pi\cdot k_v}{N_v})}{\sin(\frac{2\pi\cdot k_v}{N_v})}
-%  \begin{pmatrix}
-%    \cos(2\pi\cdot \frac{k}{\Nclass})\\
-%    \sin(2\pi\cdot \frac{k}{\Nclass})
-%  \end{pmatrix}
-%%\end{gathered}
-%\end{equation*}
-$
-\speed_\kernelind = 
-  ( 
-    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
-    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
-  )
-$ (\seeFig{motionclouds}-(a) for an illustration).
-For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content.
-This method provides a rich dataset of textured movies for which we know the ground truth for motion.
-%
-%%%-----------------------------------------------------------------
-%: fig:motion_clouds
-%%%-----------------------------------------------------------------
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.95\linewidth]{figures/motion_clouds.pdf}
-    \caption{{\bf Role of image parameters in the motion detection accuracy.} {\bf (a)} The motion direction represented as the plain red vector, other possible motion directions are represented in light red.
-{\bf (b)} A screenshot of one generated naturalistic textured stimulus at a specific time.
-% $\timev_\arank$.
-    {\bf (c)} The corresponding ON (in red) and OFF (in blue) event stream generated from the stimuli on (b) and constituting the input to the spiking neural network.}
-    \label{fig:motion_clouds}
-\end{figure}
-%%%-----------------------------------------------------------------
 
-To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames.
-If, on a specific pixel at that specific timestamp, the absolute value of this gradient exceeds a threshold, an event is generated.
-The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive.
-This signed threshold value is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is threfore similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motionclouds}-(c)).
-The goal here is to infer the correct motion solely by observing these events.
 %

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Results: 3.3

@@
 As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the number of spikes times the number of synapses.
-This scales with the number of nonzero synaptic weights.
+Indeed, the computations are dominated by the convolution operation.
+In a dense setting, this corresponds for all voxels in the ouput to a sum over all voxels in the inputs for all weights in the kernel.
+If the support of information is sparse, then computations can be performed only on those events.
+Also, if we set some weights of the kernels to zero, then the sum can be skipped for thse adresses.
+Knowing the sparseness of the input, the total number of computations thus scales with the number of nonzero synaptic weights.
+
 To assess the robustness of the classification as a function of the computational load, we will prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold.
-
 In Figure~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations, or active weights, per decision for each neuron of the layer.
@@
 While less computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
-By pruning weights, we observe that the evolution of accuracy as a function of the log percentage of active weights follows a sigmoid curve.
-Half-saturation level is reached at about $3.5\times 10^{-3}\%$ of active weights, corresponding to a total amount of $6$ computations per decision.
+
+By pruning weights, we observe that the evolution of accuracy as a function of the log percentage of active weights fits well a sigmoid curve.
+Half-saturation level is reached at about $3.5\times 10^{-3}\%$ of active weights, corresponding in our setting to a total amount of $6$ computations per decision.
 Compared to the full kernels, the accuracy of our method is maintained to its top performances when dividing the number of computations by a factor up to about $200$.
@@
     \includegraphics[width=0.95\linewidth]{figures/accuracy.pdf}
-    \caption{Accuracy as a function of the number of computation load for the heterogeneous delays model (in blue) and for a method using 2D time surfaces (in red).
+    \caption{Accuracy as a function of the number of computation load for the heterogeneous delays model (blue line) and for a method using 2D time surfaces (red dot)~\citep{grimaldi_robust_2022}.
 The relative computational load (on a log axis) is controlled by changing the percentage of active weights relative to the dense convolution kernel.

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Results: 3.3

@@
 This comes as a similar trend as that observed in the primary visual areas~\citep{Priebe2006}.
-Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively ressemble a more textured stimuls) than for a grating-like stimulation, reminiscent to the behavioural response of humans' eye movemnts to such stimuli~\citep{Simoncini2012}.
+\note{we should really do something more serious here...} Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively ressemble a more textured stimuls) than for a grating-like stimulation, reminiscent to the behavioural response of humans' eye movemnts to such stimuli~\citep{Simoncini2012}.
 Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Discussion

@@
 It predicts the emergent states of a recurrently connected neural network~\citep{brunel_phase_2000}.
-Several models have increased the scale of the system to include more neurons, such as The Virtual Brain (TVB)~\citep{sanz_leon_virtual_2013}, which mixes different scales and includes spiking neuron networks with (analog) neural mass models.
+Several models have increased the scale of the system to include more neurons, such as The Virtual Brain (TVB)~\citep{sanz_leon_virtual_2013}, which mixes different scales and includes spiking neural networks with (analog) neural mass models.
 However, these methods are mainly descriptive with applications in neuroscience but few spiking neurall models have had an application in computer vision.
-Another class of approaches have designed normative models of spiking neuron networks that aim to have applications in machine learning.
+Another class of approaches have designed normative models of spiking neural networks that aim to have applications in machine learning.
 One of these is the SpikeNet algorithm which uses a radical approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
@@
 The HOTS model~\citet{lagorce_hots_2017} represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events.
-In a recent study, we have introduced a spiking neuron network as a classification model applied to event-based streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model.
+In a recent study, we have introduced a spiking neural network as a classification model applied to event-based streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model.
 By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is, in an online fashion.
@@
 %
-The activation function of our spiking neuron is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
+The activation function of our spiking neural is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
 It transforms this list of weights into a probability with the following formula:
@@
 %%%-----------------------------------------------------------------
-Here, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification \note{to change if we compare to ghosh instead of HOTs}.
-The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification.
+In this paper, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification.
+The learned model bears many similarities with neurobilogical anatomical observations but also with the results from behavioral results.
+\note{to change if we compare to ghosh instead of HOTs if we follow that route} The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification.
 This shows that we may use the precise timing of spikes to enhance neural computations.
@@
 %%%-----------------------------------------------------------------
-To synthesize, we have introduced in this paper a spiking neural network model with heterogeneous delays that we have trained and evaluated on a complex motion detection task.
+To recap, we have introduced in this paper a spiking neural network model with heterogeneous delays that we have trained and evaluated on a complex motion detection task.
 The model was defined to optimally detect event-driven spatiotemporal motifs.
@@
 Let us highlight some innovations in the contributions presented in this paper.
-First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs, whereas~\citet{ghosh_spatiotemporal_2019} use a correlation-based heuristic, which we have proven to be less efficient.
+First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs, whereas~\citet{ghosh_spatiotemporal_2019} use a correlation-based heuristic, which we have observed to be less efficient.
 Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking motif.
 Another novelty is that the model simultaneously learns the weights and the delays, while for example the polychronization model~\citep{izhikevich_polychronization_2006} only learns the weights using STDP while the delays are randomly drawn and their values frozen.
-Moreover, the model is evaluated on a realistic task, while models such as tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}.
-The main contribution is to provide a model that is suitable for any type of spatio-temporal spiking motifs and that can be trained in a supervised way by providing a dataset of supervision pairs.
+Moreover, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}.
+Another main contribution is to provide a model that is suitable for learning any type of spatio-temporal spiking motifs and that can be trained in a supervised way by providing a dataset of supervision pairs.
 This allows for a more flexible definition of the model using this properly labeled dataset instead of relying on a careful description of the physical rules governing a task, e.g., the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}.
@@
 %%%-----------------------------------------------------------------
-We have identified several limits to our model, the most important being now detailed.
+We have identified several limits to our model, the most important we will now detail.
 First, the complete framework is based on a discrete binning of time that is incompatible with the continuous nature of biological time.

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Indeed, the variety of synaptic delays on the dendritic tree allows to synchronize synaptic inputs as they reach the basal dendritic tree.
-We show this can be formalized as a time-invariant logistic regression which can be trained using labelled data.
+We show this can be formalized as a time-invariant logistic regression which can be trained using labeled data.
 We apply this model to solve the specific computer vision problem of motion detection, and demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of event-based cameras.
@@
 %%%-----------------------------------------------------------------
-Machine learning, and computer vision in particular, have undergone a major transformation in modern times, transitioning from an active actor in science fiction movies to become mainstream tools found in the smartphones in everyone's pocket.
-Less visible, but even more crucial to our economy and security, these automated tools now dominate the range of services that equip virtually most sectors of our daily lives.
+Machine learning, and computer vision in particular, have undergone a major transformation in modern times, transitioning from an active character in science fiction movies to become mainstream tools found in the smartphones in everyone's pocket.
+Less visible, yet crucial to our economy and security, these automated tools now dominate the range of services that equip virtually most sectors of our daily lives.
 This trend is profoundly pronounced in the field of medicine, as it provides practitioners with new diagnostic tools to treat people faster and more effectively.
@@
 This could raise important questions about the reliability of the results of any automated diagnostic model.
-More generally, these models do not explain why they produce a given output for a class of inputs.
+More generally, these models do give an output for any possible input but they lack an explanation as to why they produced it.
 This lack of explainability is a hindrance to the advancement of these methodologies as well as to their acceptance for most safety-relevant tasks.
@@
 As an example, the best performing model in the Imagenet challenge uses 2100 million parameters.
-Training the model typically cost over 2500 GFLOPS and about 1 joule to infer a single image.
-Less power-hungry solutions have emerged, such as Mobile Net, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significative percentage of the energy consumed by an electric car under standard conditions.
+Training the model typically consumes kilowatts of power and about 1 joule to infer a decision per single image.
+Less power-hungry solutions have emerged, such as Mobile Net, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significant percentage of the energy consumed by an electric car under standard conditions.
 
+% The brain achieves all these on a low computational cost and power budget (i.e., tens of watts), whereas the state-of-the-art computer vision models require multiple graphics processors consuming kilowatts of power
 Neuroscience could provide revolutionary answers to these two closely related problems.
-Indeed, the human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, estimated at 5W at rest.
+Indeed, the human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, about tens of watts.
 This system is the result of millions of years of natural selection, and a striking difference between biological neural networks and artificial ones is the representation they use.
@@
 Another class of approaches have designed normative models of spiking neural networks that aim to have applications in machine learning.
-One of these is the SpikeNet algorithm which uses a radical approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
+One of these is the SpikeNet algorithm which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
 Another type of spiking neural network using precise spike timing attempts to determine the structure of the network in order to minimize a cost function.
 This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extanded in novel gradient-based methods.
-Surrogate gradient for instance is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}.
-However, the performance of SNNs is still lower than that of analog networks and the question of the advantage of using spikes remains open.
+The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}.
+However, the performance of SNNs is still lagging that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
 
@@
 The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations.
-The HOTS model~\citet{lagorce_hots_2017} represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events.
+It also offer the possiblity to use the precise timing of events.
+In the HOTS model for instance, the event stream induces spatio-temporal relationships between events which are represented by building ``time surfaces'', 2D images computed using the time difference to the last recorded events~\citep{lagorce_hots_2017}.
 In a recent study, we have introduced a spiking neural network as a classification model applied to event-based streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model.
@@
 A series of similar algorithms are applied to real-life applications such as optical flow~\citep{benosman_asynchronous_2012}, depth estimation~\citep{dardelet_event-by-event_2021}, or gesture recognition~\citep{maro_event-based_2020}.
-Recently, such an algorithm was used in the CASPR mission on the ISS, illustrating, given the stringent selection processes used by the space sector, that such algorithms have reached maturity for industry.
+Recently, such an algorithm was used in the CASPR mission on the International Space Station, illustrating, given the stringent selection processes used by the space sector, that such algorithms have reached maturity for industry.
 
@@
 This paper explores the possibility that prototypical spiking motifs may be a useful representation used in neurobiology and that we may transfer this hypothesis to neuromorphic algorithms.
-Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
-Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, heterogeneous delays.
-In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay in the computational process.
-As a consequence, this extends our previous model by encoding spiking motifs as 3D images.
+Our spiking neural network model starts as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
+Here, we propose to extend our model to a layer of spiking neurons which include heterogeneous delays in addition to weights.
+In particular, one afferent may be connected with multiple delays.
+Crucially, we will explicitly use the delay in the computational process.
+As a consequence, this extends our previous MLR model by encoding spiking motifs as spatio-temporal patterns.
 The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier.
@@
 This will open perspectives for future SNNs.
-In particular, as neuromorphic devices are by design good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is transferable to this type of hardware.
+In particular, as neuromorphic devices are by design good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is perfectly fit to be transfered to this type of hardware.
 %
@@
 By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory.
-Solving is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina.
+Solving this task is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina.
 The task is thus to guess the motion as accurately as possible (see \seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays.

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %
-%● Spike based spatiotemporal processing and its implications for neuromorphic vision.
+%● Spike based spatio-temporal processing and its implications for neuromorphic vision.
 %
@@
 %%%-----------------------------------------------------------------
-\note{c'est la meme expe faite par baudot et modelisée par kremkow}
-Let us now define a trajectory inspired by the biological movements of the eyes.
+Let us now define a procedure of animating of natural scene by virtual eye-movements  similar to that which was used in neurobiological~\citep{baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
+First, let's define a trajectory inspired by the biological movements of the eyes.
 Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the field of vision.
@@
 This approach first defines a finite set of possible 2D motions in polar coordinates.
-Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of N-V-phi=8 linearly spaced motion directions and N-Y-velocity=7 different velocities (see inset of \seeFig{motion_task}-(a)).
-Note that the velocities are sampled on a geometric scale between $1/4$ and $v-speed_base = 4$.
-Next, we define one trajectory of gaze as segments whose duration is drawn from a Poisson distribution with an average block length = $20~\ms$ - similar to a Lévy flight~\citep[p.
+Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of % N-V-phi
+$8$ linearly spaced motion directions and %N-Y-velocity=
+$7$ different velocities (see inset of \seeFig{motion_task}-(a)).
+Note that the velocities are sampled on a geometric scale between $1/4$ and $4$.
+Next, we define one trajectory of gaze as segments whose duration is drawn from a Poisson distribution with an average block length of $20~\ms$, similarly to a Lévy flight~\citep[p.
 289]{mandelbrot_fractal_1982}.
-Finally, the trajectory is integrated by assuming 1/ that velocities are uniformly and independently sampled from the set of different motion-sets and 2/ that motion is uniform during a time segment.
+Finally, the trajectory is integrated by assuming first that velocities are uniformly and independently sampled from the set of different motion-sets and second that motion is uniform during a time segment.
 The resulting instances yield trajectories qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(a)).
-A similar procedure of animating of natural scene by virtual eye-movements was previously in neurobiliological~\citep{Baudot2013} and computational neuroscience~\citet{Kremkow2016} studies.
 
@@
     \caption{
-    {\bf Motion detection task.} {\bf (a)} To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like random walk, and we show here on example path of the trajectory.
-{\bf (b)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (here the cross at the center of the image) such as to reproduce the effect of displacing the eye.
-{\bf (c)} Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation.
+    {\bf Motion detection task.} {\bf (a)}  {\bf (b)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (here the cross at the center of the image) such as to reproduce the effect of displacing the eye.
+To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like random walk, and we show here on example path of the trajectory.
+{\bf (b)} Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation.
 Mimicking the retina, this representation codes for proportional increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events.
@@
 Now that our motion detection task is defined in visual space, we can extend the convolution-based detection model to a 3D convolution so that the resulting model would also benefit from spatial invariance.
-The use of spatiotemporal filters on an event stream has improved the performance of CNNs for an action recognition task by~\citet{ghosh_spatiotemporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.
+The use of spatio-temporal filters on an event stream has improved the performance of CNNs for an action recognition task by~\citet{ghosh_spatio-temporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.
 For this purpose, we design 3D kernels of shape $(\Kx, \Ky, \Ktime) = (9, 9, 11)$, representing the two spatial dimensions and the range of time delays, respectively.
-The computations are performed on the spatiotemporal windows that are defined by the kernels and are "moving" around the events, i.e.
-the spatial center of the spatiotemporal window around the current event $\epsilon_\arank$.
+The computations are performed on the spatio-temporal windows that are defined by the kernels and are "moving" around the events, i.e.
+the spatial center of the spatio-temporal window around the current event $\epsilon_\arank$.
 Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
@@
 %ith the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
-We also observe that these cells show an orientation selectivity, similar to that observe in MT neurons~\citep{DeAngelis1999}.
-Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{Kremkow2016} to explain neurophysiological results~\citep{Baudot2013}.
+We also observe that these cells show an orientation selectivity, similar to that observe in MT neurons~\citep{deangelis_functional_1999}.
+Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results~\citep{baudot_animation_2013}.
 Focusing on the positive weights, a strong selectivity is observed along specific axes of motion for each of the different kernels.
@@
 Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli.
-In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds})~\citep{Vacher2018}.
+In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds})~\citep{vacher_bayesian_2018}.
 In particular, we will set the spatial size and duration similarly to the motion task defined above.
@@
 We plot here main axis of interests.
-First, as we change the mean spatial frequency of the texture, we observe a montonous decrease in accuracy.
-This comes as a similar trend as that observed in the primary visual areas~\citep{Priebe2006}.
-\note{we should really do something more serious here...} Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively ressemble a more textured stimuls) than for a grating-like stimulation, reminiscent to the behavioural response of humans' eye movemnts to such stimuli~\citep{Simoncini2012}.
+First, as we change the mean spatial frequency of the texture, we observe a monotonous decrease in accuracy.
+This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006}.
+\note{we should really do something more serious here...} Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively resemble a more textured stimulus) than for a grating-like stimulation, reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}.
 Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.
@@
 Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter.
-This shows a progressive decrease in accuracy similar to that observed in the amplitude of humans' eye movements~\citep{MansourPour2018}.
+This shows a progressive decrease in accuracy similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018}.
 %
@@
 To recap, we have introduced in this paper a spiking neural network model with heterogeneous delays that we have trained and evaluated on a complex motion detection task.
-The model was defined to optimally detect event-driven spatiotemporal motifs.
+The model was defined to optimally detect event-driven spatio-temporal motifs.
 We have shown that when the model is trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example.
@@
 Let us highlight some innovations in the contributions presented in this paper.
-First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs, whereas~\citet{ghosh_spatiotemporal_2019} use a correlation-based heuristic, which we have observed to be less efficient.
+First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs, whereas~\citet{ghosh_spatio-temporal_2019} use a correlation-based heuristic, which we have observed to be less efficient.
 Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking motif.

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 \newcommand{\ShortTitle}{
-Learning synaptic delays of spiking neurons for motion detection
+Learning heterogeneous delays of spiking neurons for motion detection
 }
@@
 Training the model typically consumes kilowatts of power and about 1 joule to infer a decision per single image.
-Less power-hungry solutions have emerged, such as Mobile Net, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significant percentage of the energy consumed by an electric car under standard conditions.
+Less power-hungry solutions have emerged, such as Mobile Net~\citep{how}, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significant percentage of the energy consumed by an electric car under standard conditions.
 
@@
 Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels.
+This threshold was found using validation and future work should automate how this may be determined for various datasets.
 Given a stimulus input, the model thus yields a spike output in the postsynaptic address space.
@@
 By taking the logit of the probability of the output given by the model, we can therefore calculate the probability of the output.
-This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error $|| \speed - ||^2$ of the velocity estimate given by :
-$
-\speed = \sum_{\class} \speed_\class \cdot Pr(\class \vert image)
-$ 
+This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error% $|| \speed - ||^2$
+of the velocity estimate given by average estimate. 
+%$\hat{\speed} = \sum_{\class} \speed_\class \cdot Pr(\class \vert image)$ 
 These calculations are performed on a different input dataset than the one used in the training or validation steps.
@@
 However, the generic theoretical model is able to include them by inserting new spikes into the list of spikes that are reaching pre-synaptic addresses.
-While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside of a homeostatic state (through a death or explosion of activity).
+While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside of a homeostatic state (through a extinction or explosion of activity).
 \note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}~\citep{perrinet_motion-based_2012}

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
full pass

@@
 This can be the identified set of neurons in a neurobiological recording.
-For event-based cameras, it is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for ON and OFF polarities).
+For event-based cameras, it is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras).
 
@@
 A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree, the so-called weight matrix.
-A more detailed description of these synaptic contacts reveals the added importance of heterogeneous delays, i.e., the precise timing from synaptic activation to its integration into the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}.
+A more detailed description of these synaptic contacts reveals the added importance of heterogeneous delays, i.e., the precise timing from one afferent neuron's firing to its arrival to the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}.
 In that heterogeneous description, we can parameterize each neuron by the set of tuples defining both the weight and the delay of each synaptic contact.

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Training the model typically consumes kilowatts of power and about 1 joule to infer a decision per single image.
-Less power-hungry solutions have emerged, such as Mobile Net~\citep{how}, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significant percentage of the energy consumed by an electric car under standard conditions.
+Less power-hungry solutions have emerged, such as Mobile Net~\citep{howard_searching_2019}, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significant percentage of the energy consumed by an electric car under standard conditions.
 

### 2022-09-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Several models have increased the scale of the system to include more neurons, such as The Virtual Brain (TVB)~\citep{sanz_leon_virtual_2013}, which mixes different scales and includes spiking neural networks with (analog) neural mass models.
-However, these methods are mainly descriptive with applications in neuroscience but few spiking neurall models have had an application in computer vision.
+However, these methods are mainly descriptive with applications in neuroscience but few spiking neural models have had an application in computer vision.
 Another class of approaches have designed normative models of spiking neural networks that aim to have applications in machine learning.
@@
 The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations.
-It also offer the possiblity to use the precise timing of events.
+It also allows using the precise timing of events.
 In the HOTS model for instance, the event stream induces spatio-temporal relationships between events which are represented by building ``time surfaces'', 2D images computed using the time difference to the last recorded events~\citep{lagorce_hots_2017}.
-In a recent study, we have introduced a spiking neural network as a classification model applied to event-based streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model.
+In a recent study, we have introduced a SNN as a classification model applied to event-based streams using Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model.
 By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is, in an online fashion.
-We have demonstrated on several benchmark datasets that it provides with efficient computations resulting in ultra-fast categorization.
+We have demonstrated on several benchmark datasets that it provides with efficient computations, resulting in ultra-fast categorization.
 Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.
-A series of similar algorithms are applied to real-life applications such as optical flow~\citep{benosman_asynchronous_2012}, depth estimation~\citep{dardelet_event-by-event_2021}, or gesture recognition~\citep{maro_event-based_2020}.
+A serie of similar algorithms are applied to real-life applications such as optical flow~\citep{benosman_asynchronous_2012}, depth estimation~\citep{dardelet_event-by-event_2021}, or gesture recognition~\citep{maro_event-based_2020}.
 Recently, such an algorithm was used in the CASPR mission on the International Space Station, illustrating, given the stringent selection processes used by the space sector, that such algorithms have reached maturity for industry.
@@
 Let us now formally define the heterogeneous model, as well as the task and how we will learn to solve it.
-First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neural network (SNN) with heterogeneous delays.
+First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays.
 Our SNN model will be tested on a computationally relevant task of motion detection in an event-based sensory input.
@@
 Solving this task is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina.
-The task is thus to guess the motion as accurately as possible (see \seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays.
+The task is thus to guess the motion as accurately as possible (\seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays.
 To this end, we will extend the SNN model to be efficient for the resolution of the task by including convolutional kernels.

### 2022-09-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Indeed, in autonomous driving, an important task is to react as quickly as possible to a possible collision, for example when a pedestrian unexpectedly crosses the street.
-As such, future systems should always be able to give a response at any time, instead of having to wait for the pipeline to provide the correct response.
+As such, future systems should always be able to give a response at any time, instead of having to wait for the pipeline to provide the correct answer.
 This online processing is also a feature of processing in biological systems and could be an important first contribution if we could translate these principles into future automated processes.
@@
 In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator.
-Using this forward model, it possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator.
+Using this forward model, it is possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator.
 It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.
@@
 This process is known to occur as early as the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step (see \seeFig{motion_task}-(b)).
+These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step (\seeFig{motion_task}-(b)).
 We will discretize the time in $1~\ms$ bins and produce movies of duration N-T = $250~\ms$.
@@
 The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions.
-As in this semi-supervised context, the label is lknown but the timing is not, we used a selection of the temporal suppport in an unsuperviesd manner.
+As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner.
 It is important to note that we weighted the cost function by considering only active cells, such that the error is only backpropagated to the spatial locations of these most active cells.
@@
 By taking the logit of the probability of the output given by the model, we can therefore calculate the probability of the output.
-This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error% $|| \speed - ||^2$
-of the velocity estimate given by average estimate. 
+This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error % $|| \speed - ||^2$
+of the velocity estimate given by the average estimate. 
 %$\hat{\speed} = \sum_{\class} \speed_\class \cdot Pr(\class \vert image)$ 
@@
 After training our model, we first analyze the weights learned for the different neurons (\seeFig{kernels}). 
-We first observed a high dependendance between the weights reaching the ON polarities and that reaching the OFF polarities.
-In particular, whenever a weight for a given position and delay is positive for one poalrity, it will be negative in the other.
+We first observed a high dependence between the weights reaching the ON polarities and that reaching the OFF polarities.
+In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative in the other.
 This property comes from the way the events are generated and that the luminance can not at the same time increase and decrease.
@@
 Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter.
-This shows a progressive decrease in accuracy similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018}.
+This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018}.
 %
@@
 Indeed, the computations are dominated by the convolution operation.
-In a dense setting, this corresponds for all voxels in the ouput to a sum over all voxels in the inputs for all weights in the kernel.
+In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel.
 If the support of information is sparse, then computations can be performed only on those events.
-Also, if we set some weights of the kernels to zero, then the sum can be skipped for thse adresses.
+Also, if we set some weights of the kernels to zero, then the sum can be skipped for those addresses.
 Knowing the sparseness of the input, the total number of computations thus scales with the number of nonzero synaptic weights.
@@
 In this paper, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification.
-The learned model bears many similarities with neurobilogical anatomical observations but also with the results from behavioral results.
+The learned model bears many similarities with neurobiological anatomical observations but also with the results from behavioral results.
 \note{to change if we compare to ghosh instead of HOTs if we follow that route} The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification.

### 2022-09-30 Grimaldi-etal-BiolCybernetics.tex
figure motion task + kernels

@@
     \includegraphics[width=0.95\linewidth]{figures/motion_task.pdf}
-    \note{TODO: redo figure according to caption}
     \caption{
-    {\bf Motion detection task.} {\bf (a)}  {\bf (b)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (here the cross at the center of the image) such as to reproduce the effect of displacing the eye.
-To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like random walk, and we show here on example path of the trajectory.
+    {\bf Motion detection task.} {\bf (a)}  {\bf (b)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (the image on the left) such as to reproduce the effect of displacing the eye.
+To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like random walk, and we show here on example path of the trajectory as a green line ($200$ steps).
 {\bf (b)} Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation.
@@
     	Representation of the weights for $3$ of the $32$ different learned kernels of the model.
-Each corresponding to the OFF polarities and selective to the different motion directions (because of the symmetry observed between the ON and OFF event streams, we observed that kernels are very similar for the ON polarities).
+Each pair of line correspond to the ON and OFF polarities respectively, with excitatory weights in warm colors.
+Delays are represented in the horizontal axis from right (zero delay) to the left (delay of $11$ steps).
+Because of the symmetry observed between the ON and OFF event streams, we observed that kernels are very similar for the ON polarities.
 These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
 %	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.
-        One sees positive (excitatory) coefficients for the specific direction of motion but also negative (inhibitory) coefficients for other directions.
+	Different kernels are selective to the different motion directions and we observe some level of orientation selectivity, where ON and OFF subfields organized in a push-pull organization. 
+%        One sees positive (excitatory) coefficients for the specific direction of motion but also negative (inhibitory) coefficients for other directions.
 	}

### 2022-09-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 The challenges are numerous in such situations and two major obstacles must be highlighted: the computational delay and the energy budget.
-First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun1998gradient}, then high definition images, such as VGG on Imagenet~\citep{Simonyan:2014cmh}.
+First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun_gradient-based_1998}, then high definition images, such as VGG on Imagenet~\citep{simonyan_very_2015}.
 This has a cost and notably the fact that the computation time increases with the size of the image.
@@
 CNNs, for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels.
-These networks have been known to mimick many propoerties of biological systems~\citep{Schrimpf2020}, yet, this score does not count rapidity or energy.
+These networks have been known to mimick many propoerties of biological systems~\citep{schrimpf_brain-score_2020}, yet, this score does not count rapidity or energy.
 On the other hand, information in a large majority of biological neural networks is represented as spikes, i.e.
@@
 Now that our motion detection task is defined in visual space, we can extend the convolution-based detection model to a 3D convolution so that the resulting model would also benefit from spatial invariance.
-The use of spatio-temporal filters on an event stream has improved the performance of CNNs for an action recognition task by~\citet{ghosh_spatio-temporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.
+The use of spatio-temporal filters on an event stream has improved the performance of CNNs for an action recognition task by~\citet{ghosh_spatiotemporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.
 For this purpose, we design 3D kernels of shape $(\Kx, \Ky, \Ktime) = (9, 9, 11)$, representing the two spatial dimensions and the range of time delays, respectively.
@@
 It is important to note that we weighted the cost function by considering only active cells, such that the error is only backpropagated to the spatial locations of these most active cells.
-This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier2007}.
+This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier_unsupervised_2007}.
 Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$).
@@
 Let us highlight some innovations in the contributions presented in this paper.
-First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs, whereas~\citet{ghosh_spatio-temporal_2019} use a correlation-based heuristic, which we have observed to be less efficient.
+First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs, whereas~\citet{ghosh_spatiotemporal_2019} use a correlation-based heuristic, which we have observed to be less efficient.
 Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking motif.
@@
 %%%-----------------------------------------------------------------
-We have identified several limits to our model, the most important we will now detail.
+We have identified several limits to our model which we will now detail.
 First, the complete framework is based on a discrete binning of time that is incompatible with the continuous nature of biological time.

### 2022-09-30 Grimaldi-etal-BiolCybernetics.tex

@@
 We apply this model to solve the specific computer vision problem of motion detection, and demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of event-based cameras.
-In particular, we quantify how its accuracy can vary with the total computational load.
+In particular, we quantify how the accuracy of the model can vary with the total computational load.
 This end-to-end event-driven computational brick could help improve the performance of future Spiking Neural Network (SNN) algorithms and their prospective use in neuromorphic chips.
@@
 %\DeclareRobustCommand
-\newcommand{\note}[1]{{\sethlcolor{yellow}\hl{#1}}}
+%\newcommand{\note}[1]{{\sethlcolor{yellow}\hl{#1}}}
+\newcommand{\note}[1]{}
 %\newcommand{\noteLP}[1]{\todo[author=\textbf{Laurent}, inline]{#1}}
@@
 The task is thus to guess the motion as accurately as possible (\seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays.
-To this end, we will extend the SNN model to be efficient for the resolution of the task by including convolutional kernels.
+To this end, we will extend the SNN model to be efficient for the resolution of the task by including spatio-temporal convolutional kernels.
 We will finally describe its implementation on conventional computers.
@@
  \presynaddr_\ranksyn^\postsynaddr, \timev=\timev_\arank + 
-\synapticdelay_\ranksyn^\postsynaddr \}_{ \arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]} )
+\synapticdelay_\ranksyn^\postsynaddr )_{ \arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]} )
 \}$. \note{formaliser le multiplexage}
@@
 When these connections are high and sparsely distributed, this firing will cause a specific temporal motif.
-Another example is given for the barn owl auditory system: When it hears the sound of a mouse, this will generate a specific spiking response in both ears, and specifically, the precise timing between these signal generated by the the left relative to the right ear can be for instance used to determine the position of the prey~\citep{goodman_spike-timing-based_2010}.
+Another example is given for the barn owl auditory system: As it hears the sound of a mouse, this sound will generate a specific spiking response in both ears, and specifically, the precise timing between the signal generated by the the left relative to the right ear can be for instance used to determine the position of the prey~\citep{goodman_spike-timing-based_2010}.
 Overall, these examples show that raster plots may be considered as a mixture of the effects of different elementary causes, and that each event triggers a specific spatio-temporal motif.
@@
 In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
-Using this dense representation, the counting defined above becomes:
-\begin{equation*}
-\mathcal{C}^\postsynaddr(a,t)
-= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
-\end{equation*}
-%
-This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
-This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
-%
-
-\note{formalize the following:}
-More generally, every cause may be considered as occurring independently and one might write the generative model for the generation of the presynaptic events as:
-
-Using this formalization, one might now deduce an optimal algorithm for the detection of such temporal motifs.
-
-By discretization of time (with here an arbitrary unitary time unit), we can also define the dendrite as a matrix giving the weight corresponding to the different delays $d \in [0, D]$ (where $D$ is the maximum delay) on different pre-synaptic addresses $a \in [1, N]$ defining the list of the $N$ dendrites.
-We will denote as $W(a, d)$ these weights.
-
-Following the observations of~\citet{izhikevich_polychronization_2006}, let us assume that such precise discharge motif defines a polychronous group (PG).
-Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
-This allows then to derive a generative model for raster plots (see Figure~\ref{fig:model}-(b)).
-
-Indeed, the probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
-Assuming that the presence of PGs is conditionally independent, this bias can be written as the combination of these factors, whose values are given by the corresponding weights.
+The probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
+Assuming that the presence of PGs conditions the probability on all efferents, the logit (inverse of the sigmoid) of this probability bias can be written as the sum of the logit of each of these factors, whose values are given by the corresponding weights.
 PGs may be activated independently at random times and  we write that $B(b, t)=1$ if $b$ is activated at $t$ (and else $B(b, t)=0$).
@@
 This formulation thus defines a simple generative model for raster plots as a combination of independent PGs.
+%
+%Using this dense representation, the counting defined above becomes:
+%\begin{equation*}
+%\mathcal{C}^\postsynaddr(a,t)
+%= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
+%\end{equation*}
+%%
+%This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
+%This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
+%%
+%
+%\note{formalize the following:}
+%More generally, every cause may be considered as occurring independently and one might write the generative model for the generation of the presynaptic events as:
+%
+%
+%Using this formalization, one might now deduce an optimal algorithm for the detection of such temporal motifs.
+%
+%
+%By discretization of time (with here an arbitrary unitary time unit), we can also define the dendrite as a matrix giving the weight corresponding to the different delays $d \in [0, D]$ (where $D$ is the maximum delay) on different pre-synaptic addresses $a \in [1, N]$ defining the list of the $N$ dendrites.
+We will denote as $W(a, d)$ these weights.
+%
+%
+%Following the observations of~\citet{izhikevich_polychronization_2006}, let us assume that such precise discharge motif defines a polychronous group (PG).
+Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
+This allows then to derive a generative model for raster plots (see Figure~\ref{fig:model}-(b)).
 
-\subsubsection{Implementation using spiking neurons with heterogeneous delays}
+%\subsubsection{Implementation using spiking neurons with heterogeneous delays}
 %
 The activation function of our spiking neural is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
-It transforms this list of weights into a probability with the following formula:
-$
-Pr(k=\postsynaddr \; \vert \; \timev) =
-\frac 1 Z
-{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
-$ 
-where $\mathcal{C}^\postsynaddr(\timev) = \sum
-\activeweights^\postsynaddr(t)
-$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
-In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
+%It transforms this list of weights into a probability with the following formula:
+%$
+%Pr(k=\postsynaddr \; \vert \; \timev) =
+%\frac 1 Z
+%{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
+%$ 
+%where $\mathcal{C}^\postsynaddr(\timev) = \sum
+%\activeweights^\postsynaddr(t)
+%$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
+%In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
+%%
 %
-
 \note{say that it is effortless in biology but difficult on conventional computers}
@@
     \includegraphics[width=0.95\linewidth]{figures/motion_clouds.pdf}
-    \caption{{\bf Role of image parameters in the motion detection accuracy.} Accuracy as a function of {\bf (a)} the mean spatial frequency, {\bf (b)} the bandwidth in orientation: from a grating-like (left) to isotropic textures (right)).
-{\bf (c)} the bandwidth in speed, from a rigid motion (left) to independent frames (right).
-Note that these accuracy is computed both in the case is necessarily perpendiular to the motion (aperture) or aribtrary (no aperture), showing that accuracy decreases in the latter case.}
+    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of {\bf (a)} the mean spatial frequency, {\bf (b)} the bandwidth in orientation: from a grating-like (left) to isotropic textures (right)), {\bf (c)} the bandwidth in speed, from a rigid motion (left) to independent frames (right).
+Note that these accuracy is computed both in the case is necessarily perpendicular to the motion (no aperture) or arbitrary (aperture), showing that accuracy decreases in the latter case.}
     \label{fig:motion_clouds}
@@
 Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter.
-This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018}.
+This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018} but also that accuracy is low for a rigid motion which lacks variability.
 %

### 2022-09-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 }
-\newcommand{\Acknowledgments}{
-Thanks to CB, TS, ...
-This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}'').
-%HL and LP received funding from the ANR project  N° ANR-20-CE23-0021 (''\href{https://laurentperrinet.github.io/grant/anr-anr/}
-LP received funding from the ANR project  N° ANR-22-XXX-XXX (''\href{https://laurentperrinet.github.io/grant/polychronies/}{polychronies}''). 
+\newcommand{\Funding}{
+This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{APROVIS3D}{http://aprovis3d.eu/}'').
+%HL and
+LP received funding from the ANR project  N° ANR-20-CE23-0021 ``\href{AgileNeuroBot}{https://laurentperrinet.github.io/grant/anr-anr/}''
+% LP received funding 
+and from the ANR project % N° ANR-22-XXX-XXX 
+``\href{PolyChronies}{https://laurentperrinet.github.io/grant/polychronies/}''. 
 %HL received funding from a University of Montreal Artificial Intelligence scholarship. 
 %
+}
+\newcommand{\Acknowledgments}{
+Thanks to Thomas Schatz, Rosa Cossart, Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions during the elaboration of this work. %
+\Funding %
 %For the purpose of open access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript version arising from this submission.
@@
 \title[\ShortTitle]{\Title}
+\artnote{This article is published as part of the Special Issue on "What can Computer Vision learn from Visual Neuroscience?"}
 %
@@
 This particular example also illustrates some important challenges that these tools are increasingly facing.
-One of these is their sensitivity to adversarial attacks, that is, the sensitivity of the models' output to minute changes in the input.
+One of these is the sensitivity of deep-learning based algorithms to adversarial attacks, that is, the sensitivity of the models' output to minute changes in the input.
 This could raise important questions about the reliability of the results of any automated diagnostic model.
@@
 The paper is organized as follows.
-We develop a theoretically defined SNN capable of learning from heterogeneous delays that we first evaluate on a synthetic event-based dataset.
+We develop a theoretically defined SNN capable of learning heterogeneous delays from spiking motifs that we first evaluate on a synthetic event-based dataset.
 We first detail the methodology by defining the basic mechanism of spiking neurons that utilize heterogeneous delays.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Declarations}
 
-\textcolor{purple}{TBD - What the template says about it:}
+\section*{Statements and Declarations}
 
-\textcolor{purple}{Some journals require declarations to be submitted in a standardised format.
-Please check the Instructions for Authors of the journal to which you are submitting to see if you need to complete this section.
-If yes, your manuscript must contain the following sections under the heading `Declarations': Funding; Conflict of interest/Competing interests (check journal-specific guidelines for which heading to use); Ethics approval; Consent to participate; Consent for publication; Availability of data and materials; Code availability; Authors' contributions.
-}
+\paragraph{Funding}
 
-\textcolor{purple}{If any of the sections are not relevant to your manuscript, please include the heading and write `Not applicable' for that section. }
+\Funding %
 
+\paragraph{Conflict of interest}
+Not applicable.
+% The authors do not have any financial or non-financial interests directly or indirectly related to this work.
+
+\paragraph{Ethics approval}
+Not applicable.
+
+\paragraph{Consent to participate}
+Not applicable.
+
+\paragraph{Consent for publication}
+Not applicable.
+
+\paragraph{Availability of data and materials}
+Not applicable.
+
+\paragraph{Code availability}
+
+The code is publicly available online at: \url{https://github.com/SpikeAI/pyTERtorch}.
+
+\paragraph{Authors' contributions}
+
+Both authors contributed to the conceptualization and methodology design of the study, to the project's coordination and administration.
+Laurent Perrinet carried out the funding acquisition and supervision.
+Formal analysis and investigation were performed by both authors.
+Results visualization and presentation were realized by both authors.
+The the manuscript was written by Laurent Perrinet.
+Both authors have read and approved the final manuscript.
 %%%%%%%%%%%%%%%%%%%%%%%%

### 2022-09-30 Grimaldi-etal-BiolCybernetics.tex
Introduction: figure motion task
APA BST
a last pass

@@
 %%\documentclass[sn-vancouver]{sn-jnl}% Vancouver Reference Style
-%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style
+\documentclass[sn-apa]{sn-jnl}% APA Reference Style
 %%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
 %%\documentclass[sn-standardnature]{sn-jnl}% Standard Nature Portfolio Reference Style
-\documentclass[default]{sn-jnl}% Default
+%%\documentclass[default]{sn-jnl}% Default
 %%\documentclass[default,iicol]{sn-jnl}% Default with double column layout
-%\documentclass[iicol]{sn-jnl}
+%%\documentclass[iicol]{sn-jnl}
 
@@
 % ---------------
-\author*{\fnm{\FirstAG} \sur{\LastAG}}\email{\EmailAG}
+\author{\fnm{\FirstAG} \sur{\LastAG}}\email{\EmailAG}
 \author{\fnm{\FirstLP} \sur{\LastLP}}\email{\EmailLP}
@@
     \caption{
-    {\bf Motion detection task.} {\bf (a)}  {\bf (b)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (the image on the left) such as to reproduce the effect of displacing the eye.
-To mimic the effect of a saccadic-like movement, the axis of view moves following a step-like random walk, and we show here on example path of the trajectory as a green line ($200$ steps).
+    {\bf Motion detection task.} {\bf (a)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (the image on the left) such as to reproduce the effect of displacing the eye.
+To mimic the effect of a saccadic-like eye movement, the axis of view moves following a step-like random walk, and we show here on example path of the trajectory as a green line ($200$ time steps).
 {\bf (b)} Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation.
@@
     	Representation of the weights for $3$ of the $32$ different learned kernels of the model.
-Each pair of line correspond to the ON and OFF polarities respectively, with excitatory weights in warm colors.
+Each pair of line correspond to the OFF and ON polarities respectively, with excitatory weights in warm colors.
 Delays are represented in the horizontal axis from right (zero delay) to the left (delay of $11$ steps).
@@
 For example, the combination of neighboring orientations may contribute to image categorization~\citep{perrinet_edge_2015}.
-Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background~\citep{roelfsema_early_2016} and it has been shown that feedback may be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020, boutin_pooling_2022}.
+Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background~\citep{roelfsema_early_2016} and it has been shown that feedback may be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020}, Notably to explain non-linear mechanisms~\citep{boutin_pooling_2022}.
 It is currently not possible to implement these recurrent connections in our implementation (lateral or feedback), mainly due to our use of convolutions.
@@
 While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside of a homeostatic state (through a extinction or explosion of activity).
-\note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}~\citep{perrinet_motion-based_2012}
+\note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}%~\citep{perrinet_motion-based_2012}
 
@@
 \note{Say that this could be implemented by properly setting the timing of the label} As previously stated, this could be implemented using generalized coordinates (that is variables such as position which are complemented by speed, acceleration, jerk, ...) and ``neurobiologically, the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
-Our proposed heterogeneous model provides an alternate and elegant implementation solution to this problem.
+Our proposed model using heterogeneous delays provides an alternate and elegant implementation solution to this problem.
 %

### 2022-09-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
a last pass

@@
 \newcommand{\Acknowledgments}{
-Thanks to Thomas Schatz, Rosa Cossart, Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions during the elaboration of this work. %
+Thanks to Thomas Schatz, Rosa Cossart, Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work. %
 \Funding %
@@
 On the other hand, information in a large majority of biological neural networks is represented as spikes, i.e.
-prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted this spike.
+prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted this spike~\citep{paugam-moisy_computing_2012}.
 Why spikes?

### 2022-09-30 Grimaldi-etal-BiolCybernetics.tex
a last pass

@@
 However, most artificial neuronal models do not take advantage of this minute temporal dimension.
-Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of neurons with heterogeneous delays.
+Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of neurons with heterogeneous delays which we apply to the computer vision task of motion detection.
 Indeed, the variety of synaptic delays on the dendritic tree allows to synchronize synaptic inputs as they reach the basal dendritic tree.

### 2022-09-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{APROVIS3D}{http://aprovis3d.eu/}'').
-%HL and
+% Agence Nationale de la Recherche %HL and
 LP received funding from the ANR project  N° ANR-20-CE23-0021 ``\href{AgileNeuroBot}{https://laurentperrinet.github.io/grant/anr-anr/}''
@@
 %%\documentclass[sn-vancouver]{sn-jnl}% Vancouver Reference Style
-\documentclass[sn-apa]{sn-jnl}% APA Reference Style
+%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style
 %%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
 %%\documentclass[sn-standardnature]{sn-jnl}% Standard Nature Portfolio Reference Style
-%%\documentclass[default]{sn-jnl}% Default
+\documentclass[default]{sn-jnl}% Default
 %%\documentclass[default,iicol]{sn-jnl}% Default with double column layout
@@
 \usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.
-
+\AtBeginDocument{\RenewCommandCopy\qty\SI}
 \DeclareMathOperator*{\argmax}{arg\,max}
@@
 %
-\begin{figure}[t!]
+\begin{figure}%[t!]
     \centering
@@
 %%%-----------------------------------------------------------------
-\begin{figure}[h!]%[ht!]
+\begin{figure}%[h!]%[ht!]
     \centering
@@
 Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli.
-In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds})~\citep{vacher_bayesian_2018}.
+In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
 In particular, we will set the spatial size and duration similarly to the motion task defined above.
@@
 %%%-----------------------------------------------------------------
-\begin{figure}[h!]
+\begin{figure}%[h!]
     \centering

### 2022-09-30 cover_letter/coverletter.tex
cover letter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Long Lined Cover Letter
% LaTeX Template
% Version 1.0 (1/6/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Matthew J. Miller
% http://www.matthewjmiller.net/howtos/customized-cover-letter-scripts/
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt,stdletter,dateno,sigleft]{newlfm}

\usepackage{etoolbox}
\makeatletter
\patchcmd{\@zfancyhead}{\fancy@reset}{\f@nch@reset}{}{}
\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
\patchcmd{\@set@em@up}{\f@ncyorh}{\f@nch@orh}{}{}
\makeatother

%\makeatletter
%\patchcmd{\@zfancyhead}{\fancy@reset}{\f@nch@reset}{}{}
%\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
%\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
%\patchcmd{\@set@em@up}{\f@ncyorh}{\f@nch@orh}{}{}
%\makeatother

\usepackage{url}
% Extra options: 'sigleft' for a left-aligned signature, 'stdletternofrom' to remove the from address, 'letterpaper' for US letter paper - consult the newlfm class manual for more options

\usepackage{charter} % Use the Charter font for the document text

\newsavebox{\Luiuc}\sbox{\Luiuc}{\parbox[b]{1.75in}{\vspace{0.5in}
\includegraphics[width=1.2\linewidth]{amu.jpg}}
{\hspace{3.5in}
\includegraphics[width=0.3\linewidth]{CNRS-Logo.png}}
} % Company/institution logo at the top left of the page
\makeletterhead{Uiuc}{\Lheader{\usebox{\Luiuc}}}

\newlfmP{sigsize=10pt}% Slightly decrease the height of the signature field
% Print a phone number under the sender's address
%\newlfmP{addrfromphone} % Print a phone number under the sender's address
\newlfmP{addrfromemail} % Print an email address under the sender's address
%\PhrPhone{Phone} % Customize the "Telephone" text
\PhrEmail{Email} % Customize the "E-mail" text

\lthUiuc % Print the company/institution logo

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\namefrom{\textit{The authors}, Antoine Grimaldi and Laurent Perrinet } % Name

\addrfrom{
\today\\[12pt]
Laurent Perrinet\\
Institute of Neuroscience of la Timone\\
CNRS - Aix Marseille University\\
27, boulevard Jean Moulin\\
13005 Marseille, France
}

%\phonefrom{(000) 111-1111} % Phone number

\emailfrom{laurent.perrinet@univ-amu.fr} % Email address

%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING/CLOSING
%----------------------------------------------------------------------------------------

\greetto{Dear editor,} % Greeting text
\closeline{Sincerely yours,} % Closing text

%\nameto{Victor Boutin} % Addressee of the letter above the to address

%\addrto{
%Neural Network Editorial Board
%Recruitment Officer \\ % To address
%The Corporation \\
%123 Pleasant Lane \\
%City, State 12345
%}

%----------------------------------------------------------------------------------------

\begin{document}
\begin{newlfm}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------
Please find enclosed a manuscript entitled ``Learning heterogeneous delays in a layer of spiking neurons for fast motion detection'' for your consideration as an article in \emph{Biological Cybernetics}. All the authors have been involved with the work, have approved the manuscript and agreed to its submission.

Our manuscript describes a novel method inspired by neuroscience to overcome some challenges faced in computer vision, notably when dealing with large amounts of data. By using heterogeneous delays on different synapses, this novel spiking neuron method is able to detect spiking motifs and we validate" the method on synthetic and then realistic data. Results show that this method could provide a path for future spiking neural network algorithms using less energy for a similar performance as thier analog counterparts.

Our  manuscript follows our submission and presentation during the ``NeuroVision'' workshop. In addition, all the python code needed to reproduce figures and supplementary materials will be fully open-sourced once the article will be published.

\end{newlfm}
\end{document}

### 2022-10-03 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 % vim: set fenc=utf-8
+%
+% public
+% https://github.com/SpikeAI/pyTERtorch/tree/main/2022-09-30_BiolCybernetics
+% https://www.zotero.org/groups/4776796/fastmotiondetection
+
+% private
+% https://framateam.org/spikeai/channels/2022-10_icip-bordeaux
+% https://www.overleaf.com/project/630f9044c38e7a3cea81a7b2
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% 2022-09-30
+% Learning heterogeneous delays in a layer of spiking neurons for fast motion detection
+% Corresponding Author: Laurent Perrinet
+% Biological Cybernetics
+% a7f45f8e-87d6-4a30-ac7f-9f0c94e2592a   |  v.1.0
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 }
+\newcommand{\DataAvailability}{
+This works is made reproducible using the following tools. First the code reproducing all figures is available on~\href{ https://github.com/SpikeAI/pyTERtorch}{GitHub}.
+
+\noindent %The paper is available a an~\href{https://arxiv.org/abs/2205.03635}{arXiv preprint} with links to previous versions ands to the code. 
+Find also the associated~\href{https://www.zotero.org/groups/4776796/fastmotiondetection}{zotero group} used to regroup relevant literature on the subject.
+
+% public
+% https://github.com/SpikeAI/pyTERtorch/tree/main/2022-09-30_BiolCybernetics
+% https://www.zotero.org/groups/4776796/fastmotiondetection
+
+% private
+% https://framateam.org/spikeai/channels/2022-10_icip-bordeaux
+% https://www.overleaf.com/project/630f9044c38e7a3cea81a7b2
+
+%In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study.
+Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.
+If the study did not report any data, you might add ``Not applicable'' here.
+Chen
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-10-04 Grimaldi-etal-BiolCybernetics.tex
submitted rev0 on 2022-09-30

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\usepackage%[disable]	% uncomment to hide all the notes
-%	{todonotes}
-%\newcommand{\noteAG}[1]{\todo[author=\textbf{Antoine}, inline]{#1}}
+% a simpler annotation system:
 \usepackage{soulutf8}
 \usepackage{color}
-
-%\DeclareRobustCommand
-%\newcommand{\note}[1]{{\sethlcolor{yellow}\hl{#1}}}
-\newcommand{\note}[1]{}
-%\newcommand{\noteLP}[1]{\todo[author=\textbf{Laurent}, inline]{#1}}
-%\newcommand{\modif}[1]{\todo[author=\textbf{Proposition}, inline]{#1}}
-
+\newcommand{\note}[1]{{\sethlcolor{yellow}\hl{#1}}}
+%\newcommand{\note}[1]{}
+%
 \begin{document}
@@
 CNNs, for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels.
-These networks have been known to mimick many propoerties of biological systems~\citep{schrimpf_brain-score_2020}, yet, this score does not count rapidity or energy.
+These networks have been known to mimic many properties of biological systems~\citep{schrimpf_brain-score_2020}, yet, this score does not count rapidity or energy.
 On the other hand, information in a large majority of biological neural networks is represented as spikes, i.e.

### 2022-10-05 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %%%-----------------------------------------------------------------
-Let us now define a procedure of animating of natural scene by virtual eye-movements  similar to that which was used in neurobiological~\citep{baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
+Let us now define a procedure of animating of natural scene by virtual eye-movements  similar to that which was used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
 First, let's define a trajectory inspired by the biological movements of the eyes.

### 2022-10-05 cover_letter/coverletter.tex
cover

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Long Lined Cover Letter
-% LaTeX Template
-% Version 1.0 (1/6/13)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% Matthew J. Miller
-% http://www.matthewjmiller.net/howtos/customized-cover-letter-scripts/
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%!TeX TS-program = lualatex
+%!TeX encoding = UTF-8 Unicode
+%!TeX spellcheck = fr-FR
+%!BIB TS-program = biber
+% -*- coding: UTF-8; -*-
+% vim: set fenc=utf-8
 %----------------------------------------------------------------------------------------
@@
 %----------------------------------------------------------------------------------------
+\documentclass[10pt,english]{article}
 
-\documentclass[10pt,stdletter,dateno,sigleft]{newlfm}
+\usepackage{charter,graphicx}
+\usepackage[margin=1in]{geometry}
 
+\usepackage{hyperref,tabularx}
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\renewcommand{\headrulewidth}{.4pt}% Default header rule
+\renewcommand{\footrulewidth}{0pt}% No footer rule
+\fancyhf{}% Clear header/footer
+\fancypagestyle{plain}{
+  \renewcommand{\headrulewidth}{0pt}% No header rule
+  \renewcommand{\footrulewidth}{.4pt}% Default footer rule
+  \fancyhf{}% Clear header/footer
+}
+\AtBeginDocument{\thispagestyle{plain}}
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{.5\baselineskip plus 1pt minus 1pt}
+\usepackage[utf8]{inputenc}%
+\usepackage{babel}%
 \usepackage{etoolbox}
@@
 \makeatother
-
-%\makeatletter
-%\patchcmd{\@zfancyhead}{\fancy@reset}{\f@nch@reset}{}{}
-%\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
-%\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
-%\patchcmd{\@set@em@up}{\f@ncyorh}{\f@nch@orh}{}{}
-%\makeatother
-
+%\usepackage{kpfonts}
+%\usepackage{babel}
+%\usepackage{csquotes}
 \usepackage{url}
-% Extra options: 'sigleft' for a left-aligned signature, 'stdletternofrom' to remove the from address, 'letterpaper' for US letter paper - consult the newlfm class manual for more options
-
 \usepackage{charter} % Use the Charter font for the document text
-
-\newsavebox{\Luiuc}\sbox{\Luiuc}{\parbox[b]{1.75in}{\vspace{0.5in}
-\includegraphics[width=1.2\linewidth]{amu.jpg}}
-{\hspace{3.5in}
-\includegraphics[width=0.3\linewidth]{CNRS-Logo.png}}
-} % Company/institution logo at the top left of the page
-\makeletterhead{Uiuc}{\Lheader{\usebox{\Luiuc}}}
-
-\newlfmP{sigsize=10pt}% Slightly decrease the height of the signature field
-% Print a phone number under the sender's address
-%\newlfmP{addrfromphone} % Print a phone number under the sender's address
-\newlfmP{addrfromemail} % Print an email address under the sender's address
-%\PhrPhone{Phone} % Customize the "Telephone" text
-\PhrEmail{Email} % Customize the "E-mail" text
-
-\lthUiuc % Print the company/institution logo
-
 %----------------------------------------------------------------------------------------
@@
 %----------------------------------------------------------------------------------------
+%Jean-Nicolas Jérémie\\
+%Institute of Neuroscience of la Timone\\
+%CNRS - Aix Marseille University\\
+%27, boulevard Jean Moulin\\
+%13005 Marseille, France
+\newcommand{\LastName}{Grimaldi}%{Perrinet}%
+\newcommand{\FirstName}{Antoine}%{Laurent U.}%
+\newcommand{\Institute}{Institut de Neurosciences de la Timone, CNRS / Aix-Marseille Universit\'e}%
+\newcommand{\Address}{27, Bd. Jean Moulin, 13385 Marseille Cedex 5, France}%
+\newcommand{\Website}{\url{https://laurentperrinet.github.io/}}%
+\newcommand{\Email}{\url{laurent.perrinet@univ-amu.fr}}%
+%----------------------------------------------------------------------------------------
+%	YOUR DOCUMENT
+%----------------------------------------------------------------------------------------
+\usepackage[document]{ragged2e}
+\begin{document}
+%\centerline{
+\includegraphics[width=.4\textwidth]{~/metagit/blog/perrinet_curriculum-vitae_tex/troislogos.png}
+%}
+\\
+\vspace{.1\baselineskip}
+\hrulefill
+\vspace{.1\baselineskip}
 
-\namefrom{\textit{The authors}, Antoine Grimaldi and Laurent Perrinet } % Name
-
-\addrfrom{
-\today\\[12pt]
-Laurent Perrinet\\
-Institute of Neuroscience of la Timone\\
-CNRS - Aix Marseille University\\
-27, boulevard Jean Moulin\\
-13005 Marseille, France
-}
-
-%\phonefrom{(000) 111-1111} % Phone number
-
-\emailfrom{laurent.perrinet@univ-amu.fr} % Email address
-
+\begin{flushright}
+	\FirstName\  \LastName\  \\
+	\Institute\\[6pt]
+	\Address\\%[6pt]
+	\Website \\
+	  Email: \Email \\[6pt]
+\end{flushright}
+\justifying
+\vspace{1\baselineskip}
 %----------------------------------------------------------------------------------------
@@
 %----------------------------------------------------------------------------------------
+%À qui de droit,\\
+Marseille, \today\\[12pt] % Date
 
-\greetto{Dear editor,} % Greeting text
-\closeline{Sincerely yours,} % Closing text
-
-%\nameto{Victor Boutin} % Addressee of the letter above the to address
-
-%\addrto{
-%Neural Network Editorial Board
-%Recruitment Officer \\ % To address
-%The Corporation \\
-%123 Pleasant Lane \\
-%City, State 12345
-%}
-
-%----------------------------------------------------------------------------------------
-
-\begin{document}
-\begin{newlfm}
+Dear editor,
 
@@
 By using heterogeneous delays on different synapses, this novel spiking neuron method is able to detect spiking motifs and we validate" the method on synthetic and then realistic data.
-Results show that this method could provide a path for future spiking neural network algorithms using less energy for a similar performance as thier analog counterparts.
+Results show that this method could provide a path for future spiking neural network algorithms using less energy for a similar performance as their analog counterparts.
 
 Our  manuscript follows our submission and presentation during the ``NeuroVision'' workshop.
-In addition, all the python code needed to reproduce figures and supplementary materials will be fully open-sourced once the article will be published.
+In addition, all the python code needed to reproduce figures and supplementary materials will be fully open-source
 
-\end{newlfm}
+%\vspace{.2\baselineskip}
+%----------------------------------------------------------------------------------------
+Sincerely yours,%\vspace{.1cm}
+
+\FirstName\ \LastName \\
+
+%\includegraphics[width=.3\linewidth]{../RTC/signature_LuP.jpg}
+
 \end{document}

### 2022-10-11 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Importantly, each neuron fires with great variability, and up to 300 spikes per second.
-Many attempts have been made to build artificial spiking neural networks, and in particular to achieve the same performance as observed in classical CNNs with a lower energy budget.
+Many attempts have been made to build artificial spiking neural networks (SNNs), and in particular to achieve the same performance as observed in classical CNNs with a lower energy budget.
 However, none of these networks has yet surpassed the state of the art performances.
@@
 %%%-----------------------------------------------------------------
-Let us first review some existing spiking neural network models.
+Let us first review some existing SNN models.
 A first motivation for these models was to understand biological observations at the different scales of the central nervous system.
@@
 It predicts the emergent states of a recurrently connected neural network~\citep{brunel_phase_2000}.
-Several models have increased the scale of the system to include more neurons, such as The Virtual Brain (TVB)~\citep{sanz_leon_virtual_2013}, which mixes different scales and includes spiking neural networks with (analog) neural mass models.
+Several models have increased the scale of the system to include more neurons, such as The Virtual Brain (TVB)~\citep{sanz_leon_virtual_2013}, which mixes different scales and includes SNNs with (analog) neural mass models.
 However, these methods are mainly descriptive with applications in neuroscience but few spiking neural models have had an application in computer vision.
-Another class of approaches have designed normative models of spiking neural networks that aim to have applications in machine learning.
+Another class of approaches have designed normative models of SNNs that aim to have applications in machine learning.
 One of these is the SpikeNet algorithm which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
-Another type of spiking neural network using precise spike timing attempts to determine the structure of the network in order to minimize a cost function.
+Another type of SNN using precise spike timing attempts to determine the structure of the network in order to minimize a cost function.
 This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extanded in novel gradient-based methods.
@@
 This paper explores the possibility that prototypical spiking motifs may be a useful representation used in neurobiology and that we may transfer this hypothesis to neuromorphic algorithms.
-Our spiking neural network model starts as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
+Our SNN model starts as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
 Here, we propose to extend our model to a layer of spiking neurons which include heterogeneous delays in addition to weights.
@@
 Mimicking the retina, this representation codes for proportional increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events.
-This will constitute the input to the spiking neural network.}
+This will constitute the input to the SNN.}
     \label{fig:motion_task}
@@
 %%%-----------------------------------------------------------------
-To recap, we have introduced in this paper a spiking neural network model with heterogeneous delays that we have trained and evaluated on a complex motion detection task.
+To recap, we have introduced in this paper a SNN model with heterogeneous delays that we have trained and evaluated on a complex motion detection task.
 The model was defined to optimally detect event-driven spatio-temporal motifs.

### 2022-10-11 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 We first detail the methodology by defining the basic mechanism of spiking neurons that utilize heterogeneous delays.
-This will allow us to formalize the spike neuron used to learn the model's parameters and test its effectiveness.
+This will allow us to formalize the spiking neuron used to learn the model's parameters and test its effectiveness.
 We aim to apply this model to a realistic motion detection task which we will first define.
@@
 %
-As in a neurobiological recording or in the sensory signal representing the output of an event-based camera,  any generic raster plot consists a stream of \emph{spikes} (see Figure~\ref{fig:model}-(a)).
+As in a neurobiological recording or in the sensory signal representing the output of an event-based camera, any generic raster plot consists in a stream of \emph{spikes} (see Figure~\ref{fig:model}-(a)).
 This can be formalized as an ordered set $\event$ of addresses and timestamps $\{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event.
@@
 A more detailed description of these synaptic contacts reveals the added importance of heterogeneous delays, i.e., the precise timing from one afferent neuron's firing to its arrival to the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}.
-In that heterogeneous description, we can parameterize each neuron by the set of tuples defining both the weight and the delay of each synaptic contact.
-As a consequence, a set of input presynaptic spikes $\event$ will be processed by the dendrite defined by this synaptic set and notably by the respective delays, which will multiplex in time all events.
+In that description, we can parameterize each neuron by the set of tuples defining both the weight and the delay of each synaptic contact.
+As a consequence, a set of input presynaptic spikes $\event$ will be processed by the dendrites defined by this synaptic set and notably by the respective delays, which will multiplex in time all events.
 

### 2022-10-12 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree, the so-called weight matrix.
-A more detailed description of these synaptic contacts reveals the added importance of heterogeneous delays, i.e., the precise timing from one afferent neuron's firing to its arrival to the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}.
+A more detailed description of these synaptic contacts reveals the added importance of heterogeneous delays, i.e., the precise timing from one afferent neuron's firing to its arrival in the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}.
 In that description, we can parameterize each neuron by the set of tuples defining both the weight and the delay of each synaptic contact.
@@
 This defines the full connectivity of the heterogeneous model.
-As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic address space is transformed into a new event stream $\event_\postsynaddrspace$ which is defined by
+As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic address space is transformed by the synapse $\ranksyn$ of neuron $\postsynaddr$ into a new event stream $\event_\presynaddrspace_\ranksyn$ which is defined by
 $\{ (\presynaddr_\arank = 
@@
 \synapticdelay_\ranksyn^\postsynaddr )_{ \arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]} )
-\}$. \note{formaliser le multiplexage}
-When post-synaptic neurons are activated on their soma by this discharge motif the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.
+\}$. This new stream of events \note{formaliser le multiplexage}
+When post-synaptic neurons are activated on their soma by this spatio-temporal motif the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.
 
@@
 Indeed, any spike in the presynaptic address space is generated by sensory neurons (for instance photoreceptors in the retina, sensors in a CMOS chip) or by afferent spiking neurons.
-In the latter case, these are connected to the spiking cell by a set of weights and delays, whose structure is relatively stable relative to the coding time-scale.
+In the latter case, these are connected to the spiking cell by a set of weights and delays, whose structure is stable relatively to the coding timescale.
 When these connections are high and sparsely distributed, this firing will cause a specific temporal motif.
-Another example is given for the barn owl auditory system: As it hears the sound of a mouse, this sound will generate a specific spiking response in both ears, and specifically, the precise timing between the signal generated by the the left relative to the right ear can be for instance used to determine the position of the prey~\citep{goodman_spike-timing-based_2010}.
+Another example is given for the barn owl auditory system: As it hears the sound of a mouse, this sound will generate a specific spiking response in both ears, and specifically, the precise timing between the signal generated by the left relative to the right ear can be for instance used to determine the position of the prey~\citep{goodman_spike-timing-based_2010}.
 Overall, these examples show that raster plots may be considered as a mixture of the effects of different elementary causes, and that each event triggers a specific spatio-temporal motif.

### 2022-10-12 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner.
-It is important to note that we weighted the cost function by considering only active cells, such that the error is only backpropagated to the spatial locations of these most active cells.
+It is important to note that we weighted the cost function by considering only active cells, such that the error is only back propagated to the spatial locations of these most active cells.
 This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier_unsupervised_2007}.

### 2022-10-25 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 
+\todo{try also a quantization of the weights (binary?)}
 %
@@
 Let us highlight some innovations in the contributions presented in this paper.
-First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs, whereas~\citet{ghosh_spatiotemporal_2019} use a correlation-based heuristic, which we have observed to be less efficient.
+First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs, whereas~\citet{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we have observed to be less efficient.
 Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking motif.

### 2022-10-28 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %Call for papers: Biological Cybernetics Special Issue: What can Computer Vision learn from Visual Neuroscience?
+%
+% https://www.springer.com/journal/422/updates/20421374
 %
@@
 This could raise important questions about the reliability of the results of any automated diagnostic model.
-More generally, these models do give an output for any possible input but they lack an explanation as to why they produced it.
-This lack of explainability is a hindrance to the advancement of these methodologies as well as to their acceptance for most safety-relevant tasks.
+More generally, these models do give an output for any possible input, but they lack an explanation as to why they produced it.
+This lack of explainability is a hindrance to the advancement of these methodologies, as well as to their acceptance for most safety-relevant tasks.
 Having said that, we may remember that these tools are inherited from theoretical models originally designed to understand the brain.
@@
 
-The challenges are numerous in such situations and two major obstacles must be highlighted: the computational delay and the energy budget.
+The challenges are numerous in such situations, and two major obstacles must be highlighted: the computational delay and the energy budget.
 First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun_gradient-based_1998}, then high definition images, such as VGG on Imagenet~\citep{simonyan_very_2015}.
@@
 Many attempts have been made to build artificial spiking neural networks (SNNs), and in particular to achieve the same performance as observed in classical CNNs with a lower energy budget.
-However, none of these networks has yet surpassed the state of the art performances.
-Here, we will try to focus on one core property of spiking neurons inspired by neurobiology to progress in this direction.
+However, none of these networks has yet surpassed the state-of-the-art performances.
+Here, we will try to focus on one core property of spiking neurons, inspired by neurobiology to progress in this direction.
 %
@@
 Another class of approaches have designed normative models of SNNs that aim to have applications in machine learning.
-One of these is the SpikeNet algorithm which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
+One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
 Another type of SNN using precise spike timing attempts to determine the structure of the network in order to minimize a cost function.
@@
 %%%-----------------------------------------------------------------
-We have identified several limits to our model which we will now detail.
+We have identified several limits to our model, which we will now detail.
 First, the complete framework is based on a discrete binning of time that is incompatible with the continuous nature of biological time.
@@
 However, the generic theoretical model is able to include them by inserting new spikes into the list of spikes that are reaching pre-synaptic addresses.
-While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside of a homeostatic state (through a extinction or explosion of activity).
+While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside a homeostatic state (through an extinction or explosion of activity).
 \note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}%~\citep{perrinet_motion-based_2012}
@@
 Such recurrent activity would be essential for implementing predictive or anticipatory processes.
-This is essential in a neural system as it contains multiple different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}.
+This is essential in a neural system, as it contains multiple different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}.
 This has been previously modeled to explain for example the flash-lag illusion~\citep{khoei_flash-lag_2017}.
@@
 Studying this dependence will allow us to dissociate these two forms of visual information and to enable us to integrate them.
-In particular it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka, the aperture problem), thus allowing us to implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
+In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka, the aperture problem), thus allowing us to implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
 

### 2022-11-10 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 \newcommand{\DataAvailability}{
-This works is made reproducible using the following tools. First the code reproducing all figures is available on~\href{ https://github.com/SpikeAI/pyTERtorch}{GitHub}.
+This works is made reproducible using the following tools. First the code reproducing all figures is available on~\href{ https://github.com/SpikeAI/pyTERtorch}{GitHub}. % todo: make a new repo
 
@@
 The challenges are numerous in such situations, and two major obstacles must be highlighted: the computational delay and the energy budget.
-First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun_gradient-based_1998}, then high definition images, such as VGG on Imagenet~\citep{simonyan_very_2015}.
+First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun_gradient-based_1998}, then high definition images, such as VGG on ImageNet~\citep{simonyan_very_2015}.
 This has a cost and notably the fact that the computation time increases with the size of the image.
@@
 However, the transition from images to videos, then to multiple cameras implies a further level in the change of scale in the amount of data processed.
-It also raises an issue that is often not addressed when it comes to image categorization.
+It also raises an issue that is often not addressed when it comes to image categorization, that is, of the right timing to take a decision.
 Indeed, in autonomous driving, an important task is to react as quickly as possible to a possible collision, for example when a pedestrian unexpectedly crosses the street.
@@
 In addition, the need to solve increasingly complex tasks has required the construction of increasingly complex and deep architectures.
-As an example, the best performing model in the Imagenet challenge uses 2100 million parameters.
+As an example, the best performing model in the ImageNet challenge uses 2100 million parameters.
 Training the model typically consumes kilowatts of power and about 1 joule to infer a decision per single image.

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
methods

@@
 % LP received funding 
-and from the ANR project % N° ANR-22-XXX-XXX 
+and from A*Midex grant number AMX-21-RID-025
 ``\href{PolyChronies}{https://laurentperrinet.github.io/grant/polychronies/}''. 
@@
 %%%-----------------------------------------------------------------
-Let us now formally define the heterogeneous model, as well as the task and how we will learn to solve it.
+Let us now formally define the spiking model with  heterogeneous delays, as well as the task and how we will learn to solve it.
 First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays.
@@
     \centering
-    \includegraphics[width=0.490\linewidth]{figures/figure_1a_k.pdf}
-    \includegraphics[width=0.490\linewidth]{figures/figure_1c.pdf}
-    \includegraphics[width=0.490\linewidth]{figures/figure_1b.pdf}
-    \includegraphics[width=0.490\linewidth]{figures/figure_1a.pdf}
+    \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
+    \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
+    \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
+    \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
 	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
 	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs).
-	    {\bf (b)}~The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
-	    {\bf (c)}~We will define a layer of neurons that will detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).
+	    {\bf (b)}~We show these motifs, each identified at the top by a different color.
+Each neurons and $71$ different possible delays is assigned the evidence of activation (red) or deactivation (blue).
+%The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
+	    {\bf (c)}~The activation in time of the different motifs is then used to define a generative model for drawing a raster plot on the multi-unit address space.
+        % We will define a layer of neurons that will detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).
 	    Note that each afferent may be connected with multiple weights at different delays.
@@
 %These PG motifs were efficiently learned by a logistic regression.
-	    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
+	    % {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
 	    {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG.
+The original raster plot can be annotated with each identified spiking motif.
 	     }
@@
     \centering
-    \includegraphics[width=0.490\linewidth]{figures/figure_N_PGs.pdf}
-    \includegraphics[width=0.490\linewidth]{figures/figure_N_PG_time.pdf}
+    \includegraphics[width=0.490\linewidth]{figures/THC_N_PGs.pdf}
+    \includegraphics[width=0.490\linewidth]{figures/THC_N_PG_time.pdf}
 	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
methods

@@
 %%%-----------------------------------------------------------------
-Let us now formally define the spiking model with  heterogeneous delays, as well as the task and how we will learn to solve it.
+Let us now formally define the spiking neural network (SNN) model with heterogeneous delays, as well as the task and how we will learn to solve it.
 First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays.
-Our SNN model will be tested on a computationally relevant task of motion detection in an event-based sensory input.
+Our SNN model will be tested on a computationally relevant task of detecting motion detection from an event-based sensory input.
 By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory.
@@
 	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
-	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs).
-	    {\bf (b)}~We show these motifs, each identified at the top by a different color.
-Each neurons and $71$ different possible delays is assigned the evidence of activation (red) or deactivation (blue).
-%The generative model is defined as a set of motifs (here $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay.
+	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``spiking motifs''.
+	    {\bf (b)}~We show some randomly drawn motifs, each identified at the top by a different color.
+To each neurons and $71$ different possible delays is assigned an evidence of activation (red) or deactivation (blue).
 	    {\bf (c)}~The activation in time of the different motifs is then used to define a generative model for drawing a raster plot on the multi-unit address space.
@@
 This defines the full connectivity of the heterogeneous model.
-As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic address space is transformed by the synapse $\ranksyn$ of neuron $\postsynaddr$ into a new event stream $\event_\presynaddrspace_\ranksyn$ which is defined by
+As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic address space is transformed by the synapse $\ranksyn$ of neuron $\postsynaddr$ into a new event stream $\event_{\presynaddrspace, \ranksyn}$ which is defined by
 $\{ (\presynaddr_\arank = 
@@
 
-\todo{try also a quantization of the weights (binary?)}
+\note{try also a quantization of the weights (binary?)}
 %

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
methods

@@
 	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
-	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``spiking motifs''.
-	    {\bf (b)}~We show some randomly drawn motifs, each identified at the top by a different color.
-To each neurons and $71$ different possible delays is assigned an evidence of activation (red) or deactivation (blue).
+	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of precise spiking motifs.
+	    {\bf (b)}~We show the motifs used in this example, each identified at the top by a different color.
+To each of the $10$ synapses and $71$ different possible delays is assigned an evidence of activation (red) or deactivation (blue).
 	    {\bf (c)}~The activation in time of the different motifs is then used to define a generative model for drawing a raster plot on the multi-unit address space.
-        % We will define a layer of neurons that will detect these PGs such that the addresses are the synapses of these neurons (which we therefore call here synapses).
 Note that each afferent may be connected with multiple weights at different delays.
 The propagation of the afferent information through these delay may generate at each time step a synchronous motif on a subset of synapses.
-%These PG motifs were efficiently learned by a logistic regression.
-	    % {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
-	    {\bf (d)}~Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes are now associated with a PG.
-The original raster plot can be annotated with each identified spiking motif.
+	    {\bf (d)}~The output of the model provides with the predicted probability of occurrence of each motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of the occurrence of motifs.
+Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes can be annotated with each identified spiking motif.
 	    }

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
2.1

@@
 	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
-	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of precise spiking motifs.
+	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific neural addresses and at specific times, one may consider that this information consists of the repeated occurrence of precise spiking motifs.
 	    {\bf (b)}~We show the motifs used in this example, each identified at the top by a different color.
-To each of the $10$ synapses and $71$ different possible delays is assigned an evidence of activation (red) or deactivation (blue).
-	    {\bf (c)}~The activation in time of the different motifs is then used to define a generative model for drawing a raster plot on the multi-unit address space.
+To each of the $10$ neural adresses and $71$ different possible delays is assigned an evidence of activation (red) or deactivation (blue).
 Note that each afferent may be connected with multiple weights at different delays.
-The propagation of the afferent information through these delay may generate at each time step a synchronous motif on a subset of synapses.
-	    {\bf (d)}~The output of the model provides with the predicted probability of occurrence of each motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of the occurrence of motifs.
+	    {\bf (c)}~The activation in time of the different motifs is then used to define a generative model for drawing a raster plot on the multi-unit address space: The propagation of the afferent information through these delays generates the raster plot seen in (a).
+	    {\bf (d)}~The inversion of the generative model provides with a model which gives the predicted probability of occurrence of each motif at any time and which may be used to generate a spike as a Bernoulli trial.
+The detection model yields in this particular case with an exact identification of the occurrence of motifs.
 Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes can be annotated with each identified spiking motif.
@@
 %
-As in a neurobiological recording or in the sensory signal representing the output of an event-based camera, any generic raster plot consists in a stream of \emph{spikes} (see Figure~\ref{fig:model}-(a)).
-This can be formalized as an ordered set $\event$ of addresses and timestamps $\{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event.
-That event has a time of occurrence $\timev_\arank$  and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
+In neurobiological recordings or in the sensory signal obtained from an event-based camera, any generic raster plot consists of a stream of \emph{spikes} (see Figure~\ref{fig:model}-(a)).
+This can be formalized as a list (an ordered set) of neural addresses and timestamps  $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events.
+Each event has a time of occurrence $\timev_\arank$  and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
 This defines an address space $\presynaddrspace$ which consists of the set of possible addresses.
@@
 
-Formally, let's define a layer of heterogeneous spiking neurons $\postsynaddr \in \postsynaddrspace$ by first describing how each neuron connects to presynaptic afferent from $\presynaddrspace$.
+Let's formalize a layer of spiking neurons with heterogeneous delays (HD-SNN).
+Each neuron $\postsynaddr \in \postsynaddrspace$  connects to presynaptic afferent from $\presynaddrspace$.
 In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
2.2

@@
 
-This raster plot is generated by spiking neurons which are defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree, or on the sensory processes (such as the pixel on the sensor of a camera, or the photoreceptors in the retina).
+% This raster plot is generated by spiking neurons which are defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree, or on the sensory processes (such as the pixel on the sensor of a camera, or the photoreceptors in the retina).
 A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree, the so-called weight matrix.
@@
 Each neuron $\postsynaddr \in \postsynaddrspace$  connects to presynaptic afferent from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses, and each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+In biology, a single cortical neuron has generally several thousands of synapses.
+Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
 A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
@@
 For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
-This defines the full connectivity of the heterogeneous model.
+This defines the full connectivity of the HD-SNN model.
 As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic address space is transformed by the synapse $\ranksyn$ of neuron $\postsynaddr$ into a new event stream $\event_{\presynaddrspace, \ranksyn}$ which is defined by
@@
 
+\note{ TODO: show why this formula implements a polychrony detector }
+
+\note{ éclaircir }
+
 Taking the argument the other way around, one may form a generative model for generic raster plots.
@@
 Another example is given for the barn owl auditory system: As it hears the sound of a mouse, this sound will generate a specific spiking response in both ears, and specifically, the precise timing between the signal generated by the left relative to the right ear can be for instance used to determine the position of the prey~\citep{goodman_spike-timing-based_2010}.
-Overall, these examples show that raster plots may be considered as a mixture of the effects of different elementary causes, and that each event triggers a specific spatio-temporal motif.
+Overall, these examples show that raster plots may be considered as a mixture of the effects of different elementary causes, and that each event triggers a specific spatio-temporal spiking motif.
 %
@@
 
-\note{ TODO: show why this formula implements a polychrony detector }
-From the perspective of simulating such event-based computations on standard chips, it is useful to transform this sparse representation into a dense representation.
-As such, we may first write any event-based input as the boolean matrix $A \in \{0, 1 \}^\presynaddrspace$.
+From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers, it is useful to transform this event-based representation into a dense representation.
+Indeed, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{N\times T}$, where $N$ is the number of neurons in $\presynaddrspace$ and $T$ is the number of time bins.
 In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
 The probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
-Assuming that the presence of PGs conditions the probability on all efferents, the logit (inverse of the sigmoid) of this probability bias can be written as the sum of the logit of each of these factors, whose values are given by the corresponding weights.
-PGs may be activated independently at random times and  we write that $B(b, t)=1$ if $b$ is activated at $t$ (and else $B(b, t)=0$).
+Assuming that the presence of spiking motifs conditions the probability on all efferents, the logit (inverse of the sigmoid) of this probability bias can be written as the sum of the logit of each of these factors, whose values are defined by the corresponding weights.
+Spiking motifs may be activated independently and at random times, such that  we write this activity as $B(b, t)=1$ if $b$ is activated at $t$ and else $B(b, t)=0$.
 We can thus write the probability bias as the accumulated evidence given these factors as
@@
 
+% 
+This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$.
+This assumes that we know the spiking motifs as defined by the $W_b$ matrices.
+The underlying metric is the binary cross-entropy, as used in the logistic regression model.
+In particular, if we consider kernels with similar decreasing exponential time profile, we can prove that this is similar to the method of~\citet{berens_fast_2012}.
+In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator.
+Using this forward model, it is possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator.
+It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.
+
 %\subsubsection{Implementation using spiking neurons with heterogeneous delays}
 %
-The activation function of our spiking neural is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
+% The activation function of our spiking neural is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
 %It transforms this list of weights into a probability with the following formula:
@@
 %
-\note{say that it is effortless in biology but difficult on conventional computers}
-In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+% \note{say that it is effortless in biology but difficult on conventional computers}
+% In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
 The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the heterogeneous delays model.
@@
 	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
-	    {\bf (a)}~Accuracy of PG detection as a function of the number $M$ of kernels.
-	    {\bf (b)}~Accuracy of PG detection as a function of the temporal depth $D$ of kernels among $M=144$ kernels.
+	    {\bf (a)}~Accuracy of detection as a function of the number of kernels.
+	    {\bf (b)}~Accuracy of detection as a function of the temporal depth $D$ of kernels among $M=144$ kernels.
 	    }
@@
 \end{figure}
-% 
-This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$.
-This assumes that we know the PGs as defined by the $W_b$ matrices.
-The underlying metric is the binary cross-entropy, as used in the logistic regression model.
-In particular, if we consider PG kernels with similar decreasing exponential time profile, we can prove that this is similar to the method of~\citet{berens_fast_2012}.
-In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator.
-Using this forward model, it is possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator.
-It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.
 
-To quantify the efficiency of this operation, we generated $M=55$ synthetic PGs as random independent kernels over $128$ presynaptic inputs and $D=127$ possible delays.
-We drew random independent instances of $B$ with a length of $T=1000$ time steps and with on average $2.0$ occurrences each.
+To quantify the efficiency of this operation, we generated $M=144$ synthetic spiking motifs as random independent kernels over $128$ presynaptic inputs and $D=71$ possible delays.
+We drew random independent instances of $B$ with a length of $T=1000$ time steps and with on average $1.0$ occurrences in each draw.
 This allowed us to generate raster plots which we use to infer $\hat{B}$.
@@
 We show in Figure~\ref{fig:model_results}~(a) that the accuracy of finding the right PG is still above $80\%$ accuracy with more than $1364$ overlapping PGs.
-Moreover, we show in Figure~\ref{fig:model_results}~(b) that (with $M=55$ PGs fixed) the accuracy increases notably as the temporal depth $D$ of the PG kernel increased, demonstrating quantitatively the potential of heterogeneous delays.
+Moreover, we show in Figure~\ref{fig:model_results}~(b) that (with $M=144$ PGs fixed) the accuracy increases notably as the temporal depth $D$ of the PG kernel increased, demonstrating quantitatively the computational advantage of using heterogeneous delays.
 These results were obtained while assuming that we know $W$.
 However, this is in general not the case, for instance when observing the raster plot of a population of neurons.
-In the following, we will define a generic visual task and determine a learning algorithm to solve it.
+In the following, we will define a generic visual task and determine a learning algorithm to solve that challenging task.
 %Inspired by the k-means algorithm, it is possible to devise a self-supervised learning algorithm.

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
2.3 drawing motion_task from 2022-11-10_MotionDetection

@@
 In animals with a fovea, this is particularly useful as it allows to move the area with the highest density of photoreceptors in the environment, for example at a point of interest.
-For example, saccades are rapid movements of the eyeball that reposition the center of vision.
-In humans, these are very frequent (on average one every 3 per second~\citep{dandekar_neural_2012}).
+A specific mechanism to do so are called saccades which are rapide eye movements of the eyeball that reposition the center of vision.
+In humans, these are very frequent (on average one to 3 each second~\citep{dandekar_neural_2012}).
 They are produced very rapidly (about $80~\ms$) and display a whole range of speeds.
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{figures/motion_clouds.pdf}
+    \includegraphics[width=0.95\linewidth]{2022-09-27_MotionDetection_motion_clouds.png}%motion_clouds.pdf}
     \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of {\bf (a)} the mean spatial frequency, {\bf (b)} the bandwidth in orientation: from a grating-like (left) to isotropic textures (right)), {\bf (c)} the bandwidth in speed, from a rigid motion (left) to independent frames (right).

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
2.1

@@
 \newcommand{\presynaddr}{a} % pre address
-\newcommand{\postsynaddr}{n} % post address
+\newcommand{\postsynaddr}{b} % post address
 \newcommand{\numevent}{N_{ev}} % total number of events
@@
 \newcommand{\bias}{b} % bias for the MLR model
-\newcommand{\synapse}{\sigma} % synapse
+\newcommand{\synapse}{\mathcal{S}} % synapse
 \newcommand{\synapticweight}{w} % synaptic weight
@@
 In neurobiological recordings or in the sensory signal obtained from an event-based camera, any generic raster plot consists of a stream of \emph{spikes} (see Figure~\ref{fig:model}-(a)).
-This can be formalized as a list (an ordered set) of neural addresses and timestamps  $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events.
+This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events.
+Events are typically ordered by their time of occurrence.
 Each event has a time of occurrence $\timev_\arank$  and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
 This defines an address space $\presynaddrspace$ which consists of the set of possible addresses.
-This can be the identified set of neurons in a neurobiological recording.
+In a neurobiological recording, this can be the identified set of neurons.
 For event-based cameras, it is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras).
@@
 Note that a neuron may contact an afferent neuron with multiple different delays.
-For each neuron $\postsynaddr$, we thus define a set of $\Nsyn^\postsynaddr$ synapses, as  $\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.
+Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
+as  $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a presynaptic address $\postsynaddr_\ranksyn$,  a weight $\synapticweight^\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
 This defines the full connectivity of the HD-SNN model.
-As a consequence, an event stream $\event_\presynaddrspace$ on the presynaptic address space is transformed by the synapse $\ranksyn$ of neuron $\postsynaddr$ into a new event stream $\event_{\presynaddrspace, \ranksyn}$ which is defined by
-$\{ (\presynaddr_\arank = 
- \presynaddr_\ranksyn^\postsynaddr, \timev=\timev_\arank + 
-\synapticdelay_\ranksyn^\postsynaddr )_{ \arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]} )
-\}$. This new stream of events \note{formaliser le multiplexage}
-When post-synaptic neurons are activated on their soma by this spatio-temporal motif the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.
+Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]} $, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]}$.
+As a consequence, an event stream which evokes neurons in the presynaptic address space is multiplexed by the synapses into a new event stream which is defined by the union of the sets generated by each emitting field from the presynaptic space:
+$ \cup_{\arank \in [1,\numevent]} \{ \{(\postsynaddr_\ranksyn, \synapticweight_\ranksyn, \timev_\arank + \synapticdelay_\ranksyn) \}_{ \ranksyn \in \synapse_{\presynaddr_\arank}} \}$.
+This new stream of events is by nature ordered in time as events reach the soma of post-synaptic neurons.
+In particular, when post-synaptic neurons are activated on their soma by this spatio-temporal motif, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.
+% \note{ TODO: show why this formula implements a polychrony detector }
 
-\note{ TODO: show why this formula implements a polychrony detector }
+% \note{ éclaircir }
 
-\note{ éclaircir }
-
-Taking the argument the other way around, one may form a generative model for generic raster plots.
+Taking the argument the other way around, one may from a generative model for generic raster plots.
 Indeed, any spike in the presynaptic address space is generated by sensory neurons (for instance photoreceptors in the retina, sensors in a CMOS chip) or by afferent spiking neurons.
@@
 Finally, we obtain the raster plot $A\in \{0, 1\}^{N\times T}$ by drawing spikes using independent Bernoulli trials $A \sim \mathcal{B}(p)$.
-Note that, depending on the shape of the kernels, the generative model can model a Poisson process, generate rhythmic activity or more generally propagating waves.
+Note that, depending on the shape of the kernels, the generative model can model a discretized Poisson process, generate rhythmic activity or more generally propagating waves.
 This formulation thus defines a simple generative model for raster plots as a combination of independent PGs.
@@
 Using this forward model, it is possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator.
+Equivalently, this consists in using the emitting field $\synapse_\presynaddr$ of presynaptic neurons in place of the receptive field $\synapse^\postsynaddr$ of postsynaptic neurons.
 It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.
+This assumption hold as long as the kernels are uncorrelated, a condition which is met here numerically by chosing a relatively sparse set of synapses (approximately $1\%$ of active synapses).
 
@@
 This method provides a rich dataset of textured movies for which we know the ground truth for motion.
-
 %%%-----------------------------------------------------------------
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{2022-09-27_MotionDetection_motion_clouds.png}%motion_clouds.pdf}
+    \includegraphics[width=0.95\linewidth]{figures/motion_clouds.pdf} % 2022-09-27_MotionDetection_motion_clouds.png}%
     \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of {\bf (a)} the mean spatial frequency, {\bf (b)} the bandwidth in orientation: from a grating-like (left) to isotropic textures (right)), {\bf (c)} the bandwidth in speed, from a rigid motion (left) to independent frames (right).
@@
     \includegraphics[width=0.95\linewidth]{figures/accuracy.pdf}
-    \caption{Accuracy as a function of the number of computation load for the heterogeneous delays model (blue line) and for a method using 2D time surfaces (red dot)~\citep{grimaldi_robust_2022}.
+    \caption{Accuracy as a function of the number of computation load for the model with heterogeneous delays (blue line) and for a method using 2D time surfaces (red dot)~\citep{grimaldi_robust_2022}.
 The relative computational load (on a log axis) is controlled by changing the percentage of active weights relative to the dense convolution kernel.
@@
 \paragraph{Code availability}
-
+%
 The code is publicly available online at: \url{https://github.com/SpikeAI/pyTERtorch}.
-
+%   
 \paragraph{Authors' contributions}
-
+%
 Both authors contributed to the conceptualization and methodology design of the study, to the project's coordination and administration.

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
pass

@@
 However, most artificial neuronal models do not take advantage of this minute temporal dimension.
-Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of neurons with heterogeneous delays which we apply to the computer vision task of motion detection.
+Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous delays which we apply to the computer vision task of motion detection.
 Indeed, the variety of synaptic delays on the dendritic tree allows to synchronize synaptic inputs as they reach the basal dendritic tree.

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
pass: intro

@@
 %%%-----------------------------------------------------------------
-Machine learning, and computer vision in particular, have undergone a major transformation in modern times, transitioning from an active character in science fiction movies to become mainstream tools found in the smartphones in everyone's pocket.
+Machine learning, and computer vision in particular, have undergone a major transformation in modern times, transitioning from an eccentric character in science fiction movies to become mainstream tools found in the smartphones in everyone's pocket.
 Less visible, yet crucial to our economy and security, these automated tools now dominate the range of services that equip virtually most sectors of our daily lives.
@@
 
-The challenges are numerous in such situations, and two major obstacles must be highlighted: the computational delay and the energy budget.
+The challenges are numerous in such situations, and two major obstacles must be highlighted: the computational processing time and the energy budget.
 First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun_gradient-based_1998}, then high definition images, such as VGG on ImageNet~\citep{simonyan_very_2015}.
-This has a cost and notably the fact that the computation time increases with the size of the image.
+This has a cost and notably the fact that the computation time increases with the size (in number of pixels) of the image.
 This phenomenon has been mitigated by the increase in available computing power, notably thanks to the parallel processing implemented in GPU cards.
@@
 As an example, the best performing model in the ImageNet challenge uses 2100 million parameters.
-Training the model typically consumes kilowatts of power and about 1 joule to infer a decision per single image.
+Training the model typically consumes many kilowatts hours of energy and about 1 joule to infer a decision per single image.
 Less power-hungry solutions have emerged, such as Mobile Net~\citep{howard_searching_2019}, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significant percentage of the energy consumed by an electric car under standard conditions.

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
pass: intro

@@
 
-This paper explores the possibility that prototypical spiking motifs may be a useful representation used in neurobiology and that we may transfer this hypothesis to neuromorphic algorithms.
-Our SNN model starts as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
+This paper explores the possibility that prototypical spiking motifs may be a useful representation used in neurobiology and that we may transfer this hypothesis to neuromorphic algorithms, notably by using Heterogeneous Delays in Spiking Neural Networks (HD-SNN).
+Our HD-SNN model starts as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
 Here, we propose to extend our model to a layer of spiking neurons which include heterogeneous delays in addition to weights.
@@
 Crucially, we will explicitly use the delay in the computational process.
-As a consequence, this extends our previous MLR model by encoding spiking motifs as spatio-temporal patterns.
+As a consequence, this extends our previous MLR model~\citep{grimaldi_robust_2022} by explictly encoding spiking motifs as spatio-temporal patterns.
 The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier.
@@
 The paper is organized as follows.
-We develop a theoretically defined SNN capable of learning heterogeneous delays from spiking motifs that we first evaluate on a synthetic event-based dataset.
+We develop a theoretically defined HD-SNN capable of learning heterogeneous delays that we first evaluate on a synthetic event-based dataset.
 We first detail the methodology by defining the basic mechanism of spiking neurons that utilize heterogeneous delays.
@@
 This will open perspectives for future SNNs.
-In particular, as neuromorphic devices are by design good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is perfectly fit to be transfered to this type of hardware.
+In particular, as neuromorphic devices are by design good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is perfectly fit to be transfered to this type of hardware and to obtain significant gains in the energy which is used.
 %

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
pass: results

@@
 %%%-----------------------------------------------------------------
-Let us now formally define the spiking neural network (SNN) model with heterogeneous delays, as well as the task and how we will learn to solve it.
+Let us now formally define the Heterogeneous Delays Spiking Neural Network (HD-SNN) model, as well as the task and how we will learn to solve it.
 First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays.
-Our SNN model will be tested on a computationally relevant task of detecting motion detection from an event-based sensory input.
+Our HD-SNN model will be tested on a computationally relevant task of detecting motion detection from an event-based sensory input.
 By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory.
@@
 The task is thus to guess the motion as accurately as possible (\seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays.
-To this end, we will extend the SNN model to be efficient for the resolution of the task by including spatio-temporal convolutional kernels.
+To this end, we will extend the HD-SNN model to be efficient for the resolution of the task by including spatio-temporal convolutional kernels.
 We will finally describe its implementation on conventional computers.
@@
 A specific mechanism to do so are called saccades which are rapide eye movements of the eyeball that reposition the center of vision.
-In humans, these are very frequent (on average one to 3 each second~\citep{dandekar_neural_2012}).
+In humans, these are very frequent (on average $2$ each second~\citep{dandekar_neural_2012}).
 They are produced very rapidly (about $80~\ms$) and display a whole range of speeds.
@@
     \caption{
-    {\bf Motion detection task.} {\bf (a)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view (the image on the left) such as to reproduce the effect of displacing the eye.
-To mimic the effect of a saccadic-like eye movement, the axis of view moves following a step-like random walk, and we show here on example path of the trajectory as a green line ($200$ time steps).
-{\bf (b)} Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation.
+    {\bf Motion detection task.} {\bf (Left)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view  such as to reproduce the effect of displacing the eye.
+To mimic the effect of a saccadic-like eye movement, the axis of view moves following a step-like random walk, and we show here on example path of the trajectory as a yellow line ($200$ time steps).
+{\bf (Right)}~Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation.
 Mimicking the retina, this representation codes for proportional increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events.
-This will constitute the input to the SNN.}
+This will constitute the input to the SNN.
+Note the change of direction between the second and third frames.}
     \label{fig:motion_task}
@@
 These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step (\seeFig{motion_task}-(b)).
-We will discretize the time in $1~\ms$ bins and produce movies of duration N-T = $250~\ms$.
+We will discretize the time in $1~\ms$ bins and produce movies of duration $N_T = 250~\ms$.
 To avoid border effects, we will position a draw of the complete trajectory at random in the image space so that the sub-image is translated using the position given by the trajectory at each time step.
@@
 
-\note{ TODO: formula / say it is driven by image gradients / give scaling rules }
+% \note{ TODO: formula / say it is driven by image gradients / give scaling rules }
 To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames.
@@
 This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
-\note{ the task is full-field rigid translations, the model resolves it locally (analogy to the input from V1 to MT) }
+% \note{ the task is full-field rigid translations, the model resolves it locally (analogy to the input from V1 to MT) }
 %
@@
 Now that our motion detection task is defined in visual space, we can extend the convolution-based detection model to a 3D convolution so that the resulting model would also benefit from spatial invariance.
-The use of spatio-temporal filters on an event stream has improved the performance of CNNs for an action recognition task by~\citet{ghosh_spatiotemporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.
+The use of spatio-temporal filters on an event stream has improved the performance of CNNs for an action recognition task in a model by~\citet{ghosh_spatiotemporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.
 For this purpose, we design 3D kernels of shape $(\Kx, \Ky, \Ktime) = (9, 9, 11)$, representing the two spatial dimensions and the range of time delays, respectively.
-The computations are performed on the spatio-temporal windows that are defined by the kernels and are "moving" around the events, i.e.
+The computations are performed on the spatio-temporal windows that are defined by the kernels and are ``moving'' around the events, i.e.
 the spatial center of the spatio-temporal window around the current event $\epsilon_\arank$.
@@
     %\vspace{-3cm}
-    \includegraphics[width=\linewidth]{figures/motion_kernels.pdf}
+    \includegraphics[width=\linewidth]{figures/motion_kernels.png}%pdf}
     }
@@
     \caption{
-    	Representation of the weights for $3$ of the $32$ different learned kernels of the model.
+    	Representation of the weights for $3$ of the $32$ different learned kernels of the model as learned on natural scenes.
 Each pair of line correspond to the OFF and ON polarities respectively, with excitatory weights in warm colors.

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
GO!

@@
 %
-\note{kernel size has to be adapted to the characteristics of motion see Samonds %~\citep{samonds}
-say that we have tested }
-\note{tester la meme expe faite par baudot et modelisée par kremkow en montrant des raster plots du modele avec NI versus Gratings }
+% \note{kernel size has to be adapted to the characteristics of motion see Samonds %~\citep{samonds}
+% say that we have tested }
+% \note{tester la meme expe faite par baudot et modelisée par kremkow en montrant des raster plots du modele avec NI versus Gratings }
 
@@
 This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006}.
-\note{we should really do something more serious here...} Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively resemble a more textured stimulus) than for a grating-like stimulation, reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}.
+%\note{we should really do something more serious here...}
+Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively resemble a more textured stimulus) than for a grating-like stimulation, reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}.
 Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
pass: conclusion

@@
     \centering
-    \includegraphics[width=0.95\linewidth]{figures/motion_clouds.pdf} % 2022-09-27_MotionDetection_motion_clouds.png}%
+    \includegraphics[width=0.99\linewidth]{figures/2022-09-27_MotionDetection_motion_clouds.png}%motion_clouds.pdf} % 2022-09-27_MotionDetection_motion_clouds.png}%
     \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of {\bf (a)} the mean spatial frequency, {\bf (b)} the bandwidth in orientation: from a grating-like (left) to isotropic textures (right)), {\bf (c)} the bandwidth in speed, from a rigid motion (left) to independent frames (right).
-Note that these accuracy is computed both in the case is necessarily perpendicular to the motion (no aperture) or arbitrary (aperture), showing that accuracy decreases in the latter case.}
+Note that these accuracy is computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion (no aperture) or arbitrary (aperture), showing that accuracy decreases in the latter case.}
     \label{fig:motion_clouds}
@@
 
-\note{try also a quantization of the weights (binary?)}
+% \note{try also a quantization of the weights (binary?)}
 %
@@
 The learned model bears many similarities with neurobiological anatomical observations but also with the results from behavioral results.
-\note{to change if we compare to ghosh instead of HOTs if we follow that route} The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification.
+% \note{to change if we compare to ghosh instead of HOTs if we follow that route}
+The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification.
 This shows that we may use the precise timing of spikes to enhance neural computations.
@@
 Such an architecture could provide promising speed gains for the calculations.
-\note{pourrait parler du poster bernstein?
+% \note{pourrait parler du poster bernstein?
 de nadafian?}
@@
 While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside a homeostatic state (through an extinction or explosion of activity).
-\note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}%~\citep{perrinet_motion-based_2012}
+% \note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}%~\citep{perrinet_motion-based_2012}
 
@@
 This has been previously modeled to explain for example the flash-lag illusion~\citep{khoei_flash-lag_2017}.
-\note{Say that this could be implemented by properly setting the timing of the label} As previously stated, this could be implemented using generalized coordinates (that is variables such as position which are complemented by speed, acceleration, jerk, ...) and ``neurobiologically, the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
+%\note{Say that this could be implemented by properly setting the timing of the label}
+As previously stated, this could be implemented using generalized coordinates (that is variables such as position which are complemented by speed, acceleration, jerk, \ldots) and ``neurobiologically, the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
 Our proposed model using heterogeneous delays provides an alternate and elegant implementation solution to this problem.
@@
 %%%-----------------------------------------------------------------
-\note{la tache ne met pas en avant les capacites des motifs hetero synaptiques =trop simple et peut se faire en HOTS}
+% \note{la tache ne met pas en avant les capacites des motifs hetero synaptiques =trop simple et peut se faire en HOTS}
 In the definition of our task, we highlighted how the generation of events depends on the spatial gradient in each image.
@@
 This could be initiated by training the model on a dataset with labels providing local orientation.
-\note{we can say we tried and that it works well}.
+%\note{we can say we tried and that it works well}.
 Studying this dependence will allow us to dissociate these two forms of visual information and to enable us to integrate them.

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 $8$ linearly spaced motion directions and %N-Y-velocity=
-$7$ different velocities (see inset of \seeFig{motion_task}-(a)).
+$7$ different velocities (\seeFig{motion_task}-(a)-inset).
 Note that the velocities are sampled on a geometric scale between $1/4$ and $4$.
@@
 Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels.
-This threshold was found using validation and future work should automate how this may be determined for various datasets.
+This threshold was found using cross-validation and future work should automate how this may be determined for various datasets.
 Given a stimulus input, the model thus yields a spike output in the postsynaptic address space.
@@
 %
+Since the model is fully differentiable, we can now implement a semi-supervised learning rule.
+The loss function of the MLR model is the binary cross-entropy on the output of the classification layer that can be extended to space-time convolutions.
+Supervision was implemented using the input binary events as defined above and the labels as the desired output.
+The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions.
+As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner.
+It is important to note that we weighted the cost function by considering only active cells, such that the error is only back propagated to the spatial locations of these most active cells.
+This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier_unsupervised_2007}.
+Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$).
+We tested the effect of different parameters on a validation set to quantify the role of each parameter.
 Since the model is fully differentiable, we can now implement a semi-supervised learning rule.
@@
 As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner.
-It is important to note that we weighted the cost function by considering only active cells, such that the error is only back propagated to the spatial locations of these most active cells.
+It is important to note that we weighted the cost function by considering only the  most active cells, such that the error is only back propagated to the spatial locations of these most active cells.
 This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier_unsupervised_2007}.
@@
 %ith the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
-We also observe that these cells show an orientation selectivity, similar to that observe in MT neurons~\citep{deangelis_functional_1999}.
+We also observe that these cells show an orientation selectivity, similar to that observed in MT neurons~\citep{deangelis_functional_1999}.
 Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results~\citep{baudot_animation_2013}.

### 2022-11-30 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}.
-However, the performance of SNNs is still lagging that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
+However, the performance of SNNs is still lagging that of firing rate-based networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
 
@@
 %
-\subsection{Accuracy for the motion detection task}
+\subsection{Accuracy efficiency trade-off}% for the motion detection task}
 %%%-----------------------------------------------------------------

### 2023-01-18 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 % public
-% https://github.com/SpikeAI/pyTERtorch/tree/main/2022-09-30_BiolCybernetics
+% https://github.com/SpikeAI/pyTERtorch/
 % https://www.zotero.org/groups/4776796/fastmotiondetection
@@
 % https://www.overleaf.com/project/630f9044c38e7a3cea81a7b2
+
+% 2023-01-13 decision: Revise
+% 
+% REVIEWER REPORTS
+
+% Reviewer Comments:
+
+% Reviewer 1
+% 1.There is a lack of an intuitive figure to show the HD-SNN model used in this paper.
+
+% 2.The MLR model used in the experimental part was presented in authors’ another work. Therefore, the changes on the MLR model are suggested to be highlighted and contrasted graphically.
+
+% 3.The title of section 2.2.1 is ’’a generative model for rater plots’’, but actually this section introduces basic terms and symbols.
+Section 2.2.2 is named ‘’detecting spiking motifs’’, but most of the content seems to introduce the generation process of raster plots, while only the last paragraph introduces the detection of motif.
+It is recommended that the content of these two sections be reorganized and that references to Figure 1(b)(c)(d) be added where appropriate during the description.
+
+% 4.The adjustment of weights and delays needs more detailed elaboration, preferably given in mathematical formulas.
+
+% 5.In Figure 4, it is claimed that the kernels are very similar for the ON polarities, and different kernels are selective to the different motion directions.
+However, in terms of these two observations, I think they are not very significant, so I suggest a more intuitive way to reflect them, such as quantifying the similarity between the ON kernels, and the angle between the kernels and the vector of the standard motion directions.
+
+% 6.There are few experiments and lack of comparative experiments.
+Since the model has an improvement based on MLR model, it is suggested to increase the comparison with MLR model and HOTS model on other datasets, such as N-MNIST and other neuromorphic datasets, so as to show the advantages of the proposed model in computational consumption.
+
+% 7.Some sections are verbose, such as the discussion.
+
+% 8.There are some minor errors in the text, and it is recommended to check the full text: 
+% 1)In the first paragraph on page 8, "a presynaptic address b_s" should be "a postsynaptic address b_s"; 
+% 2)Whether W_b in the first formula should be K_b? And d in it lacks description. 
+% 3)"PGs" should be given in its full name when it first appears. 
+% 4)There is no subgraph label of (a), (b) and (c) in Fig. 3, but the expression of Fig. 3-(a)/(b)/(c) appears many times. 
+% 5)The content of paragraph 1 on page 13 is repeated.
+
+% Reviewer 2
+% Key results: Please summarize what you consider to be the outstanding features of the work.
+% Using a single layer of spiking neurons with heterogeneous delays to learn spatio-temporal  spiking motifs
+
+% Validity: Does the manuscript have flaws which should prohibit its publication? If so, please provide details.
+
+% The work presented here is very similar to the author's other work,  "Learning hetero-synaptic delays for motion detection in a single layer of spiking neurons" 10.1109/ICIP46576.2022.9897394.
+This must be addressed and cited with an explanation of how this work presented is an extension of this previous publication.
+Abstract, Fig 6 and the main result is presented in the previous work.
+Arguably some of the figures from the previous publication would be nice to see in this publication as it helps with the visualisation of the problem.
+One suggestion for the extension to make the work unique would be to extend with some event-camera data, as it is mentioned multiple times within the publication.
+% There are also a number of citations that could have been added that did appear in the other publication, while also including other motion detection/ segmentation or previous work on event-based optical flow, which is a very similar problem.
+
+% Originality and significance: If the conclusions are not original, please provide relevant references.
+
+% See above...
 
@@
 Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
-as  $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a presynaptic address $\postsynaddr_\ranksyn$,  a weight $\synapticweight^\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
+as  $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$,  a weight $\synapticweight^\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
 This defines the full connectivity of the HD-SNN model.

### 2023-01-19 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 However, most artificial neuronal models do not take advantage of this minute temporal dimension.
-Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous delays which we apply to the computer vision task of motion detection.
+Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays which we apply to the computer vision task of motion detection.
 Indeed, the variety of synaptic delays on the dendritic tree allows to synchronize synaptic inputs as they reach the basal dendritic tree.
@@
 %%%-----------------------------------------------------------------
-Let us now formally define the Heterogeneous Delays Spiking Neural Network (HD-SNN) model, as well as the task and how we will learn to solve it.
+Let us now formally define the Heterogeneous Delays Spiking Neural Network (HD-SNN) model, as well as the task to solve and how we will learn to solve it.
 First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays.
@@
 %
-In neurobiological recordings or in the sensory signal obtained from an event-based camera, any generic raster plot consists of a stream of \emph{spikes} (see Figure~\ref{fig:model}-(a)).
+In neurobiological recordings or in the sensory signal obtained from an event-based camera, any generic raster plot consists of a stream of \emph{spikes} or \emph{events} (see Figure~\ref{fig:model}-(a)).
 This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events.

### 2023-01-19 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 \subsection{Recent challenges in machine learning and computer vision}
-%%%-----------------------------------------------------------------
-Machine learning, and computer vision in particular, have undergone a major transformation in modern times, transitioning from an eccentric character in science fiction movies to become mainstream tools found in the smartphones in everyone's pocket.
+
+The human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, about tens of watts.
+This system is the result of millions of years of natural selection, and a striking difference between biological neural networks and artificial ones is the representation they use.
+Convolutional Neural Networks (CNNs), for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels.
+These networks have been known to mimic many properties of biological systems~\citep{schrimpf_brain-score_2020}, yet, this score does not account for rapidity or energy consumption.
+On the other hand, information in a large majority of biological neural networks is represented as \emph{spikes}, i.e.
+prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted this spike~\citep{paugam-moisy_computing_2012}.
+While CNNs obtain state-of-the-art performances for some computer vision tasks like image recognition, they do not make use of the dynamics inherent to the way we sense our natural environment.
+The third generation of artificial neural networks, i.e.
+Spiking Neural Networks (SNNs), include this temporal dimension in their computations.
+This new kind of algorithms are interesting for computational neuroscience as they are a better model of biological neurons.
+Another class of approaches have designed normative models of SNNs that aim to have applications in machine learning.
+One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
+Another type of SNN using precise spike timing attempts to determine the structure of the network in order to minimize a cost function.
+This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extended in novel gradient-based methods.
+The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}.
+However, the performance of SNNs is still lagging that of firing rate-based networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
+
+%%%-----------------------------------------------------------------
+%Machine learning, and computer vision in particular, have undergone a major transformation in modern times, transitioning from an eccentric character in science fiction movies to become mainstream tools found in the smartphones in everyone's pocket.
 Less visible, yet crucial to our economy and security, these automated tools now dominate the range of services that equip virtually most sectors of our daily lives.
@@
 
-The challenges are numerous in such situations, and two major obstacles must be highlighted: the computational processing time and the energy budget.
+%The challenges are numerous in such situations, and two major obstacles must be highlighted: the computational processing time and the energy budget.
 First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun_gradient-based_1998}, then high definition images, such as VGG on ImageNet~\citep{simonyan_very_2015}.
@@
 %
-Second, a major constraint for future systems is energy cost.
+%Second, a major constraint for future systems is energy cost.
 As we have seen, fast computation time has been maintained for high-dimensional images, but at the cost of higher energy consumption.
@@
 % The brain achieves all these on a low computational cost and power budget (i.e., tens of watts), whereas the state-of-the-art computer vision models require multiple graphics processors consuming kilowatts of power
-Neuroscience could provide revolutionary answers to these two closely related problems.
+%Neuroscience could provide revolutionary answers to these two closely related problems.
 Indeed, the human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, about tens of watts.

### 2023-01-19 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 % !TEX encoding = UTF-8 Unicode
-%!TeX TS-program = pdfLaTeX
-%!TeX spellcheck = en-US
-%!BIB TS-program = bibtex
+% !TeX TS-program = pdfLaTeX
+% !TeX spellcheck = en-US
+% !BIB TS-program = bibtex
 % -*- coding: UTF-8; -*-
@@
 
+In a review paper, we report experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues~\cite{grimaldi_precise_2022}.
+These spatiotemporal patterns of neural activity can be an interesting representation used by neurons to perform cognitive tasks.
+
 %%%-----------------------------------------------------------------

### 2023-01-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 \label{sec:intro}
-% >>>>>>>>>>> LuP is here <<<<<<<<<<<
-%
-\subsection{Recent challenges in machine learning and computer vision}
 
@@
 This system is the result of millions of years of natural selection, and a striking difference between biological neural networks and artificial ones is the representation they use.
-Convolutional Neural Networks (CNNs), for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels.
+Convolutional Neural Networks (CNNs), for example, represent information passing from one layer to another as tensors storing visual information densely over the visual topography, while representing different properties in different channels.
 These networks have been known to mimic many properties of biological systems~\citep{schrimpf_brain-score_2020}, yet, this score does not account for rapidity or energy consumption.
@@
 Spiking Neural Networks (SNNs), include this temporal dimension in their computations.
-This new kind of algorithms are interesting for computational neuroscience as they are a better model of biological neurons.
+This new kind of algorithms are interesting for computational neuroscience as they represent a better model of biological neurons.
 Another class of approaches have designed normative models of SNNs that aim to have applications in machine learning.
@@
 In a review paper, we report experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues~\cite{grimaldi_precise_2022}.
-These spatiotemporal patterns of neural activity can be an interesting representation used by neurons to perform cognitive tasks.
-
-%%%-----------------------------------------------------------------
-%Machine learning, and computer vision in particular, have undergone a major transformation in modern times, transitioning from an eccentric character in science fiction movies to become mainstream tools found in the smartphones in everyone's pocket.
-Less visible, yet crucial to our economy and security, these automated tools now dominate the range of services that equip virtually most sectors of our daily lives.
-This trend is profoundly pronounced in the field of medicine, as it provides practitioners with new diagnostic tools to treat people faster and more effectively.
-This particular example also illustrates some important challenges that these tools are increasingly facing.
-One of these is the sensitivity of deep-learning based algorithms to adversarial attacks, that is, the sensitivity of the models' output to minute changes in the input.
-This could raise important questions about the reliability of the results of any automated diagnostic model.
-More generally, these models do give an output for any possible input, but they lack an explanation as to why they produced it.
-This lack of explainability is a hindrance to the advancement of these methodologies, as well as to their acceptance for most safety-relevant tasks.
-Having said that, we may remember that these tools are inherited from theoretical models originally designed to understand the brain.
-While the first artificial neural networks were mainly applied to toy-like problems, they also offered a growing number of applications in machine learning and computer vision, but also more generally for task automation.
-In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates, and today they reach higher performances than humans for key image categorization challenges.
-One of the challenges of these models is to be able to scale them up for future applications, for example in autonomous driving.
-Indeed, in such applications, the input does not consist of a single image (say a few million pixels) but of a continuous stream of images from several cameras.
-
-%The challenges are numerous in such situations, and two major obstacles must be highlighted: the computational processing time and the energy budget.
-First, most existing solutions are built incrementally from models built to process low definition images, such as Le Net for MNIST~\citep{lecun_gradient-based_1998}, then high definition images, such as VGG on ImageNet~\citep{simonyan_very_2015}.
-This has a cost and notably the fact that the computation time increases with the size (in number of pixels) of the image.
-This phenomenon has been mitigated by the increase in available computing power, notably thanks to the parallel processing implemented in GPU cards.
-However, the transition from images to videos, then to multiple cameras implies a further level in the change of scale in the amount of data processed.
-It also raises an issue that is often not addressed when it comes to image categorization, that is, of the right timing to take a decision.
-Indeed, in autonomous driving, an important task is to react as quickly as possible to a possible collision, for example when a pedestrian unexpectedly crosses the street.
-As such, future systems should always be able to give a response at any time, instead of having to wait for the pipeline to provide the correct answer.
-This online processing is also a feature of processing in biological systems and could be an important first contribution if we could translate these principles into future automated processes.
-%
-%Second, a major constraint for future systems is energy cost.
-As we have seen, fast computation time has been maintained for high-dimensional images, but at the cost of higher energy consumption.
-In addition, the need to solve increasingly complex tasks has required the construction of increasingly complex and deep architectures.
-As an example, the best performing model in the ImageNet challenge uses 2100 million parameters.
-Training the model typically consumes many kilowatts hours of energy and about 1 joule to infer a decision per single image.
-Less power-hungry solutions have emerged, such as Mobile Net~\citep{howard_searching_2019}, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be a significant percentage of the energy consumed by an electric car under standard conditions.
-
-% The brain achieves all these on a low computational cost and power budget (i.e., tens of watts), whereas the state-of-the-art computer vision models require multiple graphics processors consuming kilowatts of power
-%Neuroscience could provide revolutionary answers to these two closely related problems.
-Indeed, the human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, about tens of watts.
-This system is the result of millions of years of natural selection, and a striking difference between biological neural networks and artificial ones is the representation they use.
-CNNs, for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels.
-These networks have been known to mimic many properties of biological systems~\citep{schrimpf_brain-score_2020}, yet, this score does not count rapidity or energy.
-On the other hand, information in a large majority of biological neural networks is represented as spikes, i.e.
-prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted this spike~\citep{paugam-moisy_computing_2012}.
-Why spikes?
-It is not yet known why most biological systems have evolved to use this representation, but one major advantage seems to be some form of efficiency.
-In the human brain, each of the approximately 90 trillion nerve cells emits on average one spike per second which fan out to an average of 10000 synapses.
-Importantly, each neuron fires with great variability, and up to 300 spikes per second.
-Many attempts have been made to build artificial spiking neural networks (SNNs), and in particular to achieve the same performance as observed in classical CNNs with a lower energy budget.
-However, none of these networks has yet surpassed the state-of-the-art performances.
-Here, we will try to focus on one core property of spiking neurons, inspired by neurobiology to progress in this direction.
-%
-\subsection{What can neuroscience bring to computer vision?}% and the heterogeneous delays in spiking neurons} ~\citep{}
-%%%-----------------------------------------------------------------
-Let us first review some existing SNN models.
-A first motivation for these models was to understand biological observations at the different scales of the central nervous system.
-For example, the Hodgkin-Huxley model describes the dynamics of the membrane potential of a single nerve cell and specifically explains the role of the different voltage-dependent ion channels.
-It also explains how the model can produce a spike.
-Larger scale models introduce an simpler dynamics and an explicit spiking mechanism, for example in the Integrate-and-Fire model.
-It predicts the emergent states of a recurrently connected neural network~\citep{brunel_phase_2000}.
-Several models have increased the scale of the system to include more neurons, such as The Virtual Brain (TVB)~\citep{sanz_leon_virtual_2013}, which mixes different scales and includes SNNs with (analog) neural mass models.
-However, these methods are mainly descriptive with applications in neuroscience but few spiking neural models have had an application in computer vision.
-Another class of approaches have designed normative models of SNNs that aim to have applications in machine learning.
-One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
-Another type of SNN using precise spike timing attempts to determine the structure of the network in order to minimize a cost function.
-This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extanded in novel gradient-based methods.
-The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}.
-However, the performance of SNNs is still lagging that of firing rate-based networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
+These spatiotemporal patterns of neural activity can be a useful representation used by neurons to perform the computations needed to solve different cognitive tasks.
+In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking motifs~\citep{abeles_role_1982}.
+While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code.
+Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
+All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
+In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.
+Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays.
+This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
 
@@
 
-These computational results suggest that the temporal representation underlying processing in spiking neurons may be a key ingredient in the efficiency of the algorithms, but one may wonder how this can be supported by neurobiological data.
-In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking motifs~\citep{abeles_role_1982}.
-While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code.
-Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
-All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
-In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.
-Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays.
-This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
-
 This paper explores the possibility that prototypical spiking motifs may be a useful representation used in neurobiology and that we may transfer this hypothesis to neuromorphic algorithms, notably by using Heterogeneous Delays in Spiking Neural Networks (HD-SNN).
@@
 %
-%
-\subsection{Outline}
 %%%-----------------------------------------------------------------

### 2023-01-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
-In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}.
+In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian}.
 Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays.
 This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
+
+This paper proposes additional experiments on a recent model of spiking neuron with heterogeneous synaptic delays.
 

### 2023-01-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 
-This paper proposes additional experiments on a recent model of spiking neuron with heterogeneous synaptic delays.
+The present paper proposes additional experiments on a recent model of spiking neurons with heterogeneous synaptic delays (i.e.
+time between the arrival of an afferent spike at a specific synapse and the contribution of the associated postsynaptic potential to the neuron's soma).
+This model was trained to solve a motion detection task on a synthetic event-based dataset.
+Once trained, its event-driven computations could be reduced drastically through the pruning of synapses, while maintaining top performance for classification.
+This showed that we may use the precise timing of spikes to enhance neural computations.
+In the different kernels obtained after training, one could observe spiking motifs linked to the different motion directions
 

### 2023-01-31 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %
+% Guest Editors:
+% 
+% Kexin Chen - Department of Cognitive Sciences, University of California, Irvine
+% Hirak J. Kashyap - Department of Computer Science, University of California, Irvine
+% Jeffrey L. Krichmar - Department of Cognitive Sciences, Department of Computer Science, University of California, Irvine
+% Xiumin Li - College of Automation, Chongqing University, Chongqing, China
+% 
+% Kexin Chen, Hirak Kashyap, Jeffrey Krichmar, Xiumin Li
+% NeuroVision Organizing Team
+% 
+% Kexin Chen <kexinc3@uci.edu>
+%
+% 
 % public
@@
 
+% preprint server =  https://www.researchsquare.com/login?redirect=/article/rs-2120999/private/timeline
+% 
 % 2023-01-13 decision: Revise
@@
 This new kind of algorithms are interesting for computational neuroscience as they represent a better model of biological neurons.
-Another class of approaches have designed normative models of SNNs that aim to have applications in machine learning.
+Some approaches have designed normative models of SNNs that aim to have applications in machine learning.
 One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
@@
 The present paper proposes additional experiments on a recent model of spiking neurons with heterogeneous synaptic delays (i.e.
-time between the arrival of an afferent spike at a specific synapse and the contribution of the associated postsynaptic potential to the neuron's soma).
+time between the arrival of an afferent spike at a specific synapse and the contribution of the associated postsynaptic potential to the neuron's soma)~\cite{}.
 This model was trained to solve a motion detection task on a synthetic event-based dataset.

### 2023-01-31 Grimaldi-etal-BiolCybernetics.tex
intro

@@
 \usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.
-\AtBeginDocument{\RenewCommandCopy\qty\SI}
+%\AtBeginDocument{\RenewCommandCopy\qty\SI}
 \DeclareMathOperator*{\argmax}{arg\,max}
@@
 All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
-In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian}.
+In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020}.
 Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays.
@@
 The present paper proposes additional experiments on a recent model of spiking neurons with heterogeneous synaptic delays (i.e.
-time between the arrival of an afferent spike at a specific synapse and the contribution of the associated postsynaptic potential to the neuron's soma)~\cite{}.
+time between the arrival of an afferent spike at a specific synapse and the contribution of the associated postsynaptic potential to the neuron's soma)~\cite{grimaldi_learning_2022}.
 This model was trained to solve a motion detection task on a synthetic event-based dataset.
 Once trained, its event-driven computations could be reduced drastically through the pruning of synapses, while maintaining top performance for classification.
-This showed that we may use the precise timing of spikes to enhance neural computations.
-In the different kernels obtained after training, one could observe spiking motifs linked to the different motion directions
-
-A new generation of algorithms has emerged with the advent of so-called silicon retinas, which mimics the output of the retina by representing visual information using spikes.
-%A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance.
-The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations.
-It also allows using the precise timing of events.
-In the HOTS model for instance, the event stream induces spatio-temporal relationships between events which are represented by building ``time surfaces'', 2D images computed using the time difference to the last recorded events~\citep{lagorce_hots_2017}.
-In a recent study, we have introduced a SNN as a classification model applied to event-based streams using Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model.
-By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is, in an online fashion.
-We have demonstrated on several benchmark datasets that it provides with efficient computations, resulting in ultra-fast categorization.
-Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.
-A serie of similar algorithms are applied to real-life applications such as optical flow~\citep{benosman_asynchronous_2012}, depth estimation~\citep{dardelet_event-by-event_2021}, or gesture recognition~\citep{maro_event-based_2020}.
-Recently, such an algorithm was used in the CASPR mission on the International Space Station, illustrating, given the stringent selection processes used by the space sector, that such algorithms have reached maturity for industry.
-
-This paper explores the possibility that prototypical spiking motifs may be a useful representation used in neurobiology and that we may transfer this hypothesis to neuromorphic algorithms, notably by using Heterogeneous Delays in Spiking Neural Networks (HD-SNN).
-Our HD-SNN model starts as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights.
-Here, we propose to extend our model to a layer of spiking neurons which include heterogeneous delays in addition to weights.
-In particular, one afferent may be connected with multiple delays.
-Crucially, we will explicitly use the delay in the computational process.
-As a consequence, this extends our previous MLR model~\citep{grimaldi_robust_2022} by explictly encoding spiking motifs as spatio-temporal patterns.
-The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier.
-In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these heterogeneous delays.
-%
-%%%-----------------------------------------------------------------
-In this work, we study the emergence of spatio-temporal spiking motifs when training a single layer of spiking neurons on a semi-supervised classification task (\seeFig{motion_task}).
-The paper is organized as follows.
-We develop a theoretically defined HD-SNN capable of learning heterogeneous delays that we first evaluate on a synthetic event-based dataset.
-We first detail the methodology by defining the basic mechanism of spiking neurons that utilize heterogeneous delays.
-This will allow us to formalize the spiking neuron used to learn the model's parameters and test its effectiveness.
-We aim to apply this model to a realistic motion detection task which we will first define.
-Then, we will show how the model can be adapted to solve the motion detection task.
-In the results section, we will detail the emergent structure of the model solving the task and show in particular the similarities with neurobiological observations.
-We will also test the network with classical stimuli used in neuroscience, and in particular with frequency or contrast responses.
-We will also study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.
+This showed that formal neurons can make use the precise timing of spikes to enhance neural computations.
+In the present paper, we extend the motion detection to a more ecological task.
+Instead of synthetic textures that were used to generate stimuli for which motion is controlled, we use natural images and random movements that mimic biological saccades.
+%%%-----------------------------------------------------------------
+Through an ecological motion detection task, we study the emergence of spatio-temporal spiking motifs when training a single layer of spiking neurons on a semi-supervised classification task (\seeFig{motion_task}).
+First, we develop a theoretically defined Heterogeneous Delays in Spiking Neural Network (HD-SNN) model capable of learning heterogenous synpatic delays.
+Then, we define the ecological cognitive task the model has to solve with the different datasets it will be tested on.
+We will study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.
 This will allow us to show how such a model can provide an efficient solution to the energy/accuracy trade-off.
-Finally, we will conclude by highlighting the main contributions of this paper, while defining some limitations.
-This will open perspectives for future SNNs.
-In particular, as neuromorphic devices are by design good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is perfectly fit to be transfered to this type of hardware and to obtain significant gains in the energy which is used.
 %

### 2023-02-03 Grimaldi-etal-BiolCybernetics.tex
figure comments

@@
 %
-\subsection{Detecting event-based motifs using spiking neurons with heterogeneous delays}% model as a Logistic Regression with a temporal convolution}
-%%%-----------------------------------------------------------------
-%
-\begin{figure}%[t!]
-    \centering
-    \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
-    \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
-    \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
-    \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
-	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
-	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific neural addresses and at specific times, one may consider that this information consists of the repeated occurrence of precise spiking motifs.
-	    {\bf (b)}~We show the motifs used in this example, each identified at the top by a different color.
-To each of the $10$ neural adresses and $71$ different possible delays is assigned an evidence of activation (red) or deactivation (blue).
-Note that each afferent may be connected with multiple weights at different delays.
-	    {\bf (c)}~The activation in time of the different motifs is then used to define a generative model for drawing a raster plot on the multi-unit address space: The propagation of the afferent information through these delays generates the raster plot seen in (a).
-	    {\bf (d)}~The inversion of the generative model provides with a model which gives the predicted probability of occurrence of each motif at any time and which may be used to generate a spike as a Bernoulli trial.
-The detection model yields in this particular case with an exact identification of the occurrence of motifs.
-Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes can be annotated with each identified spiking motif.
-	    }
-    \label{fig:model}
-\end{figure}
-% 
-\subsubsection{A generative model for raster plots}
-%
-In neurobiological recordings or in the sensory signal obtained from an event-based camera, any generic raster plot consists of a stream of \emph{spikes} or \emph{events} (see Figure~\ref{fig:model}-(a)).
-This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events.
-Events are typically ordered by their time of occurrence.
-Each event has a time of occurrence $\timev_\arank$  and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
-This defines an address space $\presynaddrspace$ which consists of the set of possible addresses.
-In a neurobiological recording, this can be the identified set of neurons.
-For event-based cameras, it is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras).
-
-% This raster plot is generated by spiking neurons which are defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree, or on the sensory processes (such as the pixel on the sensor of a camera, or the photoreceptors in the retina).
-A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree, the so-called weight matrix.
-A more detailed description of these synaptic contacts reveals the added importance of heterogeneous delays, i.e., the precise timing from one afferent neuron's firing to its arrival in the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}.
-In that description, we can parameterize each neuron by the set of tuples defining both the weight and the delay of each synaptic contact.
-As a consequence, a set of input presynaptic spikes $\event$ will be processed by the dendrites defined by this synaptic set and notably by the respective delays, which will multiplex in time all events.
-
-Let's formalize a layer of spiking neurons with heterogeneous delays (HD-SNN).
-Each neuron $\postsynaddr \in \postsynaddrspace$  connects to presynaptic afferent from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses.
-Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-Note that a neuron may contact an afferent neuron with multiple different delays.
-Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
-as  $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$,  a weight $\synapticweight^\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
-This defines the full connectivity of the HD-SNN model.
-Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]} $, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]}$.
-As a consequence, an event stream which evokes neurons in the presynaptic address space is multiplexed by the synapses into a new event stream which is defined by the union of the sets generated by each emitting field from the presynaptic space:
-$ \cup_{\arank \in [1,\numevent]} \{ \{(\postsynaddr_\ranksyn, \synapticweight_\ranksyn, \timev_\arank + \synapticdelay_\ranksyn) \}_{ \ranksyn \in \synapse_{\presynaddr_\arank}} \}$.
-This new stream of events is by nature ordered in time as events reach the soma of post-synaptic neurons.
-In particular, when post-synaptic neurons are activated on their soma by this spatio-temporal motif, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.
-% \note{ TODO: show why this formula implements a polychrony detector }
-
-% \note{ éclaircir }
-
-Taking the argument the other way around, one may from a generative model for generic raster plots.
-Indeed, any spike in the presynaptic address space is generated by sensory neurons (for instance photoreceptors in the retina, sensors in a CMOS chip) or by afferent spiking neurons.
-In the latter case, these are connected to the spiking cell by a set of weights and delays, whose structure is stable relatively to the coding timescale.
-When these connections are high and sparsely distributed, this firing will cause a specific temporal motif.
-Another example is given for the barn owl auditory system: As it hears the sound of a mouse, this sound will generate a specific spiking response in both ears, and specifically, the precise timing between the signal generated by the left relative to the right ear can be for instance used to determine the position of the prey~\citep{goodman_spike-timing-based_2010}.
-Overall, these examples show that raster plots may be considered as a mixture of the effects of different elementary causes, and that each event triggers a specific spatio-temporal spiking motif.
-%
-\subsubsection{Detecting spiking motifs}
-%: Detection model
-
-From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers, it is useful to transform this event-based representation into a dense representation.
-Indeed, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{N\times T}$, where $N$ is the number of neurons in $\presynaddrspace$ and $T$ is the number of time bins.
-In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
-The probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
-Assuming that the presence of spiking motifs conditions the probability on all efferents, the logit (inverse of the sigmoid) of this probability bias can be written as the sum of the logit of each of these factors, whose values are defined by the corresponding weights.
-Spiking motifs may be activated independently and at random times, such that  we write this activity as $B(b, t)=1$ if $b$ is activated at $t$ and else $B(b, t)=0$.
-We can thus write the probability bias as the accumulated evidence given these factors as
-\begin{equation*}
-p(t, a) = \sigma\big(W_0 + \sum_{b, t} B(b, t) \cdot W_b(a, t-d) \big)  
-\end{equation*}
-where $\sigma$ is the sigmoid function.
-We will further assume that the weights are balanced (their mean is zero) and that $W_0$ is a bias such that $p_0=\sigma(W_0)$ is the average background firing rate.
-Conveniently, one can write this summation as a one-dimensional temporal convolution operator such that we may simply write
-\begin{equation*}
-p = \sigma(W_0 + B \ast W )
-\end{equation*}
-where  $p\in [ 0, 1]^{N\times T}$ and $B\in \{0, 1\}^{M\times T}$ is the raster plot corresponding to the temporal activation of the PGs.
-Finally, we obtain the raster plot $A\in \{0, 1\}^{N\times T}$ by drawing spikes using independent Bernoulli trials $A \sim \mathcal{B}(p)$.
-Note that, depending on the shape of the kernels, the generative model can model a discretized Poisson process, generate rhythmic activity or more generally propagating waves.
-This formulation thus defines a simple generative model for raster plots as a combination of independent PGs.
-%
-%Using this dense representation, the counting defined above becomes:
-%\begin{equation*}
-%\mathcal{C}^\postsynaddr(a,t)
-%= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
-%\end{equation*}
-%%
-%This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
-%This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
-%%
-%
-%\note{formalize the following:}
-%More generally, every cause may be considered as occurring independently and one might write the generative model for the generation of the presynaptic events as:
-%
-%
-%Using this formalization, one might now deduce an optimal algorithm for the detection of such temporal motifs.
-%
-%
-%By discretization of time (with here an arbitrary unitary time unit), we can also define the dendrite as a matrix giving the weight corresponding to the different delays $d \in [0, D]$ (where $D$ is the maximum delay) on different pre-synaptic addresses $a \in [1, N]$ defining the list of the $N$ dendrites.
-We will denote as $W(a, d)$ these weights.
-%
-%
-%Following the observations of~\citet{izhikevich_polychronization_2006}, let us assume that such precise discharge motif defines a polychronous group (PG).
-Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
-This allows then to derive a generative model for raster plots (see Figure~\ref{fig:model}-(b)).
-
-% 
-This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$.
-This assumes that we know the spiking motifs as defined by the $W_b$ matrices.
-The underlying metric is the binary cross-entropy, as used in the logistic regression model.
-In particular, if we consider kernels with similar decreasing exponential time profile, we can prove that this is similar to the method of~\citet{berens_fast_2012}.
-In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator.
-Using this forward model, it is possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator.
-Equivalently, this consists in using the emitting field $\synapse_\presynaddr$ of presynaptic neurons in place of the receptive field $\synapse^\postsynaddr$ of postsynaptic neurons.
-It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.
-This assumption hold as long as the kernels are uncorrelated, a condition which is met here numerically by chosing a relatively sparse set of synapses (approximately $1\%$ of active synapses).
-
-%\subsubsection{Implementation using spiking neurons with heterogeneous delays}
-%
-% The activation function of our spiking neural is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
-%It transforms this list of weights into a probability with the following formula:
-%$
-%Pr(k=\postsynaddr \; \vert \; \timev) =
-%\frac 1 Z
-%{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
-%$ 
-%where $\mathcal{C}^\postsynaddr(\timev) = \sum
-%\activeweights^\postsynaddr(t)
-%$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
-%In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
-%%
-%
-% \note{say that it is effortless in biology but difficult on conventional computers}
-% In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
-The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the heterogeneous delays model.
-%
-\subsubsection{Performance on synthetic data}
-%
-\begin{figure}%[t!]
-    \centering
-    \includegraphics[width=0.490\linewidth]{figures/THC_N_PGs.pdf}
-    \includegraphics[width=0.490\linewidth]{figures/THC_N_PG_time.pdf}
-	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
-	    {\bf (a)}~Accuracy of detection as a function of the number of kernels.
-	    {\bf (b)}~Accuracy of detection as a function of the temporal depth $D$ of kernels among $M=144$ kernels.
-	    }
-    \label{fig:model_results}
-\end{figure}
-
-To quantify the efficiency of this operation, we generated $M=144$ synthetic spiking motifs as random independent kernels over $128$ presynaptic inputs and $D=71$ possible delays.
-We drew random independent instances of $B$ with a length of $T=1000$ time steps and with on average $1.0$ occurrences in each draw.
-This allowed us to generate raster plots which we use to infer $\hat{B}$.
-We compute the accuracy as the rate of true positive detections (both for inferring the address and its exact timing) and observe on average $\approx 98\%$ correct detections.
-We further extended this result by showing how the accuracy would evolve as a function of the number of simultaneous PGs, while keeping the same frequency of occurrence.
-We show in Figure~\ref{fig:model_results}~(a) that the accuracy of finding the right PG is still above $80\%$ accuracy with more than $1364$ overlapping PGs.
-Moreover, we show in Figure~\ref{fig:model_results}~(b) that (with $M=144$ PGs fixed) the accuracy increases notably as the temporal depth $D$ of the PG kernel increased, demonstrating quantitatively the computational advantage of using heterogeneous delays.
-These results were obtained while assuming that we know $W$.
-However, this is in general not the case, for instance when observing the raster plot of a population of neurons.
-In the following, we will define a generic visual task and determine a learning algorithm to solve that challenging task.
-%Inspired by the k-means algorithm, it is possible to devise a self-supervised learning algorithm.
-Our preliminary results show that it is possible to retrieve PGs embedded in the data, yet that further analysis is necessary to improve the convergence of the algorithm.
-In particular, it seems promising to use a sparseness constraint in the inference mechanism such as to remove spurious correlations in the inference.
-
 \subsection{Task definition: fast motion detection}
@@
 The resulting instances yield trajectories qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(a)).
-
 %%%-----------------------------------------------------------------
@@
     \caption{
-    {\bf Motion detection task.} {\bf (Left)} We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view  such as to reproduce the effect of displacing the eye.
-To mimic the effect of a saccadic-like eye movement, the axis of view moves following a step-like random walk, and we show here on example path of the trajectory as a yellow line ($200$ time steps).
-{\bf (Right)}~Recording the dynamics of the sub-image as a function of time, it generates a naturalistic movie which may be transformed to an event-based representation.
-Mimicking the retina, this representation codes for proportional increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events.
-This will constitute the input to the SNN.
+    {\bf - Motion detection task.} {\bf (Left)}~We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view such as to reproduce the effect of displacing the eye.
+To mimic the effect of a saccadic-like eye movement, the axis of view moves following a step-like random walk, and we show here an example path of the trajectory as a green line ($200$ time steps).
+{\bf (Right)}~Snapshot of the synthesized event stream at different time steps.
+The dynamics of the sub-image as a function of time generates a naturalistic movie which may be transformed into an event-based representation.
+Mimicking the retina, this representation codes for proportional and pixel-wise increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events.
+On the bottom left corner of the images, the translation vector are plotted in red among the different possible classes of movements.
 Note the change of direction between the second and third frames.}
@@
 %
+%\subsection{Detecting event-based motifs using spiking neurons with heterogeneous delays}% model as a Logistic Regression with a temporal convolution}
+%%%%-----------------------------------------------------------------
+%%
+%\begin{figure}%[t!]
+%    \centering
+%    \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
+%    \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
+%    \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
+%    \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
+%	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
+%	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific neural addresses and at specific times, one may consider that this information consists of the repeated occurrence of precise spiking motifs.
+%	    {\bf (b)}~We show the motifs used in this example, each identified at the top by a different color.
+To each of the $10$ neural adresses and $71$ different possible delays is assigned an evidence of activation (red) or deactivation (blue).
+Note that each afferent may be connected with multiple weights at different delays.
+%	    {\bf (c)}~The activation in time of the different motifs is then used to define a generative model for drawing a raster plot on the multi-unit address space: The propagation of the afferent information through these delays generates the raster plot seen in (a).
+%	    {\bf (d)}~The inversion of the generative model provides with a model which gives the predicted probability of occurrence of each motif at any time and which may be used to generate a spike as a Bernoulli trial.
+The detection model yields in this particular case with an exact identification of the occurrence of motifs.
+Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes can be annotated with each identified spiking motif.
+%	    }
+%    \label{fig:model}
+%\end{figure}
+%% 
+%\subsubsection{A generative model for raster plots}
+%%
+%In neurobiological recordings or in the sensory signal obtained from an event-based camera, any generic raster plot consists of a stream of \emph{spikes} or \emph{events} (see Figure~\ref{fig:model}-(a)).
+This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events.
+Events are typically ordered by their time of occurrence.
+Each event has a time of occurrence $\timev_\arank$  and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
+This defines an address space $\presynaddrspace$ which consists of the set of possible addresses.
+In a neurobiological recording, this can be the identified set of neurons.
+For event-based cameras, it is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras).
+%
+%% This raster plot is generated by spiking neurons which are defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree, or on the sensory processes (such as the pixel on the sensor of a camera, or the photoreceptors in the retina).
+A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree, the so-called weight matrix.
+A more detailed description of these synaptic contacts reveals the added importance of heterogeneous delays, i.e., the precise timing from one afferent neuron's firing to its arrival in the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}.
+In that description, we can parameterize each neuron by the set of tuples defining both the weight and the delay of each synaptic contact.
+As a consequence, a set of input presynaptic spikes $\event$ will be processed by the dendrites defined by this synaptic set and notably by the respective delays, which will multiplex in time all events.
+%
+%Let's formalize a layer of spiking neurons with heterogeneous delays (HD-SNN).
+Each neuron $\postsynaddr \in \postsynaddrspace$  connects to presynaptic afferent from $\presynaddrspace$.
+In biology, a single cortical neuron has generally several thousands of synapses.
+Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+Note that a neuron may contact an afferent neuron with multiple different delays.
+Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
+%as  $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$,  a weight $\synapticweight^\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
+This defines the full connectivity of the HD-SNN model.
+Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]} $, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]}$.
+As a consequence, an event stream which evokes neurons in the presynaptic address space is multiplexed by the synapses into a new event stream which is defined by the union of the sets generated by each emitting field from the presynaptic space:
+%$ \cup_{\arank \in [1,\numevent]} \{ \{(\postsynaddr_\ranksyn, \synapticweight_\ranksyn, \timev_\arank + \synapticdelay_\ranksyn) \}_{ \ranksyn \in \synapse_{\presynaddr_\arank}} \}$.
+This new stream of events is by nature ordered in time as events reach the soma of post-synaptic neurons.
+In particular, when post-synaptic neurons are activated on their soma by this spatio-temporal motif, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.
+%% \note{ TODO: show why this formula implements a polychrony detector }
+%
+%% \note{ éclaircir }
+%
+%Taking the argument the other way around, one may from a generative model for generic raster plots.
+Indeed, any spike in the presynaptic address space is generated by sensory neurons (for instance photoreceptors in the retina, sensors in a CMOS chip) or by afferent spiking neurons.
+In the latter case, these are connected to the spiking cell by a set of weights and delays, whose structure is stable relatively to the coding timescale.
+When these connections are high and sparsely distributed, this firing will cause a specific temporal motif.
+Another example is given for the barn owl auditory system: As it hears the sound of a mouse, this sound will generate a specific spiking response in both ears, and specifically, the precise timing between the signal generated by the left relative to the right ear can be for instance used to determine the position of the prey~\citep{goodman_spike-timing-based_2010}.
+Overall, these examples show that raster plots may be considered as a mixture of the effects of different elementary causes, and that each event triggers a specific spatio-temporal spiking motif.
+%%
+%\subsubsection{Detecting spiking motifs}
+%%: Detection model
+%
+%From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers, it is useful to transform this event-based representation into a dense representation.
+Indeed, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{N\times T}$, where $N$ is the number of neurons in $\presynaddrspace$ and $T$ is the number of time bins.
+In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
+%The probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
+Assuming that the presence of spiking motifs conditions the probability on all efferents, the logit (inverse of the sigmoid) of this probability bias can be written as the sum of the logit of each of these factors, whose values are defined by the corresponding weights.
+Spiking motifs may be activated independently and at random times, such that  we write this activity as $B(b, t)=1$ if $b$ is activated at $t$ and else $B(b, t)=0$.
+We can thus write the probability bias as the accumulated evidence given these factors as
+%\begin{equation*}
+%p(t, a) = \sigma\big(W_0 + \sum_{b, t} B(b, t) \cdot W_b(a, t-d) \big)  
+%\end{equation*}
+%where $\sigma$ is the sigmoid function.
+We will further assume that the weights are balanced (their mean is zero) and that $W_0$ is a bias such that $p_0=\sigma(W_0)$ is the average background firing rate.
+Conveniently, one can write this summation as a one-dimensional temporal convolution operator such that we may simply write
+%\begin{equation*}
+%p = \sigma(W_0 + B \ast W )
+%\end{equation*}
+%where  $p\in [ 0, 1]^{N\times T}$ and $B\in \{0, 1\}^{M\times T}$ is the raster plot corresponding to the temporal activation of the PGs.
+Finally, we obtain the raster plot $A\in \{0, 1\}^{N\times T}$ by drawing spikes using independent Bernoulli trials $A \sim \mathcal{B}(p)$.
+Note that, depending on the shape of the kernels, the generative model can model a discretized Poisson process, generate rhythmic activity or more generally propagating waves.
+This formulation thus defines a simple generative model for raster plots as a combination of independent PGs.
+%
+%Using this dense representation, the counting defined above becomes:
+%\begin{equation*}
+%\mathcal{C}^\postsynaddr(a,t)
+%= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
+%\end{equation*}
+%%
+%This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
+%This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
+%%
+%
+%\note{formalize the following:}
+%More generally, every cause may be considered as occurring independently and one might write the generative model for the generation of the presynaptic events as:
+%
+%
+%Using this formalization, one might now deduce an optimal algorithm for the detection of such temporal motifs.
+%
+%
+%By discretization of time (with here an arbitrary unitary time unit), we can also define the dendrite as a matrix giving the weight corresponding to the different delays $d \in [0, D]$ (where $D$ is the maximum delay) on different pre-synaptic addresses $a \in [1, N]$ defining the list of the $N$ dendrites.
+We will denote as $W(a, d)$ these weights.
+%
+%
+%Following the observations of~\citet{izhikevich_polychronization_2006}, let us assume that such precise discharge motif defines a polychronous group (PG).
+Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
+This allows then to derive a generative model for raster plots (see Figure~\ref{fig:model}-(b)).
+
+% 
+%This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$.
+This assumes that we know the spiking motifs as defined by the $W_b$ matrices.
+The underlying metric is the binary cross-entropy, as used in the logistic regression model.
+In particular, if we consider kernels with similar decreasing exponential time profile, we can prove that this is similar to the method of~\citet{berens_fast_2012}.
+In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator.
+Using this forward model, it is possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator.
+Equivalently, this consists in using the emitting field $\synapse_\presynaddr$ of presynaptic neurons in place of the receptive field $\synapse^\postsynaddr$ of postsynaptic neurons.
+It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.
+This assumption hold as long as the kernels are uncorrelated, a condition which is met here numerically by chosing a relatively sparse set of synapses (approximately $1\%$ of active synapses).
+
+%\subsubsection{Implementation using spiking neurons with heterogeneous delays}
+%
+% The activation function of our spiking neural is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
+%It transforms this list of weights into a probability with the following formula:
+%$
+%Pr(k=\postsynaddr \; \vert \; \timev) =
+%\frac 1 Z
+%{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
+%$ 
+%where $\mathcal{C}^\postsynaddr(\timev) = \sum
+%\activeweights^\postsynaddr(t)
+%$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
+%In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
+%%
+%
+% \note{say that it is effortless in biology but difficult on conventional computers}
+% In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the heterogeneous delays model.
+%
+%\subsubsection{Performance on synthetic data}
+%%
+%\begin{figure}%[t!]
+%    \centering
+%    \includegraphics[width=0.490\linewidth]{figures/THC_N_PGs.pdf}
+%    \includegraphics[width=0.490\linewidth]{figures/THC_N_PG_time.pdf}
+%	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
+%	    {\bf (a)}~Accuracy of detection as a function of the number of kernels.
+%	    {\bf (b)}~Accuracy of detection as a function of the temporal depth $D$ of kernels among $M=144$ kernels.
+%	    }
+%    \label{fig:model_results}
+%\end{figure}
+%
+%To quantify the efficiency of this operation, we generated $M=144$ synthetic spiking motifs as random independent kernels over $128$ presynaptic inputs and $D=71$ possible delays.
+We drew random independent instances of $B$ with a length of $T=1000$ time steps and with on average $1.0$ occurrences in each draw.
+This allowed us to generate raster plots which we use to infer $\hat{B}$.
+We compute the accuracy as the rate of true positive detections (both for inferring the address and its exact timing) and observe on average $\approx 98\%$ correct detections.
+We further extended this result by showing how the accuracy would evolve as a function of the number of simultaneous PGs, while keeping the same frequency of occurrence.
+We show in Figure~\ref{fig:model_results}~(a) that the accuracy of finding the right PG is still above $80\%$ accuracy with more than $1364$ overlapping PGs.
+Moreover, we show in Figure~\ref{fig:model_results}~(b) that (with $M=144$ PGs fixed) the accuracy increases notably as the temporal depth $D$ of the PG kernel increased, demonstrating quantitatively the computational advantage of using heterogeneous delays.
+These results were obtained while assuming that we know $W$.
+However, this is in general not the case, for instance when observing the raster plot of a population of neurons.
+In the following, we will define a generic visual task and determine a learning algorithm to solve that challenging task.
+%Inspired by the k-means algorithm, it is possible to devise a self-supervised learning algorithm.
+Our preliminary results show that it is possible to retrieve PGs embedded in the data, yet that further analysis is necessary to improve the convergence of the algorithm.
+In particular, it seems promising to use a sparseness constraint in the inference mechanism such as to remove spurious correlations in the inference.
+%
 \subsection{Extension as 3D convolutions for motion detection}
 %%%-----------------------------------------------------------------
-%
 Now that our motion detection task is defined in visual space, we can extend the convolution-based detection model to a 3D convolution so that the resulting model would also benefit from spatial invariance.
@@
 %
+\begin{figure}%[h!]%[ht!]
+    \centering
+    \includegraphics[width=\linewidth]{figures/HD-SNN.png}
+    \caption{
+    \textbf{- HD-SNN illustration.
+From left to right:} A 3D representation of the event stream as output of the DVS.
+A 3D convolution is applied, on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis.
+There are as many kernels as the possible number of motion classes in the recording.
+For clarity, only $4$ kernels are illustrated and some weights are not represented in the voxel grid.
+For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model.
+The output of the MLR gives a probability for each class associated to a specific kernel to emit an output event, or spike.
+With an appropriate spiking mechanism, here the winner-takes-all mechanism, we obtained, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class.
+}
+    \label{fig:model}
+\end{figure}
+%
 The output of the MLR model corresponds to an event with the highest probability class, keeping the same timing as the input event.
-It is important to note that we include the prior knowledge that the event-based input is spatio-temporally clustered and that these clusters are sparse.
+It is important to note that we include the prior knowledge that the event-based input is spatiotemporally clustered and that these clusters are sparse.
 Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels.
@@
 We tested the effect of different parameters on a validation set to quantify the role of each parameter.
-Since the model is fully differentiable, we can now implement a semi-supervised learning rule.
-The loss function of the MLR model is the binary cross-entropy on the output of the classification layer and can be extended to space-time convolutions.
-Supervision was implemented using the input binary events as defined above and the labels as the desired output.
-The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions.
-As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner.
-It is important to note that we weighted the cost function by considering only the  most active cells, such that the error is only back propagated to the spatial locations of these most active cells.
-This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier_unsupervised_2007}.
-Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$).
-We tested the effect of different parameters on a validation set to quantify the role of each parameter.
-
 %

### 2023-02-04 Grimaldi-etal-BiolCybernetics.tex
modif legend

@@
 Through an ecological motion detection task, we study the emergence of spatio-temporal spiking motifs when training a single layer of spiking neurons on a semi-supervised classification task (\seeFig{motion_task}).
-First, we develop a theoretically defined Heterogeneous Delays in Spiking Neural Network (HD-SNN) model capable of learning heterogenous synpatic delays.
-Then, we define the ecological cognitive task the model has to solve with the different datasets it will be tested on.
+First, we define the ecological cognitive task the model has to solve with the different datasets it will be tested on.
+Then, we develop a theoretically defined Heterogeneous Delays in Spiking Neural Network (HD-SNN) model capable of learning heterogenous synaptic delays.
 We will study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.
@@
 %%%-----------------------------------------------------------------
-Let us now formally define the Heterogeneous Delays Spiking Neural Network (HD-SNN) model, as well as the task to solve and how we will learn to solve it.
-First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays.
-Our HD-SNN model will be tested on a computationally relevant task of detecting motion detection from an event-based sensory input.
-By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory.
-Solving this task is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina.
-The task is thus to guess the motion as accurately as possible (\seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays.
-To this end, we will extend the HD-SNN model to be efficient for the resolution of the task by including spatio-temporal convolutional kernels.
-We will finally describe its implementation on conventional computers.
 %
@@
     \caption{
-    {\bf - Motion detection task.} {\bf (Left)}~We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view such as to reproduce the effect of displacing the eye.
+    {\bf - Motion detection task.} {\bf (Left)}~We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view.
 To mimic the effect of a saccadic-like eye movement, the axis of view moves following a step-like random walk, and we show here an example path of the trajectory as a green line ($200$ time steps).

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
local changes laptop

@@
 %%%-----------------------------------------------------------------
-Let us now define a procedure of animating of natural scene by virtual eye-movements similar to that which was used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
+%Let us now define a procedure of animating of natural scene by virtual eye-movements similar to that which was used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
+We want to test our model on a motion detection task on a binary stream of events.
+This type of signal, typically recorded by a Dynamic Vision Sensor (DVS), is a good representation of the information sent by the retinal ganglion cells through the optic nerve.
+With binary events, the visual system is able to capture the dynamics of a natural scene and perform various cognitive tasks in a fast and efficient fashion.
+To train and test our model, we need to define a dataset for which we control for motion direction and velocity.
 First, let's define a trajectory inspired by the biological movements of the eyes.
@@
 In animals with a fovea, this is particularly useful as it allows to move the area with the highest density of photoreceptors in the environment, for example at a point of interest.
-A specific mechanism to do so are called saccades which are rapide eye movements of the eyeball that reposition the center of vision.
+A specific mechanism to do so are called saccades which are rapid eye movements of the eyeball that reposition the center of vision.
 In humans, these are very frequent (on average $2$ each second~\citep{dandekar_neural_2012}).
@@
 %%%-----------------------------------------------------------------
-Now that our motion detection task is defined in visual space, we can extend the convolution-based detection model to a 3D convolution so that the resulting model would also benefit from spatial invariance.
+\subsubsection{The general model}
+\subsubsection{Applied on this detection task}
+Now that our motion detection task is defined in visual space, we can define the model that will be trained to solve the task.
 The use of spatio-temporal filters on an event stream has improved the performance of CNNs for an action recognition task in a model by~\citet{ghosh_spatiotemporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %%%-----------------------------------------------------------------
-Through an ecological motion detection task, we study the emergence of spatio-temporal spiking motifs when training a single layer of spiking neurons on a semi-supervised classification task (\seeFig{motion_task}).
+Through an ecological motion detection task, we study the emergence of spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task.
 First, we define the ecological cognitive task the model has to solve with the different datasets it will be tested on.

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
methods modif

--- 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
@@
 \affil{\orgdiv{\Department}, \orgname{\Affiliation}\\ \orgaddress{\street{\Street}, \city{\City}, \postcode{\PostCode}, \country{\Country}}}
-
 %%==================================%%
@@
 Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of % N-V-phi
-$8$ linearly spaced motion directions and %N-Y-velocity=
-$7$ different velocities (\seeFig{motion_task}-(a)-inset).
-Note that the velocities are sampled on a geometric scale between $1/4$ and $4$.
+$8$ linearly spaced motion directions.
 Next, we define one trajectory of gaze as segments whose duration is drawn from a Poisson distribution with an average block length of $20~\ms$, similarly to a Lévy flight~\citep[p.
@@
 Finally, the trajectory is integrated by assuming first that velocities are uniformly and independently sampled from the set of different motion-sets and second that motion is uniform during a time segment.
-The resulting instances yield trajectories qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(a)).
+The resulting instances yield trajectories qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(Left)).
 %%%-----------------------------------------------------------------

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 This process is known to occur as early as the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step (\seeFig{motion_task}-(b)).
+These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step.
 We will discretize the time in $1~\ms$ bins and produce movies of duration $N_T = 250~\ms$.
@@
 The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
-Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that a velocity of $V=1$ corresponds to a motion of one pixel per frame (i.e., per time bin).
+Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that it corresponds to a motion of one pixel per frame (i.e., per time bin).
 
-% \note{ TODO: formula / say it is driven by image gradients / give scaling rules }
 To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames.
@@
 This signed threshold value is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is as a consequence similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion_task}-(c)).
+When applied to the whole movie, the event stream is as a consequence similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion_task}-(Right)).
 The goal here is to infer the correct motion solely by observing these events.

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 The goal here is to infer the correct motion solely by observing these events.
-This sensory signal representing the output of this event-based camera forms a discrete stream of events, $\event$ as defined above.   
-Each event has an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
-This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).
 % \note{ the task is full-field rigid translations, the model resolves it locally (analogy to the input from V1 to MT) }
@@
 %
-\subsection{Extension as 3D convolutions for motion detection}
+\subsection{Description of the model}
 %%%-----------------------------------------------------------------
 \subsubsection{The general model}
+
+In neurobiological recordings or in the sensory signal obtained from an event-based camera, any generic raster plot consists of a stream of \emph{spikes} or \emph{events}.
+This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events.
+Events are typically ordered by their time of occurrence.
+Each event has a time of occurrence $\timev_\arank$ and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
+This defines an address space $\presynaddrspace$ which consists of the set of possible addresses.
+In a neurobiological recording, this can be the identified set of neurons.
+For event-based cameras, it is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras).
+
+Let's formalize a layer of spiking neurons with heterogeneous delays (HD-SNN).
+Each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferent from $\presynaddrspace$.
+In biology, a single cortical neuron has generally several thousands of synapses.
+Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+Note that a neuron may contact an afferent neuron with multiple different delays.
+Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
+as  $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
+This defines the full connectivity of the HD-SNN model.
+Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]} $, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]}$.
+As a consequence, a postsynaptic neuron $\postsynaddr$ receives, on its receptive field, an event stream which is multiplexed by the synapses into a new event stream defined by the union of the sets generated by each emitting field from the presynaptic space:
+$ \cup_{\arank \in [1,\numevent]} \{ \{(\postsynaddr_\ranksyn, \synapticweight_\ranksyn, \timev_\arank + \synapticdelay_\ranksyn) \}_{ \ranksyn \in \synapse_{\presynaddr_\arank}} \}$.
+This new stream of events is by nature ordered in time as events reach the soma of post-synaptic neurons.
+In particular, when post-synaptic neurons are activated on their soma by this spatio-temporal motif, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.
+
 \subsubsection{Applied on this detection task}

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]} $, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]}$.
-As a consequence, a postsynaptic neuron $\postsynaddr$ receives, on its receptive field, an event stream which is multiplexed by the synapses into a new event stream defined by the union of the sets generated by each emitting field from the presynaptic space:
+As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field.
+into a new event stream defined by the union of the sets generated by each emitting field from the presynaptic space:
 $ \cup_{\arank \in [1,\numevent]} \{ \{(\postsynaddr_\ranksyn, \synapticweight_\ranksyn, \timev_\arank + \synapticdelay_\ranksyn) \}_{ \ranksyn \in \synapse_{\presynaddr_\arank}} \}$.

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field.
-into a new event stream defined by the union of the sets generated by each emitting field from the presynaptic space:
-$ \cup_{\arank \in [1,\numevent]} \{ \{(\postsynaddr_\ranksyn, \synapticweight_\ranksyn, \timev_\arank + \synapticdelay_\ranksyn) \}_{ \ranksyn \in \synapse_{\presynaddr_\arank}} \}$.
+It results in a weighted event stream:
+$\event_\postsynaddr = \{(\presynaddr_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}$.
+Unlike the binary stream of events as input, once integrated by the synapses, the postsynaptic neurons receives a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank$.
 This new stream of events is by nature ordered in time as events reach the soma of post-synaptic neurons.

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
abstract

--- 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
@@
 The response of a biological neuron depends on the precise timing of afferent spikes.
-This temporal aspect of the neuronal code is essential in understanding information processing in neurobiology and applies particularly well to the output of neuromorphic hardware such as event-based cameras.
-However, most artificial neuronal models do not take advantage of this minute temporal dimension.
-Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays which we apply to the computer vision task of motion detection.
-Indeed, the variety of synaptic delays on the dendritic tree allows to synchronize synaptic inputs as they reach the basal dendritic tree.
-We show this can be formalized as a time-invariant logistic regression which can be trained using labeled data.
-We apply this model to solve the specific computer vision problem of motion detection, and demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of event-based cameras.
-In particular, we quantify how the accuracy of the model can vary with the total computational load.
-This end-to-end event-driven computational brick could help improve the performance of future Spiking Neural Network (SNN) algorithms and their prospective use in neuromorphic chips.
+This temporal aspect of the neural code is essential to understand information processing in neurobiology and applies particularly well to the output of neuromorphic hardware such as event-based cameras.
+However, most artificial neural models do not take advantage of this important temporal dimension of the neural code.
+Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays, which we apply to the computer vision task of motion detection.
+Indeed, the diversity of synaptic delays on the dendritic tree allows for the synchronization of synaptic inputs as they reach the basal dendritic tree.
+We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
+We apply this model to solve the specific computer vision problem of motion detection.
+We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of event-based cameras.
+In particular, we quantify how the accuracy of the model can vary as a function of the overall computational load.
+This end-to-end, event-driven computational building block could improve the performance of future spiking neural network (SNN) algorithms and neuromorphic chips.
 }

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex

@@
 Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]} $, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]}$.
-As a consequence, a postsynaptic neuron $\postsynaddr$ receives, on its receptive field, an event stream which is multiplexed by the synapses into a new event stream defined by the union of the sets generated by each emitting field from the presynaptic space:
-$ \cup_{\arank \in [1,\numevent]} \{ \{(\postsynaddr_\ranksyn, \synapticweight_\ranksyn, \timev_\arank + \synapticdelay_\ranksyn) \}_{ \ranksyn \in \synapse_{\presynaddr_\arank}} \}$.
+As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field.
+It results in a weighted event stream:
+$\event_\postsynaddr = \{(\presynaddr_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}$.
+Unlike the binary stream of events as input, once integrated by the synapses, the postsynaptic neurons receives a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank$.
 This new stream of events is by nature ordered in time as events reach the soma of post-synaptic neurons.

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
intro : a

@@
 \noindent %The paper is available a an~\href{https://arxiv.org/abs/2205.03635}{arXiv preprint} with links to previous versions ands to the code. 
-Find also the associated~\href{https://www.zotero.org/groups/4776796/fastmotiondetection}{zotero group} used to regroup relevant literature on the subject.
+Find also the associated~\href{https://www.zotero.org/groups/4776796/fastmotiondetection}{zotero group} used to gather relevant literature on the subject.
 
@@
 
-The human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, about tens of watts.
-This system is the result of millions of years of natural selection, and a striking difference between biological neural networks and artificial ones is the representation they use.
-Convolutional Neural Networks (CNNs), for example, represent information passing from one layer to another as tensors storing visual information densely over the visual topography, while representing different properties in different channels.
-These networks have been known to mimic many properties of biological systems~\citep{schrimpf_brain-score_2020}, yet, this score does not account for rapidity or energy consumption.
-On the other hand, information in a large majority of biological neural networks is represented as \emph{spikes}, i.e.
-prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted this spike~\citep{paugam-moisy_computing_2012}.
-While CNNs obtain state-of-the-art performances for some computer vision tasks like image recognition, they do not make use of the dynamics inherent to the way we sense our natural environment.
-The third generation of artificial neural networks, i.e.
-Spiking Neural Networks (SNNs), include this temporal dimension in their computations.
-This new kind of algorithms are interesting for computational neuroscience as they represent a better model of biological neurons.
-Some approaches have designed normative models of SNNs that aim to have applications in machine learning.
+The human brain has the remarkable property of being able to react at any time while consuming a reasonable amount of energy, about tens of watts.
+This system is the result of millions of years of natural selection, and a striking difference between biological and artificial neural networks is the representation they use.
+Convolutional neural networks (CNNs), for example, represent the flow of information from one layer to another as tensors, storing visual information densely across the visual topography, with different properties represented in different channels.
+These networks are known to mimic many properties of biological systems, such that each can be attributed a ``Barin Score''~\citep{schrimpf_brain-score_2020}, yet this score takes no account of speed or energy consumption.
+CNNs have achieved state-of-the-art performance for some computer vision tasks, such as image recognition.
+However, they do not take advantage of the dynamics inherent in the way we perceive our natural environment.
+On the other hand, in the vast majority of biological neural networks, information is represented as \emph{spikes}, i.e.
+prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted the spike~\citep{paugam-moisy_computing_2012}.
+The third generation of artificial neural networks, called Spiking Neural Networks (SNNs), incorporate this temporal dimension into how they compute.
+By providing a better model of biological neurons, these new algorithms are of interest to computational neuroscience.
+Some approaches have developed normative models of SNNs that aim to have applications in machine learning.
 One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
-Another type of SNN using precise spike timing attempts to determine the structure of the network in order to minimize a cost function.
-This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extended in novel gradient-based methods.
-The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}.
-However, the performance of SNNs is still lagging that of firing rate-based networks and the question of the advantage of using spikes in machine learning and computer vision remains open.
+Another type of SNN that uses accurate timing of spikes attempts to determine the structure of the network to minimize a cost function.
+This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been recently extended in gradient-based methods.
+This surrogate gradient method is now widely used in methods that attempt to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
+However, the performance of SNNs still lags behind that of firing rate based networks.
+The question of the advantage of using spikes in machine learning and computer vision remains open.
 

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
intro : b

@@
 
-In a review paper, we report experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues~\cite{grimaldi_precise_2022}.
-These spatiotemporal patterns of neural activity can be a useful representation used by neurons to perform the computations needed to solve different cognitive tasks.
-In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking motifs~\citep{abeles_role_1982}.
-While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code.
-Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
-All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
-In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020}.
-Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays.
-This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
+In a recent review paper, we reported experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues~\cite{grimaldi_precise_2023}.
+These spatio-temporal motifs of neural activity may be useful representations used by neurons to perform computations for various cognitive tasks.
+In 1982, Abeles asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs~\citep{abeles_role_1982}.
+While the first possibility favors the theory of rate coding, the second one emphasizes the function of temporal precision in the neural code.
+Since then, numerous studies have demonstrated the emergence of synchronicity of activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, it's use for efficient coding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
+All these findings and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
+In neural models, the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows for the efficient exploitation or detection of these spatio-temporal motifs embedded in the spike train.
+In particular,~\citep{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, but precise, relative spiking delays.
+Due to the variety of configurations and the possible coexistence of multiple superimposed motifs, this representation has a much higher information capacity than a firing rate-based approach to neural coding.
 

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
intro : c

@@
 
-The present paper proposes additional experiments on a recent model of spiking neurons with heterogeneous synaptic delays (i.e.
-time between the arrival of an afferent spike at a specific synapse and the contribution of the associated postsynaptic potential to the neuron's soma)~\cite{grimaldi_learning_2022}.
-This model was trained to solve a motion detection task on a synthetic event-based dataset.
-Once trained, its event-driven computations could be reduced drastically through the pruning of synapses, while maintaining top performance for classification.
-This showed that formal neurons can make use the precise timing of spikes to enhance neural computations.
-In the present paper, we extend the motion detection to a more ecological task.
-Instead of synthetic textures that were used to generate stimuli for which motion is controlled, we use natural images and random movements that mimic biological saccades.
-%%%-----------------------------------------------------------------
-Through an ecological motion detection task, we study the emergence of spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task.
+The present paper proposes additional experiments extending a recent model of spiking neurons with heterogeneous synaptic delays (i.e., the time between the arrival of an afferent spike at a given synapse and the contribution of the associated postsynaptic potential to the neuron's soma)~\cite{grimaldi_learning_2022}.
+This particular model was trained to solve a motion detection task on a synthetic event-based dataset generated by moving parameterized textures.
+Once trained, the volume of event-driven computations could be drastically reduced by pruning the synapses, while maintaining top performance for the classification task.
+This was a demonstration that formal neurons can exploit the precise timing of spikes to improve neural computation.
+In the present work, we extend motion detection to a more ecological task.
+Instead of the synthetic textures that have been used to generate motion-driven stimuli, we use natural images and random movements that mimic biological saccades.
+Using an ecological motion detection task, we study the emergence of spatio-temporal spiking motifs when a single layer of spiking neurons is trained on a supervised classification task.
 First, we define the ecological cognitive task the model has to solve with the different datasets it will be tested on.
-Then, we develop a theoretically defined Heterogeneous Delays in Spiking Neural Network (HD-SNN) model capable of learning heterogenous synaptic delays.
-We will study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.
-This will allow us to show how such a model can provide an efficient solution to the energy/accuracy trade-off.
+Then, we develop a theoretically defined Heterogeneous Delays in Spiking Neural Network (HD-SNN) model capable of learning heterogeneous synaptic delays.
+We will investigate the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.
+In this way, we will be able to show how such a model can provide an efficient solution to the trade-off between energy and accuracy.
 %

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
intro : pass

@@
 Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays, which we apply to the computer vision task of motion detection.
-Indeed, the diversity of synaptic delays on the dendritic tree allows for the synchronization of synaptic inputs as they reach the basal dendritic tree.
+Indeed, the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangments of synaptic inputs as they reach the basal dendritic tree.
 We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
@@
 In particular, we quantify how the accuracy of the model can vary as a function of the overall computational load.
-This end-to-end, event-driven computational building block could improve the performance of future spiking neural network (SNN) algorithms and neuromorphic chips.
+This end-to-end, event-driven computational building block could improve the performance of future spiking neural network (SNN) algorithms and in particular their implementation in neuromorphic chips.
 }
@@
 Convolutional neural networks (CNNs), for example, represent the flow of information from one layer to another as tensors, storing visual information densely across the visual topography, with different properties represented in different channels.
-These networks are known to mimic many properties of biological systems, such that each can be attributed a ``Barin Score''~\citep{schrimpf_brain-score_2020}, yet this score takes no account of speed or energy consumption.
+These networks are known to mimic many properties of biological systems, such that each can be attributed a ``Brain Score''~\citep{schrimpf_brain-score_2020}, yet this score takes no account of inference speed or energy consumption.
 CNNs have achieved state-of-the-art performance for some computer vision tasks, such as image recognition.
 However, they do not take advantage of the dynamics inherent in the way we perceive our natural environment.
-On the other hand, in the vast majority of biological neural networks, information is represented as \emph{spikes}, i.e.
-prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted the spike~\citep{paugam-moisy_computing_2012}.
-The third generation of artificial neural networks, called Spiking Neural Networks (SNNs), incorporate this temporal dimension into how they compute.
-By providing a better model of biological neurons, these new algorithms are of interest to computational neuroscience.
+In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
+The third generation of artificial neural networks, known as spiking neural networks (SNNs), incorporates this temporal dimension into the way they perform their computations.
+They are of interest for computational neuroscience because they provide a better model of the biological neuron.
 Some approaches have developed normative models of SNNs that aim to have applications in machine learning.
@@
 This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been recently extended in gradient-based methods.
-This surrogate gradient method is now widely used in methods that attempt to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
+This surrogate gradient method is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
 However, the performance of SNNs still lags behind that of firing rate based networks.
@@
 In a recent review paper, we reported experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues~\cite{grimaldi_precise_2023}.
-These spatio-temporal motifs of neural activity may be useful representations used by neurons to perform computations for various cognitive tasks.
+These spatio-temporal motifs within the neural activity may be useful representations to perform computations for various cognitive tasks.
 In 1982, Abeles asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs~\citep{abeles_role_1982}.
@@
 In neural models, the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows for the efficient exploitation or detection of these spatio-temporal motifs embedded in the spike train.
-In particular,~\citep{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, but precise, relative spiking delays.
+In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, but precise, relative spiking delays, i.e., the time between the arrival of an afferent spike at a given synapse and the contribution of the associated postsynaptic potential to the neuron's soma.
 Due to the variety of configurations and the possible coexistence of multiple superimposed motifs, this representation has a much higher information capacity than a firing rate-based approach to neural coding.
 
-The present paper proposes additional experiments extending a recent model of spiking neurons with heterogeneous synaptic delays (i.e., the time between the arrival of an afferent spike at a given synapse and the contribution of the associated postsynaptic potential to the neuron's soma)~\cite{grimaldi_learning_2022}.
-This particular model was trained to solve a motion detection task on a synthetic event-based dataset generated by moving parameterized textures.
+The present paper proposes additional experiments extending a recent model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}.
+This model was trained to solve a motion detection task on a synthetic event-based dataset generated by moving parameterized textures.
 Once trained, the volume of event-driven computations could be drastically reduced by pruning the synapses, while maintaining top performance for the classification task.
-This was a demonstration that formal neurons can exploit the precise timing of spikes to improve neural computation.
+This was a demonstration that formal neurons can exploit the precise timing of spikes to improve neural computations.
 In the present work, we extend motion detection to a more ecological task.
@@
 First, we define the ecological cognitive task the model has to solve with the different datasets it will be tested on.
-Then, we develop a theoretically defined Heterogeneous Delays in Spiking Neural Network (HD-SNN) model capable of learning heterogeneous synaptic delays.
+Then, we develop a theoretically defined Heterogeneous Delays Spiking Neural Network (HD-SNN) model capable of learning these heterogeneous synaptic delays.
 We will investigate the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
discussion : intro

@@
 %%%-----------------------------------------------------------------
-In this paper, we have introduced a generic SNN using heterogeneous delays and have shown how it compares favorably for a visual motion detection task with a state-of-the-art event-based algorithm used for classification.
-The learned model bears many similarities with neurobiological anatomical observations but also with the results from behavioral results.
-% \note{to change if we compare to ghosh instead of HOTs if we follow that route}
-The event-driven computations of our method can be reduced drastically through the pruning of synapses, while maintaining top performance for classification.
-This shows that we may use the precise timing of spikes to enhance neural computations.
+In this paper, we have introduced a generic SNN with heterogeneous delays and shown how it compares favorably to a state-of-the-art event-based classification algorithm for a visual motion detection task.
+The learned model shares a number of similarities with neurobiological anatomical observations, as well as with behavioral results.
+The event-based computations of our method can be drastically reduced by pruning synapses, while maintaining top classification performance.
+This shows that we can take advantage of the precise timing of spikes to improve the performance of neural computations.
 %

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
discussion : a

@@
 %%%-----------------------------------------------------------------
-To recap, we have introduced in this paper a SNN model with heterogeneous delays that we have trained and evaluated on a complex motion detection task.
-The model was defined to optimally detect event-driven spatio-temporal motifs.
-We have shown that when the model is trained on a dataset of natural images with realistic eye movements, the model learns kernels similar to those found in the early visual cortex of humans for example.
-We have then shown that the model displays similarities with the responses observed in biology.
-We have evaluated the computational cost of this model if implemented on neuromorphic hardware, showing that the use of heterogeneous delays may be a frugal computational solution for future neuromorphic hardware, but also a key to understand why spikes represent a universal component of neuronal information processing.
-
-Let us highlight some innovations in the contributions presented in this paper.
-First, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs, whereas~\citet{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we have observed to be less efficient.
-Moreover, in comparison to HOTS~\citep{lagorce_hots_2017} the weights are explainable as they directly inform on the logit (inverse sigmoid of the probability) of detecting each spatio-temporal spiking motif.
-Another novelty is that the model simultaneously learns the weights and the delays, while for example the polychronization model~\citep{izhikevich_polychronization_2006} only learns the weights using STDP while the delays are randomly drawn and their values frozen.
-Moreover, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified problems~\citep{gutig_tempotron_2006}.
-Another main contribution is to provide a model that is suitable for learning any type of spatio-temporal spiking motifs and that can be trained in a supervised way by providing a dataset of supervision pairs.
-This allows for a more flexible definition of the model using this properly labeled dataset instead of relying on a careful description of the physical rules governing a task, e.g., the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}.
+In summary, in this paper we have presented a heterogeneous delay SNN model that has been trained and evaluated on a complex motion detection task.
+The model has been defined to provide optimal detection of event-driven spatio-temporal motifs.
+We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in e.g.
+early visual cortex.
+We then demonstrated that the model shows similarities to biological responses.
+We have evaluated the computational cost of this model when implemented on neuromorphic hardware.
+We show that the use of heterogeneous delays may be an efficient computational solution for future neuromorphic hardware, but also a key to understanding why spikes are a universal component of neural information processing.
+
+We would like to highlight a few innovations in the contributions that are presented in this paper.
+First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we observed to be less efficient, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs.
+Moreover, in comparison to HOTS~\citep{lagorce_hots_2017}, the weights are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection of each spatio-temporal spike motif.
+Another novelty is that the model learns the weights and the delays simultaneously.
+For example, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn and their values are frozen during learning.
+In addition, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified toy problems~\citep{gutig_tempotron_2006}.
+Another major contribution is to provide a model that is suitable for learning any kind of spatio-temporal spiking motifs and that can be trained in a supervised manner by providing a dataset of supervision pairs.
+Instead of relying on a careful description of the physical rules governing a task, e.g.
+the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}, this allows a more flexible definition of the model using this properly labeled dataset.
 %

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
discussion : b

@@
 %%%-----------------------------------------------------------------
-We have identified several limits to our model, which we will now detail.
-First, the complete framework is based on a discrete binning of time that is incompatible with the continuous nature of biological time.
-We have used this binning to be able to efficiently implement the framework on conventional hardware, especially GPUs, and in particular to be able to use three-dimensional convolutions.
-We have tested the effect of the size of the time bin and shown that it has essentially no impact on the results presented in this paper.
-This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where the accuracy was not affected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}.
-This suggests the possibility of analytically including a precision term in the temporal value of the input spikes, a mechanism potentially implemented by the filtering that is implemented by the synaptic time constant of about $5~\ms$.
-Furthermore, it is possible to circonvent the necessity of using a time discretization by using a purely event-based scheme.
-Indeed, it is not necessary to compute voltage traces between two spikes~\citep{hanuschkin_general_2010} and it is thus possible to define a purely event-based framework.
-Such an architecture could provide promising speed gains for the calculations.
+We have identified a number of limitations of our model, which we will now discuss in detail.
+First, the entire framework is based on a discrete binning of time, which is not compatible with the continuous nature of biological time.
+We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, and especially to be able to use three-dimensional convolutions.
+We have tested the effect of the size of the time bin and shown that it has essentially no effect on the results presented in this paper.
+This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}.
+This suggests the possibility of analytically including a precision term in the temporal value of the input spikes.
+This mechanism may be implemented by the filtering implemented by the synaptic time constant of about $5~\ms$.
+Furthermore, it is possible to circumvent the need for time discretization by the use of a purely event-based scheme.
+In fact, it is not necessary to compute the voltage traces between two spikes~\citep{hanuschkin_general_2010}.
+Thus, it is possible to define a purely event-based framework.
+Such an architecture could provide promising computational speedups.
 % \note{pourrait parler du poster bernstein?
@@
 
-Another restriction is that this model is purely feed-forward.
-Therefore, the spikes generated by the postsynaptic neurons are produced solely on the basis of information contained in the classical receptive field.
-However, it is well known that neurons in a layer can interact using lateral interactions, for instance in V1 and that this can be the basis for computational principles~\citep{chavane_revisiting_2022}.
+A further limitation is that the model is purely feed-forward.
+Thus, the spikes generated by the postsynaptic neurons are based solely on information contained in the classical receptive field.
+However, it is well known that neurons in the same layer can interact with each other using lateral interactions, for example in V1, and that this can be the basis for computational principles~\citep{chavane_revisiting_2022}.
 For example, the combination of neighboring orientations may contribute to image categorization~\citep{perrinet_edge_2015}.
-Furthermore, neural information is modulated by feedback information, for example to distinguish a figure from its background~\citep{roelfsema_early_2016} and it has been shown that feedback may be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020}, Notably to explain non-linear mechanisms~\citep{boutin_pooling_2022}.
-It is currently not possible to implement these recurrent connections in our implementation (lateral or feedback), mainly due to our use of convolutions.
-However, the generic theoretical model is able to include them by inserting new spikes into the list of spikes that are reaching pre-synaptic addresses.
-While this is possible in theory, it needs to be properly adapted in practice so that these recurrent connections do not amplify neuronal activity outside a homeostatic state (through an extinction or explosion of activity).
+Furthermore, neural information is modulated by feedback information, e.g.
+to distinguish a figure from its background~\citep{roelfsema_early_2016}.
+Feedback has been shown to be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020}, especially to explain non-linear mechanisms~\citep{boutin_pooling_2022}.
+Currently, mainly due to our use of convolutions, it is not possible to implement these recurrent connections in our implementation (lateral or feedback).
+However, by inserting new spikes into the list of spikes reaching presynaptic addresses, the generic theoretical model is able to incorporate them.
+While theoretically possible, this needs to be properly adjusted in practice so that these recurrent connections do not amplify neuronal activity outside a homeostatic state (by extinction or explosion).
 % \note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}%~\citep{perrinet_motion-based_2012}

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
discussion : b

@@
 
-Such recurrent activity would be essential for implementing predictive or anticipatory processes.
-This is essential in a neural system, as it contains multiple different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}.
-This has been previously modeled to explain for example the flash-lag illusion~\citep{khoei_flash-lag_2017}.
+Such recurrent activity would be essential for the implementation of predictive or anticipatory processes.
+This is essential in a neural system because it contains several different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}.
+This has been modeled before to explain, for example, the flash-lag illusion~\citep{khoei_flash-lag_2017}.
 %\note{Say that this could be implemented by properly setting the timing of the label}
-As previously stated, this could be implemented using generalized coordinates (that is variables such as position which are complemented by speed, acceleration, jerk, \ldots) and ``neurobiologically, the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
-Our proposed model using heterogeneous delays provides an alternate and elegant implementation solution to this problem.
+As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by velocity, acceleration, jerk, \ldots), and ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
+Our proposed model using heterogeneous delays provides an alternative and elegant implementation solution to this problem.
 %

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 $\event_\postsynaddr = \{(\presynaddr_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}$.
-Unlike the binary stream of events as input, once integrated by the synapses, the postsynaptic neurons receives a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank$.
+Unlike the binary stream of events as input, once integrated by the synapses, the postsynaptic neurons receives a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank = \synapticweight_\ranksyn$ when $\presynaddr_\arank = \presynaddr_\ranksyn$.
 This new stream of events is by nature ordered in time as events reach the soma of post-synaptic neurons.

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
discussion : c

--- 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
@@
 % \note{la tache ne met pas en avant les capacites des motifs hetero synaptiques =trop simple et peut se faire en HOTS}
-In the definition of our task, we highlighted how the generation of events depends on the spatial gradient in each image.
-This gradient is defined in both horizontal and vertical dimensions and its maxima are generally oriented.
+In defining our task, we emphasized that the generation of events depends on the spatial gradient in each image.
+This gradient has both horizontal and vertical dimensions, and its maxima are generally orientation dependent.
 Taken together, these oriented edges form the contours of visual objects in the scene~\citep{koenderink_representation_1987}.
-Thus, there is a dependence within the event stream between motion and orientation information.
-It would be crucial to study this interdependence further.
-This could be initiated by training the model on a dataset with labels providing local orientation.
-%\note{we can say we tried and that it works well}.
-Studying this dependence will allow us to dissociate these two forms of visual information and to enable us to integrate them.
-In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka, the aperture problem), thus allowing us to implement recurrent predictive rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
+Thus, there is an interdependence between the information about motion and the information about orientation within the event stream.
+It would be crucial to investigate this dependency further.
+This could be initiated by training the model on a dataset with labels that provide local orientation.
+%\note{we can say that we tried it and that it works well}.
+Exploring this dependence will allow us to dissociate these two forms of visual information and enable us to integrate them.
+In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka the aperture problem).
+Thus, it will allow us to implement recurrent prediction rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
 
 The model is trained on a low-level local motion detection task, and one might wonder if it could be trained on higher-level tasks.
-An example task would be depth estimation in the visual scene.
-There are multiple sources of information for inferring depth, such as binocular disparity or changes in texture or shading, but in our case, motion parallax would be the most significant cue~\citep{rogers_motion_1979}.
-This is because objects close to an observer move relatively faster on the retina than an object at a great distance, and also that visual occlusions depend on depth ordering.
-Using this information, one can segment objects and estimate their depth~\citep{yoonessi_contribution_2011}.
-However, this would require computing first the optical flow, that is extending the framework described here for a rigid full-field motion to a more generic one where motion may vary in the visual field.
-A possible implementation is therefore to add a new layer to our model, in analogy with the hierarchical organization highlighted in the visual cortex.
+An example of such a task would be the estimation of depth in the visual scene.
+There are several sources of information for depth estimation, such as binocular disparity or changes in texture or shading, but in our case motion parallax would be the most important cue~\citep{rogers_motion_1979}.
+This is because objects that are close to an observer will move relatively faster on the retina than an object that is far away, and also because visual occlusions are dependent on the depth order.
+Using this information, it is possible to segment objects and estimate their depth~\citep{yoonessi_contribution_2011}.
+However, this would require the computation of the optical flow first, i.e., the extension of the framework described here for a rigid full-field motion to a more general one where the motion may vary in the visual field.
+A possible implementation is therefore to add a new layer to our model, analogous to the hierarchical organization highlighted in the visual cortex.
 This is theoretically possible by using the output of our model (which estimates velocity in retinotopic space) as input to a new layer of neurons that would estimate velocity in the visual field, including the depth dimension in the output supervision labels.
-This could have direct and important applications, for example in autonomous driving, to detect obstacles in a fast and robust way.
-Another extension would be to actively generate sensor movements (physically or virtually) to yield better depths estimates, notably to disambiguate uncertain estimates~\citep{nawrot_eye_2003}.
-
-In conclusion, the model we have presented provides a way to efficiently process event-based signals.
-We have shown that we can train the model in a semi-supervised manner, knowing \emph{what} output label but not knowing \emph{when} it occurs.
+This could have direct and important applications, e.g.
+in autonomous driving to detect obstacles in a fast and robust way.
+Another extension would be to actively generate sensor motion (physical or virtual) to obtain better depth estimates, especially to disambiguate uncertain estimates~\citep{nawrot_eye_2003}.
+
+In conclusion, the model that we have presented provides a way to process event-based signals in an efficient manner.
+We have shown that we can train the model semi-supervised, knowing \emph{what} output label, but not \emph{when} it occurs.
 Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data~\citep{barlow_unsupervised_1989}.
-This type of learning is thought to be predominant in the central nervous system and, assuming that the signal is sparse~\citep{olshausen_emergence_1996}, one could extend those Hebbian sparse learning schemes to spikes~\citep{perrinet_emergence_2004, masquelier_competitive_2009}.
-We anticipate that this would be particularly well suited to the exploration of neurobiological data.
-In fact, there is a substantial literature indicating that brain dynamics often organize into stereotyped sequences such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}.
-These motifs are stereotypical and robust, as they can be activated in the same motif from day to day~\citep{haimerl_internal_2019}.
-In contrast to conventional methodologies that are used to process neurobiological data, such a event-based model would be able to answer key questions regarding the representation of information in neurobiological data.
+This type of learning is thought to be prevalent in the central nervous system and, assuming the signal is sparse~\citep{olshausen_emergence_1996}, one could extend these Hebbian sparse learning schemes to spikes~\citep{perrinet_emergence_2004, masquelier_competitive_2009}.
+We expect that this would be particularly useful for exploring neurobiological data.
+Indeed, there is a large literature showing that brain dynamics often organize into stereotyped sequences, such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}.
+These patterns are stereotyped and robust, as they can be activated in the same pattern from day to day~\citep{haimerl_internal_2019}.
+In contrast to conventional methods of processing neurobiological data, such an event-based model would be able to answer key questions about the representation of information in neurobiological data, and it would open up possibilities in the field of computational neuroscience.
 Furthermore, it would open up possibilities in the field of machine learning, especially in computer vision, to address current key concerns such as robustness to attacks, scalability, interpretability, or energy consumption.

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 The model has been defined to provide optimal detection of event-driven spatio-temporal motifs.
-We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in e.g.
-early visual cortex.
+We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in the early visual cortex.
 We then demonstrated that the model shows similarities to biological responses.
-We have evaluated the computational cost of this model when implemented on neuromorphic hardware.
+We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware.
 We show that the use of heterogeneous delays may be an efficient computational solution for future neuromorphic hardware, but also a key to understanding why spikes are a universal component of neural information processing.
@@
 First, the entire framework is based on a discrete binning of time, which is not compatible with the continuous nature of biological time.
-We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, and especially to be able to use three-dimensional convolutions.
-We have tested the effect of the size of the time bin and shown that it has essentially no effect on the results presented in this paper.
+We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, to be able to use fast three-dimensional convolutions.
+We have tested the effect of the width of the time bin and shown that it has essentially no effect on the results presented in this paper.
 This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}.
@@
 Currently, mainly due to our use of convolutions, it is not possible to implement these recurrent connections in our implementation (lateral or feedback).
-However, by inserting new spikes into the list of spikes reaching presynaptic addresses, the generic theoretical model is able to incorporate them.
+However, by inserting new spikes into the list of spikes reaching presynaptic addresses, the generic model is able to incorporate them.
 While theoretically possible, this needs to be properly adjusted in practice so that these recurrent connections do not amplify neuronal activity outside a homeostatic state (by extinction or explosion).
@@
 %\note{Say that this could be implemented by properly setting the timing of the label}
-As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by velocity, acceleration, jerk, \ldots), and ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
+As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by velocity, acceleration, jerk,~\ldots), and ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
 Our proposed model using heterogeneous delays provides an alternative and elegant implementation solution to this problem.

### 2023-02-07 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %%%-----------------------------------------------------------------
-In summary, in this paper we have presented a heterogeneous delay SNN model that has been trained and evaluated on a complex motion detection task.
+In summary, in this paper we have presented a HD-SNN model that has been trained and evaluated on a complex motion detection task.
 The model has been defined to provide optimal detection of event-driven spatio-temporal motifs.
@@
 Results visualization and presentation were realized by both authors.
-The the manuscript was written by Laurent Perrinet.
+The manuscript was written by Laurent Perrinet.
 Both authors have read and approved the final manuscript.

### 2023-02-08 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
on 2.2.2

@@
 and from A*Midex grant number AMX-21-RID-025
-``\href{PolyChronies}{https://laurentperrinet.github.io/grant/polychronies/}''. 
+``\href{Polychronies}{https://laurentperrinet.github.io/grant/polychronies/}''. 
 %HL received funding from a University of Montreal Artificial Intelligence scholarship. 
@@
 \newcommand{\Acknowledgments}{
-Thanks to Thomas Schatz, Rosa Cossart, Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work. %
+Thanks to Thomas Schatz, Rosa Cossart, Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work.
+%Thanks to HL  for participating to Mustache November ©.
+%
 \Funding %
@@
 $\event_\postsynaddr = \{(\presynaddr_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}$.
-Unlike the binary stream of events as input, once integrated by the synapses, the postsynaptic neurons receives a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank = \synapticweight_\ranksyn$ when $\presynaddr_\arank = \presynaddr_\ranksyn$.
+Unlike the binary stream of events from the signal as input, the postsynaptic neurons receives a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank = \synapticweight_\ranksyn$ when $\presynaddr_\arank = \presynaddr_\ranksyn$.
 This new stream of events is by nature ordered in time as events reach the soma of post-synaptic neurons.
 In particular, when post-synaptic neurons are activated on their soma by this spatio-temporal motif, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.
+This model is defined in this subsection in all generality and the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions.
+In the next subsection we describe an implementation of our model adapted for the motion detection task.
+
+\note{I can make a small figure with event streams of different colors for the input one, the multiplexed one etc and the response in the bottom}
 
 \subsubsection{Applied on this detection task}
+
+From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation.
+Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{N\times T}$, where $N$ is the number of neurons in $\presynaddrspace$ and $T$ is the number of time bins.
+In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel^\postsynaddr(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$.
+Therefore, if we make the approximation that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train as input can be formalized by a 3D convolution operation.
+The longest synaptic delay defines the depth of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented.
+Using this dense representation, the integration of the spike train as input by the synapses of neuron $\postsynaddr$ is:
+\begin{equation*}
+\forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
+\mathcal{C}^\postsynaddr(\presynaddr,\timev)
+= \sum_{\ranksyn  \in \synapse^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
+\end{equation*}
+%
+This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
+This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
+%
+
+\note{formalize the following:}
+More generally, every cause may be considered as occurring independently and one might write the generative model for the generation of the presynaptic events as:
+
+Using this formalization, one might now deduce an optimal algorithm for the detection of such temporal motifs.
+
+By discretization of time (with here an arbitrary unitary time unit), we can also define the dendrite as a matrix giving the weight corresponding to the different delays $d \in [0, D]$ (where $D$ is the maximum delay) on different pre-synaptic addresses $a \in [1, N]$ defining the list of the $N$ dendrites.
+We will denote as $W(a, d)$ these weights.
+
+Following the observations of~\citet{izhikevich_polychronization_2006}, let us assume that such precise discharge motif defines a polychronous group (PG).
+Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
+This allows then to derive a generative model for raster plots (see Figure~\ref{fig:model}-(b)).
+
+This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$.
+This assumes that we know the spiking motifs as defined by the $W_b$ matrices.
+The underlying metric is the binary cross-entropy, as used in the logistic regression model.
+In particular, if we consider kernels with similar decreasing exponential time profile, we can prove that this is similar to the method of~\citet{berens_fast_2012}.
+In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator.
+Using this forward model, it is possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator.
+Equivalently, this consists in using the emitting field $\synapse_\presynaddr$ of presynaptic neurons in place of the receptive field $\synapse^\postsynaddr$ of postsynaptic neurons.
+It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.
+This assumption hold as long as the kernels are uncorrelated, a condition which is met here numerically by chosing a relatively sparse set of synapses (approximately $1\%$ of active synapses).
+
+\subsubsection{Implementation using spiking neurons with heterogeneous delays}
+
+The activation function of our spiking neural is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
+It transforms this list of weights into a probability with the following formula:
+$
+Pr(k=\postsynaddr \; \vert \; \timev) =
+\frac 1 Z
+{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
+$ 
+where $\mathcal{C}^\postsynaddr(\timev) = \sum
+\activeweights^\postsynaddr(t)
+$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
+In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
+%
+
+ \note{say that it is effortless in biology but difficult on conventional computers}
+ In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the heterogeneous delays model.
+
 Now that our motion detection task is defined in visual space, we can define the model that will be trained to solve the task.
@@
 For clarity, only $4$ kernels are illustrated and some weights are not represented in the voxel grid.
-For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model.
-The output of the MLR gives a probability for each class associated to a specific kernel to emit an output event, or spike.
+For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e.
+softmax function).
+The output of the MLR gives a probability for each class, associated to a specific kernel, to emit an output event, or spike.
 With an appropriate spiking mechanism, here the winner-takes-all mechanism, we obtained, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class.
@@
 Results visualization and presentation were realized by both authors.
-The manuscript was written by Laurent Perrinet.
+The manuscript was written by both authors.
 Both authors have read and approved the final manuscript.

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Using this dense representation, the integration of the spike train as input by the synapses of neuron $\postsynaddr$ is:
-\begin{equation*}
+%
+\begin{equation}\label{eq:kernel_b}
+\forall \presynaddr \in \presynaddrspace^{\synapse^\postsynaddr}, \forall \timev \in T, \; \;
+\mathcal{C}^\postsynaddr(\presynaddr,\timev)
+= \sum_{\ranksyn  \in [1, N_{\synapse^\postsynaddr}]} \kernel^\postsynaddr(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
+\end{equation}
+%
+where $\presynaddrspace^{\synapse^\postsynaddr}$ is the address space of the receptive field of neuron $\postsynaddr$ and $N_{\synapse^\postsynaddr}$ is the number of synapses belonging to neuron $\postsynaddr$.
+To benefit from invariance to location observed in CNN and to simplify the implementation of the model, one assumes that kernels should be similar across different positions and define a convolution operator.
+The characteristics of the kernels are shared among one class of neurons such that $\forall i,j \in [1, N_\class]^2, \; \kernel^i_\class = \kernel^j_\class$.
+The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$:
+%
+\begin{equation}\label{eq:conv}
 \forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
-\mathcal{C}^\postsynaddr(\presynaddr,\timev)
-= \sum_{\ranksyn  \in \synapse^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
-\end{equation*}
-%
-This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
-This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
-%
-
-\note{formalize the following:}
-More generally, every cause may be considered as occurring independently and one might write the generative model for the generation of the presynaptic events as:
-
-Using this formalization, one might now deduce an optimal algorithm for the detection of such temporal motifs.
-
-By discretization of time (with here an arbitrary unitary time unit), we can also define the dendrite as a matrix giving the weight corresponding to the different delays $d \in [0, D]$ (where $D$ is the maximum delay) on different pre-synaptic addresses $a \in [1, N]$ defining the list of the $N$ dendrites.
-We will denote as $W(a, d)$ these weights.
-
-Following the observations of~\citet{izhikevich_polychronization_2006}, let us assume that such precise discharge motif defines a polychronous group (PG).
-Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
-This allows then to derive a generative model for raster plots (see Figure~\ref{fig:model}-(b)).
-
+\mathcal{C}_\class(\presynaddr,\timev)
+= \sum_{\ranksyn  \in [1, N_{\synapse^\postsynaddr}]} \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
+\end{equation}
+%
+This shows that $\mathcal{C}_\class$ is a spatiotemporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$.
+This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
+%
 This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$.

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
3D conv

@@
 \newcommand{\arank}{r} % address index
-\newcommand{\bias}{b} % bias for the MLR model
+\newcommand{\bias}{\beta} % bias for the MLR model
 \newcommand{\synapse}{\mathcal{S}} % synapse
@@
 %
-This shows that $\mathcal{C}_\class$ is a spatiotemporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$.
+This shows that $\mathcal{C}_\class$ is a spatiotemporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration).
+Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
 This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
 %
-This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$.
+The same type of spatiotemporal filtering is used as a preprocessing stage of a pattern recognition algorithm~\cite{ghosh_spatiotemporal_2019}.
+And in \cite{sekikawa_constant_2018}, an efficient 3D convolutional algorithm succeeds a motion estimation task.
+By assuming a locally constant velocity the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.
+Here, we keep the analogy with spiking neurons and we try to use as few heuristics as possible to observe the emergence of spiking motifs in the spatiotemporal kernels.
+If so, this precise spatiotemporal patterns prove to be of interest for neural computations and one can assume that biological neurons make use of this information as well.
+
+The 3D convolution represents the linear integration of the spike train as input.
+The activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
+It transforms this list of weights into a probability with the following formula:
+\begin{equation}\label{eq:mlr}
+\forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \forall \class \in [1, N_\class], \; \;
+Pr(k=\class \; \vert \; \presynaddr, \timev) =
+\frac 1 Z
+{\exp  (\mathcal{C}_\class(\presynaddr, \timev) +\bias_\class) }
+\end{equation} 
+where $\bias^\postsynaddr$ is the bias linked to class $\class$. 
+In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
+%This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$.
 This assumes that we know the spiking motifs as defined by the $W_b$ matrices.
@@
 $
-Pr(k=\postsynaddr \; \vert \; \timev) =
+\forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
+Pr(k=\class \; \vert \; \timev) =
 \frac 1 Z

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
after lunch... where were I?

@@
 \end{equation} 
-where $\bias^\postsynaddr$ is the bias linked to class $\class$. 
+where $\bias_\class$ is the bias linked to class $\class$ and $Z$ is the partition function. 
+In our MLR model, a probability value, for each class (i.e.
+each motion direction), is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the HD-SNN model.
 In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
-%This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$.
-This assumes that we know the spiking motifs as defined by the $W_b$ matrices.
-The underlying metric is the binary cross-entropy, as used in the logistic regression model.
-In particular, if we consider kernels with similar decreasing exponential time profile, we can prove that this is similar to the method of~\citet{berens_fast_2012}.
-In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator.
-Using this forward model, it is possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator.
-Equivalently, this consists in using the emitting field $\synapse_\presynaddr$ of presynaptic neurons in place of the receptive field $\synapse^\postsynaddr$ of postsynaptic neurons.
-It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.
-This assumption hold as long as the kernels are uncorrelated, a condition which is met here numerically by chosing a relatively sparse set of synapses (approximately $1\%$ of active synapses).
-
-\subsubsection{Implementation using spiking neurons with heterogeneous delays}
-
-The activation function of our spiking neural is a softmax function implementing a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
-It transforms this list of weights into a probability with the following formula:
-$
-\forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
-Pr(k=\class \; \vert \; \timev) =
-\frac 1 Z
-{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
-$ 
-where $\mathcal{C}^\postsynaddr(\timev) = \sum
-\activeweights^\postsynaddr(t)
-$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
-In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
-%
-
- \note{say that it is effortless in biology but difficult on conventional computers}
- In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
-The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the heterogeneous delays model.
-
-Now that our motion detection task is defined in visual space, we can define the model that will be trained to solve the task.
-The use of spatio-temporal filters on an event stream has improved the performance of CNNs for an action recognition task in a model by~\citet{ghosh_spatiotemporal_2019} and here we will use a similar strategy, but focus on using an event-based representation.
-For this purpose, we design 3D kernels of shape $(\Kx, \Ky, \Ktime) = (9, 9, 11)$, representing the two spatial dimensions and the range of time delays, respectively.
-The computations are performed on the spatio-temporal windows that are defined by the kernels and are ``moving'' around the events, i.e.
-the spatial center of the spatio-temporal window around the current event $\epsilon_\arank$.
-Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
-%
+The spiking output of the model corresponds to an event with the highest probability class, keeping the same timing as the input event.
+
+\note{describe the heuristics inspired by neuroscience}
+
+\note{describe the parameters of the model}
+
+\note{describe the learning -> analogy with hebbian learning from TPAMI}
+
 \begin{figure}%[h!]%[ht!]
@@
 The output of the MLR gives a probability for each class, associated to a specific kernel, to emit an output event, or spike.
-With an appropriate spiking mechanism, here the winner-takes-all mechanism, we obtained, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class.
+With an appropriate spiking mechanism, here a winner-takes-all, we obtain, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class.
 }

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
method first draft

@@
 %
-\subsection{Task definition: fast motion detection}
+\subsection{Task definition: fast motion detection}\label{sec:task}
 %%%-----------------------------------------------------------------
@@
 Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{N\times T}$, where $N$ is the number of neurons in $\presynaddrspace$ and $T$ is the number of time bins.
-In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel^\postsynaddr(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$.
+In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ of size $(N_{tb}, N_x, N_y)$, where $N_{tb}$ is the number of time bins, $N_x$ and $N_y$ are the number of pixels in both spatial dimensions.
+For each neuron $\postsynaddr$, $\kernel^\postsynaddr$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel^\postsynaddr(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$.
 Therefore, if we make the approximation that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train as input can be formalized by a 3D convolution operation.
@@
 
+Once explained the general framework used to solve the ecological task described in section~\ref{sec:task}, we can add some heuristics based on neuroscientific observations to constrain our model and its strategies to solve the task.
+We set \note{arbitrarily} the size of the kernels to $(9, 17, 17)$ and define as many classes as the number of motion direction: $N_\class = 8$.
+Because receptive fields of biological neurons are mainly circular and can be more or less elongated, we apply a circular mask on the spatial dimensions of the kernels.
+And because the synaptic delay is linked, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one, we apply a linear mask on the temporal axis such that the connectivity depends on the delay: the lower the delay the lower the circular mask radius (see Figure~\ref{fig:kernels} for an illustration).
+The second heuristics is based on the observation that biological neurons have a sparse spiking activity.
+From this observation, we set an output firing rate for the layer of spiking neurons in order to make a sparse selection of the events at which the model makes a decision.
+It results in the selection of the coefficients with the maximum likelihood value.
+Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels.
+This threshold was found using cross-validation and future work should automate how this may be determined for various datasets.
+Given a stimulus input, the model thus yields a spike output in the postsynaptic address space.
+
 \note{describe the heuristics inspired by neuroscience}
@@
 \note{describe the parameters of the model}
-
-\note{describe the learning -> analogy with hebbian learning from TPAMI}
 
@@
 %
-The output of the MLR model corresponds to an event with the highest probability class, keeping the same timing as the input event.
-It is important to note that we include the prior knowledge that the event-based input is spatiotemporally clustered and that these clusters are sparse.
-Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels.
-This threshold was found using cross-validation and future work should automate how this may be determined for various datasets.
-Given a stimulus input, the model thus yields a spike output in the postsynaptic address space.
-
-%
 Since the model is fully differentiable, we can now implement a semi-supervised learning rule.
-The loss function of the MLR model is the binary cross-entropy on the output of the classification layer that can be extended to space-time convolutions.
-Supervision was implemented using the input binary events as defined above and the labels as the desired output.
+The loss function of the MLR model is the binary cross-entropy on the output of the classification layer.
+Supervision was implemented using the input binary events as defined above and the motion direction labels as the desired output.
 The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions.
@@
 Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$).
-We tested the effect of different parameters on a validation set to quantify the role of each parameter.
+We demonstrated in a previous work that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
 %
@@
 By taking the logit of the probability of the output given by the model, we can therefore calculate the probability of the output.
-This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error % $|| \speed - ||^2$
-of the velocity estimate given by the average estimate. 
-%$\hat{\speed} = \sum_{\class} \speed_\class \cdot Pr(\class \vert image)$ 
+This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error of the velocity estimate given by the average estimate.
 These calculations are performed on a different input dataset than the one used in the training or validation steps.
-These metrics will be useful to compare our methods with other methods on our defined task.
 The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/pyTERtorch}.
+\note{à rendre publique}
 %

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
methods

--- 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
@@
 %Let us now define a procedure of animating of natural scene by virtual eye-movements similar to that which was used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
-We want to test our model on a motion detection task on a binary stream of events.
-This type of signal, typically recorded by a Dynamic Vision Sensor (DVS), is a good representation of the information sent by the retinal ganglion cells through the optic nerve.
+We want to test our model on a motion detection task on a realistic stream of events.
+This type of signal, typically recorded by a Dynamic Vision Sensor (DVS), is a good representation of the information and inspired by that sent by the retinal ganglion cells through the optic nerve.
 With binary events, the visual system is able to capture the dynamics of a natural scene and perform various cognitive tasks in a fast and efficient fashion.
@@
 %%%-----------------------------------------------------------------
-\subsubsection{The general model}
+\subsubsection{The heterogeneous synaptic delay model}
 

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
methods : A

@@
 %Let us now define a procedure of animating of natural scene by virtual eye-movements similar to that which was used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
-We want to test our model on a motion detection task on a realistic stream of events.
-This type of signal, typically recorded by a Dynamic Vision Sensor (DVS), is a good representation of the information and inspired by that sent by the retinal ganglion cells through the optic nerve.
-With binary events, the visual system is able to capture the dynamics of a natural scene and perform various cognitive tasks in a fast and efficient fashion.
-To train and test our model, we need to define a dataset for which we control for motion direction and velocity.
-First, let's define a trajectory inspired by the biological movements of the eyes.
-Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the field of vision.
-In animals with a fovea, this is particularly useful as it allows to move the area with the highest density of photoreceptors in the environment, for example at a point of interest.
-A specific mechanism to do so are called saccades which are rapid eye movements of the eyeball that reposition the center of vision.
-In humans, these are very frequent (on average $2$ each second~\citep{dandekar_neural_2012}).
-They are produced very rapidly (about $80~\ms$) and display a whole range of speeds.
-At a more microscopic scale, the human gaze moves with an incessant drift similar to a Brownian-like trajectory~\citep{poletti_head-eye_2015}.
-To maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.
-This approach first defines a finite set of possible 2D motions in polar coordinates.
-Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of % N-V-phi
-$8$ linearly spaced motion directions.
-Next, we define one trajectory of gaze as segments whose duration is drawn from a Poisson distribution with an average block length of $20~\ms$, similarly to a Lévy flight~\citep[p.
+The goal is to test the model in a motion detection task using a realistic data stream.
+This type of signal is typically recorded by a Dynamic Vision Sensor (DVS).
+It is a good representation of the information and is inspired by the information sent by retinal ganglion cells through the optic nerve.
+Binary events allow the visual system to capture the dynamics of the natural scene and carry out different cognitive tasks quickly and effectively.
+To train and test our model, we need to define a dataset for which we control the direction and speed of motion.
+First of all, let's define a trajectory that is inspired by the biological movements of the eyes.
+Indeed, these movements allow us to dynamically actuate the centre of vision, or gaze, in the visual field.
+In animals with a fovea, this is particularly useful as it allows the area with the highest density of photoreceptors in the environment to be moved, for example to a point of interest.
+A specific mechanism for this is called saccades, which are rapid eye movements that reposition the centre of vision.
+In humans, saccades are very frequent (on average $2 per second~\citep{dandekar_neural_2012}).
+They are generated very quickly (about $80~\ms$) and they have a whole range of speeds.
+On a more microscopic scale, the human gaze moves with a continuous drift similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
+In order to maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.
+This approach first defines a finite set of possible 2D movements in polar coordinates.
+Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of $8$ linearly spaced motion directions.
+Next, we define a gaze trajectory as segments whose duration is drawn from a Poisson distribution with a mean block length of $20~\ms$, similar to a Lévy flight~\citep[p.
 289]{mandelbrot_fractal_1982}.
-Finally, the trajectory is integrated by assuming first that velocities are uniformly and independently sampled from the set of different motion-sets and second that motion is uniform during a time segment.
-The resulting instances yield trajectories qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(Left)).
+Finally, the trajectory is integrated, assuming first that the velocities are sampled uniformly and independently from the set of different motion sets, and second that the motion is uniform during a time segment.
+The resulting instances yield trajectories that are qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(left)).
 %%%-----------------------------------------------------------------

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
methods : caption

@@
     \caption{
-    {\bf - Motion detection task.} {\bf (Left)}~We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view.
-To mimic the effect of a saccadic-like eye movement, the axis of view moves following a step-like random walk, and we show here an example path of the trajectory as a green line ($200$ time steps).
-{\bf (Right)}~Snapshot of the synthesized event stream at different time steps.
-The dynamics of the sub-image as a function of time generates a naturalistic movie which may be transformed into an event-based representation.
-Mimicking the retina, this representation codes for proportional and pixel-wise increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events.
-On the bottom left corner of the images, the translation vector are plotted in red among the different possible classes of movements.
-Note the change of direction between the second and third frames.}
+    {\bf - Motion detection task.} {\bf (left)}~We use large natural images ($512\times512$) in which an aperture ($128\times128$) extracts a cropped  image around the axis of view.
+To mimic the effect of a saccadic eye movement, the view axis moves according to a stepwise random walk, and we show an example path of the trajectory as a green line ($200$ time steps).
+{\bf (right)}~Snapshot of the synthetic event stream at different time steps.
+The dynamics of the sub-image as a function of time generates a naturalistic film.
+This can be transformed into an event-based representation.
+Mimicking the retina, this representation encodes proportional and pixel-wise increases or decreases in luminance in the image, respectively ON (red) and OFF (blue) events.
+In the lower left corner of the images, the translation vector is shown in red.
+It is one of the possible classes of motion.
+Note the change in the direction of motion between the second and the third image.}
     \label{fig:motion_task}

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
methods : A 2

@@
 %%%-----------------------------------------------------------------
-Once these eye movement trajectories are generated, we can apply them to a visual scene.
+Once these eye trajectories are generated, we can apply them to a visual scene.
 For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images~\citep{olshausen_emergence_1996}.
-Note that these are pre-processed to equalize the energy in each frequency band (i.e., whitened).
-This process is known to occur as early as the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step.
-We will discretize the time in $1~\ms$ bins and produce movies of duration $N_T = 250~\ms$.
-To avoid border effects, we will position a draw of the complete trajectory at random in the image space so that the sub-image is translated using the position given by the trajectory at each time step.
-The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
-Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that it corresponds to a motion of one pixel per frame (i.e., per time bin).
+Note that these have been pre-processed to equalize (i.e., whiten) the energy in each frequency band.
+This process is known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
+These images are $512 \times 512$ in size, and we will extract sub-images of size $128 \times 128$ positioned around the center of gaze at each time step.
+We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 250~\ms$.
+To avoid boundary effects, we will randomly position a plot of the full trajectory in image space so that the subimage is translated using the position given by the trajectory at each time step.
+The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a subpixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
+Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity to correspond to a movement of one pixel per frame (i.e., per time bin).
 

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
methods : caption

@@
     \caption{
-    {\bf - Motion detection task.} {\bf (left)}~We use large natural images ($512\times512$) in which an aperture ($128\times128$) extracts a cropped  image around the axis of view.
-To mimic the effect of a saccadic eye movement, the view axis moves according to a stepwise random walk, and we show an example path of the trajectory as a green line ($200$ time steps).
+        {\bf - Motion Detection Task.} {\bf (left)}~We use large natural images ($512\times512$) in which an aperture ($128\times128$) extracts a cropped image around the view axis.
+To mimic the effect of a saccadic eye movement, the view axis moves according to a stepwise random walk.
+We show an example path of the trajectory for $200$ time steps (green line).
 {\bf (right)}~Snapshot of the synthetic event stream at different time steps.
-The dynamics of the sub-image as a function of time generates a naturalistic film.
+The dynamics of the sub-image as a function of time produces a naturalistic movie.
 This can be transformed into an event-based representation.
-Mimicking the retina, this representation encodes proportional and pixel-wise increases or decreases in luminance in the image, respectively ON (red) and OFF (blue) events.
+Mimicking the retina, this representation encodes proportional and pixel-wise increases or decreases in luminance in the image, i.e.
+ON (red) and OFF (blue) events.
 In the lower left corner of the images, the translation vector is shown in red.
 It is one of the possible classes of motion.
-Note the change in the direction of motion between the second and the third image.}
+Note the change in the direction of motion between the second and third image.}
     \label{fig:motion_task}

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
methods : A 1

@@
 %%%-----------------------------------------------------------------
-%Let us now define a procedure of animating of natural scene by virtual eye-movements similar to that which was used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
 The goal is to test the model in a motion detection task using a realistic data stream.
@@
 It is a good representation of the information and is inspired by the information sent by retinal ganglion cells through the optic nerve.
-Binary events allow the visual system to capture the dynamics of the natural scene and carry out different cognitive tasks quickly and effectively.
-To train and test our model, we need to define a dataset for which we control the direction and speed of motion.
-First of all, let's define a trajectory that is inspired by the biological movements of the eyes.
-Indeed, these movements allow us to dynamically actuate the centre of vision, or gaze, in the visual field.
-In animals with a fovea, this is particularly useful as it allows the area with the highest density of photoreceptors in the environment to be moved, for example to a point of interest.
-A specific mechanism for this is called saccades, which are rapid eye movements that reposition the centre of vision.
+Binary events allow the visual system to capture the dynamics of the natural scene and perform various cognitive tasks quickly and effectively.
+To train and test our model, we need to define a data set for which we control the direction and speed of motion.
+Let us now define a procedure for animating a natural scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
+First, let's define a trajectory inspired by the biological movements of the eyes.
+Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the visual field.
+In animals with a fovea, this is particularly useful because it allows the area with the highest density of photoreceptors in the environment to be moved, for example to a point of interest.
+A specific mechanism for this is called saccades, which are rapid eye movements that reposition the center of vision.
 In humans, saccades are very frequent (on average $2 per second~\citep{dandekar_neural_2012}).
-They are generated very quickly (about $80~\ms$) and they have a whole range of speeds.
+They are generated very quickly (about $80~\ms$) and have a wide range of speeds.
 On a more microscopic scale, the human gaze moves with a continuous drift similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
-In order to maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.
-This approach first defines a finite set of possible 2D movements in polar coordinates.
-Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of $8$ linearly spaced motion directions.
+In order to maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.This approach first defines a finite set of possible 2D movements in polar coordinates.
+Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $8$ linearly spaced movement directions and 3 geometrically spaced speeds.
 Next, we define a gaze trajectory as segments whose duration is drawn from a Poisson distribution with a mean block length of $20~\ms$, similar to a Lévy flight~\citep[p.

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
methods : A 1
response to reviewer

--- 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
@@
 % https://www.zotero.org/groups/4776796/fastmotiondetection
-
-% private
-% https://framateam.org/spikeai/channels/2022-10_icip-bordeaux
-% https://www.overleaf.com/project/630f9044c38e7a3cea81a7b2
-
-% preprint server =  https://www.researchsquare.com/login?redirect=/article/rs-2120999/private/timeline
-% 
-% 2023-01-13 decision: Revise
-% 
-% REVIEWER REPORTS
-
-% Reviewer Comments:
-
-% Reviewer 1
-% 1.There is a lack of an intuitive figure to show the HD-SNN model used in this paper.
-
-% 2.The MLR model used in the experimental part was presented in authors’ another work. Therefore, the changes on the MLR model are suggested to be highlighted and contrasted graphically.
-
-% 3.The title of section 2.2.1 is ’’a generative model for rater plots’’, but actually this section introduces basic terms and symbols.
-Section 2.2.2 is named ‘’detecting spiking motifs’’, but most of the content seems to introduce the generation process of raster plots, while only the last paragraph introduces the detection of motif.
-It is recommended that the content of these two sections be reorganized and that references to Figure 1(b)(c)(d) be added where appropriate during the description.
-
-% 4.The adjustment of weights and delays needs more detailed elaboration, preferably given in mathematical formulas.
-
-% 5.In Figure 4, it is claimed that the kernels are very similar for the ON polarities, and different kernels are selective to the different motion directions.
-However, in terms of these two observations, I think they are not very significant, so I suggest a more intuitive way to reflect them, such as quantifying the similarity between the ON kernels, and the angle between the kernels and the vector of the standard motion directions.
-
-% 6.There are few experiments and lack of comparative experiments.
-Since the model has an improvement based on MLR model, it is suggested to increase the comparison with MLR model and HOTS model on other datasets, such as N-MNIST and other neuromorphic datasets, so as to show the advantages of the proposed model in computational consumption.
-
-% 7.Some sections are verbose, such as the discussion.
-
-% 8.There are some minor errors in the text, and it is recommended to check the full text: 
-% 1)In the first paragraph on page 8, "a presynaptic address b_s" should be "a postsynaptic address b_s"; 
-% 2)Whether W_b in the first formula should be K_b? And d in it lacks description. 
-% 3)"PGs" should be given in its full name when it first appears. 
-% 4)There is no subgraph label of (a), (b) and (c) in Fig. 3, but the expression of Fig. 3-(a)/(b)/(c) appears many times. 
-% 5)The content of paragraph 1 on page 13 is repeated.
-
-% Reviewer 2
-% Key results: Please summarize what you consider to be the outstanding features of the work.
-% Using a single layer of spiking neurons with heterogeneous delays to learn spatio-temporal  spiking motifs
-
-% Validity: Does the manuscript have flaws which should prohibit its publication? If so, please provide details.
-
-% The work presented here is very similar to the author's other work,  "Learning hetero-synaptic delays for motion detection in a single layer of spiking neurons" 10.1109/ICIP46576.2022.9897394.
-This must be addressed and cited with an explanation of how this work presented is an extension of this previous publication.
-Abstract, Fig 6 and the main result is presented in the previous work.
-Arguably some of the figures from the previous publication would be nice to see in this publication as it helps with the visualisation of the problem.
-One suggestion for the extension to make the work unique would be to extend with some event-camera data, as it is mentioned multiple times within the publication.
-% There are also a number of citations that could have been added that did appear in the other publication, while also including other motion detection/ segmentation or previous work on event-based optical flow, which is a very similar problem.
-
-% Originality and significance: If the conclusions are not original, please provide relevant references.
-
-% See above...
 
@@
 }
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \usepackage[utf8]{inputenc}
+%
+\usepackage{lineno}
+\linenumbers
 
@@
 %Let us now define a procedure of animating of natural scene by virtual eye-movements similar to that which was used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
-The goal is to test the model in a motion detection task using a realistic data stream.
-This type of signal is typically recorded by a Dynamic Vision Sensor (DVS).
-It is a good representation of the information and is inspired by the information sent by retinal ganglion cells through the optic nerve.
-Binary events allow the visual system to capture the dynamics of the natural scene and carry out different cognitive tasks quickly and effectively.
-To train and test our model, we need to define a dataset for which we control the direction and speed of motion.
-First of all, let's define a trajectory that is inspired by the biological movements of the eyes.
-Indeed, these movements allow us to dynamically actuate the centre of vision, or gaze, in the visual field.
-In animals with a fovea, this is particularly useful as it allows the area with the highest density of photoreceptors in the environment to be moved, for example to a point of interest.
-A specific mechanism for this is called saccades, which are rapid eye movements that reposition the centre of vision.
-In humans, saccades are very frequent (on average $2 per second~\citep{dandekar_neural_2012}).
-They are generated very quickly (about $80~\ms$) and they have a whole range of speeds.
-On a more microscopic scale, the human gaze moves with a continuous drift similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
-In order to maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.
-This approach first defines a finite set of possible 2D movements in polar coordinates.
-Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of $8$ linearly spaced motion directions.
-Next, we define a gaze trajectory as segments whose duration is drawn from a Poisson distribution with a mean block length of $20~\ms$, similar to a Lévy flight~\citep[p.
+We want to test our model on a motion detection task on a binary stream of events.
+This type of signal, typically recorded by a Dynamic Vision Sensor (DVS), is a good representation of the information sent by the retinal ganglion cells through the optic nerve.
+With binary events, the visual system is able to capture the dynamics of a natural scene and perform various cognitive tasks in a fast and efficient fashion.
+To train and test our model, we need to define a dataset for which we control for motion direction and velocity.
+First, let's define a trajectory inspired by the biological movements of the eyes.
+Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the field of vision.
+In animals with a fovea, this is particularly useful as it allows to move the area with the highest density of photoreceptors in the environment, for example at a point of interest.
+A specific mechanism to do so are called saccades which are rapid eye movements of the eyeball that reposition the center of vision.
+In humans, these are very frequent (on average $2$ each second~\citep{dandekar_neural_2012}).
+They are produced very rapidly (about $80~\ms$) and display a whole range of speeds.
+At a more microscopic scale, the human gaze moves with an incessant drift similar to a Brownian-like trajectory~\citep{poletti_head-eye_2015}.
+To maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.
+This approach first defines a finite set of possible 2D motions in polar coordinates.
+Based on the distribution of biological motions, we simplified it by selecting a set of eye movements as the Cartesian product of % N-V-phi
+$8$ linearly spaced motion directions.
+Next, we define one trajectory of gaze as segments whose duration is drawn from a Poisson distribution with an average block length of $20~\ms$, similarly to a Lévy flight~\citep[p.
 289]{mandelbrot_fractal_1982}.
-Finally, the trajectory is integrated, assuming first that the velocities are sampled uniformly and independently from the set of different motion sets, and second that the motion is uniform during a time segment.
-The resulting instances yield trajectories that are qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(left)).
+Finally, the trajectory is integrated by assuming first that velocities are uniformly and independently sampled from the set of different motion-sets and second that motion is uniform during a time segment.
+The resulting instances yield trajectories qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(Left)).
 %%%-----------------------------------------------------------------
@@
     \caption{
-        {\bf - Motion Detection Task.} {\bf (left)}~We use large natural images ($512\times512$) in which an aperture ($128\times128$) extracts a cropped image around the view axis.
-To mimic the effect of a saccadic eye movement, the view axis moves according to a stepwise random walk.
-We show an example path of the trajectory for $200$ time steps (green line).
-{\bf (right)}~Snapshot of the synthetic event stream at different time steps.
-The dynamics of the sub-image as a function of time produces a naturalistic movie.
-This can be transformed into an event-based representation.
-Mimicking the retina, this representation encodes proportional and pixel-wise increases or decreases in luminance in the image, i.e.
-ON (red) and OFF (blue) events.
-In the lower left corner of the images, the translation vector is shown in red.
-It is one of the possible classes of motion.
-Note the change in the direction of motion between the second and third image.}
+    {\bf - Motion detection task.} {\bf (Left)}~We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view.
+To mimic the effect of a saccadic-like eye movement, the axis of view moves following a step-like random walk, and we show here an example path of the trajectory as a green line ($200$ time steps).
+{\bf (Right)}~Snapshot of the synthesized event stream at different time steps.
+The dynamics of the sub-image as a function of time generates a naturalistic movie which may be transformed into an event-based representation.
+Mimicking the retina, this representation codes for proportional and pixel-wise increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events.
+On the bottom left corner of the images, the translation vector are plotted in red among the different possible classes of movements.
+Note the change of direction between the second and third frames.}
     \label{fig:motion_task}
@@
 %%%-----------------------------------------------------------------
-Once these eye trajectories are generated, we can apply them to a visual scene.
+Once these eye movement trajectories are generated, we can apply them to a visual scene.
 For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images~\citep{olshausen_emergence_1996}.
-Note that these have been pre-processed to equalize (i.e., whiten) the energy in each frequency band.
-This process is known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $512 \times 512$ in size, and we will extract sub-images of size $128 \times 128$ positioned around the center of gaze at each time step.
-We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 250~\ms$.
-To avoid boundary effects, we will randomly position a plot of the full trajectory in image space so that the subimage is translated using the position given by the trajectory at each time step.
-The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a subpixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
-Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity to correspond to a movement of one pixel per frame (i.e., per time bin).
+Note that these are pre-processed to equalize the energy in each frequency band (i.e., whitened).
+This process is known to occur as early as the retino-thalamic pathway~\citep{dan_efficient_1996}.
+These images are $512 \times 512$ in size and we will extract sub-images of size $128 \times 128$ which will be positioned around the center of the gaze at each time step.
+We will discretize the time in $1~\ms$ bins and produce movies of duration $N_T = 250~\ms$.
+To avoid border effects, we will position a draw of the complete trajectory at random in the image space so that the sub-image is translated using the position given by the trajectory at each time step.
+The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
+Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity such that it corresponds to a motion of one pixel per frame (i.e., per time bin).
 
@@
 Using this dense representation, the integration of the spike train as input by the synapses of neuron $\postsynaddr$ is:
-%
+
 \begin{equation}\label{eq:kernel_b}
@@
 \end{equation}
-%
+
 where $\presynaddrspace^{\synapse^\postsynaddr}$ is the address space of the receptive field of neuron $\postsynaddr$ and $N_{\synapse^\postsynaddr}$ is the number of synapses belonging to neuron $\postsynaddr$.
@@
 The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$:
-%
+
 \begin{equation}\label{eq:conv}
@@
 \end{equation}
-%
+
 This shows that $\mathcal{C}_\class$ is a spatiotemporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration).
@@
 
+\note{plus de biblio sur la motion detection/segmentation ou optique flow avec les events? (reviewer 2)}
+
 The 3D convolution represents the linear integration of the spike train as input.
@@
 It transforms this list of weights into a probability with the following formula:
+
 \begin{equation}\label{eq:mlr}
@@
 \end{equation}
+
 where $\bias_\class$ is the bias linked to class $\class$ and $Z$ is the partition function. 
@@
 Once explained the general framework used to solve the ecological task described in section~\ref{sec:task}, we can add some heuristics based on neuroscientific observations to constrain our model and its strategies to solve the task.
+Note that the general framework is the same as the one presented in~\cite{grimaldi_learning_2022} and that, appart from the task to solve and a deeper analysis of the results, these priors, inspired by neuroscience, are the main methodological differences with this previous study.
 We set \note{arbitrarily} the size of the kernels to $(9, 17, 17)$ and define as many classes as the number of motion direction: $N_\class = 8$.

### 2023-02-09 cover_letter/2023-02-10_coverletter.tex
response to reviewer

%!TeX TS-program = lualatex
%!TeX encoding = UTF-8 Unicode
%!TeX spellcheck = fr-FR
%!BIB TS-program = biber
% -*- coding: UTF-8; -*-
% vim: set fenc=utf-8
%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------
\documentclass[10pt,english]{article}

\usepackage{charter,graphicx}
\usepackage[margin=1in]{geometry}

\usepackage{hyperref,tabularx}
\usepackage{fancyhdr}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{.4pt}% Default header rule
\renewcommand{\footrulewidth}{0pt}% No footer rule
\fancyhf{}% Clear header/footer
\fancypagestyle{plain}{
  \renewcommand{\headrulewidth}{0pt}% No header rule
  \renewcommand{\footrulewidth}{.4pt}% Default footer rule
  \fancyhf{}% Clear header/footer
}
\usepackage%[disable]	% uncomment to hide all the notes
	{todonotes}
\newcommand{\noteRev}[1]{\todo[color=blue!10, inline]{#1}}

\AtBeginDocument{\thispagestyle{plain}}

\setlength{\parindent}{0pt}
\setlength{\parskip}{.5\baselineskip plus 1pt minus 1pt}
\usepackage[utf8]{inputenc}%
\usepackage{babel}%
\usepackage{etoolbox}
\makeatletter
\patchcmd{\@zfancyhead}{\fancy@reset}{\f@nch@reset}{}{}
\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
\patchcmd{\@set@em@up}{\f@ncyolh}{\f@nch@olh}{}{}
\patchcmd{\@set@em@up}{\f@ncyorh}{\f@nch@orh}{}{}
\makeatother
%\usepackage{kpfonts}
%\usepackage{babel}
%\usepackage{csquotes}
\usepackage{url}
\usepackage{charter} % Use the Charter font for the document text
%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------
%Jean-Nicolas Jérémie\\
%Institute of Neuroscience of la Timone\\
%CNRS - Aix Marseille University\\
%27, boulevard Jean Moulin\\
%13005 Marseille, France
\newcommand{\LastName}{Grimaldi}%{Perrinet}%
\newcommand{\FirstName}{Antoine}%{Laurent U.}%
\newcommand{\Institute}{Institut de Neurosciences de la Timone, CNRS / Aix-Marseille Universit\'e}%
\newcommand{\Address}{27, Bd. Jean Moulin, 13385 Marseille Cedex 5, France}%
\newcommand{\Website}{\url{https://laurentperrinet.github.io/}}%
\newcommand{\Email}{\url{antoine.grimaldi@univ-amu.fr}}%
%----------------------------------------------------------------------------------------
%	YOUR DOCUMENT
%----------------------------------------------------------------------------------------
\usepackage[document]{ragged2e}
\begin{document}
%\centerline{
\includegraphics[width=.4\textwidth]{~/Documents/troislogos.png}
%}
\\
\vspace{.1\baselineskip}
\hrulefill
\vspace{.1\baselineskip}

\begin{flushright}
	\FirstName\  \LastName\  \\
	\Institute\\[6pt]
	\Address\\%[6pt]
	\Website \\
	  Email: \Email \\[6pt]
\end{flushright}
\justifying
\vspace{1\baselineskip}
%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING/CLOSING
%----------------------------------------------------------------------------------------
%À qui de droit,\\
Marseille, 
February 10, 2023%\today
\\[12pt] % Date

Dear editor,

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------
Please find enclosed the revised manuscript entitled ``Learning heterogeneous delays in a layer of spiking neurons for fast motion detection'' for your consideration as an article in \emph{Biological Cybernetics}. All the authors have been involved with the work, have approved the manuscript and agreed to its revision. The original submission ID is a7f45f8e-87d6-4a30-ac7f-9f0c94e2592a.

Our manuscript describes a novel method inspired by neuroscience to overcome some challenges faced in computer vision, notably when dealing with large amounts of data. By using heterogeneous delays on different synapses, this novel spiking neuron method is able to detect spiking motifs and we validate the method on synthetic event-based data. Results show that this method could provide a path for future spiking neural network algorithms using less energy for a similar performance as their analog counterparts. This work is the extension of a previous study on a simpler event-based dataset. Our contribution here is to offer a better formal description of the model, an application on a more ecological task and testing of the influence of various parameters of the validation dataset.

Given the substantial modifications we did to the original manuscript to reduce and clarify the different sections we do not provide a tracked changes version of the revision (such a document can be provided if needed). We provide a point by point response to the reviewers with references to the manuscript.

%\vspace{.2\baselineskip}
%----------------------------------------------------------------------------------------
Sincerely yours,%\vspace{.1cm}

\FirstName\ \LastName \\

\newpage

\textbf{REVIEWER REPORTS}

\textbf{Reviewer 1}

 1.There is a lack of an intuitive figure to show the HD-SNN model used in this paper.

 \noteRev{We provide such a figure in the new manuscript: Figure 3, P9}

 2.The MLR model used in the experimental part was presented in authors’ another work. Therefore, the changes on the MLR model are suggested to be highlighted and contrasted graphically.

 \noteRev{The main differences with the previous study are in the task to solve (section 2.1) and in the analysis of the results. The response of the model is tested on different sets of motion cloud stimuli with a control of the parameters of such stimuli generation (section 3.2). Regarding the MLR model, main differences with the previous ones are reported in the methods (P8, L269).}

 3.The title of section 2.2.1 is ’’a generative model for rater plots’’, but actually this section introduces basic terms and symbols. Section 2.2.2 is named ‘’detecting spiking motifs’’, but most of the content seems to introduce the generation process of raster plots, while only the last paragraph introduces the detection of motif. It is recommended that the content of these two sections be reorganized and that references to Figure 1(b)(c)(d) be added where appropriate during the description.

 \noteRev{We removed these sections to have a more concise explanation of the model.}

 4.The adjustment of weights and delays needs more detailed elaboration, preferably given in mathematical formulas.

 \noteRev{We improved the mathematical formulation of the model in sections 2.2.1 and 2.2.2 and hope that it clarifies the general framework we use. Regarding training, it is described more clearly in P8, L295. Demonstration of the mechanism and the analogy with Hebbian learning was done in~\cite{grimaldi_robust_2022} and was not reproduced here. We hope this new version offers a clearer methodology.}

 5.In Figure 4, it is claimed that the kernels are very similar for the ON polarities, and different kernels are selective to the different motion directions. However, in terms of these two observations, I think they are not very significant, so I suggest a more intuitive way to reflect them, such as quantifying the similarity between the ON kernels, and the angle between the kernels and the vector of the standard motion directions.

 \noteRev{We gave a better explanation of the symmetry assumption in the methods (P8, L293) an how this new model uses this invariance in its computation like observed in complex cells. Indded, we don't want the polarity to have an impact on the classification and then this observation does not apply to the revised manuscript as it is a prior and not an emergent property.}

 6.There are few experiments and lack of comparative experiments. Since the model has an improvement based on MLR model, it is suggested to increase the comparison with MLR model and HOTS model on other datasets, such as N-MNIST and other neuromorphic datasets, so as to show the advantages of the proposed model in computational consumption.

 \noteRev{The reviewer is making a good point but regarding the deadline for the revision and some objectives we focused on to improve the manuscript we could not apply the model on DVS recordings.}

 7.Some sections are verbose, such as the discussion.

 \noteRev{We significantly changed the introduction (section 1), the methods (section 2) and the discussion (section 4) to address this comment. We hope to provide a more concise and clear explanation of our study.}

 8.There are some minor errors in the text, and it is recommended to check the full text: 
 1)In the first paragraph on page 8, "a presynaptic address $b_s$" should be "a postsynaptic address $b_s$"; 
 2)Whether $W_b$ in the first formula should be $K_b$? And d in it lacks description. 
 3)"PGs" should be given in its full name when it first appears. 
 4)There is no subgraph label of (a), (b) and (c) in Fig. 3, but the expression of Fig. 3-(a)/(b)/(c) appears many times. 
 5)The content of paragraph 1 on page 13 is repeated.

 \noteRev{We corrected these minor errors and thank the reviewer for reporting them.}

 \noteRev{Overall, we thank Reviewer 1 for its comments that allowed us to focus on reshaping the manuscript and give clearer and shorter explanation of the work we submit in your journal. Because we did major modifications in the manuscript, we do not provide a tracked changes version of the revision but we can easily provide one if needed. }

\textbf{Reviewer 2}

 Key results: Please summarize what you consider to be the outstanding features of the work.
 Using a single layer of spiking neurons with heterogeneous delays to learn spatio-temporal  spiking motifs

 Validity: Does the manuscript have flaws which should prohibit its publication? If so, please provide details.

 The work presented here is very similar to the author's other work,  "Learning hetero-synaptic delays for motion detection in a single layer of spiking neurons" 10.1109/ICIP46576.2022.9897394. This must be addressed and cited with an explanation of how this work presented is an extension of this previous publication. Abstract, Fig 6 and the main result is presented in the previous work. Arguably some of the figures from the previous publication would be nice to see in this publication as it helps with the visualisation of the problem. One suggestion for the extension to make the work unique would be to extend with some event-camera data, as it is mentioned multiple times within the publication. 
 There are also a number of citations that could have been added that did appear in the other publication, while also including other motion detection/ segmentation or previous work on event-based optical flow, which is a very similar problem.

 Originality and significance: If the conclusions are not original, please provide relevant references.

 See above...

\noteRev{We thank Reviewer 2 for its comments and indeed this previous work was not mentioned in the original manuscript. As mentioned in the cover letter and in the previous responses, we add this citation and describe the main differences with the submitted study: ecological task (section 2.1), more complete formulation of the model (section 2.2) and list of the priors used in this particular model (P8, L269). We also cite studies that use similar methods (P7, L243). }

\end{document}

### 2023-02-09 cover_letter/coverletter.tex
cover
Update on Overleaf.
response to reviewer

@@
 %\centerline{
-\includegraphics[width=.4\textwidth]{~/metagit/blog/perrinet_curriculum-vitae_tex/troislogos.png}
+\includegraphics[width=.4\textwidth]{~/Documents/troislogos.png}
 %}
@@
 %À qui de droit,\\
-Marseille, \today\\[12pt] % Date
+Marseille, 
+February 10, 2023%\today
+\\[12pt] % Date
 
@@
 %----------------------------------------------------------------------------------------
-Please find enclosed a manuscript entitled ``Learning heterogeneous delays in a layer of spiking neurons for fast motion detection'' for your consideration as an article in \emph{Biological Cybernetics}.
+Please find enclosed manuscript entitled ``Learning heterogeneous delays in a layer of spiking neurons for fast motion detection'' for your consideration as an article in \emph{Biological Cybernetics}.
 All the authors have been involved with the work, have approved the manuscript and agreed to its submission.

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex

--- 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
@@
 A specific mechanism for this is called saccades, which are rapid eye movements that reposition the center of vision.
-In humans, saccades are very frequent (on average $2 per second~\citep{dandekar_neural_2012}).
+In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}).
 They are generated very quickly (about $80~\ms$) and have a wide range of speeds.

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
Methods : 0

@@
 %%%-----------------------------------------------------------------
+The goal is to test the model in a motion detection task using a realistic event-based data stream.
+This type of signal is typically recorded by a Dynamic Vision Sensor (DVS) and is inspired by the signal sent by retinal ganglion cells through the optic nerve.
+Events are binary and allow to capture the dynamics of the natural scene and perform various cognitive tasks quickly and effectively.
+Here, we will first describe the definition of the task and then the model we use to infer motion.
 %%%-----------------------------------------------------------------
@@
 %%%-----------------------------------------------------------------
-The goal is to test the model in a motion detection task using a realistic data stream.
-This type of signal is typically recorded by a Dynamic Vision Sensor (DVS).
-It is a good representation of the information and is inspired by the information sent by retinal ganglion cells through the optic nerve.
-Binary events allow the visual system to capture the dynamics of the natural scene and perform various cognitive tasks quickly and effectively.
-To train and test our model, we need to define a data set for which we control the direction and speed of motion.
+To train and test our model, we need to define a dataset for which we control the direction and speed of motion.
 Let us now define a procedure for animating a natural scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
Methods : A

@@
 The resulting instances yield trajectories that are qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(left)).
+
 %%%-----------------------------------------------------------------
@@
     \caption{
-        {\bf - Motion Detection Task.} {\bf (left)}~We use large natural images ($512\times512$) in which an aperture ($128\times128$) extracts a cropped image around the view axis.
+        {\bf - Motion Detection Task.} {\bf (Left)}~We use large natural images ($512\times512$) in which an aperture ($128\times128$) extracts a cropped image around the view axis.
 To mimic the effect of a saccadic eye movement, the view axis moves according to a stepwise random walk.
 We show an example path of the trajectory for $200$ time steps (green line).
-{\bf (right)}~Snapshot of the synthetic event stream at different time steps.
+{\bf (Right)}~Snapshot of the synthetic event stream at different time steps.
 The dynamics of the sub-image as a function of time produces a naturalistic movie.
 This can be transformed into an event-based representation.
-Mimicking the retina, this representation encodes proportional and pixel-wise increases or decreases in luminance in the image, i.e.
+Mimicking the retina, this representation encodes pixel-per-pixel proportional increases or decreases in luminance in the image, i.e.
 ON (red) and OFF (blue) events.
-In the lower left corner of the images, the translation vector is shown in red.
-It is one of the possible classes of motion.
+In the lower left corner of the images, the translation vector is shown in red as one of the possible classes of motion.
 Note the change in the direction of motion between the second and third image.}
@@
 We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 250~\ms$.
-To avoid boundary effects, we will randomly position a plot of the full trajectory in image space so that the subimage is translated using the position given by the trajectory at each time step.
+To avoid boundary effects, we will randomly position the full trajectory in image space such that the subimage is translated using the position given by the trajectory at each time step without touching the borders.
 The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a subpixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Once these eye trajectories are generated, we can apply them to a visual scene.
-For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images~\citep{olshausen_emergence_1996}.
-Note that these have been pre-processed to equalize (i.e., whiten) the energy in each frequency band.
+% For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images~\citep{olshausen_emergence_1996}.
+For this purpose, we selected a database of 600 natural images that were previously used to study the statistics of natural images~\citep{perrinet_edge_2015}.
+Note that these have been pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band.
 This process is known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $512 \times 512$ in size, and we will extract sub-images of size $128 \times 128$ positioned around the center of gaze at each time step.
+These images are $256 \times 256$ in size, and we will extract sub-images of size $128 \times 128$ positioned around the center of gaze at each time step.
 We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 250~\ms$.
@@
 The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/pyTERtorch}.
-\note{à rendre publique}
+%\note{à rendre publique}
 %

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
Methods : 2

--- 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
@@
 
-\subsubsection{Applied on this detection task}
+\subsubsection{Application to the motion detection task}
 
@@
 \end{equation}
-
 where $\presynaddrspace^{\synapse^\postsynaddr}$ is the address space of the receptive field of neuron $\postsynaddr$ and $N_{\synapse^\postsynaddr}$ is the number of synapses belonging to neuron $\postsynaddr$.

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
Methods : 2

@@
 \newcommand{\Ny}{N_\text{Y}}
-\newcommand{\Ntime}{N_\text{t}}
+\newcommand{\Ntime}{N_\text{T}}
 \newcommand{\kernel}{K} % convolution kernel
@@
 %
-\subsection{Description of the model}
-%%%-----------------------------------------------------------------
-\subsubsection{The heterogeneous synaptic delay model}
+% \subsection{Description of the model}
+%%%-----------------------------------------------------------------
+\subsection{The Heterogeneous Delays SNN (HD-SNN) model}
 
@@
 
-Let's formalize a layer of spiking neurons with heterogeneous delays (HD-SNN).
+Let's formalize the heterogeneous delays spiking neural network (HD-SNN).
 Each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferent from $\presynaddrspace$.
@@
 
-\subsubsection{Application to the motion detection task}
+\subsection{Application of HD-SNN to the motion detection task}
 
@@
 Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{N\times T}$, where $N$ is the number of neurons in $\presynaddrspace$ and $T$ is the number of time bins.
-In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ of size $(N_{tb}, N_x, N_y)$, where $N_{tb}$ is the number of time bins, $N_x$ and $N_y$ are the number of pixels in both spatial dimensions.
+In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ of size $(\Ntime, \Nx, \Ny)$, where $\Ntime$ is the number of time bins, $\Nx$ and $\Ny$ are the number of pixels in both spatial dimensions.
 For each neuron $\postsynaddr$, $\kernel^\postsynaddr$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel^\postsynaddr(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$.

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
Methods : 3

@@
 Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{N\times T}$, where $N$ is the number of neurons in $\presynaddrspace$ and $T$ is the number of time bins.
-In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ of size $(\Ntime, \Nx, \Ny)$, where $\Ntime$ is the number of time bins, $\Nx$ and $\Ny$ are the number of pixels in both spatial dimensions.
+In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented as a kernel by the dense matrix $\kernel^\postsynaddr$ of size $(\Ntime, \Nx, \Ny)$, where $\Ntime$ is the number of delays and $\Nx$ and $\Ny$ are the number of pixels in both spatial dimensions.
 For each neuron $\postsynaddr$, $\kernel^\postsynaddr$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel^\postsynaddr(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$.
@@
 Supervision was implemented using the input binary events as defined above and the motion direction labels as the desired output.
-The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions.
+The labels were defined at each time point as a one-hot encoding of the current motion in the channel corresponding to the current motion for all positions.
 As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner.

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
Methods : 2.3

@@
 \newcommand{\Ntime}{N_\text{T}}
-\newcommand{\kernel}{K} % convolution kernel
+\newcommand{\kernel}{\mathcal{K}} % convolution kernel
 %\newcommand{\kernelind}{\mathbf{k}} % indice of the kernel
 \newcommand{\kernelind}{k} % indice of the kernel
-\newcommand{\Kx}{K_\text{x}}
-\newcommand{\Ky}{K_\text{y}}
-\newcommand{\Ktime}{K_\text{t}}
+\newcommand{\Kx}{K_\text{X}}
+\newcommand{\Ky}{K_\text{Y}}
+\newcommand{\Ktime}{K_\text{T}}
 \newcommand{\classiflayer}{\mathbf{C}}
@@
 %
-\subsection{Task definition: fast motion detection}\label{sec:task}
+\subsection{Task definition: motion detection in a synthetic naturalistic event stream}\label{sec:task}
 %%%-----------------------------------------------------------------
@@
 This process is known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $256 \times 256$ in size, and we will extract sub-images of size $128 \times 128$ positioned around the center of gaze at each time step.
-We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 250~\ms$.
+These images are $256 \times 256$ in size, and we will extract sub-images of size $64 \times 64$ positioned around the center of gaze at each time step.
+We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 128~\ms$.
 To avoid boundary effects, we will randomly position the full trajectory in image space such that the subimage is translated using the position given by the trajectory at each time step without touching the borders.
-The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a subpixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
+The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
 Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity to correspond to a movement of one pixel per frame (i.e., per time bin).
@@
 This defines the full connectivity of the HD-SNN model.
+
 Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]} $, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]}$.
@@
 
-\note{I can make a small figure with event streams of different colors for the input one, the multiplexed one etc and the response in the bottom}
+% \note{I can make a small figure with event streams of different colors for the input one, the multiplexed one etc and the response in the bottom}
 
@@
 From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation.
-Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{N\times T}$, where $N$ is the number of neurons in $\presynaddrspace$ and $T$ is the number of time bins.
-In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented as a kernel by the dense matrix $\kernel^\postsynaddr$ of size $(\Ntime, \Nx, \Ny)$, where $\Ntime$ is the number of delays and $\Nx$ and $\Ny$ are the number of pixels in both spatial dimensions.
-For each neuron $\postsynaddr$, $\kernel^\postsynaddr$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel^\postsynaddr(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$.
+Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{\Nx \times \Ny \times \Ntime}$, defined in space coordinates $x$ and $y$ and time $t$.
+To benefit from invariance to location observed in CNN and to simplify the implementation of the model, one assumes that kernels should be similar across different positions and define a convolution operator.
+% The characteristics of the kernels are shared among one class of neurons such that $\forall i,j \in [1, N_\class]^2, \; \kernel^i_\class = \kernel^j_\class$.
+Here, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented as a kernel for each class $\class$ of the supervision task by the dense matrix $\kernel_\class$ of size $(\Kx, \Ky, \Ktime)$, where $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions and  $\Ktime$ is the number of delays.
+To keep an analogy with the HD-SNN model, for each neuron $\postsynaddr$ of position $(x, y)$ and class $\class$, $\kernel_\class$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$.
 Therefore, if we make the approximation that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train as input can be formalized by a 3D convolution operation.
 The longest synaptic delay defines the depth of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented.
-Using this dense representation, the integration of the spike train as input by the synapses of neuron $\postsynaddr$ is:
-
+Using this dense representation, the processing of $A$ by the model can be written as:
 \begin{equation}\label{eq:kernel_b}
-\forall \presynaddr \in \presynaddrspace^{\synapse^\postsynaddr}, \forall \timev \in T, \; \;
-\mathcal{C}^\postsynaddr(\presynaddr,\timev)
-= \sum_{\ranksyn  \in [1, N_{\synapse^\postsynaddr}]} \kernel^\postsynaddr(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
+\forall x, y, \forall \timev \in T, \; \;
+\mathcal{C}_\class(x, y, \timev)
+= \sum_{\delta_x, \delta_y, \synapticdelay} \kernel_\class(\delta_x, \delta_y, \synapticdelay) \cdot A(x - \delta_x, y - \delta_y, \timev-\synapticdelay)
 \end{equation}
-where $\presynaddrspace^{\synapse^\postsynaddr}$ is the address space of the receptive field of neuron $\postsynaddr$ and $N_{\synapse^\postsynaddr}$ is the number of synapses belonging to neuron $\postsynaddr$.
-To benefit from invariance to location observed in CNN and to simplify the implementation of the model, one assumes that kernels should be similar across different positions and define a convolution operator.
-The characteristics of the kernels are shared among one class of neurons such that $\forall i,j \in [1, N_\class]^2, \; \kernel^i_\class = \kernel^j_\class$.
-The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$:
-
-\begin{equation}\label{eq:conv}
-\forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
-\mathcal{C}_\class(\presynaddr,\timev)
-= \sum_{\ranksyn  \in [1, N_{\synapse^\postsynaddr}]} \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
-\end{equation}
-
+% The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$:
+% \begin{equation}\label{eq:conv}
+% \forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
+% \mathcal{C}_\class(\presynaddr,\timev)
+% = \sum_{\ranksyn  \in [1, N_{\synapse^\postsynaddr}]} \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
+% \end{equation}
 This shows that $\mathcal{C}_\class$ is a spatiotemporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration).
@@
 This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
-%
+
 The same type of spatiotemporal filtering is used as a preprocessing stage of a pattern recognition algorithm~\cite{ghosh_spatiotemporal_2019}.
 And in \cite{sekikawa_constant_2018}, an efficient 3D convolutional algorithm succeeds a motion estimation task.
-By assuming a locally constant velocity the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.
-Here, we keep the analogy with spiking neurons and we try to use as few heuristics as possible to observe the emergence of spiking motifs in the spatiotemporal kernels.
+By assuming a locally constant velocity, the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.
+Here, we keep the analogy with spiking neurons, and we try to use as few heuristics as possible to observe the emergence of spiking motifs in the spatiotemporal kernels.
 If so, this precise spatiotemporal patterns prove to be of interest for neural computations and one can assume that biological neurons make use of this information as well.
-
-\note{plus de biblio sur la motion detection/segmentation ou optique flow avec les events? (reviewer 2)}
-
+% \note{plus de biblio sur la motion detection/segmentation ou optique flow avec les events? (reviewer 2)}
 The 3D convolution represents the linear integration of the spike train as input.
 The activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
-It transforms this list of weights into a probability with the following formula:
-
+In our MLR model, a probability value, for each class (i.e.
+each motion direction), is predicted for each event at position $x, y$ and at time $\timev$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the HD-SNN model.
+Using the kernels, it transforms the input raster plot into a probability with the following formula:
 \begin{equation}\label{eq:mlr}
-\forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \forall \class \in [1, N_\class], \; \;
-Pr(k=\class \; \vert \; \presynaddr, \timev) =
-\frac 1 Z
-{\exp  (\mathcal{C}_\class(\presynaddr, \timev) +\bias_\class) }
+    \forall x, y, \forall \timev \in T, \forall \class \in [1, N_\class], \; \;
+Pr(k=\class \; \vert \;  x, y, \timev) =
+\frac {\exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class) }{Z(x, y, \timev)}
 \end{equation} 
-
-where $\bias_\class$ is the bias linked to class $\class$ and $Z$ is the partition function. 
-In our MLR model, a probability value, for each class (i.e.
-each motion direction), is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
-The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the HD-SNN model.
+where $\bias_\class$ is the bias linked to class $\class$ and $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function. 
 In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
-The spiking output of the model corresponds to an event with the highest probability class, keeping the same timing as the input event.
+The spiking output of the model corresponds to an event with the highest probability class.
 
@@
 
-\note{describe the heuristics inspired by neuroscience}
-
-\note{describe the parameters of the model}
+% \note{describe the heuristics inspired by neuroscience}
+
+% \note{describe the parameters of the model}
 
@@
 \end{figure}
+
+\subsection{Supervised learning of the motion detection task}
 %

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Methods : 2.3

@@
 In the next subsection we describe an implementation of our model adapted for the motion detection task.
-
+%
 % \note{I can make a small figure with event streams of different colors for the input one, the multiplexed one etc and the response in the bottom}
-
+%
 \subsection{Application of HD-SNN to the motion detection task}
-
+%
 From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation.
-Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{\Nx \times \Ny \times \Ntime}$, defined in space coordinates $x$ and $y$ and time $t$.
+Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{\Nx \times \Ny \times \Ntime \times \Npol}$, defined in space coordinates $x$ and $y$, time $t$ and polarities $p$.
 To benefit from invariance to location observed in CNN and to simplify the implementation of the model, one assumes that kernels should be similar across different positions and define a convolution operator.
 % The characteristics of the kernels are shared among one class of neurons such that $\forall i,j \in [1, N_\class]^2, \; \kernel^i_\class = \kernel^j_\class$.
-Here, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented as a kernel for each class $\class$ of the supervision task by the dense matrix $\kernel_\class$ of size $(\Kx, \Ky, \Ktime)$, where $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions and  $\Ktime$ is the number of delays.
+Here, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented as a kernel for each class $\class$ of the supervision task by the dense matrix $\kernel_\class$ of size $(\Npol, \Kx, \Ky, \Ktime)$, where $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions and  $\Ktime$ is the number of delays.
 To keep an analogy with the HD-SNN model, for each neuron $\postsynaddr$ of position $(x, y)$ and class $\class$, $\kernel_\class$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$.
@@
 \begin{equation}\label{eq:kernel_b}
-\forall x, y, \forall \timev \in T, \; \;
+\forall x, y, \timev, \; \;
 \mathcal{C}_\class(x, y, \timev)
-= \sum_{\delta_x, \delta_y, \synapticdelay} \kernel_\class(\delta_x, \delta_y, \synapticdelay) \cdot A(x - \delta_x, y - \delta_y, \timev-\synapticdelay)
+= \sum_{p, \delta_x, \delta_y, \synapticdelay} \kernel_\class(p, \delta_x, \delta_y, \synapticdelay) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay)
 \end{equation}
@@
 In our MLR model, a probability value, for each class (i.e.
-each motion direction), is predicted for each event at position $x, y$ and at time $\timev$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+each motion direction), is predicted for each event at position $x, y$ and at time $\timev$ as a softmax function of the linear combination of the list of events.
+% on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
 The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the HD-SNN model.
@@
 \begin{equation}\label{eq:mlr}
-    \forall x, y, \forall \timev \in T, \forall \class \in [1, N_\class], \; \;
+    \forall x, y, \timev, \forall \class \in [1, N_\class], \; \;
 Pr(k=\class \; \vert \;  x, y, \timev) =
@@
 \end{equation} 
-where $\bias_\class$ is the bias linked to class $\class$ and $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function. 
+where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function and  $\bias_\class$ is the bias linked to class $\class$. 
 In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
methods : figure izhikevich

@@
 \subsection{The Heterogeneous Delays SNN (HD-SNN) model}
+%----------------------------%
+%
+\begin{figure}%[t!]
+    \centering
+    \includegraphics[width=0.980\linewidth]{figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
+      \caption{\textbf{Core mechanism of the HD-SNN model.} \textit{(Left)}~In this example inspired by the polychronization model~\citep{izhikevich_polychronization_2006}, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
+\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
+      %at these different delays, and these may not be sufficient to generate a spike in either neuron.
+      \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from post-synaptic neuron \textit{e}.
+Figure reproduced from~\cite{grimaldi_precise_2023} under a CC-BY licence.
+       }
+    \label{fig:izhikevich}
+  \end{figure}
+  % 
+  %----------------------------%
 
@@
 Let's formalize the heterogeneous delays spiking neural network (HD-SNN).
-Each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferent from $\presynaddrspace$.
+Each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferents from $\presynaddrspace$.
 In biology, a single cortical neuron has generally several thousands of synapses.
@@
 
-Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]} $, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]}$.
+Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$.
 As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field.
-It results in a weighted event stream:
-$\event_\postsynaddr = \{(\presynaddr_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}$.
-Unlike the binary stream of events from the signal as input, the postsynaptic neurons receives a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank = \synapticweight_\ranksyn$ when $\presynaddr_\arank = \presynaddr_\ranksyn$.
-This new stream of events is by nature ordered in time as events reach the soma of post-synaptic neurons.
+It results in a weighted event stream for each  postsynaptic neuron $\postsynaddr$:
+\begin{equation}\label{eq:stream_b}
+\event_\postsynaddr = \{(\presynaddr_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}
+\end{equation}
+Unlike the binary stream of events from the signal as input, the postsynaptic neurons receive a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank = \synapticweight_\ranksyn$ when $\presynaddr_\arank = \presynaddr_\ranksyn$.
+This new stream of events is ordered in time in biology as events reach the soma of post-synaptic neurons but should be re-orderes in simulations.
 In particular, when post-synaptic neurons are activated on their soma by this spatio-temporal motif, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
methods

@@
 Events are typically ordered by their time of occurrence.
-Each event has a time of occurrence $\timev_\arank$ and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
+Each event has a time of occurrence $\timev_\arank$ and an associated address $\presynaddr_\arank$.
 This defines an address space $\presynaddrspace$ which consists of the set of possible addresses.
 In a neurobiological recording, this can be the identified set of neurons.
-For event-based cameras, it is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras).
+For event-based cameras, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$ where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras) and $(\Nx, \Ny)$ is the size of the sensor in pixels.
+As such, an address $\presynaddr_\arank$ is typically in the form $(\polev_\arank, x_\arank, y_\arank)$ for event-based cameras.
 
@@
 Unlike the binary stream of events from the signal as input, the postsynaptic neurons receive a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank = \synapticweight_\ranksyn$ when $\presynaddr_\arank = \presynaddr_\ranksyn$.
-This new stream of events is ordered in time in biology as events reach the soma of post-synaptic neurons but should be re-orderes in simulations.
-In particular, when post-synaptic neurons are activated on their soma by this spatio-temporal motif, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.
+This new stream of events is ordered in time in biology as events reach the soma of post-synaptic neurons but should be reordered in simulations.
+Crucially, when post-synaptic neurons are activated on their soma by a specific spatio-temporal motif which is imprinted in the set of synapses, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner (see Figure~\ref{fig:izhikevich}).
 This model is defined in this subsection in all generality and the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions.
@@
 From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation.
-Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{\Nx \times \Ny \times \Ntime \times \Npol}$, defined in space coordinates $x$ and $y$, time $t$ and polarities $p$.
+Indeed, by discretization of time, we may transform any event-based input as a boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$, defined for all polarities $p$, times $t$ and space coordinates $x$ and $y$.
+The values are by definition equal to zero except at the occurences of events: $\forall \numevent \in \mathbb{N}$, $A()=1$
+
+$\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events.
+Events are typically ordered by their time of occurrence.
+Each event has a time of occurrence $\timev_\arank$ and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
+
 To benefit from invariance to location observed in CNN and to simplify the implementation of the model, one assumes that kernels should be similar across different positions and define a convolution operator.
 % The characteristics of the kernels are shared among one class of neurons such that $\forall i,j \in [1, N_\class]^2, \; \kernel^i_\class = \kernel^j_\class$.
-Here, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented as a kernel for each class $\class$ of the supervision task by the dense matrix $\kernel_\class$ of size $(\Npol, \Kx, \Ky, \Ktime)$, where $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions and  $\Ktime$ is the number of delays.
+% Here, we will consider that heterogeneous delays are limited in range
+In particular, the whole synaptic set can be represented as a one kernel for each class $\class$ of the supervision task as the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
 To keep an analogy with the HD-SNN model, for each neuron $\postsynaddr$ of position $(x, y)$ and class $\class$, $\kernel_\class$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$.

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
methods : 2.2

@@
 We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
-We apply this model to solve the specific computer vision problem of motion detection.
 We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of event-based cameras.
@@
 In the lower left corner of the images, the translation vector is shown in red as one of the possible classes of motion.
-Note the change in the direction of motion between the second and third image.}
+Note the change in the direction of motion between the second and third image and that due to the aperture problems, contours parallel to the motion emit less spikes.}
     \label{fig:motion_task}
@@
 \begin{equation}\label{eq:stream_b}
-\event_\postsynaddr = \{(\presynaddr_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}
+\event_\postsynaddr = \{(\presynaddr_\arank, \synapticweight_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}
 \end{equation}
-Unlike the binary stream of events from the signal as input, the postsynaptic neurons receive a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank = \synapticweight_\ranksyn$ when $\presynaddr_\arank = \presynaddr_\ranksyn$.
+% Unlike the binary stream of events from the signal as input, the postsynaptic neurons receive a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank = \synapticweight_\ranksyn$ when $\presynaddr_\arank = \presynaddr_\ranksyn$.
 This new stream of events is ordered in time in biology as events reach the soma of post-synaptic neurons but should be reordered in simulations.
-Crucially, when post-synaptic neurons are activated on their soma by a specific spatio-temporal motif which is imprinted in the set of synapses, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner (see Figure~\ref{fig:izhikevich}).
-This model is defined in this subsection in all generality and the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions.
-In the next subsection we describe an implementation of our model adapted for the motion detection task.
+Crucially, when post-synaptic neurons are activated on their soma by a specific spatio-temporal motif which is imprinted in the set of synapses, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner while the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions (see Figure~\ref{fig:izhikevich}).
+We defined the HD-SNN model in this subsection in all generality and in the next subsection we describe an implementation of our model adapted for the motion detection task.
 %

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
methods : 2.2

@@
 % NOTATIONS
-% Pixel world
+\newcommand{\ms}{\si{\milli\second}}%
+\newcommand{\us}{\si{\micro\second}}%
 \newcommand{\presynaddr}{a} % pre address
@@
 \newcommand{\layer}{\mathbf{L}} % layer
-\newcommand{\ms}{\si{\milli\second}}%
-\newcommand{\us}{\si{\micro\second}}%
-\newcommand{\timecontext}{T} % time context (cf HOTS) matrice gathering last event times
+% \newcommand{\timecontext}{T} % time context (cf HOTS) matrice gathering last event times
 \newcommand{\current}{I} % post synaptic current
@@
 \newcommand{\gain}{\gamma} % homeostatic gain
-\newcommand{\simil}{\beta} % similarity value
+% \newcommand{\simil}{\beta} % similarity value
 \newcommand{\Nclass}{N_\text{class}} % number of classes for MLR:
@@
 Indeed, by discretization of time, we may transform any event-based input as a boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$, defined for all polarities $p$, times $t$ and space coordinates $x$ and $y$.
-The values are by definition equal to zero except at the occurences of events: $\forall \numevent \in \mathbb{N}$, $A()=1$
-
-$\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events.
-Events are typically ordered by their time of occurrence.
-Each event has a time of occurrence $\timev_\arank$ and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
-
-To benefit from invariance to location observed in CNN and to simplify the implementation of the model, one assumes that kernels should be similar across different positions and define a convolution operator.
+The values are by definition equal to zero except at the occurrences of events: $\forall \numevent \in \mathbb{N}$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
+First, one can note that by using a discretization, the computational brick used in equation~\eqref{eq:kernel_b} corresponds to a temporal convolution which transforms the input $A$ using one kernel per postsynaptic neuron~\citep{grimaldi_learning_2022}.
+To benefit from invariance to location observed in images and exploited in Convolutional Neural Networks, one may further assume that synaptic motifs should be similar across different positions such that we may define a spatio-temporal convolution operator.
 % The characteristics of the kernels are shared among one class of neurons such that $\forall i,j \in [1, N_\class]^2, \; \kernel^i_\class = \kernel^j_\class$.
@@
 To keep an analogy with the HD-SNN model, for each neuron $\postsynaddr$ of position $(x, y)$ and class $\class$, $\kernel_\class$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$.
-Therefore, if we make the approximation that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train as input can be formalized by a 3D convolution operation.
+%Therefore, if we make the approximation that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train as input can be formalized by a 3D convolution operation.
 The longest synaptic delay defines the depth of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented.
@@
 \end{equation}
-% The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$:
+% The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$: equation~\eqref{eq:stream_b}
 % \begin{equation}\label{eq:conv}
@@
 % \end{equation}
-This shows that $\mathcal{C}_\class$ is a spatiotemporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration).
+This shows that $\mathcal{C}_\class$ is a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration).
 Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
@@
 
-The same type of spatiotemporal filtering is used as a preprocessing stage of a pattern recognition algorithm~\cite{ghosh_spatiotemporal_2019}.
+The same type of spatio-temporal filtering is used as a preprocessing stage of a pattern recognition algorithm~\cite{ghosh_spatio-temporal_2019}.
 And in \cite{sekikawa_constant_2018}, an efficient 3D convolutional algorithm succeeds a motion estimation task.
 By assuming a locally constant velocity, the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.
-Here, we keep the analogy with spiking neurons, and we try to use as few heuristics as possible to observe the emergence of spiking motifs in the spatiotemporal kernels.
-If so, this precise spatiotemporal patterns prove to be of interest for neural computations and one can assume that biological neurons make use of this information as well.
+Here, we keep the analogy with spiking neurons, and we try to use as few heuristics as possible to observe the emergence of spiking motifs in the spatio-temporal kernels.
+If so, this precise spatio-temporal patterns prove to be of interest for neural computations and one can assume that biological neurons make use of this information as well.
 % \note{plus de biblio sur la motion detection/segmentation ou optique flow avec les events? (reviewer 2)}
@@
 We would like to highlight a few innovations in the contributions that are presented in this paper.
-First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we observed to be less efficient, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs.
+First, whereas~\citep{ghosh_spatio-temporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we observed to be less efficient, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs.
 Moreover, in comparison to HOTS~\citep{lagorce_hots_2017}, the weights are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection of each spatio-temporal spike motif.

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Since then, numerous studies have demonstrated the emergence of synchronicity of activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, it's use for efficient coding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
-All these findings and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
+All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
 In neural models, the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows for the efficient exploitation or detection of these spatio-temporal motifs embedded in the spike train.

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 While the first possibility favors the theory of rate coding, the second one emphasizes the function of temporal precision in the neural code.
-Since then, numerous studies have demonstrated the emergence of synchronicity of activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, it's use for efficient coding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
+Since then, numerous studies have demonstrated the emergence of synchronicity of activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
 All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
@@
 In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, but precise, relative spiking delays, i.e., the time between the arrival of an afferent spike at a given synapse and the contribution of the associated postsynaptic potential to the neuron's soma.
-Due to the variety of configurations and the possible coexistence of multiple superimposed motifs, this representation has a much higher information capacity than a firing rate-based approach to neural coding.
+Due to the variety of configurations and the possible coexistence of multiple superimposed motifs, representation with polychronous groups has a much higher information capacity than a firing rate-based approach to neural coding.
 
@@
 Let us now define a procedure for animating a natural scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
-First, let's define a trajectory inspired by the biological movements of the eyes.
+First, we draw a trajectory inspired by the biological movements of the eyes.
 Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the visual field.
@@
 We show an example path of the trajectory for $200$ time steps (green line).
-{\bf (Right)}~Snapshot of the synthetic event stream at different time steps.
+{\bf (Right)}~Snapshots of the synthetic event stream at different time steps.
 The dynamics of the sub-image as a function of time produces a naturalistic movie.
@@
 ON (red) and OFF (blue) events.
-In the lower left corner of the images, the translation vector is shown in red as one of the possible classes of motion.
+In the lower left corner of the snapshots, the translation vector is shown in red as one of the possible classes of motion.
 Note the change in the direction of motion between the second and third image and that due to the aperture problems, contours parallel to the motion emit less spikes.}
@@
 To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames.
-If, on a specific pixel at that specific timestamp, the absolute value of this gradient exceeds a threshold, an event is generated.
+For a specific pixel and timestamp, an event is generated if the absolute value of this gradient exceeds a threshold.
 The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive.

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
 This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
-
-The same type of spatio-temporal filtering is used as a preprocessing stage of a pattern recognition algorithm~\cite{ghosh_spatio-temporal_2019}.
-And in \cite{sekikawa_constant_2018}, an efficient 3D convolutional algorithm succeeds a motion estimation task.
-By assuming a locally constant velocity, the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.
+The same type of spatio-temporal filtering is used as a preprocessing stage of a pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
+Also,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm which implements a motion estimation task.
+By assuming a locally constant velocity, the authors assume the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.
+
 Here, we keep the analogy with spiking neurons, and we try to use as few heuristics as possible to observe the emergence of spiking motifs in the spatio-temporal kernels.
@@
 We would like to highlight a few innovations in the contributions that are presented in this paper.
-First, whereas~\citep{ghosh_spatio-temporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we observed to be less efficient, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs.
+First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we observed to be less efficient, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs.
 Moreover, in comparison to HOTS~\citep{lagorce_hots_2017}, the weights are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection of each spatio-temporal spike motif.

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
2.1

@@
 In order to maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.This approach first defines a finite set of possible 2D movements in polar coordinates.
-Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $8$ linearly spaced movement directions and 3 geometrically spaced speeds.
+Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $8$ linearly spaced movement directions \note{and 3 geometrically spaced speeds}.
 Next, we define a gaze trajectory as segments whose duration is drawn from a Poisson distribution with a mean block length of $20~\ms$, similar to a Lévy flight~\citep[p.
@@
 
-To transform each movie into events, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames.
+To transform each movie into events as the ones recorded by a DVS, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames.
 For a specific pixel and timestamp, an event is generated if the absolute value of this gradient exceeds a threshold.
@@
 This signed threshold value is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is as a consequence similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion_task}-(Right)).
+When applied to the whole movie, the event stream is similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion_task}-(Right)).
 The goal here is to infer the correct motion solely by observing these events.
@@
 
-% \note{describe the heuristics inspired by neuroscience}
+\note{describe the heuristics inspired by neuroscience, remaining complex cells}
 

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
Update Grimaldi-etal-BiolCybernetics.tex
methods

@@
 This defines an address space $\presynaddrspace$ which consists of the set of possible addresses.
-In a neurobiological recording, this can be the identified set of neurons.
+In a neurobiological recording, this can be the identified set of presynaptic neurons.
 For event-based cameras, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$ where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras) and $(\Nx, \Ny)$ is the size of the sensor in pixels.
@@
 
-Let's formalize the heterogeneous delays spiking neural network (HD-SNN).
-Each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferents from $\presynaddrspace$.
+In the HD-SNN, each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferents from $\presynaddrspace$.
 In biology, a single cortical neuron has generally several thousands of synapses.
@@
 It results in a weighted event stream for each  postsynaptic neuron $\postsynaddr$:
+
 \begin{equation}\label{eq:stream_b}
@@
 \end{equation}
+
 % Unlike the binary stream of events from the signal as input, the postsynaptic neurons receive a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank = \synapticweight_\ranksyn$ when $\presynaddr_\arank = \presynaddr_\ranksyn$.
@@
 Indeed, by discretization of time, we may transform any event-based input as a boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$, defined for all polarities $p$, times $t$ and space coordinates $x$ and $y$.
-The values are by definition equal to zero except at the occurrences of events: $\forall \numevent \in \mathbb{N}$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
+The values are by definition equal to zero except at the occurrences of events: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
 First, one can note that by using a discretization, the computational brick used in equation~\eqref{eq:kernel_b} corresponds to a temporal convolution which transforms the input $A$ using one kernel per postsynaptic neuron~\citep{grimaldi_learning_2022}.
@@
 Using this dense representation, the processing of $A$ by the model can be written as:
+
 \begin{equation}\label{eq:kernel_b}
@@
 \end{equation}
+
 % The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$: equation~\eqref{eq:stream_b}
@@
 Using the kernels, it transforms the input raster plot into a probability with the following formula:
+
 \begin{equation}\label{eq:mlr}
@@
 \end{equation}
+
 where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function and  $\bias_\class$ is the bias linked to class $\class$. 
@@
 It results in the selection of the coefficients with the maximum likelihood value.
-Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels.
+Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels \note{change accordingly to the model's parameters}.
 This threshold was found using cross-validation and future work should automate how this may be determined for various datasets.
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/HD-SNN.png}
+    \includegraphics[width=\linewidth]{figures/conv_HD-SNN.pdf}
+    % \includegraphics[width=\linewidth]{figures/HD-SNN.png}
     \caption{
@@
 This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier_unsupervised_2007}.
-Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$).
+Simulations are performed with the \note{PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$)}.
 We demonstrated in a previous work that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
methods : caption
Update on Overleaf.

@@
 The model has been defined to provide optimal detection of event-driven spatio-temporal motifs.
-We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in the early visual cortex.
-We then demonstrated that the model shows similarities to biological responses.
+We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in the early visual cortex~\cite{kremkow_push-pull_2016}.
+%We then demonstrated that the model shows similarities to biological responses.
 We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware.

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
merge

@@
 %%%-----------------------------------------------------------------
-In summary, in this paper we have presented a HD-SNN model that has been trained and evaluated on a complex motion detection task.
+The HD-SNN model that has been trained and evaluated on a complex motion detection task.
 The model has been defined to provide optimal detection of event-driven spatio-temporal motifs.
@@
 Instead of relying on a careful description of the physical rules governing a task, e.g.
-the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}, this allows a more flexible definition of the model using this properly labeled dataset.
+the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}, this allows a more flexible definition of the model using this properly labelled dataset.
 %

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
4.1

@@
 First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we observed to be less efficient, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs.
-Moreover, in comparison to HOTS~\citep{lagorce_hots_2017}, the weights are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection of each spatio-temporal spike motif.
+Moreover, in comparison to a representation with time surfaces, the weights are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection of each spatio-temporal spike motif.
 Another novelty is that the model learns the weights and the delays simultaneously.

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
Methods : caption

@@
 
-% The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$: equation~\eqref{eq:stream_b}
-% \begin{equation}\label{eq:conv}
-% \forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
-% \mathcal{C}_\class(\presynaddr,\timev)
-% = \sum_{\ranksyn  \in [1, N_{\synapse^\postsynaddr}]} \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
-% \end{equation}
+%The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$: equation~\eqref{eq:stream_b}
+%\begin{equation}\label{eq:conv}
+%\forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
+%\mathcal{C}_\class(\presynaddr,\timev)
+%= \sum_{\ranksyn  \in [1, N_{\synapse^\postsynaddr}]} \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
+%\end{equation}
 This shows that $\mathcal{C}_\class$ is a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration).
@@
     \caption{
-    \textbf{- HD-SNN illustration.
+    % \textbf{- HD-SNN illustration.
 From left to right:} A 3D representation of the event stream as output of the DVS.
@@
 With an appropriate spiking mechanism, here a winner-takes-all, we obtain, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class.
+}
+    \textbf{Application of HD-SNN to the motion detection task.} \textit{(Left)}~We plot here as  a raster plot a slice of the input event stream (showing ON spikes in red and OFF spikes in blue for each vertical position and time for one given horizontal position in the middle of the sensor), displaying a pure translation motion upwards.
+A spatio-temporal convolution is applied on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis, and summed over the two polarities.
+At any given time (represented by the blue dot), we show the extent of the kernels (green dashed frame).
+\textit{(Middle)}~There are as many kernels as the possible number of motion classes in the recording and we show the activation resulting from the convolution for this specific motion upwards.
+The activation can be above chance (warm colors) or below chance (cold colors).
+For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e.
+softmax function).
+The output of the MLR gives a probability for each class, associated to a specific kernel.
+\textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all, we obtain as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class (here red is associated to a motion upwards).
+Most of the motion is accuratly detected.
 }
@@
 %%%-----------------------------------------------------------------
-
 Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection.

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Methods : caption through  DeepWrite

@@
 }
-    \textbf{Application of HD-SNN to the motion detection task.} \textit{(Left)}~We plot here as  a raster plot a slice of the input event stream (showing ON spikes in red and OFF spikes in blue for each vertical position and time for one given horizontal position in the middle of the sensor), displaying a pure translation motion upwards.
-A spatio-temporal convolution is applied on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis, and summed over the two polarities.
-At any given time (represented by the blue dot), we show the extent of the kernels (green dashed frame).
-\textit{(Middle)}~There are as many kernels as the possible number of motion classes in the recording and we show the activation resulting from the convolution for this specific motion upwards.
+    \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot here as a raster plot a slice of the input event stream (showing ON spikes in red and OFF spikes in blue for each vertical position and time for a given horizontal position at the center of the sensor), with a pure translation motion upwards.
+A spatio-temporal convolution is applied to the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis.
+The convolution is summed over the two polarities.
+At any given time (represented by the blue point), we show the extent of the kernels (green dashed frame).
+\textit{(Middle)}~There are as many kernels as there are possible motion classes in the recording, and we show the activation resulting from the convolution for this particular motion.
 The activation can be above chance (warm colors) or below chance (cold colors).
-For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e.
-softmax function).
-The output of the MLR gives a probability for each class, associated to a specific kernel.
-\textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all, we obtain as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class (here red is associated to a motion upwards).
-Most of the motion is accuratly detected.
+For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the nonlinearity of the MLR model (i.e., the softmax function).
+The output of the MLR gives a probability for each class associated with a particular kernel.
+\textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated to a thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated to a specific motion class (here red is associated to an upward motion).
+Most of the motion is detected accurately.
 }

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Methods : 2.4

@@
 Since the model is fully differentiable, we can now implement a semi-supervised learning rule.
-The loss function of the MLR model is the binary cross-entropy on the output of the classification layer.
-Supervision was implemented using the input binary events as defined above and the motion direction labels as the desired output.
+The loss function of the MLR model is the binary cross entropy on the output of the classification layer.
+Supervision was implemented using the binary input events as defined above and the motion direction labels as the desired output.
 The labels were defined at each time point as a one-hot encoding of the current motion in the channel corresponding to the current motion for all positions.
-As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner.
-It is important to note that we weighted the cost function by considering only active cells, such that the error is only back propagated to the spatial locations of these most active cells.
-This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier_unsupervised_2007}.
+Since in this semi-supervised context the label is known but the timing is not, we used a selection of the temporal support in an unsupervised manner.
+It is important to note that we weighted the cost function by considering only active cells, so that the error is only propagated back to the spatial locations of these most active cells.
+This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}.
 Simulations are performed with the \note{PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$)}.
-We demonstrated in a previous work that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
-%
+We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
+
 Finally, the output of the MLR model is an event-based representation predicting at each position and at each time the probability of each motion.

### 2023-02-10 Grimaldi-etal-BiolCybernetics.tex
Methods : 2.4

@@
 
-Finally, the output of the MLR model is an event-based representation predicting at each position and at each time the probability of each motion.
-Such an output gives a form of optical flow that can be exploited for non-rigid motions, but we have defined here, for simplicity, an evaluation method that applies to our task with a full field motion.
-We have shown above that when different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities.
-By taking the logit of the probability of the output given by the model, we can therefore calculate the probability of the output.
-This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error of the velocity estimate given by the average estimate.
-These calculations are performed on a different input dataset than the one used in the training or validation steps.
+Finally, the output of the MLR model is an event-based representation that predicts the probability of each motion at each position and time.
+Such an output provides a form of optical flow that can be exploited for non-rigid motion, but we have defined here, for simplicity, an evaluation method that applies to our full-field motion task.
+We have shown above that if different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities.
+Thus, by taking the logit of the probability of the output given by the model, we can calculate the probability of the output.
+This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error of the velocity estimate given by the mean estimate.
+These calculations are performed on a different input data set than the one used in the training or validation steps.
 The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/pyTERtorch}.

### 2023-02-18 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% public
+% https://github.com/SpikeAI/pyTERtorch/
+% https://www.zotero.org/groups/4776796/fastmotiondetection
+%
+% private
+% https://framateam.org/spikeai/channels/fast-motion-biolcyb
+% https://www.overleaf.com/project/630f9044c38e7a3cea81a7b2
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 % Guest Editors:
@@
 % https://www.zotero.org/groups/4776796/fastmotiondetection
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 However, most artificial neural models do not take advantage of this important temporal dimension of the neural code.
-Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays, which we apply to the computer vision task of motion detection.
-Indeed, the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangments of synaptic inputs as they reach the basal dendritic tree.
+Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
+The model uses the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of synaptic inputs as they reach the basal dendritic tree.
 We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
-We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of event-based cameras.
-In particular, we quantify how the accuracy of the model can vary as a function of the overall computational load.
+We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or to event-based cameras and for which we will characterize the accuracy of the model in detecting visual motion.
+In particular, we quantify how the accuracy can vary as a function of the overall computational load showing it is still efficient at very low workloads.
 This end-to-end, event-driven computational building block could improve the performance of future spiking neural network (SNN) algorithms and in particular their implementation in neuromorphic chips.
@@
 LP received funding from the ANR project  N° ANR-20-CE23-0021 ``\href{AgileNeuroBot}{https://laurentperrinet.github.io/grant/anr-anr/}''
+
 % LP received funding 
@@
 \newcommand{\Acknowledgments}{
-Thanks to Thomas Schatz, Rosa Cossart, Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work.
-%Thanks to HL  for participating to Mustache November ©.
-%
+Thanks to % Thomas Schatz, Rosa Cossart,
+Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work. %Thanks to HL  for participating to Mustache November ©. %
 \Funding %
@@
 \newcommand{\DataAvailability}{
-This works is made reproducible using the following tools. First the code reproducing all figures is available on~\href{ https://github.com/SpikeAI/pyTERtorch}{GitHub}. % todo: make a new repo
-
-\noindent %The paper is available a an~\href{https://arxiv.org/abs/2205.03635}{arXiv preprint} with links to previous versions ands to the code. 
+This works is made reproducible using the following tools. First the code reproducing all figures is available on~\href{ https://github.com/SpikeAI/pyTERtorch}{GitHub}. % todo: open-up the repo
 Find also the associated~\href{https://www.zotero.org/groups/4776796/fastmotiondetection}{zotero group} used to gather relevant literature on the subject.
-
-% public
-% https://github.com/SpikeAI/pyTERtorch/tree/main/2022-09-30_BiolCybernetics
-% https://www.zotero.org/groups/4776796/fastmotiondetection
-
-% private
-% https://framateam.org/spikeai/channels/2022-10_icip-bordeaux
-% https://www.overleaf.com/project/630f9044c38e7a3cea81a7b2
-
-%In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study.
-Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.
-If the study did not report any data, you might add ``Not applicable'' here.
-Chen
-}
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                                 %%
+%% Please do not use \input{...} to include other tex files.       %%
+%% Submit your LaTeX manuscript as one .tex document.              %%
+%%                                                                 %%
+%% All additional figures and files should be attached             %%
+%% separately and not embedded in the \TeX\ document itself.       %%
+%%                                                                 %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing
+
+%%=======================================================%%
+%% to print line numbers in the margin use lineno option %%
+%%=======================================================%%
+
+%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+
+%%======================================================%%
+%% to compile with pdflatex/xelatex use pdflatex option %%
+%%======================================================%%
+
+%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+
+%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
+%%\documentclass[sn-mathphys]{sn-jnl}% Math and Physical Sciences Reference Style
+%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
+%%\documentclass[sn-vancouver]{sn-jnl}% Vancouver Reference Style
+%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style
+%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
+%%\documentclass[sn-standardnature]{sn-jnl}% Standard Nature Portfolio Reference Style
+\documentclass[default]{sn-jnl}% Default
+%%\documentclass[default,iicol]{sn-jnl}% Default with double column layout
+%%\documentclass[iicol]{sn-jnl}
+
+\usepackage[T1]{fontenc}
+%\usepackage{textcomp}
+\usepackage[utf8]{inputenc}
+%
+\usepackage{lineno}
+\linenumbers
+
+%%%% Standard Packages
+%%<additional latex packages if required can be included here>
+%\usepackage[lofdepth,lotdepth,position=top]{subfig}
+%\usepackage{tabularx}
+%%%%
+
+%%%%%=============================================================================%%%%
+%%%%  Remarks: This template is provided to aid authors with the preparation
+%%%%  of original research articles intended for submission to journals published 
+%%%%  by Springer Nature. The guidance has been prepared in partnership with 
+%%%%  production teams to conform to Springer Nature technical requirements. 
+%%%%  Editorial and presentation requirements differ among journal portfolios and 
+%%%%  research disciplines. You may find sections in this template are irrelevant 
+%%%%  to your work and are empowered to omit any such section if allowed by the 
+%%%%  journal you intend to submit to. The submission guidelines and policies 
+%%%%  of the journal take precedence. A detailed User Manual is available in the 
+%%%%  template package for technical guidance.
+%%%%%=============================================================================%%%%
+
+\jyear{2023}%
+
+%% as per the requirement new theorem styles can be included as shown below
+\theoremstyle{thmstyleone}%
+\newtheorem{theorem}{Theorem}%  meant for continuous numbers
+%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
+%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
+\newtheorem{proposition}[theorem]{Proposition}% 
+%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
+
+\theoremstyle{thmstyletwo}%
+\newtheorem{example}{Example}%
+\newtheorem{remark}{Remark}%
+
+\theoremstyle{thmstylethree}%
+\newtheorem{definition}{Definition}%
+
+\raggedbottom
+
+\newcommand{\seeFig}[1]{see Fig.~\ref{fig:#1}}%{see Figure~\ref{fig:#1}}
+%% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+%\usepackage{cite}
+%\usepackage[usenames,dvipsnames]{xcolor} %The package starts from the basic facilities of the color package, and provides easy driver-independent access to several kinds of color tints, shades, tones, and mixes of arbitrary colors.
+%\usepackage{tcolorbox} % define boxes that can be customed, has to be loaded after xcolor package
+\usepackage{amsmath,amssymb,amsfonts} % amsmath and amssymb packages, useful for mathematical formulas and symbols
+%%\usepackage{float} %allows me to use [H] with figures to insert figures in multicol env (poster)
+%\usepackage{ulem} %The package provides an \ul (underline) command which will break over line ends; this technique may be used to replace \em (both in that form and as the \emph command), so as to make output look as if it comes from a typewrite.
+Caution: with this package, bibliography will make unbreakable and underlined journal titles
+%\usepackage[switch]{lineno} % add line numbers
+%\modulolinenumbers[5]
+\usepackage{physics} %The  goal  of  this  package  is  to  make  typesetting  equations  for  physics  simpler,  faster,  and  more  human-readable.
+%\usepackage{lineno} %Adds line numbers to selected paragraphs with reference possible through the LaTeX \ref and \pageref cross reference mechanism.
+%\usepackage{hyperref} %The hyperref package is used to handle cross-referencing commands in LaTeX to produce hypertext links in the document
+%\hypersetup{
+%    colorlinks = false,
+%    hidelinks
+%    %linkbordercolor = {white},
+%}
+%\usepackage{array} %An extended implementation of the array and tabular environments which extends the options for column formats, and provides “programmable” format specifications.
+%\usepackage{multicol} %Multicol defines a multicols environment which typesets text in multiple columns
+%%\usepackage{wrapfig} %Allows figures or tables to have text wrapped around them. wrapfig is used when calling multicols
+% \usepackage[position=top]{subfig} %The package provides support for the manipulation and reference of small or ‘sub’ figures and tables within a single figure or table environment.
+Linked to subfloat.
+% for units
+\usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.
+%\AtBeginDocument{\RenewCommandCopy\qty\SI}
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                                                 %%
-%% Please do not use \input{...} to include other tex files.       %%
-%% Submit your LaTeX manuscript as one .tex document.              %%
-%%                                                                 %%
-%% All additional figures and files should be attached             %%
-%% separately and not embedded in the \TeX\ document itself.       %%
-%%                                                                 %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing
-
-%%=======================================================%%
-%% to print line numbers in the margin use lineno option %%
-%%=======================================================%%
-
-%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
-
-%%======================================================%%
-%% to compile with pdflatex/xelatex use pdflatex option %%
-%%======================================================%%
-
-%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
-
-%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
-%%\documentclass[sn-mathphys]{sn-jnl}% Math and Physical Sciences Reference Style
-%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
-%%\documentclass[sn-vancouver]{sn-jnl}% Vancouver Reference Style
-%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style
-%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
-%%\documentclass[sn-standardnature]{sn-jnl}% Standard Nature Portfolio Reference Style
-\documentclass[default]{sn-jnl}% Default
-%%\documentclass[default,iicol]{sn-jnl}% Default with double column layout
-%%\documentclass[iicol]{sn-jnl}
-
-\usepackage[T1]{fontenc}
-%\usepackage{textcomp}
-\usepackage[utf8]{inputenc}
-%
-\usepackage{lineno}
-\linenumbers
-
-%%%% Standard Packages
-%%<additional latex packages if required can be included here>
-%\usepackage[lofdepth,lotdepth,position=top]{subfig}
-%\usepackage{tabularx}
-%%%%
-
-%%%%%=============================================================================%%%%
-%%%%  Remarks: This template is provided to aid authors with the preparation
-%%%%  of original research articles intended for submission to journals published 
-%%%%  by Springer Nature. The guidance has been prepared in partnership with 
-%%%%  production teams to conform to Springer Nature technical requirements. 
-%%%%  Editorial and presentation requirements differ among journal portfolios and 
-%%%%  research disciplines. You may find sections in this template are irrelevant 
-%%%%  to your work and are empowered to omit any such section if allowed by the 
-%%%%  journal you intend to submit to. The submission guidelines and policies 
-%%%%  of the journal take precedence. A detailed User Manual is available in the 
-%%%%  template package for technical guidance.
-%%%%%=============================================================================%%%%
-
-\jyear{2022}%
-
-%% as per the requirement new theorem styles can be included as shown below
-\theoremstyle{thmstyleone}%
-\newtheorem{theorem}{Theorem}%  meant for continuous numbers
-%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
-%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
-\newtheorem{proposition}[theorem]{Proposition}% 
-%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
-
-\theoremstyle{thmstyletwo}%
-\newtheorem{example}{Example}%
-\newtheorem{remark}{Remark}%
-
-\theoremstyle{thmstylethree}%
-\newtheorem{definition}{Definition}%
-
-\raggedbottom
-
-\newcommand{\seeFig}[1]{see Fig.~\ref{fig:#1}}%{see Figure~\ref{fig:#1}}
-%% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-%\usepackage{cite}
-%\usepackage[usenames,dvipsnames]{xcolor} %The package starts from the basic facilities of the color package, and provides easy driver-independent access to several kinds of color tints, shades, tones, and mixes of arbitrary colors.
-%\usepackage{tcolorbox} % define boxes that can be customed, has to be loaded after xcolor package
-\usepackage{amsmath,amssymb,amsfonts} % amsmath and amssymb packages, useful for mathematical formulas and symbols
-%%\usepackage{float} %allows me to use [H] with figures to insert figures in multicol env (poster)
-%\usepackage{ulem} %The package provides an \ul (underline) command which will break over line ends; this technique may be used to replace \em (both in that form and as the \emph command), so as to make output look as if it comes from a typewrite.
-Caution: with this package, bibliography will make unbreakable and underlined journal titles
-%\usepackage[switch]{lineno} % add line numbers
-%\modulolinenumbers[5]
-\usepackage{physics} %The  goal  of  this  package  is  to  make  typesetting  equations  for  physics  simpler,  faster,  and  more  human-readable.
-%\usepackage{lineno} %Adds line numbers to selected paragraphs with reference possible through the LaTeX \ref and \pageref cross reference mechanism.
-%\usepackage{hyperref} %The hyperref package is used to handle cross-referencing commands in LaTeX to produce hypertext links in the document
-%\hypersetup{
-%    colorlinks = false,
-%    hidelinks
-%    %linkbordercolor = {white},
-%}
-%\usepackage{array} %An extended implementation of the array and tabular environments which extends the options for column formats, and provides “programmable” format specifications.
-%\usepackage{multicol} %Multicol defines a multicols environment which typesets text in multiple columns
-%%\usepackage{wrapfig} %Allows figures or tables to have text wrapped around them. wrapfig is used when calling multicols
-% \usepackage[position=top]{subfig} %The package provides support for the manipulation and reference of small or ‘sub’ figures and tables within a single figure or table environment.
-Linked to subfloat.
-% for units
-\usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.
-%\AtBeginDocument{\RenewCommandCopy\qty\SI}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\DeclareMathOperator*{\argmin}{arg\,min}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:intro}
-
+%
 The human brain has the remarkable property of being able to react at any time while consuming a reasonable amount of energy, about tens of watts.
-This system is the result of millions of years of natural selection, and a striking difference between biological and artificial neural networks is the representation they use.
-Convolutional neural networks (CNNs), for example, represent the flow of information from one layer to another as tensors, storing visual information densely across the visual topography, with different properties represented in different channels.
+This system is the result of millions of years of natural selection, and a striking difference with artificial neural networks is the representation both use.
+Artificial convolutional neural networks (CNNs), for example, represent the flow of information from one layer to another as tensors, storing visual information densely across the visual topography, with different properties represented in different channels.
 These networks are known to mimic many properties of biological systems, such that each can be attributed a ``Brain Score''~\citep{schrimpf_brain-score_2020}, yet this score takes no account of inference speed or energy consumption.
@@
 Another type of SNN that uses accurate timing of spikes attempts to determine the structure of the network to minimize a cost function.
-This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been recently extended in gradient-based methods.
-This surrogate gradient method is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
-However, the performance of SNNs still lags behind that of firing rate based networks.
+This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been recently extended.
+This method uses a surrogate gradient and is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
+However, the performance of SNNs still lags behind that of firing-rate based networks.
 The question of the advantage of using spikes in machine learning and computer vision remains open.
@@
 These spatio-temporal motifs within the neural activity may be useful representations to perform computations for various cognitive tasks.
-In 1982, Abeles asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs~\citep{abeles_role_1982}.
-While the first possibility favors the theory of rate coding, the second one emphasizes the function of temporal precision in the neural code.
-Since then, numerous studies have demonstrated the emergence of synchronicity of activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
+Notably, \citet{abeles_role_1982} asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs.
+While the first possibility favors the firing-rate coding theory, the second one emphasizes the function of temporal precision in the neural code.
+Since then, numerous studies have demonstrated the emergence of synchronicity of activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding using spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
 All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
@@
 %
+%%%-----------------------------------------------------------------
+%%%-----------------------------------------------------------------
 \section{Methods}
@@
 %%%-----------------------------------------------------------------
-The goal is to test the model in a motion detection task using a realistic event-based data stream.
+Our objective in this paper is to test whether the HD-SNN model is capable to efficiently learn a motion detection task as defined by a realistic event-based data stream.
 This type of signal is typically recorded by a Dynamic Vision Sensor (DVS) and is inspired by the signal sent by retinal ganglion cells through the optic nerve.
-Events are binary and allow to capture the dynamics of the natural scene and perform various cognitive tasks quickly and effectively.
-Here, we will first describe the definition of the task and then the model we use to infer motion.
+Events are binary and should be sufficient to perform the motion detection task quickly and effectively.
+Here, we will first describe the definition of the task and then of the HD-SNN model we use to infer motion.
 %%%-----------------------------------------------------------------
@@
 %%%-----------------------------------------------------------------
-To train and test our model, we need to define a dataset for which we control the direction and speed of motion.
-Let us now define a procedure for animating a natural scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
+To train and test our model, we need to define a visual dataset for which we control the direction and speed of motion.
+Let us now define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
 First, we draw a trajectory inspired by the biological movements of the eyes.
@@
 In animals with a fovea, this is particularly useful because it allows the area with the highest density of photoreceptors in the environment to be moved, for example to a point of interest.
-A specific mechanism for this is called saccades, which are rapid eye movements that reposition the center of vision.
+A specific mechanism for this function are saccades, which are rapid eye movements that reposition the center of vision.
 In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}).
@@
 On a more microscopic scale, the human gaze moves with a continuous drift similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
-In order to maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.This approach first defines a finite set of possible 2D movements in polar coordinates.
+In order to maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.
+This approach first defines a finite set of possible 2D movements in polar coordinates.
 Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $8$ linearly spaced movement directions \note{and 3 geometrically spaced speeds}.
@@
     \caption{
-        {\bf - Motion Detection Task.} {\bf (Left)}~We use large natural images ($512\times512$) in which an aperture ($128\times128$) extracts a cropped image around the view axis.
-To mimic the effect of a saccadic eye movement, the view axis moves according to a stepwise random walk.
+        {\bf - Motion Detection Task.} {\bf (Left)}~We use large natural images ($256\times256$) in which an aperture ($128\times128$) extracts a cropped image around the view axis.
+To mimic the effect of a saccadic eye movement, the view axis moves according to a step-wise random walk.
 We show an example path of the trajectory for $200$ time steps (green line).
@@
 This can be transformed into an event-based representation.
-Mimicking the retina, this representation encodes pixel-per-pixel proportional increases or decreases in luminance in the image, i.e.
-ON (red) and OFF (blue) events.
+Mimicking the retina, this representation encodes proportional increases or decreases in luminance, i.e.
+ON (red) and OFF (blue) events, in each pixel of the image.
 In the lower left corner of the snapshots, the translation vector is shown in red as one of the possible classes of motion.
-Note the change in the direction of motion between the second and third image and that due to the aperture problems, contours parallel to the motion emit less spikes.}
+Note the change in the direction of motion between the second and third image and that due to the aperture problem, contours parallel to the motion emit relatively less spikes.}
     \label{fig:motion_task}
@@
 % For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images~\citep{olshausen_emergence_1996}.
-For this purpose, we selected a database of 600 natural images that were previously used to study the statistics of natural images~\citep{perrinet_edge_2015}.
+For this purpose, we selected a database of $600$ natural images that were previously used to study the statistics of natural images~\citep{perrinet_edge_2015}.
 Note that these have been pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band.
 This process is known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $256 \times 256$ in size, and we will extract sub-images of size $64 \times 64$ positioned around the center of gaze at each time step.
-We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 128~\ms$.
+These images are $256 \times 256$ in size, and we will extract sub-images of size $128 \times 128$ positioned around the center of gaze at each time step.
+We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 200~\ms$.
 To avoid boundary effects, we will randomly position the full trajectory in image space such that the subimage is translated using the position given by the trajectory at each time step without touching the borders.
-The translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
-Note that the magnitude of the displacement is relative to the time bin, and we have defined the velocity to correspond to a movement of one pixel per frame (i.e., per time bin).
+At each point in time, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
+Note that the magnitude of the displacement is relative to the time bin, and we have defined the unit velocity to correspond to a movement of one pixel per frame (i.e., per time bin).
 
@@
     \includegraphics[width=0.980\linewidth]{figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
-      \caption{\textbf{Core mechanism of the HD-SNN model.} \textit{(Left)}~In this example inspired by the polychronization model~\citep{izhikevich_polychronization_2006}, three presynaptic neurons denoted \textit{b}, \textit{c} and, \textit{d} are fully connected to two post-synaptic neurons \textit{a} and \textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for \textit{a} and  $8$, $5$, and $1~\ms$ for \textit{e}.
-\textit{(Middle)}~If three synchronous pulses are emitted from presynaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
+  \caption{\textbf{Core mechanism of the HD-SNN model.} \textit{(Left)}~In this example inspired by the polychronization model~\citep{izhikevich_polychronization_2006}, three pre-synaptic neurons denoted~\textit{b},~\textit{c} and~\textit{d} are fully connected to two post-synaptic neurons~\textit{a} and~\textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for~\textit{a} and $8$, $5$, and $1~\ms$ for~\textit{e} (in this schematics, we consider unitary synaptic weights).
+\textit{(Middle)}~If three synchronous pulses are emitted from pre-synaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
   %at these different delays, and these may not be sufficient to generate a spike in either neuron.
-      \textit{(Right)}~If the pulses are emitted from presynaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron \textit{a} at the same time (here, at $t=10~\ms$),  the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from post-synaptic neuron \textit{e}.
-Figure reproduced from~\cite{grimaldi_precise_2023} under a CC-BY licence.
-       }
-    \label{fig:izhikevich}
-  \end{figure}
+  \textit{(Right)}~If the pulses are emitted from pre-synaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron~\textit{a} at the same time (here, at $t=10~\ms$), the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from the post-synaptic neuron~\textit{e}.
+Figure reproduced from~\cite{grimaldi_precise_2023} under a CC-BY licence.}%
+    \label{fig:izhikevich}%
+\end{figure}%
 % 
 %----------------------------%
-
-In neurobiological recordings or in the sensory signal obtained from an event-based camera, any generic raster plot consists of a stream of \emph{spikes} or \emph{events}.
+In a neurobiological recording or in the sensory signal obtained from an event-based camera, the input consists of a stream of \emph{spikes} or \emph{events}.
 This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events.
@@
 In a neurobiological recording, this can be the identified set of presynaptic neurons.
-For event-based cameras, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$ where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras) and $(\Nx, \Ny)$ is the size of the sensor in pixels.
+For neuromorphic hardware, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$ where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels.
 As such, an address $\presynaddr_\arank$ is typically in the form $(\polev_\arank, x_\arank, y_\arank)$ for event-based cameras.
 
-In the HD-SNN, each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferents from $\presynaddrspace$.
+In the HD-SNN model, each neuron $\postsynaddr \in \postsynaddrspace$ connects to pre-synaptic afferents from $\presynaddrspace$.
 In biology, a single cortical neuron has generally several thousands of synapses.
-Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the pre-synaptic neuron's soma to that of the post-synaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a pre-synaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
 Note that a neuron may contact an afferent neuron with multiple different delays.
 Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
-as  $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
+as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a pre-synaptic address $\presynaddr_\ranksyn$, a post-synaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
 This defines the full connectivity of the HD-SNN model.
 
-Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$.
-As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field.
-It results in a weighted event stream for each  postsynaptic neuron $\postsynaddr$:
-
+Of interest is to define the receptive field of a post-synaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$, or the emitting field of a pre-synaptic neuron $\synapse_\presynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$.
+As a consequence, a post-synaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field.
+It results in a weighted event stream for each  post-synaptic neuron $\postsynaddr$:
 \begin{equation}\label{eq:stream_b}
@@
 \end{equation}
-
-% Unlike the binary stream of events from the signal as input, the postsynaptic neurons receive a weighted stream of event such that for each event $\arank$ its value is $\synapticweight_\arank = \synapticweight_\ranksyn$ when $\presynaddr_\arank = \presynaddr_\ranksyn$.
-This new stream of events is ordered in time in biology as events reach the soma of post-synaptic neurons but should be reordered in simulations.
+In biology, this new stream of events is naturally ordered in time as events reach the soma of post-synaptic neurons.
+However, it should be properly reordered in simulations.
 Crucially, when post-synaptic neurons are activated on their soma by a specific spatio-temporal motif which is imprinted in the set of synapses, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner while the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions (see Figure~\ref{fig:izhikevich}).
@@
 %
-% \note{I can make a small figure with event streams of different colors for the input one, the multiplexed one etc and the response in the bottom}
-%
-\subsection{Application of HD-SNN to the motion detection task}
-%
-From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation.
-Indeed, by discretization of time, we may transform any event-based input as a boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$, defined for all polarities $p$, times $t$ and space coordinates $x$ and $y$.
+\subsection{Application of the HD-SNN model to the motion detection task}
+%
+Such event-based model can be adapted to computer vision tasks.
+In particular, from the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation.
+Indeed, by discretization of time, we may transform any event-based input from an event-based camera as a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$, defined for all polarities $p$, times $t$ and space coordinates $x$ and $y$.
 The values are by definition equal to zero except at the occurrences of events: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
@@
 % The characteristics of the kernels are shared among one class of neurons such that $\forall i,j \in [1, N_\class]^2, \; \kernel^i_\class = \kernel^j_\class$.
+
 % Here, we will consider that heterogeneous delays are limited in range
-In particular, the whole synaptic set can be represented as a one kernel for each class $\class$ of the supervision task as the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
+In particular, the whole synaptic set can be represented as one kernel for each class $\class$ of the supervision task as the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
 To keep an analogy with the HD-SNN model, for each neuron $\postsynaddr$ of position $(x, y)$ and class $\class$, $\kernel_\class$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$.
@@
 Using this dense representation, the processing of $A$ by the model can be written as:
-
 \begin{equation}\label{eq:kernel_b}
@@
 \end{equation}
-
 %The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$: equation~\eqref{eq:stream_b}
@@
 %\end{equation}
-This shows that $\mathcal{C}_\class$ is a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration).
+This shows that $\mathcal{C}_\class$ is the result of a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration).
 Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
 This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
-The same type of spatio-temporal filtering is used as a preprocessing stage of a pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
+The same type of spatio-temporal filtering is used as a pre-processing stage of a pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
 Also,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm which implements a motion estimation task.
@@
 
-Here, we keep the analogy with spiking neurons, and we try to use as few heuristics as possible to observe the emergence of spiking motifs in the spatio-temporal kernels.
+Here, we keep the analogy with spiking neurons, and we try to use as few heuristics as possible to observe the detection of spiking motifs using the spatio-temporal kernels.
 If so, this precise spatio-temporal patterns prove to be of interest for neural computations and one can assume that biological neurons make use of this information as well.
 % \note{plus de biblio sur la motion detection/segmentation ou optique flow avec les events? (reviewer 2)}
-The 3D convolution represents the linear integration of the spike train as input.
-The activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
-In our MLR model, a probability value, for each class (i.e.
-each motion direction), is predicted for each event at position $x, y$ and at time $\timev$ as a softmax function of the linear combination of the list of events.
+The 3D convolution represents the linear integration of the spike train as the linear input to the neuron.
+Then, the activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
+In our MLR model, a probability value for each class (i.e.
+each motion direction) is predicted for each position $x, y$ and time $\timev$ as a softmax function of the linear combination of the list of events.
 % on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
 The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the HD-SNN model.
-Using the kernels, it transforms the input raster plot into a probability with the following formula:
-
+Formally, using the kernels, it transforms the input raster plot into a probability with the following formula:
 \begin{equation}\label{eq:mlr}
@@
 \end{equation} 
-
 where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function and  $\bias_\class$ is the bias linked to class $\class$. 
@@
 
-Once explained the general framework used to solve the ecological task described in section~\ref{sec:task}, we can add some heuristics based on neuroscientific observations to constrain our model and its strategies to solve the task.
-Note that the general framework is the same as the one presented in~\cite{grimaldi_learning_2022} and that, appart from the task to solve and a deeper analysis of the results, these priors, inspired by neuroscience, are the main methodological differences with this previous study.
-We set \note{arbitrarily} the size of the kernels to $(9, 17, 17)$ and define as many classes as the number of motion direction: $N_\class = 8$.
-Because receptive fields of biological neurons are mainly circular and can be more or less elongated, we apply a circular mask on the spatial dimensions of the kernels.
-And because the synaptic delay is linked, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one, we apply a linear mask on the temporal axis such that the connectivity depends on the delay: the lower the delay the lower the circular mask radius (see Figure~\ref{fig:kernels} for an illustration).
+Once the general framework used to solve the ecological task described in section~\ref{sec:task} explained, we can add some heuristics based on neuro-scientific observations to constrain our model and its strategies to solve the task.
+Note that the general framework is the same as the one presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel heuristics, inspired by neuroscience, are the main methodological differences with that previous study.
+In our simulations, We set the size of the kernels to $(9, 17, 17)$ and define as many classes as the number of motions (directions and speeds): $N_\class = 8 \times 3$.
+To avoid introducing biases in the directions, we apply a circular mask on the spatial dimensions of the kernels.
+And because the synaptic delay is linked, by physical principles, to the distance between the pre-synaptic neuron and the postsynaptic one, we apply a linear mask on the temporal axis such that the connectivity depends on the delay: the lower the delay, the lower the circular mask radius (see Figure~\ref{fig:kernels} for an illustration).
 The second heuristics is based on the observation that biological neurons have a sparse spiking activity.
@@
 It results in the selection of the coefficients with the maximum likelihood value.
-Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels \note{change accordingly to the model's parameters}.
+Therefore, we will only select a given proportion of the most active cells, which we set here to $0.1\%$ of all output voxels \note{change accordingly to the model's parameters}.
 This threshold was found using cross-validation and future work should automate how this may be determined for various datasets.
@@
 The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/pyTERtorch}.
-%\note{à rendre publique}
+%
 %
@@
 Because of the symmetry observed between the ON and OFF event streams, we observed that kernels are very similar for the ON polarities.
-These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
+These weights are associated to a specific delay on the \textit{delays} axis and to a pre-synaptic address defined on the two other axes.
 %	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.
@@
 We also observe that these cells show an orientation selectivity, similar to that observed in MT neurons~\citep{deangelis_functional_1999}.
-Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results~\citep{baudot_animation_2013}.
+Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results observed after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
 Focusing on the positive weights, a strong selectivity is observed along specific axes of motion for each of the different kernels.
@@
 If one focuses on the interpretation of these kernels in terms of spatio-temporal motifs embedded in the event stream, it can lead to interesting outcomes.
-In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions.
+In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a pre-synaptic activity associated to the different motion directions.
 Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spike motif.

### 2023-02-19 Grimaldi-etal-BiolCybernetics.tex
new figures

@@
 This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{APROVIS3D}{http://aprovis3d.eu/}''). 
-% Agence Nationale de la Recherche %HL and
 LP received funding from the ANR project  N° ANR-20-CE23-0021 ``\href{AgileNeuroBot}{https://laurentperrinet.github.io/grant/anr-anr/}''
-
-% LP received funding 
 and from A*Midex grant number AMX-21-RID-025
 ``\href{Polychronies}{https://laurentperrinet.github.io/grant/polychronies/}''. 
-%HL received funding from a University of Montreal Artificial Intelligence scholarship. 
-%
 }
@@
 Thanks to % Thomas Schatz, Rosa Cossart,
-Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work. %Thanks to HL  for participating to Mustache November ©. %
+Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work. 
 \Funding %
@@
 \jyear{2023}%
-
 %% as per the requirement new theorem styles can be included as shown below
@@
 \newtheorem{theorem}{Theorem}%  meant for continuous numbers
-%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
-%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
 \newtheorem{proposition}[theorem]{Proposition}% 
-%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
-
+%
 \theoremstyle{thmstyletwo}%
@@
 \usepackage{physics} %The  goal  of  this  package  is  to  make  typesetting  equations  for  physics  simpler,  faster,  and  more  human-readable.
-%\usepackage{lineno} %Adds line numbers to selected paragraphs with reference possible through the LaTeX \ref and \pageref cross reference mechanism.
 %\usepackage{hyperref} %The hyperref package is used to handle cross-referencing commands in LaTeX to produce hypertext links in the document
@@
 %%%-----------------------------------------------------------------
-% Title.
-% ------
 \title[\ShortTitle]{\Title}
 \artnote{This article is published as part of the Special Issue on "What can Computer Vision learn from Visual Neuroscience?"}
-%
-% Single address.
 % ---------------
@@
 \affil{\orgdiv{\Department}, \orgname{\Affiliation}\\ \orgaddress{\street{\Street}, \city{\City}, \postcode{\PostCode}, \country{\Country}}}
-%%==================================%%
-%% sample for unstructured abstract %%
-%%==================================%%
-
+%
 \abstract{
@@
 
-%\ninept
-%
-%\abstract{Abstract}
-%
 \keywords{\Keywords}
@@
 \maketitle
-%%%%%%%%% BODY TEXT
+%
 \section{Introduction}
-%%%-----------------------------------------------------------------
 %%%-----------------------------------------------------------------
@@
     \centering
-    \includegraphics[width=0.980\linewidth]{figures/izhikevich.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
+    \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
   \caption{\textbf{Core mechanism of the HD-SNN model.} \textit{(Left)}~In this example inspired by the polychronization model~\citep{izhikevich_polychronization_2006}, three pre-synaptic neurons denoted~\textit{b},~\textit{c} and~\textit{d} are fully connected to two post-synaptic neurons~\textit{a} and~\textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for~\textit{a} and $8$, $5$, and $1~\ms$ for~\textit{e} (in this schematics, we consider unitary synaptic weights).
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/conv_HD-SNN.pdf}
+    \includegraphics[width=\linewidth]{figures/HDSNN_conv.pdf}
     % \includegraphics[width=\linewidth]{figures/HD-SNN.png}
@@
 In this case, the number of computations is greatly reduced compared to~\citep{grimaldi_robust_2022}, thus demonstrating the efficiency of the presented method.
-
 %
@@
 \end{figure}
-
-% \note{try also a quantization of the weights (binary?)}
 %
 \section{Discussion}
-%%%-----------------------------------------------------------------
 %%%-----------------------------------------------------------------
@@
 %
-% Acknowledgement hidden for review.
 \Acknowledgments
@@
 Not applicable.
-% The authors do not have any financial or non-financial interests directly or indirectly related to this work.
 

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Once these eye trajectories are generated, we can apply them to a visual scene.  
-% For this purpose, we selected a database of 10 natural grayscale images that are commonly used to study the statistics of natural images~\citep{olshausen_emergence_1996}.
 For this purpose, we selected a database of $600$ natural images that were previously used to study the statistics of natural images~\citep{perrinet_edge_2015}.
@@
 For a specific pixel and timestamp, an event is generated if the absolute value of this gradient exceeds a threshold.
-The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive.
+The event has either an OFF or ON polarity respectively, whether the gradient is negative or positive.
 This signed threshold value is then subtracted from the residual gradient image.
@@
 % \note{ the task is full-field rigid translations, the model resolves it locally (analogy to the input from V1 to MT) }
-%
-%\subsection{Detecting event-based motifs using spiking neurons with heterogeneous delays}% model as a Logistic Regression with a temporal convolution}
-%%%%-----------------------------------------------------------------
-%%
-%\begin{figure}%[t!]
-%    \centering
-%    \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
-%    \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
-%    \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
-%    \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
-%	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
-%	    {\bf (a)}~Given a generic raster plot defined by a set of spikes occurring on specific neural addresses and at specific times, one may consider that this information consists of the repeated occurrence of precise spiking motifs.
-%	    {\bf (b)}~We show the motifs used in this example, each identified at the top by a different color.
-To each of the $10$ neural adresses and $71$ different possible delays is assigned an evidence of activation (red) or deactivation (blue).
-Note that each afferent may be connected with multiple weights at different delays.
-%	    {\bf (c)}~The activation in time of the different motifs is then used to define a generative model for drawing a raster plot on the multi-unit address space: The propagation of the afferent information through these delays generates the raster plot seen in (a).
-%	    {\bf (d)}~The inversion of the generative model provides with a model which gives the predicted probability of occurrence of each motif at any time and which may be used to generate a spike as a Bernoulli trial.
-The detection model yields in this particular case with an exact identification of the occurrence of motifs.
-Knowing the results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic examples, all spikes can be annotated with each identified spiking motif.
-%	    }
-%    \label{fig:model}
-%\end{figure}
-%% 
-%\subsubsection{A generative model for raster plots}
-%%
-%In neurobiological recordings or in the sensory signal obtained from an event-based camera, any generic raster plot consists of a stream of \emph{spikes} or \emph{events} (see Figure~\ref{fig:model}-(a)).
-This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events.
-Events are typically ordered by their time of occurrence.
-Each event has a time of occurrence $\timev_\arank$  and an associated address $\presynaddr_\arank$ (which is typically in the form $(x_\arank, y_\arank, \polev_\arank)$ for event-based cameras).
-This defines an address space $\presynaddrspace$ which consists of the set of possible addresses.
-In a neurobiological recording, this can be the identified set of neurons.
-For event-based cameras, it is denoted by $[1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras).
-%
-%% This raster plot is generated by spiking neurons which are defined on the one hand by the equations governing the evolution of its membrane potential dynamics on its soma and on the other hand by the characterization of the synaptic contacts on its dendritic tree, or on the sensory processes (such as the pixel on the sensor of a camera, or the photoreceptors in the retina).
-A classical characterization consists in detailing the synaptic weights of each synaptic contact on the dendritic tree, the so-called weight matrix.
-A more detailed description of these synaptic contacts reveals the added importance of heterogeneous delays, i.e., the precise timing from one afferent neuron's firing to its arrival in the soma, and how this changes the network's dynamics~\citep{izhikevich_polychronization_2006}.
-In that description, we can parameterize each neuron by the set of tuples defining both the weight and the delay of each synaptic contact.
-As a consequence, a set of input presynaptic spikes $\event$ will be processed by the dendrites defined by this synaptic set and notably by the respective delays, which will multiplex in time all events.
-%
-%Let's formalize a layer of spiking neurons with heterogeneous delays (HD-SNN).
-Each neuron $\postsynaddr \in \postsynaddrspace$  connects to presynaptic afferent from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses.
-Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-Note that a neuron may contact an afferent neuron with multiple different delays.
-Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
-%as  $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$,  a weight $\synapticweight^\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
-This defines the full connectivity of the HD-SNN model.
-Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]} $, or the emitting field of a presynaptic neuron $\synapse_\presynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]}$.
-As a consequence, an event stream which evokes neurons in the presynaptic address space is multiplexed by the synapses into a new event stream which is defined by the union of the sets generated by each emitting field from the presynaptic space:
-%$ \cup_{\arank \in [1,\numevent]} \{ \{(\postsynaddr_\ranksyn, \synapticweight_\ranksyn, \timev_\arank + \synapticdelay_\ranksyn) \}_{ \ranksyn \in \synapse_{\presynaddr_\arank}} \}$.
-This new stream of events is by nature ordered in time as events reach the soma of post-synaptic neurons.
-In particular, when post-synaptic neurons are activated on their soma by this spatio-temporal motif, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner.
-%% \note{ TODO: show why this formula implements a polychrony detector }
-%
-%% \note{ éclaircir }
-%
-%Taking the argument the other way around, one may from a generative model for generic raster plots.
-Indeed, any spike in the presynaptic address space is generated by sensory neurons (for instance photoreceptors in the retina, sensors in a CMOS chip) or by afferent spiking neurons.
-In the latter case, these are connected to the spiking cell by a set of weights and delays, whose structure is stable relatively to the coding timescale.
-When these connections are high and sparsely distributed, this firing will cause a specific temporal motif.
-Another example is given for the barn owl auditory system: As it hears the sound of a mouse, this sound will generate a specific spiking response in both ears, and specifically, the precise timing between the signal generated by the left relative to the right ear can be for instance used to determine the position of the prey~\citep{goodman_spike-timing-based_2010}.
-Overall, these examples show that raster plots may be considered as a mixture of the effects of different elementary causes, and that each event triggers a specific spatio-temporal spiking motif.
-%%
-%\subsubsection{Detecting spiking motifs}
-%%: Detection model
-%
-%From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers, it is useful to transform this event-based representation into a dense representation.
-Indeed, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{N\times T}$, where $N$ is the number of neurons in $\presynaddrspace$ and $T$ is the number of time bins.
-In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$.
-%The probability of firing of a neuron $a$ at a given time $t$ can be understood as a Bernoulli trial whose (only) parameter is a bias $p(t, a) \in [0, 1]$.
-Assuming that the presence of spiking motifs conditions the probability on all efferents, the logit (inverse of the sigmoid) of this probability bias can be written as the sum of the logit of each of these factors, whose values are defined by the corresponding weights.
-Spiking motifs may be activated independently and at random times, such that  we write this activity as $B(b, t)=1$ if $b$ is activated at $t$ and else $B(b, t)=0$.
-We can thus write the probability bias as the accumulated evidence given these factors as
-%\begin{equation*}
-%p(t, a) = \sigma\big(W_0 + \sum_{b, t} B(b, t) \cdot W_b(a, t-d) \big)  
-%\end{equation*}
-%where $\sigma$ is the sigmoid function.
-We will further assume that the weights are balanced (their mean is zero) and that $W_0$ is a bias such that $p_0=\sigma(W_0)$ is the average background firing rate.
-Conveniently, one can write this summation as a one-dimensional temporal convolution operator such that we may simply write
-%\begin{equation*}
-%p = \sigma(W_0 + B \ast W )
-%\end{equation*}
-%where  $p\in [ 0, 1]^{N\times T}$ and $B\in \{0, 1\}^{M\times T}$ is the raster plot corresponding to the temporal activation of the PGs.
-Finally, we obtain the raster plot $A\in \{0, 1\}^{N\times T}$ by drawing spikes using independent Bernoulli trials $A \sim \mathcal{B}(p)$.
-Note that, depending on the shape of the kernels, the generative model can model a discretized Poisson process, generate rhythmic activity or more generally propagating waves.
-This formulation thus defines a simple generative model for raster plots as a combination of independent PGs.
-%
-%Using this dense representation, the counting defined above becomes:
-%\begin{equation*}
-%\mathcal{C}^\postsynaddr(a,t)
-%= \sum_{\presynaddr, \synapticdelay_\ranksyn^\postsynaddr} \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn^\postsynaddr)
-%\end{equation*}
-%%
-%This shows that $\mathcal{C}^\postsynaddr$ is a temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast A$.
-%This well-known computation defines a time-invariant, differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
-%%
-%
-%\note{formalize the following:}
-%More generally, every cause may be considered as occurring independently and one might write the generative model for the generation of the presynaptic events as:
-%
-%
-%Using this formalization, one might now deduce an optimal algorithm for the detection of such temporal motifs.
-%
-%
-%By discretization of time (with here an arbitrary unitary time unit), we can also define the dendrite as a matrix giving the weight corresponding to the different delays $d \in [0, D]$ (where $D$ is the maximum delay) on different pre-synaptic addresses $a \in [1, N]$ defining the list of the $N$ dendrites.
-We will denote as $W(a, d)$ these weights.
-%
-%
-%Following the observations of~\citet{izhikevich_polychronization_2006}, let us assume that such precise discharge motif defines a polychronous group (PG).
-Assuming that we know there exists $M$ such groups, we will define as $b \in [1, M]$ the address of a PG and as $W_b$ the corresponding weight matrices.
-This allows then to derive a generative model for raster plots (see Figure~\ref{fig:model}-(b)).
-
-% 
-%This generative model defined above allows to determine this inference model for guessing sources $B$ when observing a raster plot $A$.
-This assumes that we know the spiking motifs as defined by the $W_b$ matrices.
-The underlying metric is the binary cross-entropy, as used in the logistic regression model.
-In particular, if we consider kernels with similar decreasing exponential time profile, we can prove that this is similar to the method of~\citet{berens_fast_2012}.
-In our specific case, the difference is that the regression is performed in both dendritic and delay space by extending the summation using a temporal convolution operator.
-Using this forward model, it is possible to estimate the logit (inverse of a sigmoid) $\hat{B}(b, t)$ for the presence of a PG of address $b$ and at time $t$ by using the transpose convolution operator.
-Equivalently, this consists in using the emitting field $\synapse_\presynaddr$ of presynaptic neurons in place of the receptive field $\synapse^\postsynaddr$ of postsynaptic neurons.
-It thus comes that when observing $A$, then one may infer $\hat{B} = A \ast W^T$ and select the most activated items.
-This assumption hold as long as the kernels are uncorrelated, a condition which is met here numerically by chosing a relatively sparse set of synapses (approximately $1\%$ of active synapses).
-
-%\subsubsection{Implementation using spiking neurons with heterogeneous delays}
-%
-% The activation function of our spiking neural is a softmax function implementing a form of  Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
-%It transforms this list of weights into a probability with the following formula:
-%$
-%Pr(k=\postsynaddr \; \vert \; \timev) =
-%\frac 1 Z
-%{\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }
-%$ 
-%where $\mathcal{C}^\postsynaddr(\timev) = \sum
-%\activeweights^\postsynaddr(t)
-%$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$. 
-%In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
-%%
-%
-% \note{say that it is effortless in biology but difficult on conventional computers}
-% In our MLR model with $\Nclass=\Nspeed$ classes, a probability value is predicted for each event at address $\presynaddr_\arank$ and at time $\timev_\arank$ as a softmax function of the linear combination of the list of events on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
-The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the heterogeneous delays model.
-%
-%\subsubsection{Performance on synthetic data}
-%%
-%\begin{figure}%[t!]
-%    \centering
-%    \includegraphics[width=0.490\linewidth]{figures/THC_N_PGs.pdf}
-%    \includegraphics[width=0.490\linewidth]{figures/THC_N_PG_time.pdf}
-%	    \caption{Detecting event-based motifs using spiking neurons with heterogeneous delays. 
-%	    {\bf (a)}~Accuracy of detection as a function of the number of kernels.
-%	    {\bf (b)}~Accuracy of detection as a function of the temporal depth $D$ of kernels among $M=144$ kernels.
-%	    }
-%    \label{fig:model_results}
-%\end{figure}
-%
-%To quantify the efficiency of this operation, we generated $M=144$ synthetic spiking motifs as random independent kernels over $128$ presynaptic inputs and $D=71$ possible delays.
-We drew random independent instances of $B$ with a length of $T=1000$ time steps and with on average $1.0$ occurrences in each draw.
-This allowed us to generate raster plots which we use to infer $\hat{B}$.
-We compute the accuracy as the rate of true positive detections (both for inferring the address and its exact timing) and observe on average $\approx 98\%$ correct detections.
-We further extended this result by showing how the accuracy would evolve as a function of the number of simultaneous PGs, while keeping the same frequency of occurrence.
-We show in Figure~\ref{fig:model_results}~(a) that the accuracy of finding the right PG is still above $80\%$ accuracy with more than $1364$ overlapping PGs.
-Moreover, we show in Figure~\ref{fig:model_results}~(b) that (with $M=144$ PGs fixed) the accuracy increases notably as the temporal depth $D$ of the PG kernel increased, demonstrating quantitatively the computational advantage of using heterogeneous delays.
-These results were obtained while assuming that we know $W$.
-However, this is in general not the case, for instance when observing the raster plot of a population of neurons.
-In the following, we will define a generic visual task and determine a learning algorithm to solve that challenging task.
-%Inspired by the k-means algorithm, it is possible to devise a self-supervised learning algorithm.
-Our preliminary results show that it is possible to retrieve PGs embedded in the data, yet that further analysis is necessary to improve the convergence of the algorithm.
-In particular, it seems promising to use a sparseness constraint in the inference mechanism such as to remove spurious correlations in the inference.
-%
-% \subsection{Description of the model}
 %%%-----------------------------------------------------------------
@@
     \centering
-    \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}%png}% https://www.overleaf.com/5625872443qpcwrkssgbsf
-  \caption{\textbf{Core mechanism of the HD-SNN model.} \textit{(Left)}~In this example inspired by the polychronization model~\citep{izhikevich_polychronization_2006}, three pre-synaptic neurons denoted~\textit{b},~\textit{c} and~\textit{d} are fully connected to two post-synaptic neurons~\textit{a} and~\textit{e}, with different delays of respectively $1$, $5$, and $9~\ms$ for~\textit{a} and $8$, $5$, and $1~\ms$ for~\textit{e} (in this schematics, we consider unitary synaptic weights).
-\textit{(Middle)}~If three synchronous pulses are emitted from pre-synaptic neurons, this will generate post-synaptic potentials that will reach \textit{a} and \textit{e} asynchronously because of the heterogeneous delays, and they may not be sufficient to reach the membrane threshold in either of the post-synaptic neurons, therefore no spike will be emitted as this is not sufficient to reach the membrane threshold of the post synaptic neuron, so no output spike is emitted.
-  %at these different delays, and these may not be sufficient to generate a spike in either neuron.
-  \textit{(Right)}~If the pulses are emitted from pre-synaptic neurons such that, taking into account the delays, they reach the post-synaptic neuron~\textit{a} at the same time (here, at $t=10~\ms$), the post-synaptic potentials evoked by the three pre-synaptic neurons sum up, causing the voltage threshold to be crossed and thus to the emission of an output spike (red color), while none is emitted from the post-synaptic neuron~\textit{e}.
-Figure reproduced from~\cite{grimaldi_precise_2023} under a CC-BY licence.}%
+    \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}
+  \caption{\textbf{Core mechanism of the HD-SNN model.} \textit{(Top-Left)}~Two spiking motifs are emitted by four presynaptic neurons.
+Once integrated by the synapses of the postsynaptic neuron~\textit{(Right)}, the spiking motifs are shifted in time by the synaptic delays and weighted by the synaptic weights~\textit{(Middle-Left)}.
+When they reach the soma of the postsynaptic neuron, the different spikes contribute to a modification of its membrane potential according to an activation function.
+In this example, we use the same activation function as a Leaky Integrate and Fire neuron~\textit{(Bottom-Left)}.
+The first spiking motif is synchronized by the synaptic delays and leads to a sudden increase of the membrane potential of the postsynaptic neuron.
+An output spike is emitted when reaching the threshold.
+  \textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses).
+}%
     \label{fig:izhikevich}%
@@
 It results in a weighted event stream for each  post-synaptic neuron $\postsynaddr$:
+
 \begin{equation}\label{eq:stream_b}
@@
 \end{equation}
+
 In biology, this new stream of events is naturally ordered in time as events reach the soma of post-synaptic neurons.
@@
 Crucially, when post-synaptic neurons are activated on their soma by a specific spatio-temporal motif which is imprinted in the set of synapses, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner while the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions (see Figure~\ref{fig:izhikevich}).
-We defined the HD-SNN model in this subsection in all generality and in the next subsection we describe an implementation of our model adapted for the motion detection task.
+We defined the HD-SNN model in this subsection in all generality, and in the next subsection, we describe an implementation of our model adapted for the motion detection task.
 %
@@
 \end{equation}
+
 %The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$: equation~\eqref{eq:stream_b}

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 This approach first defines a finite set of possible 2D movements in polar coordinates.
-Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $8$ linearly spaced movement directions \note{and 3 geometrically spaced speeds}.
-Next, we define a gaze trajectory as segments whose duration is drawn from a Poisson distribution with a mean block length of $20~\ms$, similar to a Lévy flight~\citep[p.
+Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $8$ linearly spaced movement directions and 3 geometrically spaced speeds.
+Next, we define a gaze trajectory as segments whose duration is drawn from a Poisson distribution with a mean block length of $24~\ms$, similar to a Lévy flight~\citep[p.
 289]{mandelbrot_fractal_1982}.
@@
     \caption{
-        {\bf - Motion Detection Task.} {\bf (Left)}~We use large natural images ($256\times256$) in which an aperture ($128\times128$) extracts a cropped image around the view axis.
+        {\bf - Motion Detection Task.} {\bf (Left)}~We use large natural images ($256\times256$) in which an aperture ($64\times64$) extracts a cropped image around the view axis.
 To mimic the effect of a saccadic eye movement, the view axis moves according to a step-wise random walk.
-We show an example path of the trajectory for $200$ time steps (green line).
+We show an example path of the trajectory for $128$ time steps (green line).
 {\bf (Right)}~Snapshots of the synthetic event stream at different time steps.
@@
 This process is known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $256 \times 256$ in size, and we will extract sub-images of size $128 \times 128$ positioned around the center of gaze at each time step.
-We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 200~\ms$.
+These images are $256 \times 256$ in size, and we will extract sub-images of size $64 \times 64$ positioned around the center of gaze at each time step.
+We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 128~\ms$.
 To avoid boundary effects, we will randomly position the full trajectory in image space such that the subimage is translated using the position given by the trajectory at each time step without touching the borders.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Finally, the trajectory is integrated, assuming first that the velocities are sampled uniformly and independently from the set of different motion sets, and second that the motion is uniform during a time segment.
-The resulting instances yield trajectories that are qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-(left)).
+The resulting instances yield trajectories that are qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
 
@@
     \caption{
-        {\bf - Motion Detection Task.} {\bf (Left)}~We use large natural images ($256\times256$) in which an aperture ($64\times64$) extracts a cropped image around the view axis.
+        {\bf - Motion Detection Task.} \textit{(Left)}~We use large natural images ($256\times256$) in which an aperture ($64\times64$) extracts a cropped image around the view axis.
 To mimic the effect of a saccadic eye movement, the view axis moves according to a step-wise random walk.
 We show an example path of the trajectory for $128$ time steps (green line).
-{\bf (Right)}~Snapshots of the synthetic event stream at different time steps.
+\textit{(Right)}~Snapshots of the synthetic event stream at different time steps.
 The dynamics of the sub-image as a function of time produces a naturalistic movie.
@@
 This signed threshold value is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion_task}-(Right)).
+When applied to the whole movie, the event stream is similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion_task}-\textit{(Right)}).
 The goal here is to infer the correct motion solely by observing these events.
@@
 In a neurobiological recording or in the sensory signal obtained from an event-based camera, the input consists of a stream of \emph{spikes} or \emph{events}.
-This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events.
+This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{(Top-Left} for an illustration).
 Events are typically ordered by their time of occurrence.
@@
 
-In the HD-SNN model, each neuron $\postsynaddr \in \postsynaddrspace$ connects to pre-synaptic afferents from $\presynaddrspace$.
+In the HD-SNN model, each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferents from $\presynaddrspace$.
 In biology, a single cortical neuron has generally several thousands of synapses.
-Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the pre-synaptic neuron's soma to that of the post-synaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a pre-synaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
 Note that a neuron may contact an afferent neuron with multiple different delays.
 Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
-as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a pre-synaptic address $\presynaddr_\ranksyn$, a post-synaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
-This defines the full connectivity of the HD-SNN model.
+as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
+This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{(Right)} for an illustration).
 
-Of interest is to define the receptive field of a post-synaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$, or the emitting field of a pre-synaptic neuron $\synapse_\presynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$.
-As a consequence, a post-synaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field.
-It results in a weighted event stream for each  post-synaptic neuron $\postsynaddr$:
-
+Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$, or the emitting field of a presynaptic neuron $\synapse_\presynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$.
+As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field.
+It results in a weighted event stream (\seeFig{izhikevich}-\textit{(Middle-Left}) for each  postsynaptic neuron $\postsynaddr$:
+%
 \begin{equation}\label{eq:stream_b}
@@
 \end{equation}
-
-In biology, this new stream of events is naturally ordered in time as events reach the soma of post-synaptic neurons.
+%
+In biology, this new stream of events is naturally ordered in time as events reach the soma of postsynaptic neurons.
 However, it should be properly reordered in simulations.
-Crucially, when post-synaptic neurons are activated on their soma by a specific spatio-temporal motif which is imprinted in the set of synapses, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner while the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions (see Figure~\ref{fig:izhikevich}).
+Crucially, when postsynaptic neurons are activated on their soma by a specific spatio-temporal motif which is imprinted in the set of synapses, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner while the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions (\seeFig{izhikevich}).
 We defined the HD-SNN model in this subsection in all generality, and in the next subsection, we describe an implementation of our model adapted for the motion detection task.
@@
 Using this dense representation, the processing of $A$ by the model can be written as:
-
+%
 \begin{equation}\label{eq:kernel_b}
@@
 \end{equation}
-
+%
 %The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$: equation~\eqref{eq:stream_b}
@@
 %\end{equation}
-This shows that $\mathcal{C}_\class$ is the result of a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (see Figure~\ref{fig:model} for an illustration).
+This shows that $\mathcal{C}_\class$ is the result of a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration).
 Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
@@
 Formally, using the kernels, it transforms the input raster plot into a probability with the following formula:
+%
 \begin{equation}\label{eq:mlr}
@@
 \end{equation} 
+%
 where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function and  $\bias_\class$ is the bias linked to class $\class$. 
@@
 To avoid introducing biases in the directions, we apply a circular mask on the spatial dimensions of the kernels.
-And because the synaptic delay is linked, by physical principles, to the distance between the pre-synaptic neuron and the postsynaptic one, we apply a linear mask on the temporal axis such that the connectivity depends on the delay: the lower the delay, the lower the circular mask radius (see Figure~\ref{fig:kernels} for an illustration).
+And because the synaptic delay is linked, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one, we apply a linear mask on the temporal axis such that the connectivity depends on the delay: the lower the delay, the lower the circular mask radius (\seeFig{fig:kernels} for an illustration).
 The second heuristics is based on the observation that biological neurons have a sparse spiking activity.
@@
 Because of the symmetry observed between the ON and OFF event streams, we observed that kernels are very similar for the ON polarities.
-These weights are associated to a specific delay on the \textit{delays} axis and to a pre-synaptic address defined on the two other axes.
+These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
 %	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.
@@
 If one focuses on the interpretation of these kernels in terms of spatio-temporal motifs embedded in the event stream, it can lead to interesting outcomes.
-In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a pre-synaptic activity associated to the different motion directions.
+In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions.
 Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spike motif.
@@
 events away in the past.
-For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Figure~\ref{fig:kernels}.
+For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Fig.~\ref{fig:kernels}.
 
@@
 To assess the robustness of the classification as a function of the computational load, we will prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold.
-In Figure~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations, or active weights, per decision for each neuron of the layer.
+In Fig.~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations, or active weights, per decision for each neuron of the layer.
 As a comparison and to account for the gain in performance by using heterogeneous delays, we provide the accuracy obtained with a MLR model using 2D time surface (in red) as in~\citep{grimaldi_robust_2022}.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 To benefit from invariance to location observed in images and exploited in Convolutional Neural Networks, one may further assume that synaptic motifs should be similar across different positions such that we may define a spatio-temporal convolution operator.
-% The characteristics of the kernels are shared among one class of neurons such that $\forall i,j \in [1, N_\class]^2, \; \kernel^i_\class = \kernel^j_\class$.
+%The characteristics of the kernels are shared among one class of neurons such that $\forall i,j \in [1, N_\class]^2, \; \kernel^i_\class = \kernel^j_\class$.
 
@@
 In particular, the whole synaptic set can be represented as one kernel for each class $\class$ of the supervision task as the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
-To keep an analogy with the HD-SNN model, for each neuron $\postsynaddr$ of position $(x, y)$ and class $\class$, $\kernel_\class$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$.
+To keep an analogy with the HD-SNN model, for neuron $\postsynaddr$ of position $(x_\postsynaddr, y_\postsynaddr)$ and class $\class$, $\kernel_\class$ gives the weights as a function of relative addresses and synaptic delays: $\forall {\delta_x \in [1, \Kx], \delta_y \in [1, \Ky], \delta_\timev \in [1, \Ktime], \polev \in [1, \Npol]}, \; \kernel_\class(\polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$.
+The connectivity of neuron $\postsynaddr$ is defined locally, around it position $(x_\postsynaddr, y_\postsynaddr)$.
 %Therefore, if we make the approximation that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train as input can be formalized by a 3D convolution operation.
@@
 %
-%The computational brick used in equation~\eqref{eq:kernel_b} can be extended to every position, in space and time, of the boolean matrix $A$: equation~\eqref{eq:stream_b}
-%\begin{equation}\label{eq:conv}
-%\forall \presynaddr \in \presynaddrspace, \forall \timev \in T, \; \;
-%\mathcal{C}_\class(\presynaddr,\timev)
-%= \sum_{\ranksyn  \in [1, N_{\synapse^\postsynaddr}]} \kernel_\class(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) \cdot A(\presynaddr, \timev-\synapticdelay_\ranksyn)
-%\end{equation}
+where $\delta_x$ and $\delta_y$ are the relative adresses of the synapses inside a kernel and $\delta_\timev$ is the synaptic delay. 
 This shows that $\mathcal{C}_\class$ is the result of a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration).
@@
 Here, we keep the analogy with spiking neurons, and we try to use as few heuristics as possible to observe the detection of spiking motifs using the spatio-temporal kernels.
-If so, this precise spatio-temporal patterns prove to be of interest for neural computations and one can assume that biological neurons make use of this information as well.
+If so, this precise spatio-temporal patterns prove to be of interest for neural computations and one can infer that biological neurons make use of this information as well.
 % \note{plus de biblio sur la motion detection/segmentation ou optique flow avec les events? (reviewer 2)}
 The 3D convolution represents the linear integration of the spike train as the linear input to the neuron.
-Then, the activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Take-All network~\citep{nessler_bayesian_2013}.
+Then, the activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Takes-All network~\citep{nessler_bayesian_2013}.
 In our MLR model, a probability value for each class (i.e.
@@
 Note that the general framework is the same as the one presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel heuristics, inspired by neuroscience, are the main methodological differences with that previous study.
-In our simulations, We set the size of the kernels to $(9, 17, 17)$ and define as many classes as the number of motions (directions and speeds): $N_\class = 8 \times 3$.
+In our simulations, We set the size of the kernels to $(21, 17, 17)$ and define as many classes as the number of motions (directions and speeds): $N_\class = 8 \times 3$.
 To avoid introducing biases in the directions, we apply a circular mask on the spatial dimensions of the kernels.
-And because the synaptic delay is linked, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one, we apply a linear mask on the temporal axis such that the connectivity depends on the delay: the lower the delay, the lower the circular mask radius (\seeFig{fig:kernels} for an illustration).
+And because the synaptic delay is linked, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one, we apply a linear mask on the temporal axis such that the connectivity depends on the delay: the lower the delay, the lower the circular mask radius (\seeFig{kernels} for an illustration).
 The second heuristics is based on the observation that biological neurons have a sparse spiking activity.
@@
 }
-    \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot here as a raster plot a slice of the input event stream (showing ON spikes in red and OFF spikes in blue for each vertical position and time for a given horizontal position at the center of the sensor), with a pure translation motion upwards.
+    \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot here as a raster plot a 2D representation of the input event stream (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time), with a pure translation motion upwards.
 A spatio-temporal convolution is applied to the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 \note{describe the heuristics inspired by neuroscience, remaining complex cells}
+\note{describe how we compute accuracy}
 
@@
 }
-    \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot here as a raster plot a 2D representation of the input event stream (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time), with a pure translation motion upwards.
-A spatio-temporal convolution is applied to the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis.
+    \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot here as a raster plot a 2D representation of the input event stream (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time).
+A spatio-temporal convolution is applied to the dense representation of the input with $2$ distinct convolutional kernels (the green and orange kernels) that will define the output.
 The convolution is summed over the two polarities.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 \newcommand{\DataAvailability}{
-This works is made reproducible using the following tools. First the code reproducing all figures is available on~\href{ https://github.com/SpikeAI/pyTERtorch}{GitHub}. % todo: open-up the repo
+This works is made reproducible using the following tools.
+First the code reproducing all figures is available on~\href{ https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}{GitHub}.
+% todo: open-up the repo
 Find also the associated~\href{https://www.zotero.org/groups/4776796/fastmotiondetection}{zotero group} used to gather relevant literature on the subject.
@@
     \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot here as a raster plot a 2D representation of the input event stream (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time).
-A spatio-temporal convolution is applied to the dense representation of the input with $2$ distinct convolutional kernels (the green and orange kernels) that will define the output.
+A spatio-temporal convolution is applied to the dense representation of the input with $2$ distinct convolutional kernels (the green and orange kernels) that will define the output channels.
 The convolution is summed over the two polarities.
+If you have two axes $X$ and $Y$ to represent the presynaptic addresses as for the pixel grid of a DVS, you obtain a 3D convolution.
 At any given time (represented by the blue point), we show the extent of the kernels (green dashed frame).

### 2023-02-20 cover_letter/2023-02-20_coverletter.tex
cover_letter

--- 2023-02-09 cover_letter/2023-02-10_coverletter.tex
@@
 %\centerline{
-\includegraphics[width=.4\textwidth]{~/Documents/troislogos.png}
+\includegraphics[width=.4\textwidth]{troislogos.png}
 %}
@@
 Marseille, 
-February 10, 2023%\today
+% February 20th, 2023%
+\today
 \\[12pt] % Date
@@
 Please find enclosed the revised manuscript entitled ``Learning heterogeneous delays in a layer of spiking neurons for fast motion detection'' for your consideration as an article in \emph{Biological Cybernetics}.
-All the authors have been involved with the work, have approved the manuscript and agreed to its revision.
+Both authors have been involved with the work, have approved the manuscript and agreed to its revision.
 The original submission ID is a7f45f8e-87d6-4a30-ac7f-9f0c94e2592a.
 
-Our manuscript describes a novel method inspired by neuroscience to overcome some challenges faced in computer vision, notably when dealing with large amounts of data.
+Our manuscript describes a novel method inspired by neuroscience to overcome some challenges faced in computer vision, notably when dealing with large amounts of data, by using an event-based representation.
 By using heterogeneous delays on different synapses, this novel spiking neuron method is able to detect spiking motifs and we validate the method on synthetic event-based data.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex

@@
 If you have two axes $X$ and $Y$ to represent the presynaptic addresses as for the pixel grid of a DVS, you obtain a 3D convolution.
-At any given time (represented by the blue point), we show the extent of the kernels (green dashed frame).
-\textit{(Middle)}~There are as many kernels as there are possible motion classes in the recording, and we show the activation resulting from the convolution for this particular motion.
-The activation can be above chance (warm colors) or below chance (cold colors).
-For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the nonlinearity of the MLR model (i.e., the softmax function).
-The output of the MLR gives a probability for each class associated with a particular kernel.
-\textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated to a thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated to a specific motion class (here red is associated to an upward motion).
-Most of the motion is detected accurately.
-}
+We restrict the illustration to a 2D representation and to 2 possible classes (green and orange) that are associated with different motion directions.
+\textit{(Middle)}~For each position (address, time), one can compute the activation resulting from the convolution.
+The output of the convolution is processed by the nonlinearity of the MLR model (i.e., the softmax function).
+The output of the MLR gives a probability for each class associated with a particular kernel (colored bars in the highlighted pixel).
+\textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated to a thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated to a specific motion class.
+Note that the position of the output spikes does not systematically correspond to the position of the input spikes but only when enough evidence is obtained.}
     \label{fig:model}

### 2023-02-20 cover_letter/2023-02-20_coverletter.tex
cover_letter

@@
-%!TeX TS-program = lualatex
-%!TeX encoding = UTF-8 Unicode
-%!TeX spellcheck = fr-FR
-%!BIB TS-program = biber
+% !TEX encoding = UTF-8 Unicode
+% !TeX TS-program = pdfLaTeX
+% !TeX spellcheck = en-US
+% !BIB TS-program = bibtex
 % -*- coding: UTF-8; -*-
@@
 Results show that this method could provide a path for future spiking neural network algorithms using less energy for a similar performance as their analog counterparts.
-This work is the extension of a previous study on a simpler event-based dataset.
+This work is the extension of a previous study on a simpler event-based dataset with a simplified model.
 Our contribution here is to offer a better formal description of the model, an application on a more ecological task and testing of the influence of various parameters of the validation dataset.
@@
 
-\textbf{Reviewer 1}
+In the following we will give response to reviewers in \noteRev{framed boxes.}
 
- 1.There is a lack of an intuitive figure to show the HD-SNN model used in this paper.
+\textbf{Reviewer \#1}
+
+ \noteRev{Overall, we would like to thank Reviewer 1 for its comments that allowed us to focus on reshaping the manuscript and give clearer and shorter explanation of the work we submit in your journal.
+Because we did major modifications in the manuscript, we do not provide a tracked changes version of the revision but we can easily provide one if needed.
+}
+
+\begin{quote}
+ 1. There is a lack of an intuitive figure to show the HD-SNN model used in this paper.
+\end{quote}
 
@@
 
- 2.The MLR model used in the experimental part was presented in authors’ another work. Therefore, the changes on the MLR model are suggested to be highlighted and contrasted graphically.
+\begin{quote}
+ 2. The MLR model used in the experimental part was presented in authors’ another work. Therefore, the changes on the MLR model are suggested to be highlighted and contrasted graphically.
+\end{quote}
 
@@
 
- 3.The title of section 2.2.1 is ’’a generative model for rater plots’’, but actually this section introduces basic terms and symbols.
+\begin{quote}
+ 3.
+The title of section 2.2.1 is ’’a generative model for rater plots’’, but actually this section introduces basic terms and symbols.
 Section 2.2.2 is named ‘’detecting spiking motifs’’, but most of the content seems to introduce the generation process of raster plots, while only the last paragraph introduces the detection of motif.
 It is recommended that the content of these two sections be reorganized and that references to Figure 1(b)(c)(d) be added where appropriate during the description.
+\end{quote}
 
- \noteRev{We removed these sections to have a more concise explanation of the model.}
+ \noteRev{We removed these sections to have a more concise explanation of the model and focused on resolving the motion detection task.
+We hope the manuscript is now more concise and clear to the point.}
 
- 4.The adjustment of weights and delays needs more detailed elaboration, preferably given in mathematical formulas.
+\begin{quote}
+ 4. The adjustment of weights and delays needs more detailed elaboration, preferably given in mathematical formulas.
+\end{quote}
 
@@
 
- 5.In Figure 4, it is claimed that the kernels are very similar for the ON polarities, and different kernels are selective to the different motion directions.
+\begin{quote}
+ 5.
+In Figure 4, it is claimed that the kernels are very similar for the ON polarities, and different kernels are selective to the different motion directions.
 However, in terms of these two observations, I think they are not very significant, so I suggest a more intuitive way to reflect them, such as quantifying the similarity between the ON kernels, and the angle between the kernels and the vector of the standard motion directions.
+\end{quote}
 
@@
 
- 6.There are few experiments and lack of comparative experiments.
+\begin{quote}
+ 6.
+There are few experiments and lack of comparative experiments.
 Since the model has an improvement based on MLR model, it is suggested to increase the comparison with MLR model and HOTS model on other datasets, such as N-MNIST and other neuromorphic datasets, so as to show the advantages of the proposed model in computational consumption.
+\end{quote}
 
@@
 
- 7.Some sections are verbose, such as the discussion.
+\begin{quote}
+ 7. Some sections are verbose, such as the discussion.
+\end{quote}
 
@@
 
- 8.There are some minor errors in the text, and it is recommended to check the full text: 
+\begin{quote}
+ 8. There are some minor errors in the text, and it is recommended to check the full text: 
  1)In the first paragraph on page 8, "a presynaptic address $b_s$" should be "a postsynaptic address $b_s$"; 
@@
  5)The content of paragraph 1 on page 13 is repeated. 
+\end{quote}
 
@@
 
- \noteRev{Overall, we thank Reviewer 1 for its comments that allowed us to focus on reshaping the manuscript and give clearer and shorter explanation of the work we submit in your journal.
-Because we did major modifications in the manuscript, we do not provide a tracked changes version of the revision but we can easily provide one if needed.
-}
+\textbf{Reviewer \#2}
 
-\textbf{Reviewer 2}
+\begin{quote}
+ {\it Key results: Please summarize what you consider to be the outstanding features of the work.}
+ Using a single layer of spiking neurons with heterogeneous delays to learn spatio-temporal  spiking motifs
+\end{quote} 
+\noteRev{We thank Reviewer 2 for its comments. We have thoroughly reworked the manuscript to take them into account.}
 
- Key results: Please summarize what you consider to be the outstanding features of the work.
- Using a single layer of spiking neurons with heterogeneous delays to learn spatio-temporal  spiking motifs
-
- Validity: Does the manuscript have flaws which should prohibit its publication? If so, please provide details.
-
+\begin{quote}
  The work presented here is very similar to the author's other work,  "Learning hetero-synaptic delays for motion detection in a single layer of spiking neurons" 10.1109/ICIP46576.2022.9897394.
@@
  There are also a number of citations that could have been added that did appear in the other publication, while also including other motion detection/ segmentation or previous work on event-based optical flow, which is a very similar problem.
+\end{quote}
 
- Originality and significance: If the conclusions are not original, please provide relevant references.
-
- See above...
-
-\noteRev{We thank Reviewer 2 for its comments and indeed this previous work was not mentioned in the original manuscript.
-As mentioned in the cover letter and in the previous responses, we add this citation and describe the main differences with the submitted study: ecological task (section 2.1), more complete formulation of the model (section 2.2) and list of the priors used in this particular model (P8, L269).
+\noteRev{Indeed this previous work was not mentioned in the original manuscript as this 4-pages paper was not yet published in the proceedings of the conference in which it was presented.
+As mentioned in this cover letter and in the previous responses to reviewer \#1, we now added this citation~\cite{grimaldi_2022} and describe the main differences with the submitted study: the use of a more ecological task (section 2.1), a complete formulation of the model (section 2.2) and the ameliroations provided in this particular model (P8, L269).
 We also cite studies that use similar methods (P7, L243).
+We also cover more extensive results, notably by testing the learned model with parameterized textures.
 }
 
+\bibliographystyle{apalike}
+\bibliography{../FastMotionDetection}
+
 \end{document}

### 2023-02-20 cover_letter/2023-02-20_coverletter.tex
cover_letter

@@
 \end{quote} 
-\noteRev{We thank Reviewer 2 for its comments. We have thoroughly reworked the manuscript to take them into account.}
+\noteRev{We would like to thank reviewer \#2 for his comments. We have thoroughly revised the manuscript to take them into account.}
 
@@
 
-\noteRev{Indeed this previous work was not mentioned in the original manuscript as this 4-pages paper was not yet published in the proceedings of the conference in which it was presented.
-As mentioned in this cover letter and in the previous responses to reviewer \#1, we now added this citation~\cite{grimaldi_2022} and describe the main differences with the submitted study: the use of a more ecological task (section 2.1), a complete formulation of the model (section 2.2) and the ameliroations provided in this particular model (P8, L269).
-We also cite studies that use similar methods (P7, L243).
-We also cover more extensive results, notably by testing the learned model with parameterized textures.
-}
+\noteRev{In fact, there was no mention of this previous work in the original manuscript, as this 4-page paper had not yet been published in the proceedings of the conference at which it had been presented.
+As mentioned in the responses to reviewer \#1 in this cover letter, we have now added this citation~\cite{grimaldi_learning_2022}.
+In addition, in this submitted manuscript we describe the main differences from that previous study: the use of a more ecological task (Section 2.1), a complete formulation of the model (Section 2.2), and the improvements made in this particular model (P8, L269).
+We also provide citations to studies using similar methods (P7, L243).
+Further results are presented, especially the testing of the learned model with parametric textures.}
 

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %%%-----------------------------------------------------------------
-Our objective in this paper is to test whether the HD-SNN model is capable to efficiently learn a motion detection task as defined by a realistic event-based data stream.
-This type of signal is typically recorded by a Dynamic Vision Sensor (DVS) and is inspired by the signal sent by retinal ganglion cells through the optic nerve.
-Events are binary and should be sufficient to perform the motion detection task quickly and effectively.
-Here, we will first describe the definition of the task and then of the HD-SNN model we use to infer motion.
+In this paper, we aim to test whether the HD-SNN model is capable of efficiently learning a motion detection task, which is defined by a realistic event-based data stream.
+This type of signal is typically captured by a Dynamic Vision Sensor (DVS) and is inspired by the signal that is sent from the retinal ganglion cells through the optic nerve.
+The events are binary in nature and should be sufficient to perform the task of motion detection in a fast and efficient manner.
+Here we will first describe the task definition and then the HD-SNN model we use for motion inference and how we train it.
 %%%-----------------------------------------------------------------
@@
 To mimic the effect of a saccadic eye movement, the view axis moves according to a step-wise random walk.
-We show an example path of the trajectory for $128$ time steps (green line).
-\textit{(Right)}~Snapshots of the synthetic event stream at different time steps.
-The dynamics of the sub-image as a function of time produces a naturalistic movie.
-This can be transformed into an event-based representation.
+We show such a trajectory with a length of $128$ time steps (green line).
+\textit{(Right)}~Snapshots of the synthetic event stream at different time steps (start and end frames denoted by a white and black dot, respectively).
+The dynamics of the cropped image translated according to the trajectory as a function of time produces a naturalistic movie which is then transformed into an event-based representation.
 Mimicking the retina, this representation encodes proportional increases or decreases in luminance, i.e.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 ON (red) and OFF (blue) events, in each pixel of the image.
-In the lower left corner of the snapshots, the translation vector is shown in red as one of the possible classes of motion.
-Note the change in the direction of motion between the second and third image and that due to the aperture problem, contours parallel to the motion emit relatively less spikes.}
+In the lower left corner of the snapshots, the translation vector is shown in red as one of the possible classes of motions.
+Note the change in the direction of motion between the third and fourth image, and also that due to the aperture problem, contours parallel to the motion emit relatively less spikes.}
     \label{fig:motion_task}

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
@@
         {\bf - Motion Detection Task.} \textit{(Left)}~We use large natural images ($256\times256$) in which an aperture ($64\times64$) extracts a cropped image around the view axis.
-To mimic the effect of a saccadic eye movement, the view axis moves according to a step-wise random walk.
+To mimic the effect of a saccadic eye movement, the view axis moves according to a stepwise random walk.
 We show such a trajectory with a length of $128$ time steps (green line).
-\textit{(Right)}~Snapshots of the synthetic event stream at different time steps (start and end frames denoted by a white and black dot, respectively).
-The dynamics of the cropped image translated according to the trajectory as a function of time produces a naturalistic movie which is then transformed into an event-based representation.
+\textit{(Right)}~Snapshots of the synthetic event stream at different time steps (start and end frames marked by a white and black dot, respectively).
+The dynamics of the cropped image translated according to the trajectory as a function of time produces a naturalistic movie, which is then transformed into an event-based representation.
 Mimicking the retina, this representation encodes proportional increases or decreases in luminance, i.e.
@@
 In the lower left corner of the snapshots, the translation vector is shown in red as one of the possible classes of motion.
-Note the change in the direction of motion between the second and third image and that due to the aperture problem, contours parallel to the motion emit relatively less spikes.}
+Note the change in the direction of motion between the third and fourth image, and also that, due to the aperture problem, contours parallel to the motion emit relatively fewer spikes.}
     \label{fig:motion_task}

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 
-To transform each movie into events as the ones recorded by a DVS, we compute a gradient image (initialized at zero) by adding the gradient of the pixels' intensity over two successive frames.
+To transform each movie into events as the ones recorded by a DVS, we compute a gradient image (initialized at zero) by computing the temporal gradient of the pixels' intensity over two successive frames.
 For a specific pixel and timestamp, an event is generated if the absolute value of this gradient exceeds a threshold.
@@
 The first spiking motif is synchronized by the synaptic delays and leads to a sudden increase of the membrane potential of the postsynaptic neuron.
-An output spike is emitted when reaching the threshold.
+An output spike is emitted when the membrane potential reaches the threshold, and it is then resetted.
 \textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses).

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
     \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}
-    \caption{\textbf{Core mechanism of the HD-SNN model.} \textit{(Top-Left)}~Two spiking motifs are emitted by four presynaptic neurons.
+    \caption{\textbf{The core mechanism of the HD SNN model.} \textit{(Top-Left)}~Two spiking motifs are emitted from four presynaptic neurons.
 Once integrated by the synapses of the postsynaptic neuron~\textit{(Right)}, the spiking motifs are shifted in time by the synaptic delays and weighted by the synaptic weights~\textit{(Middle-Left)}.
 When they reach the soma of the postsynaptic neuron, the different spikes contribute to a modification of its membrane potential according to an activation function.
-In this example, we use the same activation function as a Leaky Integrate and Fire neuron~\textit{(Bottom-Left)}.
-The first spiking motif is synchronized by the synaptic delays and leads to a sudden increase of the membrane potential of the postsynaptic neuron.
-An output spike is emitted when the membrane potential reaches the threshold, and it is then resetted.
+In this example we use the same activation function as for a Leaky Integrate and Fire neuron~\textit{(Bottom-Left)}.
+The first spiking motif is synchronized by the synaptic delays and causes a sudden rise in the membrane potential of the postsynaptic neuron.
+An output spike is emitted when the membrane potential reaches the threshold and it is then reset.
 \textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses).
@@
 
-In the HD-SNN model, each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferents from $\presynaddrspace$.
+In the HD-SNN model, each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferent neurons from $\presynaddrspace$.
 In biology, a single cortical neuron has generally several thousands of synapses.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %
-Such event-based model can be adapted to computer vision tasks.
+Indeed, such an event-based model can be easily adapted to common computer vision tasks.
 In particular, from the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation.
@@
 
-% Here, we will consider that heterogeneous delays are limited in range
 In particular, the whole synaptic set can be represented as one kernel for each class $\class$ of the supervision task as the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
@@
 
-Here, we keep the analogy with spiking neurons, and we try to use as few heuristics as possible to observe the detection of spiking motifs using the spatio-temporal kernels.
+Here, we keep the analogy with spiking neurons and we try to observe the detection of spiking motifs using the spatio-temporal kernels.
 If so, this precise spatio-temporal patterns prove to be of interest for neural computations and one can infer that biological neurons make use of this information as well.
@@
 
-Once the general framework used to solve the ecological task described in section~\ref{sec:task} explained, we can add some heuristics based on neuro-scientific observations to constrain our model and its strategies to solve the task.
+Now that this general framework explained, we can add some heuristics based on neuro-scientific observations to constrain our model and its strategies to solve the task  used to solve the ecological task described in section~\ref{sec:task}.
 Note that the general framework is the same as the one presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel heuristics, inspired by neuroscience, are the main methodological differences with that previous study.
@@
 To avoid introducing biases in the directions, we apply a circular mask on the spatial dimensions of the kernels.
-And because the synaptic delay is linked, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one, we apply a linear mask on the temporal axis such that the connectivity depends on the delay: the lower the delay, the lower the circular mask radius (\seeFig{kernels} for an illustration).
-The second heuristics is based on the observation that biological neurons have a sparse spiking activity.
+First, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both arrangements of polarities.
+This is due to the fact that our whitened images show a symmetry in the luminance profiles, and that an image with inverted contrast is indistinguishable.
+As this arrangement of polarities is independent of the speed, we added a mechanism gathering the linear values for the movie and the movie with ON and OFF cells flipped, and kept only the maximal value for each voxel.
+This is similar to the computation operated for complex cells in the primary visual cortex.
+Second, we included a prior in the motions that could be selected, as there is a prior in natural scenes for slow speeds~\citep{vacher_bayesian_2018}.
+As here we aim at capturing the possible convergence of events' trajectories converging to each voxel, we thus apply a mask on the spatio-temporal kernels such that the connectivity is bounded by the delay: the lower the delay, the lower the circular mask radius (\seeFig{kernels} for an illustration).
+This is compatible with neuroscientific principles because of the limits in conduction delay along horizontal connections, the synaptic delay is linked, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one.
+
+% The second heuristics is based on the observation that biological neurons have a sparse spiking activity.
 From this observation, we set an output firing rate for the layer of spiking neurons in order to make a sparse selection of the events at which the model makes a decision.
@@
 Given a stimulus input, the model thus yields a spike output in the postsynaptic address space.
-
-\note{describe the heuristics inspired by neuroscience, remaining complex cells}
+% \note{describe the heuristics inspired by neuroscience, remaining complex cells}
 \note{describe how we compute accuracy}

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 
-Now that this general framework explained, we can add some heuristics based on neuro-scientific observations to constrain our model and its strategies to solve the task  used to solve the ecological task described in section~\ref{sec:task}.
-Note that the general framework is the same as the one presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel heuristics, inspired by neuroscience, are the main methodological differences with that previous study.
-In our simulations, We set the size of the kernels to $(21, 17, 17)$ and define as many classes as the number of motions (directions and speeds): $N_\class = 8 \times 3$.
-To avoid introducing biases in the directions, we apply a circular mask on the spatial dimensions of the kernels.
-First, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both arrangements of polarities.
-This is due to the fact that our whitened images show a symmetry in the luminance profiles, and that an image with inverted contrast is indistinguishable.
-As this arrangement of polarities is independent of the speed, we added a mechanism gathering the linear values for the movie and the movie with ON and OFF cells flipped, and kept only the maximal value for each voxel.
-This is similar to the computation operated for complex cells in the primary visual cortex.
+Now that this general framework has been explained, we can add some heuristics, based on neuroscientific observations, to constrain our model and its strategies for solving the task that is used to solve the ecological task described in section~\ref{sec:task}.
+Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel heuristics inspired by neuroscience are the main methodological differences with that previous study.
+In our simulations, we set the size of the kernels to $(21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = 8 \times 3$.
+To avoid introducing biases in the directions, we apply a circular mask to the spatial dimensions of the kernels.
+First, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities arrangements.
+This is due to the fact that our whitened images have a symmetry in the luminance profiles, and an image with inverted contrast is indistinguishable.
+Since this arrangement of polarities is independent of speed, we added a mechanism that collects the linear values for the movie and the movie with the ON and OFF cells flipped, keeping only the maximum value for each voxel.
+This is similar to the computation done for complex cells in primary visual cortex.
 Second, we included a prior in the motions that could be selected, as there is a prior in natural scenes for slow speeds~\citep{vacher_bayesian_2018}.
-As here we aim at capturing the possible convergence of events' trajectories converging to each voxel, we thus apply a mask on the spatio-temporal kernels such that the connectivity is bounded by the delay: the lower the delay, the lower the circular mask radius (\seeFig{kernels} for an illustration).
-This is compatible with neuroscientific principles because of the limits in conduction delay along horizontal connections, the synaptic delay is linked, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one.
+Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we apply a mask to the spatiotemporal kernels so that the connectivity is limited by the delay: the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration).
+This is consistent with neuroscientific principles because, due to the limits of the conduction delay along horizontal connections, the synaptic delay is related, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one.
 

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel heuristics inspired by neuroscience are the main methodological differences with that previous study.
-In our simulations, we set the size of the kernels to $(21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = 8 \times 3$.
+First, in our simulations, we set the size of the kernels to $(21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = 8 \times 3$.
 To avoid introducing biases in the directions, we apply a circular mask to the spatial dimensions of the kernels.
-First, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities arrangements.
+Furthermore, we included a prior in the motions that could be selected, as there is a prior in natural scenes for slow speeds~\citep{vacher_bayesian_2018}.
+Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we thus apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration).
+This is consistent with neuroscientific principles because, due to the limits of the conduction delay along horizontal connections, the synaptic delay is related, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one.
+Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities arrangements.
 This is due to the fact that our whitened images have a symmetry in the luminance profiles, and an image with inverted contrast is indistinguishable.
@@
 This is similar to the computation done for complex cells in primary visual cortex.
-Second, we included a prior in the motions that could be selected, as there is a prior in natural scenes for slow speeds~\citep{vacher_bayesian_2018}.
-Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we apply a mask to the spatiotemporal kernels so that the connectivity is limited by the delay: the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration).
-This is consistent with neuroscientific principles because, due to the limits of the conduction delay along horizontal connections, the synaptic delay is related, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one.
 
@@
 % \note{describe the heuristics inspired by neuroscience, remaining complex cells}
-\note{describe how we compute accuracy}
 
@@
 We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
+
+\note{describe how we compute accuracy}
 

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-Note that a neuron may contact an afferent neuron with multiple different delays.
+Note that a neuron may contact the same afferent neuron with different delays through different synaptic connections.
 Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
 as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
-This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{(Right)} for an illustration).
+This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{(Right)} for an illustration of the connectivity of one neuron with synaptic weights and delays).
 

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %
-Indeed, such an event-based model can be easily adapted to common computer vision tasks.
+We can define a specific implementation of this model to adapt it to common computer vision tasks.
 In particular, from the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation.
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{figures/accuracy.pdf}
-    \caption{Accuracy as a function of the number of computation load for the model with heterogeneous delays (blue line) and for a method using 2D time surfaces (red dot)~\citep{grimaldi_robust_2022}.
-The relative computational load (on a log axis) is controlled by changing the percentage of active weights relative to the dense convolution kernel.
-We observe a similar accuracy than HOTS, yet that our model could achieve a similar accuracy with significantly fewer coefficients.}
+    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
+    \caption{Accuracy as a function of the number of computation load for the HD-SNN model (blue dots) with errorbars giving the 5\% - 95\% quantiles  and a sigmoid fit (blue line).
+The relative computational load (on a log axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
+As we prune coefficients, we observe a stable accuracy value, with a drop observed with approximately 25 times less computations.}
     \label{fig:accuracy}

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
     \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
-    \caption{Accuracy as a function of the number of computation load for the HD-SNN model (blue dots) with errorbars giving the 5\% - 95\% quantiles  and a sigmoid fit (blue line).
-The relative computational load (on a log axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
-As we prune coefficients, we observe a stable accuracy value, with a drop observed with approximately 25 times less computations.}
+    \caption{Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
+The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
+As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about 25 times fewer computations.}
     \label{fig:accuracy}

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 % public
-% https://github.com/SpikeAI/pyTERtorch/
+% https://github.com/SpikeAI/pyTERtorch/ (for dev)
+% https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN 
 % https://www.zotero.org/groups/4776796/fastmotiondetection
@@
 % Kexin Chen <kexinc3@uci.edu>
-%
-% 
-% public
-% https://github.com/SpikeAI/pyTERtorch/
-% https://www.zotero.org/groups/4776796/fastmotiondetection
 %
@@
 \newcommand{\DataAvailability}{
-This works is made reproducible using the following tools.
-First the code reproducing all figures is available on~\href{ https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}{GitHub}.
-% todo: open-up the repo
+This works is made reproducible.
+The code reproducing the manuscript and all figures is available on~\href{ https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}{GitHub}.
+It also contains supplementary figures and results.
 Find also the associated~\href{https://www.zotero.org/groups/4776796/fastmotiondetection}{zotero group} used to gather relevant literature on the subject.
@@
 %
+\begin{linenomath*}
 \begin{equation}\label{eq:stream_b}
@@
 \end{equation}
+\end{linenomath*}
 %
@@
 %
-We can define a specific implementation of this model to adapt it to common computer vision tasks.
+It is possible to define a specific implementation of this model in order to adapt it to common tasks in computer vision.
 In particular, from the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation.
-Indeed, by discretization of time, we may transform any event-based input from an event-based camera as a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$, defined for all polarities $p$, times $t$ and space coordinates $x$ and $y$.
-The values are by definition equal to zero except at the occurrences of events: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
-First, one can note that by using a discretization, the computational brick used in equation~\eqref{eq:kernel_b} corresponds to a temporal convolution which transforms the input $A$ using one kernel per postsynaptic neuron~\citep{grimaldi_learning_2022}.
-To benefit from invariance to location observed in images and exploited in Convolutional Neural Networks, one may further assume that synaptic motifs should be similar across different positions such that we may define a spatio-temporal convolution operator.
-%The characteristics of the kernels are shared among one class of neurons such that $\forall i,j \in [1, N_\class]^2, \; \kernel^i_\class = \kernel^j_\class$.
+Indeed, by discretizing time, we can transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.
+The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
+First, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:kernel_b} corresponds to a temporal convolution that transforms the input $A$ using one kernel per postsynaptic neuron~\citep{grimaldi_learning_2022}.
+To take advantage of the position invariance observed in images and exploited in convolutional neural networks, we can further assume that synaptic motifs should be similar across different positions, so we can define a spatio-temporal convolutional operator.
 
+Therefore, if we make the approximation at any given time that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
+The longest synaptic delay defines the depth $\Ktime$ of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented.
 In particular, the whole synaptic set can be represented as one kernel for each class $\class$ of the supervision task as the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
@@
 The connectivity of neuron $\postsynaddr$ is defined locally, around it position $(x_\postsynaddr, y_\postsynaddr)$.
-%Therefore, if we make the approximation that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train as input can be formalized by a 3D convolution operation.
-The longest synaptic delay defines the depth of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented.
-Using this dense representation, the processing of $A$ by the model can be written as:
+Using this dense representation, the processing of $A$ by the model can be written as: %
 %
+\begin{linenomath*}
 \begin{equation}\label{eq:kernel_b}
@@
 \end{equation}
+\end{linenomath*}
 %
-where $\delta_x$ and $\delta_y$ are the relative adresses of the synapses inside a kernel and $\delta_\timev$ is the synaptic delay. 
+where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses inside a kernel and $\delta_\timev$ is the synaptic delay. 
 This shows that $\mathcal{C}_\class$ is the result of a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration).
@@
 This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
-The same type of spatio-temporal filtering is used as a pre-processing stage of a pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
+A similar type of spatio-temporal filtering is used as a pre-processing stage for a pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
 Also,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm which implements a motion estimation task.
@@
 %
+\begin{linenomath*}
 \begin{equation}\label{eq:mlr}
@@
 \end{equation} 
+\end{linenomath*}
 %
@@
 This is similar to the computation done for complex cells in primary visual cortex.
-
-% The second heuristics is based on the observation that biological neurons have a sparse spiking activity.
-From this observation, we set an output firing rate for the layer of spiking neurons in order to make a sparse selection of the events at which the model makes a decision.
-It results in the selection of the coefficients with the maximum likelihood value.
-Therefore, we will only select a given proportion of the most active cells, which we set here to $0.1\%$ of all output voxels \note{change accordingly to the model's parameters}.
-This threshold was found using cross-validation and future work should automate how this may be determined for various datasets.
-Given a stimulus input, the model thus yields a spike output in the postsynaptic address space.
-% \note{describe the heuristics inspired by neuroscience, remaining complex cells}
-
-% \note{describe the parameters of the model}
 
@@
 %
-Since the model is fully differentiable, we can now implement a semi-supervised learning rule.
+Since the model is fully differentiable, we can now implement a supervised learning rule.
 The loss function of the MLR model is the binary cross entropy on the output of the classification layer.
@@
 The labels were defined at each time point as a one-hot encoding of the current motion in the channel corresponding to the current motion for all positions.
-Since in this semi-supervised context the label is known but the timing is not, we used a selection of the temporal support in an unsupervised manner.
-It is important to note that we weighted the cost function by considering only active cells, so that the error is only propagated back to the spatial locations of these most active cells.
-This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}.
+Note that in this context the label is known but the position is not, mainly due to the sparse spatial content of natural images.
+However, the supervised optimization of this multinomial logistic regression model will involve adjusting the kernels' weights.
+As a consequence, the error is only propagated back to the spatial locations of these most active cells.
+This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation.
 Simulations are performed with the \note{PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$)}.
 We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
-
-\note{describe how we compute accuracy}
 
@@
 These calculations are performed on a different input data set than the one used in the training or validation steps.
-The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/pyTERtorch}.
+The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}.
 %
@@
 %
-The code is publicly available online at: \url{https://github.com/SpikeAI/pyTERtorch}.
+\DataAvailability
 %   

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Since the model is fully differentiable, we can now implement a supervised learning rule.
-The loss function of the MLR model is the binary cross entropy on the output of the classification layer.
-Supervision was implemented using the binary input events as defined above and the motion direction labels as the desired output.
+This rule was implemented using the binary input events as inputs and the corresponding motion direction labels as the desired output.
+The loss function of the MLR model is the binary cross entropy at the output of the classification layer.
 The labels were defined at each time point as a one-hot encoding of the current motion in the channel corresponding to the current motion for all positions.
-Note that in this context the label is known but the position is not, mainly due to the sparse spatial content of natural images.
-However, the supervised optimization of this multinomial logistic regression model will involve adjusting the kernels' weights.
-As a consequence, the error is only propagated back to the spatial locations of these most active cells.
+Note that in this context, the label is known but the position is not, mainly due to the sparse spatial content of natural images.
+However, the supervised optimization of this MLR model will involve adjusting the weights of the kernels.
+As a result, the error is only propagated back to the spatial locations of these most active cells.
 This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation.
@@
 We have shown above that if different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities.
-Thus, by taking the logit of the probability of the output given by the model, we can calculate the probability of the output.
-This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error of the velocity estimate given by the mean estimate.
+Thus, by taking the mean logit of the probability of the output given by the model at all positions for any given time, we can calculate the probability of the output at this time.
+This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted).
 These calculations are performed on a different input data set than the one used in the training or validation steps.
@@
     %\vspace{-3cm}
-    \includegraphics[width=\linewidth]{figures/motion_kernels.png}%pdf}
+    \includegraphics[width=\linewidth]{figures/motion_kernels.pdf}
     }

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
     \caption{
-    	Representation of the weights for $3$ of the $32$ different learned kernels of the model as learned on natural scenes.
-Each pair of line correspond to the OFF and ON polarities respectively, with excitatory weights in warm colors.
-Delays are represented in the horizontal axis from right (zero delay) to the left (delay of $11$ steps).
-Because of the symmetry observed between the ON and OFF event streams, we observed that kernels are very similar for the ON polarities.
-These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
+    	Representation of the weights for $8$ directions (among the $8 \times 3$ different kernels of the model) as learned on natural scenes.
+Directions are shown as a red arrow on the right.
+The spatio-temporal kernels are represented as a sequence of images at different delays according to the horizontal axis from right (zero delay) to the left (delay of $11$ steps).
+Each image correspond to the weights at a given delay with excitatory and inhibitory weights in warm and cold colors, respectively.
+Because of the symmetry observed between the ON and OFF event streams, we observed that kernels are very similar for the ON polarities, and these are not shown here.
+% These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
 %	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
@@
     \caption{
-    	Representation of the weights for $3$ of the $32$ different learned kernels of the model as learned on natural scenes.
-Each pair of line correspond to the OFF and ON polarities respectively, with excitatory weights in warm colors.
-Delays are represented in the horizontal axis from right (zero delay) to the left (delay of $11$ steps).
-Because of the symmetry observed between the ON and OFF event streams, we observed that kernels are very similar for the ON polarities.
-These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
+    	Representation of the weights for $8$ directions for one speed (among the $8 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
+Directions are shown as red arrows on the left.
+The spatio-temporal kernels are represented as a sequence of images at different delays along the horizontal axis from the far right (delay of one step) to the left (delay of $12$ steps).
+Each image corresponds to the weights at a given delay, with excitatory and inhibitory weights in warm and cold colors, respectively.
+Due to the observed symmetry between the ON and OFF event streams, we observed that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective to the different motion directions and we observe a slight level of orientation selectivity perpendicular to the orientation.
+%      % These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
 %	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.
-	Different kernels are selective to the different motion directions and we observe some level of orientation selectivity, where ON and OFF subfields organized in a push-pull organization. 
-%        One sees positive (excitatory) coefficients for the specific direction of motion but also negative (inhibitory) coefficients for other directions.
+One sees positive (excitatory) coefficients for the specific direction of motion but also negative (inhibitory) coefficients for other directions.
 	}

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %%%-----------------------------------------------------------------
-To test our model, we will quantify its ability to categorize different motions.
+To test our model, we will quantify its ability to categorize different motions, that is on event streams for which the ground truth motion is known at every instant in time.
 Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli.
@@
     \centering
-    \includegraphics[width=0.99\linewidth]{figures/2022-09-27_MotionDetection_motion_clouds.png}%motion_clouds.pdf} % 2022-09-27_MotionDetection_motion_clouds.png}%
-    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of {\bf (a)} the mean spatial frequency, {\bf (b)} the bandwidth in orientation: from a grating-like (left) to isotropic textures (right)), {\bf (c)} the bandwidth in speed, from a rigid motion (left) to independent frames (right).
-Note that these accuracy is computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion (no aperture) or arbitrary (aperture), showing that accuracy decreases in the latter case.}
+    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
+    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of (from left to right) the mean spatial frequency, the bandwidth in orientation (from a grating-like (left) to isotropic textures (right)), the bandwidth in spatial frequency, the bandwidth in speed, from a rigid motion (left) to independent frames (right).
+Note that these accuracy is computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) or in the arbitrary case where orientation is independent of direction ('aperture').}
     \label{fig:motion_clouds}

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
-In particular, we will set the spatial size and duration similarly to the motion task defined above.
-This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions $
-\speed_\kernelind = 
-  ( 
-    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
-    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
-  )
-$.
+In particular, we will set the spatial extent and duration of the generated movies similarly to the motion task defined above.
+This procedure defines a set of textures with different spatial properties and different motions picked from the same set of $8$ directions and $3$ speeds.
 For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content to provide with some naturalistic variability.
@@
 First, as we change the mean spatial frequency of the texture, we observe a monotonous decrease in accuracy.
-This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006}.
-%\note{we should really do something more serious here...}
-Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively resemble a more textured stimulus) than for a grating-like stimulation, reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}.
+This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006} and revels the most informative scales learned by our model.
+We observe the tendency for different spatial frequency bandwidth.
+Notably, the accuracy is better for a large (which qualitatively resemble a more textured stimulus) than , reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}.
 Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 We observe the tendency for different spatial frequency bandwidth.
-Notably, the accuracy is better for a large (which qualitatively resemble a more textured stimulus) than , reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}.
+Notably, the accuracy is better for a grating-like stimulationfor a large (which qualitatively resemble a more textured stimulus) than , reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}.
 Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
@@
 We observe the tendency for different spatial frequency bandwidth.
-Notably, the accuracy is better for a large (which qualitatively resemble a more textured stimulus) than , reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}.
+Notably, the accuracy is better for a grating-like stimulation than , reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}.
 Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 We observe the tendency for different spatial frequency bandwidth.
-Notably, the accuracy is better for a grating-like stimulation than , reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}.
+Notably, the accuracy is better for a grating-like stimulation than for a large (which qualitatively resemble a more textured stimulus), reminiscent to the behavioral response of humans to such stimuli~\citep{simoncini_more_2012}.
 Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the visual field.
-In animals with a fovea, this is particularly useful because it allows the area with the highest density of photoreceptors in the environment to be moved, for example to a point of interest.
+In animals with a fovea, this is particularly useful because it allows the area with the highest density of photoreceptors in the environment to be moved, for example, to a point of interest.
 A specific mechanism for this function are saccades, which are rapid eye movements that reposition the center of vision.
@@
 289]{mandelbrot_fractal_1982}.
-Finally, the trajectory is integrated, assuming first that the velocities are sampled uniformly and independently from the set of different motion sets, and second that the motion is uniform during a time segment.
+Finally, the trajectory is integrated, assuming first that the velocities are sampled uniformly and independently of the set of different motion sets, and second that the motion is uniform during a time segment.
 The resulting instances yield trajectories that are qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
@@
 For a specific pixel and timestamp, an event is generated if the absolute value of this gradient exceeds a threshold.
-The event has either an OFF or ON polarity respectively, whether the gradient is negative or positive.
+The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive.
 This signed threshold value is then subtracted from the residual gradient image.
@@
 When they reach the soma of the postsynaptic neuron, the different spikes contribute to a modification of its membrane potential according to an activation function.
-In this example we use the same activation function as for a Leaky Integrate and Fire neuron~\textit{(Bottom-Left)}.
+In this example, we use the same activation function as for a Leaky Integrate and Fire neuron~\textit{(Bottom-Left)}.
 The first spiking motif is synchronized by the synaptic delays and causes a sudden rise in the membrane potential of the postsynaptic neuron.
@@
 This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted).
-These calculations are performed on a different input data set than the one used in the training or validation steps.
+These calculations are performed on a different input dataset than the one used in the training or validation steps.
 The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}.
@@
 Directions are shown as red arrows on the left.
-The spatio-temporal kernels are represented as a sequence of images at different delays along the horizontal axis from the far right (delay of one step) to the left (delay of $12$ steps).
+The spatio-temporal kernels are represented as slices of spatial weights at different delays.
+Delays varry along the horizontal axis from the far right (delay of one step) to the left (delay of $12$ steps).
 Each image corresponds to the weights at a given delay, with excitatory and inhibitory weights in warm and cold colors, respectively.
-Due to the observed symmetry between the ON and OFF event streams, we observed that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective to the different motion directions and we observe a slight level of orientation selectivity perpendicular to the orientation.
+Due to the observed symmetry between the ON and OFF event streams, we report that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective to the different motion directions and we observe a slight level of orientation selectivity perpendicular to the orientation.
 %      % These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
@@
     \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
-    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of (from left to right) the mean spatial frequency, the bandwidth in orientation (from a grating-like (left) to isotropic textures (right)), the bandwidth in spatial frequency, the bandwidth in speed, from a rigid motion (left) to independent frames (right).
-Note that these accuracy is computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) or in the arbitrary case where orientation is independent of direction ('aperture').}
+    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of (from left to right) the mean spatial frequency, the bandwidth in spatial frequency (from a grating-like (left) to isotropic textures (right)), the bandwidth in orientation (from isotropic textures (left) to grating-like (right)), the bandwidth in speed (from a rigid motion (left) to independent frames (right)).
+Note that these accuracy is computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) or in the generic case where orientation is independent of direction ('aperture').}
     \label{fig:motion_clouds}
@@
 Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection.
-We observed that the distribution of the kernels' weights is sparse, with most values near zero.
+We observed that the distribution of the kernels' weights is sparse, with most values near zero (\seeFig.~\ref{fig:kernels}).
 As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the number of spikes times the number of synapses.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 The spatio-temporal kernels are represented as slices of spatial weights at different delays.
-Delays varry along the horizontal axis from the far right (delay of one step) to the left (delay of $12$ steps).
+Delays vary along the horizontal axis from the far right (delay of one step) to the left (delay of $12$ steps).
 Each image corresponds to the weights at a given delay, with excitatory and inhibitory weights in warm and cold colors, respectively.
@@
 In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions.
-Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spike motif.
+Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spiking motif.
 Each voxel in the 3D kernels defines a specific timestamp and a specific address.
 Consequently, our model is able to detect precise spatio-temporal motifs embedded in the spike train and associated to the different motion directions.
-The cone shape for the positive weights distribution highlights a loss of precision for longer delays, i.e.
+The cone shape for the positive weights' distribution highlights a loss of precision for longer delays, i.e.
 events away in the past.
@@
 %%%-----------------------------------------------------------------
+%%%-----------------------------------------------------------------
+%: fig:motion_clouds
+%%%-----------------------------------------------------------------
+\begin{figure}%[h!]
+    \centering
+    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
+    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of (from left to right) the mean spatial frequency, the bandwidth in spatial frequency (from a grating-like (left) to isotropic textures (right)), the bandwidth in orientation (from isotropic textures (left) to grating-like (right)), the bandwidth in speed (from a rigid motion (left) to independent frames (right)).
+Note that these accuracies are computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) or in the generic case where orientation is independent of direction ('aperture').}
+    \label{fig:motion_clouds}
+\end{figure}
+%%%-----------------------------------------------------------------
 To test our model, we will quantify its ability to categorize different motions, that is on event streams for which the ground truth motion is known at every instant in time.
@@
 This method provides a rich dataset of textured movies for which we know the ground truth for motion.
-%%%-----------------------------------------------------------------
-%: fig:motion_clouds
-%%%-----------------------------------------------------------------
-\begin{figure}%[h!]
-    \centering
-    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
-    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of (from left to right) the mean spatial frequency, the bandwidth in spatial frequency (from a grating-like (left) to isotropic textures (right)), the bandwidth in orientation (from isotropic textures (left) to grating-like (right)), the bandwidth in speed (from a rigid motion (left) to independent frames (right)).
-Note that these accuracy is computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) or in the generic case where orientation is independent of direction ('aperture').}
-    \label{fig:motion_clouds}
-\end{figure}
-%%%-----------------------------------------------------------------
 
@@
 Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection.
-We observed that the distribution of the kernels' weights is sparse, with most values near zero (\seeFig.~\ref{fig:kernels}).
-As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the number of spikes times the number of synapses.
-Indeed, the computations are dominated by the convolution operation.
+We observed that the distribution of the kernels' weights is sparse, with most values near zero (\seeFig{kernels}).
+As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the computations used for the convolution operation.
 In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel.
@@
 Also, if we set some weights of the kernels to zero, then the sum can be skipped for those addresses.
-Knowing the sparseness of the input, the total number of computations thus scales with the number of nonzero synaptic weights.
+Knowing the sparseness of the input, the total number of computations thus scales with the number of spikes multiplied by the number of nonzero synaptic weights.
+% , which is proportional to the number of spikes multiplied by the number of synapses
 

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Each image corresponds to the weights at a given delay, with excitatory and inhibitory weights in warm and cold colors, respectively.
-Due to the observed symmetry between the ON and OFF event streams, we report that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective to the different motion directions and we observe a slight level of orientation selectivity perpendicular to the orientation.
+Due to the observed symmetry between the ON and OFF event streams, we report that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective to the different motion directions and we observe for all kernels an orientation preference perpendicular to the orientation.
 %      % These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
@@
 After training our model, we first analyze the weights learned for the different neurons (\seeFig{kernels}).
-We first observed a high dependence between the weights reaching the ON polarities and that reaching the OFF polarities.
+We first observe that these cells are very localized along each trajectory, yet show an orientation selectivity, similar to that observed in MT neurons~\citep{deangelis_functional_1999}.
+Focusing on the positive weights, a strong selectivity is observed along specific axes of motion for each of the different kernels.
+These directions can be easily associated to the direction of motion controlled in the natural images.
+For instance, the first kernel shows a strong selectivity to a downward motion direction.
+We also observe observed a high dependence between the weights reaching the ON polarities and that reaching the OFF polarities.
 In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative in the other.
 This property comes from the way the events are generated and that the luminance can not at the same time increase and decrease.
+Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results observed after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
 %ith the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
-We also observe that these cells show an orientation selectivity, similar to that observed in MT neurons~\citep{deangelis_functional_1999}.
-Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results observed after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
-Focusing on the positive weights, a strong selectivity is observed along specific axes of motion for each of the different kernels.
-These directions can be easily associated to the direction of motion controlled in the natural images.
-For instance, the first kernel shows a strong selectivity to horizontal motion directions.
 %
@@
 
-We plot here main axis of interests.
+We observe some facts of interest.
 First, as we change the mean spatial frequency of the texture, we observe a monotonous decrease in accuracy.
@@
 As a comparison and to account for the gain in performance by using heterogeneous delays, we provide the accuracy obtained with a MLR model using 2D time surface (in red) as in~\citep{grimaldi_robust_2022}.
-This latter method is based on delays from the last recorded events and uses fewer computations (in our case $15\times15$) than the dense 3D kernels without any pruning ($15\times15\times8$).
+This latter method is based on delays from the last recorded events and uses fewer computations (in our case $17\times17$) than the dense 3D kernels without any pruning ($21\times17\times17$).
 While less computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
@@
 By pruning weights, we observe that the evolution of accuracy as a function of the log percentage of active weights fits well a sigmoid curve.
-Half-saturation level is reached at about $3.5\times 10^{-3}\%$ of active weights, corresponding in our setting to a total amount of $6$ computations per decision.
-Compared to the full kernels, the accuracy of our method is maintained to its top performances when dividing the number of computations by a factor up to about $200$.
+Half-saturation level is reached at a ratio of about $3.2\times 10^{-2}$ of active weights, corresponding in our setting to a total amount of $180$ computations per decision.
+Compared to the full kernels, the accuracy of our method is maintained to its top performances when dividing the number of computations by a factor up to about $31$.
 In this case, the number of computations is greatly reduced compared to~\citep{grimaldi_robust_2022}, thus demonstrating the efficiency of the presented method.
@@
 %%%-----------------------------------------------------------------
-The HD-SNN model that has been trained and evaluated on a complex motion detection task.
+The HD-SNN model that has been trained and evaluated on a naturalistic motion detection task.
 The model has been defined to provide optimal detection of event-driven spatio-temporal motifs.
-We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in the early visual cortex~\cite{kremkow_push-pull_2016}.
+We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in the early visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
 %We then demonstrated that the model shows similarities to biological responses.
@@
 We would like to highlight a few innovations in the contributions that are presented in this paper.
-First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we observed to be less efficient, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs.
-Moreover, in comparison to a representation with time surfaces, the weights are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection of each spatio-temporal spike motif.
+First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs.
+Moreover, in comparison to a representation with that algorithm, the parameters of the model (weights and delays) are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection for each spatio-temporal spike motif.
 Another novelty is that the model learns the weights and the delays simultaneously.

### 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Thanks to % Thomas Schatz, Rosa Cossart,
-Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work. 
+Salvatore Giancani, Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work. 
 \Funding %
@@
 A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-Note that a neuron may contact the same afferent neuron with different delays through different synaptic connections.
+Note that a neuron may contact another afferent neuron with different delays through different synaptic connections.
 Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
@@
 This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation.
-Simulations are performed with the \note{PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$)}.
+Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$).
 We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
@@
 These calculations are performed on a different input dataset than the one used in the training or validation steps.
-The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}.
+The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN} (see Data Availability).
 %
@@
 events away in the past.
-For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Fig.~\ref{fig:kernels}.
+%For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Fig.~\ref{fig:kernels}.
 
@@
 Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter.
-This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018} but also that accuracy is low for a rigid motion which lacks variability.
+This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018}.
+% but also that accuracy is low for a rigid motion which lacks variability.
 %
@@
 By pruning weights, we observe that the evolution of accuracy as a function of the log percentage of active weights fits well a sigmoid curve.
-Half-saturation level is reached at a ratio of about $3.2\times 10^{-2}$ of active weights, corresponding in our setting to a total amount of $180$ computations per decision.
+Half-saturation level is reached at a ratio of about $3.2\times 10^{-2}$ of active weights, corresponding in our setting to a total amount of about $180$ computations per decision.
 Compared to the full kernels, the accuracy of our method is maintained to its top performances when dividing the number of computations by a factor up to about $31$.
@@
 In conclusion, the model that we have presented provides a way to process event-based signals in an efficient manner.
-We have shown that we can train the model semi-supervised, knowing \emph{what} output label, but not \emph{when} it occurs.
+We have shown that we can train the model using a supervised rule, knowing \emph{what} output label, but not \emph{where} it occurs.
 Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data~\citep{barlow_unsupervised_1989}.

### 2023-02-20 cover_letter/2023-02-20_coverletter.tex
cover_letter
Update on Overleaf.

@@
 \usepackage[margin=1in]{geometry}
+
+\usepackage{natbib}
 
@@
 The response of the model is tested on different sets of motion cloud stimuli with a control of the parameters of such stimuli generation (section 3.2).
-Regarding the MLR model, main differences with the previous ones are reported in the methods (P8, L269).}
+Regarding the MLR model, main differences with the previous ones are reported in the methods (P8, L285).}
 
@@
  \noteRev{We improved the mathematical formulation of the model in sections 2.2.1 and 2.2.2 and hope that it clarifies the general framework we use.
-Regarding training, it is described more clearly in P8, L295.
+Regarding training, it is described more clearly in P9, L313.
 Demonstration of the mechanism and the analogy with Hebbian learning was done in~\cite{grimaldi_robust_2022} and was not reproduced here.
@@
 
- \noteRev{We gave a better explanation of the symmetry assumption in the methods (P8, L293) an how this new model uses this invariance in its computation like observed in complex cells.
+ \noteRev{We gave a better explanation of the symmetry assumption in the methods (P9, L303) an how this new model uses this invariance in its computation like observed in complex cells.
 Indded, we don't want the polarity to have an impact on the classification and then this observation does not apply to the revised manuscript as it is a prior and not an emergent property.}
@@
 As mentioned in the responses to reviewer \#1 in this cover letter, we have now added this citation~\cite{grimaldi_learning_2022}.
-In addition, in this submitted manuscript we describe the main differences from that previous study: the use of a more ecological task (Section 2.1), a complete formulation of the model (Section 2.2), and the improvements made in this particular model (P8, L269).
-We also provide citations to studies using similar methods (P7, L243).
+In addition, in this submitted manuscript we describe the main differences from that previous study: the use of a more ecological task (Section 2.1), a complete formulation of the model (Section 2.2), and the improvements made in this particular model (P8, L285).
+We also provide citations to studies using similar methods (P8, L259).
 Further results are presented, especially the testing of the learned model with parametric textures.}
 
-\bibliographystyle{apalike}
-\bibliography{../FastMotionDetection}
+\bibliographystyle{plainnat}
+\bibliography{FastMotionDetection}
 

### 2023-03-01 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 We observe the tendency for different spatial frequency bandwidth.
-Notably, the accuracy is better for a grating-like stimulation than for a large (which qualitatively resemble a more textured stimulus), reminiscent to the behavioral response of humans to such stimuli~\citep{simoncini_more_2012}.
+Notably, the accuracy is better for a grating-like stimulation than for a large (which qualitatively resemble a more textured stimulus), reminiscent to the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}.
 Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.

### 2023-03-22 2023-03-28_HBP_summit.tex
Update on Overleaf.

--- 2023-02-20 Grimaldi-etal-BiolCybernetics.tex
@@
-% !TEX encoding = UTF-8 Unicode
-% !TeX TS-program = pdfLaTeX
-% !TeX spellcheck = en-US
-% !BIB TS-program = bibtex
+% !TeX TS-program = LuaLatex
+% !TeX spellcheck = en
+% !BIB TS-program = biber
+% !TeX encoding = UTF-8 Unicode
 % -*- coding: UTF-8; -*-
 % vim: set fenc=utf-8
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% public
-% https://github.com/SpikeAI/pyTERtorch/ (for dev)
-% https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN 
-% https://www.zotero.org/groups/4776796/fastmotiondetection
-%
-% private
-% https://framateam.org/spikeai/channels/fast-motion-biolcyb
-% https://www.overleaf.com/project/630f9044c38e7a3cea81a7b2
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Guest Editors:
-% 
-% Kexin Chen - Department of Cognitive Sciences, University of California, Irvine
-% Hirak J. Kashyap - Department of Computer Science, University of California, Irvine
-% Jeffrey L. Krichmar - Department of Cognitive Sciences, Department of Computer Science, University of California, Irvine
-% Xiumin Li - College of Automation, Chongqing University, Chongqing, China
-% 
-% Kexin Chen, Hirak Kashyap, Jeffrey Krichmar, Xiumin Li
-% NeuroVision Organizing Team
-% 
-% Kexin Chen <kexinc3@uci.edu>
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% 2022-09-30
-% Learning heterogeneous delays in a layer of spiking neurons for fast motion detection
-% Corresponding Author: Laurent Perrinet
-% Biological Cybernetics
-% a7f45f8e-87d6-4a30-ac7f-9f0c94e2592a   |  v.1.0
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%Call for papers: Biological Cybernetics Special Issue: What can Computer Vision learn from Visual Neuroscience?
-%
-% https://www.springer.com/journal/422/updates/20421374
-%
-%Computer vision still struggles to solve many problems as quickly and accurately as our brain does, despite its tremendous progress using deep learning over the last decade.
-Current challenges include geometry understanding, dynamic scene understanding, few shot learning of novel objects, anomaly and counterfeit detection, stable reproduction of imaginary visuals, and more.
-The brain achieves all these on a low computational cost and power budget (i.e., tens of watts), whereas the state-of-the-art computer vision models require multiple graphics processors consuming kilowatts of power.
-Hence, new inspirations from the brain could make computer vision more efficient, robust, and capable of continuous learning and adaptation.
-Although visual neuroscience literature provides knowledge about neuronal selectivities and early processing of visual inputs, computational modelling of neuronal and network behaviours with some degree of biological constraints and characteristics is an important tool to address many of the challenges facing computer vision.
-With new neuroscientific insights about deep cortical vision processing using the latest stimulation and recording technologies, it is important that we strive to understand the underlying computations and learning mechanisms, so that artificial vision systems with similar efficiency can be developed.
-This special issue invites original research and review articles related to topics in biological vision that can potentially benefit computer vision systems.
-The following is a non-exhaustive list of topics in biological vision that can also benefit computer vision systems.
-%
-%● Active vision’s role in visual search, scene understanding, social interactions, etc.
-%
-%● Learning in the visual system. Learning in biology is continual, few-shot, and adversarially robust.
-%
-%● The roles of recurrent and top-down connections in the visual cortex.
-%
-%● Spike based spatio-temporal processing and its implications for neuromorphic vision.
-%
-%● Motion perception in dynamic environments.
-%
-%● Neural coding schemes in the visual system (e.g., sparse coding, predictive coding, and temporal coding.)
-%
-%● The roles of attention mechanisms in biological vision.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% il y a aussi tikzposter qui est bien cf https://www.overleaf.com/learn/latex/Posters
+\documentclass[landscape,a0]{a0poster}
+%\documentclass[landscape,a0, draft]{a0poster}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\doc}{poster}
+
 \newcommand{\Title}{
-Learning heterogeneous delays in a layer of spiking neurons for fast motion detection
+Learning heterogeneous delays of spiking neurons for motion detection
+%Learning hetero-synaptic delays for motion detection in a population of Spiking Neurons
+%FAST MOTION DETECTION USING HETERO-SYNAPTIC DELAYS IN A LAYER OF SPIKING NEURONS
 }
-\newcommand{\ShortTitle}{
-Learning heterogeneous delays of spiking neurons for motion detection
-}
-\newcommand{\FirstAG}{Antoine}
-\newcommand{\LastAG}{Grimaldi}
-\newcommand{\AuthorAG}{\FirstAG \LastAG}
-\newcommand{\FirstLP}{Laurent U}
-\newcommand{\LastLP}{Perrinet}
-\newcommand{\AuthorLP}{\FirstLP \LastLP}
+\newcommand{\AuthorAG}{Antoine Grimaldi}
+\newcommand{\AuthorCB}{Camille Besnainou}
+\newcommand{\AuthorHL}{Hugo Ladret}
+\newcommand{\AuthorLP}{Laurent U Perrinet}
 \newcommand{\EmailLP}{laurent.perrinet@univ-amu.fr}
 \newcommand{\EmailAG}{antoine.grimaldi@univ-amu.fr}
+\newcommand{\EmailHL}{hugo.ladret@univ-amu.fr}
 \newcommand{\orcidLP}{0000-0002-9536-010X}
+\newcommand{\orcidHL}{0000-0001-7999-3751}
 \newcommand{\orcidAG}{0000-0002-3107-4788}
-\newcommand{\Department}{Institut de Neurosciences de la Timone}% (UMR 7289)}%
+\newcommand{\Address}{Institut de Neurosciences de la Timone (UMR 7289); Aix Marseille Univ, CNRS; Marseille, France}%
 \newcommand{\Affiliation}{Aix Marseille Univ, CNRS}%
-\newcommand{\Street}{27 boulevard Jean Moulin}%
-\newcommand{\PostCode}{13005}%
-\newcommand{\City}{Marseille}%
-\newcommand{\Country}{France}%
+\newcommand{\CityLP}{Marseille, France}%
 \newcommand{\WebsiteLP}{https://laurentperrinet.github.io}%
@@
 }
-\newcommand{\Funding}{
-This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{APROVIS3D}{http://aprovis3d.eu/}''). 
-LP received funding from the ANR project  N° ANR-20-CE23-0021 ``\href{AgileNeuroBot}{https://laurentperrinet.github.io/grant/anr-anr/}''
-and from A*Midex grant number AMX-21-RID-025
-``\href{Polychronies}{https://laurentperrinet.github.io/grant/polychronies/}''. 
+\newcommand{\Acknowledgments}{AG and LP received funding from the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}'').
+HL and LP received funding from the ANR project  N° ANR-20-CE23-0021 (''\href{https://laurentperrinet.github.io/grant/anr-anr/}{AgileNeuroBot}'').
+%HL received funding from a University of Montreal Artificial Intelligence scholarship.
+For the purpose of open access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript version arising from this submission.
 }
-\newcommand{\Acknowledgments}{
-Thanks to % Thomas Schatz, Rosa Cossart,
-Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work. 
-\Funding %
-%For the purpose of open access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript version arising from this submission.
-This research was funded, in whole or in part, by [Organisation name, Grant #].
-A CC-BY public copyright license has been applied by the authors to the present document and will be applied to all subsequent versions up to the Author Accepted Manuscript arising from this submission, in accordance with the grant’s open access conditions.
-}
-\newcommand{\DataAvailability}{
-This works is made reproducible.
-The code reproducing the manuscript and all figures is available on~\href{ https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}{GitHub}.
-It also contains supplementary figures and results.
-Find also the associated~\href{https://www.zotero.org/groups/4776796/fastmotiondetection}{zotero group} used to gather relevant literature on the subject.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                                                 %%
-%% Please do not use \input{...} to include other tex files.       %%
-%% Submit your LaTeX manuscript as one .tex document.              %%
-%%                                                                 %%
-%% All additional figures and files should be attached             %%
-%% separately and not embedded in the \TeX\ document itself.       %%
-%%                                                                 %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%\documentclass[referee,sn-basic]{sn-jnl}% referee option is meant for double line spacing
-
-%%=======================================================%%
-%% to print line numbers in the margin use lineno option %%
-%%=======================================================%%
-
-%%\documentclass[lineno,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
-
-%%======================================================%%
-%% to compile with pdflatex/xelatex use pdflatex option %%
-%%======================================================%%
-
-%%\documentclass[pdflatex,sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
-
-%%\documentclass[sn-basic]{sn-jnl}% Basic Springer Nature Reference Style/Chemistry Reference Style
-%%\documentclass[sn-mathphys]{sn-jnl}% Math and Physical Sciences Reference Style
-%%\documentclass[sn-aps]{sn-jnl}% American Physical Society (APS) Reference Style
-%%\documentclass[sn-vancouver]{sn-jnl}% Vancouver Reference Style
-%%\documentclass[sn-apa]{sn-jnl}% APA Reference Style
-%%\documentclass[sn-chicago]{sn-jnl}% Chicago-based Humanities Reference Style
-%%\documentclass[sn-standardnature]{sn-jnl}% Standard Nature Portfolio Reference Style
-\documentclass[default]{sn-jnl}% Default
-%%\documentclass[default,iicol]{sn-jnl}% Default with double column layout
-%%\documentclass[iicol]{sn-jnl}
-
-\usepackage[T1]{fontenc}
-%\usepackage{textcomp}
-\usepackage[utf8]{inputenc}
-%
-\usepackage{lineno}
-\linenumbers
-
-%%%% Standard Packages
-%%<additional latex packages if required can be included here>
-%\usepackage[lofdepth,lotdepth,position=top]{subfig}
-%\usepackage{tabularx}
-%%%%
-
-%%%%%=============================================================================%%%%
-%%%%  Remarks: This template is provided to aid authors with the preparation
-%%%%  of original research articles intended for submission to journals published 
-%%%%  by Springer Nature. The guidance has been prepared in partnership with 
-%%%%  production teams to conform to Springer Nature technical requirements. 
-%%%%  Editorial and presentation requirements differ among journal portfolios and 
-%%%%  research disciplines. You may find sections in this template are irrelevant 
-%%%%  to your work and are empowered to omit any such section if allowed by the 
-%%%%  journal you intend to submit to. The submission guidelines and policies 
-%%%%  of the journal take precedence. A detailed User Manual is available in the 
-%%%%  template package for technical guidance.
-%%%%%=============================================================================%%%%
-
-\jyear{2023}%
-%% as per the requirement new theorem styles can be included as shown below
-\theoremstyle{thmstyleone}%
-\newtheorem{theorem}{Theorem}%  meant for continuous numbers
-\newtheorem{proposition}[theorem]{Proposition}% 
-%
-\theoremstyle{thmstyletwo}%
-\newtheorem{example}{Example}%
-\newtheorem{remark}{Remark}%
-
-\theoremstyle{thmstylethree}%
-\newtheorem{definition}{Definition}%
-
-\raggedbottom
-
-\newcommand{\seeFig}[1]{see Fig.~\ref{fig:#1}}%{see Figure~\ref{fig:#1}}
-%% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-%\usepackage{cite}
-%\usepackage[usenames,dvipsnames]{xcolor} %The package starts from the basic facilities of the color package, and provides easy driver-independent access to several kinds of color tints, shades, tones, and mixes of arbitrary colors.
-%\usepackage{tcolorbox} % define boxes that can be customed, has to be loaded after xcolor package
-\usepackage{amsmath,amssymb,amsfonts} % amsmath and amssymb packages, useful for mathematical formulas and symbols
-%%\usepackage{float} %allows me to use [H] with figures to insert figures in multicol env (poster)
-%\usepackage{ulem} %The package provides an \ul (underline) command which will break over line ends; this technique may be used to replace \em (both in that form and as the \emph command), so as to make output look as if it comes from a typewrite.
-Caution: with this package, bibliography will make unbreakable and underlined journal titles
-%\usepackage[switch]{lineno} % add line numbers
-%\modulolinenumbers[5]
-\usepackage{physics} %The  goal  of  this  package  is  to  make  typesetting  equations  for  physics  simpler,  faster,  and  more  human-readable.
-%\usepackage{hyperref} %The hyperref package is used to handle cross-referencing commands in LaTeX to produce hypertext links in the document
-%\hypersetup{
-%    colorlinks = false,
-%    hidelinks
-%    %linkbordercolor = {white},
-%}
-%\usepackage{array} %An extended implementation of the array and tabular environments which extends the options for column formats, and provides “programmable” format specifications.
-%\usepackage{multicol} %Multicol defines a multicols environment which typesets text in multiple columns
-%%\usepackage{wrapfig} %Allows figures or tables to have text wrapped around them. wrapfig is used when calling multicols
-% \usepackage[position=top]{subfig} %The package provides support for the manipulation and reference of small or ‘sub’ figures and tables within a single figure or table environment.
-Linked to subfloat.
-% for units
-\usepackage{siunitx}%The siunitx package provides a  set  of  tools  for  authors  to  typeset  quantities  in  aconsistent  way.
-%\AtBeginDocument{\RenewCommandCopy\qty\SI}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\DeclareMathOperator*{\argmin}{arg\,min}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% NOTATIONS
-\newcommand{\ms}{\si{\milli\second}}%
-\newcommand{\us}{\si{\micro\second}}%
 \newcommand{\presynaddr}{a} % pre address
-\newcommand{\postsynaddr}{b} % post address
+\newcommand{\postsynaddr}{n} % post address
 \newcommand{\numevent}{N_{ev}} % total number of events
@@
 \newcommand{\arank}{r} % address index
-\newcommand{\bias}{\beta} % bias for the MLR model
-\newcommand{\synapse}{\mathcal{S}} % synapse
+\newcommand{\bias}{b} % bias for the MLR model
+\newcommand{\synapse}{\sigma} % synapse
 \newcommand{\synapticweight}{w} % synaptic weight
-\newcommand{\synapticdelay}{\delta} % synaptic delay
+\newcommand{\synapticdelay}{d} % synaptic delay
 \newcommand{\ranksyn}{s} % synapse index
@@
 \newcommand{\layer}{\mathbf{L}} % layer
-% \newcommand{\timecontext}{T} % time context (cf HOTS) matrice gathering last event times
+\newcommand{\ms}{\si{\milli\second}}%
+\newcommand{\us}{\si{\micro\second}}%
+\newcommand{\timecontext}{T} % time context (cf HOTS) matrice gathering last event times
 \newcommand{\current}{I} % post synaptic current
@@
 \newcommand{\gain}{\gamma} % homeostatic gain
-% \newcommand{\simil}{\beta} % similarity value
+\newcommand{\simil}{\beta} % similarity value
 \newcommand{\Nclass}{N_\text{class}} % number of classes for MLR:
@@
 \newcommand{\Ny}{N_\text{Y}}
-\newcommand{\Ntime}{N_\text{T}}
-\newcommand{\kernel}{\mathcal{K}} % convolution kernel
+\newcommand{\Ntime}{N_\text{t}}
+\newcommand{\kernel}{K} % convolution kernel
 %\newcommand{\kernelind}{\mathbf{k}} % indice of the kernel
 \newcommand{\kernelind}{k} % indice of the kernel
-\newcommand{\Kx}{K_\text{X}}
-\newcommand{\Ky}{K_\text{Y}}
-\newcommand{\Ktime}{K_\text{T}}
+\newcommand{\Kx}{K_\text{x}}
+\newcommand{\Ky}{K_\text{y}}
+\newcommand{\Ktime}{K_\text{t}}
 \newcommand{\classiflayer}{\mathbf{C}}
@@
 \newcommand{\decision}{\hat{y}}
-\newcommand{\colorsec}{black}
-\newcommand{\colorsubsec}{black}
 \newcommand{\speed}{v}
 \newcommand{\Nspeed}{N_v}
-% Example definitions.
-% --------------------
-\def\x{{\mathbf x}}
-\def\L{{\cal L}}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% a simpler annotation system:
-\usepackage{soulutf8}
-\usepackage{color}
-\newcommand{\note}[1]{{\sethlcolor{yellow}\hl{#1}}}
-%\newcommand{\note}[1]{}
 %
+\newcommand{\colorsec}{red}
+\newcommand{\colorsubsec}{gray}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{fontspec}
+\usepackage{titlesec}
+\titleformat*{\section}{\huge\bfseries}
+\usepackage{float} %allows me to use [H] with figures to insert figures in multicol env (poster)
+\usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{hyperref}
+\usepackage{multicol}
+\usepackage[utf8]{inputenc}
+\defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+\setmainfont{Alegreya}
+\DeclareMathOperator*{\argmax}{arg\,max}
+\usepackage[
+%style=chem-acs,
+style=numeric,						% numeric style for reference list
+citestyle=numeric-comp,
+%style=alphabetic-verb,
+giveninits=false,
+maxbibnames=1,
+%firstinits=true,
+%style=apa,
+%maxcitenames=1,
+%maxnames=3,
+%minnames=1,
+%maxbibnames=99,
+dateabbrev=true,
+giveninits=true,
+%uniquename=init,
+url=false,
+doi=false,
+isbn=false,
+eprint=false,
+texencoding=utf8,
+bibencoding=utf8,
+autocite=superscript,
+backend=biber,
+%sorting=none,
+sorting=none,
+sortcites=false,
+%articletitle=false
+]{biblatex}%
+
+\bibliography{references.bib}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\columnsep=2cm % This is the amount of white space between the columns in the poster
+\columnseprule=2pt % This is the thickness of the black line between the columns in the poster
+\def\columnseprulecolor{\color{black}}
+\usepackage{multirow}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\sffamily
+\usepackage[most]{tcolorbox}
+\newtcbox{\mybox}{size = normal, on line}% boxsep=3mm,
+%outer arc=1mm, arc=3mm,
+%boxrule=2mm,colback=red!10!white, colframe=red!50!black, arc is angular}
 \begin{document}
-%%%-----------------------------------------------------------------
-\title[\ShortTitle]{\Title}
-\artnote{This article is published as part of the Special Issue on "What can Computer Vision learn from Visual Neuroscience?"}
-% ---------------
-\author{\fnm{\FirstAG} \sur{\LastAG}}\email{\EmailAG}
-\author{\fnm{\FirstLP} \sur{\LastLP}}\email{\EmailLP}
-\affil{\orgdiv{\Department}, \orgname{\Affiliation}\\ \orgaddress{\street{\Street}, \city{\City}, \postcode{\PostCode}, \country{\Country}}}
+%----------------------------------------------------------------------------------------
+%	POSTER HEADER
+%----------------------------------------------------------------------------------------
+% The header is divided into two boxes:
+% The first is 75% wide and houses the title, subtitle, names, university/organization and contact information
+% The second is 25% wide and houses a logo for your university/organization or a photo of you
+% The widths of these boxes can be easily edited to accommodate your content as you see fit
+\vspace*{-2cm}
 %
-\abstract{
+\begin{minipage}[b]{0.175\textwidth}
+%\hspace*{2cm}
+\includegraphics[width=.4\textwidth]{figures/int_logo.pdf}
+\includegraphics[width=.6\textwidth]{figures/amu_logo.pdf}\\
+%\vspace{-.5cm}
+\begin{center}
+\color{blue}
+\Large
+\href{mailto://\EmailAG}{\EmailAG} \hspace{.2cm}
+\end{center}
+\vspace{1cm}
+\end{minipage}
+%
+\begin{minipage}[b]{.65\textwidth}
+\begin{center}
+\VeryHuge
+\textbf{\color{\colorsec}
+\Title
+} \\ % \color{Black}
+\end{center}
+\large
+\centering \hspace{.5cm} \textbf{
+\AuthorAG , \hspace{.1cm}
+%\AuthorCB , \hspace{.1cm}
+%\AuthorHL  \hspace{.1cm}
+\& \hspace{.1cm}
+\AuthorLP
+%\hspace{.5cm}
+ -
+%
+\large
+%\hspace{.1cm}
+\Address} \\
+%\vspace{1cm}
+\end{minipage}
+%
+\begin{minipage}[t]{0.175\textwidth}
+%
+%\vskip{-8cm}
+\vspace{-9cm}
+%\begin{tabular} {c c}
+%\includegraphics[height=.20\textwidth]{figures/cnrs_logo.pdf} & \includegraphics[height=.20\textwidth]{figures/brain-logo-240.jpg} \\
+%\multicolumn{2}{c}{\includegraphics[height=.15\textwidth]{figures/chistera_logo.pdf}}\\
+%\end{tabular}
+\includegraphics[height=.2\textwidth]{figures/cnrs_logo.pdf}\hspace{1cm} \includegraphics[height=.2\textwidth]{figures/chistera_logo.pdf}\hspace{1cm}
+%\includegraphics[height=.15\textwidth]{figures/brain-logo-240.jpg}
+%\vskip{-2cm}
+%\hspace*{5cm}
+%\includegraphics[height=.15\textwidth]{figures/chistera_logo.pdf}\\
+%\vspace*{3cm}
+%\hspace{3cm}
+%\includegraphics[height=.20\textwidth]{figures/cnrs_logo.pdf}
+%\hspace{1cm}
+%\includegraphics[height=.20\textwidth]{figures/brain-logo-240.jpg}
+%\\
+\vspace{.25cm}
+
+\begin{center}
+\color{blue}
+\Large
+\href{https://laurentperrinet.github.io/publication/grimaldi-22-icip/}{https://laurentperrinet.github.io
+\\ /publication/grimaldi-22-icip/} \end{center}
+\vspace{2cm}
+
+\end{minipage}
+%
+\vspace*{-1.5cm}
+\hrule
+\vspace{5pt}
+\hrule
+%----------------------------------------------------------------------------------------
+\begin{multicols}{3} % This is how many columns your poster will be broken into, a portrait poster is generally split into 2 columns
+%----------------------------------------------------------------------------------------
+%	ABSTRACT
+%----------------------------------------------------------------------------------------
+\section*{\color{\colorsec}Abstract \vspace{.3cm}  \hrule}
+\vspace{-2.5cm}
 \Abstract
-}
-
-\keywords{\Keywords}
+%----------------------------------------------------------------------------------------
+%	MLR
+%----------------------------------------------------------------------------------------
+\section*{\color{\colorsec}Heterogeneous delays model\hrule}
 %
-\maketitle
+\vspace*{-2.5cm}
+\subsection*{\color{\colorsubsec}Illustration}
 %
-\section{Introduction}
-%%%-----------------------------------------------------------------
-\label{sec:intro}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}
+    \caption{\textbf{The core mechanism of the HD SNN model.} \textit{(Top-Left)}~Two spiking motifs are emitted from four presynaptic neurons.
+Once integrated by the synapses of the postsynaptic neuron~\textit{(Right)}, the spiking motifs are shifted in time by the synaptic delays and weighted by the synaptic weights~\textit{(Middle-Left)}.
+When they reach the soma of the postsynaptic neuron, the different spikes contribute to a modification of its membrane potential according to an activation function.
+In this example, we use the same activation function as for a Leaky Integrate and Fire neuron~\textit{(Bottom-Left)}.
+The first spiking motif is synchronized by the synaptic delays and causes a sudden rise in the membrane potential of the postsynaptic neuron.
+An output spike is emitted when the membrane potential reaches the threshold and it is then reset.
+\textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses).
+}%
+    \label{fig:izhikevich}%
+\end{figure}
+%TODO: changer le temps d'occurence des spikes
+%\vspace*{1.5cm}
+\subsection*{\color{\colorsubsec}Mathematical formalism}
 %
-The human brain has the remarkable property of being able to react at any time while consuming a reasonable amount of energy, about tens of watts.
-This system is the result of millions of years of natural selection, and a striking difference with artificial neural networks is the representation both use.
-Artificial convolutional neural networks (CNNs), for example, represent the flow of information from one layer to another as tensors, storing visual information densely across the visual topography, with different properties represented in different channels.
-These networks are known to mimic many properties of biological systems, such that each can be attributed a ``Brain Score''~\citep{schrimpf_brain-score_2020}, yet this score takes no account of inference speed or energy consumption.
-CNNs have achieved state-of-the-art performance for some computer vision tasks, such as image recognition.
-However, they do not take advantage of the dynamics inherent in the way we perceive our natural environment.
-In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
-The third generation of artificial neural networks, known as spiking neural networks (SNNs), incorporates this temporal dimension into the way they perform their computations.
-They are of interest for computational neuroscience because they provide a better model of the biological neuron.
-Some approaches have developed normative models of SNNs that aim to have applications in machine learning.
-One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
-Another type of SNN that uses accurate timing of spikes attempts to determine the structure of the network to minimize a cost function.
-This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been recently extended.
-This method uses a surrogate gradient and is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
-However, the performance of SNNs still lags behind that of firing-rate based networks.
-The question of the advantage of using spikes in machine learning and computer vision remains open.
-
-In a recent review paper, we reported experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues~\cite{grimaldi_precise_2023}.
-These spatio-temporal motifs within the neural activity may be useful representations to perform computations for various cognitive tasks.
-Notably, \citet{abeles_role_1982} asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs.
-While the first possibility favors the firing-rate coding theory, the second one emphasizes the function of temporal precision in the neural code.
-Since then, numerous studies have demonstrated the emergence of synchronicity of activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding using spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
-All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
-In neural models, the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows for the efficient exploitation or detection of these spatio-temporal motifs embedded in the spike train.
-In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, but precise, relative spiking delays, i.e., the time between the arrival of an afferent spike at a given synapse and the contribution of the associated postsynaptic potential to the neuron's soma.
-Due to the variety of configurations and the possible coexistence of multiple superimposed motifs, representation with polychronous groups has a much higher information capacity than a firing rate-based approach to neural coding.
-
-The present paper proposes additional experiments extending a recent model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}.
-This model was trained to solve a motion detection task on a synthetic event-based dataset generated by moving parameterized textures.
-Once trained, the volume of event-driven computations could be drastically reduced by pruning the synapses, while maintaining top performance for the classification task.
-This was a demonstration that formal neurons can exploit the precise timing of spikes to improve neural computations.
-In the present work, we extend motion detection to a more ecological task.
-Instead of the synthetic textures that have been used to generate motion-driven stimuli, we use natural images and random movements that mimic biological saccades.
-Using an ecological motion detection task, we study the emergence of spatio-temporal spiking motifs when a single layer of spiking neurons is trained on a supervised classification task.
-First, we define the ecological cognitive task the model has to solve with the different datasets it will be tested on.
-Then, we develop a theoretically defined Heterogeneous Delays Spiking Neural Network (HD-SNN) model capable of learning these heterogeneous synaptic delays.
-We will investigate the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.
-In this way, we will be able to show how such a model can provide an efficient solution to the trade-off between energy and accuracy.
+\textbf{Event stream:}
+%The spikes reaching our SNN can be formalized as:
+$$
+\mybox{$\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$}
+$$
+where $\numevent \in \mathbb{N} $ is the total number of events, $\timev_\arank$ is the time occurrence of event number $\arank$ and $\presynaddr_\arank$ an associated address , which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
+\\
 %
-%%%-----------------------------------------------------------------
-%%%-----------------------------------------------------------------
-\section{Methods}
-\label{sec:methods}
-%%%-----------------------------------------------------------------
-In this paper, we aim to test whether the HD-SNN model is capable of efficiently learning a motion detection task, which is defined by a realistic event-based data stream.
-This type of signal is typically captured by a Dynamic Vision Sensor (DVS) and is inspired by the signal that is sent from the retinal ganglion cells through the optic nerve.
-The events are binary in nature and should be sufficient to perform the task of motion detection in a fast and efficient manner.
-Here we will first describe the task definition and then the HD-SNN model we use for motion inference and how we train it.
-%%%-----------------------------------------------------------------
+\\
+\textbf{Spiking neuron with hetero-synaptic delays:}
 %
-\subsection{Task definition: motion detection in a synthetic naturalistic event stream}\label{sec:task}
-%%%-----------------------------------------------------------------
-To train and test our model, we need to define a visual dataset for which we control the direction and speed of motion.
-Let us now define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
-First, we draw a trajectory inspired by the biological movements of the eyes.
-Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the visual field.
-In animals with a fovea, this is particularly useful because it allows the area with the highest density of photoreceptors in the environment to be moved, for example, to a point of interest.
-A specific mechanism for this function are saccades, which are rapid eye movements that reposition the center of vision.
-In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}).
-They are generated very quickly (about $80~\ms$) and have a wide range of speeds.
-On a more microscopic scale, the human gaze moves with a continuous drift similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
-In order to maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.
-This approach first defines a finite set of possible 2D movements in polar coordinates.
-Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $8$ linearly spaced movement directions and 3 geometrically spaced speeds.
-Next, we define a gaze trajectory as segments whose duration is drawn from a Poisson distribution with a mean block length of $24~\ms$, similar to a Lévy flight~\citep[p.
-289]{mandelbrot_fractal_1982}.
-Finally, the trajectory is integrated, assuming first that the velocities are sampled uniformly and independently of the set of different motion sets, and second that the motion is uniform during a time segment.
-The resulting instances yield trajectories that are qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
-
-%%%-----------------------------------------------------------------
-%: fig:motion_task
-%%%-----------------------------------------------------------------
-\begin{figure}%[h!]%[ht!]
+$$
+\mybox{$\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$}
+$$
+is a set of $\Nsyn^\postsynaddr$ synapses, associated to neuron $\postsynaddr$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.\\
+%
+\\
+\textbf{Active weights:}
+%
+$$
+\mybox{$\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}$}
+$$
+%
+\textbf{Activation function (Multinomial Logistic Regression):}
+%
+$$
+\mybox{$Pr(k=\postsynaddr \; | \; \timev) =\frac 1 Z \cdot {\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }$}
+$$
+where $\mathcal{C}^\postsynaddr(\timev) = \sum \activeweights^\postsynaddr(t)
+$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$ and $Z= \sum_{c=1}^{\Nclass} Pr(k=c \; | \; \timev)$.\\
+%
+\\
+\textbf{Temporal convolution:}\\
+In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$ (see Figure~\ref{fig:kernels}).
+Using this dense representation, the counting $\mathcal{C}^\postsynaddr(\timev)$ defined above can be computed as a temporal convolution of the dense representation of the event stream $E$ with the dense kernels formed by the set of synapses:
+$$
+\mybox{$\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast E$}
+$$
+%----------------------------------------------------------------------------------------
+%	MOTION CLOUDS
+%----------------------------------------------------------------------------------------
+\section*{\color{\colorsec}Event-based input \vspace{.2cm}  \hrule}
+%
+\vspace{-2.5cm}
+To test our model, we will quantify its ability to categorize different motions.
+In that order, we will first define a set of synthetic stimuli, natural images with a rigid motion or also \textit{Motion Clouds}~\cite{leon2012motion}, which are natural-like random textures for which we can control for velocity, among other parameters.
+%
+\begin{figure}[H]
     \centering
@@
 \end{figure}
-%%%-----------------------------------------------------------------
-Once these eye trajectories are generated, we can apply them to a visual scene.  
-For this purpose, we selected a database of $600$ natural images that were previously used to study the statistics of natural images~\citep{perrinet_edge_2015}.
-Note that these have been pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band.
-This process is known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $256 \times 256$ in size, and we will extract sub-images of size $64 \times 64$ positioned around the center of gaze at each time step.
-We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 128~\ms$.
-To avoid boundary effects, we will randomly position the full trajectory in image space such that the subimage is translated using the position given by the trajectory at each time step without touching the borders.
-At each point in time, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
-Note that the magnitude of the displacement is relative to the time bin, and we have defined the unit velocity to correspond to a movement of one pixel per frame (i.e., per time bin).
-
-To transform each movie into events as the ones recorded by a DVS, we compute a gradient image (initialized at zero) by computing the temporal gradient of the pixels' intensity over two successive frames.
-For a specific pixel and timestamp, an event is generated if the absolute value of this gradient exceeds a threshold.
-The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive.
-This signed threshold value is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion_task}-\textit{(Right)}).
-The goal here is to infer the correct motion solely by observing these events.
-% \note{ the task is full-field rigid translations, the model resolves it locally (analogy to the input from V1 to MT) }
-%%%-----------------------------------------------------------------
-\subsection{The Heterogeneous Delays SNN (HD-SNN) model}
-%----------------------------%
 %
-\begin{figure}%[t!]
-    \centering
-    \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}
-    \caption{\textbf{The core mechanism of the HD SNN model.} \textit{(Top-Left)}~Two spiking motifs are emitted from four presynaptic neurons.
-Once integrated by the synapses of the postsynaptic neuron~\textit{(Right)}, the spiking motifs are shifted in time by the synaptic delays and weighted by the synaptic weights~\textit{(Middle-Left)}.
-When they reach the soma of the postsynaptic neuron, the different spikes contribute to a modification of its membrane potential according to an activation function.
-In this example, we use the same activation function as for a Leaky Integrate and Fire neuron~\textit{(Bottom-Left)}.
-The first spiking motif is synchronized by the synaptic delays and causes a sudden rise in the membrane potential of the postsynaptic neuron.
-An output spike is emitted when the membrane potential reaches the threshold and it is then reset.
-\textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses).
-}%
-    \label{fig:izhikevich}%
-\end{figure}%
+\vspace{1cm}
+\hrule
 %
-%----------------------------%
-In a neurobiological recording or in the sensory signal obtained from an event-based camera, the input consists of a stream of \emph{spikes} or \emph{events}.
-This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{(Top-Left} for an illustration).
-Events are typically ordered by their time of occurrence.
-Each event has a time of occurrence $\timev_\arank$ and an associated address $\presynaddr_\arank$.
-This defines an address space $\presynaddrspace$ which consists of the set of possible addresses.
-In a neurobiological recording, this can be the identified set of presynaptic neurons.
-For neuromorphic hardware, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$ where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels.
-As such, an address $\presynaddr_\arank$ is typically in the form $(\polev_\arank, x_\arank, y_\arank)$ for event-based cameras.
-
-In the HD-SNN model, each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferent neurons from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses.
-Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-Note that a neuron may contact the same afferent neuron with different delays through different synaptic connections.
-Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
-as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
-This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{(Right)} for an illustration of the connectivity of one neuron with synaptic weights and delays).
-
-Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$, or the emitting field of a presynaptic neuron $\synapse_\presynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$.
-As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field.
-It results in a weighted event stream (\seeFig{izhikevich}-\textit{(Middle-Left}) for each  postsynaptic neuron $\postsynaddr$:
+%----------------------------------------------------------------------------------------
+%	REFERENCES
+%----------------------------------------------------------------------------------------
+\vspace{-1cm}
+\printbibliography
+%----------------------------------------------------------------------------------------
+%	RESULTS
+%----------------------------------------------------------------------------------------
+\section*{\color{\colorsec}Results \vspace{.2cm}\hrule}
+% ------------------------------------------------------
+\vspace{-2.5cm}
+\subsection*{\color{\colorsubsec}Learning hetero-synaptic delays}
 %
-\begin{linenomath*}
-\begin{equation}\label{eq:stream_b}
-\event_\postsynaddr = \{(\presynaddr_\arank, \synapticweight_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}
-\end{equation}
-\end{linenomath*}
-%
-In biology, this new stream of events is naturally ordered in time as events reach the soma of postsynaptic neurons.
-However, it should be properly reordered in simulations.
-Crucially, when postsynaptic neurons are activated on their soma by a specific spatio-temporal motif which is imprinted in the set of synapses, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner while the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions (\seeFig{izhikevich}).
-We defined the HD-SNN model in this subsection in all generality, and in the next subsection, we describe an implementation of our model adapted for the motion detection task.
-%
-\subsection{Application of the HD-SNN model to the motion detection task}
-%
-It is possible to define a specific implementation of this model in order to adapt it to common tasks in computer vision.
-In particular, from the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation.
-Indeed, by discretizing time, we can transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.
-The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
-First, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:kernel_b} corresponds to a temporal convolution that transforms the input $A$ using one kernel per postsynaptic neuron~\citep{grimaldi_learning_2022}.
-To take advantage of the position invariance observed in images and exploited in convolutional neural networks, we can further assume that synaptic motifs should be similar across different positions, so we can define a spatio-temporal convolutional operator.
-
-Therefore, if we make the approximation at any given time that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
-The longest synaptic delay defines the depth $\Ktime$ of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented.
-In particular, the whole synaptic set can be represented as one kernel for each class $\class$ of the supervision task as the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
-To keep an analogy with the HD-SNN model, for neuron $\postsynaddr$ of position $(x_\postsynaddr, y_\postsynaddr)$ and class $\class$, $\kernel_\class$ gives the weights as a function of relative addresses and synaptic delays: $\forall {\delta_x \in [1, \Kx], \delta_y \in [1, \Ky], \delta_\timev \in [1, \Ktime], \polev \in [1, \Npol]}, \; \kernel_\class(\polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$.
-The connectivity of neuron $\postsynaddr$ is defined locally, around it position $(x_\postsynaddr, y_\postsynaddr)$.
-Using this dense representation, the processing of $A$ by the model can be written as: %
-%
-\begin{linenomath*}
-\begin{equation}\label{eq:kernel_b}
-\forall x, y, \timev, \; \;
-\mathcal{C}_\class(x, y, \timev)
-= \sum_{p, \delta_x, \delta_y, \synapticdelay} \kernel_\class(p, \delta_x, \delta_y, \synapticdelay) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay)
-\end{equation}
-\end{linenomath*}
-%
-where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses inside a kernel and $\delta_\timev$ is the synaptic delay. 
-This shows that $\mathcal{C}_\class$ is the result of a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration).
-Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
-This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
-A similar type of spatio-temporal filtering is used as a pre-processing stage for a pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
-Also,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm which implements a motion estimation task.
-By assuming a locally constant velocity, the authors assume the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.
-
-Here, we keep the analogy with spiking neurons and we try to observe the detection of spiking motifs using the spatio-temporal kernels.
-If so, this precise spatio-temporal patterns prove to be of interest for neural computations and one can infer that biological neurons make use of this information as well.
-% \note{plus de biblio sur la motion detection/segmentation ou optique flow avec les events? (reviewer 2)}
-The 3D convolution represents the linear integration of the spike train as the linear input to the neuron.
-Then, the activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Takes-All network~\citep{nessler_bayesian_2013}.
-In our MLR model, a probability value for each class (i.e.
-each motion direction) is predicted for each position $x, y$ and time $\timev$ as a softmax function of the linear combination of the list of events.
-% on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
-The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the HD-SNN model.
-Formally, using the kernels, it transforms the input raster plot into a probability with the following formula:
-%
-\begin{linenomath*}
-\begin{equation}\label{eq:mlr}
-    \forall x, y, \timev, \forall \class \in [1, N_\class], \; \;
-Pr(k=\class \; \vert \;  x, y, \timev) =
-\frac {\exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class) }{Z(x, y, \timev)}
-\end{equation} 
-\end{linenomath*}
-%
-where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function and  $\bias_\class$ is the bias linked to class $\class$. 
-In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
-The spiking output of the model corresponds to an event with the highest probability class.
-
-Now that this general framework has been explained, we can add some heuristics, based on neuroscientific observations, to constrain our model and its strategies for solving the task that is used to solve the ecological task described in section~\ref{sec:task}.
-Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel heuristics inspired by neuroscience are the main methodological differences with that previous study.
-First, in our simulations, we set the size of the kernels to $(21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = 8 \times 3$.
-To avoid introducing biases in the directions, we apply a circular mask to the spatial dimensions of the kernels.
-Furthermore, we included a prior in the motions that could be selected, as there is a prior in natural scenes for slow speeds~\citep{vacher_bayesian_2018}.
-Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we thus apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration).
-This is consistent with neuroscientific principles because, due to the limits of the conduction delay along horizontal connections, the synaptic delay is related, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one.
-Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities arrangements.
-This is due to the fact that our whitened images have a symmetry in the luminance profiles, and an image with inverted contrast is indistinguishable.
-Since this arrangement of polarities is independent of speed, we added a mechanism that collects the linear values for the movie and the movie with the ON and OFF cells flipped, keeping only the maximum value for each voxel.
-This is similar to the computation done for complex cells in primary visual cortex.
-
-\begin{figure}%[h!]%[ht!]
-    \centering
-    \includegraphics[width=\linewidth]{figures/HDSNN_conv.pdf}
-    % \includegraphics[width=\linewidth]{figures/HD-SNN.png}
-    \caption{
-    % \textbf{- HD-SNN illustration.
-From left to right:} A 3D representation of the event stream as output of the DVS.
-A 3D convolution is applied, on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis.
-There are as many kernels as the possible number of motion classes in the recording.
-For clarity, only $4$ kernels are illustrated and some weights are not represented in the voxel grid.
-For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e.
-softmax function).
-The output of the MLR gives a probability for each class, associated to a specific kernel, to emit an output event, or spike.
-With an appropriate spiking mechanism, here a winner-takes-all, we obtain, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class.
-}
-    \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot here as a raster plot a 2D representation of the input event stream (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time).
-A spatio-temporal convolution is applied to the dense representation of the input with $2$ distinct convolutional kernels (the green and orange kernels) that will define the output channels.
-The convolution is summed over the two polarities.
-If you have two axes $X$ and $Y$ to represent the presynaptic addresses as for the pixel grid of a DVS, you obtain a 3D convolution.
-We restrict the illustration to a 2D representation and to 2 possible classes (green and orange) that are associated with different motion directions.
-\textit{(Middle)}~For each position (address, time), one can compute the activation resulting from the convolution.
-The output of the convolution is processed by the nonlinearity of the MLR model (i.e., the softmax function).
-The output of the MLR gives a probability for each class associated with a particular kernel (colored bars in the highlighted pixel).
-\textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated to a thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated to a specific motion class.
-Note that the position of the output spikes does not systematically correspond to the position of the input spikes but only when enough evidence is obtained.}
-    \label{fig:model}
-\end{figure}
-
-\subsection{Supervised learning of the motion detection task}
-%
-Since the model is fully differentiable, we can now implement a supervised learning rule.
-This rule was implemented using the binary input events as inputs and the corresponding motion direction labels as the desired output.
-The loss function of the MLR model is the binary cross entropy at the output of the classification layer.
-The labels were defined at each time point as a one-hot encoding of the current motion in the channel corresponding to the current motion for all positions.
-Note that in this context, the label is known but the position is not, mainly due to the sparse spatial content of natural images.
-However, the supervised optimization of this MLR model will involve adjusting the weights of the kernels.
-As a result, the error is only propagated back to the spatial locations of these most active cells.
-This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation.
-Simulations are performed with the \note{PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$)}.
-We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
-
-Finally, the output of the MLR model is an event-based representation that predicts the probability of each motion at each position and time.
-Such an output provides a form of optical flow that can be exploited for non-rigid motion, but we have defined here, for simplicity, an evaluation method that applies to our full-field motion task.
-We have shown above that if different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities.
-Thus, by taking the mean logit of the probability of the output given by the model at all positions for any given time, we can calculate the probability of the output at this time.
-This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted).
-These calculations are performed on a different input dataset than the one used in the training or validation steps.
-The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}.
-%
-%
-\section{Results}
-\label{sec:results}
-%%%-----------------------------------------------------------------
-%: fig:kernels
-%%%-----------------------------------------------------------------
-\subsection{Kernels learned for motion detection}
-%%%-----------------------------------------------------------------
-\begin{figure*}[ht!]
+\vspace{-1cm}
+\begin{figure}[H]
     {\centering
@@
     \label{fig:kernels}
-\end{figure*} 
-%
-%
-After training our model, we first analyze the weights learned for the different neurons (\seeFig{kernels}).
-We first observe that these cells are very localized along each trajectory, yet show an orientation selectivity, similar to that observed in MT neurons~\citep{deangelis_functional_1999}.
-Focusing on the positive weights, a strong selectivity is observed along specific axes of motion for each of the different kernels.
-These directions can be easily associated to the direction of motion controlled in the natural images.
-For instance, the first kernel shows a strong selectivity to a downward motion direction.
-We also observe observed a high dependence between the weights reaching the ON polarities and that reaching the OFF polarities.
-In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative in the other.
-This property comes from the way the events are generated and that the luminance can not at the same time increase and decrease.
-Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results observed after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
-%ith the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
-%
-%This qualitative look at the 3D kernels allows the reader to infer for the $8$ different motion directions used to generate our synthetic event streams.
-
-If one focuses on the interpretation of these kernels in terms of spatio-temporal motifs embedded in the event stream, it can lead to interesting outcomes.
-In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions.
-Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spiking motif.
-Each voxel in the 3D kernels defines a specific timestamp and a specific address.
-Consequently, our model is able to detect precise spatio-temporal motifs embedded in the spike train and associated to the different motion directions.
-The cone shape for the positive weights' distribution highlights a loss of precision for longer delays, i.e.
-events away in the past.
-For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Fig.~\ref{fig:kernels}.
-
-%
-% \note{kernel size has to be adapted to the characteristics of motion see Samonds %~\citep{samonds}
-% say that we have tested }
-% \note{tester la meme expe faite par baudot et modelisée par kremkow en montrant des raster plots du modele avec NI versus Gratings }
-
-%%%-----------------------------------------------------------------
-%
-\subsection{Testing with natural-like textures}
-%%%-----------------------------------------------------------------
-%%%-----------------------------------------------------------------
-%: fig:motion_clouds
-%%%-----------------------------------------------------------------
-\begin{figure}%[h!]
+\end{figure} 
+% ------------------------------------------------------
+\subsection*{\color{\colorsubsec}Accuracy as a function of the number of computations}
+% ------------------------------------------------------
+\begin{figure}[H]
     \centering
@@
 \end{figure}
-%%%-----------------------------------------------------------------
-To test our model, we will quantify its ability to categorize different motions, that is on event streams for which the ground truth motion is known at every instant in time.
-Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli.
-In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
-In particular, we will set the spatial extent and duration of the generated movies similarly to the motion task defined above.
-This procedure defines a set of textures with different spatial properties and different motions picked from the same set of $8$ directions and $3$ speeds.
-For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content to provide with some naturalistic variability.
-This method provides a rich dataset of textured movies for which we know the ground truth for motion.
-
-We observe some facts of interest.
-First, as we change the mean spatial frequency of the texture, we observe a monotonous decrease in accuracy.
-This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006} and revels the most informative scales learned by our model.
-We observe the tendency for different spatial frequency bandwidth.
-Notably, the accuracy is better for a grating-like stimulation than for a large (which qualitatively resemble a more textured stimulus), reminiscent to the behavioral response of humans to such stimuli~\citep{simoncini_more_2012}.
-Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.
-When the stimulus is grating-like and that the orientation is arbitrary with respect to the direction of motion, the system is faced with the aperture problem and see a decrease of accuracy.
-This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion.
-Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter.
-This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018} but also that accuracy is low for a rigid motion which lacks variability.
+%----------------------------------------------------------------------------------------
+%	CONCLUSIONS
+%----------------------------------------------------------------------------------------
+\section*{\color{\colorsec}Conclusion \vspace{.2cm} \hrule}
+\vspace{-2.5cm}
 %
-\subsection{Accuracy efficiency trade-off}% for the motion detection task}
-%%%-----------------------------------------------------------------
-Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection.
-We observed that the distribution of the kernels' weights is sparse, with most values near zero (\seeFig{kernels}).
-As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the computations used for the convolution operation.
-In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel.
-If the support of information is sparse, then computations can be performed only on those events.
-Also, if we set some weights of the kernels to zero, then the sum can be skipped for those addresses.
-Knowing the sparseness of the input, the total number of computations thus scales with the number of spikes multiplied by the number of nonzero synaptic weights.
-% , which is proportional to the number of spikes multiplied by the number of synapses
-
-To assess the robustness of the classification as a function of the computational load, we will prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold.
-In Fig.~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations, or active weights, per decision for each neuron of the layer.
-As a comparison and to account for the gain in performance by using heterogeneous delays, we provide the accuracy obtained with a MLR model using 2D time surface (in red) as in~\citep{grimaldi_robust_2022}.
-This latter method is based on delays from the last recorded events and uses fewer computations (in our case $17\times17$) than the dense 3D kernels without any pruning ($21\times17\times17$).
-While less computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
-
-By pruning weights, we observe that the evolution of accuracy as a function of the log percentage of active weights fits well a sigmoid curve.
-Half-saturation level is reached at a ratio of about $3.2\times 10^{-2}$ of active weights, corresponding in our setting to a total amount of $180$ computations per decision.
-Compared to the full kernels, the accuracy of our method is maintained to its top performances when dividing the number of computations by a factor up to about $31$.
-In this case, the number of computations is greatly reduced compared to~\citep{grimaldi_robust_2022}, thus demonstrating the efficiency of the presented method.
+We have introduced a generic SNN using hetero-synaptic delays and shown how it compares favorably with a state-of-the-art event-based algorithm used for classification~\cite{Lagorce17}.
+This shows that we may use the precise timing of a spike to enhance neural computations.
 %
-%%%-----------------------------------------------------------------
-%: fig:accuracy
-%%%-----------------------------------------------------------------
-\begin{figure}%[h!]
-    \centering
-    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
-    \caption{Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
-The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
-As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about 25 times fewer computations.}
-    \label{fig:accuracy}
-\end{figure}
+One advantage of our model is the generality of the approach.
+Indeed, this supervised learning scheme can be extended to a novel task by defining a new set of supervision pairs (for instance supervised by local orientation) which would lead to the emergence of new kernels adapted to this new task.
+This constitutes a major advantage over other algorithms which derive event-based algorithms from specific physical rules.
+%(see for instance~\cite{Benosman12} for computing the optic flow using the luminance conservation rule).
+We aim at extending the application of this model on more generic datasets acquired in natural conditions for progressively more complex tasks such as time-to-contact maps, but also to neurophysiological data.
 %
-\section{Discussion}
-%%%-----------------------------------------------------------------
-In this paper, we have introduced a generic SNN with heterogeneous delays and shown how it compares favorably to a state-of-the-art event-based classification algorithm for a visual motion detection task.
-The learned model shares a number of similarities with neurobiological anatomical observations, as well as with behavioral results.
-The event-based computations of our method can be drastically reduced by pruning synapses, while maintaining top classification performance.
-This shows that we can take advantage of the precise timing of spikes to improve the performance of neural computations.
+\vspace{1cm}
+\hrule
+%----------------------------------------------------------------------------------------
+%	ACKNOWLEDGEMENTS
+%----------------------------------------------------------------------------------------
+\vspace{-1cm}
+\section*{Acknowledgments}
+\vspace{-1cm}
+\Acknowledgments
+%----------------------------------------------------------------------------------------
+%\renewcommand*{\bibfont}{\normalfont\fontsize{20}{12}\selectfont}
 %
-\subsection{Synthesis and main contributions}
-%%%-----------------------------------------------------------------
-The HD-SNN model that has been trained and evaluated on a naturalistic motion detection task.
-The model has been defined to provide optimal detection of event-driven spatio-temporal motifs.
-We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in the early visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
-%We then demonstrated that the model shows similarities to biological responses.
-We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware.
-We show that the use of heterogeneous delays may be an efficient computational solution for future neuromorphic hardware, but also a key to understanding why spikes are a universal component of neural information processing.
-
-We would like to highlight a few innovations in the contributions that are presented in this paper.
-First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs.
-Moreover, in comparison to a representation with that algorithm, the parameters of the model (weights and delays) are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection for each spatio-temporal spike motif.
-Another novelty is that the model learns the weights and the delays simultaneously.
-For example, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn and their values are frozen during learning.
-In addition, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified toy problems~\citep{gutig_tempotron_2006}.
-Another major contribution is to provide a model that is suitable for learning any kind of spatio-temporal spiking motifs and that can be trained in a supervised manner by providing a dataset of supervision pairs.
-Instead of relying on a careful description of the physical rules governing a task, e.g.
-the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}, this allows a more flexible definition of the model using this properly labelled dataset.
+%\printbibliography
+\end{multicols}
 %
-\subsection{Main limits}
-%%%-----------------------------------------------------------------
-We have identified a number of limitations of our model, which we will now discuss in detail.
-First, the entire framework is based on a discrete binning of time, which is not compatible with the continuous nature of biological time.
-We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, to be able to use fast three-dimensional convolutions.
-We have tested the effect of the width of the time bin and shown that it has essentially no effect on the results presented in this paper.
-This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}.
-This suggests the possibility of analytically including a precision term in the temporal value of the input spikes.
-This mechanism may be implemented by the filtering implemented by the synaptic time constant of about $5~\ms$.
-Furthermore, it is possible to circumvent the need for time discretization by the use of a purely event-based scheme.
-In fact, it is not necessary to compute the voltage traces between two spikes~\citep{hanuschkin_general_2010}.
-Thus, it is possible to define a purely event-based framework.
-Such an architecture could provide promising computational speedups.
-% \note{pourrait parler du poster bernstein?
-de nadafian?}
-
-A further limitation is that the model is purely feed-forward.
-Thus, the spikes generated by the postsynaptic neurons are based solely on information contained in the classical receptive field.
-However, it is well known that neurons in the same layer can interact with each other using lateral interactions, for example in V1, and that this can be the basis for computational principles~\citep{chavane_revisiting_2022}.
-For example, the combination of neighboring orientations may contribute to image categorization~\citep{perrinet_edge_2015}.
-Furthermore, neural information is modulated by feedback information, e.g.
-to distinguish a figure from its background~\citep{roelfsema_early_2016}.
-Feedback has been shown to be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020}, especially to explain non-linear mechanisms~\citep{boutin_pooling_2022}.
-Currently, mainly due to our use of convolutions, it is not possible to implement these recurrent connections in our implementation (lateral or feedback).
-However, by inserting new spikes into the list of spikes reaching presynaptic addresses, the generic model is able to incorporate them.
-While theoretically possible, this needs to be properly adjusted in practice so that these recurrent connections do not amplify neuronal activity outside a homeostatic state (by extinction or explosion).
-% \note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}%~\citep{perrinet_motion-based_2012}
-
-Such recurrent activity would be essential for the implementation of predictive or anticipatory processes.
-This is essential in a neural system because it contains several different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}.
-This has been modeled before to explain, for example, the flash-lag illusion~\citep{khoei_flash-lag_2017}.
-%\note{Say that this could be implemented by properly setting the timing of the label}
-As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by velocity, acceleration, jerk,~\ldots), and ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
-Our proposed model using heterogeneous delays provides an alternative and elegant implementation solution to this problem.
-%
-\subsection{Perspectives}
-%%%-----------------------------------------------------------------
-% \note{la tache ne met pas en avant les capacites des motifs hetero synaptiques =trop simple et peut se faire en HOTS}
-In defining our task, we emphasized that the generation of events depends on the spatial gradient in each image.
-This gradient has both horizontal and vertical dimensions, and its maxima are generally orientation dependent.
-Taken together, these oriented edges form the contours of visual objects in the scene~\citep{koenderink_representation_1987}.
-Thus, there is an interdependence between the information about motion and the information about orientation within the event stream.
-It would be crucial to investigate this dependency further.
-This could be initiated by training the model on a dataset with labels that provide local orientation.
-%\note{we can say that we tried it and that it works well}.
-Exploring this dependence will allow us to dissociate these two forms of visual information and enable us to integrate them.
-In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka the aperture problem).
-Thus, it will allow us to implement recurrent prediction rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
-
-The model is trained on a low-level local motion detection task, and one might wonder if it could be trained on higher-level tasks.
-An example of such a task would be the estimation of depth in the visual scene.
-There are several sources of information for depth estimation, such as binocular disparity or changes in texture or shading, but in our case motion parallax would be the most important cue~\citep{rogers_motion_1979}.
-This is because objects that are close to an observer will move relatively faster on the retina than an object that is far away, and also because visual occlusions are dependent on the depth order.
-Using this information, it is possible to segment objects and estimate their depth~\citep{yoonessi_contribution_2011}.
-However, this would require the computation of the optical flow first, i.e., the extension of the framework described here for a rigid full-field motion to a more general one where the motion may vary in the visual field.
-A possible implementation is therefore to add a new layer to our model, analogous to the hierarchical organization highlighted in the visual cortex.
-This is theoretically possible by using the output of our model (which estimates velocity in retinotopic space) as input to a new layer of neurons that would estimate velocity in the visual field, including the depth dimension in the output supervision labels.
-This could have direct and important applications, e.g.
-in autonomous driving to detect obstacles in a fast and robust way.
-Another extension would be to actively generate sensor motion (physical or virtual) to obtain better depth estimates, especially to disambiguate uncertain estimates~\citep{nawrot_eye_2003}.
-
-In conclusion, the model that we have presented provides a way to process event-based signals in an efficient manner.
-We have shown that we can train the model semi-supervised, knowing \emph{what} output label, but not \emph{when} it occurs.
-Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data~\citep{barlow_unsupervised_1989}.
-This type of learning is thought to be prevalent in the central nervous system and, assuming the signal is sparse~\citep{olshausen_emergence_1996}, one could extend these Hebbian sparse learning schemes to spikes~\citep{perrinet_emergence_2004, masquelier_competitive_2009}.
-We expect that this would be particularly useful for exploring neurobiological data.
-Indeed, there is a large literature showing that brain dynamics often organize into stereotyped sequences, such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}.
-These patterns are stereotyped and robust, as they can be activated in the same pattern from day to day~\citep{haimerl_internal_2019}.
-In contrast to conventional methods of processing neurobiological data, such an event-based model would be able to answer key questions about the representation of information in neurobiological data, and it would open up possibilities in the field of computational neuroscience.
-Furthermore, it would open up possibilities in the field of machine learning, especially in computer vision, to address current key concerns such as robustness to attacks, scalability, interpretability, or energy consumption.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%
-\backmatter
-%
-\bmhead{Acknowledgments}
-%
-\Acknowledgments
-%
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section*{Statements and Declarations}
-
-\paragraph{Funding}
-
-\Funding %
-
-\paragraph{Conflict of interest}
-Not applicable.
-
-\paragraph{Ethics approval}
-Not applicable.
-
-\paragraph{Consent to participate}
-Not applicable.
-
-\paragraph{Consent for publication}
-Not applicable.
-
-\paragraph{Availability of data and materials}
-Not applicable.
-
-\paragraph{Code availability}
-%
-\DataAvailability
-%   
-\paragraph{Authors' contributions}
-%
-Both authors contributed to the conceptualization and methodology design of the study, to the project's coordination and administration.
-Laurent Perrinet carried out the funding acquisition and supervision.
-Formal analysis and investigation were performed by both authors.
-Results visualization and presentation were realized by both authors.
-The manuscript was written by both authors.
-Both authors have read and approved the final manuscript.
-%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{apalike}
-\bibliography{FastMotionDetection}
 \end{document}
-%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-03-27 2023-03-28_HBP_summit.tex
Update on Overleaf.

@@
 \newcommand{\Abstract}{
-The response of a biological neuron depends on the precise timing of afferent spikes.
+
+In a recent review paper, we reported experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues and neuromorphic data~\cite{grimaldi_precise_2023}.
+
+Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
+
+We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
+We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or to event-based cameras and for which we will characterize the accuracy of the model in detecting visual motion.
+
+%The response of a biological neuron depends on the precise timing of afferent spikes.
 This temporal aspect of the neural code is essential to understand information processing in neurobiology and applies particularly well to the output of neuromorphic hardware such as event-based cameras.

### 2023-03-27 2023-03-28_HBP_summit.tex
new figure

@@
 \newcommand{\Abstract}{
+\begin{itemize}
+\item In a recent review paper, we reported experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues and neuromorphic data~\cite{grimaldi_precise_2023}.
 
-In a recent review paper, we reported experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues and neuromorphic data~\cite{grimaldi_precise_2023}.
+\item Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
 
-Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
-
-We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
+\item We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
 We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or to event-based cameras and for which we will characterize the accuracy of the model in detecting visual motion.
-
+\end{itemize}
 %The response of a biological neuron depends on the precise timing of afferent spikes.
@@
 \newcommand{\bias}{b} % bias for the MLR model
-\newcommand{\synapse}{\sigma} % synapse
+\newcommand{\synapse}{\mathcal{S}} % synapse
 \newcommand{\synapticweight}{w} % synaptic weight
-\newcommand{\synapticdelay}{d} % synaptic delay
+\newcommand{\synapticdelay}{\delta} % synaptic delay
 \newcommand{\ranksyn}{s} % synapse index
@@
 
-\bibliography{references.bib}
+\bibliography{FastMotionDetection.bib}
 
@@
 $$
-\mybox{$\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$}
+\mybox{$\event = \{(\presynaddr_\arank, 1, \timev_\arank)\}_{\arank \in [1,\numevent]}$}
 $$
@@
 $$
-\mybox{$\synapse^\postsynaddr = \{(\presynaddr^\postsynaddr_\ranksyn, \synapticweight^\postsynaddr_\ranksyn, \synapticdelay^\postsynaddr_\ranksyn)\}_{\ranksyn \in [1,\Nsyn^\postsynaddr]}$}
+\mybox{$\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$}
 $$
-is a set of $\Nsyn^\postsynaddr$ synapses, associated to neuron $\postsynaddr$, where each synapse $\synapse^\postsynaddr_\ranksyn$ is associated to a weight $\synapticweight^\postsynaddr_\ranksyn$, a delay $\synapticdelay^\postsynaddr_\ranksyn$ and a presynaptic address $\presynaddr^\postsynaddr_\ranksyn$.\\
+is a set of $\Nsyn$ synapses, where each synapse $\synapse_\ranksyn$ is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$ and a delay $\synapticdelay_\ranksyn$.
+We can define the synapses of the receptive field of neuron $\postsynaddr$ as such: $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$.
+They transform the event stream as input into: 
+$$
+\mybox{$\event_\postsynaddr = \{(\presynaddr_\arank, \synapticweight_\arank, \timev_\arank+\synapticdelay_\ranksyn) \| \presynaddr_\arank = \presynaddr_\ranksyn \}_{\arank \in [1,\numevent], \ranksyn \in \synapse^\postsynaddr}$}
+$$
+
+%
+\columnbreak
+%
+\noindent
+\textbf{Temporal convolution:}
+
+\noindent By discretizing time, we can transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ where the ones define the events.
+Using this dense representation, the processing of $A$ by neuron $\postsynaddr$ can be written as: %
+%
+$$
+\mybox{$\forall \timev, \; \;
+\mathcal{C}_\postsynaddr(x_\postsynaddr, y_\postsynaddr, \timev)
+= \sum_{p, \delta_x, \delta_y, \synapticdelay_\timev} \kernel_\postsynaddr(p, \delta_x, \delta_y, \synapticdelay_\timev) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay_\timev)
+$}
+$$
+%
+where $\kernel_\postsynaddr$ gives the weights as a function of relative addresses, $\delta_x$ and $\delta_y$, and synaptic delays $\delta_\timev$.\\
 %
 \\
-\textbf{Active weights:}
-%
+\textbf{Multinomial logistic regression:}
+
+To take advantage of the position invariance observed in images and exploited in convolutional neural networks, we can further assume that synaptic motifs should be similar across different positions, so we can define a spatio-temporal convolutional operator: $\mathcal{C}_\class = \kernel_\class \ast A$, where $\class$ combines different postsynaptic neurons sharing the same class.
+Then, the activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression, in analogy to a spiking Winner-Takes-All network~\cite{nessler_bayesian_2013}:
 $$
-\mybox{$\activeweights^\postsynaddr(t) = \{\synapticweight_\ranksyn^\postsynaddr | \presynaddr_\arank = \presynaddr_\ranksyn^\postsynaddr \; \text{and} \; \timev=\timev_\arank + \synapticdelay_\ranksyn^\postsynaddr\}_{\arank \in [1,\numevent]\text{, }\ranksyn \in [1,\Nsyn^\postsynaddr]}$}
+\mybox{
+$ \forall x, y, \timev, \forall \class \in [1, N_\class], \; \;
+Pr(k=\class \; \vert \;  x, y, \timev) =
+\frac {\exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class) }{Z(x, y, \timev)}
+$
+}
 $$
 %
-\textbf{Activation function (Multinomial Logistic Regression):}
-%
-$$
-\mybox{$Pr(k=\postsynaddr \; | \; \timev) =\frac 1 Z \cdot {\exp  (\mathcal{C}^\postsynaddr(\timev) +\bias^\postsynaddr) }$}
-$$
-where $\mathcal{C}^\postsynaddr(\timev) = \sum \activeweights^\postsynaddr(t)
-$ is the sum of the synaptic weights and $\bias^\postsynaddr$ is the bias linked to neuron $\postsynaddr$ and $Z= \sum_{c=1}^{\Nclass} Pr(k=c \; | \; \timev)$.\\
-%
-\\
-\textbf{Temporal convolution:}\\
-In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in [1,\Nsyn^\postsynaddr]}, \kernel^\postsynaddr(\presynaddr_\ranksyn^\postsynaddr, \synapticdelay_\ranksyn^\postsynaddr) = \synapticweight_\ranksyn^\postsynaddr$ (see Figure~\ref{fig:kernels}).
-Using this dense representation, the counting $\mathcal{C}^\postsynaddr(\timev)$ defined above can be computed as a temporal convolution of the dense representation of the event stream $E$ with the dense kernels formed by the set of synapses:
-$$
-\mybox{$\mathcal{C}^\postsynaddr = \kernel^\postsynaddr \ast E$}
-$$
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{figures/HDSNN_conv.pdf}
+    % \includegraphics[width=\linewidth]{figures/HD-SNN.png}
+    \caption{
+    % \textbf{- HD-SNN illustration.
+From left to right:} A 3D representation of the event stream as output of the DVS.
+A 3D convolution is applied, on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis.
+There are as many kernels as the possible number of motion classes in the recording.
+For clarity, only $4$ kernels are illustrated and some weights are not represented in the voxel grid.
+For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e.
+softmax function).
+The output of the MLR gives a probability for each class, associated to a specific kernel, to emit an output event, or spike.
+With an appropriate spiking mechanism, here a winner-takes-all, we obtain, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class.
+}
+    \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot here as a raster plot a 2D representation of the input event stream (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time).
+A spatio-temporal convolution is applied to the dense representation of the input with $2$ distinct convolutional kernels (the green and orange kernels) that will define the output channels.
+The convolution is summed over the two polarities.
+If you have two axes $X$ and $Y$ to represent the presynaptic addresses as for the pixel grid of a DVS, you obtain a 3D convolution.
+We restrict the illustration to a 2D representation and to 2 possible classes (green and orange) that are associated with different motion directions.
+\textit{(Middle)}~For each position (address, time), one can compute the activation resulting from the convolution.
+The output of the convolution is processed by the nonlinearity of the MLR model (i.e., the softmax function).
+The output of the MLR gives a probability for each class associated with a particular kernel (colored bars in the highlighted pixel).
+\textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated to a thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated to a specific motion class.
+Note that the position of the output spikes does not systematically correspond to the position of the input spikes but only when enough evidence is obtained.}
+    \label{fig:model}
+\end{figure}
 %----------------------------------------------------------------------------------------
@@
 %
-We have introduced a generic SNN using hetero-synaptic delays and shown how it compares favorably with a state-of-the-art event-based algorithm used for classification~\cite{Lagorce17}.
+We have introduced a generic SNN using hetero-synaptic delays and shown how it compares favorably with a state-of-the-art event-based algorithm used for classification~\cite{lagorce_hots_2017}.
 This shows that we may use the precise timing of a spike to enhance neural computations.

### 2023-03-28 2023-03-28_HBP_summit.tex
Update on Overleaf.

@@
 \newcommand{\Acknowledgments}{AG and LP received funding from the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{http://aprovis3d.eu/}{APROVIS3D}'').
-HL and LP received funding from the ANR project  N° ANR-20-CE23-0021 (''\href{https://laurentperrinet.github.io/grant/anr-anr/}{AgileNeuroBot}'').
+LP received funding from the ANR project  N° ANR-20-CE23-0021 (''\href{https://laurentperrinet.github.io/grant/anr-anr/}{AgileNeuroBot}'').
 %HL received funding from a University of Montreal Artificial Intelligence scholarship.
@@
 \Large
-\href{https://laurentperrinet.github.io/publication/grimaldi-22-icip/}{https://laurentperrinet.github.io
-\\ /publication/grimaldi-22-icip/} \end{center}
+\href{https://laurentperrinet.github.io/publication/grimaldi-22-bc/}{https://laurentperrinet.github.io
+\\ /publication/grimaldi-22-bc/} \end{center}
 \vspace{2cm}
@@
 %----------------------------------------------------------------------------------------
-\section*{\color{\colorsec}Heterogeneous delays model\hrule}
+\section*{\color{\colorsec}Heterogeneous delays model (HD-SNN)\hrule}
 %
@@
     \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}
-    \caption{\textbf{The core mechanism of the HD SNN model.} \textit{(Top-Left)}~Two spiking motifs are emitted from four presynaptic neurons.
+    \caption{\textbf{The core mechanism of the HD-SNN model.} \textit{(Top-Left)}~Two spiking motifs are emitted from four presynaptic neurons.
 Once integrated by the synapses of the postsynaptic neuron~\textit{(Right)}, the spiking motifs are shifted in time by the synaptic delays and weighted by the synaptic weights~\textit{(Middle-Left)}.
@@
 \\
-\textbf{Multinomial logistic regression:}
+\textbf{Multinomial logistic regression (MLR):}
 
-To take advantage of the position invariance observed in images and exploited in convolutional neural networks, we can further assume that synaptic motifs should be similar across different positions, so we can define a spatio-temporal convolutional operator: $\mathcal{C}_\class = \kernel_\class \ast A$, where $\class$ combines different postsynaptic neurons sharing the same class.
-Then, the activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression, in analogy to a spiking Winner-Takes-All network~\cite{nessler_bayesian_2013}:
+\noindent We can further assume that synaptic motifs should be similar across different positions, so we can define a spatio-temporal convolutional operator: $\mathcal{C}_\class = \kernel_\class \ast A$, where $\class$ combines different postsynaptic neurons sharing the same class.
+Then, the activation function of our model is a softmax function implementing a form of MLR, in analogy to a spiking Winner-Takes-All network~\cite{nessler_bayesian_2013}:
 $$
@@
 $$
+%
+\vspace{1cm}
 %
@@
 }
-    \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot here as a raster plot a 2D representation of the input event stream (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time).
-A spatio-temporal convolution is applied to the dense representation of the input with $2$ distinct convolutional kernels (the green and orange kernels) that will define the output channels.
-The convolution is summed over the two polarities.
-If you have two axes $X$ and $Y$ to represent the presynaptic addresses as for the pixel grid of a DVS, you obtain a 3D convolution.
-We restrict the illustration to a 2D representation and to 2 possible classes (green and orange) that are associated with different motion directions.
-\textit{(Middle)}~For each position (address, time), one can compute the activation resulting from the convolution.
-The output of the convolution is processed by the nonlinearity of the MLR model (i.e., the softmax function).
-The output of the MLR gives a probability for each class associated with a particular kernel (colored bars in the highlighted pixel).
-\textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated to a thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated to a specific motion class.
-Note that the position of the output spikes does not systematically correspond to the position of the input spikes but only when enough evidence is obtained.}
+    \textbf{Applying HD-SNN to the motion detection task.} \textit{(Left)}~We plot here a 2D representation of the input event stream.
+A spatio-temporal convolution is applied to the dense representation of the input with $2$ distinct convolutional kernels (the green and orange kernels) that will define the output channels (if you have two axes $X$ and $Y$ to represent the presynaptic addresses as for the pixel grid of a DVS, you obtain a 3D convolution).
+\textit{(Middle)}~The output of the convolution is processed by the nonlinearity of the MLR model (i.e., the softmax function).
+\textit{(Right)}~By adding a spiking mechanism, we obtain a new spike train with the spikes associated to the different motion classes.}
     \label{fig:model}
@@
 %
-\vspace{-2.5cm}
-To test our model, we will quantify its ability to categorize different motions.
-In that order, we will first define a set of synthetic stimuli, natural images with a rigid motion or also \textit{Motion Clouds}~\cite{leon2012motion}, which are natural-like random textures for which we can control for velocity, among other parameters.
+\vspace{-1cm}
 %
@@
 We show such a trajectory with a length of $128$ time steps (green line).
-\textit{(Right)}~Snapshots of the synthetic event stream at different time steps (start and end frames marked by a white and black dot, respectively).
+\textit{(Right)}~Snapshots of the synthetic event stream at different time steps.
 The dynamics of the cropped image translated according to the trajectory as a function of time produces a naturalistic movie, which is then transformed into an event-based representation.
@@
 ON (red) and OFF (blue) events, in each pixel of the image.
-In the lower left corner of the snapshots, the translation vector is shown in red as one of the possible classes of motion.
-Note the change in the direction of motion between the third and fourth image, and also that, due to the aperture problem, contours parallel to the motion emit relatively fewer spikes.}
+In the lower left corner of the snapshots, the translation vector is shown in red as one of the possible classes of motion.}
     \label{fig:motion_task}
@@
 %
-\vspace{1cm}
-\hrule
+%\vspace{1cm}
+\columnbreak
 %
-%----------------------------------------------------------------------------------------
-%	REFERENCES
-%----------------------------------------------------------------------------------------
-\vspace{-1cm}
-\printbibliography
 %----------------------------------------------------------------------------------------
@@
 \vspace{-2.5cm}
-\subsection*{\color{\colorsubsec}Learning hetero-synaptic delays}
+\subsection*{\color{\colorsubsec}Learning heterogeneous delays}
 %
@@
 \begin{figure}[H]
-    {\centering
+    \centering
     %\vspace{-3cm}
-    \includegraphics[width=\linewidth]{figures/motion_kernels.pdf}
-    }
+    \includegraphics[width=.8\linewidth]{figures/motion_kernels.pdf}
     %\vspace{-.5cm}
     \caption{
-    	Representation of the weights for $8$ directions for one speed (among the $8 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
+    	\textbf{Representation of the weights} for $8$ directions for one speed (among the $8 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
 Directions are shown as red arrows on the left.
 The spatio-temporal kernels are represented as slices of spatial weights at different delays.
-Delays vary along the horizontal axis from the far right (delay of one step) to the left (delay of $12$ steps).
-Each image corresponds to the weights at a given delay, with excitatory and inhibitory weights in warm and cold colors, respectively.
-Due to the observed symmetry between the ON and OFF event streams, we report that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective to the different motion directions and we observe for all kernels an orientation preference perpendicular to the orientation.
+Due to the observed symmetry between the ON and OFF event streams, we report that the kernels for the OFF polarities are very similar and are not shown here.
+Different kernels are selective to the different motion directions and we observe for all kernels an orientation preference perpendicular to the orientation.
 %      % These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
@@
     \centering
-    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
-    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of (from left to right) the mean spatial frequency, the bandwidth in spatial frequency (from a grating-like (left) to isotropic textures (right)), the bandwidth in orientation (from isotropic textures (left) to grating-like (right)), the bandwidth in speed (from a rigid motion (left) to independent frames (right)).
-Note that these accuracies are computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) or in the generic case where orientation is independent of direction ('aperture').}
-    \label{fig:motion_clouds}
+    \includegraphics[width=0.6\linewidth]{figures/quant_accuracy.pdf}
+    \caption{\textbf{Accuracy as a function of computational load} for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
+The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
+As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about 25 times fewer computations.}
+    \label{fig:accuracy}
 \end{figure}
@@
 %----------------------------------------------------------------------------------------
-\section*{\color{\colorsec}Conclusion \vspace{.2cm} \hrule}
-\vspace{-2.5cm}
+%\section*{\color{\colorsec}Conclusion \vspace{.2cm} \hrule}
+%\vspace{-2.5cm}
 %
-We have introduced a generic SNN using hetero-synaptic delays and shown how it compares favorably with a state-of-the-art event-based algorithm used for classification~\cite{lagorce_hots_2017}.
+%We have introduced a generic SNN using hetero-synaptic delays and shown how it compares favorably with a state-of-the-art event-based algorithm used for classification~\cite{lagorce_hots_2017}.
 This shows that we may use the precise timing of a spike to enhance neural computations.
 %
-One advantage of our model is the generality of the approach.
-Indeed, this supervised learning scheme can be extended to a novel task by defining a new set of supervision pairs (for instance supervised by local orientation) which would lead to the emergence of new kernels adapted to this new task.
-This constitutes a major advantage over other algorithms which derive event-based algorithms from specific physical rules.
+%One advantage of our model is the generality of the approach.
+%Indeed, this supervised learning scheme can be extended to a novel task by defining a new set of supervision pairs (for instance supervised by local orientation) which would lead to the emergence of new kernels adapted to this new task.
+%This constitutes a major advantage over other algorithms which derive event-based algorithms from specific physical rules.
 %(see for instance~\cite{Benosman12} for computing the optic flow using the luminance conservation rule).
-We aim at extending the application of this model on more generic datasets acquired in natural conditions for progressively more complex tasks such as time-to-contact maps, but also to neurophysiological data.
+%We aim at extending the application of this model on more generic datasets acquired in natural conditions for progressively more complex tasks such as time-to-contact maps, but also to neurophysiological data.
 %
@@
 \hrule
+%	REFERENCES
+%----------------------------------------------------------------------------------------
+%\vspace{-1cm}
+\printbibliography
+%----------------------------------------------------------------------------------------
 %----------------------------------------------------------------------------------------
@@
 %----------------------------------------------------------------------------------------
-\vspace{-1cm}
+%\hrule
+%\vspace{-1cm}
 \section*{Acknowledgments}

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
Update Grimaldi-etal-BiolCybernetics.tex

@@
 %● The roles of attention mechanisms in biological vision.
+
+% 2023-05-20: rev1 - final
+% Reviewer 1
+% The author's organization of the paper is much clearer than last time, and I suggest making the following modifications:
+% 1.
+For the natural-like texture images used, the author carried out some operations, such as changing texture parameters, etc.
+I suggest that the authors give 1-2 examples of these images corresponding to the different subgraphs in Figure 5, such as isotropic textures and grating-like textures .
+% 
+% 2. Experiments that assess accuracy should be repeated many times (and give deviation) to eliminate data uncertainty.
+% 
+% 3. I would still recommend that the authors add experiments on other data sets, or compare with other works  on the current data set.
+%
+% 4. There are two “observe” in line 355.
+%
+% In order to process your paper, we require:
+%
+% • A point-by-point response to the comments, including a description of any additional experiments that were carried out and a detailed rebuttal of any criticisms or requested revisions that you disagreed with.
+% 
+% This must be uploaded as a 'Point-by-point response to reviewers' file. All changes to the manuscript must be highlighted or indicated by using tracked changes.
+% 
+% At this stage, please also ensure that you have replaced your initial-submission image files with production quality figures.
+These should be supplied at 300 dpi resolution for .jpeg and .tiff or as .eps files.
+Figures should not include Figure number labels in the image.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Each image corresponds to the weights at a given delay, with excitatory and inhibitory weights in warm and cold colors, respectively.
-Due to the observed symmetry between the ON and OFF event streams, we report that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective to the different motion directions and we observe for all kernels an orientation preference perpendicular to the orientation.
+Due to the symmetry between the ON and OFF event streams, we report that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective to the different motion directions, and we observe for all kernels an orientation preference perpendicular to the orientation.
 %      % These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
@@
 After training our model, we first analyze the weights learned for the different neurons (\seeFig{kernels}).
-We first observe that these cells are very localized along each trajectory, yet show an orientation selectivity, similar to that observed in MT neurons~\citep{deangelis_functional_1999}.
+We first notice that these cells are very localized along each trajectory, yet show an orientation selectivity, similar to that detected in MT neurons~\citep{deangelis_functional_1999}.
 Focusing on the positive weights, a strong selectivity is observed along specific axes of motion for each of the different kernels.
-These directions can be easily associated to the direction of motion controlled in the natural images.
+These directions can be easily associated to the direction of motion controlled in the natural images and show a typical anisotropic profile~\citep{kaplan_anisotropic_2013}.
 For instance, the first kernel shows a strong selectivity to a downward motion direction.
-We also observe observed a high dependence between the weights reaching the ON polarities and that reaching the OFF polarities.
+We also observed a high dependence between the weights reaching the ON polarities and that reaching the OFF polarities.
 In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative in the other.
 This property comes from the way the events are generated and that the luminance can not at the same time increase and decrease.
-Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results observed after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
+Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results obtained after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
 %ith the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
@@
 We observe some facts of interest.
-First, as we change the mean spatial frequency of the texture, we observe a monotonous decrease in accuracy.
-This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006} and revels the most informative scales learned by our model.
+First, as we change the mean spatial frequency of the texture, we notice a monotonous decrease in accuracy.
+This comes as a similar trend as that shown in the primary visual areas~\citep{priebe_tuning_2006} and revels the most informative scales learned by our model.
 We observe the tendency for different spatial frequency bandwidth.

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
abstract

@@
 \newcommand{\Abstract}{
-The response of a biological neuron depends on the precise timing of afferent spikes.
+The response of a biological neuron depends on the precise timing of spikes coming from afferent neurons.
 This temporal aspect of the neural code is essential to understand information processing in neurobiology and applies particularly well to the output of neuromorphic hardware such as event-based cameras.
 However, most artificial neural models do not take advantage of this important temporal dimension of the neural code.
-Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
+Here, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
 The model uses the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of synaptic inputs as they reach the basal dendritic tree.
 We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
-We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or to event-based cameras and for which we will characterize the accuracy of the model in detecting visual motion.
-In particular, we quantify how the accuracy can vary as a function of the overall computational load showing it is still efficient at very low workloads.
-This end-to-end, event-driven computational building block could improve the performance of future spiking neural network (SNN) algorithms and in particular their implementation in neuromorphic chips.
+We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or to event-based cameras.
+We characterize the accuracy of the model in detecting visual motion and in particular, by pruning some weights we show it is still efficient at very low workloads.
+As a conclusion, heterogeneous delays provide with an end-to-end event-driven computational building block which could improve the performance of future spiking neural network (SNN) algorithms and in particular their applications in neuromorphic chips.
 }

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
intro

@@
 %
-The human brain has the remarkable property of being able to react at any time while consuming a reasonable amount of energy, about tens of watts.
+The human brain has the remarkable property of being able to react efficiently, at any time while consuming a reasonable amount of energy, about tens of watts.
 This system is the result of millions of years of natural selection, and a striking difference with artificial neural networks is the representation both use.
-Artificial convolutional neural networks (CNNs), for example, represent the flow of information from one layer to another as tensors, storing visual information densely across the visual topography, with different properties represented in different channels.
-These networks are known to mimic many properties of biological systems, such that each can be attributed a ``Brain Score''~\citep{schrimpf_brain-score_2020}, yet this score takes no account of inference speed or energy consumption.
+Artificial convolutional neural networks (CNNs), for example, represent the flow of information from one layer to another as tensors, storing visual information densely across the visual topography, with different translation-invariant properties represented in different channels.
 CNNs have achieved state-of-the-art performance for some computer vision tasks, such as image recognition.
-However, they do not take advantage of the dynamics inherent in the way we perceive our natural environment.
+These networks are also known to mimic many properties of biological systems, such that each can be attributed a ``Brain Score''~\citep{schrimpf_brain-score_2020}, yet this score takes no account of inference speed or energy consumption.
+Indeed, they do not take advantage of the dynamics inherent to the way we perceive our natural environment.
 In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
-The third generation of artificial neural networks, known as spiking neural networks (SNNs), incorporates this temporal dimension into the way they perform their computations.
+Spiking neural networks (SNNs), known as the third generation of artificial neural networks, incorporate this temporal dimension into the way they perform their computations.
 They are of interest for computational neuroscience because they provide a better model of the biological neuron.
@@
 In a recent review paper, we reported experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues~\cite{grimaldi_precise_2023}.
-These spatio-temporal motifs within the neural activity may be useful representations to perform computations for various cognitive tasks.
+These spatio-temporal motifs which are present within the neural activity may form useful representations to perform computations for various cognitive tasks using the synchronicity of spikes reaching the soma of a neuron.
 Notably, \citet{abeles_role_1982} asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs.
@@
 In neural models, the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows for the efficient exploitation or detection of these spatio-temporal motifs embedded in the spike train.
-In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, but precise, relative spiking delays, i.e., the time between the arrival of an afferent spike at a given synapse and the contribution of the associated postsynaptic potential to the neuron's soma.
-Due to the variety of configurations and the possible coexistence of multiple superimposed motifs, representation with polychronous groups has a much higher information capacity than a firing rate-based approach to neural coding.
+In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a spiking motif defined by a subset of neurons with different, but precise, relative spiking delays.
+Delays are defined as the time between the arrival of an afferent spike at a given synapse and the arrival time of the associated postsynaptic potential on the neuron's soma.
+Due to the variety of configurations and the possible coexistence of multiple superimposed motifs, representations using polychronous groups have mathematically a much higher information capacity than a firing rate-based approach.
 
-The present paper proposes additional experiments extending a recent model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}.
-This model was trained to solve a motion detection task on a synthetic event-based dataset generated by moving parameterized textures.
-Once trained, the volume of event-driven computations could be drastically reduced by pruning the synapses, while maintaining top performance for the classification task.
-This was a demonstration that formal neurons can exploit the precise timing of spikes to improve neural computations.
-In the present work, we extend motion detection to a more ecological task.
+The present paper proposes a real-life application extending a recent model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}.
+Indeed, the model was trained to solve a simplified motion detection task on a synthetic event-based dataset generated by moving parameterized textures.
+Once trained, the volume of event-driven computations could be drastically reduced by pruning synapses, while maintaining top performance for the classification task.
+This constituted a preliminary demonstration that formal neurons could exploit the precise timing of spikes to improve neural computations.
+Here, we first define the ecological cognitive task the model has to solve with the different datasets it will be tested on.
 Instead of the synthetic textures that have been used to generate motion-driven stimuli, we use natural images and random movements that mimic biological saccades.
-Using an ecological motion detection task, we study the emergence of spatio-temporal spiking motifs when a single layer of spiking neurons is trained on a supervised classification task.
-First, we define the ecological cognitive task the model has to solve with the different datasets it will be tested on.
 Then, we develop a theoretically defined Heterogeneous Delays Spiking Neural Network (HD-SNN) model capable of learning these heterogeneous synaptic delays.
+Using this detection task, we study the emergence of spatio-temporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner.
 We will investigate the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
intro

@@
 Spiking neural networks (SNNs), known as the third generation of artificial neural networks, incorporate this temporal dimension into the way they perform their computations.
-They are of interest for computational neuroscience because they provide a better model of the biological neuron.
+They are of interest for computational neuroscience because they provide a better model of biological neuron.
 Some approaches have developed normative models of SNNs that aim to have applications in machine learning.
@@
 Another type of SNN that uses accurate timing of spikes attempts to determine the structure of the network to minimize a cost function.
-This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been recently extended.
-This method uses a surrogate gradient and is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
-However, the performance of SNNs still lags behind that of firing-rate based networks.
+This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been recently extended using a surrogate gradient, a method which is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
+However, this performance of SNNs still lags behind that of networks using an analog, firing-rate based representation.
 The question of the advantage of using spikes in machine learning and computer vision remains open.
@@
 In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a spiking motif defined by a subset of neurons with different, but precise, relative spiking delays.
-Delays are defined as the time between the arrival of an afferent spike at a given synapse and the arrival time of the associated postsynaptic potential on the neuron's soma.
-Due to the variety of configurations and the possible coexistence of multiple superimposed motifs, representations using polychronous groups have mathematically a much higher information capacity than a firing rate-based approach.
+This delay between a pair of connected neurons is defined as the time between the emission of a spike on the afferent neuron's soma and that of its arrival on the efferent neuron's soma.
+Due to the variety of weights and delay within a population, representations using polychronous groups have mathematically a much higher information capacity than a firing rate-based approach.
 
 The present paper proposes a real-life application extending a recent model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}.
-Indeed, the model was trained to solve a simplified motion detection task on a synthetic event-based dataset generated by moving parameterized textures.
-Once trained, the volume of event-driven computations could be drastically reduced by pruning synapses, while maintaining top performance for the classification task.
-This constituted a preliminary demonstration that formal neurons could exploit the precise timing of spikes to improve neural computations.
-Here, we first define the ecological cognitive task the model has to solve with the different datasets it will be tested on.
+Indeed, this previous model was trained to solve a simplified motion detection task on a synthetic event-based dataset generated by moving parameterized textures.
+This constituted a preliminary demonstration that formal neurons could exploit the precise timing of spikes to improve neural computations thanks to heterogeneous delays.
+Here, we extend these results to a more complex and natural framework.
+We first define the ecological cognitive task the model has to solve with the different datasets it will be tested on.
 Instead of the synthetic textures that have been used to generate motion-driven stimuli, we use natural images and random movements that mimic biological saccades.
-Then, we develop a theoretically defined Heterogeneous Delays Spiking Neural Network (HD-SNN) model capable of learning these heterogeneous synaptic delays.
+Then, we develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model thanks to theoretically defined first principles and derive a learning rule to adapt these heterogeneous synaptic delays.
 Using this detection task, we study the emergence of spatio-temporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner.
 We will investigate the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.
+Indeed, once trained, the amount of event-driven computations could be drastically reduced by pruning synapses, while maintaining top performance for the classification task.
 In this way, we will be able to show how such a model can provide an efficient solution to the trade-off between energy and accuracy.

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
 %%%-----------------------------------------------------------------
-In this paper, we aim to test whether the HD-SNN model is capable of efficiently learning a motion detection task, which is defined by a realistic event-based data stream.
-This type of signal is typically captured by a Dynamic Vision Sensor (DVS) and is inspired by the signal that is sent from the retinal ganglion cells through the optic nerve.
-The events are binary in nature and should be sufficient to perform the task of motion detection in a fast and efficient manner.
-Here we will first describe the task definition and then the HD-SNN model we use for motion inference and how we train it.
+In this paper, we test whether the HD-SNN model is able to efficiently learn a motion detection task defined by a realistic event-based data stream.
+This type of signal is commonly captured by a Dynamic Vision Sensor (DVS) and is inspired by the signal sent from the retinal ganglion cells through the optic nerve to the visual cortex.
+The events are binary in nature and should be sufficient to perform the motion detection task in a fast and efficient manner.
+Here we first describe the task definition and then the HD-SNN model we use for motion inference and how we train it.
 %%%-----------------------------------------------------------------

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
 \newcommand{\Abstract}{
-The response of a biological neuron depends on the precise timing of spikes coming from afferent neurons.
-This temporal aspect of the neural code is essential to understand information processing in neurobiology and applies particularly well to the output of neuromorphic hardware such as event-based cameras.
+    The response of a biological neuron depends on the precise timing of spikes from afferent neurons.
+This temporal aspect of the neural code is essential for understanding information processing in neurobiology, and is particularly applicable to the output of neuromorphic hardware such as event-based cameras.
 However, most artificial neural models do not take advantage of this important temporal dimension of the neural code.
 Here, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
-The model uses the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of synaptic inputs as they reach the basal dendritic tree.
+The model takes advantage of the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of synaptic inputs as they reach the basal dendritic tree.
 We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
-We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or to event-based cameras.
-We characterize the accuracy of the model in detecting visual motion and in particular, by pruning some weights we show it is still efficient at very low workloads.
-As a conclusion, heterogeneous delays provide with an end-to-end event-driven computational building block which could improve the performance of future spiking neural network (SNN) algorithms and in particular their applications in neuromorphic chips.
+We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or event-based cameras.
+We characterize the accuracy of the model in detecting visual motion, and in particular, by pruning some weights, we show that it is still efficient at very low workloads.
+In conclusion, heterogeneous delays provide an end-to-end event-driven computational building block which could improve the performance of future spiking neural network (SNN) algorithms and in particular their applications in neuromorphic chips.
 }

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
intro with https://www.deepl.com/write

@@
 %
-The human brain has the remarkable property of being able to react efficiently, at any time while consuming a reasonable amount of energy, about tens of watts.
-This system is the result of millions of years of natural selection, and a striking difference with artificial neural networks is the representation both use.
-Artificial convolutional neural networks (CNNs), for example, represent the flow of information from one layer to another as tensors, storing visual information densely across the visual topography, with different translation-invariant properties represented in different channels.
+The human brain has the remarkable capacity to react efficiently at any given time while consuming a reasonable amount of energy, about tens of watts.
+This system is the result of millions of years of natural selection, and a striking difference with artificial neural networks is the representation that both use.
+Artificial convolutional neural networks (CNNs), for example, represent the flow of information from one layer to another as analog tensors, storing visual information densely across the visual topography, with different translation-invariant properties represented in different channels.
 CNNs have achieved state-of-the-art performance for some computer vision tasks, such as image recognition.
-These networks are also known to mimic many properties of biological systems, such that each can be attributed a ``Brain Score''~\citep{schrimpf_brain-score_2020}, yet this score takes no account of inference speed or energy consumption.
-Indeed, they do not take advantage of the dynamics inherent to the way we perceive our natural environment.
+These networks are also known to mimic many properties of biological systems, such that each can be assigned a ``brain score''~\citep{schrimpf_brain-score_2020}, but this score does not take into account inference speed or energy consumption.
+In fact, they do not take advantage of the dynamics inherent in the way we perceive our natural environment.
 In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
 Spiking neural networks (SNNs), known as the third generation of artificial neural networks, incorporate this temporal dimension into the way they perform their computations.
-They are of interest for computational neuroscience because they provide a better model of biological neuron.
-Some approaches have developed normative models of SNNs that aim to have applications in machine learning.
-One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
-Another type of SNN that uses accurate timing of spikes attempts to determine the structure of the network to minimize a cost function.
-This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been recently extended using a surrogate gradient, a method which is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
-However, this performance of SNNs still lags behind that of networks using an analog, firing-rate based representation.
+They are of interest to computational neuroscience because they provide a better model of the biological neuron.
+One is the SpikeNet algorithm, which takes a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
+Another type of SNN that uses the exact timing of spikes tries to determine the structure of the network to minimize a cost function.
+This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and was recently extended using a surrogate gradient, a method that is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
+However, the performance of SNNs still lags behind that of networks using an analog, firing rate-based representation.
 The question of the advantage of using spikes in machine learning and computer vision remains open.
 
-In a recent review paper, we reported experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues~\cite{grimaldi_precise_2023}.
-These spatio-temporal motifs which are present within the neural activity may form useful representations to perform computations for various cognitive tasks using the synchronicity of spikes reaching the soma of a neuron.
-Notably, \citet{abeles_role_1982} asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs.
-While the first possibility favors the firing-rate coding theory, the second one emphasizes the function of temporal precision in the neural code.
-Since then, numerous studies have demonstrated the emergence of synchronicity of activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding using spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
-All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
-In neural models, the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows for the efficient exploitation or detection of these spatio-temporal motifs embedded in the spike train.
+In a recent review, we reported experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues~\cite{grimaldi_precise_2023}.
+These spatiotemporal motifs present in neural activity may form useful representations to perform computations for various cognitive tasks using the synchrony of spikes reaching the soma of a neuron.
+In particular, \citet{abeles_role_1982} asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs.
+While the first possibility favors the firing rate coding theory, the second emphasizes the function of temporal precision in the neural code.
+Since then, numerous studies have demonstrated the emergence of synchronous activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding using spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008}, or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
+All these findings and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatiotemporal spiking motifs in biological spike trains.
+In neural models, the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows the efficient exploitation or detection of these spatio-temporal motifs embedded in the spike train.
 In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a spiking motif defined by a subset of neurons with different, but precise, relative spiking delays.
-This delay between a pair of connected neurons is defined as the time between the emission of a spike on the afferent neuron's soma and that of its arrival on the efferent neuron's soma.
-Due to the variety of weights and delay within a population, representations using polychronous groups have mathematically a much higher information capacity than a firing rate-based approach.
+This delay between a pair of connected neurons is defined as the time between the emission of a spike at the soma of the afferent neuron and its arrival at the soma of the efferent neuron.
+Due to the variety of weights and delays within a population, representations using polychronous groups have a mathematically much higher information capacity than a firing rate-based approach.
 
-The present paper proposes a real-life application extending a recent model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}.
+The present paper proposes a real-world application that extends a recent model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}.
 Indeed, this previous model was trained to solve a simplified motion detection task on a synthetic event-based dataset generated by moving parameterized textures.
-This constituted a preliminary demonstration that formal neurons could exploit the precise timing of spikes to improve neural computations thanks to heterogeneous delays.
-Here, we extend these results to a more complex and natural framework.
-We first define the ecological cognitive task the model has to solve with the different datasets it will be tested on.
+This was a first demonstration that formal neurons can exploit the precise timing of spikes to improve neural computation thanks to heterogeneous delays.
+Here, we extend these results to a more complex and natural setting.
+First, we define the ecological cognitive task that the model must solve with the different datasets on which it will be tested.
 Instead of the synthetic textures that have been used to generate motion-driven stimuli, we use natural images and random movements that mimic biological saccades.
-Then, we develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model thanks to theoretically defined first principles and derive a learning rule to adapt these heterogeneous synaptic delays.
+We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model thanks to theoretically defined first principles and derive a learning rule to adapt heterogeneous synaptic delays.
 Using this detection task, we study the emergence of spatio-temporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner.
-We will investigate the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.
-Indeed, once trained, the amount of event-driven computations could be drastically reduced by pruning synapses, while maintaining top performance for the classification task.
-In this way, we will be able to show how such a model can provide an efficient solution to the trade-off between energy and accuracy.
+We investigate the efficiency of the motion detection mechanism and, in particular, its resilience to synaptic weight pruning.
+Indeed, once trained, the amount of event-driven computation could be drastically reduced by removing weak synapses while maintaining peak performance for the classification task.
+In this way, we will be able to show how such a model can provide an efficient solution to the energy/accuracy tradeoff.
 %

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
methods

@@
 %%%-----------------------------------------------------------------
-To train and test our model, we need to define a visual dataset for which we control the direction and speed of motion.
-Let us now define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
+To train and test our model, we need to define a visual dataset for which we supervise the direction and speed of motion.
+For this, we define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
 First, we draw a trajectory inspired by the biological movements of the eyes.
@@
 A specific mechanism for this function are saccades, which are rapid eye movements that reposition the center of vision.
-In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}).
-They are generated very quickly (about $80~\ms$) and have a wide range of speeds.
+In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}), they are generated very quickly (about $80~\ms$) and have a wide range of speeds.
 On a more microscopic scale, the human gaze moves with a continuous drift similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
@@
 This approach first defines a finite set of possible 2D movements in polar coordinates.
-Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $8$ linearly spaced movement directions and 3 geometrically spaced speeds.
+Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $XXX8$ linearly spaced movement directions and $XXX3$ geometrically spaced speeds.
 Next, we define a gaze trajectory as segments whose duration is drawn from a Poisson distribution with a mean block length of $24~\ms$, similar to a Lévy flight~\citep[p.
 289]{mandelbrot_fractal_1982}.
-Finally, the trajectory is integrated, assuming first that the velocities are sampled uniformly and independently of the set of different motion sets, and second that the motion is uniform during a time segment.
+Finally, the trajectory is integrated, assuming first that the velocities are sampled uniformly and independently from the set of different motions, and second that the motion is uniform during a time segment.
 The resulting instances yield trajectories that are qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
methods A

@@
 To train and test our model, we need to define a visual dataset for which we supervise the direction and speed of motion.
-For this, we define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
+To achieve this, we define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
 First, we draw a trajectory inspired by the biological movements of the eyes.
@@
 In animals with a fovea, this is particularly useful because it allows the area with the highest density of photoreceptors in the environment to be moved, for example, to a point of interest.
-A specific mechanism for this function are saccades, which are rapid eye movements that reposition the center of vision.
-In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}), they are generated very quickly (about $80~\ms$) and have a wide range of speeds.
+A specific mechanism for this function is saccades, which are rapid eye movements that reposition the center of vision.
+In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}), very fast (about $80~\ms$), and have a wide range of speeds.
 On a more microscopic scale, the human gaze moves with a continuous drift similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
-In order to maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.
+To preserve the full generality of the task, we define eye movements using such a form of random walk~\citep{engbert_integrated_2011}.
 This approach first defines a finite set of possible 2D movements in polar coordinates.
-Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $XXX8$ linearly spaced movement directions and $XXX3$ geometrically spaced speeds.
+Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $XXX8$ linearly spaced directions and $XXX3$ geometrically spaced speeds.
 Next, we define a gaze trajectory as segments whose duration is drawn from a Poisson distribution with a mean block length of $24~\ms$, similar to a Lévy flight~\citep[p.
 289]{mandelbrot_fractal_1982}.
-Finally, the trajectory is integrated, assuming first that the velocities are sampled uniformly and independently from the set of different motions, and second that the motion is uniform during a time segment.
-The resulting instances yield trajectories that are qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
+Finally, assuming first that the velocities are sampled uniformly and independently from the set of different motions, and second that the motion is uniform during a time segment,  the trajectory is generated by integrating the motion sequence.
+This generative model yield trajectories that are qualitatively very similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
 
@@
 In the lower left corner of the snapshots, the translation vector is shown in red as one of the possible classes of motion.
-Note the change in the direction of motion between the third and fourth image, and also that, due to the aperture problem, contours parallel to the motion emit relatively fewer spikes.}
+Note the change in the direction of motion between the third and fourth image, and also that, due to the aperture problem, contours parallel to the motion emit relatively fewer spikes.
+        }
     \label{fig:motion_task}
@@
 For this purpose, we selected a database of $600$ natural images that were previously used to study the statistics of natural images~\citep{perrinet_edge_2015}.
-Note that these have been pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band.
+Note that these were pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band.
 This process is known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $256 \times 256$ in size, and we will extract sub-images of size $64 \times 64$ positioned around the center of gaze at each time step.
+These images are $256 \times 256$ in size, and we will extract subimages of size $64 \times 64$ positioned around the center of gaze at each time step.
 We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 128~\ms$.
-To avoid boundary effects, we will randomly position the full trajectory in image space such that the subimage is translated using the position given by the trajectory at each time step without touching the borders.
-At each point in time, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
+To avoid boundary effects, we will randomly position the full trajectory in image space so that the subimage is translated using the position given by the trajectory at each time step without touching the boundaries.
+At each time step, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a subpixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
 Note that the magnitude of the displacement is relative to the time bin, and we have defined the unit velocity to correspond to a movement of one pixel per frame (i.e., per time bin).
 
-To transform each movie into events as the ones recorded by a DVS, we compute a gradient image (initialized at zero) by computing the temporal gradient of the pixels' intensity over two successive frames.
-For a specific pixel and timestamp, an event is generated if the absolute value of this gradient exceeds a threshold.
-The event has either an OFF or ON polarity, respectively whether the gradient is negative or positive.
-This signed threshold value is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion_task}-\textit{(Right)}).
-The goal here is to infer the correct motion solely by observing these events.
-% \note{ the task is full-field rigid translations, the model resolves it locally (analogy to the input from V1 to MT) }
-%%%-----------------------------------------------------------------
+To transform each movie into events, as recorded by a DVS, we compute a gradient image (initialized at zero) by computing the temporal gradient of the pixels' intensity over two successive frames.
+For a given pixel and time stamp, an event is generated when the absolute value of this gradient exceeds a threshold.
+The event has either an OFF or ON polarity, depending on whether the gradient is negative or positive.
+The signed threshold is then subtracted from the residual gradient image.
+When applied to the whole movie, the event stream is similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, i.e.
+a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF), and their time $\timev_\arank$ (\seeFig{motion_task}-\textit{(Right)}).
+The goal here is to infer the correct motion solely by observing these events.%%%-----------------------------------------------------------------
 \subsection{The Heterogeneous Delays SNN (HD-SNN) model}

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Methods : A

@@
     \caption{
-        {\bf - Motion Detection Task.} \textit{(Left)}~We use large natural images ($256\times256$) in which an aperture ($64\times64$) extracts a cropped image around the view axis.
-To mimic the effect of a saccadic eye movement, the view axis moves according to a stepwise random walk.
+        {\bf - Motion Detection Task.} \textit{(Left)}~We use large natural images ($1024\times1024$) in which an aperture ($128\times128$) extracts a cropped image around the view axis.
+To mimic the effect of a saccadic eye movement, this axis moves according to a jagged random walk.
 We show such a trajectory with a length of $128$ time steps (green line).
@@
 The dynamics of the cropped image translated according to the trajectory as a function of time produces a naturalistic movie, which is then transformed into an event-based representation.
-Mimicking the retina, this representation encodes proportional increases or decreases in luminance, i.e.
-ON (red) and OFF (blue) events, in each pixel of the image.
+Mimicking the retina, this representation encodes in each pixel of the image quantal increases or decreases in luminance, i.e.
+ON (red) and OFF (blue) events.
 In the lower left corner of the snapshots, the translation vector is shown in red as one of the possible classes of motion.

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Methods : B

@@
 In a neurobiological recording or in the sensory signal obtained from an event-based camera, the input consists of a stream of \emph{spikes} or \emph{events}.
-This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{(Top-Left} for an illustration).
-Events are typically ordered by their time of occurrence.
+This can be formalized as a list of tuples, each tuple representing neural addresses and timestamps.
+We defined this list as $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{(Top-Left} for an illustration).
+Events are usually ordered by their time of occurrence.
 Each event has a time of occurrence $\timev_\arank$ and an associated address $\presynaddr_\arank$.
-This defines an address space $\presynaddrspace$ which consists of the set of possible addresses.
-In a neurobiological recording, this can be the identified set of presynaptic neurons.
-For neuromorphic hardware, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$ where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels.
-As such, an address $\presynaddr_\arank$ is typically in the form $(\polev_\arank, x_\arank, y_\arank)$ for event-based cameras.
+This defines an address space $\presynaddrspace$, which consists of the set of possible addresses.
+In a neurobiological recording, this may be the identified set of all presynaptic neurons.
+For neuromorphic hardware, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$, where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities encoded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels.
+Thus, an address $\presynaddr_\arank$ is typically in the form $(\polev_\arank, x_\arank, y_\arank)$ for event-based cameras.
 
-In the HD-SNN model, each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferent neurons from $\presynaddrspace$.
-In biology, a single cortical neuron has generally several thousands of synapses.
-Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
-Note that a neuron may contact another afferent neuron with different delays through different synaptic connections.
-Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
-as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
-This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{(Right)} for an illustration of the connectivity of one neuron with synaptic weights and delays).
+In the HD-SNN model, neurons $\postsynaddr \in \postsynaddrspace$ are connected to presynaptic afferent neurons from $\presynaddrspace$.
+In biology, a single cortical neuron typically has several thousand synapses.
+Each synapse can be defined by its synaptic weight and its delay, the time it takes for a spike to travel from the soma of the presynaptic neuron to the soma of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$, but also by the set of possible delays.
+Note that a neuron can contact another afferent neuron with different delays through different synaptic connections.
+By scanning all $\postsynaddr$ neurons, we define the full set of $\Nsyn$ synapses,
+as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated with a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
+This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{(Right)} for an illustration of the connectivity of a neuron with synaptic weights and delays).
 
@@
 As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field.
-It results in a weighted event stream (\seeFig{izhikevich}-\textit{(Middle-Left}) for each  postsynaptic neuron $\postsynaddr$:
+This results in a weighted event stream (\seeFig{izhikevich}-\textit{(Middle-Left}) for each  postsynaptic neuron $\postsynaddr$:
 %
@@
 In biology, this new stream of events is naturally ordered in time as events reach the soma of postsynaptic neurons.
-However, it should be properly reordered in simulations.
-Crucially, when postsynaptic neurons are activated on their soma by a specific spatio-temporal motif which is imprinted in the set of synapses, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner while the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions (\seeFig{izhikevich}).
-We defined the HD-SNN model in this subsection in all generality, and in the next subsection, we describe an implementation of our model adapted for the motion detection task.
+In simulations, however, it should be properly reordered.
+The activation function of the HD-SNN neurons can be selected from the full range of spiking neuron response functions (\seeFig{izhikevich}).
+Crucially, when postsynaptic neurons are activated at their soma by a specific spatiotemporal motif imprinted in the synaptic set, the discharge probability increases, and especially when these spikes converge at the soma in a synchronous manner.
+We have defined the HD-SNN model in all generality in this subsection, and in the next subsection we describe an implementation of our model adapted to the motion detection task.
 %

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Methods : C

@@
 The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
-First, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:kernel_b} corresponds to a temporal convolution that transforms the input $A$ using one kernel per postsynaptic neuron~\citep{grimaldi_learning_2022}.
-To take advantage of the position invariance observed in images and exploited in convolutional neural networks, we can further assume that synaptic motifs should be similar across different positions, so we can define a spatio-temporal convolutional operator.
+First, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:kernel_b} corresponds to a temporal convolution that transforms the input $A$ with each kernels and positions assigned to the postsynaptic neurons~\citep{grimaldi_learning_2022}.
+This takes advantage of the position invariance observed in images and exploited in convolutional neural networks, and we can further assume that synaptic motifs should be similar across different positions.
+As a consequence, we can define a spatio-temporal convolutional operator, in which layer of neurons assigned to specific kernels form channels.
 
 Therefore, if we make the approximation at any given time that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
-The longest synaptic delay defines the depth $\Ktime$ of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented.
-In particular, the whole synaptic set can be represented as one kernel for each class $\class$ of the supervision task as the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
+The longest synaptic delay defines the depth $\Ktime$ of the kernel $\kernel^\postsynaddr$ such that all possible delays associated to the different presynaptic addresses are represented.
+In particular, for each class $\class$ of the supervision task, the whole synaptic set can be represented as one kernel which is represented by the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
 To keep an analogy with the HD-SNN model, for neuron $\postsynaddr$ of position $(x_\postsynaddr, y_\postsynaddr)$ and class $\class$, $\kernel_\class$ gives the weights as a function of relative addresses and synaptic delays: $\forall {\delta_x \in [1, \Kx], \delta_y \in [1, \Ky], \delta_\timev \in [1, \Ktime], \polev \in [1, \Npol]}, \; \kernel_\class(\polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$.
-The connectivity of neuron $\postsynaddr$ is defined locally, around it position $(x_\postsynaddr, y_\postsynaddr)$.
+By convenienece, the connectivity of neuron $\postsynaddr$ is defined locally, around its position $(x_\postsynaddr, y_\postsynaddr)$.
 Using this dense representation, the processing of $A$ by the model can be written as: %

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Methods : C

@@
 
-Therefore, if we make the approximation at any given time that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
-The longest synaptic delay defines the depth $\Ktime$ of the kernel $\kernel^\postsynaddr$ such that all possible delays associated to the different presynaptic addresses are represented.
-In particular, for each class $\class$ of the supervision task, the whole synaptic set can be represented as one kernel which is represented by the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
-To keep an analogy with the HD-SNN model, for neuron $\postsynaddr$ of position $(x_\postsynaddr, y_\postsynaddr)$ and class $\class$, $\kernel_\class$ gives the weights as a function of relative addresses and synaptic delays: $\forall {\delta_x \in [1, \Kx], \delta_y \in [1, \Ky], \delta_\timev \in [1, \Ktime], \polev \in [1, \Npol]}, \; \kernel_\class(\polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$.
-By convenienece, the connectivity of neuron $\postsynaddr$ is defined locally, around its position $(x_\postsynaddr, y_\postsynaddr)$.
-Using this dense representation, the processing of $A$ by the model can be written as: %
+Therefore, assuming that a neuron integrates only in the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
+The longest synaptic delay defines the depth $\Ktime$ of the kernel $\kernel^\postsynaddr$, so that all possible delays associated with the different presynaptic addresses are represented.
+In particular, for each class $\class$ of the supervision task, the entire synaptic set can be represented as a kernel, which is represented by the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
+To keep the analogy with the HD-SNN model, for a neuron $\postsynaddr$ of position $(x_\postsynaddr, y_\postsynaddr)$ and class $\class$, $\kernel_\class$ gives the weights as a function of relative addresses and synaptic delays: $\forall {\delta_x \in [1, \Kx], \delta_y \in [1, \Ky], \delta_\timev \in [1, \Ktime], \polev \in [1, \Npol]}, \; \kernel_\class(\polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$.
+For convenience, the connectivity of the neuron $\postsynaddr$ is defined locally around its position $(x_\postsynaddr, y_\postsynaddr)$.
+Using this dense representation, the model's processing of $A$ can be written as %.
 %

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
velocities

@@
 289]{mandelbrot_fractal_1982}.
-Finally, assuming first that the velocities are sampled uniformly and independently from the set of different motions, and second that the motion is uniform during a time segment,  the trajectory is generated by integrating the motion sequence.
+Finally, assuming first that the speeds are sampled uniformly and independently from the set of different motions, and second that the motion is uniform during a time segment,  the trajectory is generated by integrating the motion sequence.
 This generative model yield trajectories that are qualitatively very similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
@@
 At each time step, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a subpixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
-Note that the magnitude of the displacement is relative to the time bin, and we have defined the unit velocity to correspond to a movement of one pixel per frame (i.e., per time bin).
+Note that the magnitude of the displacement is relative to the time bin, and we have defined the unit speed to correspond to a movement of one pixel per frame (i.e., per time bin).
 
@@
 %
-where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses inside a kernel and $\delta_\timev$ is the synaptic delay. 
+where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses within a kernel and $\delta_\timev$ is the synaptic delay. 
 This shows that $\mathcal{C}_\class$ is the result of a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses: $\mathcal{C}_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration).
-Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
-This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
-A similar type of spatio-temporal filtering is used as a pre-processing stage for a pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
-Also,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm which implements a motion estimation task.
-By assuming a locally constant velocity, the authors assume the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.
+Note that to stay within the framework of a causal computation, the kernels are shifted in time so that only past information gives an answer at the present time.
+This well-known computation defines a differentiable measure, which is very efficiently implemented for GPUs, and which we will use to learn the classification of different motifs in the event stream.
+A similar type of spatiotemporal filtering is used as a preprocessing stage for an existing pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
+In addition,~\citep{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm that implements a motion estimation task.
+By assuming a locally constant motion, the authors assume that the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the motion.
 
-Here, we keep the analogy with spiking neurons and we try to observe the detection of spiking motifs using the spatio-temporal kernels.
+Here, we keep the analogy with spiking neurons, and we try to observe the detection of spiking motifs using the spatio-temporal kernels.
 If so, this precise spatio-temporal patterns prove to be of interest for neural computations and one can infer that biological neurons make use of this information as well.
-% \note{plus de biblio sur la motion detection/segmentation ou optique flow avec les events? (reviewer 2)}
 The 3D convolution represents the linear integration of the spike train as the linear input to the neuron.
-Then, the activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Takes-All network~\citep{nessler_bayesian_2013}.
-In our MLR model, a probability value for each class (i.e.
-each motion direction) is predicted for each position $x, y$ and time $\timev$ as a softmax function of the linear combination of the list of events.
-% on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
-The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the HD-SNN model.
-Formally, using the kernels, it transforms the input raster plot into a probability with the following formula:
+The activation function of our model is then a softmax function that implements a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}.
+In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a softmax function of the linear combination of the list of events.
+Formally, using the kernels, the input raster plot is transformed into a probability with the following formula:
 %
@@
 %
-where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function and  $\bias_\class$ is the bias linked to class $\class$. 
-In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
+where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function, and $\bias_\class$ is the bias associated with the class $\class$.
+In particular, we expect that some specific motifs will become tightly synchronized as they reach the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike.
 The spiking output of the model corresponds to an event with the highest probability class.
@@
 Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel heuristics inspired by neuroscience are the main methodological differences with that previous study.
-First, in our simulations, we set the size of the kernels to $(21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = 8 \times 3$.
+First, in our simulations, we set the size of the kernels to $(\Ktime, \Kx, \Ky) =  (21, 17, 17)$ and define as many classes as the number of motions (directions and speeds): $N_\class = XXX8 \times XXX3$.
 To avoid introducing biases in the directions, we apply a circular mask to the spatial dimensions of the kernels.
@@
 Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli.
-In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
+In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for motion, among other parameters (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
 In particular, we will set the spatial extent and duration of the generated movies similarly to the motion task defined above.
 This procedure defines a set of textures with different spatial properties and different motions picked from the same set of $8$ directions and $3$ speeds.
-For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content to provide with some naturalistic variability.
+For any given motion, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content to provide with some naturalistic variability.
 This method provides a rich dataset of textured movies for which we know the ground truth for motion.
@@
 %\note{Say that this could be implemented by properly setting the timing of the label}
-As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by velocity, acceleration, jerk,~\ldots), and ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
+As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by motion, acceleration, jerk,~\ldots), and ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
 Our proposed model using heterogeneous delays provides an alternative and elegant implementation solution to this problem.
@@
 A possible implementation is therefore to add a new layer to our model, analogous to the hierarchical organization highlighted in the visual cortex.
-This is theoretically possible by using the output of our model (which estimates velocity in retinotopic space) as input to a new layer of neurons that would estimate velocity in the visual field, including the depth dimension in the output supervision labels.
+This is theoretically possible by using the output of our model (which estimates motion in retinotopic space) as input to a new layer of neurons that would estimate motion in the visual field, including the depth dimension in the output supervision labels.
 This could have direct and important applications, e.g.

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Methods : C

@@
 
-Now that this general framework has been explained, we can add some heuristics, based on neuroscientific observations, to constrain our model and its strategies for solving the task that is used to solve the ecological task described in section~\ref{sec:task}.
-Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel heuristics inspired by neuroscience are the main methodological differences with that previous study.
-First, in our simulations, we set the size of the kernels to $(\Ktime, \Kx, \Ky) =  (21, 17, 17)$ and define as many classes as the number of motions (directions and speeds): $N_\class = XXX8 \times XXX3$.
+Now that this general framework has been explained, we can add some heuristics, based on neuroscientific observations, to constrain our model and its strategies for solving the task used to solve the ecological task described in section~\ref{sec:task}.
+Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel neuroscience-inspired heuristics are the main methodological differences with that previous study.
+First, in our simulations, we set the size of the kernels to $(\Ktime, \Kx, \Ky) = (21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = XXX8 \times XXX3$.
 To avoid introducing biases in the directions, we apply a circular mask to the spatial dimensions of the kernels.
-Furthermore, we included a prior in the motions that could be selected, as there is a prior in natural scenes for slow speeds~\citep{vacher_bayesian_2018}.
-Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we thus apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration).
-This is consistent with neuroscientific principles because, due to the limits of the conduction delay along horizontal connections, the synaptic delay is related, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one.
-Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities arrangements.
-This is due to the fact that our whitened images have a symmetry in the luminance profiles, and an image with inverted contrast is indistinguishable.
-Since this arrangement of polarities is independent of speed, we added a mechanism that collects the linear values for the movie and the movie with the ON and OFF cells flipped, keeping only the maximum value for each voxel.
+We also included a prior in the selectable motions, as there is a prior for slow speeds in natural scenes~\citep{vacher_bayesian_2018}.
+Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration).
+This is consistent with neuroscientific principles because, due to the limits of conduction delay along horizontal connections, the synaptic delay is physically related to the distance between the presynaptic and postsynaptic neurons.
+Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities.
+This is due to the fact that our whitened images have a symmetry in the luminance profiles, and an image with inverted contrast is indistinguishable from a standard one.
+Since this arrangement of polarities is independent of motion, we added a mechanism that collects the linear values for the movie and the movie with the ON and OFF cells flipped, keeping only the maximum value for each voxel.
 This is similar to the computation done for complex cells in primary visual cortex.
@@
     \includegraphics[width=\linewidth]{figures/HDSNN_conv.pdf}
-    % \includegraphics[width=\linewidth]{figures/HD-SNN.png}
     \caption{
-    % \textbf{- HD-SNN illustration.
-From left to right:} A 3D representation of the event stream as output of the DVS.
-A 3D convolution is applied, on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis.
-There are as many kernels as the possible number of motion classes in the recording.
-For clarity, only $4$ kernels are illustrated and some weights are not represented in the voxel grid.
-For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e.
-softmax function).
-The output of the MLR gives a probability for each class, associated to a specific kernel, to emit an output event, or spike.
-With an appropriate spiking mechanism, here a winner-takes-all, we obtain, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class.
-}
-    \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot here as a raster plot a 2D representation of the input event stream (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time).
-A spatio-temporal convolution is applied to the dense representation of the input with $2$ distinct convolutional kernels (the green and orange kernels) that will define the output channels.
+        \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot a 2D representation of the input event stream as a raster plot (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time).
+A spatio-temporal convolution is applied to the dense representation of the input with $2$ different convolution kernels (the green and orange kernels), which will define the output channels.
 The convolution is summed over the two polarities.
-If you have two axes $X$ and $Y$ to represent the presynaptic addresses as for the pixel grid of a DVS, you obtain a 3D convolution.
-We restrict the illustration to a 2D representation and to 2 possible classes (green and orange) that are associated with different motion directions.
-\textit{(Middle)}~For each position (address, time), one can compute the activation resulting from the convolution.
+Since we have two axes $X$ and $Y$ to represent the presynaptic addresses, like the pixel grid of a DVS, this results in a 3D convolution.
+Here we simplify the illustration to a 2D representation and to 2 possible classes (green and orange) associated with different directions of motion.
+\textit{(Middle)}~For each position (address, time) one can compute the activation resulting from the convolution.
 The output of the convolution is processed by the nonlinearity of the MLR model (i.e., the softmax function).
 The output of the MLR gives a probability for each class associated with a particular kernel (colored bars in the highlighted pixel).
-\textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated to a thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated to a specific motion class.
-Note that the position of the output spikes does not systematically correspond to the position of the input spikes but only when enough evidence is obtained.}
+\textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated with thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated with a particular motion class.
+Note that the position of the output spikes does not systematically correspond to the position of the input spikes, but only when enough evidence is obtained.
+    }
     \label{fig:model}
@@
 Since the model is fully differentiable, we can now implement a supervised learning rule.
-This rule was implemented using the binary input events as inputs and the corresponding motion direction labels as the desired output.
+This rule was implemented using the binary input events as inputs and the corresponding motion's labels as the desired output.
 The loss function of the MLR model is the binary cross entropy at the output of the classification layer.
@@
 Note that in this context, the label is known but the position is not, mainly due to the sparse spatial content of natural images.
-However, the supervised optimization of this MLR model will involve adjusting the weights of the kernels.
+However, the supervised optimization of this MLR model will adjust the weights of the kernels.
 As a result, the error is only propagated back to the spatial locations of these most active cells.
 This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation.
-Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$).
+Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies, each of size $TTT \times XXX \times YYY$ and a learning rate of $10^{-5}$).
 We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
@@
 We have shown above that if different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities.
-Thus, by taking the mean logit of the probability of the output given by the model at all positions for any given time, we can calculate the probability of the output at this time.
+Thus, by taking the mean logit of the probability of the output given by the model at all positions for any given time, we can calculate the probability of the output at that time.
 This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted).
-These calculations are performed on a different input dataset than the one used in the training or validation steps.
+These calculations are performed on a different input data set than the one used in the training or validation steps.
 The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN} (see Data Availability).

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Another pass

@@
     \caption{
-    	Representation of the weights for $8$ directions for one speed (among the $8 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
-Directions are shown as red arrows on the left.
-The spatio-temporal kernels are represented as slices of spatial weights at different delays.
+    	Representation of the weights for $XXX8$ directions for a single speed (among the $XXX8 \times XXX3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
+The directions are shown as red arrows in the left insets.
+The spatiotemporal kernels are shown as slices of spatial weights at different delays.
 Delays vary along the horizontal axis from the far right (delay of one step) to the left (delay of $12$ steps).
 Each image corresponds to the weights at a given delay, with excitatory and inhibitory weights in warm and cold colors, respectively.
-Due to the symmetry between the ON and OFF event streams, we report that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective to the different motion directions, and we observe for all kernels an orientation preference perpendicular to the orientation.
-%      % These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
-%	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.
-One sees positive (excitatory) coefficients for the specific direction of motion but also negative (inhibitory) coefficients for other directions.
+Due to the symmetry between the ON and OFF event streams, we report that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective for the different motion directions, and we observe a slight orientation preference perpendicular to direction for all kernels.
 	}
@@
 After training our model, we first analyze the weights learned for the different neurons (\seeFig{kernels}).
-We first notice that these cells are very localized along each trajectory, yet show an orientation selectivity, similar to that detected in MT neurons~\citep{deangelis_functional_1999}.
-Focusing on the positive weights, a strong selectivity is observed along specific axes of motion for each of the different kernels.
-These directions can be easily associated to the direction of motion controlled in the natural images and show a typical anisotropic profile~\citep{kaplan_anisotropic_2013}.
-For instance, the first kernel shows a strong selectivity to a downward motion direction.
-We also observed a high dependence between the weights reaching the ON polarities and that reaching the OFF polarities.
-In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative in the other.
-This property comes from the way the events are generated and that the luminance can not at the same time increase and decrease.
+We first notice that these cells are very localized along each trajectory, but show an orientation selectivity similar to that found for MT neurons~\citep{deangelis_functional_1999}.
+Focusing on the positive weights, a strong selectivity along specific motion axes is observed for each of the different kernels.
+These directions can be easily associated with the direction of motion controlled in the natural images and show a typical anisotropic profile~\citep{kaplan_anisotropic_2013}.
+For example, the first kernel shows a strong selectivity for a downward direction of motion.
+Note that as the delays become larger, the coefficients become lower and less localized, consistent with the average diffusion of information included in the generative model and with the diffusion introduced in motion-based prediction models~\citep{perrinet_motion-based_2012,khoei_flash-lag_2017}.
+We also observed a strong dependence between the weights reaching the ON polarities and those reaching the OFF polarities.
+In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative for the other.
+This property is due to the way the events are generated and the fact that the luminance cannot increase and decrease at the same time.
 Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results obtained after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
-%ith the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
-%
-%This qualitative look at the 3D kernels allows the reader to infer for the $8$ different motion directions used to generate our synthetic event streams.
 
@@
 Our proposed model using heterogeneous delays provides an alternative and elegant implementation solution to this problem.
-%
 \subsection{Perspectives}

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Results : A

@@
 events away in the past.
-%For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Fig.~\ref{fig:kernels}.
-
 %
-% \note{kernel size has to be adapted to the characteristics of motion see Samonds %~\citep{samonds}
-% say that we have tested }
-% \note{tester la meme expe faite par baudot et modelisée par kremkow en montrant des raster plots du modele avec NI versus Gratings }
-
 %%%-----------------------------------------------------------------
@@
     \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
-    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of (from left to right) the mean spatial frequency, the bandwidth in spatial frequency (from a grating-like (left) to isotropic textures (right)), the bandwidth in orientation (from isotropic textures (left) to grating-like (right)), the bandwidth in speed (from a rigid motion (left) to independent frames (right)).
-Note that these accuracies are computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) or in the generic case where orientation is independent of direction ('aperture').}
+    \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from a grating (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to grating (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)).
+Note that these accuracies are computed both in the case where the orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) and in the generic case where the orientation is independent of direction ('aperture').
+    }
     \label{fig:motion_clouds}
@@
 %%%-----------------------------------------------------------------
-To test our model, we will quantify its ability to categorize different motions, that is on event streams for which the ground truth motion is known at every instant in time.
+To further test our model, we will quantify its ability to categorize different motions, that is on event streams for which the ground truth motion is known at every instant in time.
 Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli.

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Results : B + C

@@
 %%%-----------------------------------------------------------------
-To further test our model, we will quantify its ability to categorize different motions, that is on event streams for which the ground truth motion is known at every instant in time.
-Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli.
-In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for motion, among other parameters (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
-In particular, we will set the spatial extent and duration of the generated movies similarly to the motion task defined above.
-This procedure defines a set of textures with different spatial properties and different motions picked from the same set of $8$ directions and $3$ speeds.
-For any given motion, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content to provide with some naturalistic variability.
-This method provides a rich dataset of textured movies for which we know the ground truth for motion.
+To further test our model, we will quantify its ability to categorize different motions, i.e., on event streams for which the ground truth motion is known at each instant.
+Before applying the model to natural images, we will first test the model on simpler, parameterized stimuli.
+For this purpose, we will use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control, among other parameters, the motion (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
+In particular, we will set the spatial extent and duration of the generated movies similar to the motion task defined above.
+This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $XXX$ directions and $YYY3$ speeds.
+For each motion, we also varied the texture parameters, such as mean and variance of orientation or spatial frequency content, to provide some naturalistic variability.
+This method provides a rich dataset of textured movies for which we know the ground truth for the motion.
 
-We observe some facts of interest.
-First, as we change the mean spatial frequency of the texture, we notice a monotonous decrease in accuracy.
-This comes as a similar trend as that shown in the primary visual areas~\citep{priebe_tuning_2006} and revels the most informative scales learned by our model.
-We observe the tendency for different spatial frequency bandwidth.
-Notably, the accuracy is better for a grating-like stimulation than for a large (which qualitatively resemble a more textured stimulus), reminiscent to the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}.
+We observe some interesting facts.
+First, as we change the mean spatial frequency of the texture, we observe a monotonic decrease in accuracy.
+This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006} and reveals the most informative scales learned by our model.
+We observe the trend for different spatial frequency bandwidths.
+In particular, the accuracy is better for a grating-like stimulus than for a large one (which qualitatively resembles a more textured stimulus), reminiscent of the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}.
 Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.
-When the stimulus is grating-like and that the orientation is arbitrary with respect to the direction of motion, the system is faced with the aperture problem and see a decrease of accuracy.
+When the stimulus is grating-like and the orientation is arbitrary with respect to the direction of motion, the system faces the aperture problem and experiences a decrease in accuracy.
 This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion.
 Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter.
-This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018}.
-% but also that accuracy is low for a rigid motion which lacks variability.
+This shows a progressive decrease in accuracy, similar to that observed in the amplitude of human eye movements~\citep{mansour_pour_speed_2018}.
 %
@@
 %%%-----------------------------------------------------------------
-Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection.
-We observed that the distribution of the kernels' weights is sparse, with most values near zero (\seeFig{kernels}).
-As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the computations used for the convolution operation.
+Once our MLR is trained, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection.
+We observed that the distribution of the kernel weights is sparse, with most values close to zero (\seeFig{kernels}).
+As shown in the formalization of our event-based model, the computational cost of our model, if implemented on a neuromorphic chip, would be dominated by the computations used for the convolution operation.
 In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel.
-If the support of information is sparse, then computations can be performed only on those events.
+If the information support is sparse, then computations can only be performed on those events.
 Also, if we set some weights of the kernels to zero, then the sum can be skipped for those addresses.
-Knowing the sparseness of the input, the total number of computations thus scales with the number of spikes multiplied by the number of nonzero synaptic weights.
-% , which is proportional to the number of spikes multiplied by the number of synapses
+Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of non-zero synaptic weights.
 
-To assess the robustness of the classification as a function of the computational load, we will prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold.
-In Fig.~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations, or active weights, per decision for each neuron of the layer.
-As a comparison and to account for the gain in performance by using heterogeneous delays, we provide the accuracy obtained with a MLR model using 2D time surface (in red) as in~\citep{grimaldi_robust_2022}.
-This latter method is based on delays from the last recorded events and uses fewer computations (in our case $17\times17$) than the dense 3D kernels without any pruning ($21\times17\times17$).
-While less computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
+To assess the robustness of the classification as a function of computational load, we will prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold.
+In Fig.~\ref{fig:accuracy} we plot the classification accuracy as a function of the relative number of computations or active weights per decision for each neuron in the layer.
+For comparison, and to take into account the performance gain from using heterogeneous delays, we show the accuracy obtained with an MLR model using a 2D time surface (in red) as in~\citep{grimaldi_robust_2022}.
+This latter method is based on the delays of the last recorded events and requires fewer computations (in our case $17\times17$) than the dense 3D kernels without any pruning ($21\times17\times17$).
+While fewer computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
 
-By pruning weights, we observe that the evolution of accuracy as a function of the log percentage of active weights fits well a sigmoid curve.
-Half-saturation level is reached at a ratio of about $3.2\times 10^{-2}$ of active weights, corresponding in our setting to a total amount of about $180$ computations per decision.
-Compared to the full kernels, the accuracy of our method is maintained to its top performances when dividing the number of computations by a factor up to about $31$.
+By pruning the weights, we observe that the evolution of accuracy as a function of the log percentage of active weights fits well to a sigmoid curve.
+Half saturation is reached at a ratio of about $3.2\ times 10^{-2}$ of active weights, which in our setting corresponds to a total amount of about $180$ computations per decision.
+Compared to the full kernels, the accuracy of our method is maintained to its top performance when the number of computations is divided by a factor up to about $31$.
 In this case, the number of computations is greatly reduced compared to~\citep{grimaldi_robust_2022}, thus demonstrating the efficiency of the presented method.
@@
     \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
-    \caption{Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
+    \caption{
+        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
 The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
-As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about 25 times fewer computations.}
+As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about 25 times fewer computations.
+        }
     \label{fig:accuracy}

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
spatiotemporal

@@
 %
-%● Spike based spatio-temporal processing and its implications for neuromorphic vision.
+%● Spike based spatiotemporal processing and its implications for neuromorphic vision.
 %
@@
 All these findings and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatiotemporal spiking motifs in biological spike trains.
-In neural models, the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows the efficient exploitation or detection of these spatio-temporal motifs embedded in the spike train.
+In neural models, the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows the efficient exploitation or detection of these spatiotemporal motifs embedded in the spike train.
 In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a spiking motif defined by a subset of neurons with different, but precise, relative spiking delays.
@@
 We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model thanks to theoretically defined first principles and derive a learning rule to adapt heterogeneous synaptic delays.
-Using this detection task, we study the emergence of spatio-temporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner.
+Using this detection task, we study the emergence of spatiotemporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner.
 We investigate the efficiency of the motion detection mechanism and, in particular, its resilience to synaptic weight pruning.
@@
 This takes advantage of the position invariance observed in images and exploited in convolutional neural networks, and we can further assume that synaptic motifs should be similar across different positions.
-As a consequence, we can define a spatio-temporal convolutional operator, in which layer of neurons assigned to specific kernels form channels.
+As a consequence, we can define a spatiotemporal convolutional operator, in which layer of neurons assigned to specific kernels form channels.
 
@@
 where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses within a kernel and $\delta_\timev$ is the synaptic delay. 
-This shows that $\mathcal{C}_\class$ is the result of a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses: $\mathcal{C}_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration).
+This shows that $\mathcal{C}_\class$ is the result of a spatiotemporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses: $\mathcal{C}_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration).
 Note that to stay within the framework of a causal computation, the kernels are shifted in time so that only past information gives an answer at the present time.
@@
 
-Here, we keep the analogy with spiking neurons, and we try to observe the detection of spiking motifs using the spatio-temporal kernels.
-If so, this precise spatio-temporal patterns prove to be of interest for neural computations and one can infer that biological neurons make use of this information as well.
+Here, we keep the analogy with spiking neurons, and we try to observe the detection of spiking motifs using the spatiotemporal kernels.
+If so, this precise spatiotemporal patterns prove to be of interest for neural computations and one can infer that biological neurons make use of this information as well.
 The 3D convolution represents the linear integration of the spike train as the linear input to the neuron.
@@
         \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot a 2D representation of the input event stream as a raster plot (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time).
-A spatio-temporal convolution is applied to the dense representation of the input with $2$ different convolution kernels (the green and orange kernels), which will define the output channels.
+A spatiotemporal convolution is applied to the dense representation of the input with $2$ different convolution kernels (the green and orange kernels), which will define the output channels.
 The convolution is summed over the two polarities.
@@
 
-If one focuses on the interpretation of these kernels in terms of spatio-temporal motifs embedded in the event stream, it can lead to interesting outcomes.
+If one focuses on the interpretation of these kernels in terms of spatiotemporal motifs embedded in the event stream, it can lead to interesting outcomes.
 In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions.
-Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spiking motif.
+Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatiotemporal spiking motif.
 Each voxel in the 3D kernels defines a specific timestamp and a specific address.
-Consequently, our model is able to detect precise spatio-temporal motifs embedded in the spike train and associated to the different motion directions.
+Consequently, our model is able to detect precise spatiotemporal motifs embedded in the spike train and associated to the different motion directions.
 The cone shape for the positive weights' distribution highlights a loss of precision for longer delays, i.e.
@@
 %%%-----------------------------------------------------------------
-Once our MLR is trained, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection.
+Once our MLR model is trained, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection.
 We observed that the distribution of the kernel weights is sparse, with most values close to zero (\seeFig{kernels}).
@@
 The HD-SNN model that has been trained and evaluated on a naturalistic motion detection task.
-The model has been defined to provide optimal detection of event-driven spatio-temporal motifs.
+The model has been defined to provide optimal detection of event-driven spatiotemporal motifs.
 We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in the early visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
@@
 We would like to highlight a few innovations in the contributions that are presented in this paper.
-First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs.
-Moreover, in comparison to a representation with that algorithm, the parameters of the model (weights and delays) are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection for each spatio-temporal spike motif.
+First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs.
+Moreover, in comparison to a representation with that algorithm, the parameters of the model (weights and delays) are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection for each spatiotemporal spike motif.
 Another novelty is that the model learns the weights and the delays simultaneously.
@@
 In addition, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified toy problems~\citep{gutig_tempotron_2006}.
-Another major contribution is to provide a model that is suitable for learning any kind of spatio-temporal spiking motifs and that can be trained in a supervised manner by providing a dataset of supervision pairs.
+Another major contribution is to provide a model that is suitable for learning any kind of spatiotemporal spiking motifs and that can be trained in a supervised manner by providing a dataset of supervision pairs.
 Instead of relying on a careful description of the physical rules governing a task, e.g.

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Results : C

@@
 %%%-----------------------------------------------------------------
-Once our MLR model is trained, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection.
+After training our MLR model, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection.
 We observed that the distribution of the kernel weights is sparse, with most values close to zero (\seeFig{kernels}).
@@
 If the information support is sparse, then computations can only be performed on those events.
-Also, if we set some weights of the kernels to zero, then the sum can be skipped for those addresses.
+Also, if we set some weights of the kernels to zero, then the sum for those addresses can be dropped.
 Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of non-zero synaptic weights.
@@
 For comparison, and to take into account the performance gain from using heterogeneous delays, we show the accuracy obtained with an MLR model using a 2D time surface (in red) as in~\citep{grimaldi_robust_2022}.
-This latter method is based on the delays of the last recorded events and requires fewer computations (in our case $17\times17$) than the dense 3D kernels without any pruning ($21\times17\times17$).
+The latter method is based on the delays of the last recorded events and requires fewer computations (in our case $17\times17$) than the dense 3D kernels without any pruning ($21\times17\times17$).
 While fewer computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
@@
 By pruning the weights, we observe that the evolution of accuracy as a function of the log percentage of active weights fits well to a sigmoid curve.
-Half saturation is reached at a ratio of about $3.2\ times 10^{-2}$ of active weights, which in our setting corresponds to a total amount of about $180$ computations per decision.
-Compared to the full kernels, the accuracy of our method is maintained to its top performance when the number of computations is divided by a factor up to about $31$.
-In this case, the number of computations is greatly reduced compared to~\citep{grimaldi_robust_2022}, thus demonstrating the efficiency of the presented method.
+Half saturation is reached at a ratio of about $3.2\ times 10^{-2}$ of active weights, which in our setting corresponds to a total of about $180$ computations per decision.
+Compared to the full kernels, the accuracy of our method is maintained relative to its peak performance when the number of computations is divided by a factor up to about $31$.
+In this case, the number of computations is greatly reduced compared to~\citep{grimaldi_robust_2022}, demonstrating the efficiency of the presented method.
 %

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Discussion

@@
 %%%-----------------------------------------------------------------
-In this paper, we have introduced a generic SNN with heterogeneous delays and shown how it compares favorably to a state-of-the-art event-based classification algorithm for a visual motion detection task.
-The learned model shares a number of similarities with neurobiological anatomical observations, as well as with behavioral results.
+In this paper, we have introduced a generic SNN with heterogeneous delays and shown how it compares favorably with state-of-the-art event-based classification algorithms for a visual motion detection task.
+The learned model has a number of similarities with neurobiological anatomical observations as well as behavioral results.
 The event-based computations of our method can be drastically reduced by pruning synapses, while maintaining top classification performance.
-This shows that we can take advantage of the precise timing of spikes to improve the performance of neural computations.
+This shows that we can use the precise timing of spikes to improve the performance of neural computations.
 %

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Discussion : A

@@
 %%%-----------------------------------------------------------------
-The HD-SNN model that has been trained and evaluated on a naturalistic motion detection task.
-The model has been defined to provide optimal detection of event-driven spatiotemporal motifs.
-We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in the early visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
-%We then demonstrated that the model shows similarities to biological responses.
+The HD-SNN model trained and evaluated on a naturalistic motion detection task.
+The model was defined to provide optimal detection of event-driven spatiotemporal motifs.
+We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in early visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
 We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware.
@@
 
-We would like to highlight a few innovations in the contributions that are presented in this paper.
-First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs.
-Moreover, in comparison to a representation with that algorithm, the parameters of the model (weights and delays) are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection for each spatiotemporal spike motif.
-Another novelty is that the model learns the weights and the delays simultaneously.
+We would like to highlight a few innovations in the contributions presented in this paper.
+First, while~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs.
+Moreover, compared to a representation using this algorithm, the parameters of the model (weights and delays) are explicable, as they directly inform about the logit (inverse sigmoid of the probability) of detection for each spatiotemporal spike motif.
+Another novelty is that the model learns the weights and delays simultaneously.
 For example, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn and their values are frozen during learning.
 In addition, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified toy problems~\citep{gutig_tempotron_2006}.
-Another major contribution is to provide a model that is suitable for learning any kind of spatiotemporal spiking motifs and that can be trained in a supervised manner by providing a dataset of supervision pairs.
+Another major contribution is to provide a model that is suitable for learning any kind of spatiotemporal spiking motif and that can be trained in a supervised manner by providing a dataset of supervised pairs.
 Instead of relying on a careful description of the physical rules governing a task, e.g.
-the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}, this allows a more flexible definition of the model using this properly labelled dataset.
+the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}, this allows a more flexible definition of the model using this properly labeled dataset.
 %

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Discussion : B

@@
 We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, to be able to use fast three-dimensional convolutions.
-We have tested the effect of the width of the time bin and shown that it has essentially no effect on the results presented in this paper.
+We have tested the effect of the width of the time bin and observed it has essentially no effect on the results presented in this paper.
 This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}.
@@
 Such an architecture could provide promising computational speedups.
-% \note{pourrait parler du poster bernstein?
-de nadafian?}
 
-A further limitation is that the model is purely feed-forward.
-Thus, the spikes generated by the postsynaptic neurons are based solely on information contained in the classical receptive field.
-However, it is well known that neurons in the same layer can interact with each other using lateral interactions, for example in V1, and that this can be the basis for computational principles~\citep{chavane_revisiting_2022}.
+Another limitation is that the model is purely feed-forward.
+Thus, the spikes generated by the postsynaptic neurons are based solely on the information contained in the classical receptive field.
+However, it is known that neurons in the same layer can interact with each other through lateral interactions, for example in V1, and that this can be the basis for computational principles~\citep{chavane_revisiting_2022}.
 For example, the combination of neighboring orientations may contribute to image categorization~\citep{perrinet_edge_2015}.
@@
 However, by inserting new spikes into the list of spikes reaching presynaptic addresses, the generic model is able to incorporate them.
-While theoretically possible, this needs to be properly adjusted in practice so that these recurrent connections do not amplify neuronal activity outside a homeostatic state (by extinction or explosion).
-% \note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}%~\citep{perrinet_motion-based_2012}
+While this is theoretically possible, it must be properly tuned in practice so that these recurrent connections do not amplify neuronal activity outside of a homeostatic state (by extinction or explosion).
 
@@
 This has been modeled before to explain, for example, the flash-lag illusion~\citep{khoei_flash-lag_2017}.
-%\note{Say that this could be implemented by properly setting the timing of the label}
 As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by motion, acceleration, jerk,~\ldots), and ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
 Our proposed model using heterogeneous delays provides an alternative and elegant implementation solution to this problem.
+%
 \subsection{Perspectives}

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Discussion : C

@@
 %%%-----------------------------------------------------------------
-% \note{la tache ne met pas en avant les capacites des motifs hetero synaptiques =trop simple et peut se faire en HOTS}
+%
 In defining our task, we emphasized that the generation of events depends on the spatial gradient in each image.
@@
 Taken together, these oriented edges form the contours of visual objects in the scene~\citep{koenderink_representation_1987}.
-Thus, there is an interdependence between the information about motion and the information about orientation within the event stream.
+Thus, there is an interdependence between motion information and orientation information within the event stream.
 It would be crucial to investigate this dependency further.
 This could be initiated by training the model on a dataset with labels that provide local orientation.
-%\note{we can say that we tried it and that it works well}.
-Exploring this dependence will allow us to dissociate these two forms of visual information and enable us to integrate them.
+Exploring this dependence will allow us to dissociate and integrate these two forms of visual information.
 In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka the aperture problem).
@@
 The model is trained on a low-level local motion detection task, and one might wonder if it could be trained on higher-level tasks.
-An example of such a task would be the estimation of depth in the visual scene.
+An example of such a task would be depth estimation in the visual scene.
 There are several sources of information for depth estimation, such as binocular disparity or changes in texture or shading, but in our case motion parallax would be the most important cue~\citep{rogers_motion_1979}.
-This is because objects that are close to an observer will move relatively faster on the retina than an object that is far away, and also because visual occlusions are dependent on the depth order.
+This is because objects that are close to an observer move relatively faster on the retina than an object that is far away, and also because visual occlusions depend on depth order.
 Using this information, it is possible to segment objects and estimate their depth~\citep{yoonessi_contribution_2011}.
-However, this would require the computation of the optical flow first, i.e., the extension of the framework described here for a rigid full-field motion to a more general one where the motion may vary in the visual field.
-A possible implementation is therefore to add a new layer to our model, analogous to the hierarchical organization highlighted in the visual cortex.
+However, this would first require the computation of the optic flow, i.e., the extension of the framework described here for a rigid full-field motion to a more general one where the motion may vary in the visual field.
+One possible implementation is to add a new layer to our model, analogous to the hierarchical organization highlighted in the visual cortex.
 This is theoretically possible by using the output of our model (which estimates motion in retinotopic space) as input to a new layer of neurons that would estimate motion in the visual field, including the depth dimension in the output supervision labels.
@@
 
-In conclusion, the model that we have presented provides a way to process event-based signals in an efficient manner.
-We have shown that we can train the model using a supervised rule, knowing \emph{what} output label, but not \emph{where} it occurs.
+In conclusion, the model we have presented provides a way to efficiently process event-based signals.
+We have shown that we can train the model using a supervised rule that knows \emph{what} output label, but not \emph{where} it occurs.
 Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data~\citep{barlow_unsupervised_1989}.
@@
 These patterns are stereotyped and robust, as they can be activated in the same pattern from day to day~\citep{haimerl_internal_2019}.
-In contrast to conventional methods of processing neurobiological data, such an event-based model would be able to answer key questions about the representation of information in neurobiological data, and it would open up possibilities in the field of computational neuroscience.
-Furthermore, it would open up possibilities in the field of machine learning, especially in computer vision, to address current key concerns such as robustness to attacks, scalability, interpretability, or energy consumption.
+In contrast to conventional methods of processing neurobiological data, such an event-based model would be able to answer key questions about the representation of information in neurobiological data, and it would open possibilities in the field of computational neuroscience.
+Furthermore, it would open possibilities in the field of machine learning, especially in computer vision, to address current key concerns such as robustness to attacks, scalability, interpretability, or energy consumption.
 %

### 2023-06-02 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 We investigate the efficiency of the motion detection mechanism and, in particular, its resilience to synaptic weight pruning.
-Indeed, once trained, the amount of event-driven computation could be drastically reduced by removing weak synapses while maintaining peak performance for the classification task.
+Indeed, once trained, the amount of event-driven computations could be drastically reduced by removing weak synapses while maintaining peak performance for the classification task.
 In this way, we will be able to show how such a model can provide an efficient solution to the energy/accuracy tradeoff.

### 2023-06-02 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 In particular, we will set the spatial extent and duration of the generated movies similar to the motion task defined above.
-This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $XXX$ directions and $YYY3$ speeds.
+This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $XXX8$ directions and $YYY3$ speeds.
 For each motion, we also varied the texture parameters, such as mean and variance of orientation or spatial frequency content, to provide some naturalistic variability.

### 2023-06-08 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 First, we define the ecological cognitive task that the model must solve with the different datasets on which it will be tested.
-Instead of the synthetic textures that have been used to generate motion-driven stimuli, we use natural images and random movements that mimic biological saccades.
-We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model thanks to theoretically defined first principles and derive a learning rule to adapt heterogeneous synaptic delays.
+Instead of the synthetic textures that had been used to generate motion-driven stimuli, we here use natural images and random-walk movements that mimic biological saccades.
+We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model from theoretically defined first principles and derive a learning rule to adapt heterogeneous synaptic delays.
 Using this detection task, we study the emergence of spatiotemporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner.
@@
 %%%-----------------------------------------------------------------
-In this paper, we test whether the HD-SNN model is able to efficiently learn a motion detection task defined by a realistic event-based data stream.
+In this paper, we test whether the HD-SNN model is able to efficiently learn to solve a motion detection task defined by a realistic event-based data stream.
 This type of signal is commonly captured by a Dynamic Vision Sensor (DVS) and is inspired by the signal sent from the retinal ganglion cells through the optic nerve to the visual cortex.
@@
 %%%-----------------------------------------------------------------
-To train and test our model, we need to define a visual dataset for which we supervise the direction and speed of motion.
+To train and validate our model which uses supervised learning, we need to define a visual dataset for which we explictly know the ground truth direction and speed of motion.
 To achieve this, we define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
@@
 To mimic the effect of a saccadic eye movement, this axis moves according to a jagged random walk.
-We show such a trajectory with a length of $128$ time steps (green line).
+We show such a trajectory with a length of $256$ time steps (green line).
 \textit{(Right)}~Snapshots of the synthetic event stream at different time steps (start and end frames marked by a white and black dot, respectively).
@@
 This process is known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $256 \times 256$ in size, and we will extract subimages of size $64 \times 64$ positioned around the center of gaze at each time step.
+These images are $1024 \times 1024$ in size, and we will extract subimages of size $128 \times 128$ positioned around the center of gaze at each time step.
 We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 128~\ms$.
@@
 In a neurobiological recording, this may be the identified set of all presynaptic neurons.
-For neuromorphic hardware, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$, where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities encoded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels.
+For neuromorphic hardware like the output of a DVS, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$, where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities encoded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels.
 Thus, an address $\presynaddr_\arank$ is typically in the form $(\polev_\arank, x_\arank, y_\arank)$ for event-based cameras.
@@
 After training our model, we first analyze the weights learned for the different neurons (\seeFig{kernels}).
-We first notice that these cells are very localized along each trajectory, but show an orientation selectivity similar to that found for MT neurons~\citep{deangelis_functional_1999}.
+We first notice that these cells are very localized around linear trajectories in the space-delay space, and also show an orientation selectivity perpendicular to motion, similarly to what is found for MT neurons~\citep{deangelis_functional_1999}.
 Focusing on the positive weights, a strong selectivity along specific motion axes is observed for each of the different kernels.
 These directions can be easily associated with the direction of motion controlled in the natural images and show a typical anisotropic profile~\citep{kaplan_anisotropic_2013}.
+The negative weights surround these positive weights, providing an enhanced response.
 For example, the first kernel shows a strong selectivity for a downward direction of motion.
-Note that as the delays become larger, the coefficients become lower and less localized, consistent with the average diffusion of information included in the generative model and with the diffusion introduced in motion-based prediction models~\citep{perrinet_motion-based_2012,khoei_flash-lag_2017}.
+Note that as the delays become larger, two effects can be remarked.
+First, coefficients become lower which is consistent with the fact that trajectories are defined in a piecewise fashion, such that such decrease provides with an optimal integration with respect to this limited memory.
+Second, coefficients become less localized compared to the kerenl's spatial profile at short delays, consistent with the average diffusion of information included in the generative model and with the diffusion introduced in motion-based prediction models~\citep{perrinet_motion-based_2012,khoei_flash-lag_2017}.
 We also observed a strong dependence between the weights reaching the ON polarities and those reaching the OFF polarities.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 \newcommand{\WebsiteLP}{https://laurentperrinet.github.io}%
-\newcommand{\Abstract}{
-    The response of a biological neuron depends on the precise timing of spikes from afferent neurons.
-This temporal aspect of the neural code is essential for understanding information processing in neurobiology, and is particularly applicable to the output of neuromorphic hardware such as event-based cameras.
-However, most artificial neural models do not take advantage of this important temporal dimension of the neural code.
-Here, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
-The model takes advantage of the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of synaptic inputs as they reach the basal dendritic tree.
-We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
-We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or event-based cameras.
-We characterize the accuracy of the model in detecting visual motion, and in particular, by pruning some weights, we show that it is still efficient at very low workloads.
-In conclusion, heterogeneous delays provide an end-to-end event-driven computational building block which could improve the performance of future spiking neural network (SNN) algorithms and in particular their applications in neuromorphic chips.
-}
 \newcommand{\Keywords}{time code, event-based computations, spiking neural networks, motion detection, efficient coding, logistic regression
@@
 \abstract{
-\Abstract
+The response of a biological neuron depends on the precise timing of spikes from afferent neurons.
+This temporal aspect of the neural code is essential for understanding information processing in neurobiology, and is particularly applicable to the output of neuromorphic hardware such as event-based cameras.
+However, most artificial neural models do not take advantage of this important temporal dimension of the neural code.
+Here, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
+The model takes advantage of the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of synaptic inputs as they reach the basal dendritic tree.
+We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
+We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or event-based cameras.
+We characterize the accuracy of the model in detecting visual motion, and in particular, by pruning some weights, we show that it is still efficient at very low workloads.
+In conclusion, heterogeneous delays provide an end-to-end event-driven computational building block which could improve the performance of future spiking neural network (SNN) algorithms and in particular their applications in neuromorphic chips.
 }
@@
 After training our model, we first analyze the weights learned for the different neurons (\seeFig{kernels}).
-We first notice that these cells are very localized around linear trajectories in the space-delay space, and also show an orientation selectivity perpendicular to motion, similarly to what is found for MT neurons~\citep{deangelis_functional_1999}.
+We first notice by following each spatial pattern from the shortest delay (right) to the longest that the selectivity of these cells are very localized and are transported around linear trajectories in the space-delay space.
+These show also a relative selectivity to the orientation perpendicular to motion, similarly to what is found for MT neurons~\citep{deangelis_functional_1999}.
+This reflects the way events are generated and in particular the so-called aperture problem which implies that an line moving along its axis would generate no change in luminance and therefore generate no event.
 Focusing on the positive weights, a strong selectivity along specific motion axes is observed for each of the different kernels.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

--- 2023-06-08 Grimaldi-etal-BiolCybernetics.tex
@@
 \abstract{
-\Abstract
+The response of a biological neuron depends on the precise timing of spikes from afferent neurons.
+This temporal aspect of the neural code is essential for understanding information processing in neurobiology, and is particularly applicable to the output of neuromorphic hardware such as event-based cameras.
+However, most artificial neural models do not take advantage of this important temporal dimension of the neural code.
+Here, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
+The model takes advantage of the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of synaptic inputs as they reach the basal dendritic tree.
+We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
+We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or event-based cameras.
+In order to characterize the robustness of the model in detecting visual motion, we test its efficiency by pruning some weights and show it is still efficient at very low workloads.
+In conclusion, heterogeneous delays provide an end-to-end event-driven computational building block which could improve the performance of future spiking neural network (SNN) algorithms and in particular their applications in neuromorphic chips.
 }

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex

--- 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
@@
 We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or event-based cameras.
-We characterize the accuracy of the model in detecting visual motion, and in particular, by pruning some weights, we show that it is still efficient at very low workloads.
+In order to characterize the robustness of the model in detecting visual motion, we test its efficiency by pruning some weights and show it is still efficient at very low workloads.
 In conclusion, heterogeneous delays provide an end-to-end event-driven computational building block which could improve the performance of future spiking neural network (SNN) algorithms and in particular their applications in neuromorphic chips.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %
-After training our model, we first analyze the weights learned for the different neurons (\seeFig{kernels}).
-We first notice by following each spatial pattern from the shortest delay (right) to the longest that the selectivity of these cells are very localized and are transported around linear trajectories in the space-delay space.
+After training our model, let us first analyze the weights learned for the different neurons (\seeFig{kernels}).
+We first notice by following each spatial pattern from the shortest delay (right) to the longest that the selectivity of these cells are very localized and are transported along linear trajectories in the space-delay space.
 These show also a relative selectivity to the orientation perpendicular to motion, similarly to what is found for MT neurons~\citep{deangelis_functional_1999}.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
@@
 %
-After training our model, we first analyze the weights learned for the different neurons (\seeFig{kernels}).
-We first notice that these cells are very localized around linear trajectories in the space-delay space, and also show an orientation selectivity perpendicular to motion, similarly to what is found for MT neurons~\citep{deangelis_functional_1999}.
-Focusing on the positive weights, a strong selectivity along specific motion axes is observed for each of the different kernels.
-These directions can be easily associated with the direction of motion controlled in the natural images and show a typical anisotropic profile~\citep{kaplan_anisotropic_2013}.
-The negative weights surround these positive weights, providing an enhanced response.
-For example, the first kernel shows a strong selectivity for a downward direction of motion.
+After training our model, let us first analyze the weights learned for the different neurons (\seeFig{kernels}).
+We first notice by following each spatial pattern from the shortest delay (right) to the longest (left) that the selectivity of these cells are very localized and are transported along linear trajectories in the space-delay space.
+Focusing on the positive weights, a strong selectivity along specific motion axes is observed for each of the different kernels and these directions can be easily associated with the direction of the corresponding motion.
+The first kernel, for example, shows a strong selectivity for a downward direction of motion.
+The opposite weights surround these positive weights, providing an enhanced response.
+We also observed a strong dependence between the weights reaching the ON polarities and those reaching the OFF polarities.
+In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative for the other.
+This property is due to the way events are generated and the fact that the luminance cannot increase and decrease at the same time.
+Interestingly, the relative organization of the receptive fields that we observe is in quadrature of phase and follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results obtained after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
+Finally, we observe that these receptive fields show also a relative selectivity to the orientation perpendicular to motion, similarly to what is found for MT neurons~\citep{deangelis_functional_1999}.
+This reflects the way events are generated and in particular the so-called aperture problem which implies that an line moving along its axis would generate no change in luminance and therefore generate no event~\citep{perrinet_motion-based_2012}.
+
+If one now widen our focus on the interpretation of these kernels in terms of spatiotemporal motifs embedded in the event stream, these show a prototypical anisotropic profile~\citep{kaplan_anisotropic_2013} adapted to motion detection.
+In~\citep{grimaldi_robust_2022}, a link was drawn between event-based MLR training and Hebbian learning, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions.
+Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatiotemporal spiking motif.
+Each voxel in the 3D kernels defines a specific property by associating a weight to a position and a delay.
+Consequently, our model is able to detect precise spatiotemporal motifs embedded in the spike train and associated to the different motion directions.
 Note that as the delays become larger, two effects can be remarked.
 First, coefficients become lower which is consistent with the fact that trajectories are defined in a piecewise fashion, such that such decrease provides with an optimal integration with respect to this limited memory.
-Second, coefficients become less localized compared to the kerenl's spatial profile at short delays, consistent with the average diffusion of information included in the generative model and with the diffusion introduced in motion-based prediction models~\citep{perrinet_motion-based_2012,khoei_flash-lag_2017}.
-We also observed a strong dependence between the weights reaching the ON polarities and those reaching the OFF polarities.
-In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative for the other.
-This property is due to the way the events are generated and the fact that the luminance cannot increase and decrease at the same time.
-Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results obtained after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
-
-If one focuses on the interpretation of these kernels in terms of spatiotemporal motifs embedded in the event stream, it can lead to interesting outcomes.
-In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions.
-Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatiotemporal spiking motif.
-Each voxel in the 3D kernels defines a specific timestamp and a specific address.
-Consequently, our model is able to detect precise spatiotemporal motifs embedded in the spike train and associated to the different motion directions.
+Second, coefficients become less localized compared to the kernel's spatial profile at short delays, consistent with the average diffusion of information included in the generative model and with the diffusion introduced in motion-based prediction models~\citep{perrinet_motion-based_2012,khoei_flash-lag_2017}.
 The cone shape for the positive weights' distribution highlights a loss of precision for longer delays, i.e.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 First, while~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs.
-Moreover, compared to a representation using this algorithm, the parameters of the model (weights and delays) are explicable, as they directly inform about the logit (inverse sigmoid of the probability) of detection for each spatiotemporal spike motif.
+Moreover, compared to a representation using this algorithm, the parameters of the model (weights and delays) are explicable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif.
+This evidence is measured by the logit of the probability, that is, the inverse sigmoid of the probability.
 Another novelty is that the model learns the weights and delays simultaneously.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
 Here, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
-The model takes advantage of the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of synaptic inputs as they reach the basal dendritic tree.
-We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
-We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or event-based cameras.
+The model takes advantage of the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of temporally precise synaptic inputs as they reach the basal dendritic tree.
+We show that this process can be formalized as a time-invariant logistic regression that can be trained on labeled data.
+We demonstrate its application to  naturalistic videos transformed into event streams similar to the output of the retina or event-based cameras.
 In order to characterize the robustness of the model in detecting visual motion, we test its efficiency by pruning some weights and show it is still efficient at very low workloads.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
correc Miles: abstract + intro

@@
 \abstract{
-The response of a biological neuron depends on the precise timing of spikes from afferent neurons.
-This temporal aspect of the neural code is essential for understanding information processing in neurobiology, and is particularly applicable to the output of neuromorphic hardware such as event-based cameras.
-However, most artificial neural models do not take advantage of this important temporal dimension of the neural code.
-Here, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
-The model takes advantage of the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of temporally precise synaptic inputs as they reach the basal dendritic tree.
+The precise timing of spikes from afferent neurons determines the response of a biological neuron.
+This temporal aspect of the neural activity is essential for understanding information processing in neurobiology, and is particularly applicable to the output of neuromorphic hardware such as event-based cameras.
+Yet, most artificial neural models ignore this important temporal dimension of the neural activity.
+Here, we develop a model for efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
+The model exploits the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of temporally precise synaptic inputs as they reach the basal dendritic tree.
 We show that this process can be formalized as a time-invariant logistic regression that can be trained on labeled data.
-We demonstrate its application to  naturalistic videos transformed into event streams similar to the output of the retina or event-based cameras.
-In order to characterize the robustness of the model in detecting visual motion, we test its efficiency by pruning some weights and show it is still efficient at very low workloads.
-In conclusion, heterogeneous delays provide an end-to-end event-driven computational building block which could improve the performance of future spiking neural network (SNN) algorithms and in particular their applications in neuromorphic chips.
+We demonstrate its application to naturalistic videos transformed into event streams similar to the output of the biological retina or event-based cameras.
+To characterize the robustness of the model in detecting visual motion, we test its efficiency by pruning some weights and show that it is still efficient at very low workloads.
+In conclusion, heterogeneous delays provide an end-to-end, event-driven computational building block that could improve the performance of future spiking neural network (SNN) algorithms and especially their applications in neuromorphic chips.
 }
@@
 One is the SpikeNet algorithm, which takes a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
-Another type of SNN that uses the exact timing of spikes tries to determine the structure of the network to minimize a cost function.
-This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and was recently extended using a surrogate gradient, a method that is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
+Another type of SNN that uses the exact timing of spikes tries to determine the structure of the network to minimize a cost function, as implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002}.
+This was recently extended using the surrogate gradient method which is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
 However, the performance of SNNs still lags behind that of networks using an analog, firing rate-based representation.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
correc Miles: intro

@@
 %
-The human brain has the remarkable capacity to react efficiently at any given time while consuming a reasonable amount of energy, about tens of watts.
+The human brain has the remarkable capacity to react efficiently at any given time while consuming a reasonable amount of energy, in the order of tens of watts.
 This system is the result of millions of years of natural selection, and a striking difference with artificial neural networks is the representation that both use.
-Artificial convolutional neural networks (CNNs), for example, represent the flow of information from one layer to another as analog tensors, storing visual information densely across the visual topography, with different translation-invariant properties represented in different channels.
+For instance, the convolutional neural networks (CNNs) which are used for processing images represent the flow of information from one layer to another using tensors of floats, storing visual information densely across the visual topography, with different translation-invariant properties represented in different channels.
 CNNs have achieved state-of-the-art performance for some computer vision tasks, such as image recognition.
-These networks are also known to mimic many properties of biological systems, such that each can be assigned a ``brain score''~\citep{schrimpf_brain-score_2020}, but this score does not take into account inference speed or energy consumption.
-In fact, they do not take advantage of the dynamics inherent in the way we perceive our natural environment.
-In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
+These networks are also known to mimic many properties of biological systems, such that each can be assigned a ``brain score''~\citep{schrimpf_brain-score_2020}.
+However, this score does not take into account key aspects of biological systems, such as inference speed or energy consumption.
+In particular, they do not take advantage of the dynamics inherent in the way we perceive our natural environment.
+In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing (binary) events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
 Spiking neural networks (SNNs), known as the third generation of artificial neural networks, incorporate this temporal dimension into the way they perform their computations.
 They are of interest to computational neuroscience because they provide a better model of the biological neuron.
-One is the SpikeNet algorithm, which takes a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
+One example of a SNN is the SpikeNet algorithm, which takes a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
 Another type of SNN that uses the exact timing of spikes tries to determine the structure of the network to minimize a cost function, as implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002}.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
correc Miles: intro

@@
 CNNs have achieved state-of-the-art performance for some computer vision tasks, such as image recognition.
-These networks are also known to mimic many properties of biological systems, such that each can be assigned a ``brain score''~\citep{schrimpf_brain-score_2020}.
-However, this score does not take into account key aspects of biological systems, such as inference speed or energy consumption.
-In particular, they do not take advantage of the dynamics inherent in the way we perceive our natural environment.
+These networks are also known to mimic several properties of biological systems, such that each can be assigned a ``brain score''~\citep{schrimpf_brain-score_2020}.
+However, this score does not take into account key aspects of the efficiency of biological systems, such as inference speed or energy consumption.
 In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing (binary) events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
 Spiking neural networks (SNNs), known as the third generation of artificial neural networks, incorporate this temporal dimension into the way they perform their computations.
-They are of interest to computational neuroscience because they provide a better model of the biological neuron.
-One example of a SNN is the SpikeNet algorithm, which takes a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
-Another type of SNN that uses the exact timing of spikes tries to determine the structure of the network to minimize a cost function, as implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002}.
+One example of a SNN is the SpikeNet algorithm, which takes a purely temporal approach by encoding information using at most one spike per neuron~\citep{delorme_spikenet_1999}.
+Alternatively, the SNN implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} that uses the exact timing of spikes and learns the structure of the network by minimizing a specifically defined cost function.
 This was recently extended using the surrogate gradient method which is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
correc Miles: intro

@@
 In a recent review, we reported experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues~\cite{grimaldi_precise_2023}.
+In particular, \citet{abeles_role_1982} asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs.
+While the first hypothesis favors the firing rate coding theory, the second emphasizes on the importance of temporal precision in neural activity.
+Since then, numerous studies have demonstrated the emergence of synchronous activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding using precise spike timings~\citep{perrinet_coding_2004, gollisch_rapid_2008}, or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
+All these findings and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatiotemporal spiking motifs in biological spike trains.
+In neural models, the definition of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows the efficient detection of these spatiotemporal motifs embedded in the spike train.
 These spatiotemporal motifs present in neural activity may form useful representations to perform computations for various cognitive tasks using the synchrony of spikes reaching the soma of a neuron.
-In particular, \citet{abeles_role_1982} asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs.
-While the first possibility favors the firing rate coding theory, the second emphasizes the function of temporal precision in the neural code.
-Since then, numerous studies have demonstrated the emergence of synchronous activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding using spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008}, or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
-All these findings and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatiotemporal spiking motifs in biological spike trains.
-In neural models, the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows the efficient exploitation or detection of these spatiotemporal motifs embedded in the spike train.
 In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a spiking motif defined by a subset of neurons with different, but precise, relative spiking delays.
 This delay between a pair of connected neurons is defined as the time between the emission of a spike at the soma of the afferent neuron and its arrival at the soma of the efferent neuron.
-Due to the variety of weights and delays within a population, representations using polychronous groups have a mathematically much higher information capacity than a firing rate-based approach.
+Importantly, due to the variety of weights and delays within a population, representations using polychronous groups have in theory a much higher information capacity than a firing rate-based approach.
 

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
correc Miles: intro

@@
 
-The present paper proposes a real-world application that extends a recent model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}.
-Indeed, this previous model was trained to solve a simplified motion detection task on a synthetic event-based dataset generated by moving parameterized textures.
-This was a first demonstration that formal neurons can exploit the precise timing of spikes to improve neural computation thanks to heterogeneous delays.
-Here, we extend these results to a more complex and natural setting.
+The present paper proposes a real-world application that extends a recently proposed model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}.
+This model was trained to solve a simplified motion detection task on a synthetic event-based dataset generated by moving parameterized textures and provides with a first demonstration that formal neurons can exploit the precise timing of spikes to detect motion thanks to heterogeneous delays.
+Here, we extend these results to a much more complex and natural setting.
 First, we define the ecological cognitive task that the model must solve with the different datasets on which it will be tested.
-Instead of the synthetic textures that had been used to generate motion-driven stimuli, we here use natural images and random-walk movements that mimic biological saccades.
-We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model from theoretically defined first principles and derive a learning rule to adapt heterogeneous synaptic delays.
-Using this detection task, we study the emergence of spatiotemporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner.
+Instead of the synthetic textures that had been used previously, we here use natural scenes synthesized using natural images that are translated using biologically-inspired saccades.
+We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model from efficiency principles and derive a learning rule to adapt heterogeneous synaptic delays using a gradient descent.
+Applied to this detection task, we study the emergence of spatiotemporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner.
 We investigate the efficiency of the motion detection mechanism and, in particular, its resilience to synaptic weight pruning.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
correc Miles: intro

@@
 Instead of the synthetic textures that had been used previously, we here use natural scenes synthesized using natural images that are translated using biologically-inspired saccades.
-We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model from efficiency principles and derive a learning rule to adapt heterogeneous synaptic delays using a gradient descent.
+We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model from efficiency principles and derive a learning rule to adapt the weights of each heterogeneous synaptic delays using a gradient descent.
 Applied to this detection task, we study the emergence of spatiotemporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
correc Miles: intro

--- 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
@@
 The present paper proposes a real-world application that extends a recently proposed model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}.
-This model was trained to solve a simplified motion detection task on a synthetic event-based dataset generated by moving parameterized textures and provides with a first demonstration that formal neurons can exploit the precise timing of spikes to detect motion thanks to heterogeneous delays.
+This model was trained to solve a simplified motion detection task on a synthetic event-based dataset generated by moving parameterized textures, and provides a first demonstration that formal neurons can exploit the precise timing of spikes to detect motion thanks to heterogeneous delays.
 Here, we extend these results to a much more complex and natural setting.
 First, we define the ecological cognitive task that the model must solve with the different datasets on which it will be tested.
-Instead of the synthetic textures that had been used previously, we here use natural scenes synthesized using natural images that are translated using biologically-inspired saccades.
-We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model from efficiency principles and derive a learning rule to adapt heterogeneous synaptic delays using a gradient descent.
+Instead of the synthetic textures used previously, we use natural scenes synthesized from natural images translated by biologically inspired saccadic movements.
+We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model from efficiency principles and derive a learning rule to adapt the weights of each heterogeneous synaptic delay using gradient descent.
 Applied to this detection task, we study the emergence of spatiotemporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner.
 We investigate the efficiency of the motion detection mechanism and, in particular, its resilience to synaptic weight pruning.
-Indeed, once trained, the amount of event-driven computations could be drastically reduced by removing weak synapses while maintaining peak performance for the classification task.
+Indeed, once trained, the amount of event-driven computation could be drastically reduced by removing weak synapses while maintaining peak performance for the classification task.
 In this way, we will be able to show how such a model can provide an efficient solution to the energy/accuracy tradeoff.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
correc Miles: Methods

@@
 In this paper, we test whether the HD-SNN model is able to efficiently learn to solve a motion detection task defined by a realistic event-based data stream.
-This type of signal is commonly captured by a Dynamic Vision Sensor (DVS) and is inspired by the signal sent from the retinal ganglion cells through the optic nerve to the visual cortex.
+This type of signal is commonly captured by an event-based camera, or Dynamic Vision Sensor (DVS) and is inspired by the signal sent from the retinal ganglion cells through the optic nerve to the visual cortex.
 The events are binary in nature and should be sufficient to perform the motion detection task in a fast and efficient manner.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
         {\bf - Motion Detection Task.} \textit{(Left)}~We use large natural images ($1024\times1024$) in which an aperture ($128\times128$) extracts a cropped image around the view axis.
-To mimic the effect of a saccadic eye movement, this axis moves according to a jagged random walk.
+To mimic the effect of a saccadic eye movement, this image is translated according to a jagged random walk mimicking saccades.
 We show such a trajectory with a length of $256$ time steps (green line).
@@
 In the lower left corner of the snapshots, the translation vector is shown in red as one of the possible classes of motion.
-Note the change in the direction of motion between the third and fourth image, and also that, due to the aperture problem, contours parallel to the motion emit relatively fewer spikes.
+Note the change in the direction of motion between the third and fourth image, and also that contours parallel to the motion generate less luminance changes, the so-called aperture problem, and thus emit relatively fewer spikes.
         }

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
fig 1

@@
     \caption{
-        {\bf - Motion Detection Task.} \textit{(Left)}~We use large natural images ($1024\times1024$) in which an aperture ($128\times128$) extracts a cropped image around the view axis.
-To mimic the effect of a saccadic eye movement, this image is translated according to a jagged random walk mimicking saccades.
-We show such a trajectory with a length of $256$ time steps (green line).
+        {\bf - Motion Detection Task.} \textit{(Left)}~To generate realistic dynamic scenes, we use large natural scenes ($1024\times1024$) from which we crop an image ($128\times128$) which center is moved dynamically at each time frame.
+The center of this cropped image defines the view axis and its dynamics is defined as a jagged random walk to mimick the effect of a biologically realistic saccadic eye movements.
+We show an instance of such a trajectory with a length of $200$ time steps (green line).
 \textit{(Right)}~Snapshots of the synthetic event stream at different time steps (start and end frames marked by a white and black dot, respectively).
 The dynamics of the cropped image translated according to the trajectory as a function of time produces a naturalistic movie, which is then transformed into an event-based representation.
-Mimicking the retina, this representation encodes in each pixel of the image quantal increases or decreases in luminance, i.e.
+Mimicking the retina, this representation encodes in each pixel of the image all-or-none increases or decreases in luminance, i.e.
 ON (red) and OFF (blue) events.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
saccades

@@
 %%%-----------------------------------------------------------------
-To train and validate our model which uses supervised learning, we need to define a visual dataset for which we explictly know the ground truth direction and speed of motion.
+To train and validate our model which uses supervised learning, we need to define a visual dataset for which we explicitly know the ground truth motion, that is, direction and speed.
 To achieve this, we define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
 First, we draw a trajectory inspired by the biological movements of the eyes.
-Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the visual field.
-In animals with a fovea, this is particularly useful because it allows the area with the highest density of photoreceptors in the environment to be moved, for example, to a point of interest.
-A specific mechanism for this function is saccades, which are rapid eye movements that reposition the center of vision.
+Indeed, these movements allow us to dynamically actuate the center of vision in the visual field.
+In animals with a fovea, this is particularly useful as it allows the area with the highest density of photoreceptors to be moved, for example, to a point of interest in the environment.
+A specific mechanism for this function are saccades, which are rapid eye movements that reposition the center of vision.
 In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}), very fast (about $80~\ms$), and have a wide range of speeds.
@@
 This approach first defines a finite set of possible 2D movements in polar coordinates.
-Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $XXX8$ linearly spaced directions and $XXX3$ geometrically spaced speeds.
-Next, we define a gaze trajectory as segments whose duration is drawn from a Poisson distribution with a mean block length of $24~\ms$, similar to a Lévy flight~\citep[p.
+Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $12$ linearly spaced directions and $3$ geometrically spaced speeds.
+Next, we define a gaze trajectory as a sequence of time segments whose durations are drawn from a Poisson distribution with a mean block length of $24~\ms$, similar to a Lévy flight~\citep[p.
 289]{mandelbrot_fractal_1982}.
-Finally, assuming first that the speeds are sampled uniformly and independently from the set of different motions, and second that the motion is uniform during a time segment,  the trajectory is generated by integrating the motion sequence.
+Finally, assuming that motion in one flight is sampled uniformly and independently within the set of different motions, and  that  motion is uniform during that time segment, the trajectory is generated by integrating this motion sequence.
 This generative model yield trajectories that are qualitatively very similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
     \caption{
-        {\bf - Motion Detection Task.} \textit{(Left)}~To generate realistic dynamic scenes, we use large natural scenes ($1024\times1024$) from which we crop an image ($128\times128$) which center is moved dynamically at each time frame.
-The center of this cropped image defines the view axis and its dynamics is defined as a jagged random walk to mimick the effect of a biologically realistic saccadic eye movements.
+        {\bf - Motion Detection Task.} \textit{(Left)}~To generate realistic dynamic scenes, we use large natural scenes ($1024\times1024$) from which we crop an image ($128\times128$) which center is moved dynamically at each time frame according to biologically realistic saccadic eye movements.
+The center of this cropped image defines the view axis and its dynamics is defined as a jagged random walk to mimick the effect of a sequence of saccades.
 We show an instance of such a trajectory with a length of $200$ time steps (green line).
-\textit{(Right)}~Snapshots of the synthetic event stream at different time steps (start and end frames marked by a white and black dot, respectively).
-The dynamics of the cropped image translated according to the trajectory as a function of time produces a naturalistic movie, which is then transformed into an event-based representation.
-Mimicking the retina, this representation encodes in each pixel of the image all-or-none increases or decreases in luminance, i.e.
+\textit{(Right)}~The dynamics of the cropped image translated according to the trajectory as a function of time produces a naturalistic movie, which is then transformed into an event-based representation.
+We show snapshots of the resulting synthetic event stream at different time steps (start and end frames marked by a white and black dot, respectively).
+Mimicking the retina, this representation encodes in each pixel all-or-none increases or decreases in luminance, i.e.
 ON (red) and OFF (blue) events.
-In the lower left corner of the snapshots, the translation vector is shown in red as one of the possible classes of motion.
+In the lower left corner of the snapshots, we show the corresponding instantaneous motion vector (red arrow).
 Note the change in the direction of motion between the third and fourth image, and also that contours parallel to the motion generate less luminance changes, the so-called aperture problem, and thus emit relatively fewer spikes.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
will
will

@@
 Once these eye trajectories are generated, we can apply them to a visual scene.
-For this purpose, we selected a database of $600$ natural images that were previously used to study the statistics of natural images~\citep{perrinet_edge_2015}.
+For this purpose, we selected a database of $600$ natural images that were previously used to study the statistics of natural images~\citep{vanHateren1998}.
 Note that these were pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band.
 This process is known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $1024 \times 1024$ in size, and we will extract subimages of size $128 \times 128$ positioned around the center of gaze at each time step.
-We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 128~\ms$.
-To avoid boundary effects, we will randomly position the full trajectory in image space so that the subimage is translated using the position given by the trajectory at each time step without touching the boundaries.
+These images are $1024 \times 1024$ in size, and we extract subimages of size $128 \times 128$ positioned around the center of gaze at each time step.
+We discretize time in $1~\ms$ bins and produce movies of duration $N_T = 200~\ms$.
+To avoid boundary effects, we  randomly position the full trajectory in image space so that the subimage is translated using the position given by the trajectory at each time step without touching the boundaries.
 At each time step, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a subpixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
@@
 Note that to stay within the framework of a causal computation, the kernels are shifted in time so that only past information gives an answer at the present time.
-This well-known computation defines a differentiable measure, which is very efficiently implemented for GPUs, and which we will use to learn the classification of different motifs in the event stream.
+This well-known computation defines a differentiable measure, which is very efficiently implemented for GPUs, and which we use to learn the classification of different motifs in the event stream.
 A similar type of spatiotemporal filtering is used as a preprocessing stage for an existing pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
@@
 where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function, and $\bias_\class$ is the bias associated with the class $\class$.
-In particular, we expect that some specific motifs will become tightly synchronized as they reach the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike.
+In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike.
 The spiking output of the model corresponds to an event with the highest probability class.
@@
         \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot a 2D representation of the input event stream as a raster plot (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time).
-A spatiotemporal convolution is applied to the dense representation of the input with $2$ different convolution kernels (the green and orange kernels), which will define the output channels.
+A spatiotemporal convolution is applied to the dense representation of the input with $2$ different convolution kernels (the green and orange kernels), which define the output channels.
 The convolution is summed over the two polarities.
@@
 Note that in this context, the label is known but the position is not, mainly due to the sparse spatial content of natural images.
-However, the supervised optimization of this MLR model will adjust the weights of the kernels.
+However, the supervised optimization of this MLR model adjusts the weights of the kernels.
 As a result, the error is only propagated back to the spatial locations of these most active cells.
@@
 If one now widen our focus on the interpretation of these kernels in terms of spatiotemporal motifs embedded in the event stream, these show a prototypical anisotropic profile~\citep{kaplan_anisotropic_2013} adapted to motion detection.
-In~\citep{grimaldi_robust_2022}, a link was drawn between event-based MLR training and Hebbian learning, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions.
+In~\citep{grimaldi_robust_2022}, a link was drawn between event-based MLR training and Hebbian learning, allowing to say that the present model learns its weights according to a presynaptic activity associated to the different motion directions.
 Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatiotemporal spiking motif.
@@
 %%%-----------------------------------------------------------------
-To further test our model, we will quantify its ability to categorize different motions, i.e., on event streams for which the ground truth motion is known at each instant.
-Before applying the model to natural images, we will first test the model on simpler, parameterized stimuli.
-For this purpose, we will use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control, among other parameters, the motion (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
-In particular, we will set the spatial extent and duration of the generated movies similar to the motion task defined above.
+To further test our model, we quantify its ability to categorize different motions, i.e., on event streams for which the ground truth motion is known at each instant.
+Before applying the model to natural images, we first test the model on simpler, parameterized stimuli.
+For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control, among other parameters, the motion (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
+In particular, we set the spatial extent and duration of the generated movies similar to the motion task defined above.
 This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $XXX8$ directions and $YYY3$ speeds.
@@
 
-To assess the robustness of the classification as a function of computational load, we will prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold.
+To assess the robustness of the classification as a function of computational load, we prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold.
 In Fig.~\ref{fig:accuracy} we plot the classification accuracy as a function of the relative number of computations or active weights per decision for each neuron in the layer.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
2.1

@@
 This approach first defines a finite set of possible 2D movements in polar coordinates.
-Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $12$ linearly spaced directions and $3$ geometrically spaced speeds.
+Based on the distribution of biological movements, we simplified that set by selecting eye movements from the Cartesian product of $12$ linearly spaced directions and $3$ geometrically spaced speeds.
 Next, we define a gaze trajectory as a sequence of time segments whose durations are drawn from a Poisson distribution with a mean block length of $24~\ms$, similar to a Lévy flight~\citep[p.
 289]{mandelbrot_fractal_1982}.
-Finally, assuming that motion in one flight is sampled uniformly and independently within the set of different motions, and  that  motion is uniform during that time segment, the trajectory is generated by integrating this motion sequence.
+Finally, assuming that  motion is stationary during each segment and that motions for each flight are sampled uniformly and independently, the global trajectory is generated by integrating this motion sequence.
 This generative model yield trajectories that are qualitatively very similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
@@
 At each time step, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a subpixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
-Note that the magnitude of the displacement is relative to the time bin, and we have defined the unit speed to correspond to a movement of one pixel per frame (i.e., per time bin).
+Note that the magnitude of the displacement is relative to the time bin, and we have defined the unit speed to correspond to a movement of one pixel per frame (i.e., per time bin of $1~\ms$).
 
-To transform each movie into events, as recorded by a DVS, we compute a gradient image (initialized at zero) by computing the temporal gradient of the pixels' intensity over two successive frames.
+To transform each movie into events, as recorded by a DVS, we compute a residual gradient image which we initialize at zero.
+We then compute the temporal gradient of the pixels' intensity over two successive frames.
 For a given pixel and time stamp, an event is generated when the absolute value of this gradient exceeds a threshold.
@@
 a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF), and their time $\timev_\arank$ (\seeFig{motion_task}-\textit{(Right)}).
-The goal here is to infer the correct motion solely by observing these events.%%%-----------------------------------------------------------------
+Ultimately, the goal of the model is to infer the correct motion solely by observing these events.%%%-----------------------------------------------------------------
 \subsection{The Heterogeneous Delays SNN (HD-SNN) model}

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
2.2

@@
 In a neurobiological recording or in the sensory signal obtained from an event-based camera, the input consists of a stream of \emph{spikes} or \emph{events}.
-This can be formalized as a list of tuples, each tuple representing neural addresses and timestamps.
-We defined this list as $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{(Top-Left} for an illustration).
+Formally, this can be defined as a list of tuples, each tuple representing the neural address and timestamp.
+We denote this list as $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{(Top-Left} for an illustration).
 Events are usually ordered by their time of occurrence.
@@
 In biology, a single cortical neuron typically has several thousand synapses.
-Each synapse can be defined by its synaptic weight and its delay, the time it takes for a spike to travel from the soma of the presynaptic neuron to the soma of the postsynaptic neuron.
+Each synapse can be defined by its synaptic weight and its delay, that is, the time it takes for a spike to travel from the soma of the presynaptic neuron to the soma of the postsynaptic neuron.
 A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$, but also by the set of possible delays.

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Methods

@@
 where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function, and $\bias_\class$ is the bias associated with the class $\class$.
-In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike.
+In particular, we expect that some specific motifs may generate tightly synchronized outputs as they are integrated in the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike.
 The spiking output of the model corresponds to an event with the highest probability class.
 
-Now that this general framework has been explained, we can add some heuristics, based on neuroscientific observations, to constrain our model and its strategies for solving the task used to solve the ecological task described in section~\ref{sec:task}.
+Now that this general framework has been explained, we can include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}.
 Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel neuroscience-inspired heuristics are the main methodological differences with that previous study.

### 2023-06-15 Grimaldi-etal-BiolCybernetics.tex
Methods : full rewrite

@@
 On a more microscopic scale, the human gaze moves with a continuous drift similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
-To preserve the full generality of the task, we define eye movements using such a form of random walk~\citep{engbert_integrated_2011}.
-This approach first defines a finite set of possible 2D movements in polar coordinates.
-Based on the distribution of biological movements, we simplified that set by selecting eye movements from the Cartesian product of $12$ linearly spaced directions and $3$ geometrically spaced speeds.
-Next, we define a gaze trajectory as a sequence of time segments whose durations are drawn from a Poisson distribution with a mean block length of $24~\ms$, similar to a Lévy flight~\citep[p.
+To preserve the full generality of the task, we define eye movements using such a form of random walk~\citep{engbert_integrated_2011}: We first define a finite set of possible motions in polar coordinates as the Cartesian product of $12$ regularly spaced directions and $3$ geometrically spaced speeds.
+Next, we define a saccadic path as a sequence of time segments whose durations are drawn from a Poisson distribution with a mean block length of $24~\ms$, similar to a Lévy flight~\citep[p.
 289]{mandelbrot_fractal_1982}.
-Finally, assuming that  motion is stationary during each segment and that motions for each flight are sampled uniformly and independently, the global trajectory is generated by integrating this motion sequence.
+Finally, assuming that  motion is stationary during each segment and that motions for each flight are drawn uniformly and independently, the global trajectory is generated by integrating this motion sequence.
 This generative model yield trajectories that are qualitatively very similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
@@
     \caption{
-        {\bf - Motion Detection Task.} \textit{(Left)}~To generate realistic dynamic scenes, we use large natural scenes ($1024\times1024$) from which we crop an image ($128\times128$) which center is moved dynamically at each time frame according to biologically realistic saccadic eye movements.
-The center of this cropped image defines the view axis and its dynamics is defined as a jagged random walk to mimick the effect of a sequence of saccades.
-We show an instance of such a trajectory with a length of $200$ time steps (green line).
-\textit{(Right)}~The dynamics of the cropped image translated according to the trajectory as a function of time produces a naturalistic movie, which is then transformed into an event-based representation.
+        {\bf - Motion Detection Task.} \textit{(Left)}~To generate realistic dynamic scenes, we mimick the effect of minute saccadic eye movements (with a length of $200~\ms$, green line) over a large natural scene ($1024\times1024$).
+For this, we crop an image ($128\times128$, here shown at time step $20~\ms$) which center is moved dynamically  according to a jagged random walk.
+\textit{(Right)}~The dynamics of the cropped image translated according to the saccadic path as a function of time produces a naturalistic movie, which is then transformed into an event-based representation.
 We show snapshots of the resulting synthetic event stream at different time steps (start and end frames marked by a white and black dot, respectively).
-Mimicking the retina, this representation encodes in each pixel all-or-none increases or decreases in luminance, i.e.
+Mimicking the retina, this representation encodes at each pixel all-or-none increases or decreases in luminance, i.e.
 ON (red) and OFF (blue) events.
@@
 For this purpose, we selected a database of $600$ natural images that were previously used to study the statistics of natural images~\citep{vanHateren1998}.
-Note that these were pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band.
-This process is known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $1024 \times 1024$ in size, and we extract subimages of size $128 \times 128$ positioned around the center of gaze at each time step.
+Note that these were pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band, similarly to a process known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
+These images are $1024 \times 1024$ in size, and we crop images of size $128 \times 128$ positioned around the center of gaze at each time step.
 We discretize time in $1~\ms$ bins and produce movies of duration $N_T = 200~\ms$.
@@
 a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF), and their time $\timev_\arank$ (\seeFig{motion_task}-\textit{(Right)}).
-Ultimately, the goal of the model is to infer the correct motion solely by observing these events.%%%-----------------------------------------------------------------
+Ultimately, the goal of the model is to infer the correct motion solely by observing these events.
+%%%%-----------------------------------------------------------------
 \subsection{The Heterogeneous Delays SNN (HD-SNN) model}
+%%%-----------------------------------------------------------------
 %----------------------------%
@@
     \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}
-    \caption{\textbf{The core mechanism of the HD SNN model.} \textit{(Top-Left)}~Two spiking motifs are emitted from four presynaptic neurons.
-Once integrated by the synapses of the postsynaptic neuron~\textit{(Right)}, the spiking motifs are shifted in time by the synaptic delays and weighted by the synaptic weights~\textit{(Middle-Left)}.
-When they reach the soma of the postsynaptic neuron, the different spikes contribute to a modification of its membrane potential according to an activation function.
-In this example, we use the same activation function as for a Leaky Integrate and Fire neuron~\textit{(Bottom-Left)}.
+    \caption{\textbf{The core mechanism of the HD SNN model.}~\textit{(Left-Top)}~Four presynaptic neurons show some spiking activity in which a spiking motif is embedded (starting at time $t=50~\ms$).~\textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses).~\textit{(Left-Middle)} Each spike is weighted by the synaptic weights (height of the bar) and shifted in time by the synaptic delays on each synapse of the postsynaptic neuron (input spikes are shown in light gray for comparison).
+As a result, the spikes from the spiking motif are synchronized as they reach the soma of the postsynaptic neuron.~\textit{(Left-Bottom)} These spikes are then integrated, and the different spikes contribute to a modification of its membrane potential according to the neural activation function.
+In this example, we use the same activation function from a Leaky Integrate-and-Fire neuron.
 The first spiking motif is synchronized by the synaptic delays and causes a sudden rise in the membrane potential of the postsynaptic neuron.
-An output spike is emitted when the membrane potential reaches the threshold and it is then reset.
-\textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses).
-}%
+An output spike is emittedat time $t=75~\ms$ when the membrane potential reaches the threshold, and it is then reset.}%
     \label{fig:izhikevich}%
@@
 %----------------------------%
-In a neurobiological recording or in the sensory signal obtained from an event-based camera, the input consists of a stream of \emph{spikes} or \emph{events}.
+In this task, the input consists of a stream of \emph{spikes} or \emph{events}, a property shared in the signal obtained from an event-based camera or from a neurobiological recording of the activity of single units.
 Formally, this can be defined as a list of tuples, each tuple representing the neural address and timestamp.
-We denote this list as $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{(Top-Left} for an illustration).
+We denote this list as $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N}$ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{Left-Top} for an illustration).
 Events are usually ordered by their time of occurrence.
@@
 This defines an address space $\presynaddrspace$, which consists of the set of possible addresses.
-In a neurobiological recording, this may be the identified set of all presynaptic neurons.
-For neuromorphic hardware like the output of a DVS, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$, where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities encoded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels.
-Thus, an address $\presynaddr_\arank$ is typically in the form $(\polev_\arank, x_\arank, y_\arank)$ for event-based cameras.
+In neurobiological spiking activity, this may be the identified set of all recorded neurons.
+For neuromorphic hardware like the output of a DVS or our task, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$, where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities encoded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels.
+Thus, each address $\presynaddr_\arank$ is typically in the form $(\polev_\arank, x_\arank, y_\arank)$ for event-based cameras.
 
-In the HD-SNN model, neurons $\postsynaddr \in \postsynaddrspace$ are connected to presynaptic afferent neurons from $\presynaddrspace$.
+In the HD-SNN model, neurons $\postsynaddr \in \postsynaddrspace$ are connected to presynaptic afferent neurons from $\presynaddrspace$ using realistic synapses.
 In biology, a single cortical neuron typically has several thousand synapses.
 Each synapse can be defined by its synaptic weight and its delay, that is, the time it takes for a spike to travel from the soma of the presynaptic neuron to the soma of the postsynaptic neuron.
-A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$, but also by the set of possible delays.
 Note that a neuron can contact another afferent neuron with different delays through different synaptic connections.
-By scanning all $\postsynaddr$ neurons, we define the full set of $\Nsyn$ synapses,
-as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated with a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
+By scanning all postsynaptic neurons $\postsynaddr$, we may thus define the full set of $\Nsyn$ synapses, as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated with a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
 This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{(Right)} for an illustration of the connectivity of a neuron with synaptic weights and delays).
+% A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$, but also by the set of possible delays.
 
-Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$, or the emitting field of a presynaptic neuron $\synapse_\presynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$.
-As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field.
-This results in a weighted event stream (\seeFig{izhikevich}-\textit{(Middle-Left}) for each  postsynaptic neuron $\postsynaddr$:
+Of interest is to define the emitting field of a presynaptic neuron $\synapse_\presynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$, or also the receptive field of a postsynaptic neuron $\synapse^\postsynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$.
+In particular, when driven by a stream of spikes $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$, each incoming spike is multiplexed by the synapses of the receptive field $\synapse^\postsynaddr$ of postsynaptic neuron $\postsynaddr$.
+This results in a weighted event stream (\seeFig{izhikevich}-\textit{Left-Middle}) for each postsynaptic neuron $\postsynaddr$:
 %
@@
 In simulations, however, it should be properly reordered.
-The activation function of the HD-SNN neurons can be selected from the full range of spiking neuron response functions (\seeFig{izhikevich}).
-Crucially, when postsynaptic neurons are activated at their soma by a specific spatiotemporal motif imprinted in the synaptic set, the discharge probability increases, and especially when these spikes converge at the soma in a synchronous manner.
-We have defined the HD-SNN model in all generality in this subsection, and in the next subsection we describe an implementation of our model adapted to the motion detection task.
+Once transformed by the synaptic connectivity, this weighted event stream may be integrated, for instance as the membrane potential of a Leaky-Integrate-and-Fire neuron (\seeFig{izhikevich}-\textit{Left-Bottom}), yet the activation function of the HD-SNN neurons can be selected from the full range of spiking neuron response functions.
+Importantly, this activation function has to be such that when postsynaptic neurons are activated at their soma by a specific spatiotemporal motif imprinted in the synaptic set, and such that these spikes converge at the soma in a synchronous manner, the discharge probability should increase.
+In this subsection, we have defined the HD-SNN model in all generality, and in the next subsection we describe an implementation of our model adapted to the motion detection task.
 %
-\subsection{Application of the HD-SNN model to the motion detection task}
+% \subsection{Application of the HD-SNN model to the motion detection task}
+\subsection{Application of HD-SNN to motion detection}
 %
-It is possible to define a specific implementation of this model in order to adapt it to common tasks in computer vision.
-In particular, from the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation.
+Indeed, it is possible to adapt the HD-SNN model specifically for common computer vision tasks.
+First, neural adresses are defined such as to represent the range of possible positions and polarities.
+Then, in the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and benefit from parallel computing acceleration, it is useful to transform this event-based representation into a dense representation.
 Indeed, by discretizing time, we can transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.
 The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
-First, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:kernel_b} corresponds to a temporal convolution that transforms the input $A$ with each kernels and positions assigned to the postsynaptic neurons~\citep{grimaldi_learning_2022}.
-This takes advantage of the position invariance observed in images and exploited in convolutional neural networks, and we can further assume that synaptic motifs should be similar across different positions.
-As a consequence, we can define a spatiotemporal convolutional operator, in which layer of neurons assigned to specific kernels form channels.
-
-Therefore, assuming that a neuron integrates only in the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
+Similarly, one may discretize the connectivity of the HD-SNN model defined above.
 The longest synaptic delay defines the depth $\Ktime$ of the kernel $\kernel^\postsynaddr$, so that all possible delays associated with the different presynaptic addresses are represented.
@@
 For convenience, the connectivity of the neuron $\postsynaddr$ is defined locally around its position $(x_\postsynaddr, y_\postsynaddr)$.
-Using this dense representation, the model's processing of $A$ can be written as %.
+
+Then, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:stream_b} corresponds to a weighted reordering of the input $A$ with each kernels and positions assigned to the postsynaptic neurons~\citep{grimaldi_learning_2022}.
+We will here define the activity $B_\class$ of post-synaptic neurons as the integration of this activity in each voxel:
 %
@@
 \forall x, y, \timev, \; \;
-\mathcal{C}_\class(x, y, \timev)
+B_\class(x, y, \timev)
 = \sum_{p, \delta_x, \delta_y, \synapticdelay} \kernel_\class(p, \delta_x, \delta_y, \synapticdelay) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay)
@@
 where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses within a kernel and $\delta_\timev$ is the synaptic delay.
-This shows that $\mathcal{C}_\class$ is the result of a spatiotemporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses: $\mathcal{C}_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration).
-Note that to stay within the framework of a causal computation, the kernels are shifted in time so that only past information gives an answer at the present time.
-This well-known computation defines a differentiable measure, which is very efficiently implemented for GPUs, and which we use to learn the classification of different motifs in the event stream.
+In this formulation, we recognize that it takes advantage of the position invariance observed in images and exploited in convolutional neural networks.
+Here, we further assume that synaptic motifs should be similar across different times.
+As a consequence, this defines a 3D, spatiotemporal convolutional operator, in which layer of neurons assigned to specific kernels form channels.
+Using this dense representation, the model's processing of the inpu $A$ can be written as $\forall \class$, $B_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration).
 A similar type of spatiotemporal filtering is used as a preprocessing stage for an existing pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
 In addition,~\citep{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm that implements a motion estimation task.
-By assuming a locally constant motion, the authors assume that the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the motion.
+By assuming locally a constant motion, the authors assume that the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the motion.
+Note that to stay within the framework of a causal computation, the kernels are shifted in time so that only past information gives an answer at the present time that is, after a time equal to the depth $\Ktime$ of the kernel.
+This well-known computation defines a differentiable measure, which is very efficiently implemented for GPUs, and which we will use to learn the classification of different motifs in the event stream.
+% Therefore, assuming that a neuron integrates only in the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
 
-Here, we keep the analogy with spiking neurons, and we try to observe the detection of spiking motifs using the spatiotemporal kernels.
-If so, this precise spatiotemporal patterns prove to be of interest for neural computations and one can infer that biological neurons make use of this information as well.
+Keeping the analogy with spiking neurons, the analog activity $B$ represents the integration of synaptic activity and we will now try to define the detection of spiking motifs using the spatiotemporal kernels.
 The 3D convolution represents the linear integration of the spike train as the linear input to the neuron.
-The activation function of our model is then a softmax function that implements a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}.
+It represents the evidence accumulated from a set of sources in logistic regression.
+Since we know that at each instant, there is one and only one motion, we will define the activation function of our model as a softmax function that implements a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, and analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}.
 In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a softmax function of the linear combination of the list of events.
@@
 Pr(k=\class \; \vert \; x, y, \timev) =
-\frac {\exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class) }{Z(x, y, \timev)}
+\frac {\exp (B_\class(x, y, \timev) +\bias_\class) }{Z(x, y, \timev)}
 \end{equation} 
@@
 %
-where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function, and $\bias_\class$ is the bias associated with the class $\class$.
+where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp (B_\class(x, y, \timev) +\bias_\class)$ is the partition function, and $\bias_\class$ is the bias associated with the class $\class$. 
 In particular, we expect that some specific motifs may generate tightly synchronized outputs as they are integrated in the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike.
-The spiking output of the model corresponds to an event with the highest probability class.
+The spiking output of the model corresponds for instance to the event with the highest probability.
 
 Now that this general framework has been explained, we can include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}.
-Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel neuroscience-inspired heuristics are the main methodological differences with that previous study.
-First, in our simulations, we set the size of the kernels to $(\Ktime, \Kx, \Ky) = (21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = XXX8 \times XXX3$.
-To avoid introducing biases in the directions, we apply a circular mask to the spatial dimensions of the kernels.
+Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve  a deeper analysis of the results, these novel neuroscience-inspired heuristics are the main methodological differences with that previous study.
+First, in our simulations, we set the size of the kernels to $(\Ktime, \Kx, \Ky) = (21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = 8 \times 3$.
+To avoid introducing biases in the directions which may be learned, we apply a circular mask to the spatial dimensions of the kernels.
 We also included a prior in the selectable motions, as there is a prior for slow speeds in natural scenes~\citep{vacher_bayesian_2018}.
@@
 Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities.
-This is due to the fact that our whitened images have a symmetry in the luminance profiles, and an image with inverted contrast is indistinguishable from a standard one.
+This is due to the fact that our whitened images have a relative symmetry in the luminance profiles, that is, that an image with inverted contrast is indistinguishable from a standard one.
 Since this arrangement of polarities is independent of motion, we added a mechanism that collects the linear values for the movie and the movie with the ON and OFF cells flipped, keeping only the maximum value for each voxel.
@@
 However, the supervised optimization of this MLR model adjusts the weights of the kernels.
-As a result, the error is only propagated back to the spatial locations of these most active cells.
+As a result, we observed that the error is only propagated back to the spatial locations of these most active cells.
 This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation.
-Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies, each of size $TTT \times XXX \times YYY$ and a learning rate of $10^{-5}$).
+Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{10}$ movies, each of size $200 \times 128 \times 128$, a learning rate of $10^{-5}$ and $25$ epochs).
 We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
@@
 Thus, by taking the mean logit of the probability of the output given by the model at all positions for any given time, we can calculate the probability of the output at that time.
-This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted).
+This allows one to calculate the accuracy as the percentage of times the motion is accurately predicted at any given time step.
 These calculations are performed on a different input data set than the one used in the training or validation steps.

### 2023-06-15 Grimaldi-etal-BiolCybernetics.tex
Methods : full rewrite

@@
 % \newcommand{\simil}{\beta} % similarity value
-\newcommand{\Nclass}{N_\text{class}} % number of classes for MLR:
 \newcommand{\Nx}{N_\text{X}}
@@
 \newcommand{\Ntime}{N_\text{T}}
-\newcommand{\kernel}{\mathcal{K}} % convolution kernel
+\newcommand{\kernel}{K} % convolution kernel
+% \newcommand{\kernel}{\mathcal{K}} % convolution kernel
 %\newcommand{\kernelind}{\mathbf{k}} % indice of the kernel
@@
 \newcommand{\class}{c} % class k of the MLR
+% \newcommand{\Nclass}{N_\text{class}} % number of classes for MLR:
+\newcommand{\Nclass}{N_\class} % number of classes for MLR:
 \newcommand{\lrweights}{\theta} % matrix of MLR weights
@@
     \centering
+    % TODO: corriger post to pre
     \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}
@@
 Indeed, it is possible to adapt the HD-SNN model specifically for common computer vision tasks.
-First, neural adresses are defined such as to represent the range of possible positions and polarities.
-Then, in the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and benefit from parallel computing acceleration, it is useful to transform this event-based representation into a dense representation.
-Indeed, by discretizing time, we can transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.
+First, neural addresses are defined such as to represent the range of possible positions and polarities.
+Then, in the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and benefit from parallel computing acceleration, we transform this event-based representation into a dense representation.
+Indeed, by discretizing time, we transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.
 The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
 Similarly, one may discretize the connectivity of the HD-SNN model defined above.
-The longest synaptic delay defines the depth $\Ktime$ of the kernel $\kernel^\postsynaddr$, so that all possible delays associated with the different presynaptic addresses are represented.
-In particular, for each class $\class$ of the supervision task, the entire synaptic set can be represented as a kernel, which is represented by the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
-To keep the analogy with the HD-SNN model, for a neuron $\postsynaddr$ of position $(x_\postsynaddr, y_\postsynaddr)$ and class $\class$, $\kernel_\class$ gives the weights as a function of relative addresses and synaptic delays: $\forall {\delta_x \in [1, \Kx], \delta_y \in [1, \Ky], \delta_\timev \in [1, \Ktime], \polev \in [1, \Npol]}, \; \kernel_\class(\polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$.
-For convenience, the connectivity of the neuron $\postsynaddr$ is defined locally around its position $(x_\postsynaddr, y_\postsynaddr)$.
+The longest synaptic delay defines the depth $\Ktime$ of the kernel, so that all possible delays associated with the different presynaptic addresses are represented.
+In particular, for each class $\class$ of the supervision task, the entire synaptic set can be represented as a kernel, which is represented by the dense matrix $\kernel$ of size $(\Nclass, \Npol, \Ktime, \Kx, \Ky)$, where $\Nclass$ is the number of classes, $\Ktime$ the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
+To keep the analogy with the HD-SNN model, $\kernel$ gives the weights as a function of class $\class$, polarity, synaptic delay and relative position: $\forall {\polev \in [1, \Npol], \delta_\timev \in [1, \Ktime], \delta_x \in [1, \Kx], \delta_y \in [1, \Ky]}, \; \kernel(\class, \polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$.
+Such a kernel defines a dense representation of the synaptic set.
 
 Then, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:stream_b} corresponds to a weighted reordering of the input $A$ with each kernels and positions assigned to the postsynaptic neurons~\citep{grimaldi_learning_2022}.
-We will here define the activity $B_\class$ of post-synaptic neurons as the integration of this activity in each voxel:
+We will here define the activity $B$ of post-synaptic neurons as the integration of this activity in each voxel and for each channel $\class$:
 %
@@
 \forall x, y, \timev, \; \;
-B_\class(x, y, \timev)
-= \sum_{p, \delta_x, \delta_y, \synapticdelay} \kernel_\class(p, \delta_x, \delta_y, \synapticdelay) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay)
+B(\class, x, y, \timev)
+= \sum_{p, \delta_x, \delta_y, \synapticdelay} \kernel(\class, p, \delta_x, \delta_y, \synapticdelay) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay)
 \end{equation}
@@
 In this formulation, we recognize that it takes advantage of the position invariance observed in images and exploited in convolutional neural networks.
-Here, we further assume that synaptic motifs should be similar across different times.
+Here, we further assume that synaptic motifs should be similar across different times as defined in the temporal convolution.
 As a consequence, this defines a 3D, spatiotemporal convolutional operator, in which layer of neurons assigned to specific kernels form channels.
-Using this dense representation, the model's processing of the inpu $A$ can be written as $\forall \class$, $B_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration).
+Using this dense representation, the model's processing of the input $A$ can be written as layer-wise convolution as used in CNNs: $B = \kernel \ast A$ (\seeFig{model} for an illustration).
 A similar type of spatiotemporal filtering is used as a preprocessing stage for an existing pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
-In addition,~\citep{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm that implements a motion estimation task.
+In addition,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm that implements a motion estimation task.
 By assuming locally a constant motion, the authors assume that the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the motion.
-Note that to stay within the framework of a causal computation, the kernels are shifted in time so that only past information gives an answer at the present time that is, after a time equal to the depth $\Ktime$ of the kernel.
+For convenience, the connectivity of the neuron $\postsynaddr$ is defined locally around its position $(x_\postsynaddr, y_\postsynaddr)$.
+Note also that to stay within the framework of a causal computation, the kernels are shifted in time so that only past information gives an answer at the present time that is, after a time equal to the depth $\Ktime$ of the kernel.
 This well-known computation defines a differentiable measure, which is very efficiently implemented for GPUs, and which we will use to learn the classification of different motifs in the event stream.
@@
 \begin{equation}\label{eq:mlr}
-\forall x, y, \timev, \forall \class \in [1, N_\class], \; \;
+\forall x, y, \timev, \forall \class \in [1, \Nclass], \; \;
 Pr(k=\class \; \vert \; x, y, \timev) =
@@
 %
-where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp (B_\class(x, y, \timev) +\bias_\class)$ is the partition function, and $\bias_\class$ is the bias associated with the class $\class$. 
+where $Z(x, y, \timev)=\sum_{\class \in [1, \Nclass]} \exp (B_\class(x, y, \timev) +\bias_\class)$ is the partition function, and $\bias_\class$ is a scalar representing the bias associated with the class $\class$.
 In particular, we expect that some specific motifs may generate tightly synchronized outputs as they are integrated in the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike.
@@
 
-Now that this general framework has been explained, we can include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}.
+Now that this general framework has been explained, we may include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}.
 Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve  a deeper analysis of the results, these novel neuroscience-inspired heuristics are the main methodological differences with that previous study.
-First, in our simulations, we set the size of the kernels to $(\Ktime, \Kx, \Ky) = (21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = 8 \times 3$.
+First, in our simulations, we define as many classes as the number of motions (directions and velocities): $\Nclass = 8 \times 3$ and set the size of the model's kernel to $(\Nclass, \Npol, \Ktime, \Kx, \Ky) = (24, 2, 21, 17, 17)$ (that is, $291312$ parameters).
 To avoid introducing biases in the directions which may be learned, we apply a circular mask to the spatial dimensions of the kernels.
@@
 \textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated with thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated with a particular motion class.
-Note that the position of the output spikes does not systematically correspond to the position of the input spikes, but only when enough evidence is obtained.
+Note that the position of the output spikes does not systematically correspond to the position of the input spikes, but only when enough evidence is reached.
     }
@@
 This rule was implemented using the binary input events as inputs and the corresponding motion's labels as the desired output.
-The loss function of the MLR model is the binary cross entropy at the output of the classification layer.
+The loss function of the MLR model is the binary cross entropy of the output of the classification layer knowing the ground truth.
 The labels were defined at each time point as a one-hot encoding of the current motion in the channel corresponding to the current motion for all positions.
-Note that in this context, the label is known but the position is not, mainly due to the sparse spatial content of natural images.
+Note that in this context, the label is known, but the position is not, mainly due to the sparse spatial content of natural images.
 However, the supervised optimization of this MLR model adjusts the weights of the kernels.

### 2023-06-15 Grimaldi-etal-BiolCybernetics.tex
Methods : full rewrite

@@
 To keep the analogy with the HD-SNN model, $\kernel$ gives the weights as a function of class $\class$, polarity, synaptic delay and relative position: $\forall {\polev \in [1, \Npol], \delta_\timev \in [1, \Ktime], \delta_x \in [1, \Kx], \delta_y \in [1, \Ky]}, \; \kernel(\class, \polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$.
-Such a kernel defines a dense representation of the synaptic set.
+In our simulations, we define as many classes as the number of motions (directions and velocities): $\Nclass = 8 \times 3$ and set the size of the model's kernel to $(\Nclass, \Npol, \Ktime, \Kx, \Ky) = (24, 2, 21, 17, 17)$ (that is, $291312$ parameters).
+Such a kernel defines a dense representation of the full synaptic set.
 
 Then, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:stream_b} corresponds to a weighted reordering of the input $A$ with each kernels and positions assigned to the postsynaptic neurons~\citep{grimaldi_learning_2022}.
-We will here define the activity $B$ of post-synaptic neurons as the integration of this activity in each voxel and for each channel $\class$:
+Let's define \emph{evidence} as the logit of a probability, that is, the inverse sigmoid of that probability.
+By this definition of evidence, logistic regression takes advantage of the fact that if different independent observations (here, the estimated motion at different spatial locations and timings) share a common cause (here, the rigid local motion of the image on the receptive field), then an optimal estimate of the evidence of this motion is the sum of the evidences from the independent sources.
+Interpreting the weights of the kernel as evidences (also called factors in logistic regression), we may therefore define the activity $B$ of post-synaptic neurons as the integration of this activity in each voxel and for each channel $\class$ in order to infer the evidence of each motion:
 %
@@
 
-Keeping the analogy with spiking neurons, the analog activity $B$ represents the integration of synaptic activity and we will now try to define the detection of spiking motifs using the spatiotemporal kernels.
+Keeping the analogy with spiking neurons, the analog activity $B$ represents the integration of synaptic activity, and we will now try to define the detection of spiking motifs using the spatiotemporal kernels.
 The 3D convolution represents the linear integration of the spike train as the linear input to the neuron.
@@
 Now that this general framework has been explained, we may include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}.
-Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve  a deeper analysis of the results, these novel neuroscience-inspired heuristics are the main methodological differences with that previous study.
-First, in our simulations, we define as many classes as the number of motions (directions and velocities): $\Nclass = 8 \times 3$ and set the size of the model's kernel to $(\Nclass, \Npol, \Ktime, \Kx, \Ky) = (24, 2, 21, 17, 17)$ (that is, $291312$ parameters).
-To avoid introducing biases in the directions which may be learned, we apply a circular mask to the spatial dimensions of the kernels.
+Note that the general framework is an extension to that presented in~\cite{grimaldi_learning_2022}, in particular by including the more complex task to solve, the deeper analysis of the results, and these novel neuroscience-inspired heuristics.
+First, to avoid introducing biases in the directions which may be learned, we apply a circular mask to the spatial dimensions of the kernels.
 We also included a prior in the selectable motions, as there is a prior for slow speeds in natural scenes~\citep{vacher_bayesian_2018}.
@@
 %
-Since the model is fully differentiable, we can now implement a supervised learning rule.
+Since the model is fully differentiable, we can now implement a supervised learning rule to learn the weights of this kernel.
 This rule was implemented using the binary input events as inputs and the corresponding motion's labels as the desired output.
@@
 Such an output provides a form of optical flow that can be exploited for non-rigid motion, but we have defined here, for simplicity, an evaluation method that applies to our full-field motion task.
-We have shown above that if different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities.
-Thus, by taking the mean logit of the probability of the output given by the model at all positions for any given time, we can calculate the probability of the output at that time.
+Using the properties of logistic regression, by taking the mean evidence represented in the output given by the model at all positions for any given time, and using the sigmoid function, we can derive the probability of each motion at that time.
 This allows one to calculate the accuracy as the percentage of times the motion is accurately predicted at any given time step.
@@
 First, while~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs.
-Moreover, compared to a representation using this algorithm, the parameters of the model (weights and delays) are explicable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif.
-This evidence is measured by the logit of the probability, that is, the inverse sigmoid of the probability.
+Moreover, compared to a representation using this algorithm, the parameters of the model (weights and delays) are explicable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif, where we define \emph{evidence} as the logit of the probability, that is, the inverse sigmoid of the probability.
 Another novelty is that the model learns the weights and delays simultaneously.

### 2023-06-15 Grimaldi-etal-BiolCybernetics.tex
correc Miles: Results

@@
     \caption{
-    	Representation of the weights for $XXX8$ directions for a single speed (among the $XXX8 \times XXX3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
+    	Representation of the weights for $8$ directions for a single speed (among the $8 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
 The directions are shown as red arrows in the left insets.
@@
 %
-After training our model, let us first analyze the weights learned for the different neurons (\seeFig{kernels}).
+After training our model, let us first examine the emerging weights learned for the different motions (\seeFig{kernels}).
 We first notice by following each spatial pattern from the shortest delay (right) to the longest (left) that the selectivity of these cells are very localized and are transported along linear trajectories in the space-delay space.
@@
 Interestingly, the relative organization of the receptive fields that we observe is in quadrature of phase and follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results obtained after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
-Finally, we observe that these receptive fields show also a relative selectivity to the orientation perpendicular to motion, similarly to what is found for MT neurons~\citep{deangelis_functional_1999}.
+Finally, we observe that these receptive fields show also a relative selectivity to the orientation perpendicular to motion, similarly to what is found for neurons in cortical area MT which is known to be selective to visual motions~\citep{deangelis_functional_1999}.
 This reflects the way events are generated and in particular the so-called aperture problem which implies that an line moving along its axis would generate no change in luminance and therefore generate no event~\citep{perrinet_motion-based_2012}.
@@
 Note that as the delays become larger, two effects can be remarked.
-First, coefficients become lower which is consistent with the fact that trajectories are defined in a piecewise fashion, such that such decrease provides with an optimal integration with respect to this limited memory.
+First, coefficients become lower which is consistent with the fact that trajectories are defined in a piece wise fashion, such that such decrease provides with an optimal integration with respect to this limited memory~\citep{Pasturel2020}.
 Second, coefficients become less localized compared to the kernel's spatial profile at short delays, consistent with the average diffusion of information included in the generative model and with the diffusion introduced in motion-based prediction models~\citep{perrinet_motion-based_2012,khoei_flash-lag_2017}.
-The cone shape for the positive weights' distribution highlights a loss of precision for longer delays, i.e.
+This conic shape along the motion's trajectory highlights a loss of precision for longer delays, i.e.
 events away in the past.
@@
     \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
+    % TODO : show prototypes of textures on top
     \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from a grating (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to grating (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)).
@@
 %%%-----------------------------------------------------------------
-To further test our model, we quantify its ability to categorize different motions, i.e., on event streams for which the ground truth motion is known at each instant.
-Before applying the model to natural images, we first test the model on simpler, parameterized stimuli.
+To test our model, we will quantify its ability to categorize different motions, i.e., on event streams for which the ground truth motion is known at each instant.
+Before applying the model to natural scenes that we defined in the motion detection task, we first test the model on simpler, parameterized stimuli.
 For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control, among other parameters, the motion (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
@@
 We observe some interesting facts.
-First, as we change the mean spatial frequency of the texture, we observe a monotonic decrease in accuracy.
+First, as we change the mean spatial frequency of the texture, we observe a broadly tuned response in accuracy.
 This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006} and reveals the most informative scales learned by our model.
-We observe the trend for different spatial frequency bandwidths.
+We observe a similar trend for different spatial frequency bandwidths.
 In particular, the accuracy is better for a grating-like stimulus than for a large one (which qualitatively resembles a more textured stimulus), reminiscent of the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}.
@@
 %
-\subsection{Accuracy efficiency trade-off}% for the motion detection task}
+\subsection{Accuracy versus efficiency trade-off}% for the motion detection task}
 %%%-----------------------------------------------------------------
@@
 If the information support is sparse, then computations can only be performed on those events.
-Also, if we set some weights of the kernels to zero, then the sum for those addresses can be dropped.
+Also, if we set some weights of the kernels to zero, then the sum for those addresses can be dropped and computations can be performed only on those events.
 Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of non-zero synaptic weights.

### 2023-06-15 Grimaldi-etal-BiolCybernetics.tex
correc Miles: Results

@@
 %%%-----------------------------------------------------------------
-The HD-SNN model trained and evaluated on a naturalistic motion detection task.
-The model was defined to provide optimal detection of event-driven spatiotemporal motifs.
-We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in early visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
+The HD-SNN model was trained and evaluated on a naturalistic motion detection task with realistic eye movements.
+The model was defined to provide optimal detection of event-driven spatiotemporal motifs and learns kernels similar to those found in the visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
 We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware.
@@
 Moreover, compared to a representation using this algorithm, the parameters of the model (weights and delays) are explicable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif, where we define \emph{evidence} as the logit of the probability, that is, the inverse sigmoid of the probability.
-Another novelty is that the model learns the weights and delays simultaneously.
-For example, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn and their values are frozen during learning.
+Another novelty is that the model learns simultaneously weights and delays.
+In contrast, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn and their values are frozen during learning.
 In addition, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified toy problems~\citep{gutig_tempotron_2006}.
@@
 We have identified a number of limitations of our model, which we will now discuss in detail.
-First, the entire framework is based on a discrete binning of time, which is not compatible with the continuous nature of biological time.
+First, this implementation of the HD-SNN model is based on a discrete binning of time, which is not compatible with the continuous nature of biological time.
 We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, to be able to use fast three-dimensional convolutions.
@@
 This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}.
-This suggests the possibility of analytically including a precision term in the temporal value of the input spikes.
+This suggests the advantage of analytically including a precision term in the temporal value of the input spikes.
 This mechanism may be implemented by the filtering implemented by the synaptic time constant of about $5~\ms$.
 Furthermore, it is possible to circumvent the need for time discretization by the use of a purely event-based scheme.
-In fact, it is not necessary to compute the voltage traces between two spikes~\citep{hanuschkin_general_2010}.
-Thus, it is possible to define a purely event-based framework.
+In fact, it is possible to derive event-triggered computations of the continuous activity of the SNN~\citep{hanuschkin_general_2010} and thus to define a purely event-based framework.
 Such an architecture could provide promising computational speedups.
@@
 For example, the combination of neighboring orientations may contribute to image categorization~\citep{perrinet_edge_2015}.
-Furthermore, neural information is modulated by feedback information, e.g.
+Furthermore, neural information may be modulated by feedback information, e.g.
 to distinguish a figure from its background~\citep{roelfsema_early_2016}.
@@
 
-Such recurrent activity would be essential for the implementation of predictive or anticipatory processes.
+Such recurrent activity would be essential for the implementation of predictive or anticipatory processes~\citep{Benvenuti2020}.
 This is essential in a neural system because it contains several different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}.
 This has been modeled before to explain, for example, the flash-lag illusion~\citep{khoei_flash-lag_2017}.
-As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by motion, acceleration, jerk,~\ldots), and ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
+As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by motion, acceleration, jerk,~\ldots), and  knowing that ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
 Our proposed model using heterogeneous delays provides an alternative and elegant implementation solution to this problem.
@@
 However, this would first require the computation of the optic flow, i.e., the extension of the framework described here for a rigid full-field motion to a more general one where the motion may vary in the visual field.
-One possible implementation is to add a new layer to our model, analogous to the hierarchical organization highlighted in the visual cortex.
+One possible implementation is to add a new layer to our model, analogous to the hierarchical organization which is prevalent in the visual cortex.
 This is theoretically possible by using the output of our model (which estimates motion in retinotopic space) as input to a new layer of neurons that would estimate motion in the visual field, including the depth dimension in the output supervision labels.

### 2023-06-15 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
Update Grimaldi-etal-BiolCybernetics.tex
conclusion

@@
 % https://github.com/SpikeAI/pyTERtorch/ (for dev)
-% https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN 
+% https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN (TODO: to become public)
 % https://www.zotero.org/groups/4776796/fastmotiondetection
@@
 The HD-SNN model was trained and evaluated on a naturalistic motion detection task with realistic eye movements.
-The model was defined to provide optimal detection of event-driven spatiotemporal motifs and learns kernels similar to those found in the visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
+The model was defined such as to provide an optimal detection of  spatiotemporal motifs and learns kernels similar to those found in the visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
 We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware.
@@
 We would like to highlight a few innovations in the contributions presented in this paper.
-First, while~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs.
-Moreover, compared to a representation using this algorithm, the parameters of the model (weights and delays) are explicable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif, where we define \emph{evidence} as the logit of the probability, that is, the inverse sigmoid of the probability.
+First, while~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of spatiotemporal spiking motifs.
+Moreover, compared to classical CNN solutions, the parameters of the model (weights and delays) are explicable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif, where we define \emph{evidence} as the logit of the probability, that is, the inverse sigmoid of the probability.
 Another novelty is that the model learns simultaneously weights and delays.
-In contrast, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn and their values are frozen during learning.
+In contrast, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn at initialization and their values are frozen during learning.
 In addition, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified toy problems~\citep{gutig_tempotron_2006}.
@@
 First, this implementation of the HD-SNN model is based on a discrete binning of time, which is not compatible with the continuous nature of biological time.
-We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, to be able to use fast three-dimensional convolutions.
+We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, in particular to be able to use fast, differentiable three-dimensional convolutions.
 We have tested the effect of the width of the time bin and observed it has essentially no effect on the results presented in this paper.
 This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}.
-This suggests the advantage of analytically including a precision term in the temporal value of the input spikes.
+This suggests the potential advantage of analytically including an additional precision term to the temporal value of input spikes.
 This mechanism may be implemented by the filtering implemented by the synaptic time constant of about $5~\ms$.
@@
 Thus, the spikes generated by the postsynaptic neurons are based solely on the information contained in the classical receptive field.
-However, it is known that neurons in the same layer can interact with each other through lateral interactions, for example in V1, and that this can be the basis for computational principles~\citep{chavane_revisiting_2022}.
+However, it is known that neurons in the same layer can interact with each other through lateral interactions, for example in V1, and that this can be the basis for more complex computational principles~\citep{chavane_revisiting_2022}.
 For example, the combination of neighboring orientations may contribute to image categorization~\citep{perrinet_edge_2015}.
@@
 Currently, mainly due to our use of convolutions, it is not possible to implement these recurrent connections in our implementation (lateral or feedback).
-However, by inserting new spikes into the list of spikes reaching presynaptic addresses, the generic model is able to incorporate them.
-While this is theoretically possible, it must be properly tuned in practice so that these recurrent connections do not amplify neuronal activity outside of a homeostatic state (by extinction or explosion).
+However, by inserting new spikes into the list of spikes reaching presynaptic addresses, the generic HD-SNN model is able to incorporate them.
+While this is theoretically possible, it must be properly tuned in practice so that these recurrent connections do not bring neuronal activity outside a homeostatic state (by extinction or explosion).
 
 Such recurrent activity would be essential for the implementation of predictive or anticipatory processes~\citep{Benvenuti2020}.
-This is essential in a neural system because it contains several different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}.
+This is essential in a neural system because it contains several delays that require temporal alignment~\citep{hogendoorn_predictive_2019}.
 This has been modeled before to explain, for example, the flash-lag illusion~\citep{khoei_flash-lag_2017}.
@@
 
-In conclusion, the model we have presented provides a way to efficiently process event-based signals.
+In conclusion, the HD-SNN model that we have presented provides a way to efficiently process event-based signals.
 We have shown that we can train the model using a supervised rule that knows \emph{what} output label, but not \emph{where} it occurs.

### 2023-06-15 Grimaldi-etal-BiolCybernetics.tex
Perrinet2002

@@
 While the first hypothesis favors the firing rate coding theory, the second emphasizes on the importance of temporal precision in neural activity.
-Since then, numerous studies have demonstrated the emergence of synchronous activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding using precise spike timings~\citep{perrinet_coding_2004, gollisch_rapid_2008}, or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
+Since then, numerous studies have demonstrated the emergence of synchronous activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding using precise spike timings~\citep{Perrinet2002,perrinet_coding_2004, gollisch_rapid_2008}, or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
 All these findings and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatiotemporal spiking motifs in biological spike trains.

### 2023-06-15 Grimaldi-etal-BiolCybernetics.tex
Perrinet2007

@@
 On a more microscopic scale, the human gaze moves with a continuous drift similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
-To preserve the full generality of the task, we define eye movements using such a form of random walk~\citep{engbert_integrated_2011}: We first define a finite set of possible motions in polar coordinates as the Cartesian product of $12$ regularly spaced directions and $3$ geometrically spaced speeds.
+To preserve the full generality of the task, we define eye movements using such a form of random walk~\citep{engbert_integrated_2011}: We first define a finite set of possible motions in polar coordinates as the Cartesian product of the $12$ regularly spaced directions and $3$ geometrically spaced speeds.
 Next, we define a saccadic path as a sequence of time segments whose durations are drawn from a Poisson distribution with a mean block length of $24~\ms$, similar to a Lévy flight~\citep[p.
@@
 We show snapshots of the resulting synthetic event stream at different time steps (start and end frames marked by a white and black dot, respectively).
-Mimicking the retina, this representation encodes at each pixel all-or-none increases or decreases in luminance, i.e.
-ON (red) and OFF (blue) events.
+Mimicking the response of ganglion cells in the retina, this representation encodes at each pixel all-or-none increases or decreases in luminance, i.e.
+ON (red) and OFF (blue) spikes.
 In the lower left corner of the snapshots, we show the corresponding instantaneous motion vector (red arrow).
-Note the change in the direction of motion between the third and fourth image, and also that contours parallel to the motion generate less luminance changes, the so-called aperture problem, and thus emit relatively fewer spikes.
+Note the change in the direction of motion between the third and fourth image, and also that contours parallel to the motion generate less luminance changes, the so-called aperture problem, and thus produce relatively fewer spikes.
         }
@@
 In particular, we set the spatial extent and duration of the generated movies similar to the motion task defined above.
-This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $XXX8$ directions and $YYY3$ speeds.
+This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $12$ directions and $3$ speeds.
 For each motion, we also varied the texture parameters, such as mean and variance of orientation or spatial frequency content, to provide some naturalistic variability.
@@
 First, as we change the mean spatial frequency of the texture, we observe a broadly tuned response in accuracy.
-This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006} and reveals the most informative scales learned by our model.
+This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006,Perrinet2007} and reveals the most informative scales learned by our model.
 We observe a similar trend for different spatial frequency bandwidths.

### 2023-06-15 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex
methods

@@
     \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}
-    \caption{\textbf{The core mechanism of the HD SNN model.}~\textit{(Left-Top)}~Four presynaptic neurons show some spiking activity in which a spiking motif is embedded (starting at time $t=50~\ms$).~\textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses).~\textit{(Left-Middle)} Each spike is weighted by the synaptic weights (height of the bar) and shifted in time by the synaptic delays on each synapse of the postsynaptic neuron (input spikes are shown in light gray for comparison).
-As a result, the spikes from the spiking motif are synchronized as they reach the soma of the postsynaptic neuron.~\textit{(Left-Bottom)} These spikes are then integrated, and the different spikes contribute to a modification of its membrane potential according to the neural activation function.
-In this example, we use the same activation function from a Leaky Integrate-and-Fire neuron.
+    \caption{\textbf{Core mechanism of the HD-SNN model.} \textit{(Left-Top)}~Four presynaptic neurons show some spiking activity in which a spiking motif is embedded (starting at time $t=50~\ms$).
+\textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses).
+\textit{(Left-Middle)}~Each spike is weighted by the synaptic weights (height of the bar) and shifted in time according to the synaptic delays on each respective synapse (input spikes are shown in light gray for comparison).
+As a result, the spikes from the spiking motif are synchronized as they reach the soma of the postsynaptic neuron.
+\textit{(Left-Bottom)}~These spikes are then integrated, and the different spikes contribute to a modification of its membrane potential according to the neural activation function.
+In this example, we use the activation function of a Leaky Integrate-and-Fire neuron.
 The first spiking motif is synchronized by the synaptic delays and causes a sudden rise in the membrane potential of the postsynaptic neuron.
-An output spike is emittedat time $t=75~\ms$ when the membrane potential reaches the threshold, and it is then reset.}%
+An output spike is emitted at time $t=75~\ms$ when the membrane potential reaches the threshold, and it is then reset.}%
     \label{fig:izhikevich}%
@@
 By scanning all postsynaptic neurons $\postsynaddr$, we may thus define the full set of $\Nsyn$ synapses, as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated with a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
-This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{(Right)} for an illustration of the connectivity of a neuron with synaptic weights and delays).
+This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{Right} for an illustration of the connectivity of a neuron with synaptic weights and delays).
 % A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$, but also by the set of possible delays.
@@
 To keep the analogy with the HD-SNN model, $\kernel$ gives the weights as a function of class $\class$, polarity, synaptic delay and relative position: $\forall {\polev \in [1, \Npol], \delta_\timev \in [1, \Ktime], \delta_x \in [1, \Kx], \delta_y \in [1, \Ky]}, \; \kernel(\class, \polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$.
-In our simulations, we define as many classes as the number of motions (directions and velocities): $\Nclass = 8 \times 3$ and set the size of the model's kernel to $(\Nclass, \Npol, \Ktime, \Kx, \Ky) = (24, 2, 21, 17, 17)$ (that is, $291312$ parameters).
+In our simulations, we define as many classes as the number of motions (directions and velocities): $\Nclass = 12 \times 3$ and set the size of the model's kernel to $(\Nclass, \Npol, \Ktime, \Kx, \Ky) = (36, 2, 21, 17, 17)$ (that is, $436968$ parameters).
 Such a kernel defines a dense representation of the full synaptic set.
@@
 As a consequence, this defines a 3D, spatiotemporal convolutional operator, in which layer of neurons assigned to specific kernels form channels.
-Using this dense representation, the model's processing of the input $A$ can be written as layer-wise convolution as used in CNNs: $B = \kernel \ast A$ (\seeFig{model} for an illustration).
+Using this dense representation, the model's processing of the input $A$ can be written as layer-wise convolution similar to that used in CNNs: $B = \kernel \ast A$ (\seeFig{model} for an illustration).
 A similar type of spatiotemporal filtering is used as a preprocessing stage for an existing pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
@@
 %
+% TODO : change to logistic regression
 \begin{linenomath*}
@@
 
-Now that this general framework has been explained, we may include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}.
+Now that this general framework has been described formally, we may include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}.
 Note that the general framework is an extension to that presented in~\cite{grimaldi_learning_2022}, in particular by including the more complex task to solve, the deeper analysis of the results, and these novel neuroscience-inspired heuristics.
@@
 Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration).
-This is consistent with neuroscientific principles because, due to the limits of conduction delay along horizontal connections, the synaptic delay is physically related to the distance between the presynaptic and postsynaptic neurons.
+This is consistent with neuroscientific principles because, due to the limits of conduction delay, the synaptic delay is physically related to the distance between the presynaptic and postsynaptic neurons.
 Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities.
@@
 %
-Since the model is fully differentiable, we can now implement a supervised learning rule to learn the weights of this kernel.
+Since the model is fully differentiable, we can now implement a supervised learning rule to learn the weights of the model's kernel.
 This rule was implemented using the binary input events as inputs and the corresponding motion's labels as the desired output.
@@
 Such an output provides a form of optical flow that can be exploited for non-rigid motion, but we have defined here, for simplicity, an evaluation method that applies to our full-field motion task.
-Using the properties of logistic regression, by taking the mean evidence represented in the output given by the model at all positions for any given time, and using the sigmoid function, we can derive the probability of each motion at that time.
+Using the properties of logistic regression, by taking the mean evidence represented in the output given by the model at all positions for any given time, and using the sigmoid function, we can derive each motion's probability at that time.
 This allows one to calculate the accuracy as the percentage of times the motion is accurately predicted at any given time step.
@@
     \caption{
-    	Representation of the weights for $8$ directions for a single speed (among the $8 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
+    	Representation of the weights for $XX$ directions for a single speed (among the $12 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
 The directions are shown as red arrows in the left insets.

### 2023-06-16 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 In this paper, we test whether the HD-SNN model is able to efficiently learn to solve a motion detection task defined by a realistic event-based data stream.
-This type of signal is commonly captured by an event-based camera, or Dynamic Vision Sensor (DVS) and is inspired by the signal sent from the retinal ganglion cells through the optic nerve to the visual cortex.
+This type of signal is commonly captured by an event-based camera, or Dynamic Vision Sensor (DVS), and is inspired by the signal sent from the retinal ganglion cells through the optic nerve to the visual cortex.
 The events are binary in nature and should be sufficient to perform the motion detection task in a fast and efficient manner.
@@
 To achieve this, we define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
-First, we draw a trajectory inspired by the biological movements of the eyes.
+First, we draw a trajectory inspired by biological eye movements.
 Indeed, these movements allow us to dynamically actuate the center of vision in the visual field.
 In animals with a fovea, this is particularly useful as it allows the area with the highest density of photoreceptors to be moved, for example, to a point of interest in the environment.
-A specific mechanism for this function are saccades, which are rapid eye movements that reposition the center of vision.
+A specific mechanism for this function is saccades, which are rapid eye movements that reposition the center of vision.
 In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}), very fast (about $80~\ms$), and have a wide range of speeds.
@@
 Finally, assuming that  motion is stationary during each segment and that motions for each flight are drawn uniformly and independently, the global trajectory is generated by integrating this motion sequence.
-This generative model yield trajectories that are qualitatively very similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
+This generative model yields trajectories that are qualitatively very similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
 
@@
     \caption{
-        {\bf - Motion Detection Task.} \textit{(Left)}~To generate realistic dynamic scenes, we mimick the effect of minute saccadic eye movements (with a length of $200~\ms$, green line) over a large natural scene ($1024\times1024$).
-For this, we crop an image ($128\times128$, here shown at time step $20~\ms$) which center is moved dynamically  according to a jagged random walk.
-\textit{(Right)}~The dynamics of the cropped image translated according to the saccadic path as a function of time produces a naturalistic movie, which is then transformed into an event-based representation.
+        {\bf - Motion Detection Task.} \textit{(Left)}~To generate realistic dynamic scenes, we mimic the effect of minute saccadic eye movements (with a length of $200~\ms$, green line) over a large natural scene ($1024\times1024$).
+For this, we crop an image ($128\times128$, for which we show luminance contrasts at time step $20~\ms$).
+The center of this aperture is moved dynamically according to a jagged random walk.
+\textit{(Right)}~The dynamics of the cropped image translated according to the saccadic path as a function of time produce a naturalistic movie, which is then transformed into an event-based representation.
 We show snapshots of the resulting synthetic event stream at different time steps (start and end frames marked by a white and black dot, respectively).
@@
 In the lower left corner of the snapshots, we show the corresponding instantaneous motion vector (red arrow).
-Note the change in the direction of motion between the third and fourth image, and also that contours parallel to the motion generate less luminance changes, the so-called aperture problem, and thus produce relatively fewer spikes.
+Note the change in the direction of motion between the third and fourth image, and also that contours parallel to the motion generate fewer luminance changes, the so-called aperture problem, and thus produce relatively fewer spikes.
         }
@@
 We discretize time in $1~\ms$ bins and produce movies of duration $N_T = 200~\ms$.
-To avoid boundary effects, we randomly position the full trajectory in image space so that the subimage is translated using the position given by the trajectory at each time step without touching the boundaries.
+To avoid boundary effects, we randomly position the full trajectory in image space so that the sub-image is translated using the position given by the trajectory at each time step without touching the boundaries.
 At each time step, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a subpixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
@@
 As a result, the spikes from the spiking motif are synchronized as they reach the soma of the postsynaptic neuron.
-\textit{(Left-Bottom)}~These spikes are then integrated, and the different spikes contribute to a modification of its membrane potential according to the neural activation function.
+\textit{(Left-Bottom)}~These spikes are then integrated, and contribute to a modification of the membrane potential of the output neuron according to the neural activation function.
 In this example, we use the activation function of a Leaky Integrate-and-Fire neuron.
@@
 B(\class, x, y, \timev)
-= \sum_{p, \delta_x, \delta_y, \synapticdelay} \kernel(\class, p, \delta_x, \delta_y, \synapticdelay) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay)
+= \sum_{p, \delta_x, \delta_y, \synapticdelay_\timev} \kernel(\class, p, \delta_x, \delta_y, \synapticdelay_\timev) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay_\timev)
 \end{equation}

### 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
abstract

@@
 The precise timing of spikes from afferent neurons determines the response of a biological neuron.
-This temporal aspect of the neural activity is essential for understanding information processing in neurobiology, and is particularly applicable to the output of neuromorphic hardware such as event-based cameras.
-Yet, most artificial neural models ignore this important temporal dimension of the neural activity.
-Here, we develop a model for efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
-The model exploits the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of temporally precise synaptic inputs as they reach the basal dendritic tree.
+This temporal aspect of neural activity is essential for understanding information processing in neurobiology, and is particularly applicable to the output of neuromorphic hardware such as event-based cameras.
+However, most artificial neural models ignore this important temporal dimension of neural activity.
+Here, we develop a model for efficiently detecting temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
+The model exploits the property that the diversity of synaptic delays on the dendritic tree allows specific arrangements of temporally precise synaptic inputs to synchronize as they reach the basal dendritic tree.
 We show that this process can be formalized as a time-invariant logistic regression that can be trained on labeled data.
-We demonstrate its application to naturalistic videos transformed into event streams similar to the output of the biological retina or event-based cameras.
+We demonstrate its application to naturalistic videos transformed into event streams, similar to the output of the biological retina or event-based cameras.
 To characterize the robustness of the model in detecting visual motion, we test its efficiency by pruning some weights and show that it is still efficient at very low workloads.
-In conclusion, heterogeneous delays provide an end-to-end, event-driven computational building block that could improve the performance of future spiking neural network (SNN) algorithms and especially their applications in neuromorphic chips.
+In conclusion, heterogeneous delays provide an end-to-end, event-driven computational building block that could improve the performance of future spiking neural network (SNN) algorithms and, in particular, their applications in neuromorphic chips.
 }
@@
 
+% TODO citer Levy2021 "Communication consumes 35 times more energy than computation in the human cortex" 
 Finally, the output of the MLR model is an event-based representation that predicts the probability of each motion at each position and time.

### 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
 To characterize the robustness of the model in detecting visual motion, we test its efficiency by pruning some weights and show that it is still efficient at very low workloads.
-In conclusion, heterogeneous delays provide an end-to-end, event-driven computational building block that could improve the performance of future spiking neural network (SNN) algorithms and, in particular, their applications in neuromorphic chips.
+In conclusion, by providing an end-to-end, event-driven computational building block, the learning of heterogeneous delays could improve the performance of future spiking neural network (SNN) algorithms, and in particular their applications in neuromorphic chips.
 }
@@
 %
-The human brain has the remarkable capacity to react efficiently at any given time while consuming a reasonable amount of energy, in the order of tens of watts.
+The human brain has the remarkable capacity to react efficiently at any given time while consuming a reasonable amount of energy, in the order of 20 watts.
 This system is the result of millions of years of natural selection, and a striking difference with artificial neural networks is the representation that both use.
@@
 These networks are also known to mimic several properties of biological systems, such that each can be assigned a ``brain score''~\citep{schrimpf_brain-score_2020}.
-However, this score does not take into account key aspects of the efficiency of biological systems, such as inference speed or energy consumption.
+However, this score does not take into account key aspects of the efficiency of biological systems, such as inference speed or energy consumption, which is about hundreds of watts on a standard GPU.
 In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing (binary) events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
@@
 
-In a recent review, we reported experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues~\cite{grimaldi_precise_2023}.
+In a recent review, we reported previous theoretical and  experimental evidence for the use of precise spiking motifs in biological neural networks~\cite{grimaldi_precise_2023}.
 In particular, \citet{abeles_role_1982} asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs.

### 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
caption Motion Detection Task figure

@@
     \caption{
-        {\bf - Motion Detection Task.} \textit{(Left)}~To generate realistic dynamic scenes, we mimic the effect of minute saccadic eye movements (with a length of $200~\ms$, green line) over a large natural scene ($1024\times1024$).
-For this, we crop an image ($128\times128$, for which we show luminance contrasts at time step $20~\ms$).
-The center of this aperture is moved dynamically according to a jagged random walk.
-\textit{(Right)}~The dynamics of the cropped image translated according to the saccadic path as a function of time produce a naturalistic movie, which is then transformed into an event-based representation.
-We show snapshots of the resulting synthetic event stream at different time steps (start and end frames marked by a white and black dot, respectively).
-Mimicking the response of ganglion cells in the retina, this representation encodes at each pixel all-or-none increases or decreases in luminance, i.e.
-ON (red) and OFF (blue) spikes.
+        {\bf Motion Detection Task.} To generate realistic event-based dynamic scenes, we mimic the effect of minute saccadic eye movements on a large natural scene ($1024\times1024$) by extracting an image ($128\times128$) which center is moving dynamically according to a jagged random walk.
+\textit{(Left)}~We show an instance of this trajectory (with a length of $200~\ms$, green line) superimposed on the luminance contrasts observed at time step $20~\ms$.
+\textit{( Right)}~The dynamics of this image translated according to the saccadic trajectory produces a naturalistic movie, which is then transformed into an event-based representation.
+We show snapshots of the resulting synthetic event stream at different time steps (start and end frames are marked by a white and black dot, respectively, in the left inset).
+Mimicking the response of ganglion cells in the retina, this representation encodes at each pixel all-or-none increases or decreases in luminance, i.e., ON (red) and OFF (blue) spikes.
 In the lower left corner of the snapshots, we show the corresponding instantaneous motion vector (red arrow).
-Note the change in the direction of motion between the third and fourth image, and also that contours parallel to the motion generate fewer luminance changes, the so-called aperture problem, and thus produce relatively fewer spikes.
+Note the change in the direction of motion between the third and fourth frames, and also that contours parallel to the motion produce fewer luminance changes, the so-called aperture problem, and thus relatively fewer spikes.
         }

### 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Since then, numerous studies have demonstrated the emergence of synchronous activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding using precise spike timings~\citep{Perrinet2002,perrinet_coding_2004, gollisch_rapid_2008}, or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
-All these findings and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatiotemporal spiking motifs in biological spike trains.
+All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatiotemporal spiking motifs in biological spike trains.
 In neural models, the definition of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows the efficient detection of these spatiotemporal motifs embedded in the spike train.

### 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
HD-SNN

--- 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
@@
 The events are binary in nature and should be sufficient to perform the motion detection task in a fast and efficient manner.
-Here we first describe the task definition and then the HD-SNN model we use for motion inference and how we train it.
+Here, we first describe the task definition, then the HD-SNN model we use for motion inference, and finally how we train it.
 %%%-----------------------------------------------------------------
@@
 In animals with a fovea, this is particularly useful as it allows the area with the highest density of photoreceptors to be moved, for example, to a point of interest in the environment.
-A specific mechanism for this function is saccades, which are rapid eye movements that reposition the center of vision.
+A specific mechanism for this function are saccades, which are rapid eye movements that reposition the center of vision.
 In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}), very fast (about $80~\ms$), and have a wide range of speeds.
-On a more microscopic scale, the human gaze moves with a continuous drift similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
+On a more microscopic scale, the human gaze moves with minute micro-saccades which trajectories are similar a Brownian trajectory~\citep{poletti_head-eye_2015}.
 To preserve the full generality of the task, we define eye movements using such a form of random walk~\citep{engbert_integrated_2011}: We first define a finite set of possible motions in polar coordinates as the Cartesian product of the $12$ regularly spaced directions and $3$ geometrically spaced speeds.
@@
 We discretize time in $1~\ms$ bins and produce movies of duration $N_T = 200~\ms$.
-To avoid boundary effects, we randomly position the full trajectory in image space so that the sub-image is translated using the position given by the trajectory at each time step without touching the boundaries.
+To avoid boundary effects, we randomly position the full trajectory in image space so that the sub-image is translated using the position given by the trajectory at each time step and without touching the boundaries.
 At each time step, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a subpixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
@@
 We denote this list as $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N}$ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{Left-Top} for an illustration).
+Each event has a time of occurrence $\timev_\arank$ and an associated address $\presynaddr_\arank$.
 Events are usually ordered by their time of occurrence.
-Each event has a time of occurrence $\timev_\arank$ and an associated address $\presynaddr_\arank$.
 This defines an address space $\presynaddrspace$, which consists of the set of possible addresses.
@@
 Importantly, this activation function has to be such that when postsynaptic neurons are activated at their soma by a specific spatiotemporal motif imprinted in the synaptic set, and such that these spikes converge at the soma in a synchronous manner, the discharge probability should increase.
-In this subsection, we have defined the HD-SNN model in all generality, and in the next subsection we describe an implementation of our model adapted to the motion detection task.
+In this subsection, we have briefly defined the HD-SNN model in all generality (see~\citep{Perrinet23ICANN} for a more detailed description and treatment), and in the next subsection we describe an implementation of our model adapted to the motion detection task.
 %
@@
 %
-Indeed, it is possible to adapt the HD-SNN model specifically for common computer vision tasks.
-First, neural addresses are defined such as to represent the range of possible positions and polarities.
-Then, in the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and benefit from parallel computing acceleration, we transform this event-based representation into a dense representation.
+In fact, it is possible to adapt the HD-SNN model specifically for common computer vision tasks.
+First, neural addresses are defined to represent the range of possible positions and polarities.
+In order to simulate such event-based computations on standard CPU- or GPU-based computers and to benefit from parallel computing acceleration, we transform this event-based representation into a dense representation.
 Indeed, by discretizing time, we transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.

### 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
@@
 This delay between a pair of connected neurons is defined as the time between the emission of a spike at the soma of the afferent neuron and its arrival at the soma of the efferent neuron.
-Importantly, due to the variety of weights and delays within a population, representations using polychronous groups have in theory a much higher information capacity than a firing rate-based approach.
+Importantly, due to the variety of weights and delays within a population, representations using polychronous groups have, in theory, a much higher information capacity than a firing rate-based approach.
 

### 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
 In particular, for each class $\class$ of the supervision task, the entire synaptic set can be represented as a kernel, which is represented by the dense matrix $\kernel$ of size $(\Nclass, \Npol, \Ktime, \Kx, \Ky)$, where $\Nclass$ is the number of classes, $\Ktime$ the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
-To keep the analogy with the HD-SNN model, $\kernel$ gives the weights as a function of class $\class$, polarity, synaptic delay and relative position: $\forall {\polev \in [1, \Npol], \delta_\timev \in [1, \Ktime], \delta_x \in [1, \Kx], \delta_y \in [1, \Ky]}, \; \kernel(\class, \polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$.
+To keep the analogy with the HD-SNN model, $\kernel$ gives the weights as a function of class $\class$, polarity, synaptic delay and relative position: $\forall {\polev \in [1, \Npol], \delta_\timev \in [1, \Ktime], \delta_x \in [1, \Kx], \delta_y \in [1, \Ky]}, \; \kernel(\class, \polev, \synapticdelay_\timev, \delta_x, \delta_y) = \synapticweight$.
 In our simulations, we define as many classes as the number of motions (directions and velocities): $\Nclass = 12 \times 3$ and set the size of the model's kernel to $(\Nclass, \Npol, \Ktime, \Kx, \Ky) = (36, 2, 21, 17, 17)$ (that is, $436968$ parameters).
@@
 \forall x, y, \timev, \; \;
-B(\class, x, y, \timev)
-= \sum_{p, \delta_x, \delta_y, \synapticdelay_\timev} \kernel(\class, p, \delta_x, \delta_y, \synapticdelay_\timev) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay_\timev)
+B(\class, \timev, x, y)
+= \sum_{p, \synapticdelay_\timev, \delta_x, \delta_y} \kernel(\class, p, \synapticdelay_\timev, \delta_x, \delta_y) \cdot A(p, \timev-\synapticdelay_\timev, x - \delta_x, y - \delta_y)
 \end{equation}
@@
 Here, we further assume that synaptic motifs should be similar across different times as defined in the temporal convolution.
-As a consequence, this defines a 3D, spatiotemporal convolutional operator, in which layer of neurons assigned to specific kernels form channels.
+As a consequence, this defines a 3D, spatiotemporal convolutional operator, in which the layers of neurons assigned to specific kernels form channels.
 Using this dense representation, the model's processing of the input $A$ can be written as layer-wise convolution similar to that used in CNNs: $B = \kernel \ast A$ (\seeFig{model} for an illustration).
-A similar type of spatiotemporal filtering is used as a preprocessing stage for an existing pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
+A similar type of spatiotemporal filtering was used as a preprocessing stage for an existing pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
 In addition,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm that implements a motion estimation task.
@@
 For convenience, the connectivity of the neuron $\postsynaddr$ is defined locally around its position $(x_\postsynaddr, y_\postsynaddr)$.
-Note also that to stay within the framework of a causal computation, the kernels are shifted in time so that only past information gives an answer at the present time that is, after a time equal to the depth $\Ktime$ of the kernel.
+Note also that to stay within the framework of a causal computation, the kernels are shifted in time so that only past information provides an inference at the present time that is, after a time equal to the depth $\Ktime$ of the kernel.
 This well-known computation defines a differentiable measure, which is very efficiently implemented for GPUs, and which we will use to learn the classification of different motifs in the event stream.
 % Therefore, assuming that a neuron integrates only in the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
+Such a method contrasts with classical methods for delay learning which explictly manipulate the delay as a variable and which are not directly diferentiable~\citep{nadafian_bio-plausible_2020}.
 
@@
 Since we know that at each instant, there is one and only one motion, we will define the activation function of our model as a softmax function that implements a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, and analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}.
-In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a softmax function of the linear combination of the list of events.
+In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a sigmoid function $\sigma(\beta) = \frac {1} {1 + \exp(-\beta)}$ of the linear combination of the list of events.
 Formally, using the kernels, the input raster plot is transformed into a probability with the following formula:
@@
 % TODO : change to logistic regression
-\begin{linenomath*}
+% \begin{linenomath*}
 \begin{equation}\label{eq:mlr}
@@
 Pr(k=\class \; \vert \; x, y, \timev) =
-\frac {\exp (B_\class(x, y, \timev) +\bias_\class) }{Z(x, y, \timev)}
+\sigma (B(\class, \timev, x, y) +\bias_\class) 
 \end{equation} 
-\end{linenomath*}
+% \end{linenomath*}
 %
-where $Z(x, y, \timev)=\sum_{\class \in [1, \Nclass]} \exp (B_\class(x, y, \timev) +\bias_\class)$ is the partition function, and $\bias_\class$ is a scalar representing the bias associated with the class $\class$.
+where $\bias_\class$ is a scalar representing the bias associated with the class $\class$. 
 In particular, we expect that some specific motifs may generate tightly synchronized outputs as they are integrated in the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike.

### 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
@@
 \textit{(Left)}~We show an instance of this trajectory (with a length of $200~\ms$, green line) superimposed on the luminance contrasts observed at time step $20~\ms$.
-\textit{( Right)}~The dynamics of this image translated according to the saccadic trajectory produces a naturalistic movie, which is then transformed into an event-based representation.
+\textit{( Right)}~The dynamics of this image, translated according to the saccadic trajectory, produces a naturalistic movie, which is then transformed into an event-based representation.
 We show snapshots of the resulting synthetic event stream at different time steps (start and end frames are marked by a white and black dot, respectively, in the left inset).

### 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
methods

--- 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
@@
 \newcommand{\loss}{J} % cost function for MLR
-\newcommand{\softmax}{\sigma}
 \newcommand{\actfreq}{f}
@@
 It represents the evidence accumulated from a set of sources in logistic regression.
-Since we know that at each instant, there is one and only one motion, we will define the activation function of our model as a softmax function that implements a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, and analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}.
+Since we know that at each instant, there is one and only one motion, we will define the activation function of our model as a sigmoid function that implements a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, and analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}.
 In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a sigmoid function $\sigma(\beta) = \frac {1} {1 + \exp(-\beta)}$ of the linear combination of the list of events.
@@
 Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration).
-This is consistent with neuroscientific principles because, due to the limits of conduction delay, the synaptic delay is physically related to the distance between the presynaptic and postsynaptic neurons.
+%This is consistent with neuroscientific principles because, due to the limits of conduction delay, the synaptic delay is physically related to the distance between the presynaptic and postsynaptic neurons.
 Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities.
@@
 Since we have two axes $X$ and $Y$ to represent the presynaptic addresses, like the pixel grid of a DVS, this results in a 3D convolution.
-Here we simplify the illustration to a 2D representation and to 2 possible classes (green and orange) associated with different directions of motion.
+Here we simplify the illustration to a 2D representation and to 2 possible classes (green and orange) associated to two different directions of motion.
 \textit{(Middle)}~For each position (address, time) one can compute the activation resulting from the convolution.
-The output of the convolution is processed by the nonlinearity of the MLR model (i.e., the softmax function).
+The output of the convolution is processed by the nonlinearity of the MLR model (i.e., the sigmoid function).
 The output of the MLR gives a probability for each class associated with a particular kernel (colored bars in the highlighted pixel).
@@
 The loss function of the MLR model is the binary cross entropy of the output of the classification layer knowing the ground truth.
-The labels were defined at each time point as a one-hot encoding of the current motion in the channel corresponding to the current motion for all positions.
-Note that in this context, the label is known, but the position is not, mainly due to the sparse spatial content of natural images.
+The labels were defined at each time point as a one-hot encoding of the current motion in the channel corresponding to the current motion, and applied for all positions.
+Note that in this context, the label is known, but the position of visual features is not, mainly due to the sparse spatial content of natural images.
 However, the supervised optimization of this MLR model adjusts the weights of the kernels.
@@
 This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation.
-Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{10}$ movies, each of size $200 \times 128 \times 128$, a learning rate of $10^{-5}$ and $25$ epochs).
+Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{10}$ movies, each of size $200 \times 128 \times 128$, a learning rate of $10^{-5}$ and $200$ epochs).
 We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
@@
     \caption{
-    	Representation of the weights for $XX$ directions for a single speed (among the $12 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
+    	Representation of the weights for $8$ directions for a single speed (among the $12 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
 The directions are shown as red arrows in the left insets.
@@
 Each image corresponds to the weights at a given delay, with excitatory and inhibitory weights in warm and cold colors, respectively.
-Due to the symmetry between the ON and OFF event streams, we report that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective for the different motion directions, and we observe a slight orientation preference perpendicular to direction for all kernels.
+Due to the symmetry between the ON and OFF event streams, we observed that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective for the different motion directions, and we observe a slight orientation preference perpendicular to the respective direction for all kernels.
 	}
@@
 Finally, we observe that these receptive fields show also a relative selectivity to the orientation perpendicular to motion, similarly to what is found for neurons in cortical area MT which is known to be selective to visual motions~\citep{deangelis_functional_1999}.
-This reflects the way events are generated and in particular the so-called aperture problem which implies that an line moving along its axis would generate no change in luminance and therefore generate no event~\citep{perrinet_motion-based_2012}.
+This reflects the way events are generated and in particular the so-called aperture problem which implies that a line moving along its axis would generate no change in luminance and therefore generate no event~\citep{perrinet_motion-based_2012}.
 

### 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
results

@@
 This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation.
-Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{10}$ movies, each of size $200 \times 128 \times 128$, a learning rate of $10^{-5}$ and $200$ epochs).
+Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{10}$ movies, each of size $200 \times 128 \times 128$, a learning rate of $10^{-5}$ and $100$ epochs).
 We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
@@
 After training our MLR model, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection.
-We observed that the distribution of the kernel weights is sparse, with most values close to zero (\seeFig{kernels}).
+The accuracy computed on the test set was approximately $85\%$ (with a chance level of $1 / 12/ 3 \approx 2\%$).
+Moreover, we observed that the classification performance obtained for the model using time surfaces as in~\citep{grimaldi_robust_2022} gives similar results as our method.
+
+We observed that the distribution of the kernel's weights is sparse, with most values close to zero (\seeFig{kernels}).
 As shown in the formalization of our event-based model, the computational cost of our model, if implemented on a neuromorphic chip, would be dominated by the computations used for the convolution operation.
@@
 
-To assess the robustness of the classification as a function of computational load, we prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold.
+To assess the robustness of the classification as a function of computational load, we prune the weights in $\kernel$ that are below a defined threshold.
 In Fig.~\ref{fig:accuracy} we plot the classification accuracy as a function of the relative number of computations or active weights per decision for each neuron in the layer.
-For comparison, and to take into account the performance gain from using heterogeneous delays, we show the accuracy obtained with an MLR model using a 2D time surface (in red) as in~\citep{grimaldi_robust_2022}.
-The latter method is based on the delays of the last recorded events and requires fewer computations (in our case $17\times17$) than the dense 3D kernels without any pruning ($21\times17\times17$).
-While fewer computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
+For comparison, and to take into account the performance gain from using heterogeneous delays, we show the accuracy obtained with an MLR model using a shortening strategy in which temporal width is varried by selecting only the weights corresponding to the shortest delays (in orange).
+Compared to the inference using the dense 3D kernels without any pruning ($21\times17\times17$), both methods requires fewer computations as measured by nonzero weights.
 
-By pruning the weights, we observe that the evolution of accuracy as a function of the log percentage of active weights fits well to a sigmoid curve.
-Half saturation is reached at a ratio of about $3.2\ times 10^{-2}$ of active weights, which in our setting corresponds to a total of about $180$ computations per decision.
+By pruning the weights, we observe that the evolution of accuracy as a function of the (log-)percentage of active weights fits well to a sigmoid curve for both strategies (pruning and shortening).
+While the shortening strategy yields a very rapid loss of accuracy which reaches hal-saturation when half of the weights are removed, the situation is different for pruning.
+Indeed, in that case, half saturation is reached at a ratio of about $3.2\times 10^{-2}$ of active weights, which in our setting corresponds to a total of about $180$ computations per decision.
 Compared to the full kernels, the accuracy of our method is maintained relative to its peak performance when the number of computations is divided by a factor up to about $31$.

### 2023-07-04 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
 After training our model, let us first examine the emerging weights learned for the different motions (\seeFig{kernels}).
-We first notice by following each spatial pattern from the shortest delay (right) to the longest (left) that the selectivity of these cells are very localized and are transported along linear trajectories in the space-delay space.
+We first notice by following each spatial motif from the shortest delay (right) to the longest (left) that the selectivity of these cells are very localized and are transported along linear trajectories in the space-delay space.
 Focusing on the positive weights, a strong selectivity along specific motion axes is observed for each of the different kernels and these directions can be easily associated with the direction of the corresponding motion.
@@
 The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
-As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about 25 times fewer computations.
+As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about $250$ times fewer computations.
         }
@@
 %%%-----------------------------------------------------------------
-In this paper, we have introduced a generic SNN with heterogeneous delays and shown how it compares favorably with state-of-the-art event-based classification algorithms for a visual motion detection task.
-The learned model has a number of similarities with neurobiological anatomical observations as well as behavioral results.
-The event-based computations of our method can be drastically reduced by pruning synapses, while maintaining top classification performance.
-This shows that we can use the precise timing of spikes to improve the performance of neural computations.
+In this paper, we have introduced a generic heterogeneous delay SNN.
+It is trained using supervised learning on a visual motion detection task.
+We have shown how the model performs well compared to other event-based classification algorithms for this task.
+The learned model has a number of similarities with neurobiological and behavioral observations.
+The event-based computations of our method can be drastically reduced by pruning synapses, while maintaining robust classification performance.
+This shows that we can exploit the precise timing of spikes to improve the performance of neural computations.
 %
@@
 We would like to highlight a few innovations in the contributions presented in this paper.
-First, while~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of spatiotemporal spiking motifs.
+First, while~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of spatiotemporal spiking motifs using a time-invariant logistic regression.
 Moreover, compared to classical CNN solutions, the parameters of the model (weights and delays) are explicable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif, where we define \emph{evidence} as the logit of the probability, that is, the inverse sigmoid of the probability.
@@
 We have tested the effect of the width of the time bin and observed it has essentially no effect on the results presented in this paper.
-This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}.
-This suggests the potential advantage of analytically including an additional precision term to the temporal value of input spikes.
-This mechanism may be implemented by the filtering implemented by the synaptic time constant of about $5~\ms$.
+This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017,grimaldi_robust_2022}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}.
+It suggests the potential advantage of analytically including an additional precision term to the temporal value of input spikes.
+Such a mechanism may be implemented by the filtering implemented by the synaptic time constant of about $5~\ms$.
 Furthermore, it is possible to circumvent the need for time discretization by the use of a purely event-based scheme.
@@
 Exploring this dependence will allow us to dissociate and integrate these two forms of visual information.
-In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka the aperture problem).
+In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (that is, the aperture problem).
 Thus, it will allow us to implement recurrent prediction rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
@@
 There are several sources of information for depth estimation, such as binocular disparity or changes in texture or shading, but in our case motion parallax would be the most important cue~\citep{rogers_motion_1979}.
-This is because objects that are close to an observer move relatively faster on the retina than an object that is far away, and also because visual occlusions depend on depth order.
+This is because objects that are close to an observer move on the retina relatively faster than an object that is far away, and also because visual occlusions depend on depth order.
 Using this information, it is possible to segment objects and estimate their depth~\citep{yoonessi_contribution_2011}.
@@
 In conclusion, the HD-SNN model that we have presented provides a way to efficiently process event-based signals.
-We have shown that we can train the model using a supervised rule that knows \emph{what} output label, but not \emph{where} it occurs.
+We have shown that we can train the model using a supervised rule that infers \emph{what} is the output label, but not \emph{where} it occurs.
 Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data~\citep{barlow_unsupervised_1989}.
@@
 Indeed, there is a large literature showing that brain dynamics often organize into stereotyped sequences, such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}.
-These patterns are stereotyped and robust, as they can be activated in the same pattern from day to day~\citep{haimerl_internal_2019}.
+These motifs are stereotyped and robust, as they can be activated following the same motif from day to day~\citep{haimerl_internal_2019}.
 In contrast to conventional methods of processing neurobiological data, such an event-based model would be able to answer key questions about the representation of information in neurobiological data, and it would open possibilities in the field of computational neuroscience.

### 2023-07-05 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 \textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses).
-\textit{(Left-Middle)}~Each spike is weighted by the synaptic weights (height of the bar) and shifted in time according to the synaptic delays on each respective synapse (input spikes are shown in light gray for comparison).
+\textit{(Left-Middle)}~Each spike is weighted by the synaptic weights (height of the blue bars) and shifted in time according to the synaptic delays on each respective synapse (input spikes are shown in light gray for comparison).
 As a result, the spikes from the spiking motif are synchronized as they reach the soma of the postsynaptic neuron.
@@
 This defines an address space $\presynaddrspace$, which consists of the set of possible addresses.
-In neurobiological spiking activity, this may be the identified set of all recorded neurons.
+In neurobiological spiking activity, this may be the identified set of recorded neurons.
 For neuromorphic hardware like the output of a DVS or our task, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$, where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities encoded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels.

### 2023-07-05 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 Indeed, by discretizing time, we transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.
-The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
+The values are, by definition, equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
 Similarly, one may discretize the connectivity of the HD-SNN model defined above.
@@
 where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses within a kernel and $\delta_\timev$ is the synaptic delay.
-In this formulation, we recognize that it takes advantage of the position invariance observed in images and exploited in convolutional neural networks.
+In this formulation, we recognize that it takes advantage of the position invariance observed in images and exploited in CNNs.
 Here, we further assume that synaptic motifs should be similar across different times as defined in the temporal convolution.
@@
 % Therefore, assuming that a neuron integrates only in the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
-Such a method contrasts with classical methods for delay learning which explictly manipulate the delay as a variable and which are not directly diferentiable~\citep{nadafian_bio-plausible_2020}.
+Such a method contrasts with classical methods for delay learning, which explicitly manipulate the delay as a variable and which are not directly differentiable~\citep{nadafian_bio-plausible_2020}.
 
@@
 It represents the evidence accumulated from a set of sources in logistic regression.
-Since we know that at each instant, there is one and only one motion, we will define the activation function of our model as a sigmoid function that implements a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, and analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}.
+Since we know that at each instant, there is one and only one motion, we will define the activation function of our model as a sigmoid function that implements a form of Multinomial Logistic Regression (MLR), analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}.
 In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a sigmoid function $\sigma(\beta) = \frac {1} {1 + \exp(-\beta)}$ of the linear combination of the list of events.
@@
 Now that this general framework has been described formally, we may include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}.
-Note that the general framework is an extension to that presented in~\cite{grimaldi_learning_2022}, in particular by including the more complex task to solve, the deeper analysis of the results, and these novel neuroscience-inspired heuristics.
+Note that the general framework is an extension to that presented in~\cite{grimaldi_learning_2022}, in particular by including a more complex task to solve, the deeper analysis of the results, and these novel neuroscience-inspired heuristics.
 First, to avoid introducing biases in the directions which may be learned, we apply a circular mask to the spatial dimensions of the kernels.

### 2023-07-05 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 
-If one now widen our focus on the interpretation of these kernels in terms of spatiotemporal motifs embedded in the event stream, these show a prototypical anisotropic profile~\citep{kaplan_anisotropic_2013} adapted to motion detection.
+If we now widen our focus on the interpretation of these kernels in terms of spatiotemporal motifs embedded in the event stream, these show a prototypical anisotropic profile~\citep{kaplan_anisotropic_2013} adapted to motion detection.
 In~\citep{grimaldi_robust_2022}, a link was drawn between event-based MLR training and Hebbian learning, allowing to say that the present model learns its weights according to a presynaptic activity associated to the different motion directions.
@@
 Note that as the delays become larger, two effects can be remarked.
-First, coefficients become lower which is consistent with the fact that trajectories are defined in a piece wise fashion, such that such decrease provides with an optimal integration with respect to this limited memory~\citep{Pasturel2020}.
+First, coefficients become lower which is consistent with the fact that trajectories are defined in a piece wise fashion, such that this decrease provides with an optimal integration with respect to this limited memory~\citep{Pasturel2020}.
 Second, coefficients become less localized compared to the kernel's spatial profile at short delays, consistent with the average diffusion of information included in the generative model and with the diffusion introduced in motion-based prediction models~\citep{perrinet_motion-based_2012,khoei_flash-lag_2017}.
@@
 %%%-----------------------------------------------------------------
-To test our model, we will quantify its ability to categorize different motions, i.e., on event streams for which the ground truth motion is known at each instant.
+To test our model, we will quantify its ability to categorize different motions, i.e.
+on event streams for which the ground truth motion is known at each instant.
 Before applying the model to natural scenes that we defined in the motion detection task, we first test the model on simpler, parameterized stimuli.

### 2023-07-05 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 To assess the robustness of the classification as a function of computational load, we prune the weights in $\kernel$ that are below a defined threshold.
-In Fig.~\ref{fig:accuracy} we plot the classification accuracy as a function of the relative number of computations or active weights per decision for each neuron in the layer.
-For comparison, and to take into account the performance gain from using heterogeneous delays, we show the accuracy obtained with an MLR model using a shortening strategy in which temporal width is varried by selecting only the weights corresponding to the shortest delays (in orange).
-Compared to the inference using the dense 3D kernels without any pruning ($21\times17\times17$), both methods requires fewer computations as measured by nonzero weights.
+In Fig.~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations or active weights per decision for each neuron in the layer.
+For comparison, and to take into account the performance gain from using heterogeneous delays, we show the accuracy obtained with an MLR model using a shortening strategy in which temporal width is varied by selecting only the weights corresponding to the shortest delays (in orange).
+Compared to the inference using the dense 3D kernels without any pruning ($21\times17\times17$), both methods require fewer computations as measured by non-zero weights.
 
@@
 The HD-SNN model was trained and evaluated on a naturalistic motion detection task with realistic eye movements.
-The model was defined such as to provide an optimal detection of  spatiotemporal motifs and learns kernels similar to those found in the visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
+The model is defined such as to provide an optimal detection of  spatiotemporal motifs and learns kernels similar to those found in the visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
 We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware.
@@
 First, while~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of spatiotemporal spiking motifs using a time-invariant logistic regression.
-Moreover, compared to classical CNN solutions, the parameters of the model (weights and delays) are explicable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif, where we define \emph{evidence} as the logit of the probability, that is, the inverse sigmoid of the probability.
+Moreover, compared to classical CNN solutions, the parameters of the model (weights and delays) are explainable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif, where we define \emph{evidence} as the logit of the probability, that is, the inverse sigmoid of the probability.
 Another novelty is that the model learns simultaneously weights and delays.
@@
 We have shown that we can train the model using a supervised rule that infers \emph{what} is the output label, but not \emph{where} it occurs.
-Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data~\citep{barlow_unsupervised_1989}.
+Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e.
+without any labelled data~\citep{barlow_unsupervised_1989}.
 This type of learning is thought to be prevalent in the central nervous system and, assuming the signal is sparse~\citep{olshausen_emergence_1996}, one could extend these Hebbian sparse learning schemes to spikes~\citep{perrinet_emergence_2004, masquelier_competitive_2009}.

### 2023-07-05 Grimaldi-etal-BiolCybernetics.tex
figures

--- 2023-07-05 Grimaldi-etal-BiolCybernetics.tex
@@
 \textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses).
-\textit{(Left-Middle)}~Each spike is weighted by the synaptic weights (height of the blue bars) and shifted in time according to the synaptic delays on each respective synapse (input spikes are shown in light gray for comparison).
+\textit{(Left-Middle)}~Each spike is weighted by the synaptic weights (height of the bar) and shifted in time according to the synaptic delays on each respective synapse (input spikes are shown in light gray for comparison).
 As a result, the spikes from the spiking motif are synchronized as they reach the soma of the postsynaptic neuron.
@@
 This defines an address space $\presynaddrspace$, which consists of the set of possible addresses.
-In neurobiological spiking activity, this may be the identified set of recorded neurons.
+In neurobiological spiking activity, this may be the identified set of all recorded neurons.
 For neuromorphic hardware like the output of a DVS or our task, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$, where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities encoded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels.
@@
 Indeed, by discretizing time, we transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.
-The values are, by definition, equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
+The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
 Similarly, one may discretize the connectivity of the HD-SNN model defined above.
@@
 where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses within a kernel and $\delta_\timev$ is the synaptic delay.
-In this formulation, we recognize that it takes advantage of the position invariance observed in images and exploited in CNNs.
+In this formulation, we recognize that it takes advantage of the position invariance observed in images and exploited in convolutional neural networks.
 Here, we further assume that synaptic motifs should be similar across different times as defined in the temporal convolution.
@@
 % Therefore, assuming that a neuron integrates only in the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
-Such a method contrasts with classical methods for delay learning, which explicitly manipulate the delay as a variable and which are not directly differentiable~\citep{nadafian_bio-plausible_2020}.
+Such a method contrasts with classical methods for delay learning which explictly manipulate the delay as a variable and which are not directly diferentiable~\citep{nadafian_bio-plausible_2020}.
 
@@
 It represents the evidence accumulated from a set of sources in logistic regression.
-Since we know that at each instant, there is one and only one motion, we will define the activation function of our model as a sigmoid function that implements a form of Multinomial Logistic Regression (MLR), analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}.
+Since we know that at each instant, there is one and only one motion, we will define the activation function of our model as a sigmoid function that implements a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, and analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}.
 In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a sigmoid function $\sigma(\beta) = \frac {1} {1 + \exp(-\beta)}$ of the linear combination of the list of events.
@@
 Now that this general framework has been described formally, we may include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}.
-Note that the general framework is an extension to that presented in~\cite{grimaldi_learning_2022}, in particular by including a more complex task to solve, the deeper analysis of the results, and these novel neuroscience-inspired heuristics.
+Note that the general framework is an extension to that presented in~\cite{grimaldi_learning_2022}, in particular by including the more complex task to solve, the deeper analysis of the results, and these novel neuroscience-inspired heuristics.
 First, to avoid introducing biases in the directions which may be learned, we apply a circular mask to the spatial dimensions of the kernels.

### 2023-07-06 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2023-07-05 Grimaldi-etal-BiolCybernetics.tex
@@
 To characterize the robustness of the model in detecting visual motion, we test its efficiency by pruning some weights and show that it is still efficient at very low workloads.
-In conclusion, by providing an end-to-end, event-driven computational building block, the learning of heterogeneous delays could improve the performance of future spiking neural network (SNN) algorithms, and in particular their applications in neuromorphic chips.
+In conclusion, by providing an end-to-end, event-driven computational building block, the learning of heterogeneous delays could improve the performance of future spiking neural network algorithms, and in particular their applications in neuromorphic chips.
 }

### 2023-07-06 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %
-The human brain has the remarkable capacity to react efficiently at any given time while consuming a reasonable amount of energy, in the order of 20 watts.
+The human brain has the remarkable capacity to react efficiently at any given time while consuming a reasonable amount of energy, in the order of $20$ watts.
 This system is the result of millions of years of natural selection, and a striking difference with artificial neural networks is the representation that both use.
-For instance, the convolutional neural networks (CNNs) which are used for processing images represent the flow of information from one layer to another using tensors of floats, storing visual information densely across the visual topography, with different translation-invariant properties represented in different channels.
+Indeed, our computers use digital values and for instance, the convolutional neural networks (CNNs) which are used for processing images represent the flow of information from one layer to another using tensors of floats.
+These networks store visual information densely across the visual topography, with different translation-invariant properties represented in different channels.
 CNNs have achieved state-of-the-art performance for some computer vision tasks, such as image recognition.
 These networks are also known to mimic several properties of biological systems, such that each can be assigned a ``brain score''~\citep{schrimpf_brain-score_2020}.
-However, this score does not take into account key aspects of the efficiency of biological systems, such as inference speed or energy consumption, which is about hundreds of watts on a standard GPU.
+However, this score does not take into account key aspects of the efficiency of biological systems, such as inference speed or energy consumption, which is about $360$ watts on a standard GPU (NVIDIA RTX 3090).
 In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing (binary) events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
@@
 Once these eye trajectories are generated, we can apply them to a visual scene.
-For this purpose, we selected a database of $600$ natural images that were previously used to study the statistics of natural images~\citep{vanHateren1998}.
+For this purpose, we selected a database of $100$ large-scale natural images that were previously used to study the statistics of natural images~\citep{vanHateren1998}.
 Note that these were pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band, similarly to a process known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
@@
 To avoid boundary effects, we randomly position the full trajectory in image space so that the sub-image is translated using the position given by the trajectory at each time step and without touching the boundaries.
-At each time step, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a subpixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
+At each time step, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
 Note that the magnitude of the displacement is relative to the time bin, and we have defined the unit speed to correspond to a movement of one pixel per frame (i.e., per time bin of $1~\ms$).

### 2023-07-06 Grimaldi-etal-BiolCybernetics.tex

@@
 The signed threshold is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, i.e.
+When applied to the whole movie, the event stream is similar to the output of a neuromorphic camera~\citep{Gallego2022}, i.e.
 a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF), and their time $\timev_\arank$ (\seeFig{motion_task}-\textit{(Right)}).
@@
 Events are usually ordered by their time of occurrence.
-This defines an address space $\presynaddrspace$, which consists of the set of possible addresses.
+We define the address space $\presynaddrspace$, which consists of the set of possible addresses.
 In neurobiological spiking activity, this may be the identified set of recorded neurons.
@@
 Importantly, this activation function has to be such that when postsynaptic neurons are activated at their soma by a specific spatiotemporal motif imprinted in the synaptic set, and such that these spikes converge at the soma in a synchronous manner, the discharge probability should increase.
-In this subsection, we have briefly defined the HD-SNN model in all generality (see~\citep{Perrinet23ICANN} for a more detailed description and treatment), and in the next subsection we describe an implementation of our model adapted to the motion detection task.
+In this subsection, we have briefly defined the HD-SNN model in all generality (see~\citep{Perrinet23ICANN} for a more specific description and treatment), and in the next subsection we describe an implementation of our model adapted to the motion detection task.
 %
@@
 In particular, for each class $\class$ of the supervision task, the entire synaptic set can be represented as a kernel, which is represented by the dense matrix $\kernel$ of size $(\Nclass, \Npol, \Ktime, \Kx, \Ky)$, where $\Nclass$ is the number of classes, $\Ktime$ the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
-To keep the analogy with the HD-SNN model, $\kernel$ gives the weights as a function of class $\class$, polarity, synaptic delay and relative position: $\forall {\polev \in [1, \Npol], \delta_\timev \in [1, \Ktime], \delta_x \in [1, \Kx], \delta_y \in [1, \Ky]}, \; \kernel(\class, \polev, \synapticdelay_\timev, \delta_x, \delta_y) = \synapticweight$.
-In our simulations, we define as many classes as the number of motions (directions and velocities): $\Nclass = 12 \times 3$ and set the size of the model's kernel to $(\Nclass, \Npol, \Ktime, \Kx, \Ky) = (36, 2, 21, 17, 17)$ (that is, $436968$ parameters).
+To keep the analogy with the HD-SNN model, $\kernel$ gives the synaptic set that defines the weights as a function of class $\class$, polarity, synaptic delay and relative position: $\forall {\polev \in [1, \Npol], \delta_\timev \in [1, \Ktime], \delta_x \in [1, \Kx], \delta_y \in [1, \Ky]}, \; \kernel(\class, \polev, \synapticdelay_\timev, \delta_x, \delta_y) = \synapticweight$.
+In our simulations, we define as many classes as the number of motions (directions and velocities): $\Nclass = 12 \times 3$ and set the size of the model's kernel to $(\Nclass, \Npol, \Ktime, \Kx, \Ky) = (36, 2, 21, 17, 17)$.
 Such a kernel defines a dense representation of the full synaptic set.
+%  (that is, $436968$ parameters)
 
@@
 As a consequence, this defines a 3D, spatiotemporal convolutional operator, in which the layers of neurons assigned to specific kernels form channels.
-Using this dense representation, the model's processing of the input $A$ can be written as layer-wise convolution similar to that used in CNNs: $B = \kernel \ast A$ (\seeFig{model} for an illustration).
+Using this dense representation, the model's processing of the input $A$ can be written as layer-wise convolution: $B = \kernel \ast A$ (\seeFig{model} for an illustration).
 A similar type of spatiotemporal filtering was used as a preprocessing stage for an existing pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
@@
 For convenience, the connectivity of the neuron $\postsynaddr$ is defined locally around its position $(x_\postsynaddr, y_\postsynaddr)$.
-Note also that to stay within the framework of a causal computation, the kernels are shifted in time so that only past information provides an inference at the present time that is, after a time equal to the depth $\Ktime$ of the kernel.
-This well-known computation defines a differentiable measure, which is very efficiently implemented for GPUs, and which we will use to learn the classification of different motifs in the event stream.
+Furthermore, it is important to consider that in order to adhere to the limitations of causal computation using biologically realistic neurons, synaptic delays are assigned positive values.
+This ensures that only past information contributes to the inference made at the present moment.
+In practical implementation, the kernels are temporally shifted so that the inference at the present time is solely influenced by past information.
+This temporal shift occurs after a duration equivalent to the depth of the kernel, denoted as $\Ktime$.
+The well-known convolution defines a differentiable measure, which is very efficiently implemented for GPUs, and which we will use to detect the motion direction in the event stream.
 % Therefore, assuming that a neuron integrates only in the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
@@
 
-Keeping the analogy with spiking neurons, the analog activity $B$ represents the integration of synaptic activity, and we will now try to define the detection of spiking motifs using the spatiotemporal kernels.
-The 3D convolution represents the linear integration of the spike train as the linear input to the neuron.
-It represents the evidence accumulated from a set of sources in logistic regression.
-Since we know that at each instant, there is one and only one motion, we will define the activation function of our model as a sigmoid function that implements a form of Multinomial Logistic Regression (MLR), analogous to a spiking winner-takes-all network~\citep{nessler_bayesian_2013}.
+Keeping the analogy with spiking neurons, the analog activity $B$ represents the integration of synaptic activity, and we will now try to define the detection of motion using the spatiotemporal kernels.
+Since we know that at each instant, there may be different motions, we will define the activation function of our model as a sigmoid function that implements a form of Multinomial Logistic Regression (MLR).
 In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a sigmoid function $\sigma(\beta) = \frac {1} {1 + \exp(-\beta)}$ of the linear combination of the list of events.
@@
 %
-% TODO : change to logistic regression
-% \begin{linenomath*}
 \begin{equation}\label{eq:mlr}
@@
 \end{equation} 
-% \end{linenomath*}
 %
 where $\bias_\class$ is a scalar representing the bias associated with the class $\class$.
-In particular, we expect that some specific motifs may generate tightly synchronized outputs as they are integrated in the basal dendritic tree, leading to high postsynaptic activity, making it progressively more likely to generate an output spike.
-The spiking output of the model corresponds for instance to the event with the highest probability.
+In particular, we anticipate that certain specific patterns could result in closely synchronized outputs when they are integrated within the basal dendritic tree, consequently leading to heightened postsynaptic activity.
+By utilizing this analog representation of the evidence for each potential motion at every moment, we can progressively increase the likelihood of generating an output spike.
+To determine the spiking output, we establish a firing threshold.
+We computed this threshold to ensure that neurons, on average, generate one spike per second.
+Therefore, the spiking output of the model corresponds to the motions in space and time that possess the highest probability.
 
@@
 The spatiotemporal kernels are shown as slices of spatial weights at different delays.
-Delays vary along the horizontal axis from the far right (delay of one step) to the left (delay of $12$ steps).
+Delays vary along the horizontal axis from the far right (delay of one step) to the left (up to a delay of $12$ steps, the remaining synapses being not represented).
 Each image corresponds to the weights at a given delay, with excitatory and inhibitory weights in warm and cold colors, respectively.
@@
 %
-After training our model, let us first examine the emerging weights learned for the different motions (\seeFig{kernels}).
-We first notice by following each spatial motif from the shortest delay (right) to the longest (left) that the selectivity of these cells are very localized and are transported along linear trajectories in the space-delay space.
-Focusing on the positive weights, a strong selectivity along specific motion axes is observed for each of the different kernels and these directions can be easily associated with the direction of the corresponding motion.
-The first kernel, for example, shows a strong selectivity for a downward direction of motion.
-The opposite weights surround these positive weights, providing an enhanced response.
+Once our model has been trained, we can begin by examining the learned weights for the various motions (\seeFig{kernels}).
+Notably, when we track each spatial motif from the shortest delay (on the right) to the longest delay (on the left), we observe that the cells exhibit highly localized selectivity and their preferences are conveyed along linear trajectories in the space-delay domain.
+When focusing on the positive weights, we notice a pronounced selectivity along specific motion axes for each kernel, and these directions correspond closely to the associated motion's physical direction in visual space.
+For instance, the first kernel demonstrates a robust preference for downward motion.
+The negative weights are symmetrically arranged around these positive weights, forming a center-surround profile that enhances the response.
 We also observed a strong dependence between the weights reaching the ON polarities and those reaching the OFF polarities.
@@
 Note that as the delays become larger, two effects can be remarked.
-First, coefficients become lower which is consistent with the fact that trajectories are defined in a piece wise fashion, such that this decrease provides with an optimal integration with respect to this limited memory~\citep{Pasturel2020}.
+First, coefficients become lower which is consistent with the fact that trajectories are defined in a piece wise fashion, such that this decrease provides with an optimal integration considering the gradual diminishing of evidence as time progresses~\citep{Pasturel2020}.
 Second, coefficients become less localized compared to the kernel's spatial profile at short delays, consistent with the average diffusion of information included in the generative model and with the diffusion introduced in motion-based prediction models~\citep{perrinet_motion-based_2012,khoei_flash-lag_2017}.
-This conic shape along the motion's trajectory highlights a loss of precision for longer delays, i.e.
+%This conic shape along the motion's trajectory highlights a loss of precision for longer delays, i.e.
 events away in the past.
@@
 This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006,Perrinet2007} and reveals the most informative scales learned by our model.
-We observe a similar trend for different spatial frequency bandwidths.
 In particular, the accuracy is better for a grating-like stimulus than for a large one (which qualitatively resembles a more textured stimulus), reminiscent of the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}.

### 2023-07-06 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

@@
 %
+%
 %%%-----------------------------------------------------------------
@@
 %%%-----------------------------------------------------------------
-To test our model, we will quantify its ability to categorize different motions, i.e.
+We also wanted to assess the influence of spatiotemporal parameters of the stimuli on the performance of the model.
+For this, we will quantify its ability to categorize different motions, i.e.
 on event streams for which the ground truth motion is known at each instant.
 Before applying the model to natural scenes that we defined in the motion detection task, we first test the model on simpler, parameterized stimuli.
-For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control, among other parameters, the motion (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
-In particular, we set the spatial extent and duration of the generated movies similar to the motion task defined above.
+For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control, among other parameters, the motion (\seeFig{motion_clouds})~\citep{Leon2012,vacher_bayesian_2018}.
+These stimuli allowed us to control various parameters, including motion, spatial extent, and duration.
+By matching the spatial and temporal characteristics of the generated movies with those of the motion task mentioned earlier, we created a range of textures featuring different spatial properties and motions.
 This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $12$ directions and $3$ speeds.
@@
 %
+%
 \subsection{Accuracy versus efficiency trade-off}% for the motion detection task}
@@
 After training our MLR model, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection.
-The accuracy computed on the test set was approximately $85\%$ (with a chance level of $1 / 12/ 3 \approx 2\%$).
-Moreover, we observed that the classification performance obtained for the model using time surfaces as in~\citep{grimaldi_robust_2022} gives similar results as our method.
+The accuracy computed on the test set was approximately $89\%$ (with a chance level of $1 / 12/ 3 \approx 2\%$).
+%Moreover, we observed that the classification performance obtained for the model using time surfaces as in~\citep{grimaldi_robust_2022} gives similar results as our method.
 
@@
 In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel.
-If the information support is sparse, then computations can only be performed on those events.
-Also, if we set some weights of the kernels to zero, then the sum for those addresses can be dropped and computations can be performed only on those events.
+But if the information support is sparse, then computations can be now performed only on those events.
+Specifically, if we set some weights of the kernels to zero, then the additive operation in the convolution for those addresses can be dropped.
+As a consequence, computations will be performed only on those events which were multiplexed by the pruned connectivity matrix.
 Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of non-zero synaptic weights.
 
-To assess the robustness of the classification as a function of computational load, we prune the weights in $\kernel$ that are below a defined threshold.
-In Fig.~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations or active weights per decision for each neuron in the layer.
-For comparison, and to take into account the performance gain from using heterogeneous delays, we show the accuracy obtained with an MLR model using a shortening strategy in which temporal width is varied by selecting only the weights corresponding to the shortest delays (in orange).
-Compared to the inference using the dense 3D kernels without any pruning ($21\times17\times17$), both methods require fewer computations as measured by non-zero weights.
+In order to evaluate the resilience of the classification performance with respect to computational load, we adopt a pruning approach, where we remove weights in $\kernel$ that fall below a specified threshold.
+The accuracy of classification is then plotted as a function of the relative number of computations or active weights per decision for each neuron in the layer (refer to Fig.~\ref{fig:accuracy}).
+To provide a basis for comparison and to account for the benefits of utilizing variable delays, we present the accuracy achieved by an MLR model employing a shortening strategy.
+This strategy involves adjusting the temporal width by selecting only the weights associated with the shortest delays.
+In comparison to the inference performed using the complete 3D kernels without any pruning ($21\times17\times17$), both approaches demonstrate a reduction in computational requirements as indicated by the number of non-zero weights.
 
-By pruning the weights, we observe that the evolution of accuracy as a function of the (log-)percentage of active weights fits well to a sigmoid curve for both strategies (pruning and shortening).
-While the shortening strategy yields a very rapid loss of accuracy which reaches hal-saturation when half of the weights are removed, the situation is different for pruning.
-Indeed, in that case, half saturation is reached at a ratio of about $3.2\times 10^{-2}$ of active weights, which in our setting corresponds to a total of about $180$ computations per decision.
-Compared to the full kernels, the accuracy of our method is maintained relative to its peak performance when the number of computations is divided by a factor up to about $31$.
-In this case, the number of computations is greatly reduced compared to~\citep{grimaldi_robust_2022}, demonstrating the efficiency of the presented method.
+By selectively setting certain weights to zero, we observe that the accuracy's evolution, as a function of the logarithmic percentage of active weights, aligns well with a sigmoid curve for both pruning and shortening strategies.
+The shortening strategy (depicted in orange) demonstrates a rapid decline in accuracy, reaching half-saturation when approximately one-third of the weights are removed.
+On the other hand, the pruning strategy (shown in blue) exhibits a different behavior.
+It reaches half-saturation when the ratio of active weights is approximately $3.2\times 10^{-2}$, corresponding to around 180 computations per decision in our specific scenario.
+In comparison to using the complete kernels, our method maintains accuracy close to its peak performance even when the number of computations is divided by a factor of up to around 31.
+This substantial reduction in computations showcases the robustness of the presented method.
 %
@@
 First, while~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of spatiotemporal spiking motifs using a time-invariant logistic regression.
-Moreover, compared to classical CNN solutions, the parameters of the model (weights and delays) are explainable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif, where we define \emph{evidence} as the logit of the probability, that is, the inverse sigmoid of the probability.
+Moreover, compared to classical CNN solutions, the parameters of this one-layered model (weights and delays) are explainable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif, where we define \emph{evidence} as the logit of the probability, that is, the inverse sigmoid of the probability.
 Another novelty is that the model learns simultaneously weights and delays.
@@
 We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, in particular to be able to use fast, differentiable three-dimensional convolutions.
-We have tested the effect of the width of the time bin and observed it has essentially no effect on the results presented in this paper.
-This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017,grimaldi_robust_2022}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}.
+This is consistent with the relative robustness of other event-based frameworks~\citep{lagorce_hots_2017,grimaldi_robust_2022}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $4~\ms$ on the N-MNIST dataset~\citep{grimaldi_robust_2022}.
 It suggests the potential advantage of analytically including an additional precision term to the temporal value of input spikes.

### 2023-07-06 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
 To train and validate our model which uses supervised learning, we need to define a visual dataset for which we explicitly know the ground truth motion, that is, direction and speed.
-To achieve this, we define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
+To achieve this, we define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in  studies from neurobiology~\citep{vinje_sparse_2000, baudot_animation_2013} or computational neuroscience~\citep{kremkow_push-pull_2016}.
 First, we draw a trajectory inspired by biological eye movements.
@@
 %%%-----------------------------------------------------------------
-\begin{figure}%[h!]%[ht!]
+\begin{figure}%
     \centering
@@
         {\bf Motion Detection Task.} To generate realistic event-based dynamic scenes, we mimic the effect of minute saccadic eye movements on a large natural scene ($1024\times1024$) by extracting an image ($128\times128$) which center is moving dynamically according to a jagged random walk.
-\textit{(Left)}~We show an instance of this trajectory (with a length of $200~\ms$, green line) superimposed on the luminance contrasts observed at time step $20~\ms$.
-\textit{( Right)}~The dynamics of this image, translated according to the saccadic trajectory, produces a naturalistic movie, which is then transformed into an event-based representation.
-We show snapshots of the resulting synthetic event stream at different time steps (start and end frames are marked by a white and black dot, respectively, in the left inset).
+\textit{(Left)}~We show an instance of this trajectory (with a length of $200~\ms$, green line) superimposed on the luminance contrasts observed at time step $t=15~\ms$.
+\textit{(Right)}~The dynamics of this image, translated according to the saccadic trajectory, produces a naturalistic movie, which is then transformed into an event-based representation.
+We show snapshots of the resulting synthetic event stream at different time steps (from $t=15~\ms$ to $t=19~\ms$, these frames are marked on the trajectory by a white and black dot, respectively, in the left inset).
 Mimicking the response of ganglion cells in the retina, this representation encodes at each pixel all-or-none increases or decreases in luminance, i.e., ON (red) and OFF (blue) spikes.
@@
 %----------------------------%
-In this task, the input consists of a stream of \emph{spikes} or \emph{events}, a property shared in the signal obtained from an event-based camera or from a neurobiological recording of the activity of single units.
+In this task, the input consists of a stream of events or \emph{spikes}, a representation common to the signal obtained from an event-based camera or from a neurobiological recording of the activity of single units.
 Formally, this can be defined as a list of tuples, each tuple representing the neural address and timestamp.
@@
 By scanning all postsynaptic neurons $\postsynaddr$, we may thus define the full set of $\Nsyn$ synapses, as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated with a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
-This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{Right} for an illustration of the connectivity of a neuron with synaptic weights and delays).
+This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{Right} for an illustration of the connectivity of one neuron with synaptic weights and delays).
 % A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$, but also by the set of possible delays.
@@
 %
-\begin{linenomath*}
+% \begin{linenomath*}
 \begin{equation}\label{eq:stream_b}
@@
 \end{equation}
-\end{linenomath*}
+% \end{linenomath*}
 %
@@
 First, neural addresses are defined to represent the range of possible positions and polarities.
-In order to simulate such event-based computations on standard CPU- or GPU-based computers and to benefit from parallel computing acceleration, we transform this event-based representation into a dense representation.
-Indeed, by discretizing time, we transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.
+Second, to simulate such event-based computations on standard CPU- or GPU-based computers and to benefit from parallel computing acceleration, we transform the temporal event-based representation into a dense dicretized representation.
+Indeed, by using this discretization, we transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.
 The values are, by definition, equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
@@
 The longest synaptic delay defines the depth $\Ktime$ of the kernel, so that all possible delays associated with the different presynaptic addresses are represented.
-In particular, for each class $\class$ of the supervision task, the entire synaptic set can be represented as a kernel, which is represented by the dense matrix $\kernel$ of size $(\Nclass, \Npol, \Ktime, \Kx, \Ky)$, where $\Nclass$ is the number of classes, $\Ktime$ the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
-To keep the analogy with the HD-SNN model, $\kernel$ gives the synaptic set that defines the weights as a function of class $\class$, polarity, synaptic delay and relative position: $\forall {\polev \in [1, \Npol], \delta_\timev \in [1, \Ktime], \delta_x \in [1, \Kx], \delta_y \in [1, \Ky]}, \; \kernel(\class, \polev, \synapticdelay_\timev, \delta_x, \delta_y) = \synapticweight$.
+In particular, for each class $\class$ of the supervision task, the entire synaptic set can be represented as a kernel, which is represented by the dense matrix $\kernel$ of size $(\Nclass, \Npol, \Ktime, \Kx, \Ky)$, where $\Nclass$ is the number of classes, $\Ktime$ the number of delays and $\Kx$ and $\Ky$ are the number of pixels in each spatial dimension.
+To keep the analogy with the HD-SNN model, $\kernel$ gives the synaptic set that defines the weight of all synapses $\ranksyn$ defined as a function of their class $\class$, polarity, synaptic delay and relative position: $\forall {\polev \in [1, \Npol], \delta_\timev \in [1, \Ktime], \delta_x \in [1, \Kx], \delta_y \in [1, \Ky]}, \; \kernel(\class, \polev, \synapticdelay_\timev, \delta_x, \delta_y) = \synapticweight_\ranksyn$.
 In our simulations, we define as many classes as the number of motions (directions and velocities): $\Nclass = 12 \times 3$ and set the size of the model's kernel to $(\Nclass, \Npol, \Ktime, \Kx, \Ky) = (36, 2, 21, 17, 17)$.
@@
 Using this dense representation, the model's processing of the input $A$ can be written as layer-wise convolution: $B = \kernel \ast A$ (\seeFig{model} for an illustration).
+
+The well-known convolution defines a differentiable measure, which is very efficiently implemented for GPUs, and which we will use to detect the motion direction in the event stream.
+% Therefore, assuming that a neuron integrates only in the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
 A similar type of spatiotemporal filtering was used as a preprocessing stage for an existing pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
@@
 This ensures that only past information contributes to the inference made at the present moment.
-In practical implementation, the kernels are temporally shifted so that the inference at the present time is solely influenced by past information.
+In practice, the kernels are temporally shifted so that the inference at the present time is solely influenced by past information.
 This temporal shift occurs after a duration equivalent to the depth of the kernel, denoted as $\Ktime$.
-The well-known convolution defines a differentiable measure, which is very efficiently implemented for GPUs, and which we will use to detect the motion direction in the event stream.
-% Therefore, assuming that a neuron integrates only in the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
+
 Such a method contrasts with classical methods for delay learning, which explicitly manipulate the delay as a variable and which are not directly differentiable~\citep{nadafian_bio-plausible_2020}.
-
 Keeping the analogy with spiking neurons, the analog activity $B$ represents the integration of synaptic activity, and we will now try to define the detection of motion using the spatiotemporal kernels.
 Since we know that at each instant, there may be different motions, we will define the activation function of our model as a sigmoid function that implements a form of Multinomial Logistic Regression (MLR).
-In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a sigmoid function $\sigma(\beta) = \frac {1} {1 + \exp(-\beta)}$ of the linear combination of the list of events.
+In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a sigmoid function $\sigma(\beta) = \frac {1} {1 + \exp(-\beta)}$ of the result of the convolution.
 Formally, using the kernels, the input raster plot is transformed into a probability with the following formula:
@@
 We computed this threshold to ensure that neurons, on average, generate one spike per second.
-Therefore, the spiking output of the model corresponds to the motions in space and time that possess the highest probability.
+Therefore, the spiking output of the model corresponds to the motions in space and time that represent the highest probability.
 
 Now that this general framework has been described formally, we may include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}.
-Note that the general framework is an extension to that presented in~\cite{grimaldi_learning_2022}, in particular by including a more complex task to solve, the deeper analysis of the results, and these novel neuroscience-inspired heuristics.
+Note that the general framework is an extension to that presented in~\cite{grimaldi_learning_2022}, in particular by including a more complex task, the deeper analysis of the results, and these novel neuroscience-inspired heuristics.
 First, to avoid introducing biases in the directions which may be learned, we apply a circular mask to the spatial dimensions of the kernels.
@@
 This is due to the fact that our whitened images have a relative symmetry in the luminance profiles, that is, that an image with inverted contrast is indistinguishable from a standard one.
-Since this arrangement of polarities is independent of motion, we added a mechanism that collects the linear values for the movie and the movie with the ON and OFF cells flipped, keeping only the maximum value for each voxel.
+Since this arrangement of polarities is independent of motion, we added a mechanism that collects the linear values for the movie and that with the ON and OFF cells flipped, keeping only the maximum value for each voxel.
 This is similar to the computation done for complex cells in primary visual cortex.
@@
 Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{10}$ movies, each of size $200 \times 128 \times 128$, a learning rate of $10^{-5}$ and $100$ epochs).
-We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
+%We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
 
 % TODO citer Levy2021 "Communication consumes 35 times more energy than computation in the human cortex" 
-Finally, the output of the MLR model is an event-based representation that predicts the probability of each motion at each position and time.
+Finally, the output of the MLR model is a representation that predicts the probability of each motion at each position and time.
 Such an output provides a form of optical flow that can be exploited for non-rigid motion, but we have defined here, for simplicity, an evaluation method that applies to our full-field motion task.
 Using the properties of logistic regression, by taking the mean evidence represented in the output given by the model at all positions for any given time, and using the sigmoid function, we can derive each motion's probability at that time.
-This allows one to calculate the accuracy as the percentage of times the motion is accurately predicted at any given time step.
-These calculations are performed on a different input data set than the one used in the training or validation steps.
+Taking the most probable class as the output, this allows one to calculate the accuracy as the percentage of times the motion is accurately predicted at any given time step.
+For validation, these calculations are performed on a different input dataset than the one used in the training or validation steps.
 The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN} (see Data Availability).
@@
     	Representation of the weights for $8$ directions for a single speed (among the $12 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
-The directions are shown as red arrows in the left insets.
+The directions are shown as red arrows in the left insets, where the disks correspond to the set of different possible motions.
 The spatiotemporal kernels are shown as slices of spatial weights at different delays.
@@
 For instance, the first kernel demonstrates a robust preference for downward motion.
-The negative weights are symmetrically arranged around these positive weights, forming a center-surround profile that enhances the response.
+The negative weights are symmetrically arranged around these positive weights, forming a center-surround profile that is known to enhance the response.
 We also observed a strong dependence between the weights reaching the ON polarities and those reaching the OFF polarities.
@@
 
-If we now widen our focus on the interpretation of these kernels in terms of spatiotemporal motifs embedded in the event stream, these show a prototypical anisotropic profile~\citep{kaplan_anisotropic_2013} adapted to motion detection.
+If we now widen our focus on the interpretation of these kernels in terms of spatiotemporal motifs embedded in the event stream, these show a prototypical anisotropic profile adapted to motion detection~\citep{kaplan_anisotropic_2013}.
 In~\citep{grimaldi_robust_2022}, a link was drawn between event-based MLR training and Hebbian learning, allowing to say that the present model learns its weights according to a presynaptic activity associated to the different motion directions.
@@
 events away in the past.
-%
-%
-%%%-----------------------------------------------------------------
-%
-\subsection{Testing with natural-like textures}
-%%%-----------------------------------------------------------------
-%%%-----------------------------------------------------------------
-%: fig:motion_clouds
-%%%-----------------------------------------------------------------
-\begin{figure}%[h!]
-    \centering
-    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
-    % TODO : show prototypes of textures on top
-    \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from a grating (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to grating (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)).
-Note that these accuracies are computed both in the case where the orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) and in the generic case where the orientation is independent of direction ('aperture').
-    }
-    \label{fig:motion_clouds}
-\end{figure}
-%%%-----------------------------------------------------------------
-We also wanted to assess the influence of spatiotemporal parameters of the stimuli on the performance of the model.
-For this, we will quantify its ability to categorize different motions, i.e.
-on event streams for which the ground truth motion is known at each instant.
-Before applying the model to natural scenes that we defined in the motion detection task, we first test the model on simpler, parameterized stimuli.
-For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control, among other parameters, the motion (\seeFig{motion_clouds})~\citep{Leon2012,vacher_bayesian_2018}.
-These stimuli allowed us to control various parameters, including motion, spatial extent, and duration.
-By matching the spatial and temporal characteristics of the generated movies with those of the motion task mentioned earlier, we created a range of textures featuring different spatial properties and motions.
-This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $12$ directions and $3$ speeds.
-For each motion, we also varied the texture parameters, such as mean and variance of orientation or spatial frequency content, to provide some naturalistic variability.
-This method provides a rich dataset of textured movies for which we know the ground truth for the motion.
-
-We observe some interesting facts.
-First, as we change the mean spatial frequency of the texture, we observe a broadly tuned response in accuracy.
-This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006,Perrinet2007} and reveals the most informative scales learned by our model.
-In particular, the accuracy is better for a grating-like stimulus than for a large one (which qualitatively resembles a more textured stimulus), reminiscent of the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}.
-Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.
-When the stimulus is grating-like and the orientation is arbitrary with respect to the direction of motion, the system faces the aperture problem and experiences a decrease in accuracy.
-This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion.
-Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter.
-This shows a progressive decrease in accuracy, similar to that observed in the amplitude of human eye movements~\citep{mansour_pour_speed_2018}.
 %
@@
 This substantial reduction in computations showcases the robustness of the presented method.
+%
+%
+%%%-----------------------------------------------------------------
+%
+\subsection{Testing with natural-like textures}
+%%%-----------------------------------------------------------------
+%%%-----------------------------------------------------------------
+%: fig:motion_clouds
+%%%-----------------------------------------------------------------
+\begin{figure}%[h!]
+    \centering
+    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
+    % TODO : show prototypes of textures on top
+    \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from a grating (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to grating (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)).
+Examples snapshots are shown as an illustration in the top insets.
+Note that these accuracies are computed both in the case where the orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) and in the generic case where the orientation is independent of direction ('aperture').
+    }
+    \label{fig:motion_clouds}
+\end{figure}
+%%%-----------------------------------------------------------------
+In order to assess the influence of spatiotemporal parameters of the stimuli on the performance of the model.
+For this, we will quantify its ability to categorize different motions, i.e.
+on event streams for which the ground truth motion is known at each instant.
+Before applying the model to natural scenes that we defined in the motion detection task, we first test the model on simpler, parameterized stimuli.
+For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control, among other parameters, the motion (\seeFig{motion_clouds})~\citep{Leon2012,vacher_bayesian_2018}.
+These stimuli allowed us to control various parameters, including motion, spatial extent, and duration.
+By matching the spatial and temporal characteristics of the generated movies with those of the motion task mentioned earlier, we created a range of textures featuring different spatial properties and motions.
+This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $12$ directions and $3$ speeds.
+For each motion, we also varied the texture parameters, such as mean and variance of orientation or spatial frequency content, to provide some naturalistic variability.
+This method provides a rich dataset of textured movies for which we know the ground truth for the motion.
+
+We observe some interesting facts.
+First, as we change the mean spatial frequency of the texture, we observe a broadly tuned response in accuracy.
+This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006,Perrinet2007} and reveals the most informative scales learned by our model.
+In particular, the accuracy is better for a grating-like stimulus than for a large one (which qualitatively resembles a more textured stimulus), reminiscent of the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}.
+Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.
+When the stimulus is grating-like and the orientation is arbitrary with respect to the direction of motion, the system faces the aperture problem and experiences a decrease in accuracy.
+This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion.
+Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter.
+This shows a progressive decrease in accuracy, similar to that observed in the amplitude of human eye movements~\citep{mansour_pour_speed_2018}.
 %

### 2023-07-06 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
 \newcommand{\Funding}{
-This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement N° ANR-19-CHR3-0008-03 (``\href{APROVIS3D}{http://aprovis3d.eu/}''). 
-LP received funding from the ANR project  N° ANR-20-CE23-0021 ``\href{AgileNeuroBot}{https://laurentperrinet.github.io/grant/anr-anr/}''
+This research was funded by the European Union ERA-NET CHIST-ERA 2018 research and innovation program under grant agreement ANR-19-CHR3-0008-03 (``\href{APROVIS3D}{http://aprovis3d.eu/}''). 
+LP received funding from the ANR project ANR-20-CE23-0021 (``\href{AgileNeuroBot}{https://laurentperrinet.github.io/grant/anr-anr/}'')
 and from A*Midex grant number AMX-21-RID-025
-``\href{Polychronies}{https://laurentperrinet.github.io/grant/polychronies/}''. 
+(``\href{Polychronies}{https://laurentperrinet.github.io/grant/polychronies/}''). 
 }
@@
 
-% TODO citer Levy2021 "Communication consumes 35 times more energy than computation in the human cortex" 
 Finally, the output of the MLR model is a representation that predicts the probability of each motion at each position and time.
@@
 %%%-----------------------------------------------------------------
+%
+%%%-----------------------------------------------------------------
+%: fig:accuracy
+%%%-----------------------------------------------------------------
+\begin{figure}%[h!]
+    \centering
+    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
+    \caption{
+        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
+The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
+As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about $250$ times fewer computations.
+        }
+    \label{fig:accuracy}
+\end{figure}
+%
 After training our MLR model, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection.
+For this, we quantify its ability to categorize different motions, i.e.
+on event streams for which the ground truth motion is known at each instant.
 The accuracy computed on the test set was approximately $89\%$ (with a chance level of $1 / 12/ 3 \approx 2\%$).
@@
 
-We observed that the distribution of the kernel's weights is sparse, with most values close to zero (\seeFig{kernels}).
+We also observed that the distribution of the kernel's weights is sparse, with most values close to zero (\seeFig{kernels}).
 As shown in the formalization of our event-based model, the computational cost of our model, if implemented on a neuromorphic chip, would be dominated by the computations used for the convolution operation.
@@
 Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of non-zero synaptic weights.
+This hypothesis is consistent with biological observations which have shown that communication consumes 35 times more energy than computation in the human cortex~\citep{Levy2021}.
 
-In order to evaluate the resilience of the classification performance with respect to computational load, we adopt a pruning approach, where we remove weights in $\kernel$ that fall below a specified threshold.
+In order to evaluate the resilience of the classification performance with respect to computational load, we adopt first a pruning approach, where we remove weights in $\kernel$ that fall below a specified threshold.
 The accuracy of classification is then plotted as a function of the relative number of computations or active weights per decision for each neuron in the layer (refer to Fig.~\ref{fig:accuracy}).
-To provide a basis for comparison and to account for the benefits of utilizing variable delays, we present the accuracy achieved by an MLR model employing a shortening strategy.
+To provide a basis for comparison and to account for the benefits of utilizing variable delays, we also present the accuracy achieved by an MLR model employing a shortening strategy.
 This strategy involves adjusting the temporal width by selecting only the weights associated with the shortest delays.
@@
 %%%-----------------------------------------------------------------
-In order to assess the influence of spatiotemporal parameters of the stimuli on the performance of the model.
-For this, we will quantify its ability to categorize different motions, i.e.
-on event streams for which the ground truth motion is known at each instant.
-Before applying the model to natural scenes that we defined in the motion detection task, we first test the model on simpler, parameterized stimuli.
-For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control, among other parameters, the motion (\seeFig{motion_clouds})~\citep{Leon2012,vacher_bayesian_2018}.
-These stimuli allowed us to control various parameters, including motion, spatial extent, and duration.
+In order to assess the influence of spatiotemporal parameters of the stimuli on the performance of the model, we now test the model on simpler, parameterized stimuli.
+For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control relevant parameters for motion detection, including motion direction, spatial orientation, and spatial frequency along with their respective precisions (\seeFig{motion_clouds})~\citep{Leon2012,vacher_bayesian_2018}.
 By matching the spatial and temporal characteristics of the generated movies with those of the motion task mentioned earlier, we created a range of textures featuring different spatial properties and motions.
@@
 This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006,Perrinet2007} and reveals the most informative scales learned by our model.
-In particular, the accuracy is better for a grating-like stimulus than for a large one (which qualitatively resembles a more textured stimulus), reminiscent of the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}.
+Then, by modifying the bandwidth in spatial frequency, we show that the accuracy is worse for a grating-like stimulus than for a large one (which qualitatively resembles a more textured stimulus), reminiscent of the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}.
 Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.
-When the stimulus is grating-like and the orientation is arbitrary with respect to the direction of motion, the system faces the aperture problem and experiences a decrease in accuracy.
+When the stimulus is grating-like and the orientation is arbitrary with respect to the direction of motion, the system faces the aperture problem and experiences a sharp decrease in accuracy.
 This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion.
@@
 %
-%%%-----------------------------------------------------------------
-%: fig:accuracy
-%%%-----------------------------------------------------------------
-\begin{figure}%[h!]
-    \centering
-    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
-    \caption{
-        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
-The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
-As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about $250$ times fewer computations.
-        }
-    \label{fig:accuracy}
-\end{figure}
-%
 \section{Discussion}
 %%%-----------------------------------------------------------------
-In this paper, we have introduced a generic heterogeneous delay SNN.
-It is trained using supervised learning on a visual motion detection task.
+In this paper, we have introduced a generic heterogeneous delay SNN that was trained using supervised learning on a visual motion detection task.
 We have shown how the model performs well compared to other event-based classification algorithms for this task.
@@
 Taken together, these oriented edges form the contours of visual objects in the scene~\citep{koenderink_representation_1987}.
-Thus, there is an interdependence between motion information and orientation information within the event stream.
+Thus, there is an interdependence between motion information and orientation information within the event stream, which we put in evidence by the shape of the kernels.
 It would be crucial to investigate this dependency further.
@@
 This type of learning is thought to be prevalent in the central nervous system and, assuming the signal is sparse~\citep{olshausen_emergence_1996}, one could extend these Hebbian sparse learning schemes to spikes~\citep{perrinet_emergence_2004, masquelier_competitive_2009}.
-We expect that this would be particularly useful for exploring neurobiological data.
+We expect that this would be particularly useful for exploring neurobiological data~\citep{Perrinet23ICANN}.
 Indeed, there is a large literature showing that brain dynamics often organize into stereotyped sequences, such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}.

### 2023-07-06 cover_letter/2023-07-07_coverletter.tex
cover

--- 2023-02-20 cover_letter/2023-02-20_coverletter.tex
@@
 
+We thank the reviewer for the positive assesment which encouraged us to further improve our manuscript.
 Given the substantial modifications we did to the original manuscript to reduce and clarify the different sections we do not provide a tracked changes version of the revision (such a document can be provided if needed).
-We provide a point by point response to the reviewers with references to the manuscript.
+We provide below a point by point response to the reviewers with references to the manuscript.
 
@@
 
-In the following we will give response to reviewers in \noteRev{framed boxes.}
+In the following we will give response to reviewers using: \noteRev{framed boxes.}
 
@@
 
- \noteRev{Overall, we would like to thank Reviewer 1 for its comments that allowed us to focus on reshaping the manuscript and give clearer and shorter explanation of the work we submit in your journal.
-Because we did major modifications in the manuscript, we do not provide a tracked changes version of the revision but we can easily provide one if needed.
-}
+\begin{quote}
+	The author's organization of the paper is much clearer than last time, and I suggest making the following modifications:
+	1.
+For the natural-like texture images used, the author carried out some operations, such as changing texture parameters, etc.
+I suggest that the authors give 1-2 examples of these images corresponding to the different subgraphs in Figure 5, such as isotropic textures and grating-like textures.
+\end{quote}
+
+ \noteRev{We thank the reviewer for his constructive comment.
+This encouraged us to extend our task from using natural-like textures to actual natural images.
+This was further extended by using biologically inspired eye movements.
+We have kept the use of the synthetic textures in order to assess the relevant image parameters that are relevant for motion detection and we have drawn the analogies of our model with that found in the neurophysiological and psychophysical literature.
+We have added subgraph inset to illustrate the changes in parameters to the generation of figures.}
 
 \begin{quote}
- 1. There is a lack of an intuitive figure to show the HD-SNN model used in this paper.
-\end{quote}
-
- \noteRev{We provide such a figure in the new manuscript: Figure 3, P9}
-
-\begin{quote}
- 2. The MLR model used in the experimental part was presented in authors’ another work. Therefore, the changes on the MLR model are suggested to be highlighted and contrasted graphically.
+	2. Experiments that assess accuracy should be repeated many times (and give deviation) to eliminate data uncertainty.
 \end{quote}
@@
 \begin{quote}
- 3.
-The title of section 2.2.1 is ’’a generative model for rater plots’’, but actually this section introduces basic terms and symbols.
-Section 2.2.2 is named ‘’detecting spiking motifs’’, but most of the content seems to introduce the generation process of raster plots, while only the last paragraph introduces the detection of motif.
-It is recommended that the content of these two sections be reorganized and that references to Figure 1(b)(c)(d) be added where appropriate during the description.
+	3. I would still recommend that the authors add experiments on other data sets, or compare with other works  on the current data set.
 \end{quote}
@@
 \begin{quote}
- 4. The adjustment of weights and delays needs more detailed elaboration, preferably given in mathematical formulas.
+	4. There are two “observe” in line 355.
 \end{quote}
 
- \noteRev{We improved the mathematical formulation of the model in sections 2.2.1 and 2.2.2 and hope that it clarifies the general framework we use.
-Regarding training, it is described more clearly in P9, L313.
-Demonstration of the mechanism and the analogy with Hebbian learning was done in~\cite{grimaldi_robust_2022} and was not reproduced here.
-We hope this new version offers a clearer methodology.}
-
-\begin{quote}
- 5.
-In Figure 4, it is claimed that the kernels are very similar for the ON polarities, and different kernels are selective to the different motion directions.
-However, in terms of these two observations, I think they are not very significant, so I suggest a more intuitive way to reflect them, such as quantifying the similarity between the ON kernels, and the angle between the kernels and the vector of the standard motion directions.
-\end{quote}
-
- \noteRev{We gave a better explanation of the symmetry assumption in the methods (P9, L303) an how this new model uses this invariance in its computation like observed in complex cells.
-Indded, we don't want the polarity to have an impact on the classification and then this observation does not apply to the revised manuscript as it is a prior and not an emergent property.}
-
-\begin{quote}
- 6.
-There are few experiments and lack of comparative experiments.
-Since the model has an improvement based on MLR model, it is suggested to increase the comparison with MLR model and HOTS model on other datasets, such as N-MNIST and other neuromorphic datasets, so as to show the advantages of the proposed model in computational consumption.
-\end{quote}
-
- \noteRev{The reviewer is making a good point but regarding the deadline for the revision and some objectives we focused on to improve the manuscript we could not apply the model on DVS recordings.}
-
-\begin{quote}
- 7. Some sections are verbose, such as the discussion.
-\end{quote}
-
- \noteRev{We significantly changed the introduction (section 1), the methods (section 2) and the discussion (section 4) to address this comment.
-We hope to provide a more concise and clear explanation of our study.}
-
-\begin{quote}
- 8. There are some minor errors in the text, and it is recommended to check the full text: 
- 1)In the first paragraph on page 8, "a presynaptic address $b_s$" should be "a postsynaptic address $b_s$"; 
- 2)Whether $W_b$ in the first formula should be $K_b$? And d in it lacks description. 
- 3)"PGs" should be given in its full name when it first appears. 
- 4)There is no subgraph label of (a), (b) and (c) in Fig. 3, but the expression of Fig. 3-(a)/(b)/(c) appears many times. 
- 5)The content of paragraph 1 on page 13 is repeated. 
-\end{quote}
-
- \noteRev{We corrected these minor errors and thank the reviewer for reporting them.}
-
-\textbf{Reviewer \#2}
-
-\begin{quote}
- {\it Key results: Please summarize what you consider to be the outstanding features of the work.}
- Using a single layer of spiking neurons with heterogeneous delays to learn spatio-temporal  spiking motifs
-\end{quote} 
-\noteRev{We would like to thank reviewer \#2 for his comments. We have thoroughly revised the manuscript to take them into account.}
-
-\begin{quote}
- The work presented here is very similar to the author's other work,  "Learning hetero-synaptic delays for motion detection in a single layer of spiking neurons" 10.1109/ICIP46576.2022.9897394.
-This must be addressed and cited with an explanation of how this work presented is an extension of this previous publication.
-Abstract, Fig 6 and the main result is presented in the previous work.
-Arguably some of the figures from the previous publication would be nice to see in this publication as it helps with the visualisation of the problem.
-One suggestion for the extension to make the work unique would be to extend with some event-camera data, as it is mentioned multiple times within the publication.
- There are also a number of citations that could have been added that did appear in the other publication, while also including other motion detection/ segmentation or previous work on event-based optical flow, which is a very similar problem.
-\end{quote}
-
-\noteRev{In fact, there was no mention of this previous work in the original manuscript, as this 4-page paper had not yet been published in the proceedings of the conference at which it had been presented.
-As mentioned in the responses to reviewer \#1 in this cover letter, we have now added this citation~\cite{grimaldi_learning_2022}.
-In addition, in this submitted manuscript we describe the main differences from that previous study: the use of a more ecological task (Section 2.1), a complete formulation of the model (Section 2.2), and the improvements made in this particular model (P8, L285).
-We also provide citations to studies using similar methods (P8, L259).
-Further results are presented, especially the testing of the learned model with parametric textures.}
-
-\bibliographystyle{plainnat}
-\bibliography{FastMotionDetection}
+ \noteRev{Fixed.}
 

### 2023-07-06 cover_letter/2023-07-07_coverletter.tex
$200$ test movies

@@
  \noteRev{We thank the reviewer for his constructive comment.
-This encouraged us to extend our task from using natural-like textures to actual natural images.
-This was further extended by using biologically inspired eye movements.
+While keeping the same computationnal framework, this encouraged us to extend our task from using natural-like textures to actual natural images.
+This was further extended by using biologically inspired eye movements to generate dynamical scenes with known ground truth.
 We have kept the use of the synthetic textures in order to assess the relevant image parameters that are relevant for motion detection and we have drawn the analogies of our model with that found in the neurophysiological and psychophysical literature.
@@
 
- \noteRev{The main differences with the previous study are in the task to solve (section 2.1) and in the analysis of the results.
-The response of the model is tested on different sets of motion cloud stimuli with a control of the parameters of such stimuli generation (section 3.2).
-Regarding the MLR model, main differences with the previous ones are reported in the methods (P8, L285).}
+ \noteRev{We have repeated the testing of our model using $200$ movies and report less variable estimates for the accuracy. Thanks for this suggestion.}
 
@@
 
- \noteRev{We removed these sections to have a more concise explanation of the model and focused on resolving the motion detection task.
-We hope the manuscript is now more concise and clear to the point.}
+ \noteRev{By extending the learning and testing of our model on this newly defined motion detection task, we hope to have provided with a more thorough exploration of the capabilities of our model.
+The comparison of accuracy was also performed on the synthetic textures dataset in order to provide with a direct comparison with biological observations.}
 

### 2023-07-07 Grimaldi-etal-BiolCybernetics.tex

--- 2023-07-06 Grimaldi-etal-BiolCybernetics.tex
@@
 % https://github.com/SpikeAI/pyTERtorch/ (for dev)
-% https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN (TODO: to become public)
+% https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN 
 % https://www.zotero.org/groups/4776796/fastmotiondetection
@@
 %● The roles of attention mechanisms in biological vision.
-
-% 2023-05-20: rev1 - final
-% Reviewer 1
-% The author's organization of the paper is much clearer than last time, and I suggest making the following modifications:
-% 1.
-For the natural-like texture images used, the author carried out some operations, such as changing texture parameters, etc.
-I suggest that the authors give 1-2 examples of these images corresponding to the different subgraphs in Figure 5, such as isotropic textures and grating-like textures .
-% 
-% 2. Experiments that assess accuracy should be repeated many times (and give deviation) to eliminate data uncertainty.
-% 
-% 3. I would still recommend that the authors add experiments on other data sets, or compare with other works  on the current data set.
-%
-% 4. There are two “observe” in line 355.
-%
-% In order to process your paper, we require:
-%
-% • A point-by-point response to the comments, including a description of any additional experiments that were carried out and a detailed rebuttal of any criticisms or requested revisions that you disagreed with.
-% 
-% This must be uploaded as a 'Point-by-point response to reviewers' file. All changes to the manuscript must be highlighted or indicated by using tracked changes.
-% 
-% At this stage, please also ensure that you have replaced your initial-submission image files with production quality figures.
-These should be supplied at 300 dpi resolution for .jpeg and .tiff or as .eps files.
-Figures should not include Figure number labels in the image.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand{\WebsiteLP}{https://laurentperrinet.github.io}%
+\newcommand{\Abstract}{
+The response of a biological neuron depends on the precise timing of afferent spikes.
+This temporal aspect of the neural code is essential to understand information processing in neurobiology and applies particularly well to the output of neuromorphic hardware such as event-based cameras.
+However, most artificial neural models do not take advantage of this important temporal dimension of the neural code.
+Inspired by this neuroscientific observation, we develop a model for the efficient detection of temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
+The model uses the property that the diversity of synaptic delays on the dendritic tree allows for the synchronization of specific arrangements of synaptic inputs as they reach the basal dendritic tree.
+We show that this can be formalized as a time-invariant logistic regression that can be trained on labeled data.
+We demonstrate its application to synthetic naturalistic videos transformed into event streams similar to the output of the retina or to event-based cameras and for which we will characterize the accuracy of the model in detecting visual motion.
+In particular, we quantify how the accuracy can vary as a function of the overall computational load showing it is still efficient at very low workloads.
+This end-to-end, event-driven computational building block could improve the performance of future spiking neural network (SNN) algorithms and in particular their implementation in neuromorphic chips.
+}
 \newcommand{\Keywords}{time code, event-based computations, spiking neural networks, motion detection, efficient coding, logistic regression
@@
 Thanks to % Thomas Schatz, Rosa Cossart,
-Salvatore Giancani, Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work. 
+Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work. 
 \Funding %
@@
 % \newcommand{\simil}{\beta} % similarity value
+\newcommand{\Nclass}{N_\text{class}} % number of classes for MLR:
 \newcommand{\Nx}{N_\text{X}}
@@
 \newcommand{\Ntime}{N_\text{T}}
-\newcommand{\kernel}{K} % convolution kernel
-% \newcommand{\kernel}{\mathcal{K}} % convolution kernel
+\newcommand{\kernel}{\mathcal{K}} % convolution kernel
 %\newcommand{\kernelind}{\mathbf{k}} % indice of the kernel
@@
 \newcommand{\class}{c} % class k of the MLR
-% \newcommand{\Nclass}{N_\text{class}} % number of classes for MLR:
-\newcommand{\Nclass}{N_\class} % number of classes for MLR:
 \newcommand{\lrweights}{\theta} % matrix of MLR weights
@@
 \newcommand{\loss}{J} % cost function for MLR
+\newcommand{\softmax}{\sigma}
 \newcommand{\actfreq}{f}
@@
 \abstract{
-The precise timing of spikes from afferent neurons determines the response of a biological neuron.
-This temporal aspect of neural activity is essential for understanding information processing in neurobiology, and is particularly applicable to the output of neuromorphic hardware such as event-based cameras.
-However, most artificial neural models ignore this important temporal dimension of neural activity.
-Here, we develop a model for efficiently detecting temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
-The model exploits the property that the diversity of synaptic delays on the dendritic tree allows specific arrangements of temporally precise synaptic inputs to synchronize as they reach the basal dendritic tree.
-We show that this process can be formalized as a time-invariant logistic regression that can be trained on labeled data.
-We demonstrate its application to naturalistic videos transformed into event streams, similar to the output of the biological retina or event-based cameras.
-To characterize the robustness of the model in detecting visual motion, we test its efficiency by pruning some weights and show that it is still efficient at very low workloads.
-In conclusion, by providing an end-to-end, event-driven computational building block, the learning of heterogeneous delays could improve the performance of future spiking neural network algorithms, and in particular their applications in neuromorphic chips.
+\Abstract
 }
@@
 %
-The human brain has the remarkable capacity to react efficiently at any given time while consuming a reasonable amount of energy, in the order of $20$ watts.
-This system is the result of millions of years of natural selection, and a striking difference with artificial neural networks is the representation that both use.
-Indeed, our computers use digital values and for instance, the convolutional neural networks (CNNs) which are used for processing images represent the flow of information from one layer to another using tensors of floats.
-These networks store visual information densely across the visual topography, with different translation-invariant properties represented in different channels.
+The human brain has the remarkable property of being able to react at any time while consuming a reasonable amount of energy, about tens of watts.
+This system is the result of millions of years of natural selection, and a striking difference with artificial neural networks is the representation both use.
+Artificial convolutional neural networks (CNNs), for example, represent the flow of information from one layer to another as tensors, storing visual information densely across the visual topography, with different properties represented in different channels.
+These networks are known to mimic many properties of biological systems, such that each can be attributed a ``Brain Score''~\citep{schrimpf_brain-score_2020}, yet this score takes no account of inference speed or energy consumption.
 CNNs have achieved state-of-the-art performance for some computer vision tasks, such as image recognition.
-These networks are also known to mimic several properties of biological systems, such that each can be assigned a ``brain score''~\citep{schrimpf_brain-score_2020}.
-However, this score does not take into account key aspects of the efficiency of biological systems, such as inference speed or energy consumption, which is about $360$ watts on a standard GPU (NVIDIA RTX 3090).
-In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing (binary) events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
-Spiking neural networks (SNNs), known as the third generation of artificial neural networks, incorporate this temporal dimension into the way they perform their computations.
-One example of a SNN is the SpikeNet algorithm, which takes a purely temporal approach by encoding information using at most one spike per neuron~\citep{delorme_spikenet_1999}.
-Alternatively, the SNN implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} that uses the exact timing of spikes and learns the structure of the network by minimizing a specifically defined cost function.
-This was recently extended using the surrogate gradient method which is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
-However, the performance of SNNs still lags behind that of networks using an analog, firing rate-based representation.
+However, they do not take advantage of the dynamics inherent in the way we perceive our natural environment.
+In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
+The third generation of artificial neural networks, known as spiking neural networks (SNNs), incorporates this temporal dimension into the way they perform their computations.
+They are of interest for computational neuroscience because they provide a better model of the biological neuron.
+Some approaches have developed normative models of SNNs that aim to have applications in machine learning.
+One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}.
+Another type of SNN that uses accurate timing of spikes attempts to determine the structure of the network to minimize a cost function.
+This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been recently extended.
+This method uses a surrogate gradient and is now widely used in attempts to transfer the performance of CNNs to SNNs~\citep{zenke_remarkable_2021}.
+However, the performance of SNNs still lags behind that of firing-rate based networks.
 The question of the advantage of using spikes in machine learning and computer vision remains open.
 
-In a recent review, we reported previous theoretical and  experimental evidence for the use of precise spiking motifs in biological neural networks~\cite{grimaldi_precise_2023}.
-In particular, \citet{abeles_role_1982} asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs.
-While the first hypothesis favors the firing rate coding theory, the second emphasizes on the importance of temporal precision in neural activity.
-Since then, numerous studies have demonstrated the emergence of synchronous activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding using precise spike timings~\citep{Perrinet2002,perrinet_coding_2004, gollisch_rapid_2008}, or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
-All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatiotemporal spiking motifs in biological spike trains.
-In neural models, the definition of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows the efficient detection of these spatiotemporal motifs embedded in the spike train.
-These spatiotemporal motifs present in neural activity may form useful representations to perform computations for various cognitive tasks using the synchrony of spikes reaching the soma of a neuron.
-In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a spiking motif defined by a subset of neurons with different, but precise, relative spiking delays.
-This delay between a pair of connected neurons is defined as the time between the emission of a spike at the soma of the afferent neuron and its arrival at the soma of the efferent neuron.
-Importantly, due to the variety of weights and delays within a population, representations using polychronous groups have, in theory, a much higher information capacity than a firing rate-based approach.
+In a recent review paper, we reported experimental evidence for the presence of precise spiking motifs embedded in recordings from biological neural tissues~\cite{grimaldi_precise_2023}.
+These spatio-temporal motifs within the neural activity may be useful representations to perform computations for various cognitive tasks.
+Notably, \citet{abeles_role_1982} asked whether the role of cortical neurons is the integration of synaptic inputs or rather the detection of coincidences in temporal spiking motifs.
+While the first possibility favors the firing-rate coding theory, the second one emphasizes the function of temporal precision in the neural code.
+Since then, numerous studies have demonstrated the emergence of synchronicity of activity within a neuronal population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient coding using spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}.
+All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of spatio-temporal spiking motifs in biological spike trains.
+In neural models, the integration of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows for the efficient exploitation or detection of these spatio-temporal motifs embedded in the spike train.
+In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, but precise, relative spiking delays, i.e., the time between the arrival of an afferent spike at a given synapse and the contribution of the associated postsynaptic potential to the neuron's soma.
+Due to the variety of configurations and the possible coexistence of multiple superimposed motifs, representation with polychronous groups has a much higher information capacity than a firing rate-based approach to neural coding.
 
-The present paper proposes a real-world application that extends a recently proposed model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}.
-This model was trained to solve a simplified motion detection task on a synthetic event-based dataset generated by moving parameterized textures, and provides a first demonstration that formal neurons can exploit the precise timing of spikes to detect motion thanks to heterogeneous delays.
-Here, we extend these results to a much more complex and natural setting.
-First, we define the ecological cognitive task that the model must solve with the different datasets on which it will be tested.
-Instead of the synthetic textures used previously, we use natural scenes synthesized from natural images translated by biologically inspired saccadic movements.
-We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model from efficiency principles and derive a learning rule to adapt the weights of each heterogeneous synaptic delay using gradient descent.
-Applied to this detection task, we study the emergence of spatiotemporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner.
-We investigate the efficiency of the motion detection mechanism and, in particular, its resilience to synaptic weight pruning.
-Indeed, once trained, the amount of event-driven computation could be drastically reduced by removing weak synapses while maintaining peak performance for the classification task.
-In this way, we will be able to show how such a model can provide an efficient solution to the energy/accuracy tradeoff.
+The present paper proposes additional experiments extending a recent model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}.
+This model was trained to solve a motion detection task on a synthetic event-based dataset generated by moving parameterized textures.
+Once trained, the volume of event-driven computations could be drastically reduced by pruning the synapses, while maintaining top performance for the classification task.
+This was a demonstration that formal neurons can exploit the precise timing of spikes to improve neural computations.
+In the present work, we extend motion detection to a more ecological task.
+Instead of the synthetic textures that have been used to generate motion-driven stimuli, we use natural images and random movements that mimic biological saccades.
+Using an ecological motion detection task, we study the emergence of spatio-temporal spiking motifs when a single layer of spiking neurons is trained on a supervised classification task.
+First, we define the ecological cognitive task the model has to solve with the different datasets it will be tested on.
+Then, we develop a theoretically defined Heterogeneous Delays Spiking Neural Network (HD-SNN) model capable of learning these heterogeneous synaptic delays.
+We will investigate the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning.
+In this way, we will be able to show how such a model can provide an efficient solution to the trade-off between energy and accuracy.
 %
@@
 %%%-----------------------------------------------------------------
-In this paper, we test whether the HD-SNN model is able to efficiently learn to solve a motion detection task defined by a realistic event-based data stream.
-This type of signal is commonly captured by an event-based camera, or Dynamic Vision Sensor (DVS), and is inspired by the signal sent from the retinal ganglion cells through the optic nerve to the visual cortex.
-The events are binary in nature and should be sufficient to perform the motion detection task in a fast and efficient manner.
-Here, we first describe the task definition, then the HD-SNN model we use for motion inference, and finally how we train it.
+In this paper, we aim to test whether the HD-SNN model is capable of efficiently learning a motion detection task, which is defined by a realistic event-based data stream.
+This type of signal is typically captured by a Dynamic Vision Sensor (DVS) and is inspired by the signal that is sent from the retinal ganglion cells through the optic nerve.
+The events are binary in nature and should be sufficient to perform the task of motion detection in a fast and efficient manner.
+Here we will first describe the task definition and then the HD-SNN model we use for motion inference and how we train it.
 %%%-----------------------------------------------------------------
@@
 %%%-----------------------------------------------------------------
-To train and validate our model which uses supervised learning, we need to define a visual dataset for which we explicitly know the ground truth motion, that is, direction and speed.
-To achieve this, we define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
-First, we draw a trajectory inspired by biological eye movements.
-Indeed, these movements allow us to dynamically actuate the center of vision in the visual field.
-In animals with a fovea, this is particularly useful as it allows the area with the highest density of photoreceptors to be moved, for example, to a point of interest in the environment.
+To train and test our model, we need to define a visual dataset for which we control the direction and speed of motion.
+Let us now define a procedure for animating a natural visual scene with virtual eye movements, similar to those used in neurobiological~\citep{vinje_sparse_2000, baudot_animation_2013} and computational neuroscience~\citep{kremkow_push-pull_2016} studies.
+First, we draw a trajectory inspired by the biological movements of the eyes.
+Indeed, these movements allow us to dynamically actuate the center of vision, or gaze, in the visual field.
+In animals with a fovea, this is particularly useful because it allows the area with the highest density of photoreceptors in the environment to be moved, for example to a point of interest.
 A specific mechanism for this function are saccades, which are rapid eye movements that reposition the center of vision.
-In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}), very fast (about $80~\ms$), and have a wide range of speeds.
-On a more microscopic scale, the human gaze moves with minute micro-saccades which trajectories are similar a Brownian trajectory~\citep{poletti_head-eye_2015}.
-To preserve the full generality of the task, we define eye movements using such a form of random walk~\citep{engbert_integrated_2011}: We first define a finite set of possible motions in polar coordinates as the Cartesian product of the $12$ regularly spaced directions and $3$ geometrically spaced speeds.
-Next, we define a saccadic path as a sequence of time segments whose durations are drawn from a Poisson distribution with a mean block length of $24~\ms$, similar to a Lévy flight~\citep[p.
+In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}).
+They are generated very quickly (about $80~\ms$) and have a wide range of speeds.
+On a more microscopic scale, the human gaze moves with a continuous drift similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
+In order to maintain the full generality of the task, we will define eye movements using a form of random walk~\citep{engbert_integrated_2011}.
+This approach first defines a finite set of possible 2D movements in polar coordinates.
+Based on the distribution of biological movements, we simplified it by selecting a set of eye movements as the Cartesian product of $8$ linearly spaced movement directions and 3 geometrically spaced speeds.
+Next, we define a gaze trajectory as segments whose duration is drawn from a Poisson distribution with a mean block length of $24~\ms$, similar to a Lévy flight~\citep[p.
 289]{mandelbrot_fractal_1982}.
-Finally, assuming that  motion is stationary during each segment and that motions for each flight are drawn uniformly and independently, the global trajectory is generated by integrating this motion sequence.
-This generative model yields trajectories that are qualitatively very similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
+Finally, the trajectory is integrated, assuming first that the velocities are sampled uniformly and independently from the set of different motion sets, and second that the motion is uniform during a time segment.
+The resulting instances yield trajectories that are qualitatively similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
 
@@
     \caption{
-        {\bf Motion Detection Task.} To generate realistic event-based dynamic scenes, we mimic the effect of minute saccadic eye movements on a large natural scene ($1024\times1024$) by extracting an image ($128\times128$) which center is moving dynamically according to a jagged random walk.
-\textit{(Left)}~We show an instance of this trajectory (with a length of $200~\ms$, green line) superimposed on the luminance contrasts observed at time step $20~\ms$.
-\textit{( Right)}~The dynamics of this image, translated according to the saccadic trajectory, produces a naturalistic movie, which is then transformed into an event-based representation.
-We show snapshots of the resulting synthetic event stream at different time steps (start and end frames are marked by a white and black dot, respectively, in the left inset).
-Mimicking the response of ganglion cells in the retina, this representation encodes at each pixel all-or-none increases or decreases in luminance, i.e., ON (red) and OFF (blue) spikes.
-In the lower left corner of the snapshots, we show the corresponding instantaneous motion vector (red arrow).
-Note the change in the direction of motion between the third and fourth frames, and also that contours parallel to the motion produce fewer luminance changes, the so-called aperture problem, and thus relatively fewer spikes.
-        }
+        {\bf - Motion Detection Task.} \textit{(Left)}~We use large natural images ($256\times256$) in which an aperture ($64\times64$) extracts a cropped image around the view axis.
+To mimic the effect of a saccadic eye movement, the view axis moves according to a stepwise random walk.
+We show such a trajectory with a length of $128$ time steps (green line).
+\textit{(Right)}~Snapshots of the synthetic event stream at different time steps (start and end frames marked by a white and black dot, respectively).
+The dynamics of the cropped image translated according to the trajectory as a function of time produces a naturalistic movie, which is then transformed into an event-based representation.
+Mimicking the retina, this representation encodes proportional increases or decreases in luminance, i.e.
+ON (red) and OFF (blue) events, in each pixel of the image.
+In the lower left corner of the snapshots, the translation vector is shown in red as one of the possible classes of motion.
+Note the change in the direction of motion between the third and fourth image, and also that, due to the aperture problem, contours parallel to the motion emit relatively fewer spikes.}
     \label{fig:motion_task}
@@
 Once these eye trajectories are generated, we can apply them to a visual scene.  
-For this purpose, we selected a database of $100$ large-scale natural images that were previously used to study the statistics of natural images~\citep{vanHateren1998}.
-Note that these were pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band, similarly to a process known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $1024 \times 1024$ in size, and we crop images of size $128 \times 128$ positioned around the center of gaze at each time step.
-We discretize time in $1~\ms$ bins and produce movies of duration $N_T = 200~\ms$.
-To avoid boundary effects, we randomly position the full trajectory in image space so that the sub-image is translated using the position given by the trajectory at each time step and without touching the boundaries.
-At each time step, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
-Note that the magnitude of the displacement is relative to the time bin, and we have defined the unit speed to correspond to a movement of one pixel per frame (i.e., per time bin of $1~\ms$).
+For this purpose, we selected a database of $600$ natural images that were previously used to study the statistics of natural images~\citep{perrinet_edge_2015}.
+Note that these have been pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band.
+This process is known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
+These images are $256 \times 256$ in size, and we will extract sub-images of size $64 \times 64$ positioned around the center of gaze at each time step.
+We will discretize time in $1~\ms$ bins and produce movies of duration $N_T = 128~\ms$.
+To avoid boundary effects, we will randomly position the full trajectory in image space such that the subimage is translated using the position given by the trajectory at each time step without touching the borders.
+At each point in time, the translation is computed using a coordinate roll in the horizontal and vertical dimensions, followed by a sub-pixel translation defined in Fourier space~\citep{perrinet_sparse_2015}.
+Note that the magnitude of the displacement is relative to the time bin, and we have defined the unit velocity to correspond to a movement of one pixel per frame (i.e., per time bin).
 
-To transform each movie into events, as recorded by a DVS, we compute a residual gradient image which we initialize at zero.
-We then compute the temporal gradient of the pixels' intensity over two successive frames.
-For a given pixel and time stamp, an event is generated when the absolute value of this gradient exceeds a threshold.
-The event has either an OFF or ON polarity, depending on whether the gradient is negative or positive.
-The signed threshold is then subtracted from the residual gradient image.
-When applied to the whole movie, the event stream is similar to the output of a neuromorphic camera~\citep{Gallego2022}, i.e.
-a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF), and their time $\timev_\arank$ (\seeFig{motion_task}-\textit{(Right)}).
-Ultimately, the goal of the model is to infer the correct motion solely by observing these events.
-%%%%-----------------------------------------------------------------
+To transform each movie into events as the ones recorded by a DVS, we compute a gradient image (initialized at zero) by computing the temporal gradient of the pixels' intensity over two successive frames.
+For a specific pixel and timestamp, an event is generated if the absolute value of this gradient exceeds a threshold.
+The event has either an OFF or ON polarity respectively, whether the gradient is negative or positive.
+This signed threshold value is then subtracted from the residual gradient image.
+When applied to the whole movie, the event stream is similar to the output of a neuromorphic camera~\citep{rasetto_challenges_2022}, that is, a list of events defined by $x_\arank$ and $y_\arank$ (their position on the pixel grid), their polarity $\polev_\arank$ (ON or OFF) and time $\timev_\arank$  (\seeFig{motion_task}-\textit{(Right)}).
+The goal here is to infer the correct motion solely by observing these events.
+% \note{ the task is full-field rigid translations, the model resolves it locally (analogy to the input from V1 to MT) }
+%%%-----------------------------------------------------------------
 \subsection{The Heterogeneous Delays SNN (HD-SNN) model}
-%%%-----------------------------------------------------------------
 %----------------------------%
@@
     \centering
-    % TODO: corriger post to pre
     \includegraphics[width=0.980\linewidth]{figures/HDSNN_graph.pdf}
-    \caption{\textbf{Core mechanism of the HD-SNN model.} \textit{(Left-Top)}~Four presynaptic neurons show some spiking activity in which a spiking motif is embedded (starting at time $t=50~\ms$).
+    \caption{\textbf{The core mechanism of the HD SNN model.} \textit{(Top-Left)}~Two spiking motifs are emitted from four presynaptic neurons.
+Once integrated by the synapses of the postsynaptic neuron~\textit{(Right)}, the spiking motifs are shifted in time by the synaptic delays and weighted by the synaptic weights~\textit{(Middle-Left)}.
+When they reach the soma of the postsynaptic neuron, the different spikes contribute to a modification of its membrane potential according to an activation function.
+In this example we use the same activation function as for a Leaky Integrate and Fire neuron~\textit{(Bottom-Left)}.
+The first spiking motif is synchronized by the synaptic delays and causes a sudden rise in the membrane potential of the postsynaptic neuron.
+An output spike is emitted when the membrane potential reaches the threshold and it is then reset.
 \textit{(Right)}~An illustration of a spiking neuron with different synaptic weights (represented by the thickness of the synapses) and different synaptic delays (represented by the length of the synapses).
-\textit{(Left-Middle)}~Each spike is weighted by the synaptic weights (height of the blue bars) and shifted in time according to the synaptic delays on each respective synapse (input spikes are shown in light gray for comparison).
-As a result, the spikes from the spiking motif are synchronized as they reach the soma of the postsynaptic neuron.
-\textit{(Left-Bottom)}~These spikes are then integrated, and contribute to a modification of the membrane potential of the output neuron according to the neural activation function.
-In this example, we use the activation function of a Leaky Integrate-and-Fire neuron.
-The first spiking motif is synchronized by the synaptic delays and causes a sudden rise in the membrane potential of the postsynaptic neuron.
-An output spike is emitted at time $t=75~\ms$ when the membrane potential reaches the threshold, and it is then reset.}%
+}%
     \label{fig:izhikevich}%
@@
 %----------------------------%
-In this task, the input consists of a stream of \emph{spikes} or \emph{events}, a property shared in the signal obtained from an event-based camera or from a neurobiological recording of the activity of single units.
-Formally, this can be defined as a list of tuples, each tuple representing the neural address and timestamp.
-We denote this list as $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N}$ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{Left-Top} for an illustration).
+In a neurobiological recording or in the sensory signal obtained from an event-based camera, the input consists of a stream of \emph{spikes} or \emph{events}.
+This can be formalized as a list of neural addresses and timestamps tuples $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event in the list of events (\seeFig{izhikevich}-\textit{(Top-Left} for an illustration).
+Events are typically ordered by their time of occurrence.
 Each event has a time of occurrence $\timev_\arank$ and an associated address $\presynaddr_\arank$.
-Events are usually ordered by their time of occurrence.
-We define the address space $\presynaddrspace$, which consists of the set of possible addresses.
-In neurobiological spiking activity, this may be the identified set of recorded neurons.
-For neuromorphic hardware like the output of a DVS or our task, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$, where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities encoded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels.
-Thus, each address $\presynaddr_\arank$ is typically in the form $(\polev_\arank, x_\arank, y_\arank)$ for event-based cameras.
+This defines an address space $\presynaddrspace$ which consists of the set of possible addresses.
+In a neurobiological recording, this can be the identified set of presynaptic neurons.
+For neuromorphic hardware, this can be defined as $[1, \Npol] \times [1, \Nx]\times[1, \Ny] \subset \mathbb{N}^3$ where $\Npol$ is the number of polarities ($\Npol=2$ for the ON and OFF polarities coded in event-based cameras) and $(\Nx, \Ny)$ is the height and width of the image in pixels.
+As such, an address $\presynaddr_\arank$ is typically in the form $(\polev_\arank, x_\arank, y_\arank)$ for event-based cameras.
 
-In the HD-SNN model, neurons $\postsynaddr \in \postsynaddrspace$ are connected to presynaptic afferent neurons from $\presynaddrspace$ using realistic synapses.
-In biology, a single cortical neuron typically has several thousand synapses.
-Each synapse can be defined by its synaptic weight and its delay, that is, the time it takes for a spike to travel from the soma of the presynaptic neuron to the soma of the postsynaptic neuron.
-Note that a neuron can contact another afferent neuron with different delays through different synaptic connections.
-By scanning all postsynaptic neurons $\postsynaddr$, we may thus define the full set of $\Nsyn$ synapses, as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated with a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
-This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{Right} for an illustration of the connectivity of a neuron with synaptic weights and delays).
-% A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$, but also by the set of possible delays.
+In the HD-SNN model, each neuron $\postsynaddr \in \postsynaddrspace$ connects to presynaptic afferent neurons from $\presynaddrspace$.
+In biology, a single cortical neuron has generally several thousands of synapses.
+Each synapse may be defined by its synaptic weight and its delay, that is, the time it takes for one spike to travel from the presynaptic neuron's soma to that of the postsynaptic neuron.
+A postsynaptic neuron $\postsynaddr \in \postsynaddrspace$ is then described by the synaptic weights connecting it to a presynaptic afferent from $\presynaddrspace$ but also by the set of possible delays.
+Note that a neuron may contact the same afferent neuron with different delays through different synaptic connections.
+Scanning all neurons $\postsynaddr$, we thus define the full set of $\Nsyn$ synapses,
+as $\synapse = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn)\}_{\ranksyn \in [1,\Nsyn]}$, where each synapse is associated to a presynaptic address $\presynaddr_\ranksyn$, a postsynaptic address $\postsynaddr_\ranksyn$, a weight $\synapticweight_\ranksyn$, and a delay $\synapticdelay_\ranksyn$.
+This defines the full connectivity of the HD-SNN model (\seeFig{izhikevich}-\textit{(Right)} for an illustration of the connectivity of one neuron with synaptic weights and delays).
 
-Of interest is to define the emitting field of a presynaptic neuron $\synapse_\presynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$, or also the receptive field of a postsynaptic neuron $\synapse^\postsynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$.
-In particular, when driven by a stream of spikes $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$, each incoming spike is multiplexed by the synapses of the receptive field $\synapse^\postsynaddr$ of postsynaptic neuron $\postsynaddr$.
-This results in a weighted event stream (\seeFig{izhikevich}-\textit{Left-Middle}) for each postsynaptic neuron $\postsynaddr$:
+Of interest is to define the receptive field of a postsynaptic neuron $\synapse^\postsynaddr =  \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \postsynaddr_\ranksyn=\postsynaddr\}_{\ranksyn \in [1,\Nsyn]}  \subset \synapse$, or the emitting field of a presynaptic neuron $\synapse_\presynaddr = \{(\presynaddr_\ranksyn, \postsynaddr_\ranksyn, \synapticweight_\ranksyn, \synapticdelay_\ranksyn) \| \presynaddr_\ranksyn=\presynaddr\}_{\ranksyn \in [1,\Nsyn]} \subset \synapse$.
+As a consequence, a postsynaptic neuron $\postsynaddr$ receives an event stream which is multiplexed by the synapses of its receptive field.
+It results in a weighted event stream (\seeFig{izhikevich}-\textit{(Middle-Left}) for each  postsynaptic neuron $\postsynaddr$:
 %
@@
 In biology, this new stream of events is naturally ordered in time as events reach the soma of postsynaptic neurons.
-In simulations, however, it should be properly reordered.
-Once transformed by the synaptic connectivity, this weighted event stream may be integrated, for instance as the membrane potential of a Leaky-Integrate-and-Fire neuron (\seeFig{izhikevich}-\textit{Left-Bottom}), yet the activation function of the HD-SNN neurons can be selected from the full range of spiking neuron response functions.
-Importantly, this activation function has to be such that when postsynaptic neurons are activated at their soma by a specific spatiotemporal motif imprinted in the synaptic set, and such that these spikes converge at the soma in a synchronous manner, the discharge probability should increase.
-In this subsection, we have briefly defined the HD-SNN model in all generality (see~\citep{Perrinet23ICANN} for a more specific description and treatment), and in the next subsection we describe an implementation of our model adapted to the motion detection task.
+However, it should be properly reordered in simulations.
+Crucially, when postsynaptic neurons are activated on their soma by a specific spatio-temporal motif which is imprinted in the set of synapses, the discharge probability will increase, notably when these spikes converge on the soma in a synchronous manner while the activation function of the neurons of the HD-SNN can be selected among the whole range of spiking neuron response functions (\seeFig{izhikevich}).
+We defined the HD-SNN model in this subsection in all generality, and in the next subsection, we describe an implementation of our model adapted for the motion detection task.
 %
-% \subsection{Application of the HD-SNN model to the motion detection task}
-\subsection{Application of HD-SNN to motion detection}
+\subsection{Application of the HD-SNN model to the motion detection task}
 %
-In fact, it is possible to adapt the HD-SNN model specifically for common computer vision tasks.
-First, neural addresses are defined to represent the range of possible positions and polarities.
-In order to simulate such event-based computations on standard CPU- or GPU-based computers and to benefit from parallel computing acceleration, we transform this event-based representation into a dense representation.
-Indeed, by discretizing time, we transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.
-The values are, by definition, equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
-Similarly, one may discretize the connectivity of the HD-SNN model defined above.
-The longest synaptic delay defines the depth $\Ktime$ of the kernel, so that all possible delays associated with the different presynaptic addresses are represented.
-In particular, for each class $\class$ of the supervision task, the entire synaptic set can be represented as a kernel, which is represented by the dense matrix $\kernel$ of size $(\Nclass, \Npol, \Ktime, \Kx, \Ky)$, where $\Nclass$ is the number of classes, $\Ktime$ the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
-To keep the analogy with the HD-SNN model, $\kernel$ gives the synaptic set that defines the weights as a function of class $\class$, polarity, synaptic delay and relative position: $\forall {\polev \in [1, \Npol], \delta_\timev \in [1, \Ktime], \delta_x \in [1, \Kx], \delta_y \in [1, \Ky]}, \; \kernel(\class, \polev, \synapticdelay_\timev, \delta_x, \delta_y) = \synapticweight$.
-In our simulations, we define as many classes as the number of motions (directions and velocities): $\Nclass = 12 \times 3$ and set the size of the model's kernel to $(\Nclass, \Npol, \Ktime, \Kx, \Ky) = (36, 2, 21, 17, 17)$.
-Such a kernel defines a dense representation of the full synaptic set.
-%  (that is, $436968$ parameters)
+It is possible to define a specific implementation of this model in order to adapt it to common tasks in computer vision.
+In particular, from the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation.
+Indeed, by discretizing time, we can transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.
+The values are by definition equal to zero, except when events occur: $\forall \arank \in [1,\numevent]$, $A(\polev_\arank, t_\arank, x_\arank, y_\arank)=1$.
+First, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:kernel_b} corresponds to a temporal convolution that transforms the input $A$ using one kernel per postsynaptic neuron~\citep{grimaldi_learning_2022}.
+To take advantage of the position invariance observed in images and exploited in convolutional neural networks, we can further assume that synaptic motifs should be similar across different positions, so we can define a spatio-temporal convolutional operator.
 
-Then, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:stream_b} corresponds to a weighted reordering of the input $A$ with each kernels and positions assigned to the postsynaptic neurons~\citep{grimaldi_learning_2022}.
-Let's define \emph{evidence} as the logit of a probability, that is, the inverse sigmoid of that probability.
-By this definition of evidence, logistic regression takes advantage of the fact that if different independent observations (here, the estimated motion at different spatial locations and timings) share a common cause (here, the rigid local motion of the image on the receptive field), then an optimal estimate of the evidence of this motion is the sum of the evidences from the independent sources.
-Interpreting the weights of the kernel as evidences (also called factors in logistic regression), we may therefore define the activity $B$ of post-synaptic neurons as the integration of this activity in each voxel and for each channel $\class$ in order to infer the evidence of each motion:
+Therefore, if we make the approximation at any given time that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
+The longest synaptic delay defines the depth $\Ktime$ of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented.
+In particular, the whole synaptic set can be represented as one kernel for each class $\class$ of the supervision task as the dense matrix $\kernel_\class$ of size $(\Npol, \Ktime, \Kx, \Ky)$, where  $\Ktime$ is the number of delays and $\Kx$ and $\Ky$ are the number of pixels in both spatial dimensions.
+To keep an analogy with the HD-SNN model, for neuron $\postsynaddr$ of position $(x_\postsynaddr, y_\postsynaddr)$ and class $\class$, $\kernel_\class$ gives the weights as a function of relative addresses and synaptic delays: $\forall {\delta_x \in [1, \Kx], \delta_y \in [1, \Ky], \delta_\timev \in [1, \Ktime], \polev \in [1, \Npol]}, \; \kernel_\class(\polev, \delta_x, \delta_y, \synapticdelay_\timev) = \synapticweight$.
+The connectivity of neuron $\postsynaddr$ is defined locally, around it position $(x_\postsynaddr, y_\postsynaddr)$.
+Using this dense representation, the processing of $A$ by the model can be written as: %
 %
@@
 \forall x, y, \timev, \; \;
-B(\class, \timev, x, y)
-= \sum_{p, \synapticdelay_\timev, \delta_x, \delta_y} \kernel(\class, p, \synapticdelay_\timev, \delta_x, \delta_y) \cdot A(p, \timev-\synapticdelay_\timev, x - \delta_x, y - \delta_y)
+\mathcal{C}_\class(x, y, \timev)
+= \sum_{p, \delta_x, \delta_y, \synapticdelay} \kernel_\class(p, \delta_x, \delta_y, \synapticdelay) \cdot A(p, x - \delta_x, y - \delta_y, \timev-\synapticdelay)
 \end{equation}
@@
 %
-where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses within a kernel and $\delta_\timev$ is the synaptic delay.
-In this formulation, we recognize that it takes advantage of the position invariance observed in images and exploited in CNNs.
-Here, we further assume that synaptic motifs should be similar across different times as defined in the temporal convolution.
-As a consequence, this defines a 3D, spatiotemporal convolutional operator, in which the layers of neurons assigned to specific kernels form channels.
-Using this dense representation, the model's processing of the input $A$ can be written as layer-wise convolution: $B = \kernel \ast A$ (\seeFig{model} for an illustration).
-A similar type of spatiotemporal filtering was used as a preprocessing stage for an existing pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
-In addition,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm that implements a motion estimation task.
-By assuming locally a constant motion, the authors assume that the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the motion.
-For convenience, the connectivity of the neuron $\postsynaddr$ is defined locally around its position $(x_\postsynaddr, y_\postsynaddr)$.
-Furthermore, it is important to consider that in order to adhere to the limitations of causal computation using biologically realistic neurons, synaptic delays are assigned positive values.
-This ensures that only past information contributes to the inference made at the present moment.
-In practical implementation, the kernels are temporally shifted so that the inference at the present time is solely influenced by past information.
-This temporal shift occurs after a duration equivalent to the depth of the kernel, denoted as $\Ktime$.
-The well-known convolution defines a differentiable measure, which is very efficiently implemented for GPUs, and which we will use to detect the motion direction in the event stream.
-% Therefore, assuming that a neuron integrates only in the temporal window given by its variety of synaptic delays, the integration of the spike train can be formalized by a 3D convolution operation.
-Such a method contrasts with classical methods for delay learning, which explicitly manipulate the delay as a variable and which are not directly differentiable~\citep{nadafian_bio-plausible_2020}.
+where $\delta_x$ and $\delta_y$ are the relative addresses of the synapses inside a kernel and $\delta_\timev$ is the synaptic delay. 
+This shows that $\mathcal{C}_\class$ is the result of a spatio-temporal convolution of the dense representation of the event stream with the dense kernels formed by the set of synapses:  $\mathcal{C}_\class = \kernel_\class \ast A$ (\seeFig{model} for an illustration).
+Note that to remain within the framework of a causal calculation, the kernels are shifted in time such that only past information gives an answer at the present time.
+This well-known computation defines a differentiable measure which is very efficiently implemented for GPUs and which we will use for learning the classification of different motifs in the event stream.
+A similar type of spatio-temporal filtering is used as a pre-processing stage for a pattern recognition algorithm~\citep{ghosh_spatiotemporal_2019}.
+Also,~\citet{sekikawa_constant_2018} developed an efficient 3D convolutional algorithm which implements a motion estimation task.
+By assuming a locally constant velocity, the authors assume the 3D kernel can be decomposed into a 2D kernel representing the shapes and a 3D kernel representing the velocity.
 
-Keeping the analogy with spiking neurons, the analog activity $B$ represents the integration of synaptic activity, and we will now try to define the detection of motion using the spatiotemporal kernels.
-Since we know that at each instant, there may be different motions, we will define the activation function of our model as a sigmoid function that implements a form of Multinomial Logistic Regression (MLR).
-In our MLR model, a probability value for each class (i.e., each direction of motion) is predicted for each position $x, y$ and time $\timev$ as a sigmoid function $\sigma(\beta) = \frac {1} {1 + \exp(-\beta)}$ of the linear combination of the list of events.
-Formally, using the kernels, the input raster plot is transformed into a probability with the following formula:
+Here, we keep the analogy with spiking neurons and we try to observe the detection of spiking motifs using the spatio-temporal kernels.
+If so, this precise spatio-temporal patterns prove to be of interest for neural computations and one can infer that biological neurons make use of this information as well.
+% \note{plus de biblio sur la motion detection/segmentation ou optique flow avec les events? (reviewer 2)}
+The 3D convolution represents the linear integration of the spike train as the linear input to the neuron.
+Then, the activation function of our model is a softmax function implementing a form of Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022}, in analogy to a spiking Winner-Takes-All network~\citep{nessler_bayesian_2013}.
+In our MLR model, a probability value for each class (i.e.
+each motion direction) is predicted for each position $x, y$ and time $\timev$ as a softmax function of the linear combination of the list of events.
+% on the basal dendrite of a neuron $\postsynaddr$ in association to a specific class.
+The linear combination can be defined by a set of synapses $\synapse^\postsynaddr$ as described in the HD-SNN model.
+Formally, using the kernels, it transforms the input raster plot into a probability with the following formula:
 %
+\begin{linenomath*}
 \begin{equation}\label{eq:mlr}
-\forall x, y, \timev, \forall \class \in [1, \Nclass], \; \;
+    \forall x, y, \timev, \forall \class \in [1, N_\class], \; \;
 Pr(k=\class \; \vert \;  x, y, \timev) =
-\sigma (B(\class, \timev, x, y) +\bias_\class) 
+\frac {\exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class) }{Z(x, y, \timev)}
 \end{equation} 
+\end{linenomath*}
 %
-where $\bias_\class$ is a scalar representing the bias associated with the class $\class$.
-In particular, we anticipate that certain specific patterns could result in closely synchronized outputs when they are integrated within the basal dendritic tree, consequently leading to heightened postsynaptic activity.
-By utilizing this analog representation of the evidence for each potential motion at every moment, we can progressively increase the likelihood of generating an output spike.
-To determine the spiking output, we establish a firing threshold.
-We computed this threshold to ensure that neurons, on average, generate one spike per second.
-Therefore, the spiking output of the model corresponds to the motions in space and time that possess the highest probability.
+where $Z(x, y, \timev)=\sum_{\class \in [1, N_\class]} \exp  (\mathcal{C}_\class(x, y, \timev) +\bias_\class)$ is the partition function and  $\bias_\class$ is the bias linked to class $\class$. 
+In particular, we expect that some specific motifs may become tightly synchronized as they reach the basal dendritic tree, leading to a high postsynaptic activity which makes it progressively more likely to generate an output spike.
+The spiking output of the model corresponds to an event with the highest probability class.
 
-Now that this general framework has been described formally, we may include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}.
-Note that the general framework is an extension to that presented in~\cite{grimaldi_learning_2022}, in particular by including a more complex task to solve, the deeper analysis of the results, and these novel neuroscience-inspired heuristics.
-First, to avoid introducing biases in the directions which may be learned, we apply a circular mask to the spatial dimensions of the kernels.
-We also included a prior in the selectable motions, as there is a prior for slow speeds in natural scenes~\citep{vacher_bayesian_2018}.
-Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration).
-%This is consistent with neuroscientific principles because, due to the limits of conduction delay, the synaptic delay is physically related to the distance between the presynaptic and postsynaptic neurons.
-Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities.
-This is due to the fact that our whitened images have a relative symmetry in the luminance profiles, that is, that an image with inverted contrast is indistinguishable from a standard one.
-Since this arrangement of polarities is independent of motion, we added a mechanism that collects the linear values for the movie and the movie with the ON and OFF cells flipped, keeping only the maximum value for each voxel.
+Now that this general framework has been explained, we can add some heuristics, based on neuroscientific observations, to constrain our model and its strategies for solving the task that is used to solve the ecological task described in section~\ref{sec:task}.
+Note that the general framework is similar to that presented in~\cite{grimaldi_learning_2022} and that, apart from the more complex task to solve and a deeper analysis of the results, these novel heuristics inspired by neuroscience are the main methodological differences with that previous study.
+First, in our simulations, we set the size of the kernels to $(21, 17, 17)$ and define as many classes as the number of motions (directions and velocities): $N_\class = 12 \times 3$.
+To avoid introducing biases in the directions, we apply a circular mask to the spatial dimensions of the kernels.
+Furthermore, we included a prior in the motions that could be selected, as there is a prior in natural scenes for slow speeds~\citep{vacher_bayesian_2018}.
+Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we thus apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration).
+This is consistent with neuroscientific principles because, due to the limits of the conduction delay along horizontal connections, the synaptic delay is related, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one.
+Second, we observed that moving images produced trajectories of ON and OFF spikes, and that these were present in both polarities arrangements.
+This is due to the fact that our whitened images have a symmetry in the luminance profiles, and an image with inverted contrast is indistinguishable.
+Since this arrangement of polarities is independent of speed, we added a mechanism that collects the linear values for the movie and the movie with the ON and OFF cells flipped, keeping only the maximum value for each voxel.
 This is similar to the computation done for complex cells in primary visual cortex.
@@
     \includegraphics[width=\linewidth]{figures/HDSNN_conv.pdf}
+    % \includegraphics[width=\linewidth]{figures/HD-SNN.png}
     \caption{
-        \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot a 2D representation of the input event stream as a raster plot (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time).
-A spatiotemporal convolution is applied to the dense representation of the input with $2$ different convolution kernels (the green and orange kernels), which define the output channels.
+    % \textbf{- HD-SNN illustration.
+From left to right:} A 3D representation of the event stream as output of the DVS.
+A 3D convolution is applied, on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis.
+There are as many kernels as the possible number of motion classes in the recording.
+For clarity, only $4$ kernels are illustrated and some weights are not represented in the voxel grid.
+For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e.
+softmax function).
+The output of the MLR gives a probability for each class, associated to a specific kernel, to emit an output event, or spike.
+With an appropriate spiking mechanism, here a winner-takes-all, we obtain, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class.
+}
+    \textbf{Applying HD-SNN to the task of motion detection.} \textit{(Left)}~We plot here as a raster plot a 2D representation of the input event stream (showing ON spikes in red and OFF spikes in blue for each presynaptic address and time).
+A spatio-temporal convolution is applied to the dense representation of the input with $2$ distinct convolutional kernels (the green and orange kernels) that will define the output channels.
 The convolution is summed over the two polarities.
-Since we have two axes $X$ and $Y$ to represent the presynaptic addresses, like the pixel grid of a DVS, this results in a 3D convolution.
-Here we simplify the illustration to a 2D representation and to 2 possible classes (green and orange) associated to two different directions of motion.
-\textit{(Middle)}~For each position (address, time) one can compute the activation resulting from the convolution.
-The output of the convolution is processed by the nonlinearity of the MLR model (i.e., the sigmoid function).
+If you have two axes $X$ and $Y$ to represent the presynaptic addresses as for the pixel grid of a DVS, you obtain a 3D convolution.
+We restrict the illustration to a 2D representation and to 2 possible classes (green and orange) that are associated with different motion directions.
+\textit{(Middle)}~For each position (address, time), one can compute the activation resulting from the convolution.
+The output of the convolution is processed by the nonlinearity of the MLR model (i.e., the softmax function).
 The output of the MLR gives a probability for each class associated with a particular kernel (colored bars in the highlighted pixel).
-\textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated with thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated with a particular motion class.
-Note that the position of the output spikes does not systematically correspond to the position of the input spikes, but only when enough evidence is reached.
-    }
+\textit{(Right)}~By adding a spiking mechanism, here a winner-takes-all associated to a thresholding, we obtain as output of the HD-SNN model a new spike train with the different spikes associated to a specific motion class.
+Note that the position of the output spikes does not systematically correspond to the position of the input spikes but only when enough evidence is obtained.}
     \label{fig:model}
@@
 %
-Since the model is fully differentiable, we can now implement a supervised learning rule to learn the weights of the model's kernel.
-This rule was implemented using the binary input events as inputs and the corresponding motion's labels as the desired output.
-The loss function of the MLR model is the binary cross entropy of the output of the classification layer knowing the ground truth.
-The labels were defined at each time point as a one-hot encoding of the current motion in the channel corresponding to the current motion, and applied for all positions.
-Note that in this context, the label is known, but the position of visual features is not, mainly due to the sparse spatial content of natural images.
-However, the supervised optimization of this MLR model adjusts the weights of the kernels.
-As a result, we observed that the error is only propagated back to the spatial locations of these most active cells.
+Since the model is fully differentiable, we can now implement a supervised learning rule.
+This rule was implemented using the binary input events as inputs and the corresponding motion direction labels as the desired output.
+The loss function of the MLR model is the binary cross entropy at the output of the classification layer.
+The labels were defined at each time point as a one-hot encoding of the current motion in the channel corresponding to the current motion for all positions.
+Note that in this context, the label is known but the position is not, mainly due to the sparse spatial content of natural images.
+However, the supervised optimization of this MLR model will involve adjusting the weights of the kernels.
+As a result, the error is only propagated back to the spatial locations of these most active cells.
 This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation.
-Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{10}$ movies, each of size $200 \times 128 \times 128$, a learning rate of $10^{-5}$ and $100$ epochs).
+Simulations are performed with the \note{PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$)}.
 We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
 
-% TODO citer Levy2021 "Communication consumes 35 times more energy than computation in the human cortex" 
 Finally, the output of the MLR model is an event-based representation that predicts the probability of each motion at each position and time.
 Such an output provides a form of optical flow that can be exploited for non-rigid motion, but we have defined here, for simplicity, an evaluation method that applies to our full-field motion task.
-Using the properties of logistic regression, by taking the mean evidence represented in the output given by the model at all positions for any given time, and using the sigmoid function, we can derive each motion's probability at that time.
-This allows one to calculate the accuracy as the percentage of times the motion is accurately predicted at any given time step.
+We have shown above that if different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities.
+Thus, by taking the mean logit of the probability of the output given by the model at all positions for any given time, we can calculate the probability of the output at this time.
+This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted).
 These calculations are performed on a different input data set than the one used in the training or validation steps.
-The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN} (see Data Availability).
+The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}.
 %
@@
     \caption{
-    	Representation of the weights for $8$ directions for a single speed (among the $12 \times 3$ different kernels of the model) as learned on the dataset of naturalistic scenes.
-The directions are shown as red arrows in the left insets.
-The spatiotemporal kernels are shown as slices of spatial weights at different delays.
-Delays vary along the horizontal axis from the far right (delay of one step) to the left (up to a delay of $12$ steps, the remaining synapses being not represented).
-Each image corresponds to the weights at a given delay, with excitatory and inhibitory weights in warm and cold colors, respectively.
-Due to the symmetry between the ON and OFF event streams, we observed that the kernels for the OFF polarities are very similar and are not shown here.	Different kernels are selective for the different motion directions, and we observe a slight orientation preference perpendicular to the respective direction for all kernels.
+    	Representation of the weights for $3$ of the $32$ different learned kernels of the model as learned on natural scenes.
+Each pair of line correspond to the OFF and ON polarities respectively, with excitatory weights in warm colors.
+Delays are represented in the horizontal axis from right (zero delay) to the left (delay of $11$ steps).
+Because of the symmetry observed between the ON and OFF event streams, we observed that kernels are very similar for the ON polarities.
+These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
+%	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown.
+	Different kernels are selective to the different motion directions and we observe some level of orientation selectivity, where ON and OFF subfields organized in a push-pull organization. 
+%        One sees positive (excitatory) coefficients for the specific direction of motion but also negative (inhibitory) coefficients for other directions.
 	}
@@
 %
-Once our model has been trained, we can begin by examining the learned weights for the various motions (\seeFig{kernels}).
-Notably, when we track each spatial motif from the shortest delay (on the right) to the longest delay (on the left), we observe that the cells exhibit highly localized selectivity and their preferences are conveyed along linear trajectories in the space-delay domain.
-When focusing on the positive weights, we notice a pronounced selectivity along specific motion axes for each kernel, and these directions correspond closely to the associated motion's physical direction in visual space.
-For instance, the first kernel demonstrates a robust preference for downward motion.
-The negative weights are symmetrically arranged around these positive weights, forming a center-surround profile that enhances the response.
-We also observed a strong dependence between the weights reaching the ON polarities and those reaching the OFF polarities.
-In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative for the other.
-This property is due to the way events are generated and the fact that the luminance cannot increase and decrease at the same time.
-Interestingly, the relative organization of the receptive fields that we observe is in quadrature of phase and follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results obtained after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
-Finally, we observe that these receptive fields show also a relative selectivity to the orientation perpendicular to motion, similarly to what is found for neurons in cortical area MT which is known to be selective to visual motions~\citep{deangelis_functional_1999}.
-This reflects the way events are generated and in particular the so-called aperture problem which implies that a line moving along its axis would generate no change in luminance and therefore generate no event~\citep{perrinet_motion-based_2012}.
+After training our model, we first analyze the weights learned for the different neurons (\seeFig{kernels}). 
+We first observed a high dependence between the weights reaching the ON polarities and that reaching the OFF polarities.
+In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative in the other.
+This property comes from the way the events are generated and that the luminance can not at the same time increase and decrease.
+%ith the negative weights, one can observe an anti-selectivity for directions that do not correspond to the motion to which the kernel is selective to.
+We also observe that these cells show an orientation selectivity, similar to that observed in MT neurons~\citep{deangelis_functional_1999}.
+Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results observed after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
+Focusing on the positive weights, a strong selectivity is observed along specific axes of motion for each of the different kernels.
+These directions can be easily associated to the direction of motion controlled in the natural images.
+For instance, the first kernel shows a strong selectivity to horizontal motion directions.
+%
+%This qualitative look at the 3D kernels allows the reader to infer for the $8$ different motion directions used to generate our synthetic event streams.
 
-If we now widen our focus on the interpretation of these kernels in terms of spatiotemporal motifs embedded in the event stream, these show a prototypical anisotropic profile~\citep{kaplan_anisotropic_2013} adapted to motion detection.
-In~\citep{grimaldi_robust_2022}, a link was drawn between event-based MLR training and Hebbian learning, allowing to say that the present model learns its weights according to a presynaptic activity associated to the different motion directions.
-Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatiotemporal spiking motif.
-Each voxel in the 3D kernels defines a specific property by associating a weight to a position and a delay.
-Consequently, our model is able to detect precise spatiotemporal motifs embedded in the spike train and associated to the different motion directions.
-Note that as the delays become larger, two effects can be remarked.
-First, coefficients become lower which is consistent with the fact that trajectories are defined in a piece wise fashion, such that this decrease provides with an optimal integration considering the gradual diminishing of evidence as time progresses~\citep{Pasturel2020}.
-Second, coefficients become less localized compared to the kernel's spatial profile at short delays, consistent with the average diffusion of information included in the generative model and with the diffusion introduced in motion-based prediction models~\citep{perrinet_motion-based_2012,khoei_flash-lag_2017}.
-%This conic shape along the motion's trajectory highlights a loss of precision for longer delays, i.e.
+If one focuses on the interpretation of these kernels in terms of spatio-temporal motifs embedded in the event stream, it can lead to interesting outcomes.
+In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions.
+Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatio-temporal spike motif.
+Each voxel in the 3D kernels defines a specific timestamp and a specific address.
+Consequently, our model is able to detect precise spatio-temporal motifs embedded in the spike train and associated to the different motion directions.
+The cone shape for the positive weights distribution highlights a loss of precision for longer delays, i.e.
 events away in the past.
+For the directions not coherent to the class of a training sample, an anti-Hebbian learning is also observed through the negative weights in the kernels of Fig.~\ref{fig:kernels}.
+
 %
-%
+% \note{kernel size has to be adapted to the characteristics of motion see Samonds %~\citep{samonds}
+% say that we have tested }
+% \note{tester la meme expe faite par baudot et modelisée par kremkow en montrant des raster plots du modele avec NI versus Gratings }
+
 %%%-----------------------------------------------------------------
@@
 %%%-----------------------------------------------------------------
+To test our model, we will quantify its ability to categorize different motions.
+Before applying the model on natural images, we will first test the model on simpler, parameterized stimuli.
+In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
+In particular, we will set the spatial size and duration similarly to the motion task defined above.
+This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions $
+\speed_\kernelind = 
+  ( 
+    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
+    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
+  )
+$.
+For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content to provide with some naturalistic variability.
+This method provides a rich dataset of textured movies for which we know the ground truth for motion.
 %%%-----------------------------------------------------------------
@@
     \centering
-    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
-    % TODO : show prototypes of textures on top
-    \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from a grating (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to grating (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)).
-Note that these accuracies are computed both in the case where the orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) and in the generic case where the orientation is independent of direction ('aperture').
-    }
+    \includegraphics[width=0.99\linewidth]{figures/2022-09-27_MotionDetection_motion_clouds.png}%motion_clouds.pdf} % 2022-09-27_MotionDetection_motion_clouds.png}%
+    \caption{{\bf Role of stimulus parameters in the motion detection accuracy.} Accuracy as a function of {\bf (a)} the mean spatial frequency, {\bf (b)} the bandwidth in orientation: from a grating-like (left) to isotropic textures (right)), {\bf (c)} the bandwidth in speed, from a rigid motion (left) to independent frames (right).
+Note that these accuracy is computed both in the case where orientation of the synthetic texture is necessarily perpendicular to the motion (no aperture) or arbitrary (aperture), showing that accuracy decreases in the latter case.}
     \label{fig:motion_clouds}
@@
 %%%-----------------------------------------------------------------
-We also wanted to assess the influence of spatiotemporal parameters of the stimuli on the performance of the model.
-For this, we will quantify its ability to categorize different motions, i.e.
-on event streams for which the ground truth motion is known at each instant.
-Before applying the model to natural scenes that we defined in the motion detection task, we first test the model on simpler, parameterized stimuli.
-For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control, among other parameters, the motion (\seeFig{motion_clouds})~\citep{Leon2012,vacher_bayesian_2018}.
-These stimuli allowed us to control various parameters, including motion, spatial extent, and duration.
-By matching the spatial and temporal characteristics of the generated movies with those of the motion task mentioned earlier, we created a range of textures featuring different spatial properties and motions.
-This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $12$ directions and $3$ speeds.
-For each motion, we also varied the texture parameters, such as mean and variance of orientation or spatial frequency content, to provide some naturalistic variability.
-This method provides a rich dataset of textured movies for which we know the ground truth for the motion.
 
-We observe some interesting facts.
-First, as we change the mean spatial frequency of the texture, we observe a broadly tuned response in accuracy.
-This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006,Perrinet2007} and reveals the most informative scales learned by our model.
-In particular, the accuracy is better for a grating-like stimulus than for a large one (which qualitatively resembles a more textured stimulus), reminiscent of the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}.
+We plot here main axis of interests.
+First, as we change the mean spatial frequency of the texture, we observe a monotonous decrease in accuracy.
+This comes as a similar trend as that observed in the primary visual areas~\citep{priebe_tuning_2006}.
+%\note{we should really do something more serious here...}
+Notably, the accuracy is better for a large spatial frequency bandwidth (which qualitatively resemble a more textured stimulus) than for a grating-like stimulation, reminiscent to the behavioral response of humans' eye movements to such stimuli~\citep{simoncini_more_2012}.
 Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth.
-When the stimulus is grating-like and the orientation is arbitrary with respect to the direction of motion, the system faces the aperture problem and experiences a decrease in accuracy.
+When the stimulus is grating-like and that the orientation is arbitrary with respect to the direction of motion, the system is faced with the aperture problem and see a decrease of accuracy.
 This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion.
 Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter.
-This shows a progressive decrease in accuracy, similar to that observed in the amplitude of human eye movements~\citep{mansour_pour_speed_2018}.
+This shows a progressive decrease in accuracy, similar to that observed in the amplitude of humans' eye movements~\citep{mansour_pour_speed_2018} but also that accuracy is low for a rigid motion which lacks variability.
 %
-%
-\subsection{Accuracy versus efficiency trade-off}% for the motion detection task}
+\subsection{Accuracy efficiency trade-off}% for the motion detection task}
 %%%-----------------------------------------------------------------
-After training our MLR model, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection.
-The accuracy computed on the test set was approximately $89\%$ (with a chance level of $1 / 12/ 3 \approx 2\%$).
-%Moreover, we observed that the classification performance obtained for the model using time surfaces as in~\citep{grimaldi_robust_2022} gives similar results as our method.
+Once our MLR is trained, we obtain spatio-temporal kernels corresponding to the weights associated to the heterogeneous delays of our layer of spiking neurons and which may be used for detection.
+We observed that the distribution of the kernels' weights is sparse, with most values near zero.
+As shown in the formalization of our event-based model, the computational cost of our model if implemented on a neuromorphic chip would be dominated by the number of spikes times the number of synapses.
+Indeed, the computations are dominated by the convolution operation.
+In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel.
+If the support of information is sparse, then computations can be performed only on those events.
+Also, if we set some weights of the kernels to zero, then the sum can be skipped for those addresses.
+Knowing the sparseness of the input, the total number of computations thus scales with the number of nonzero synaptic weights.
 
-We observed that the distribution of the kernel's weights is sparse, with most values close to zero (\seeFig{kernels}).
-As shown in the formalization of our event-based model, the computational cost of our model, if implemented on a neuromorphic chip, would be dominated by the computations used for the convolution operation.
-In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel.
-But if the information support is sparse, then computations can be now performed only on those events.
-Specifically, if we set some weights of the kernels to zero, then the additive operation in the convolution for those addresses can be dropped.
-As a consequence, computations will be performed only on those events which were multiplexed by the pruned connectivity matrix.
-Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of non-zero synaptic weights.
+To assess the robustness of the classification as a function of the computational load, we will prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold.
+In Fig.~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations, or active weights, per decision for each neuron of the layer.
+As a comparison and to account for the gain in performance by using heterogeneous delays, we provide the accuracy obtained with a MLR model using 2D time surface (in red) as in~\citep{grimaldi_robust_2022}.
+This latter method is based on delays from the last recorded events and uses fewer computations (in our case $15\times15$) than the dense 3D kernels without any pruning ($15\times15\times8$).
+While less computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
 
-In order to evaluate the resilience of the classification performance with respect to computational load, we adopt a pruning approach, where we remove weights in $\kernel$ that fall below a specified threshold.
-The accuracy of classification is then plotted as a function of the relative number of computations or active weights per decision for each neuron in the layer (refer to Fig.~\ref{fig:accuracy}).
-To provide a basis for comparison and to account for the benefits of utilizing variable delays, we present the accuracy achieved by an MLR model employing a shortening strategy.
-This strategy involves adjusting the temporal width by selecting only the weights associated with the shortest delays.
-In comparison to the inference performed using the complete 3D kernels without any pruning ($21\times17\times17$), both approaches demonstrate a reduction in computational requirements as indicated by the number of non-zero weights.
-
-By selectively setting certain weights to zero, we observe that the accuracy's evolution, as a function of the logarithmic percentage of active weights, aligns well with a sigmoid curve for both pruning and shortening strategies.
-The shortening strategy (depicted in orange) demonstrates a rapid decline in accuracy, reaching half-saturation when approximately one-third of the weights are removed.
-On the other hand, the pruning strategy (shown in blue) exhibits a different behavior.
-It reaches half-saturation when the ratio of active weights is approximately $3.2\times 10^{-2}$, corresponding to around 180 computations per decision in our specific scenario.
-In comparison to using the complete kernels, our method maintains accuracy close to its peak performance even when the number of computations is divided by a factor of up to around 31.
-This substantial reduction in computations showcases the robustness of the presented method.
+By pruning weights, we observe that the evolution of accuracy as a function of the log percentage of active weights fits well a sigmoid curve.
+Half-saturation level is reached at about $3.5\times 10^{-3}\%$ of active weights, corresponding in our setting to a total amount of $6$ computations per decision.
+Compared to the full kernels, the accuracy of our method is maintained to its top performances when dividing the number of computations by a factor up to about $200$.
+In this case, the number of computations is greatly reduced compared to~\citep{grimaldi_robust_2022}, thus demonstrating the efficiency of the presented method.
 %
@@
     \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
-    \caption{
-        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
+    \caption{Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
 The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
-As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about $250$ times fewer computations.
-        }
+As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about 25 times fewer computations.}
     \label{fig:accuracy}
@@
 %%%-----------------------------------------------------------------
-In this paper, we have introduced a generic heterogeneous delay SNN.
-It is trained using supervised learning on a visual motion detection task.
-We have shown how the model performs well compared to other event-based classification algorithms for this task.
-The learned model has a number of similarities with neurobiological and behavioral observations.
-The event-based computations of our method can be drastically reduced by pruning synapses, while maintaining robust classification performance.
-This shows that we can exploit the precise timing of spikes to improve the performance of neural computations.
+In this paper, we have introduced a generic SNN with heterogeneous delays and shown how it compares favorably to a state-of-the-art event-based classification algorithm for a visual motion detection task.
+The learned model shares a number of similarities with neurobiological anatomical observations, as well as with behavioral results.
+The event-based computations of our method can be drastically reduced by pruning synapses, while maintaining top classification performance.
+This shows that we can take advantage of the precise timing of spikes to improve the performance of neural computations.
 %
@@
 %%%-----------------------------------------------------------------
-The HD-SNN model was trained and evaluated on a naturalistic motion detection task with realistic eye movements.
-The model is defined such as to provide an optimal detection of  spatiotemporal motifs and learns kernels similar to those found in the visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
+The HD-SNN model that has been trained and evaluated on a complex motion detection task.
+The model has been defined to provide optimal detection of event-driven spatio-temporal motifs.
+We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in the early visual cortex~\cite{kremkow_push-pull_2016}.
+%We then demonstrated that the model shows similarities to biological responses.
 We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware.
@@
 
-We would like to highlight a few innovations in the contributions presented in this paper.
-First, while~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of spatiotemporal spiking motifs using a time-invariant logistic regression.
-Moreover, compared to classical CNN solutions, the parameters of this one-layered model (weights and delays) are explainable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif, where we define \emph{evidence} as the logit of the probability, that is, the inverse sigmoid of the probability.
-Another novelty is that the model learns simultaneously weights and delays.
-In contrast, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn at initialization and their values are frozen during learning.
+We would like to highlight a few innovations in the contributions that are presented in this paper.
+First, whereas~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, which we observed to be less efficient, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatio-temporal motifs.
+Moreover, in comparison to a representation with time surfaces, the weights are explicable as they directly inform on the logit (inverse sigmoid of the probability) of detection of each spatio-temporal spike motif.
+Another novelty is that the model learns the weights and the delays simultaneously.
+For example, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn and their values are frozen during learning.
 In addition, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified toy problems~\citep{gutig_tempotron_2006}.
-Another major contribution is to provide a model that is suitable for learning any kind of spatiotemporal spiking motif and that can be trained in a supervised manner by providing a dataset of supervised pairs.
+Another major contribution is to provide a model that is suitable for learning any kind of spatio-temporal spiking motifs and that can be trained in a supervised manner by providing a dataset of supervision pairs.
 Instead of relying on a careful description of the physical rules governing a task, e.g.
-the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}, this allows a more flexible definition of the model using this properly labeled dataset.
+the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}, this allows a more flexible definition of the model using this properly labelled dataset.
 %
@@
 We have identified a number of limitations of our model, which we will now discuss in detail.
-First, this implementation of the HD-SNN model is based on a discrete binning of time, which is not compatible with the continuous nature of biological time.
-We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, in particular to be able to use fast, differentiable three-dimensional convolutions.
-This is consistent with the relative robustness of other event-based frameworks~\citep{lagorce_hots_2017,grimaldi_robust_2022}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $4~\ms$ on the N-MNIST dataset~\citep{grimaldi_robust_2022}.
-It suggests the potential advantage of analytically including an additional precision term to the temporal value of input spikes.
-Such a mechanism may be implemented by the filtering implemented by the synaptic time constant of about $5~\ms$.
+First, the entire framework is based on a discrete binning of time, which is not compatible with the continuous nature of biological time.
+We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, to be able to use fast three-dimensional convolutions.
+We have tested the effect of the width of the time bin and shown that it has essentially no effect on the results presented in this paper.
+This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}.
+This suggests the possibility of analytically including a precision term in the temporal value of the input spikes.
+This mechanism may be implemented by the filtering implemented by the synaptic time constant of about $5~\ms$.
 Furthermore, it is possible to circumvent the need for time discretization by the use of a purely event-based scheme.
-In fact, it is possible to derive event-triggered computations of the continuous activity of the SNN~\citep{hanuschkin_general_2010} and thus to define a purely event-based framework.
+In fact, it is not necessary to compute the voltage traces between two spikes~\citep{hanuschkin_general_2010}.
+Thus, it is possible to define a purely event-based framework.
 Such an architecture could provide promising computational speedups.
+% \note{pourrait parler du poster bernstein?
+de nadafian?}
 
-Another limitation is that the model is purely feed-forward.
-Thus, the spikes generated by the postsynaptic neurons are based solely on the information contained in the classical receptive field.
-However, it is known that neurons in the same layer can interact with each other through lateral interactions, for example in V1, and that this can be the basis for more complex computational principles~\citep{chavane_revisiting_2022}.
+A further limitation is that the model is purely feed-forward.
+Thus, the spikes generated by the postsynaptic neurons are based solely on information contained in the classical receptive field.
+However, it is well known that neurons in the same layer can interact with each other using lateral interactions, for example in V1, and that this can be the basis for computational principles~\citep{chavane_revisiting_2022}.
 For example, the combination of neighboring orientations may contribute to image categorization~\citep{perrinet_edge_2015}.
-Furthermore, neural information may be modulated by feedback information, e.g.
+Furthermore, neural information is modulated by feedback information, e.g.
 to distinguish a figure from its background~\citep{roelfsema_early_2016}.
@@
 Currently, mainly due to our use of convolutions, it is not possible to implement these recurrent connections in our implementation (lateral or feedback).
-However, by inserting new spikes into the list of spikes reaching presynaptic addresses, the generic HD-SNN model is able to incorporate them.
-While this is theoretically possible, it must be properly tuned in practice so that these recurrent connections do not bring neuronal activity outside a homeostatic state (by extinction or explosion).
+However, by inserting new spikes into the list of spikes reaching presynaptic addresses, the generic model is able to incorporate them.
+While theoretically possible, this needs to be properly adjusted in practice so that these recurrent connections do not amplify neuronal activity outside a homeostatic state (by extinction or explosion).
+% \note{ dire que la prédiction permet de faire de la prediction et que ça explique les résultats de leBec}%~\citep{perrinet_motion-based_2012}
 
-Such recurrent activity would be essential for the implementation of predictive or anticipatory processes~\citep{Benvenuti2020}.
-This is essential in a neural system because it contains several delays that require temporal alignment~\citep{hogendoorn_predictive_2019}.
+Such recurrent activity would be essential for the implementation of predictive or anticipatory processes.
+This is essential in a neural system because it contains several different delays that require temporal alignment~\citep{hogendoorn_predictive_2019}.
 This has been modeled before to explain, for example, the flash-lag illusion~\citep{khoei_flash-lag_2017}.
-As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by motion, acceleration, jerk,~\ldots), and  knowing that ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
+%\note{Say that this could be implemented by properly setting the timing of the label}
+As mentioned previously, this could be implemented using generalized coordinates (i.e., variables such as position complemented by velocity, acceleration, jerk,~\ldots), and ``neurobiologically, using delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors''~\citep{perrinet_active_2014}.
 Our proposed model using heterogeneous delays provides an alternative and elegant implementation solution to this problem.
@@
 %%%-----------------------------------------------------------------
-%
+% \note{la tache ne met pas en avant les capacites des motifs hetero synaptiques =trop simple et peut se faire en HOTS}
 In defining our task, we emphasized that the generation of events depends on the spatial gradient in each image.
@@
 Taken together, these oriented edges form the contours of visual objects in the scene~\citep{koenderink_representation_1987}.
-Thus, there is an interdependence between motion information and orientation information within the event stream.
+Thus, there is an interdependence between the information about motion and the information about orientation within the event stream.
 It would be crucial to investigate this dependency further.
 This could be initiated by training the model on a dataset with labels that provide local orientation.
-Exploring this dependence will allow us to dissociate and integrate these two forms of visual information.
-In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (that is, the aperture problem).
+%\note{we can say that we tried it and that it works well}.
+Exploring this dependence will allow us to dissociate these two forms of visual information and enable us to integrate them.
+In particular, it will allow us to consider that the definition of motion is more accurate perpendicular to an oriented contour (aka the aperture problem).
 Thus, it will allow us to implement recurrent prediction rules, such as those identified to dissociate this problem~\citep{perrinet_motion-based_2012}.
@@
 The model is trained on a low-level local motion detection task, and one might wonder if it could be trained on higher-level tasks.
-An example of such a task would be depth estimation in the visual scene.
+An example of such a task would be the estimation of depth in the visual scene.
 There are several sources of information for depth estimation, such as binocular disparity or changes in texture or shading, but in our case motion parallax would be the most important cue~\citep{rogers_motion_1979}.
-This is because objects that are close to an observer move on the retina relatively faster than an object that is far away, and also because visual occlusions depend on depth order.
+This is because objects that are close to an observer will move relatively faster on the retina than an object that is far away, and also because visual occlusions are dependent on the depth order.
 Using this information, it is possible to segment objects and estimate their depth~\citep{yoonessi_contribution_2011}.
-However, this would first require the computation of the optic flow, i.e., the extension of the framework described here for a rigid full-field motion to a more general one where the motion may vary in the visual field.
-One possible implementation is to add a new layer to our model, analogous to the hierarchical organization which is prevalent in the visual cortex.
-This is theoretically possible by using the output of our model (which estimates motion in retinotopic space) as input to a new layer of neurons that would estimate motion in the visual field, including the depth dimension in the output supervision labels.
+However, this would require the computation of the optical flow first, i.e., the extension of the framework described here for a rigid full-field motion to a more general one where the motion may vary in the visual field.
+A possible implementation is therefore to add a new layer to our model, analogous to the hierarchical organization highlighted in the visual cortex.
+This is theoretically possible by using the output of our model (which estimates velocity in retinotopic space) as input to a new layer of neurons that would estimate velocity in the visual field, including the depth dimension in the output supervision labels.
 This could have direct and important applications, e.g.
@@
 
-In conclusion, the HD-SNN model that we have presented provides a way to efficiently process event-based signals.
-We have shown that we can train the model using a supervised rule that infers \emph{what} is the output label, but not \emph{where} it occurs.
-Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e.
-without any labelled data~\citep{barlow_unsupervised_1989}.
+In conclusion, the model that we have presented provides a way to process event-based signals in an efficient manner.
+We have shown that we can train the model semi-supervised, knowing \emph{what} output label, but not \emph{when} it occurs.
+Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e., without any labeled data~\citep{barlow_unsupervised_1989}.
 This type of learning is thought to be prevalent in the central nervous system and, assuming the signal is sparse~\citep{olshausen_emergence_1996}, one could extend these Hebbian sparse learning schemes to spikes~\citep{perrinet_emergence_2004, masquelier_competitive_2009}.
@@
 Indeed, there is a large literature showing that brain dynamics often organize into stereotyped sequences, such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}.
-These motifs are stereotyped and robust, as they can be activated following the same motif from day to day~\citep{haimerl_internal_2019}.
-In contrast to conventional methods of processing neurobiological data, such an event-based model would be able to answer key questions about the representation of information in neurobiological data, and it would open possibilities in the field of computational neuroscience.
-Furthermore, it would open possibilities in the field of machine learning, especially in computer vision, to address current key concerns such as robustness to attacks, scalability, interpretability, or energy consumption.
+These patterns are stereotyped and robust, as they can be activated in the same pattern from day to day~\citep{haimerl_internal_2019}.
+In contrast to conventional methods of processing neurobiological data, such an event-based model would be able to answer key questions about the representation of information in neurobiological data, and it would open up possibilities in the field of computational neuroscience.
+Furthermore, it would open up possibilities in the field of machine learning, especially in computer vision, to address current key concerns such as robustness to attacks, scalability, interpretability, or energy consumption.
 %

### 2023-07-09 Grimaldi-etal-BiolCybernetics.tex
Revert "Merge branch 'master' of https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN"
This reverts commit 84309fb14be9639f5ca19819823a52dc6e18b71a, reversing
changes made to e65e05f06cdbe320bd68dc4815d1c6afaca41bdc.
Update Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

--- 2023-07-06 Grimaldi-etal-BiolCybernetics.tex
@@
 on event streams for which the ground truth motion is known at each instant.
+The accuracy was computed on a novel dataset of $200$ novel movies.
 The accuracy computed on the test set was approximately $89\%$ (with a chance level of $1 / 12/ 3 \approx 2\%$).
@@
 On the other hand, the pruning strategy (shown in blue) exhibits a different behavior.
-It reaches half-saturation when the ratio of active weights is approximately $3.2\times 10^{-2}$, corresponding to around 180 computations per decision in our specific scenario.
+It reaches half-saturation when the ratio of active weights is approximately $2.6\times 10^{-3}$, corresponding to around $375$ less computations compared to the dense scenario.
 In comparison to using the complete kernels, our method maintains accuracy close to its peak performance even when the number of computations is divided by a factor of up to around 31.

### 2023-07-11 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
 In neural models, the definition of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows the efficient detection of these spatiotemporal motifs embedded in the spike train.
-These spatiotemporal motifs present in neural activity may form useful representations to perform computations for various cognitive tasks using the synchrony of spikes reaching the soma of a neuron.
+Such spatiotemporal motifs present in neural activity may form useful representations to perform computations for various cognitive tasks using the synchrony of spikes reaching the soma of a neuron.
 In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a spiking motif defined by a subset of neurons with different, but precise, relative spiking delays.
@@
 In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}), very fast (about $80~\ms$), and have a wide range of speeds.
-On a more microscopic scale, the human gaze moves with minute micro-saccades which trajectories are similar a Brownian trajectory~\citep{poletti_head-eye_2015}.
+On a more microscopic scale, the human gaze moves with minute micro-saccades which trajectories are similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
 To preserve the full generality of the task, we define eye movements using such a form of random walk~\citep{engbert_integrated_2011}: We first define a finite set of possible motions in polar coordinates as the Cartesian product of the $12$ regularly spaced directions and $3$ geometrically spaced speeds.
@@
 %%%%-----------------------------------------------------------------
-\subsection{The Heterogeneous Delays SNN (HD-SNN) model}
+\subsection{The HD-SNN model}
 %%%-----------------------------------------------------------------
@@
 The accuracy was computed on a novel dataset of $200$ novel movies.
-The accuracy computed on the test set was approximately $89\%$ (with a chance level of $1 / 12/ 3 \approx 2\%$).
+The accuracy computed on the test set was approximately $91\%$ (with a chance level of $1 / 12/ 3 \approx 2\%$).
 %Moreover, we observed that the classification performance obtained for the model using time surfaces as in~\citep{grimaldi_robust_2022} gives similar results as our method.
@@
 The HD-SNN model was trained and evaluated on a naturalistic motion detection task with realistic eye movements.
-The model is defined such as to provide an optimal detection of  spatiotemporal motifs and learns kernels similar to those found in the visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
+It is defined such as to provide an optimal detection of  spatiotemporal motifs and learns kernels similar to those found in the visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
 We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware.

### 2023-07-12 cover_letter/2023-07-07_coverletter.tex
Update 2023-07-07_coverletter.tex

@@
 Results show that this method could provide a path for future spiking neural network algorithms using less energy for a similar performance as their analog counterparts.
-This work is the extension of a previous study on a simpler event-based dataset with a simplified model.
-Our contribution here is to offer a better formal description of the model, an application on a more ecological task and testing of the influence of various parameters of the validation dataset.
 
 We thank the reviewer for the positive assesment which encouraged us to further improve our manuscript.
+As a result, this revision is a major overhaul of the manuscript which followed the comments from the reviewers.
+Notably, we have extended the results to natural scenes which gives a broader impact to the paper.
 Given the substantial modifications we did to the original manuscript to reduce and clarify the different sections we do not provide a tracked changes version of the revision (such a document can be provided if needed).

### 2023-07-12 cover_letter/2023-07-07_coverletter.tex
Update 2023-07-07_coverletter.tex

@@
 %----------------------------------------------------------------------------------------
-Please find enclosed the revised manuscript entitled ``Learning heterogeneous delays in a layer of spiking neurons for fast motion detection'' for your consideration as an article in \emph{Biological Cybernetics}.
-Both authors have been involved with the work, have approved the manuscript and agreed to its revision.
+Please find attached the revised manuscript entitled ``Learning heterogeneous delays in a layer of spiking neurons for fast motion detection'' for consideration as an article in \emph{Biological Cybernetics}.
+Both authors were involved in the work, approved the manuscript, and agreed to revise it.
 The original submission ID is a7f45f8e-87d6-4a30-ac7f-9f0c94e2592a.
 
-Our manuscript describes a novel method inspired by neuroscience to overcome some challenges faced in computer vision, notably when dealing with large amounts of data, by using an event-based representation.
-By using heterogeneous delays on different synapses, this novel spiking neuron method is able to detect spiking motifs and we validate the method on synthetic event-based data.
-Results show that this method could provide a path for future spiking neural network algorithms using less energy for a similar performance as their analog counterparts.
+Our manuscript describes a novel method inspired by neuroscience to overcome some challenges in computer vision, especially when dealing with large amounts of data, by using an event-based representation.
+By using heterogeneous delays at different synapses, this novel spiking neuron method is able to detect spiking motifs, and we validate the method on synthetic event-based data.
+Results show that this method could provide a path for future spiking neural network algorithms to consume less energy for similar performance as their analog counterparts.
 
-We thank the reviewer for the positive assesment which encouraged us to further improve our manuscript.
-As a result, this revision is a major overhaul of the manuscript which followed the comments from the reviewers.
-Notably, we have extended the results to natural scenes which gives a broader impact to the paper.
-Given the substantial modifications we did to the original manuscript to reduce and clarify the different sections we do not provide a tracked changes version of the revision (such a document can be provided if needed).
-We provide below a point by point response to the reviewers with references to the manuscript.
+We thank the reviewer for the positive assessment, which encouraged us to further improve our manuscript.
+As a result, this revision is a major overhaul of the manuscript that follows the reviewers' comments.
+In particular, we have extended the results to natural scenes, which gives the paper a broader impact.
+Given the substantial changes we have made to the original manuscript to reduce and clarify the various sections, we do not provide a tracked changes version of the revision (such a document can be provided if needed).
+Below we provide a point-by-point response to the reviewers with references to the manuscript.
 
@@
 
- \noteRev{We thank the reviewer for his constructive comment.
-While keeping the same computationnal framework, this encouraged us to extend our task from using natural-like textures to actual natural images.
-This was further extended by using biologically inspired eye movements to generate dynamical scenes with known ground truth.
-We have kept the use of the synthetic textures in order to assess the relevant image parameters that are relevant for motion detection and we have drawn the analogies of our model with that found in the neurophysiological and psychophysical literature.
-We have added subgraph inset to illustrate the changes in parameters to the generation of figures.}
+\noteRev{We thank the reviewer for his constructive comments.
+While keeping the same computational framework, this encouraged us to extend our task from using natural-like textures to actual natural images.
+This was further extended by using biologically inspired eye movements to generate dynamic scenes with known ground truth.
+We retained the use of synthetic textures to evaluate the relevant image parameters for motion detection, and we drew the analogies of our model to those found in the neurophysiological and psychophysical literature.
+We have added a subgraph inset to illustrate the changes in parameters for the generation of figures.}
 
@@
 
- \noteRev{We have repeated the testing of our model using $200$ movies and report less variable estimates for the accuracy. Thanks for this suggestion.}
+\noteRev{We have repeated the testing of our model using $200$ movies and report less variable estimates for accuracy. Thank you for this suggestion.}
 
@@
 
- \noteRev{By extending the learning and testing of our model on this newly defined motion detection task, we hope to have provided with a more thorough exploration of the capabilities of our model.
-The comparison of accuracy was also performed on the synthetic textures dataset in order to provide with a direct comparison with biological observations.}
+\noteRev{By extending the learning and testing of our model to this redefined motion detection task, we hope to have provided a more thorough exploration of the capabilities of our model.
+The accuracy comparison was also performed on the synthetic texture dataset to provide a direct comparison with biological observations.}
 

### 2023-07-12 cover_letter/2023-07-07_coverletter.tex
2022-10-12

@@
 
-Our manuscript describes a novel method inspired by neuroscience to overcome some challenges in computer vision, especially when dealing with large amounts of data, by using an event-based representation.
-By using heterogeneous delays at different synapses, this novel spiking neuron method is able to detect spiking motifs, and we validate the method on synthetic event-based data.
-Results show that this method could provide a path for future spiking neural network algorithms to consume less energy for similar performance as their analog counterparts.
+Our manuscript presents an innovative approach inspired by neuroscience to address challenges in computer vision, particularly when handling large datasets.
+We propose the use of an event-based representation and introduce a novel spiking neuron method that utilizes heterogeneous delays on different synapses.
+This method effectively detects spiking motifs, as demonstrated through experiments conducted on synthetic event-based data.
+Our results indicate that this approach has the potential to pave the way for future spiking neural network algorithms, offering comparable performance to their analog counterparts while consuming less energy.
 

### 2023-07-12 Grimaldi-etal-BiolCybernetics.tex
fig:activity

@@
     \centering
-    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
+    \includegraphics[width=0.95\linewidth]{figures/conv_HD-SNN.pdf}
     \caption{
-        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
-The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
-As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about $250$ times fewer computations.
+        In response to an instance of the event-based input (Left), we show the neural activity of the HD-SNN model (Right).
+For visualization purposes, we show only one spatio-temporal slice for a given vertical position ($y = 32$), for a quarter of horizontal positions.
+The input spiking activity consists of ON and OFF spikes (as in~\seeFig{motion_task}) and illustrates the switching of the naturalistic event-based stream from one motion to another.
+The output activity consists of an analog component corresponding to the evidence accumulated by the model on the spatio-temporal kernels, and a spiking part (overlaid on this activity as vertical bars) which corresponds to instants when this evidence is above the spiking threshold.
+Note that this activity is consistent with the motion of the input stream.
+The class of each output neuron is represented by a similar color as in \seeFig{motion_task}.
+Finally, an accuracy can be computed by comparing the ground truth motion in the input movie with that predicted by the model.
         }
-    \label{fig:accuracy}
+    \label{fig:activity}
 \end{figure}
@@
 
+\begin{figure}%[h!]
+    \centering
+    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
+    \caption{
+        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
+The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
+As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about $250$ times fewer computations.
+        }
+    \label{fig:accuracy}
+\end{figure}
+%
 By selectively setting certain weights to zero, we observe that the accuracy's evolution, as a function of the logarithmic percentage of active weights, aligns well with a sigmoid curve for both pruning and shortening strategies.

### 2023-07-12 Grimaldi-etal-BiolCybernetics.tex
Update Grimaldi-etal-BiolCybernetics.tex

@@
 on event streams for which the ground truth motion is known at each instant.
-The accuracy was computed on a novel dataset of $200$ novel movies.
+When applied to new instances of the input movies, the model develops a neural activity which may be used to infer the correct motion~\seeFig{activity} and from which we may deduce an accuracy value.
+This accuracy was computed on a novel dataset of $200$ novel movies.
 The accuracy computed on the test set was approximately $91\%$ (with a chance level of $1 / 12/ 3 \approx 2\%$).

### 2023-07-12 Grimaldi-etal-BiolCybernetics.tex
conv_HD-SNN

@@
 For visualization purposes, we show only one spatio-temporal slice for a given vertical position ($y = 32$), for a quarter of horizontal positions.
-The input spiking activity consists of ON and OFF spikes (as in~\seeFig{motion_task}) and illustrates the switching of the naturalistic event-based stream from one motion to another.
+The input spiking activity consists of ON and OFF spikes (as in~\seeFig{motion_task}) and illustrates the switching of the naturalistic event-based stream from one motion to another, as denoted by the colored dots above the graph (with the class of each output neuron is represented by a similar color as in \seeFig{motion_task}).
 The output activity consists of an analog component corresponding to the evidence accumulated by the model on the spatio-temporal kernels, and a spiking part (overlaid on this activity as vertical bars) which corresponds to instants when this evidence is above the spiking threshold.
 Note that this activity is consistent with the motion of the input stream.
-The class of each output neuron is represented by a similar color as in \seeFig{motion_task}.
 Finally, an accuracy can be computed by comparing the ground truth motion in the input movie with that predicted by the model.

### 2023-07-12 Grimaldi-etal-BiolCybernetics.tex
abstract

@@
 \abstract{
-The precise timing of spikes from afferent neurons determines the response of a biological neuron.
-This temporal aspect of neural activity is essential for understanding information processing in neurobiology, and is particularly applicable to the output of neuromorphic hardware such as event-based cameras.
-However, most artificial neural models ignore this important temporal dimension of neural activity.
-Here, we develop a model for efficiently detecting temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
-The model exploits the property that the diversity of synaptic delays on the dendritic tree allows specific arrangements of temporally precise synaptic inputs to synchronize as they reach the basal dendritic tree.
-We show that this process can be formalized as a time-invariant logistic regression that can be trained on labeled data.
-We demonstrate its application to naturalistic videos transformed into event streams, similar to the output of the biological retina or event-based cameras.
-To characterize the robustness of the model in detecting visual motion, we test its efficiency by pruning some weights and show that it is still efficient at very low workloads.
-In conclusion, by providing an end-to-end, event-driven computational building block, the learning of heterogeneous delays could improve the performance of future spiking neural network algorithms, and in particular their applications in neuromorphic chips.
+The precise timing of spikes emitted by neurons plays a crucial role in shaping the response of efferent biological neurons.
+This temporal aspect of neural activity holds significant importance in understanding information processing in neurobiology, especially concerning the output of neuromorphic hardware like event-based cameras.
+However, many artificial neural models disregard this critical temporal dimension of neural activity.
+In this study, we present a model designed to efficiently detect temporal spiking motifs using a layer of spiking neurons equipped with heterogeneous synaptic delays.
+Our model capitalizes on the diverse synaptic delays present on the dendritic tree, enabling specific arrangements of temporally precise synaptic inputs to synchronize upon reaching the basal dendritic tree.
+We formalize this process as a time-invariant logistic regression that can be trained using labeled data.
+To demonstrate its practical efficiency, we apply the model to naturalistic videos transformed into event streams, resembling the output of the biological retina or event-based cameras.
+To assess the robustness of the model in detecting visual motion, we conduct experiments by selectively pruning weights and demonstrate that the model remains efficient even under significantly reduced workloads.
+In conclusion, by offering a comprehensive, event-driven computational building block, heterogeneous delays has the potential to enhance the performance of future spiking neural network algorithms, particularly within the realm of neuromorphic chips.
 }

### 2023-07-12 Grimaldi-etal-BiolCybernetics.tex
methods

@@
 The precise timing of spikes emitted by neurons plays a crucial role in shaping the response of efferent biological neurons.
-This temporal aspect of neural activity holds significant importance in understanding information processing in neurobiology, especially concerning the output of neuromorphic hardware like event-based cameras.
-However, many artificial neural models disregard this critical temporal dimension of neural activity.
+This temporal dimension of neural activity holds significant importance in understanding information processing in neurobiology, especially concerning the output of neuromorphic hardware, such as event-based cameras.
+Nonetheless, many artificial neural models disregard this critical temporal dimension of neural activity.
 In this study, we present a model designed to efficiently detect temporal spiking motifs using a layer of spiking neurons equipped with heterogeneous synaptic delays.
 Our model capitalizes on the diverse synaptic delays present on the dendritic tree, enabling specific arrangements of temporally precise synaptic inputs to synchronize upon reaching the basal dendritic tree.
-We formalize this process as a time-invariant logistic regression that can be trained using labeled data.
-To demonstrate its practical efficiency, we apply the model to naturalistic videos transformed into event streams, resembling the output of the biological retina or event-based cameras.
-To assess the robustness of the model in detecting visual motion, we conduct experiments by selectively pruning weights and demonstrate that the model remains efficient even under significantly reduced workloads.
-In conclusion, by offering a comprehensive, event-driven computational building block, heterogeneous delays has the potential to enhance the performance of future spiking neural network algorithms, particularly within the realm of neuromorphic chips.
+We formalize this process as a time-invariant logistic regression, which can be trained using labeled data.
+To demonstrate its practical efficacy, we apply the model to naturalistic videos transformed into event streams, simulating the output of the biological retina or event-based cameras.
+To evaluate the robustness of the model in detecting visual motion, we conduct experiments by selectively pruning weights and demonstrate that the model remains efficient even under significantly reduced workloads.
+In conclusion, by providing a comprehensive, event-driven computational building block, the incorporation of heterogeneous delays has the potential to enhance the performance of future spiking neural network algorithms, particularly in the context of neuromorphic chips.
 }
@@
 %%%-----------------------------------------------------------------
-In this paper, we test whether the HD-SNN model is able to efficiently learn to solve a motion detection task defined by a realistic event-based data stream.
-This type of signal is commonly captured by an event-based camera, or Dynamic Vision Sensor (DVS), and is inspired by the signal sent from the retinal ganglion cells through the optic nerve to the visual cortex.
-The events are binary in nature and should be sufficient to perform the motion detection task in a fast and efficient manner.
-Here, we first describe the task definition, then the HD-SNN model we use for motion inference, and finally how we train it.
+This paper aims to investigate the capability of the HD-SNN model to effectively learn and solve a motion detection task using realistic event-based data streams, as typically captured by event-based cameras, also known as Dynamic Vision Sensors (DVS).
+DVS are designed to mimic the signal transmitted from retinal ganglion cells to the visual cortex through the optic nerve.
+The events in these data streams are binary in nature and should provide sufficient information for performing fast and efficient motion detection.
+In this study, we first outline the task definition, specifying the requirements for motion detection.
+Subsequently, we describe the HD-SNN model employed for inferring motion, highlighting its key characteristics and architecture.
+Lastly, we elaborate on the training procedure employed to train the HD-SNN model for the specific motion detection task.
 %%%-----------------------------------------------------------------
@@
     \caption{
-        In response to an instance of the event-based input (Left), we show the neural activity of the HD-SNN model (Right).
-For visualization purposes, we show only one spatio-temporal slice for a given vertical position ($y = 32$), for a quarter of horizontal positions.
-The input spiking activity consists of ON and OFF spikes (as in~\seeFig{motion_task}) and illustrates the switching of the naturalistic event-based stream from one motion to another, as denoted by the colored dots above the graph (with the class of each output neuron is represented by a similar color as in \seeFig{motion_task}).
-The output activity consists of an analog component corresponding to the evidence accumulated by the model on the spatio-temporal kernels, and a spiking part (overlaid on this activity as vertical bars) which corresponds to instants when this evidence is above the spiking threshold.
-Note that this activity is consistent with the motion of the input stream.
-Finally, an accuracy can be computed by comparing the ground truth motion in the input movie with that predicted by the model.
+        In response to a specific event-based input instance (Left), we present the neural activity of the HD-SNN model (Right).
+To aid visualization, we display a single spatio-temporal slice for a given vertical position ($y = 32$) and $10$ horizontal positions.
+The input spiking activity comprises ON and OFF spikes, as explained in Fig.~\ref{fig:motion_task}, showcasing the switching within the naturalistic event-based stream from one motion to another due to saccades.
+The dots above the graph indicate the corresponding motion class at each instant (the motion being represented by the matching color).
+The output activity consists of two components.
+Firstly, there is an analog component that corresponds to the evidence accumulated by the model on the spatio-temporal kernels.
+Secondly, there is a spiking component represented by vertical bars superimposed on the analog activity.
+These bars signify moments when the evidence surpasses the spiking threshold.
+Importantly, this activity aligns with the motion depicted in the input stream.
+Finally, it is possible to compute the accuracy by comparing the ground truth motion in the input video with the motion predicted by the model (as represented by the colored dots).
         }
@@
 on event streams for which the ground truth motion is known at each instant.
-When applied to new instances of the input movies, the model develops a neural activity which may be used to infer the correct motion~\seeFig{activity} and from which we may deduce an accuracy value.
+When applied to new instances of the input movies, the model develops a neural activity which may be used to infer the correct motion~(\seeFig{activity}) and from which we may deduce an accuracy value.
 This accuracy was computed on a novel dataset of $200$ novel movies.

### 2023-07-12 Grimaldi-etal-BiolCybernetics.tex
captions

@@
 Note that these were pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band, similarly to a process known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-These images are $1024 \times 1024$ in size, and we crop images of size $128 \times 128$ positioned around the center of gaze at each time step.
+The full-scale images are $1024 \times 1024$ in size, and we crop images of size $128 \times 128$ positioned around the center of gaze at each time step.
 We discretize time in $1~\ms$ bins and produce movies of duration $N_T = 200~\ms$.
@@
 Importantly, this activity aligns with the motion depicted in the input stream.
-Finally, it is possible to compute the accuracy by comparing the ground truth motion in the input video with the motion predicted by the model (as represented by the colored dots).
+Finally, it is possible to compute the accuracy by comparing the ground truth motion in the input video with the motion predicted by the model (as represented by the colored dots on top of the graph).
         }
@@
 The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
-As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about $250$ times fewer computations.
+If we shorten the length of the kernel by using only the weights at the shortest delays, the accuracy quickly drops.
+However, if we prune the lowest coefficients from the whole kernel, we observe a stable accuracy value, with a drop to half-saturation observed at about $375$ times fewer computations.
         }

### 2023-07-12 Grimaldi-etal-BiolCybernetics.tex
captions

@@
 
+\begin{figure}%[h!]
+    \centering
+    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
+    \caption{
+        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
+The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
+If we shorten the length of the kernel by using only the weights at the shortest delays, the accuracy quickly drops.
+However, if we prune the lowest coefficients from the whole kernel, we observe a stable accuracy value, with a drop to half-saturation observed at about $375$ times fewer computations.
+        }
+    \label{fig:accuracy}
+\end{figure}
+%
 We also observed that the distribution of the kernel's weights is sparse, with most values close to zero (\seeFig{kernels}).
@@
 
-\begin{figure}%[h!]
-    \centering
-    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
-    \caption{
-        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
-The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
-If we shorten the length of the kernel by using only the weights at the shortest delays, the accuracy quickly drops.
-However, if we prune the lowest coefficients from the whole kernel, we observe a stable accuracy value, with a drop to half-saturation observed at about $375$ times fewer computations.
-        }
-    \label{fig:accuracy}
-\end{figure}
-%
 By selectively setting certain weights to zero, we observe that the accuracy's evolution, as a function of the logarithmic percentage of active weights, aligns well with a sigmoid curve for both pruning and shortening strategies.
@@
 %%%-----------------------------------------------------------------
+In order to assess the influence of spatiotemporal parameters of the stimuli on the performance of the model, we now test the model on simpler, parameterized stimuli.
+For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control relevant parameters for motion detection, including motion direction, spatial orientation, and spatial frequency along with their respective precisions (\seeFig{motion_clouds})~\citep{Leon2012,vacher_bayesian_2018}.
+By matching the spatial and temporal characteristics of the generated movies with those of the motion task mentioned earlier, we created a range of textures featuring different spatial properties and motions.
+This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $12$ directions and $3$ speeds.
+For each motion, we also varied the texture parameters, such as mean and variance of orientation or spatial frequency content, to provide some naturalistic variability.
+This method provides a rich dataset of textured movies for which we know the ground truth for the motion.
+
 %%%-----------------------------------------------------------------
@@
     % TODO : show prototypes of textures on top
-    \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from a grating (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to grating (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)).
+    \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from gratings (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to gratings (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)).
 Examples snapshots are shown as an illustration in the top insets.
@@
 %%%-----------------------------------------------------------------
-In order to assess the influence of spatiotemporal parameters of the stimuli on the performance of the model, we now test the model on simpler, parameterized stimuli.
-For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control relevant parameters for motion detection, including motion direction, spatial orientation, and spatial frequency along with their respective precisions (\seeFig{motion_clouds})~\citep{Leon2012,vacher_bayesian_2018}.
-By matching the spatial and temporal characteristics of the generated movies with those of the motion task mentioned earlier, we created a range of textures featuring different spatial properties and motions.
-This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $12$ directions and $3$ speeds.
-For each motion, we also varied the texture parameters, such as mean and variance of orientation or spatial frequency content, to provide some naturalistic variability.
-This method provides a rich dataset of textured movies for which we know the ground truth for the motion.
-
 We observe some interesting facts.

### 2023-07-12 Grimaldi-etal-BiolCybernetics.tex
results

@@
 To determine the spiking output, we establish a firing threshold.
-We computed this threshold to ensure that neurons, on average, generate one spike per second.
+Here, we computed this threshold to ensure that neurons, on average, generate one spike per second.
 Therefore, the spiking output of the model corresponds to the motions in space and time that represent the highest probability.
@@
 Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration).
+In our simulations, we observed that including this prior accelerated the learning but was not necessary to reach convergence.
 %This is consistent with neuroscientific principles because, due to the limits of conduction delay, the synaptic delay is physically related to the distance between the presynaptic and postsynaptic neurons.
@@
 Secondly, there is a spiking component represented by vertical bars superimposed on the analog activity.
-These bars signify moments when the evidence surpasses the spiking threshold.
+These spikes signify moments when the evidence surpasses the spiking threshold.
 Importantly, this activity aligns with the motion depicted in the input stream.
@@
 This strategy involves adjusting the temporal width by selecting only the weights associated with the shortest delays.
-In comparison to the inference performed using the complete 3D kernels without any pruning ($21\times17\times17$), both approaches demonstrate a reduction in computational requirements as indicated by the number of non-zero weights.
+In comparison to the inference performed using the complete 3D kernels without any pruning ($36\times2\times21\times17\times17$), both approaches demonstrate a reduction in computational requirements as indicated by the number of non-zero weights.
 
 By selectively setting certain weights to zero, we observe that the accuracy's evolution, as a function of the logarithmic percentage of active weights, aligns well with a sigmoid curve for both pruning and shortening strategies.
-The shortening strategy (depicted in orange) demonstrates a rapid decline in accuracy, reaching half-saturation when approximately one-third of the weights are removed.
+The shortening strategy (depicted in orange) demonstrates a rapid decline in accuracy, reaching half-saturation when approximately one-third of the weights remain.
 On the other hand, the pruning strategy (shown in blue) exhibits a different behavior.

### 2023-07-12 Grimaldi-etal-BiolCybernetics.tex
Roxannnnnne!

@@
 Thanks to % Thomas Schatz, Rosa Cossart,
-Salvatore Giancani, Hugo Ladret, Camille Besnainou and Jean-Nicolas Jérémie for useful discussions prior to the elaboration of this work. 
+Salvatore Giancani, Hugo Ladret, Camille Besnainou, Jean-Nicolas Jérémie, Miles Keating, and Adrien Fois for useful discussions during the elaboration of this work. 
 \Funding %
@@
 This works is made reproducible.
-The code reproducing the manuscript and all figures is available on~\href{ https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}{GitHub}.
+The code reproducing the manuscript and all figures is available on~\href{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}{GitHub}.
 It also contains supplementary figures and results.
@@
 The precise timing of spikes emitted by neurons plays a crucial role in shaping the response of efferent biological neurons.
-This temporal dimension of neural activity holds significant importance in understanding information processing in neurobiology, especially concerning the output of neuromorphic hardware, such as event-based cameras.
+This temporal dimension of neural activity holds significant importance in understanding information processing in neurobiology, especially for the performance of neuromorphic hardware, such as event-based cameras.
 Nonetheless, many artificial neural models disregard this critical temporal dimension of neural activity.
@@
 To evaluate the robustness of the model in detecting visual motion, we conduct experiments by selectively pruning weights and demonstrate that the model remains efficient even under significantly reduced workloads.
-In conclusion, by providing a comprehensive, event-driven computational building block, the incorporation of heterogeneous delays has the potential to enhance the performance of future spiking neural network algorithms, particularly in the context of neuromorphic chips.
+In conclusion, by providing a comprehensive, event-driven computational building block, the incorporation of heterogeneous delays has the potential to greatly improve the performance of future spiking neural network algorithms, particularly in the context of neuromorphic chips.
 }
@@
 CNNs have achieved state-of-the-art performance for some computer vision tasks, such as image recognition.
-These networks are also known to mimic several properties of biological systems, such that each can be assigned a ``brain score''~\citep{schrimpf_brain-score_2020}.
-However, this score does not take into account key aspects of the efficiency of biological systems, such as inference speed or energy consumption, which is about $360$ watts on a standard GPU (NVIDIA RTX 3090).
+These networks are also known to mimic several properties of the biological visual system, such that each can be assigned a ``brain score''~\citep{schrimpf_brain-score_2020}.
+However, this score does not take into account key aspects of the efficiency of biological systems, such as inference speed (usually several times the biological time) or the energy consumption of this mesocopic model of the brain, which is about $360$ watts on a standard GPU (NVIDIA RTX 3090).
 In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing (binary) events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
@@
 %%%-----------------------------------------------------------------
-In this paper, we have introduced a generic heterogeneous delay SNN that was trained using supervised learning on a visual motion detection task.
-We have shown how the model performs well compared to other event-based classification algorithms for this task.
-The learned model has a number of similarities with neurobiological and behavioral observations.
-The event-based computations of our method can be drastically reduced by pruning synapses, while maintaining robust classification performance.
-This shows that we can exploit the precise timing of spikes to improve the performance of neural computations.
+This paper presents a novel and versatile heterogeneous delay spiking neural network (HD-SNN) that was trained using supervised learning for visual motion detection.
+We demonstrate the effectiveness of our model by comparing its performance to other event-based classification algorithms for this specific task.
+Notably, the learned model exhibits similarities with neurobiological and behavioral observations.
+One key advantage of our approach is the ability to significantly reduce the computational requirements through synapse pruning, while still maintaining robust classification performance.
+This highlights the potential to leverage the precise timing of spikes to enhance the efficiency and effectiveness of neural computations.
+Overall, our findings underscore the potential of incorporating precise spike timing in neural models and demonstrate the promising capabilities of our heterogeneous delay SNN for event-based computations, specifically in the context of visual motion detection.
 %

### 2023-07-18 Grimaldi-etal-BiolCybernetics.tex
Another pass
corrections from miles

@@
 289]{mandelbrot_fractal_1982}.
-Finally, assuming that  motion is stationary during each segment and that motions for each flight are drawn uniformly and independently, the global trajectory is generated by integrating this motion sequence.
+Finally, assuming motion is stationary during saccades, and that motions for each flight are drawn uniformly and independently, the global trajectory is generated by integrating this motion sequence.
 This generative model yields trajectories that are qualitatively very similar to those observed for human eye movements (\seeFig{motion_task}-\textit{(Left)}).
@@
 Now that this general framework has been described formally, we may include some heuristics based on neuroscientific observations to constrain our model and its strategies for solving the ecological task described in section~\ref{sec:task}.
-Note that the general framework is an extension to that presented in~\cite{grimaldi_learning_2022}, in particular by including a more complex task, the deeper analysis of the results, and these novel neuroscience-inspired heuristics.
+Note that the general framework is an extension to that presented in~\cite{grimaldi_learning_2022}, in particular by: including a more complex task, the deeper analysis of the results, and these novel neuroscience-inspired heuristics.
 First, to avoid introducing biases in the directions which may be learned, we apply a circular mask to the spatial dimensions of the kernels.

### 2023-07-21 Grimaldi-etal-BiolCybernetics.tex
sync

--- 2023-07-12 Grimaldi-etal-BiolCybernetics.tex
@@
 \abstract{
-The precise timing of spikes emitted by neurons plays a crucial role in shaping the response of efferent biological neurons.
-This temporal dimension of neural activity holds significant importance in understanding information processing in neurobiology, especially concerning the output of neuromorphic hardware, such as event-based cameras.
-Nonetheless, many artificial neural models disregard this critical temporal dimension of neural activity.
-In this study, we present a model designed to efficiently detect temporal spiking motifs using a layer of spiking neurons equipped with heterogeneous synaptic delays.
-Our model capitalizes on the diverse synaptic delays present on the dendritic tree, enabling specific arrangements of temporally precise synaptic inputs to synchronize upon reaching the basal dendritic tree.
-We formalize this process as a time-invariant logistic regression, which can be trained using labeled data.
-To demonstrate its practical efficacy, we apply the model to naturalistic videos transformed into event streams, simulating the output of the biological retina or event-based cameras.
-To evaluate the robustness of the model in detecting visual motion, we conduct experiments by selectively pruning weights and demonstrate that the model remains efficient even under significantly reduced workloads.
-In conclusion, by providing a comprehensive, event-driven computational building block, the incorporation of heterogeneous delays has the potential to enhance the performance of future spiking neural network algorithms, particularly in the context of neuromorphic chips.
+The precise timing of spikes from afferent neurons determines the response of a biological neuron.
+This temporal aspect of neural activity is essential for understanding information processing in neurobiology, and is particularly applicable to the output of neuromorphic hardware such as event-based cameras.
+However, most artificial neural models ignore this important temporal dimension of neural activity.
+Here, we develop a model for efficiently detecting temporal spiking motifs based on a layer of spiking neurons with heterogeneous synaptic delays.
+The model exploits the property that the diversity of synaptic delays on the dendritic tree allows specific arrangements of temporally precise synaptic inputs to synchronize as they reach the basal dendritic tree.
+We show that this process can be formalized as a time-invariant logistic regression that can be trained on labeled data.
+We demonstrate its application to naturalistic videos transformed into event streams, similar to the output of the biological retina or event-based cameras.
+To characterize the robustness of the model in detecting visual motion, we test its efficiency by pruning some weights and show that it is still efficient at very low workloads.
+In conclusion, by providing an end-to-end, event-driven computational building block, the learning of heterogeneous delays could improve the performance of future spiking neural network algorithms, and in particular their applications in neuromorphic chips.
 }
@@
 In neural models, the definition of heterogeneous delays~\citep{guise_bayesian_2014, zhang_supervised_2020, nadafian_bio-plausible_2020} allows the efficient detection of these spatiotemporal motifs embedded in the spike train.
-Such spatiotemporal motifs present in neural activity may form useful representations to perform computations for various cognitive tasks using the synchrony of spikes reaching the soma of a neuron.
+These spatiotemporal motifs present in neural activity may form useful representations to perform computations for various cognitive tasks using the synchrony of spikes reaching the soma of a neuron.
 In particular,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a spiking motif defined by a subset of neurons with different, but precise, relative spiking delays.
@@
 %%%-----------------------------------------------------------------
-This paper aims to investigate the capability of the HD-SNN model to effectively learn and solve a motion detection task using realistic event-based data streams, as typically captured by event-based cameras, also known as Dynamic Vision Sensors (DVS).
-DVS are designed to mimic the signal transmitted from retinal ganglion cells to the visual cortex through the optic nerve.
-The events in these data streams are binary in nature and should provide sufficient information for performing fast and efficient motion detection.
-In this study, we first outline the task definition, specifying the requirements for motion detection.
-Subsequently, we describe the HD-SNN model employed for inferring motion, highlighting its key characteristics and architecture.
-Lastly, we elaborate on the training procedure employed to train the HD-SNN model for the specific motion detection task.
+In this paper, we test whether the HD-SNN model is able to efficiently learn to solve a motion detection task defined by a realistic event-based data stream.
+This type of signal is commonly captured by an event-based camera, or Dynamic Vision Sensor (DVS), and is inspired by the signal sent from the retinal ganglion cells through the optic nerve to the visual cortex.
+The events are binary in nature and should be sufficient to perform the motion detection task in a fast and efficient manner.
+Here, we first describe the task definition, then the HD-SNN model we use for motion inference, and finally how we train it.
 %%%-----------------------------------------------------------------
@@
 In humans, saccades are very frequent (on average $2$ per second~\citep{dandekar_neural_2012}), very fast (about $80~\ms$), and have a wide range of speeds.
-On a more microscopic scale, the human gaze moves with minute micro-saccades which trajectories are similar to a Brownian trajectory~\citep{poletti_head-eye_2015}.
+On a more microscopic scale, the human gaze moves with minute micro-saccades which trajectories are similar a Brownian trajectory~\citep{poletti_head-eye_2015}.
 To preserve the full generality of the task, we define eye movements using such a form of random walk~\citep{engbert_integrated_2011}: We first define a finite set of possible motions in polar coordinates as the Cartesian product of the $12$ regularly spaced directions and $3$ geometrically spaced speeds.
@@
 Note that these were pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band, similarly to a process known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
-The full-scale images are $1024 \times 1024$ in size, and we crop images of size $128 \times 128$ positioned around the center of gaze at each time step.
+These images are $1024 \times 1024$ in size, and we crop images of size $128 \times 128$ positioned around the center of gaze at each time step.
 We discretize time in $1~\ms$ bins and produce movies of duration $N_T = 200~\ms$.
@@
 %%%%-----------------------------------------------------------------
-\subsection{The HD-SNN model}
+\subsection{The Heterogeneous Delays SNN (HD-SNN) model}
 %%%-----------------------------------------------------------------
@@
 Importantly, this activation function has to be such that when postsynaptic neurons are activated at their soma by a specific spatiotemporal motif imprinted in the synaptic set, and such that these spikes converge at the soma in a synchronous manner, the discharge probability should increase.
-In this subsection, we have briefly defined the HD-SNN model in all generality (see~\citep{Perrinet23ICANN} for a more specific description and treatment), and in the next subsection we describe an implementation of our model adapted to the motion detection task.
+In this subsection, we have briefly defined the HD-SNN model in all generality (see~\citep{KeatingPerrinet23ICANN} for a more specific description and treatment), and in the next subsection we describe an implementation of our model adapted to the motion detection task.
 %
@@
 To determine the spiking output, we establish a firing threshold.
-Here, we computed this threshold to ensure that neurons, on average, generate one spike per second.
+We computed this threshold to ensure that neurons, on average, generate one spike per second.
 Therefore, the spiking output of the model corresponds to the motions in space and time that represent the highest probability.
@@
 Since we want to capture the possible convergence of the trajectories of the events converging on each voxel, we apply a mask to the spatiotemporal kernels such that the smaller the delay, the smaller the radius of the circular mask that is applied (\seeFig{kernels} for an illustration).
-In our simulations, we observed that including this prior accelerated the learning but was not necessary to reach convergence.
 %This is consistent with neuroscientific principles because, due to the limits of conduction delay, the synaptic delay is physically related to the distance between the presynaptic and postsynaptic neurons.
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{figures/conv_HD-SNN.pdf}
+    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
     \caption{
-        In response to a specific event-based input instance (Left), we present the neural activity of the HD-SNN model (Right).
-To aid visualization, we display a single spatio-temporal slice for a given vertical position ($y = 32$) and $10$ horizontal positions.
-The input spiking activity comprises ON and OFF spikes, as explained in Fig.~\ref{fig:motion_task}, showcasing the switching within the naturalistic event-based stream from one motion to another due to saccades.
-The dots above the graph indicate the corresponding motion class at each instant (the motion being represented by the matching color).
-The output activity consists of two components.
-Firstly, there is an analog component that corresponds to the evidence accumulated by the model on the spatio-temporal kernels.
-Secondly, there is a spiking component represented by vertical bars superimposed on the analog activity.
-These spikes signify moments when the evidence surpasses the spiking threshold.
-Importantly, this activity aligns with the motion depicted in the input stream.
-Finally, it is possible to compute the accuracy by comparing the ground truth motion in the input video with the motion predicted by the model (as represented by the colored dots on top of the graph).
+        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
+The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
+As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about $250$ times fewer computations.
         }
-    \label{fig:activity}
+    \label{fig:accuracy}
 \end{figure}
@@
 on event streams for which the ground truth motion is known at each instant.
-When applied to new instances of the input movies, the model develops a neural activity which may be used to infer the correct motion~(\seeFig{activity}) and from which we may deduce an accuracy value.
-This accuracy was computed on a novel dataset of $200$ novel movies.
+The accuracy was computed on a novel dataset of $200$ novel movies.
 The accuracy computed on the test set was approximately $91\%$ (with a chance level of $1 / 12/ 3 \approx 2\%$).
@@
 
-\begin{figure}%[h!]
-    \centering
-    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
-    \caption{
-        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line).
-The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel.
-If we shorten the length of the kernel by using only the weights at the shortest delays, the accuracy quickly drops.
-However, if we prune the lowest coefficients from the whole kernel, we observe a stable accuracy value, with a drop to half-saturation observed at about $375$ times fewer computations.
-        }
-    \label{fig:accuracy}
-\end{figure}
-%
 We also observed that the distribution of the kernel's weights is sparse, with most values close to zero (\seeFig{kernels}).
@@
 This strategy involves adjusting the temporal width by selecting only the weights associated with the shortest delays.
-In comparison to the inference performed using the complete 3D kernels without any pruning ($36\times2\times21\times17\times17$), both approaches demonstrate a reduction in computational requirements as indicated by the number of non-zero weights.
+In comparison to the inference performed using the complete 3D kernels without any pruning ($21\times17\times17$), both approaches demonstrate a reduction in computational requirements as indicated by the number of non-zero weights.
 
 By selectively setting certain weights to zero, we observe that the accuracy's evolution, as a function of the logarithmic percentage of active weights, aligns well with a sigmoid curve for both pruning and shortening strategies.
-The shortening strategy (depicted in orange) demonstrates a rapid decline in accuracy, reaching half-saturation when approximately one-third of the weights remain.
+The shortening strategy (depicted in orange) demonstrates a rapid decline in accuracy, reaching half-saturation when approximately one-third of the weights are removed.
 On the other hand, the pruning strategy (shown in blue) exhibits a different behavior.
@@
 %%%-----------------------------------------------------------------
+%%%-----------------------------------------------------------------
+%: fig:motion_clouds
+%%%-----------------------------------------------------------------
+\begin{figure}%[h!]
+    \centering
+    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
+    % TODO : show prototypes of textures on top
+    \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from a grating (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to grating (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)).
+Examples snapshots are shown as an illustration in the top insets.
+Note that these accuracies are computed both in the case where the orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) and in the generic case where the orientation is independent of direction ('aperture').
+    }
+    \label{fig:motion_clouds}
+\end{figure}
+%%%-----------------------------------------------------------------
 In order to assess the influence of spatiotemporal parameters of the stimuli on the performance of the model, we now test the model on simpler, parameterized stimuli.
@@
 
-%%%-----------------------------------------------------------------
-%: fig:motion_clouds
-%%%-----------------------------------------------------------------
-\begin{figure}%[h!]
-    \centering
-    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
-    % TODO : show prototypes of textures on top
-    \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from gratings (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to gratings (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)).
-Examples snapshots are shown as an illustration in the top insets.
-Note that these accuracies are computed both in the case where the orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) and in the generic case where the orientation is independent of direction ('aperture').
-    }
-    \label{fig:motion_clouds}
-\end{figure}
-%%%-----------------------------------------------------------------
 We observe some interesting facts.
@@
 The HD-SNN model was trained and evaluated on a naturalistic motion detection task with realistic eye movements.
-It is defined such as to provide an optimal detection of  spatiotemporal motifs and learns kernels similar to those found in the visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
+The model is defined such as to provide an optimal detection of  spatiotemporal motifs and learns kernels similar to those found in the visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}.
 We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware.
@@
 This type of learning is thought to be prevalent in the central nervous system and, assuming the signal is sparse~\citep{olshausen_emergence_1996}, one could extend these Hebbian sparse learning schemes to spikes~\citep{perrinet_emergence_2004, masquelier_competitive_2009}.
-We expect that this would be particularly useful for exploring neurobiological data~\citep{Perrinet23ICANN}.
+We expect that this would be particularly useful for exploring neurobiological data~\citep{KeatingPerrinet23ICANN}.
 Indeed, there is a large literature showing that brain dynamics often organize into stereotyped sequences, such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}.

### 2023-07-21 cover_letter/2023-07-07_coverletter.tex

@@
 %13005 Marseille, France
-\newcommand{\LastName}{Grimaldi}%{Perrinet}%
-\newcommand{\FirstName}{Antoine}%{Laurent U.}%
+\newcommand{\LastName}{Perrinet}%Grimaldi}%{
+\newcommand{\FirstName}{Laurent U.}%Antoine}%{
 \newcommand{\Institute}{Institut de Neurosciences de la Timone, CNRS / Aix-Marseille Universit\'e}%
@@
 
-We thank the reviewer for the positive assessment, which encouraged us to further improve our manuscript.
-As a result, this revision is a major overhaul of the manuscript that follows the reviewers' comments.
-In particular, we have extended the results to natural scenes, which gives the paper a broader impact.
-Given the substantial changes we have made to the original manuscript to reduce and clarify the various sections, we do not provide a tracked changes version of the revision (such a document can be provided if needed).
-Below we provide a point-by-point response to the reviewers with references to the manuscript.
+We express our gratitude to the reviewer for their positive assessment, which motivated us to further enhance our manuscript.
+In response to their valuable comments, we have undertaken a major revision of the paper.
+Notably, we have expanded our experimental results to include natural scenes, thereby increasing the broader impact of our work.
+As we have extensively modified and streamlined various sections of the manuscript for improved clarity, we have not provided a tracked changes version of the revision.
+However, we are prepared to furnish such a document upon request.
+To address the reviewer's comments, we have provided a detailed response below, referring to the relevant sections of the manuscript.
 
-%\vspace{.2\baselineskip}
 %----------------------------------------------------------------------------------------
-Sincerely yours,%\vspace{.1cm}
+Sincerely yours,
 
 \FirstName\ \LastName \\
+
+\includegraphics[height=3cm]{~/quantic/RTC/signature_LuP.jpg}
 

### 2023-07-25 Grimaldi-etal-BiolCybernetics.tex
arxiv

--- 2023-07-18 Grimaldi-etal-BiolCybernetics.tex
@@
+\pdfoutput=1
 % !TEX encoding = UTF-8 Unicode

### 2023-07-25 Grimaldi-etal-BiolCybernetics.tex

@@
 Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of non-zero synaptic weights.
-This hypothesis is consistent with biological observations which have shown that communication consumes 35 times more energy than computation in the human cortex~\citep{Levy2021}.
+This hypothesis is consistent with biological observations which have shown that communication consumes $35$ times more energy than computation in the human cortex~\citep{Levy2021}.
 
@@
 It reaches half-saturation when the ratio of active weights is approximately $2.6\times 10^{-3}$, corresponding to around $375$ less computations compared to the dense scenario.
-In comparison to using the complete kernels, our method maintains accuracy close to its peak performance even when the number of computations is divided by a factor of up to around 31.
+In comparison to using the complete kernels, our method maintains accuracy close to its peak performance even when the number of computations is divided by a factor of up to around $31$.
 This substantial reduction in computations showcases the robustness of the presented method.
@@
 \bibliographystyle{apalike}
-\bibliography{FastMotionDetection}
+\bibliography{references}
 \end{document}

### 2023-09-09 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
Update Grimaldi-etal-BiolCybernetics.tex
proofs

@@
 % public
-% https://github.com/SpikeAI/pyTERtorch/ (for dev)
-% https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN (TODO: to become public)
+% https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN 
 % https://www.zotero.org/groups/4776796/fastmotiondetection
+% https://github.com/SpikeAI/pyTERtorch/ (for developpment code)
 %
@@
 % https://framateam.org/spikeai/channels/fast-motion-biolcyb
-% https://www.overleaf.com/project/630f9044c38e7a3cea81a7b2
+% https://www.overleaf.com/project/630f9044c38e7a3cea81a7b2 (OBSOLETE)
 %
@@
 \newcommand{\Acknowledgments}{
-Thanks to % Thomas Schatz, Rosa Cossart,
-Salvatore Giancani, Hugo Ladret, Camille Besnainou, Jean-Nicolas Jérémie, Miles Keating, and Adrien Fois for useful discussions during the elaboration of this work. 
+The authors thank Salvatore Giancani, Hugo Ladret, Camille Besnainou, Jean-Nicolas Jérémie, Miles Keating, and Adrien Fois for useful discussions during the elaboration of this work. 
 \Funding %
@@
 \newcommand{\DataAvailability}{
-This works is made reproducible.
-The code reproducing the manuscript and all figures is available on~\href{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}{GitHub}.
+This work is made reproducible.
+The code reproducing the manuscript and all figures is available on~\url{https://github.com/SpikeAI/2023_GrimaldiPerrinet_HeterogeneousDelaySNN}.
 It also contains supplementary figures and results.
-Find also the associated~\href{https://www.zotero.org/groups/4776796/fastmotiondetection}{zotero group} used to gather relevant literature on the subject.
+Find also the associated zotero group used to gather relevant literature on the subject at~\url{https://www.zotero.org/groups/4776796/fastmotiondetection}.
 }
@@
 %
-\usepackage{lineno}
-\linenumbers
+%\usepackage{lineno}
+%\linenumbers
+%\usepackage[switch]{lineno} % add line numbers
+%\modulolinenumbers[5]
 
@@
 Caution: with this package, bibliography will make unbreakable and underlined journal titles
-%\usepackage[switch]{lineno} % add line numbers
-%\modulolinenumbers[5]
 \usepackage{physics} %The  goal  of  this  package  is  to  make  typesetting  equations  for  physics  simpler,  faster,  and  more  human-readable.
@@
 These networks are also known to mimic several properties of the biological visual system, such that each can be assigned a ``brain score''~\citep{schrimpf_brain-score_2020}.
-However, this score does not take into account key aspects of the efficiency of biological systems, such as inference speed (usually several times the biological time) or the energy consumption of this mesocopic model of the brain, which is about $360$ watts on a standard GPU (NVIDIA RTX 3090).
+However, this score does not take into account key aspects of the efficiency of biological systems, such as inference speed (usually several times the biological time) or the energy consumption of this mesoscopic model of the brain, which is about $360$ watts on a standard GPU (NVIDIA RTX 3090).
 In the vast majority of biological neural networks, on the other hand, information is represented as \emph{spikes}, prototypical all-or-nothing (binary) events whose only parameters are their timing and the address of the neuron that fired the spike~\citep{paugam-moisy_computing_2012}.
@@
 For this purpose, we selected a database of $100$ large-scale natural images that were previously used to study the statistics of natural images~\citep{vanHateren1998}.
-Note that these were pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band, similarly to a process known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
+Note that these were pre-processed to be in grayscale and to equalize (i.e., whiten) the energy in each frequency band, similar to a process known to occur in the retino-thalamic pathway~\citep{dan_efficient_1996}.
 The full-scale images are $1024 \times 1024$ in size, and we crop images of size $128 \times 128$ positioned around the center of gaze at each time step.
@@
 Importantly, this activation function has to be such that when postsynaptic neurons are activated at their soma by a specific spatiotemporal motif imprinted in the synaptic set, and such that these spikes converge at the soma in a synchronous manner, the discharge probability should increase.
-In this subsection, we have briefly defined the HD-SNN model in all generality (see~\citep{KeatingPerrinet23ICANN} for a more specific description and treatment), and in the next subsection we describe an implementation of our model adapted to the motion detection task.
+In this subsection, we have briefly defined the HD-SNN model in all generality (see~\citep{Perrinet23ICANN} for a more specific description and treatment), and in the next subsection we describe an implementation of our model adapted to the motion detection task.
 %
@@
 First, neural addresses are defined to represent the range of possible positions and polarities.
-Second, to simulate such event-based computations on standard CPU- or GPU-based computers and to benefit from parallel computing acceleration, we transform the temporal event-based representation into a dense dicretized representation.
+Second, to simulate such event-based computations on standard CPU- or GPU-based computers and to benefit from parallel computing acceleration, we transform the temporal event-based representation into a dense discretized representation.
 Indeed, by using this discretization, we transform any event-based input from an event-based camera into a Boolean matrix $A \in \{0, 1 \}^{\Npol \times \Ntime \times \Nx \times \Ny}$ defined for all polarities $p$, times $t$, and space coordinates $x$ and $y$.
@@
 
-Then, it can be noted that by using a discretization, the computational block used in the equation~\eqref{eq:stream_b} corresponds to a weighted reordering of the input $A$ with each kernels and positions assigned to the postsynaptic neurons~\citep{grimaldi_learning_2022}.
-Let's define \emph{evidence} as the logit of a probability, that is, the inverse sigmoid of that probability.
+Then, it can be noted that by using a discretization, the computational block used in equation~\eqref{eq:stream_b} corresponds to a weighted reordering of the input $A$ with each kernels and positions assigned to the postsynaptic neurons~\citep{grimaldi_learning_2022}.
+Let us define \emph{evidence} as the logit of a probability, that is, the inverse sigmoid of that probability.
 By this definition of evidence, logistic regression takes advantage of the fact that if different independent observations (here, the estimated motion at different spatial locations and timings) share a common cause (here, the rigid local motion of the image on the receptive field), then an optimal estimate of the evidence of this motion is the sum of the evidences from the independent sources.
-Interpreting the weights of the kernel as evidences (also called factors in logistic regression), we may therefore define the activity $B$ of post-synaptic neurons as the integration of this activity in each voxel and for each channel $\class$ in order to infer the evidence of each motion:
+Interpreting the weights of the kernel as evidences (also called factors in logistic regression), we may therefore define the activity $B$ of postsynaptic neurons as the integration of this activity in each voxel and for each channel $\class$ in order to infer the evidence of each motion:
 %
-\begin{linenomath*}
+%\begin{linenomath*}
 \begin{equation}\label{eq:kernel_b}
@@
 \end{equation}
-\end{linenomath*}
+%\end{linenomath*}
 %
@@
 Interestingly, the relative organization of the receptive fields that we observe is in quadrature of phase and follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results obtained after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.
-Finally, we observe that these receptive fields show also a relative selectivity to the orientation perpendicular to motion, similarly to what is found for neurons in cortical area MT which is known to be selective to visual motions~\citep{deangelis_functional_1999}.
+Finally, we observe that these receptive fields show also a relative selectivity to the orientation perpendicular to motion, similar to what is found for neurons in cortical area MT which is known to be selective to visual motions~\citep{deangelis_functional_1999}.
 This reflects the way events are generated and in particular the so-called aperture problem which implies that a line moving along its axis would generate no change in luminance and therefore generate no event~\citep{perrinet_motion-based_2012}.
@@
 If we now widen our focus on the interpretation of these kernels in terms of spatiotemporal motifs embedded in the event stream, these show a prototypical anisotropic profile adapted to motion detection~\citep{kaplan_anisotropic_2013}.
-In~\citep{grimaldi_robust_2022}, a link was drawn between event-based MLR training and Hebbian learning, allowing to say that the present model learns its weights according to a presynaptic activity associated to the different motion directions.
+In~\citep{grimaldi_robust_2022}, a link was drawn between event-based MLR training and Hebbian learning, allowing to say that the present model learns its weights according to a presynaptic activity associated with the different motion directions.
 Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatiotemporal spiking motif.
 Each voxel in the 3D kernels defines a specific property by associating a weight to a position and a delay.
-Consequently, our model is able to detect precise spatiotemporal motifs embedded in the spike train and associated to the different motion directions.
+Consequently, our model is able to detect precise spatiotemporal motifs embedded in the spike train and associated with the different motion directions.
 Note that as the delays become larger, two effects can be remarked.
@@
 %
-\subsection{Accuracy versus efficiency trade-off}% for the motion detection task}
+\subsection{Accuracy versus efficiency tradeoff}% for the motion detection task}
 %%%-----------------------------------------------------------------
@@
         In response to a specific event-based input instance (Left), we present the neural activity of the HD-SNN model (Right).
-To aid visualization, we display a single spatio-temporal slice for a given vertical position ($y = 32$) and $10$ horizontal positions.
+To aid visualization, we display a single spatiotemporal slice for a given vertical position ($y = 32$) and $10$ horizontal positions.
 The input spiking activity comprises ON and OFF spikes, as explained in Fig.~\ref{fig:motion_task}, showcasing the switching within the naturalistic event-based stream from one motion to another due to saccades.
@@
 The output activity consists of two components.
-Firstly, there is an analog component that corresponds to the evidence accumulated by the model on the spatio-temporal kernels.
+Firstly, there is an analog component that corresponds to the evidence accumulated by the model on the spatiotemporal kernels.
 Secondly, there is a spiking component represented by vertical bars superimposed on the analog activity.
@@
 As a consequence, computations will be performed only on those events which were multiplexed by the pruned connectivity matrix.
-Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of non-zero synaptic weights.
+Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of nonzero synaptic weights.
 This hypothesis is consistent with biological observations which have shown that communication consumes $35$ times more energy than computation in the human cortex~\citep{Levy2021}.
@@
 This strategy involves adjusting the temporal width by selecting only the weights associated with the shortest delays.
-In comparison to the inference performed using the complete 3D kernels without any pruning ($36\times2\times21\times17\times17$), both approaches demonstrate a reduction in computational requirements as indicated by the number of non-zero weights.
+In comparison with the inference performed using the complete 3D kernels without any pruning ($36\times2\times21\times17\times17$), both approaches demonstrate a reduction in computational requirements as indicated by the number of nonzero weights.
 
@@
 It reaches half-saturation when the ratio of active weights is approximately $2.6\times 10^{-3}$, corresponding to around $375$ less computations compared to the dense scenario.
-In comparison to using the complete kernels, our method maintains accuracy close to its peak performance even when the number of computations is divided by a factor of up to around $31$.
+In comparison with using the complete kernels, our method maintains accuracy close to its peak performance even when the number of computations is divided by a factor of up to around $31$.
 This substantial reduction in computations showcases the robustness of the presented method.
@@
 In order to assess the influence of spatiotemporal parameters of the stimuli on the performance of the model, we now test the model on simpler, parameterized stimuli.
-For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control relevant parameters for motion detection, including motion direction, spatial orientation, and spatial frequency along with their respective precisions (\seeFig{motion_clouds})~\citep{Leon2012,vacher_bayesian_2018}.
+For this purpose, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control relevant parameters for motion detection, including motion direction, spatial orientation, and spatial frequency along with their respective precisions (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}.
 By matching the spatial and temporal characteristics of the generated movies with those of the motion task mentioned earlier, we created a range of textures featuring different spatial properties and motions.
@@
 to distinguish a figure from its background~\citep{roelfsema_early_2016}.
-Feedback has been shown to be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020}, especially to explain non-linear mechanisms~\citep{boutin_pooling_2022}.
+Feedback has been shown to be essential for building realistic models of primary visual areas~\citep{boutin_sparse_2020, boutin_effect_2020}, especially to explain nonlinear mechanisms~\citep{boutin_pooling_2022}.
 Currently, mainly due to our use of convolutions, it is not possible to implement these recurrent connections in our implementation (lateral or feedback).
@@
 Another perspective would be to extend the model to a fully self-supervised learning paradigm, i.e.
-without any labelled data~\citep{barlow_unsupervised_1989}.
+without any labeled data~\citep{barlow_unsupervised_1989}.
 This type of learning is thought to be prevalent in the central nervous system and, assuming the signal is sparse~\citep{olshausen_emergence_1996}, one could extend these Hebbian sparse learning schemes to spikes~\citep{perrinet_emergence_2004, masquelier_competitive_2009}.
-We expect that this would be particularly useful for exploring neurobiological data~\citep{KeatingPerrinet23ICANN}.
+We expect that this would be particularly useful for exploring neurobiological data~\citep{Perrinet23ICANN}.
 Indeed, there is a large literature showing that brain dynamics often organize into stereotyped sequences, such as synfire chains~\citep{ikegaya_synfire_2004}, packets~\citep{luczak_sequential_2007}, or hippocampal sequences~\citep{pastalkova_internally_2008, villette_internally_2015}.
@@
 \bibliographystyle{apalike}
-\bibliography{references}
+\bibliography{FastMotionDetection}
 \end{document}

### 2023-09-12 Grimaldi-etal-BiolCybernetics.tex
updated version for HAL + website

@@
 \title[\ShortTitle]{\Title}
-\artnote{This article is published as part of the Special Issue on "What can Computer Vision learn from Visual Neuroscience?"}
+\artnote{This article is published as part of the \textit{Biol Cyb} Special Issue on "What can Computer Vision learn from Visual Neuroscience?", see
+\url{https://laurentperrinet.github.io/publication/grimaldi-23-bc/}.
+}
 % ---------------
@@
 \author{\fnm{\FirstLP} \sur{\LastLP}}\email{\EmailLP}
-\affil{\orgdiv{\Department}, \orgname{\Affiliation}\\ \orgaddress{\street{\Street}, \city{\City}, \postcode{\PostCode}, \country{\Country}}}
+\affil{\orgdiv{\Department}, \orgname{\Affiliation}\\ \orgaddress{\street{\Street}, \city{\City}, \postcode{\PostCode}, \country{\Country}}
+}
 % 
@@
 This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation.
-Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{10}$ movies, each of size $200 \times 128 \times 128$, a learning rate of $10^{-5}$ and $100$ epochs).
+Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{10}$ movies, each of size $256 \times 128 \times 128$, a learning rate of $10^{-5}$ and $100$ epochs).
 %We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
@@
         In response to a specific event-based input instance (Left), we present the neural activity of the HD-SNN model (Right).
-To aid visualization, we display a single spatiotemporal slice for a given vertical position ($y = 32$) and $10$ horizontal positions.
+To aid visualization, we display a single spatiotemporal slice for a given vertical position (here $y = 32$) and the $10$ first horizontal positions.
 The input spiking activity comprises ON and OFF spikes, as explained in Fig.~\ref{fig:motion_task}, showcasing the switching within the naturalistic event-based stream from one motion to another due to saccades.
@@
 If we shorten the length of the kernel by using only the weights at the shortest delays, the accuracy quickly drops.
-However, if we prune the lowest coefficients from the whole kernel, we observe a stable accuracy value, with a drop to half-saturation observed at about $375$ times fewer computations.
+However, if we prune the lowest coefficients from the whole kernel, we observe a stable accuracy value, with a drop to half-saturation observed at about $670$ times fewer computations.
         }
@@
 On the other hand, the pruning strategy (shown in blue) exhibits a different behavior.
-It reaches half-saturation when the ratio of active weights is approximately $2.6\times 10^{-3}$, corresponding to around $375$ less computations compared to the dense scenario.
+It reaches half-saturation with $652$ synapses, that is, a ratio of active weights equal approximately to $1.5\times 10^{-3}$, corresponding to around $670$ less computations compared to the dense scenario.
 In comparison with using the complete kernels, our method maintains accuracy close to its peak performance even when the number of computations is divided by a factor of up to around $31$.
@@
 Examples snapshots are shown as an illustration in the top insets.
-Note that these accuracies are computed both in the case where the orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) and in the generic case where the orientation is independent of direction ('aperture').
+Note that these accuracies are computed both in the case where the orientation of the synthetic texture is necessarily perpendicular to the motion ('perpend' condition) and in the generic case where the orientation is independent of direction ('aperture').
     }

### 2023-09-14 2023-09-14_video-abstract.py
Create 2023-09-14_video-abstract.py

@@ -1,0 +1,275 @@
+"""
+# 2023-09-14_video-abstract
+
+
+ Using the template @ https://github.com/SpikeAI/2022_polychronies-review/blob/main/2022-12-23_video-abstract.py
+and using that @ https://github.com/SpikeAI/2022-09_UltraFastCat/blob/main/Jeremie-etal-Vision_video-abstract.py
+
+python 2023-09-14_video-abstract.py
+cp 2023-09-14_HDSNN_video-abstract.mp4 /Users/laurentperrinet/quantic/blog/hugo_academic/content/publication/grimaldi-23-bc/
+
+"""
+# Creating the movie using the (*excellent*) [MoviePy](http://zulko.github.io/moviepy/index.html) library:
+
+videoname = "2023-09-14_HDSNN_video-abstract"
+gifname = videoname + ".gif"
+gifname = None
+fps = 30
+from moviepy.editor import VideoFileClip, ImageClip, TextClip, CompositeVideoClip
+
+H, W = 500, 800
+H_fig, W_fig = int(H-H/(1.618*3)), int(W-W/(1.618*3))
+
+
+opt_t = dict(font="Arial", size=(W,H), method='caption')
+opt_st = dict(font="Arial", size=(W,H), method='caption')
+# opt_t = dict(size=(W,H), method='caption')
+# opt_st = dict(size=(W,H), method='caption')
+
+clip = []
+t = 0 
+
+#################################################################################
+# TITRE
+#################################################################################
+# 
+texts = ["""Learning heterogeneous delays of spiking neurons for motion detection
+
+
+
+
+
+         
+""", """Learning heterogeneous delays of spiking neurons for motion detection
+
+
+   Antoine Grimaldi and Laurent Udo Perrinet
+
+   
+
+""", """Learning heterogeneous delays of spiking neurons for motion detection
+
+
+   Antoine Grimaldi and Laurent Udo Perrinet
+
+   
+Biological Cybernetics (2023)
+"""]
+txt_opts = dict(align='center', color='white', **opt_t) #stroke_color='gray', stroke_width=.5
+duration = 1.5
+for text in texts:
+    txt = TextClip(text, fontsize=35, **txt_opts).set_start(t).set_duration(duration)
+    t += duration
+    clip.append(txt)
+
+#################################################################################
+# INTRO
+#################################################################################
+
+intro_subs = ["""The precise temporal structure of spiking neuronal
+ 
+              
+              
+              
+ 
+""", """The precise temporal structure of spiking neuronal
[…]

### 2023-09-14 2023-09-14_video-abstract.py
Update 2023-09-14_video-abstract.py

@@ -124,7 +124,7 @@ chapters = [dict(title="Visual motion detection", color='green',
                 dict(figure='figures/2023-07-12_FastMotionDetection_kernel.png', duration=7, subtitle=[
                         "This architecture is differentiable and can be...", 
                         "...trained in an unsupervised manner, yielding...", 
-                        "...spatio-temporal kernels which ressmble those...", 
+                        "...spatio-temporal kernels which ressemble those...", 
                         "...observed in neurobiology."]),
                         ]), 
            dict(title="Benchmarking on textures", color='orange',
@@ -135,13 +135,14 @@ chapters = [dict(title="Visual motion detection", color='green',
                         "...titrate the role of visual parameters such as...", 
                         "...spatial frequency, motion or orientation."]),
                 dict(figure='figures/2023-07-12_FastMotionDetection_motion_clouds.png', duration=5, subtitle=[
-                        "Results show characteristic tuning curves for these parameters...", 
+                        "Results show characteristic tuning curves for these...", 
                         "...and especially a drop of accuracy when the orientation...", 
                         "...is not perpendicular to motion, aka the 'aperture problem'."]),
                         ]), 
            dict(title="Results of the detection", color='blue',
             content=[
-                dict(figure='figures/2023-07-12_FastMotionDetection_conv_HD-SNN.png', duration=7, subtitle=[
+                # dict(figure='figures/2023-07-12_FastMotionDetection_conv_HD-SNN.png', duration=7, subtitle=[
+                dict(figure='/Users/laurentperrinet/quantic/science/Polychronies/pyTERtorch/2023-06-09_BiolCybernetics/figures/2023-07-08_FastMotionDetection_conv_HD-SNN.png', duration=7, subtitle=[
                         "The output of the heterogeneous delays SNN (HD-SNN) is...", 
                         "...a stream of events transforming the input events (left)...", 
                         "...into events informing about the most likely motion (right)...", 

### 2023-09-14 2023-09-14_video-abstract.py
Update 2023-09-14_video-abstract.py

@@ -20,6 +20,9 @@ from moviepy.editor import VideoFileClip, ImageClip, TextClip, CompositeVideoCli
 H, W = 500, 800
 H_fig, W_fig = int(H-H/(1.618*3)), int(W-W/(1.618*3))
 
+do_subtitle = False
+do_subtitle = True
+
 
 opt_t = dict(font="Arial", size=(W,H), method='caption')
 opt_st = dict(font="Arial", size=(W,H), method='caption')
@@ -194,9 +197,10 @@ for chapter in chapters:
         t_sub = t - duration
         sub_duration = duration / len(content['subtitle'])
         for subtitle in content['subtitle']:
-            sub = TextClip(subtitle, **sub_opts).set_start(t_sub).set_duration(sub_duration)
             t_sub += sub_duration
-            clip.append(sub)
+            if do_subtitle:
+                sub = TextClip(subtitle, **sub_opts).set_start(t_sub).set_duration(sub_duration)
+                clip.append(sub)
 
 #################################################################################
 #################################################################################

### 2023-09-15 Grimaldi-etal-BiolCybernetics.tex
re-run figures

@@
 In the lower left corner of the snapshots, we show the corresponding instantaneous motion vector (red arrow).
-Note the change in the direction of motion between the third and fourth frames, and also that contours parallel to the motion produce fewer luminance changes, the so-called aperture problem, and thus relatively fewer spikes.
+Note the change in the direction of motion between the second and third frames, and also that contours parallel to the motion produce fewer luminance changes, the so-called aperture problem, and thus relatively fewer spikes.
         }
@@
 This is reminiscent of previous methods that solve this problem using a winner-takes-all mechanism~\citep{masquelier_unsupervised_2007}, but is implicit in our formulation.
-Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{10}$ movies, each of size $256 \times 128 \times 128$, a learning rate of $10^{-5}$ and $100$ epochs).
+Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{10}$ movies, each of size $200 \times 128 \times 128$, a learning rate of $10^{-5}$ and $100$ epochs).
 %We have shown in a previous work on a simplified task using synthetic textures that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.
